### 2019-11-11 paper/ms.tex
Initial commit

% mnras_template.tex
%
% LaTeX template for creating an MNRAS paper
%
% v3.0 released 14 May 2015
% (version numbers match those of mnras.cls)
%
% Copyright (C) Royal Astronomical Society 2015
% Authors:
% Keith T. Smith (Royal Astronomical Society)

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.cls
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[a4paper,fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
\newcommand{\pc}{\ensuremath{\text{pc}}}
\newcommand{\kpc}{\ensuremath{\text{kpc}}}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}

\newcommand{\mwib}{\texttt{mwib}}

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Star Bar]{On Star Formation at Dynamical Resonances of the Bar}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[A. Beane et al.]{
% Angus Beane,$^{1}$\thanks{E-mail: abeane@cfa.harvard.edu}
Angus~Beane$^{1}$,\thanks{E-mail: abeane@cfa.harvard.edu}
Elena~D'Onghia$^{2,3}$,
Lars~Hernquist$^{1}$,
and Federico~Marinacci$^{4}$
\\
% List of institutions
% Center for Astrophysics {\normalfont |} Harvard \& Smithsonian, 60 Garden Street, Cambridge, MA 02138, USA
$^{1}$Center for Astrophysics {\normalfont |} Harvard \& Smithsonian, 60 Garden Street, Cambridge, MA 02138, USA\\
$^{2}$Department of Astronomy, University of Wisconsin, 475 North Charter Street, Madison, WI 53706, USA\\
$^{3}$Center for Computational Astrophysics, Flatiron Institute, 162 5th Avenue, New York, NY 10010, USA\\
$^{4}$Department of Physics \& Astronomy, University of Bologna, via Gobetti 93/2, 40129 Bologna, Italy
% $^{3}$Department of Physics \& Astronomy, University of Pennsylvania, 209 South 33rd Street, Philadelphia, PA 19104, USA
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2019}

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
Stars are known to be captured at the dynamical resonances of the Milky Way
bar, forming moving groups that are seen in the solar neighborhood. In
principle, gas can also be captured at these resonances which, if allowed to
cool, might trigger star formation -- predicting an enhanced number of
star-forming regions at these resonances. However, the efficiency of stellar
and radiation feedback may suppress the star formation process.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
Galaxy: disc -- Galaxy: kinematics and dynamics -- stars: kinematics and dynamics
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}

This is a simple template for authors to write new MNRAS papers.
See \texttt{mnras\_sample.tex} for a more complex example, and \texttt{mnras\_guide.tex}
for a full user guide.

All papers should start with an Introduction section, which sets the work
in context, cites relevant earlier studies in the field by \citet{Others2013},
and describes the problem the authors aim to solve \citep[e.g.][]{Author2012}.

\section{Methods, Observations, Simulations etc.}

Normally the next section describes the techniques the authors used.
It is frequently split into subsections, such as Section~\ref{sec:maths} below.

\subsection{Maths}
\label{sec:maths} % used for referring to this section from elsewhere

Simple mathematics can be inserted into the flow of the text e.g. $2\times3=6$
or $v=220$\,km\,s$^{-1}$, but more complicated expressions should be entered
as a numbered equation:

\begin{equation}
    x=\frac{-b\pm\sqrt{b^2-4ac}}{2a}.
	\label{eq:quadratic}
\end{equation}

Refer back to them as e.g. equation~(\ref{eq:quadratic}).

\subsection{Figures and tables}

Figures and tables should be placed at logical positions in the text. Don't
worry about the exact layout, which will be handled by the publishers.

Figures are referred to as e.g. Fig.~\ref{fig:example_figure}, and tables as
e.g. Table~\ref{tab:example_table}.

% Example figure
\begin{figure}
	% To include a figure from a file named example.*
	% Allowable file formats are eps or ps if compiling using latex
	% or pdf, png, jpg if compiling using pdflatex
	% \includegraphics[width=\columnwidth]{example}
    \caption{This is an example figure. Captions appear below each figure.
	Give enough detail for the reader to understand what they're looking at,
	but leave detailed discussion to the main body of the text.}
    \label{fig:example_figure}
\end{figure}

% Example table
\begin{table}
	\centering
	\caption{This is an example table. Captions appear above each table.
	Remember to define the quantities, symbols and units used.}
	\label{tab:example_table}
	\begin{tabular}{lccr} % four columns, alignment for each
		\hline
		A & B & C & D\\
		\hline
		1 & 2 & 3 & 4\\
		2 & 4 & 6 & 8\\
		3 & 5 & 7 & 9\\
		\hline
	\end{tabular}
\end{table}

\section{Conclusions}

The last numbered section should briefly summarise what has been done, and describe
the final conclusions which the authors draw from their work.

\section*{Acknowledgements}

The Acknowledgements section is not numbered. Here you can thank helpful
colleagues, acknowledge funding agencies, telescopes and facilities used etc.
Try to keep it short.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

% The best way to enter references is to use BibTeX:

\bibliographystyle{mnras}
\bibliography{references} % if your bibtex file is called example.bib

\appendix

\section{Two-component Toomre instability criterion}
The Toomre instability criterion for a two-component fluid was first derived
by \citet{1984ApJ...276..114J}. For a mode of wavenumber $k=2\pi/\lambda$ to
be stable against gravitational collapse, the criterion for the two fluids in
an infinitesimally thin disk is that,
\beq
Q_2(k) = \left(Q_g^{-1}(k) + Q_s^{-1}(k) \right)^{-1} > 1\text{,}
\eeq
where $Q_2(k)$ is the two-component Toomre parameter and,
\beq
Q_g(k) = \frac{\kappa^2 + k^2 c_s^2}{2\pi G k \Sigma_{g,0}}\text{,}
\eeq
where $\kappa$ is the radial epicyclic frequency, $c_s$ is the sound speed of
the gas, and $\Sigma_{g,0}$ is the surface density of the gas. $Q_s(k)$ can be
obtained by replacing $c_s$ by $\sigma_R$ (the radial velocity dispersion) and
$\Sigma_{g,0}$ by $\Sigma_{s,0}$ (the stellar surface density). Note that the
familiar one-component Toomre criterion can be obtained by minimizing $Q_g(k)$
as a function of $k$.

We denote the minimum of $Q_2(k)$ as a function of $k$ by $Q_2$. We consider
only physically plausible values of $k$ between $0.06$ and $6000\,\kpc^{-1}$
(corresponding to modes of wavelength $\lambda\sim100\,\kpc$ to $1\,\pc$). The
equivalent of the one-component Toomre stability criterion is that $Q_2>1$,
such that the disk is stable against collapse of modes of all wavelengths.

\section{\mwib{} resolution levels}
We set a series of standard resolution levels similar to those in the Aquarius
convention. Due to the isolated nature of our simulations, it is convenient to
tune the particle mass to a specific number.

\begin{table}
\begin{tabular}{cccc}
resolution level & $m_{\text{DM}}$      & $m_{\text{b}}$    & $\epsilon$        \\
                 & $[\,M_{\odot}\,]$    & $[\,M_{\odot}\,]$ & $[\,\text{pc}\,]$ \\
$5$              & $2.4\times10^6$      & $4.8\times10^5$   & $700$             \\
$4$              & $3\times10^5$        & $6\times10^4$     & $350$             \\
$3$              & $3.75 \times 10^{4}$ & $7500$            & $175$             \\
$2$              & $4690$               & $938$             & $87.5$            \\
$1$              & $586$                & $117$             & $43.8$           
\end{tabular}
\end{table}

\begin{table}
\begin{tabular}{cccc}
resolution level & $m_{\text{DM}}$      & $m_{\text{b}}$    & $\epsilon$        \\
                 & $[\,M_{\odot}\,]$    & $[\,M_{\odot}\,]$ & $[\,\text{pc}\,]$ \\
$5$h             & $2.4\times10^6$      & $4.8\times10^5$   & $700$             \\
$4$h             & $3\times10^5$        & $6\times10^4$     & $350$             \\
$3$h             & $3.75 \times 10^{4}$ & $7500$            & $175$             \\
$2$h             & $4690$               & $938$             & $87.5$            \\
$1$h             & $586$                & $117$             & $43.8$           
\end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

### 2019-12-06 paper/ms.tex
Update ms.tex
Updates to paper

@@
 \begin{abstract}
+Effective models
+
 Stars are known to be captured at the dynamical resonances of the Milky Way
@@
 \section{Introduction}
+Due to issues of large dynamic range, effective models of star formation is a
+crucial element of both high- and low-resolution numerical simulations of
+galaxy formation.
 

### 2019-12-20 plots/bar_angle/plot_pattern_speed.py
Add Elenas pattern speeds

@@ -6,6 +6,9 @@ import astropy.units as u
 
 name_list = ['lvl5', 'lvl4', 'lvl3', 'lvl5-rotbulge', 'lvl4-rotbulge']# 'lvl5-fg0.2']
 
+elena_time = [1500, 2500, 3000, 3500]
+elena_pattern_speed = [55, 40, 35, 30]
+
 n = 5
 
 fig, ax = plt.subplots(1, 1)
@@ -35,6 +38,8 @@ for name in name_list:
     l = ax.plot(time[firstkey:], pattern_speed[firstkey:], label=name)
     c = l[0].get_color()
     ax.scatter(time[firstkey:], finite_diff[firstkey:], c=c, alpha=0.2, s=1)
+
+ax.scatter(elena_time, elena_pattern_speed, s=9, c='k', label='galakos')
 
 ax.set_xlabel('time [Myr]')
 ax.set_ylabel('pattern speed [ km/s/kpc ]')

### 2019-12-24 plots/SFR/compute_SFR.py
Delete repository

@@ -1,39 +1,0 @@
-import sys
-mach = sys.platform
-if mach == 'linux':
-    sys.path.append('/n/home01/abeane/mwib_analysis')
-    base = '/n/scratchlfs/hernquist_lab/abeane/mwib_runs/arepo'
-elif mach == 'darwin':
-    sys.path.append('/Users/abeane/scratch/mwib_analysis')
-    base = '/Volumes/abeaneSSD001/mwib_runs/arepo'
-
-print(base)
-
-from mwib_analysis import mwib_io
-import numpy as np
-import h5py as h5
-from tqdm import tqdm
-import pickle
-
-import astropy.units as u 
-time_conv = (1 * u.kpc/(u.km/u.s)).to_value(u.Myr)
-
-sims_list = ['/galakos/lvl5', '/galakos/lvl4', '/galakos/lvl3-hernquist', '/galakos-fg0.2/lvl5']
-sims_list = [base+s for s in sims_list]
-name_list = ['lvl5', 'lvl4', 'lvl3', 'lvl5-fg0.2']
-
-skip = 10
-
-for sim, name in zip(sims_list, name_list):
-    indices, files = mwib_io.get_files_indices(sim+'/output/*.hdf5')
-
-    print(len(files))
-    print(sim+'/output/*.hdf5')
-
-    f = files[-1]
-    t = h5.File(f, mode='r')
-    star_birthtime = np.array(t['PartType4']['GFM_StellarFormationTime'])*time_conv
-    star_mini = np.array(t['PartType4']['GFM_InitialMass'])*1E10
-
-    pickle.dump((star_birthtime, star_mini), open('SFR_'+name+'.p', 'wb'))
-    

### 2019-12-24 plots/bar_angle/compute_bar_angle.py

@@ -1,89 +1,0 @@
-import sys
-mach = sys.platform
-if mach == 'linux':
-    sys.path.append('/n/home01/abeane/mwib_analysis')
-    base = '/n/scratchlfs/hernquist_lab/abeane/mwib_runs/arepo'
-elif mach == 'darwin':
-    sys.path.append('/Users/abeane/scratch/mwib_analysis')
-    base = '/Users/abeane/scratch/mwib_runs/arepo'
-
-from mwib_analysis import mwib_io
-import numpy as np
-import h5py as h5
-from tqdm import tqdm
-import pickle
-import re
-
-import astropy.units as u 
-time_conv = (1 * u.kpc/(u.km/u.s)).to_value(u.Myr)
-
-sims_list = ['/galakos/lvl5', '/galakos/lvl4', '/galakos/lvl3-hernquist',
-             '/galakos-rotbulge/lvl5', '/galakos-rotbulge/lvl4']
-sims_list = [base+s for s in sims_list]
-name_list = ['lvl5', 'lvl4', 'lvl3', 'lvl5-rotbulge', 'lvl4-rotbulge']
-
-Rbin = 5
-firstkey = 250
-
-def compute_bar_angle(phi, firstkey=400):
-    out = np.zeros(len(phi))
-
-    # set the first bar angle
-    first_bar_angle = phi[firstkey]/2.0
-    out[firstkey] = first_bar_angle
-    
-    # set all subsequent angles
-    for i in np.arange(firstkey+1, len(out)):
-        dphi = phi[i] - phi[i-1]
-        if dphi < 0:
-            dphi += 2.*np.pi
-        out[i] = out[i-1] + dphi/2.0
-
-    # set all previous angles to be the bar angle
-    for i in np.arange(0, firstkey):
-        out[i] = first_bar_angle
-
-    return out
-
-for sim, name in zip(sims_list, name_list):
-    out = {}
-    out['firstkey'] = firstkey
-    
-    #indices, files = mwib_io.get_files_indices(sim+'/output/*.hdf5')
-    fourier = h5.File(sim+'/analysis/fourier_components.hdf5', mode='r')
-
-    keys = list(fourier.keys())
-    keys = [k for k in keys if 'snapshot' in k]
-    indices = [int(re.findall(r'\d?\d?\d\d\d', k)[0]) for k in keys]
-    sorted_arg = np.argsort(indices)
[…]

### 2019-12-24 plots/bar_angle/plot_bar_angle.py

@@ -1,41 +1,0 @@
-import pickle
-import numpy as np
-from tqdm import tqdm
-import matplotlib.pyplot as plt
-import astropy.units as u
-
-name_list = ['lvl5', 'lvl4']#, 'lvl3', 'lvl5-fg0.2']
-
-n = 5
-
-fig, ax = plt.subplots(2, 1, figsize=(3, 6), sharex=True)
-
-for name in name_list:
-    out = pickle.load(open('bar_angle_'+name+'.p', 'rb'))
-
-    time = out['time']
-    polyfit = out['poly_fit']
-    firstkey = out['firstkey']
-
-    true_bar_angle = out['bar_angle']
-
-    bar_angle = np.zeros(len(time))
-
-    mypfit = polyfit[n]
-    for i in range(n):
-        bangle = mypfit[i] * time**(n-i) # taking the derivative of a general polynomial
-        bar_angle[firstkey:] += bangle[firstkey:] # set pattern speed to zero before firstkey
-
-    ax[0].plot(time, bar_angle, label=name)
-    ax[0].scatter(time, true_bar_angle, s=1, c='k')
-
-    ax[1].plot(time, np.diff(true_bar_angle, prepend=0), c='k')
-
-ax[1].set_xlabel('time [Myr]')
-ax[0].set_ylabel('bar angle')
-ax[1].set_ylabel('diff(bar angle)')
-
-fig.legend(frameon=False, title='resolution')
-
-fig.tight_layout()
-fig.savefig('bar_angle.pdf')

### 2019-12-24 plots/gas_fraction/compute_gas_fraction.py

@@ -1,47 +1,0 @@
-import sys
-mach = sys.platform
-if mach == 'linux':
-    sys.path.append('/n/home01/abeane/mwib_analysis')
-    base = '/n/scratchlfs/hernquist_lab/abeane/mwib_runs/arepo'
-elif mach == 'darwin':
-    sys.path.append('/Users/abeane/scratch/mwib_analysis')
-    base = '/Volumes/abeaneSSD001/mwib_runs/arepo'
-
-print(base)
-
-from mwib_analysis import mwib_io
-import numpy as np
-import h5py as h5
-from tqdm import tqdm
-import pickle
-
-sims_list = ['/galakos/lvl5', '/galakos/lvl4', '/galakos/lvl3-hernquist', '/galakos-fg0.2/lvl5']
-sims_list = [base+s for s in sims_list]
-name_list = ['lvl5', 'lvl4', 'lvl3', 'lvl5-fg0.2']
-
-skip = 10
-
-for sim, name in zip(sims_list, name_list):
-    indices, files = mwib_io.get_files_indices(sim+'/output/*.hdf5')
-    gas_fraction_list = []
-    time_list = []
-    for idx, f in zip(tqdm(indices[::skip]), files[::skip]):
-        time = mwib_io.get_time(f)
-        # dont need to subtract COM since not necessary
-        snap = mwib_io.read_snap(f, subtract_com=False)
-
-        gas_mass = np.sum(snap['PartType0']['Masses'])
-        disk_mass = np.sum(snap['PartType2']['Masses'])
-        try:
-            star_mass = np.sum(snap['PartType4']['Masses'])
-        except KeyError:
-            star_mass = 0.0
-        gas_fraction = gas_mass / (gas_mass + disk_mass + star_mass)
-
-        gas_fraction_list.append(gas_fraction)
-        time_list.append(time)
-
-    time_list = np.array(time_list)
-    gas_fraction_list = np.array(gas_fraction_list)
-    pickle.dump((time_list, gas_fraction_list), open('gas_fraction_'+name+'.p', 'wb'))
-    

### 2019-12-24 plots/movie_gas/make_movie_frames.py

@@ -1,226 +1,0 @@
-#from swiftsimio import load
-import sys
-mach = sys.platform
-if mach == 'darwin':
-    sys.path.append('/Users/abeane/scratch/mwib_analysis')
-
-# from mwib_analysis import mwib_io
-import h5py as h5
-from sphviewer.tools import QuickView
-
-from unyt import msun, kpc
-from tqdm import tqdm
-
-import pickle
-
-import numpy as np
-
-
-def grab_frame(filename, particle_type, res, fixed_hsml, center=None, extent=None, max_smooth=None):
-    # snap = mwib_io.read_snap(filename)
-    f = h5.File(filename, mode='r')
-    particle_group = {}
-    for k in ['Coordinates', 'Masses', 'Density']:
-        particle_group[k] = np.array(f[particle_type][k])
-
-    fh = f['Parameters'].attrs['GasSoftFactor']
-    volume = np.divide(particle_group['Masses'], particle_group['Density'])
-    particle_group['Smoothing_length'] = np.multiply(fh, np.power(np.multiply(volume, 3/(4.*np.pi)), 1./3.))
-
-    # convert masses to Msun
-    particle_group['Masses'] = np.multiply(particle_group['Masses'], 1E10)
-
-    if center is None:
-        center = [0, 0, 0]
-
-    # particle_group = snap[particle_type]
-    # particle_group = getattr(snap, particle_type)
-
-    try:
-        coordinates = particle_group['Coordinates']
-    except AttributeError:
-        # Dataset doesn't exist
-        print('Dataset doesnt exist')
-        return np.zeros((res, res), dtype=float)
-
-    masses = particle_group['Masses']
-
-    try:
-        # hsml = particle_group.smoothing_length.to(kpc).value
-        hsml = particle_group['Smoothing_length']
-        max_hsml = max_smooth
-    except:
-        print("Cannot find smoothing length; generating")
-        if fixed_hsml:
-            max_hsml = fixed_hsml
-            hsml = None
-        else:
-            hsml = None
-            max_hsml = None
-
-    out = QuickView(
-        coordinates,
-        mass=masses,
-        hsml=hsml,
-        r="infinity",
-        xsize=res,
-        ysize=res,
-        plot=False,
-        max_hsml=max_hsml,
-        x=center[0], y=center[1], z=center[2],
-        extent=extent
-    ).get_image()
-
-
-    return out
-
-
-def grab_all_frames(stub, start, stop, particle_type, res, fixed_hsml, center, extent, max_smooth):
-    return [
[…]

### 2019-12-24 plots/movie_gas/make_movie_video.py

@@ -1,85 +1,0 @@
-from tqdm import tqdm
-
-import pickle
-
-import numpy as np
-import matplotlib.pyplot as plt
-
-from matplotlib.animation import FuncAnimation
-
-
-def get_vmin_vmax(frames):
-    vmin = min(frames, key=lambda x: x.min()).min()
-    vmax = max(frames, key=lambda x: x.max()).max()
-
-    return vmin, vmax
-
-
-def get_animator(frames, vmin, vmax, cmap, size, res):
-    fig, ax = plt.subplots(figsize=(size, size), dpi=res//size)
-
-    fig.subplots_adjust(0, 0, 1, 1)
-    ax.axis("off")
-
-    image = ax.imshow(frames[0], cmap=cmap, origin="lower", vmin=vmin, vmax=vmax)
-
-    def animate(frame):
-        image.set_array(frames[frame])
-
-        return image
-    
-    anim = FuncAnimation(
-        fig,
-        animate,
-        tqdm(range(len(frames))),
-        fargs=[],
-        interval=1000 / 25
-    )
-
-    return anim
-
-if __name__ == "__main__":
-    import argparse as ap
-
-    parser = ap.ArgumentParser(
-        description="Makes the video of frames in the supplied pickle file."
-    )
-
-    parser.add_argument(
-        "-i",
-        "--input",
-        help="Name of the input pickle file",
-        required=True,
-        type=str,
-    )
-
-    parser.add_argument(
-        "-c",
-        "--cmap",
-        help="Colour map to use. Default: twilight.",
-        required=False,
-        type=str,
-        default="twilight"
-    )
-
-    parser.add_argument(
-        "-o",
-        "--output",
-        help="Output filename",
-        required=True,
-        type=str,
-    )
-
-    args = vars(parser.parse_args())
-
-    with open(args["input"], "rb") as handle:
-        frames = pickle.load(handle)
-
-    res = frames[0].shape[0]
-    
-    vmin, vmax = get_vmin_vmax(frames)
-
-    anim = get_animator(frames, vmin, vmax, cmap=args["cmap"], size=8, res=res)
-
-    anim.save(args["output"], dpi=res // 8)
-

### 2019-12-24 plots/spectrogram/compute_spectrogram.py

@@ -1,99 +1,0 @@
-# import numpy as np
-# import sys
-# import matplotlib.pyplot as plt
-# from tqdm import tqdm
-# import astropy.units as u
-
-# import matplotlib as mpl
-
-# sys.path.append('/Users/abeane/scratch/mwib_analysis')
-# from mwib_analysis import mwib_io, mwib_agama, disk
-
-# path_list = ['/Volumes/abeaneSSD001/mwib_runs/arepo/galakos/lvl4',
-#              '/Volumes/abeaneSSD001/mwib_runs/arepo/galakos-dry/lvl4']
-# name_list = ['SMUGGLE', 'Nbody']
-
-import sys
-mach = sys.platform
-if mach == 'linux':
-    sys.path.append('/n/home01/abeane/mwib_analysis')
-    base = '/n/scratchlfs/hernquist_lab/abeane/mwib_runs/arepo'
-elif mach == 'darwin':
-    sys.path.append('/Users/abeane/scratch/mwib_analysis')
-    base = '/Users/abeane/scratch/mwib_runs/arepo'
-
-print(base)
-
-from mwib_analysis import mwib_io, mwib_agama, disk
-import numpy as np
-import h5py as h5
-from tqdm import tqdm
-import pickle
-import astropy.units as u
-
-sims_list = ['/galakos/lvl5', '/galakos/lvl4', '/galakos/lvl3-hernquist', '/galakos-fg0.2/lvl5']
-path_list = [base+s for s in sims_list]
-name_list = ['lvl5', 'lvl4', 'lvl3', 'lvl5-fg0.2']
-
-dsnap2 = 100
-
-snaps = ['100', '300', '500', '700', '900', '1100']
-mlist = [2, 3, 4, 6, 8]
-
-idx_list_list = [np.arange(int(sp)-dsnap2, int(sp)+dsnap2) for sp in snaps]
-
-Rmin = 0.01
-Rmax = 20.
-nbins = 100
-
-xlim = [0, 20]
-
-for path, name in zip(path_list, name_list):
-    indices, files = mwib_io.get_files_indices(path + '/output/*.hdf5')
-
-    for snp, idx_list in zip(tqdm(snaps), idx_list_list):
-        for m in mlist:
-            to_dump = {}
-
-            key = int(snp)
-
-            f = files[key]
-
-            time = mwib_io.get_time(f)
-            time = "{:0.0f}".format(time)
-
-            # compute and plot the spectrogram
[…]

### 2019-12-26 paper/ms.tex
Update softening lengths
Update ms.tex
Update abstract

--- 2019-11-11 paper/ms.tex
@@
 
+\newcommand{\arepo}{\textsc{Arepo}}
+
 % Please keep new commands to a minimum, and use \newcommand not \def to avoid
@@
 \begin{abstract}
-Stars are known to be captured at the dynamical resonances of the Milky Way
-bar, forming moving groups that are seen in the solar neighborhood. In
-principle, gas can also be captured at these resonances which, if allowed to
-cool, might trigger star formation -- predicting an enhanced number of
-star-forming regions at these resonances. However, the efficiency of stellar
-and radiation feedback may suppress the star formation process.
+Long-term, secular processes can be important drivers of present-day galactic
+properties. The Milky Way is known to host one and perhaps multiple bars whose
+gravitational influence has observable consequences on both stellar and
+gaseous components. It has long been known that along the major axis of a bar
+gas can be driven towards the galactic center, fueling, e.g., the Central
+Molecular Zone and nuclear activity. On the other hand, stars are known to
+congregate near Lagrange points (i.e., minimums in the galactic potential),
+forming known moving groups such as Hercules. We report the results of
+isolated simulations incorporating both a state-of-the-art star formation and
+feedback model (in which a multi-phase interstellar medium is generated
+self-consistently) and a model of the Milky Way which forms a strong bar with
+a pattern speed matched to current observations.
+We are able to show, for the first time, that an enhancement of star formation may occur at the Lagrange points.
+We derive a collisional analogue of the theory of dynamical resonances and
 \end{abstract}
@@
 \section{Introduction}
+Introduction.
 
-This is a simple template for authors to write new MNRAS papers.
-See \texttt{mnras\_sample.tex} for a more complex example, and \texttt{mnras\_guide.tex}
-for a full user guide.
-
-All papers should start with an Introduction section, which sets the work
-in context, cites relevant earlier studies in the field by \citet{Others2013},
-and describes the problem the authors aim to solve \citep[e.g.][]{Author2012}.
-
-\section{Methods, Observations, Simulations etc.}
+\section{Methods}
+We perform our simulations using the moving-mesh code \arepo{} \citep{}
 
@@
                  & $[\,M_{\odot}\,]$    & $[\,M_{\odot}\,]$ & $[\,\text{pc}\,]$ \\
-$5$              & $2.4\times10^6$      & $4.8\times10^5$   & $700$             \\
-$4$              & $3\times10^5$        & $6\times10^4$     & $350$             \\
-$3$              & $3.75 \times 10^{4}$ & $7500$            & $175$             \\
-$2$              & $4690$               & $938$             & $87.5$            \\
-$1$              & $586$                & $117$             & $43.8$           
+$5$              & $2.4\times10^6$      & $4.8\times10^5$   & $80$             \\
+$4$              & $3\times10^5$        & $6\times10^4$     & $40$             \\
+$3$              & $3.75 \times 10^{4}$ & $7500$            & $20$             \\
+$2$              & $4690$               & $938$             & $10$            \\
+$1$              & $586$                & $117$             & $5$           
 \end{tabular}
@@
                  & $[\,M_{\odot}\,]$    & $[\,M_{\odot}\,]$ & $[\,\text{pc}\,]$ \\
-$5$h             & $2.4\times10^6$      & $4.8\times10^5$   & $700$             \\
-$4$h             & $3\times10^5$        & $6\times10^4$     & $350$             \\
-$3$h             & $3.75 \times 10^{4}$ & $7500$            & $175$             \\
-$2$h             & $4690$               & $938$             & $87.5$            \\
-$1$h             & $586$                & $117$             & $43.8$           
+$5$h             & $2.4\times10^6$      & $4.8\times10^5$   & $80$             \\
+$4$h             & $3\times10^5$        & $6\times10^4$     & $40$             \\
+$3$h             & $3.75 \times 10^{4}$ & $7500$            & $20$             \\
+$2$h             & $4690$               & $938$             & $10$            \\
+$1$h             & $586$                & $117$             & $5$           
 \end{tabular}

### 2019-12-27 paper/ms.tex
Add some more to methods

@@
 
+\newcommand{\Msun}{\ensuremath{\text{M}_{\odot}}}
+
 \newcommand{\beq}{\begin{equation}}
@@
 % Angus Beane,$^{1}$\thanks{E-mail: abeane@cfa.harvard.edu}
-Angus~Beane$^{1}$,\thanks{E-mail: abeane@cfa.harvard.edu}
+Angus~Beane$^{1}$,\thanks{E-mail: angus.beane@cfa.harvard.edu}
 Elena~D'Onghia$^{2,3}$,
@@
 \section{Methods}
-We perform our simulations using the moving-mesh code \arepo{} \citep{}
+\subsection{Numerical Methods and Physical Model}
+We perform our simulations using the moving-mesh code \arepo{}. Gravity is
+solved using a standard oct-tree algorithm \citep{1986Natur.324..446B}.
+Hydrodynamics is modelled with a finite volume solver over an unstructured
+Voronoi mesh allowed to move in a quasi-Lagrangian fashion
+\citep{2010MNRAS.401..791S, 2016MNRAS.455.1134P}. A concise summary of this
+core functionality of \arepo{} is available with its public
+release\footnote{\url{https://arepo-code.org}} \citep{2019arXiv190904667W}.
+
+In our simulations we will reach $\sim10^3\,\Msun$ mass resolution. `
 

### 2020-01-19 paper/ms.tex
Update methods section

@@
 \newcommand{\arepo}{\textsc{Arepo}}
+\newcommand{\smuggle}{\textsc{SMUGGLE}}
 
@@
 self-consistently) and a model of the Milky Way which forms a strong bar with
-a pattern speed matched to current observations.
-We are able to show, for the first time, that an enhancement of star formation may occur at the Lagrange points.
-We derive a collisional analogue of the theory of dynamical resonances and
+a pattern speed matched to current observations. We are able to show, for the
+first time, that an enhancement of star formation may occur at the Lagrange
+points. We derive a collisional analogue of the theory of dynamical resonances
+and
 \end{abstract}
@@
 
-In our simulations we will reach $\sim10^3\,\Msun$ mass resolution. `
+In our simulations we will reach $\sim10^3\,\Msun$ mass resolution. As a
+result, the cold phase of the interstellar medium (ISM) is well-resolved
+(\textcolor{red}{cite}), and an appropriate physical model which account for
+the processes occuring at these scales must be used. In this vein, we use the
+state-of-the-art model \smuggle{} \citep{2019MNRAS.489.4233M}. A detailed
+description of the model is presented in \citet{2019MNRAS.489.4233M}, but we
+summarize the salient components here.
+
+Standard cooling processes are accounted for using tabulated \textsc{cloudy}
+calculations, described in detail in \citet{2013MNRAS.436.3031V}. In addition
+to standard channels, metal line, fine-structure, and molecular cooling
+processes are also accounted for, allowing the gas to reach
+$\sim10\,\text{K}$, as well as self-shielding for high-density gas. Cosmic ray
+and photoelectric heating are also taken into account. Star formation follows
+a standard probabilistic approach \citep{2003MNRAS.339..289S}, with the star
+formation efficiency ($\epsilon$) set to $0.01$ and additional density and
+viriality conditions. Finally, stellar feedback is accounted for in three
+different channels: supernovae, young massive stellar radiation, and stellar
+winds from OB and AGB stars.
+
+The model used here is similar to the one used first by
+\citet{2011MNRAS.410.1391A, 2013ApJ...770...25A} and by the Feedback In
+Realistic Environments (FIRE) collaboration \citep{2011MNRAS.417..950H,
+2014MNRAS.445..581H, 2018MNRAS.480..800H}, with some key differences discussed
+in \citet{2019MNRAS.489.4233M}. One drawback of the \smuggle{} method is that
+radiation is not self-consistently tracked and evolved, and therefore
+radiation feedback must be implemented in an approximate manner. An extension
+of \smuggle{} which explicitly tracks the radiation field as an additional
+fluid component \citep{2019arXiv191014041K} based upon \areport{}
+\citep{2019MNRAS.485..117K} is available. We plan to use this module in the
+future.
+
+% Standard cooling processes from hydrogen and helium undergoing two-body
+% interactions, Compton cooling of cosmic microwave background photons, and
+% photoionization from a uniform UV background are accounted for using tabulated
+% \textsc{CLOUDY} calculations, described in detail in
+% \citet{2013MNRAS.436.3031V}. In addition to these processes, metal line,
+% fine-structure, and molecular cooling processes are accounted for, in order to
+% allow the gas to reach $\sim 10\,\text{K}$. At high densities, self-shielding
+% of the gas is accounted for using an analytic formula. Cosmic ray and
+% photoelectric heating are also taken into account, important for the thermal
+% stability of the cold and warm phases.
+
+% Star formation follows a standard probabilistic approach
+% \citep{2003MNRAS.339..289S}, where star particles are created and assumed to
+% be a coeval population following a \citet{2001ApJ...554.1274C} initial mass
+% function. Star formation efficiency ($\epsilon$) is assumed to be $0.01$. An
+% additional density threshold of $100\,\text{cm}^{-3}$ and a condition on a gas
+% cell to be sub-virial ($\alpha < 1$) and thus prone to gravitational collapse
+% are applied.
+
+% Finally, stellar feedback is accounted for in three different channels:
+% supernovae, young massive stellar radiation, and stellar winds from OB and AGB
+% stars.
 

### 2020-04-22 plots/SFR/plot_SFR.py
update pdf

@@ -1,0 +1,48 @@
+import matplotlib.pyplot as plt 
+import numpy as np 
+import pickle
+import astropy.units as u
+
+dt = 100 # Myr
+
+def plot_SFR(name_list, c_list, ls_list, fout, n=5):
+    fig, ax = plt.subplots(1, 1)
+
+    for name, c, ls in zip(name_list, c_list, ls_list):
+        fdata = 'data/SFR_' + name + '.p'
+        star_birthtime, star_mini = pickle.load(open(fdata, 'rb'))
+
+        bins = np.linspace(0, np.max(star_birthtime), dt)
+
+        digit = np.digitize(star_birthtime, bins)
+
+        t_list = np.array([star_birthtime[digit == i].mean() for i in range(1, len(bins))])
+        SFR_list, edges = np.histogram(star_birthtime, bins=bins, weights=star_mini)
+
+        SFR_list = SFR_list * (1/dt) * u.Msun/u.Myr
+        SFR_list = SFR_list.to_value(u.Msun/u.yr)
+
+        ax.plot(t_list, SFR_list, label=name)
+
+    ax.set_xlabel('t [Myr]')
+    ax.set_ylabel('SFR [Msun/yr]')
+    ax.legend(frameon=False)
+
+    ax.set_yscale('log')
+
+    fig.tight_layout()
+    fig.savefig(fout)
+
+if __name__ == '__main__':
+    name_list = ['fid-lvl4', 'fid-lvl3']#, 'nbody-lvl4', 'nbody-lvl3']
+    c_list = [None, None, None]
+    ls_list = [None, None, None]
+
+    plot_SFR(name_list, c_list, ls_list, 'SFR_fid.pdf')
+
+    name_list = ['fid-disp1.0-fg0.1-lvl5', 'fid-disp1.0-fg0.3-lvl5', 'fid-disp1.0-fg0.5-lvl5']
+    plot_SFR(name_list, c_list, ls_list, 'SFR_fid-fg-lvl5.pdf')
+    
+    name_list = ['fid-disp1.0-fg0.1-lvl4', 'fid-disp1.0-fg0.3-lvl4', 'fid-disp1.0-fg0.5-lvl4']
+    plot_SFR(name_list, c_list, ls_list, 'SFR_fid-fg-lvl4.pdf')
+

### 2020-04-22 plots/bar_angle/compute_bar_angle.py

@@ -1,0 +1,168 @@
+import numpy as np
+import arepo
+import sys
+from tqdm import tqdm
+import h5py as h5
+import pickle
+import re
+import astropy.units as u
+
+Rbin = 5
+firstkey = 150
+nmax = 10
+
+def get_bar_angle(phi, firstkey):
+    out = np.zeros(len(phi))
+
+    # set the first bar angle
+    first_bar_angle = phi[firstkey]/2.0
+    out[firstkey] = first_bar_angle
+    
+    # set all subsequent angles
+    for i in np.arange(firstkey+1, len(out)):
+        dphi = phi[i] - phi[i-1]
+        if dphi < 0:
+            dphi += 2.*np.pi
+        out[i] = out[i-1] + dphi/2.0
+
+    # set all previous angles to be the bar angle
+    for i in np.arange(0, firstkey):
+        out[i] = first_bar_angle
+
+    return out
+
+def get_sorted_keys(dat):
+    keys = list(dat.keys())
+    # only keep keys that are snapshot keys
+    keys = [k for k in keys if 'snapshot' in k]
+
+    # extract and sort indices
+    indices = [int(re.findall(r'\d?\d?\d\d\d', k)[0]) for k in keys]
+    sorted_arg = np.argsort(indices)
+    keys_sorted = [keys[i] for i in sorted_arg]
+
+    return keys_sorted
+
+def get_A2_angle(dat, keys):
+    Rlist = np.array([np.array(dat[k]['Rlist']) for k in keys])
+    A2r = np.array([np.array(dat[k]['A2r']) for k in keys])
+    A2i = np.array([np.array(dat[k]['A2i']) for k in keys])
+    A0 = np.array([np.array(dat[k]['A0']) for k in keys])
+
+    time_list = np.array(dat['time'])
+
+    phi = np.arctan2(A2i, A2r)
+    phi = phi[:,Rbin]
+    R_at_Rbin = Rlist[:,Rbin]
+
+    return time_list, R_at_Rbin, phi
+
+def evaluate_polynomial(pfit, n, time, bar_angle_firstkey, firstkey):
+    pfit_n = pfit[n]
[…]

### 2020-04-22 plots/fourier_component/job_fourier-lvl2.sh

@@ -1,0 +1,22 @@
+#!/bin/sh
+#SBATCH -p hernquist
+#SBATCH -J fourier 
+#SBATCH -n 4
+#SBATCH -N 1
+#SBATCH -o OUTPUT_frames.%j.out
+#SBATCH -e ERROR_frames.%j.err
+#SBATCH --exclusive
+#SBATCH --mail-user=angus.beane@cfa.harvard.edu
+#SBATCH --mail-type=BEGIN
+#SBATCH --mail-type=END
+#SBATCH --mail-type=FAIL
+#SBATCH --mem=190000
+##SBATCH -t 4-00:00           # Runtime in D-HH:MM
+#SBATCH -t 7-00:00           # Runtime in D-HH:MM
+
+source ../load-modules.sh
+
+ulimit -c unlimited
+
+python3 compute_fourier_component.py ${SLURM_NTASKS} lvl2
+

### 2020-04-22 plots/movie_5panel-fg/make_movie_5panel.py

@@ -1,0 +1,256 @@
+import matplotlib as mpl
+mpl.use('Agg')
+
+import glob
+import h5py as h5
+import numpy as np
+import matplotlib.pyplot as plt
+from tqdm import tqdm
+import sys
+import arepo
+import os
+
+import astropy.units as u
+
+from matplotlib.animation import FuncAnimation
+
+def make_movie(fout, sim, nsnap):
+
+    if 'nbody' in fout:
+        center = np.array([0, 0, 0])
+    else:
+        center = np.array([200, 200, 200])
+
+    snapbase = sim + '/output'
+    time_conv = 977.793 # converts time units to Myr
+    
+    indices = np.arange(nsnap)
+    
+    width = [30, 30, 10]
+    
+    gas_vmin = 10.**(0.0)
+    gas_vmax = 10.**(4.0)
+    #star_vmin = 10.**(-0.5)
+    #star_vmax = 10.**(3.5)
+    star_vmin = 10.**(0.0)
+    star_vmax = 10.**(4.0)
+    
+    nres = 256 
+    
+    res = [nres, nres, int(nres*width[2]/width[1])]
+    surf_xy = (width[0]/res[0])*(width[1]/res[1]) * (1000)**2
+    surf_xz = (width[0]/res[0])*(width[2]/res[2]) * (1000)**2
+    fps = 16
+    
+    range_xy = [[center[0]-width[0]/2.0, center[0]+width[0]/2.0], [center[1]-width[1]/2.0, center[1]+width[1]/2.0]]
+    range_xz = [[center[0]-width[0]/2.0, center[0]+width[0]/2.0], [center[2]-width[2]/2.0, center[2]+width[2]/2.0]]
+    
+    def animate_one_type(snap, pidx, im_xy, im_xz):
+   
+        if snap.NumPart_Total[pidx] == 0:
+            return (im_xy, im_xz)
+        
+        part = getattr(snap, 'part'+str(pidx))
+
+        xbool = np.logical_and(part.pos[:,0] < center[0]+width[0]/2.0, part.pos[:,0] > center[0]-width[0]/2.0)
+        ybool = np.logical_and(part.pos[:,1] < center[1]+width[1]/2.0, part.pos[:,1] > center[1]-width[1]/2.0)
+        zbool = np.logical_and(part.pos[:,2] < center[2]+width[2]/2.0, part.pos[:,2] > center[2]-width[2]/2.0)
[…]

### 2020-04-22 plots/movie_5panel/job_movie-lvl2.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#SBATCH -p hernquist 
+#SBATCH -J movie 
+#SBATCH -n 6
+#SBATCH -N 1
+#SBATCH -o OUTPUT_frames.%j.out
+#SBATCH -e ERROR_frames.%j.err
+##SBATCH --exclusive
+#SBATCH --mail-user=angus.beane@cfa.harvard.edu
+#SBATCH --mail-type=BEGIN
+#SBATCH --mail-type=END
+#SBATCH --mail-type=FAIL
+#SBATCH --mem-per-cpu=32000
+##SBATCH -t 4-00:00           # Runtime in D-HH:MM
+#SBATCH -t 7-00:00           # Runtime in D-HH:MM
+
+source ../load-modules.sh
+module load parallel
+
+ulimit -c unlimited
+
+python3 make_movie_5panel.py 1 nbody
+

### 2020-07-07 plots/movie_5panel/run_projections.py
Add new runs to run_projections.py

@@ -13,10 +13,16 @@ basepath = '../../runs/'
 fid_dP_sEQ_g1 = 'fid-dispPoly-fg0.1-SoftEQS0'
 fid_dP_rx_g1 = 'fid-dispPoly-fg0.1-MeshReg'
 fid_dP_rx_g1_dAGB = 'fid-dispPoly-fg0.1-MeshReg-diskAGB'
+fid_dP_rx_g1_dAGB_delay = 'fid-dispPoly-fg0.1-MeshReg-diskAGB-delay1.25'
+fid_dP_rx_g1_Rcore = 'fid-dispPoly-fg0.1-MeshReg-Rcore1.0'
+fid_dP_rx_g1_Rcore_AGB = 'fid-dispPoly-fg0.1-MeshReg-Rcore1.0-diskAGB'
+fid_dP_rx_g1_Rcore14_AGB = 'fid-dispPoly-fg0.1-MeshReg-Rcore1.4-diskAGB'
 
-pair_list = [(fid_dP_sEQ_g1, 'lvl5'), (fid_dP_sEQ_g1, 'lvl4'), (fid_dP_sEQ_g1, 'lvl3'),
-	  	     (fid_dP_rx_g1, 'lvl5'), (fid_dP_rx_g1, 'lvl4'), (fid_dP_rx_g1, 'lvl3'),
-             (fid_dP_rx_g1_dAGB, 'lvl3')]
+pair_list = [(fid_dP_sEQ_g1, 'lvl4'), (fid_dP_sEQ_g1, 'lvl3'),
+	  	     (fid_dP_rx_g1, 'lvl4'), (fid_dP_rx_g1, 'lvl3'),
+             (fid_dP_rx_g1_Rcore, 'lvl4'), (fid_dP_rx_g1_Rcore, 'lvl3'),
+             (fid_dP_rx_g1_Rcore_AGB, 'lvl4'), (fid_dP_rx_g1_Rcore_AGB, 'lvl3'),
+             (fid_dP_rx_g1_Rcore14_AGB, 'lvl4')]#, (fid_dP_rx_g1_Rcore14_AGB, 'lvl3')]
 
 name_list = [           p[0] + '-' + p[1] for p in pair_list]
 path_list = [basepath + p[0] + '/' + p[1] for p in pair_list]

### 2020-07-07 plots/movie_5panel/compute_projections.py
Improve debug info for projection comp

@@ -86,7 +86,7 @@ def construct_update_projection_hdf5(name, path, nproc=1, parttype=[0, 2, 3, 4],
                                      width=30., nres=256, output_dir='data/'):
 
     nsnap = len(glob.glob(path+'/output/snapdir*/*.0.hdf5'))
-    assert nsnap > 0
+    assert nsnap > 0,"No output files detected for name="+name
 
     fname = name + '_w' + "{:.01f}".format(width) + '_n' + str(nres) + '.hdf5' 
 

### 2020-07-07 plots/toomre/compute_toomre2.py
Update toomre plots
Includes now the toomre plots for cored profiles

@@ -125,7 +125,8 @@ if __name__ == '__main__':
 
     name_list = [           p[0] + '-' + p[1] for p in pair_list]
     path_list = [basepath + p[0] + '/' + p[1] for p in pair_list]
-                                            
-    for path, name in zip(path_list, name_list):
-        compute_two_comp_toomre(path, name)
+         
+    compute_two_comp_toomre('fid-fg0.1-Rcore1.0/', 'fid-fg0.1-Rcore1.0')
+    #for path, name in zip(path_list, name_list):
+        #compute_two_comp_toomre(path, name)
     

### 2020-07-07 plots/toomre/plot_toomre2.py

@@ -97,11 +97,17 @@ if __name__ == '__main__':
 
     fid_g1 = 'fid-disp1.0-fg0.1-lvl5'
     fid_d15_g1 = 'fid-disp1.5-fg0.1-lvl5'
-   
-    name_list = [fid_g1, fid_d15_g1]
+    fid_g1_rcore1 = 'fid-fg0.1-Rcore1.0'
+
+    name_list = [fid_g1, fid_d15_g1, fid_g1_rcore1]
 
     plot_toomre2(name_list, 'toomre_two_comp.pdf')
     plot_toomre_fR(fid_g1, 'toomre_fR.pdf')
     plot_toomre_Qm1(fid_g1, 'toomre_Qm1.pdf')
+    
+    plot_toomre_fR(fid_g1_rcore1, 'toomre_Rcore1.0_fR.pdf')
+    plot_toomre_Qm1(fid_g1_rcore1, 'toomre_Rcore1.0_Qm1.pdf')
+
+
     #plot_toomre_ratio(fid_g1, 'toomre_two_comp_ratio_'+fid_g1+'.pdf')
     

### 2020-07-07 paper/ms.tex
Add areport macro
Update ICs part of text

@@
 \newcommand{\arepo}{\textsc{Arepo}}
+\newcommand{\areport}{\textsc{Arepo-RT}}
 \newcommand{\smuggle}{\textsc{SMUGGLE}}
@@
 \section{Introduction}
-Introduction.
+Roughly two-thirds of spiral galaxies are known to host bars in the
+near-infrared, about half of which are strong
+\citep[e.g.][]{2000AJ....119..536E}.
 
@@
 future.
+
+\subsection{Initial Conditions}
+The starting point for the construction of our initial conditions is the
+N-body model GALAKOS \citep{2020ApJ...890..117D}. GALAKOS has structural
+parameters and a circular velocity curve which are in reasonable agreement
+with current data. However, GALAKOS is a particularly good starting point
+because as \citet{2020ApJ...890..117D} showed, the disk forms a strong bar
+after ~ 1 Gyr. The bar then grows, until at ~ 2.5 Gyr it has grown and slowed
+down to a pattern speed of ~40 km/s/kpc. Evidence is mounting that the Milky
+Way's bar is long and slow, just like that in GALAKOS.
+
+We convert a fraction (10\%) of the initial stellar disk of GALAKOS to a gas
+disk. As described in \citet{2005MNRAS.361..776S}, the gas disk is initially
+in hydrostatic equilibrium. We assumed an ideal EOS for the gas phase,
+consistent with our underlying physical model. We make two further
+modifications to this setup. First, we ``core'' the gas profile -- instead of
+assuming the gas surface density profile is exponential, we assume it is a
+constant value within some radius $R_{\text{core}}$. Specifically,
+\beq
+\Sigma_{\text{gas}} = \Sigma_{0,\text{gas}} \times \begin{cases} 
+      e^{-R_{\text{core}}/H} & R \leq R_{\text{core}} \\
+      e^{-R/H} & R > R_{\text{core}}\text{,}
+   \end{cases}
+\eeq
+where $H$ is the scale length of the disk and $\Sigma_{0,\text{gas}} =
+\frac{M_{\text{gas}}}{2\pi H^2}$ is the central surface density. When
+$R_{\text{core}}/H=1$ and $1.4$, $\sim10\%$ and $\sim20\%$ of the gas disk
+mass is removed, respectively. This mass is not returned to another component
+and is thus removed from the initial conditions entirely. However, since our
+assumed gas fraction is $10\%$, this only represents a total reduction in the
+disk mass by a few percent. The surface density profile of the stellar disk is
+unchanged, though its velocity distribution is suitably altered.
+
+The second modification we make is to treat the initial disk particles as star
+particles. In, e.g., \citet{2019MNRAS.489.4233M} the initial disk is assumed
+to be a purely collisionless component with no feedback contributions.
+Instead, we assume the initial disk particles are star particles so they may
+contribute gas through AGB channels. This requires assigning an age and
+metallicity to each star particle. We randomly draw the age and metallicity of
+each star according to the age-metallicity-radius relation of
+\citet{2018ApJ...865...96F}. This ignores vertical variations which is
+implemented in other models \citep[e.g.][]{2020arXiv200503646S}.
+
+Our goal is now to construct an initial disk galaxy with gas that, like
+GALAKOS, forms a strong bar. The underlying method GALAKOS uses (namely
+MakeNewDisk) can already convert a fraction of the initial stellar disk to a
+gas disk which is initially in hydrostatic equilibrium. However, if this setup
+is naive
 

### 2020-07-10 paper/ms.tex
More updates to paper

@@
 \newcommand{\Msun}{\ensuremath{\text{M}_{\odot}}}
+\newcommand{\K}{\ensuremath{\text{K}}}
 
@@
 near-infrared, about half of which are strong
-\citep[e.g.][]{2000AJ....119..536E}.
+\citep[e.g.][]{2000AJ....119..536E}. Given their ubiquity, it is no surprise
+that the Milky Way is known to host a strong bar. A full understanding of the
+origin of the Galaxy's structure clearly rests on the impact of the bar not
+just in the inner region but also near the Solar circle. Its impact on the gas
+phase is highly discussed in the inner regions of the Galaxy, but studies of
+the bar's impact on star formation at the Solar circle are less developed. In
+this work, we seek to contribute to our understanding of the latter topic.
 
@@
 
-\subsection{Initial Conditions}
+\subsection{Initial Conditions}\label{ssec:ics}
 The starting point for the construction of our initial conditions is the
@@
 
+\subsection{Gas Accretion}\label{ssec:gas_acc}
+The method described in Section~\ref{ssec:ics} will succesfully form a bar
+after a few Gyr. However, because the disk is isolated (i.e., the only gas
+present in the initial conditions is in the disk), there is no source of fresh
+gas to resupply the disk. We found in empirical tests that by the time the bar
+had fully formed, the disk had become significantly more quiescent than what
+is expected for the Milky Way, with a global SFR of <~0.5 Msun/yr and a SFR
+surface density at 8 kpc about two orders of magnitude less than expected.
+Such differences prohibit a faithful comparison to the Milky Way.
+
+As a result, it is necessary to include a source of fresh gas to resupply the
+galaxy. We found this to be quite problematic because several such techniques
+we explored would either weaken or completely destroy the bar. We found that
+it was necessary to include a somewhat \textit{ad hoc} implementation of gas
+accretion from the circumgalactic medium (CGM).
+
+Our method is inspired by the picture of a ``galactic fountain flow''
+\citep{2006MNRAS.366..449F, 2008MNRAS.386..935F}. In this picture, gas is
+launched from the central region of a spiral galaxy before falling back onto
+the disk at a larger radii. In these models, the gas has lower angular
+momentum than the disk when it rejoins the disk due to {\bf drag interactions
+with the hot corona.}
+
+We therefore have implemented a method to include ``hoses'' into AREPO. These
+hoses are circular regions of radius $R_h$ placed along the $x$-axis at $x=\pm
+R$. At a fixed timestep $\delta t$ each hose will spawn a number $N$ of gas
+particles. The spawned gas particles have a fixed mass, a uniform initial
+temperature of $10^4 \K$, and a uniform metallicity chosen to be solar.
+
+\subsection{Summary}\label{ssec:ics_summary}
+
 Our goal is now to construct an initial disk galaxy with gas that, like
 GALAKOS, forms a strong bar. The underlying method GALAKOS uses (namely
-MakeNewDisk) can already convert a fraction of the initial stellar disk to a
+MakeNewDisk) can already convert a constant fraction of the initial stellar disk to a
 gas disk which is initially in hydrostatic equilibrium. However, if this setup
-is naive
+is used na
 

### 2020-07-10 plots/bar_angle/compute_bar_angle.py
update pdf

@@ -1,0 +1,167 @@
+import numpy as np
+import arepo
+import sys
+from tqdm import tqdm
+import h5py as h5
+import pickle
+import re
+import astropy.units as u
+
+Rbin = 5
+firstkey = 150
+nmax = 10
+
+def get_bar_angle(phi, firstkey):
+    out = np.zeros(len(phi))
+
+    # set the first bar angle
+    first_bar_angle = phi[firstkey]/2.0
+    out[firstkey] = first_bar_angle
+    
+    # set all subsequent angles
+    for i in np.arange(firstkey+1, len(out)):
+        dphi = phi[i] - phi[i-1]
+        if dphi < 0:
+            dphi += 2.*np.pi
+        out[i] = out[i-1] + dphi/2.0
+
+    # set all previous angles to be the bar angle
+    for i in np.arange(0, firstkey):
+        out[i] = first_bar_angle
+
+    return out
+
+def get_sorted_keys(dat):
+    keys = list(dat.keys())
+    # only keep keys that are snapshot keys
+    keys = [k for k in keys if 'snapshot' in k]
+
+    # extract and sort indices
+    indices = [int(re.findall(r'\d?\d?\d\d\d', k)[0]) for k in keys]
+    sorted_arg = np.argsort(indices)
+    keys_sorted = [keys[i] for i in sorted_arg]
+
+    Rlist = np.array([np.array(fourier[k]['Rlist']) for k in tqdm(keys_sorted)])
+    A2r = np.array([np.array(fourier[k]['A2r']) for k in tqdm(keys_sorted)])
+    A2i = np.array([np.array(fourier[k]['A2i']) for k in tqdm(keys_sorted)])
+    A0 = np.array([np.array(fourier[k]['A0']) for k in tqdm(keys_sorted)])
+    return keys_sorted
+
+def get_A2_angle(dat, keys):
+    Rlist = np.array([np.array(dat[k]['Rlist']) for k in keys])
+    A2r = np.array([np.array(dat[k]['A2r']) for k in keys])
+    phi = np.arctan2(A2i, A2r)
+    phi = phi[:,Rbin]
+    R_at_Rbin = Rlist[:,Rbin]
[…]

### 2020-07-10 plots/fourier/plot_fourier_fn_time.py

@@ -1,0 +1,33 @@
+import pickle
+import numpy as np
+from tqdm import tqdm
+import matplotlib.pyplot as plt
+import astropy.units as u
+from scipy.ndimage import gaussian_filter1d
+
+name_list = ['lvl5', 'lvl4', 'lvl3', 'lvl5-rotbulge', 'lvl4-rotbulge']# 'lvl5-fg0.2']
+
+n = 5
+
+fig, ax = plt.subplots(1, 1)
+
+for name in name_list:
+    out = pickle.load(open('bar_angle_'+name+'.p', 'rb'))
+
+    time = out['time']
+    A2A0 = out['A2A0_max']
+    smoothed_A2A0 = gaussian_filter1d(A2A0, 5)
+
+    l = ax.plot(time, smoothed_A2A0, label=name)
+    # c = l[0].get_color()
+    # ax.scatter(time[firstkey:], finite_diff[firstkey:], c=c, alpha=0.2, s=1)
+
+ax.set_xlabel('time [Myr]')
+ax.set_ylabel('max(A2/A0)')
+
+ax.set_ylim([0, 1])
+
+fig.legend(frameon=False, title='resolution')
+
+fig.tight_layout()
+fig.savefig('fourier_fn_time.pdf')

### 2020-07-10 plots/fourier_component/job_fourier.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#SBATCH -p shared
+##SBATCH --constraint=intel
+#SBATCH -J fourier 
+#SBATCH -n 22 
+#SBATCH -N 1
+#SBATCH -o OUTPUT_frames.%j.out
+#SBATCH -e ERROR_frames.%j.err
+#SBATCH --exclusive
+#SBATCH --mail-user=angus.beane@cfa.harvard.edu
+#SBATCH --mail-type=BEGIN
+#SBATCH --mail-type=END
+#SBATCH --mail-type=FAIL
+#SBATCH --mem-per-cpu=3900
+##SBATCH -t 4-00:00           # Runtime in D-HH:MM
+#SBATCH -t 7-00:00           # Runtime in D-HH:MM
+
+source ../load-modules.sh
+
+ulimit -c unlimited
+
+python3 compute_fourier_component.py ${SLURM_NTASKS} $1
+

### 2020-07-10 plots/gas_fraction/compute_gas-fraction.py

@@ -1,0 +1,107 @@
+import numpy as np
+import arepo
+import sys
+from tqdm import tqdm
+import glob
+import pickle
+from joblib import Parallel, delayed
+
+def compute_gas_fraction(path, name, skip=50, output_dir='data/',
+                         Rmax=20, zmax=3, 
+                         R0=8.2, dR=1.0, Rcenter=6.0,
+                         gas_ptype=0, star_ptype=(2,3,4),
+                         center = np.array([200, 200, 200])):
+    # try loading snapshot
+    nsnap = len(glob.glob(path+'/output/snapdir*/*.0.hdf5'))
+
+    out = []
+    for snapnum in tqdm(np.arange(nsnap)[::skip]):
+        sn = arepo.Snapshot(path+'/output/', snapnum, combineFiles=True)
+        time = sn.header.time.as_unit(3.1536E13 * arepo.u.second).value # in Myr
+
+        gas_mass = get_mass_in_ptype(sn, gas_ptype, 0., Rmax, 0., zmax, center)
+        star_mass = get_mass_in_ptype(sn, star_ptype, 0., Rmax, 0., zmax, center)
+
+        gas_fraction_disk = gas_mass/(gas_mass + star_mass)
+
+        gas_mass = get_mass_in_ptype(sn, gas_ptype, R0-dR/2., R0+dR/2., 0., zmax, center)
+        star_mass = get_mass_in_ptype(sn, star_ptype, R0-dR/2., R0+dR/2., 0., zmax, center)
+
+        gas_fraction_R0 = gas_mass/(gas_mass + star_mass)
+        
+        gas_mass = get_mass_in_ptype(sn, gas_ptype, 0., Rcenter, 0., zmax, center)
+        star_mass = get_mass_in_ptype(sn, star_ptype, 0., Rcenter, 0., zmax, center)
+
+        gas_fraction_Rcenter = gas_mass/(gas_mass + star_mass)
+
+        out.append([time, gas_fraction_disk, gas_fraction_R0, gas_fraction_Rcenter])
+
+    pickle.dump(np.array(out), open(output_dir+'gas-fraction_'+name+'.p', 'wb'))
+
+def get_mass_in_ptype(snap, ptype, Rmin, Rmax, zmin, zmax, center):
+    if isinstance(ptype, int):
+        ptype = (ptype,)
+
[…]

### 2020-07-10 plots/gas_fraction/plot_gas-fraction.py

@@ -1,0 +1,106 @@
+import matplotlib.pyplot as plt 
+import numpy as np 
+import pickle
+import astropy.units as u
+
+time_conv = 977.793
+
+def plot_gas_fraction(pair_list, c_list, ls_list, fout, output_dir='data/', 
+                      xlim=[0, None], ylim=[None, None]):
+    name_list = [           p[0] + '-' + p[1] for p in pair_list]
+
+    fig, ax = plt.subplots(1, 1)
+
+    max_time = 0
+    for name, c, ls in zip(name_list, c_list, ls_list):
+        fdata = output_dir+'gas-fraction_'+name+'.p'
+        dat = pickle.load(open(fdata, 'rb'))
+
+        time = dat[:,0]
+        time = time - time[0]
+        gas_fraction_disk = dat[:,1]
+        gas_fraction_R0 = dat[:,2]
+        gas_fraction_Rcenter = dat[:,3]
+
+        l = ax.plot(time, gas_fraction_disk, label=name)
+        c = l[0].get_color()
+        ax.plot(time, gas_fraction_R0, ls='dashed', c=c)
+        ax.plot(time, gas_fraction_Rcenter, ls='dashdot', c=c)
+
+        if np.max(time) > max_time:
+            max_time = np.max(time)
+
+    ax.set_xlabel('t [Myr]')
+    ax.set_ylabel('gas fraction')
+    ax.legend(frameon=False)
+
+    # add band around MW values
+    # ax.axhline(0.1, c='k', alpha=0.5)
+    # ax.axhline(0.2, c='k', alpha=0.5)
+    ax.fill_between([0, max_time], 0.1, 0.2, color='k', alpha=0.3)
+
+
+    # ax.set_yscale('log')
+    ax.set_ylim(ylim)
+
+    fig.tight_layout()
+    fig.savefig(fout)
+
+if __name__ == '__main__':
+    nbody = 'fid-Nbody'
+    wet = 'fid-wet'
+    fid = 'fid'
+    fid_rdisk = 'fid-disp1.0-resetDisk'
+
[…]

### 2020-07-10 plots/movie_5panel/make_movie.py

@@ -1,0 +1,78 @@
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import numpy as np
+import h5py as h5
+from tqdm import tqdm
+
+from matplotlib.animation import FuncAnimation
+
+time_conv = 977.793 # converts time units to Myr
+
+class animate_maker(object):
+    def __init__(self, proj_hdf5, projection, parttype, nres, im, vmin, txt=None):
+        self.proj_hdf5 = proj_hdf5
+        self.projection = projection
+        self.parttype = parttype
+        self.nres = nres
+        self.base_key_list = ['PartType' + str(pt) + '/' + projection + '/snapshot_' for pt in self.parttype]
+        self.vmin = vmin
+
+        self.im = im
+        self.txt = txt
+
+    def __call__(self, frame):
+        heatmap = np.zeros((self.nres, self.nres))
+        for base_key in self.base_key_list:
+            this_ht = self.proj_hdf5[base_key + "{:03d}".format(frame)][:]
+            heatmap += this_ht
+
+        heatmap[heatmap < self.vmin] = self.vmin
+
+        self.im.set_data(heatmap.T)
+        if self.txt is None:
+            return (self.im,)
+        else:
+            time = self.proj_hdf5[self.base_key_list[0] + "{:03d}".format(frame)].attrs['Time']
+            self.txt.set_text('t='+'{0:.2f}'.format(time*time_conv)+' Myr')
+            return (self.im, self.txt)
+
+
+def make_movie(projection_file, parttype, projection, fout, plot_time=False, vmin=1E-3, vmax=10.**(0.5), fps=16):
+    assert projection in ['xy', 'xz', 'yz']
+
+    if isinstance(parttype, int):
+        parttype = [parttype]
+
+    f = h5.File(projection_file, mode='r')
+
+    width = f.attrs['width']
+    nres = f.attrs['nres']
+    maxsnap = f.attrs['maxsnap']
+
+    # initialize fig and ax, remove boundary
+    fig, ax = plt.subplots(1, figsize=(2, 2))
+    fig.subplots_adjust(0, 0, 1, 1)
+    ax.axis("off")
+
+    # initialize im
+    extent = [-width/2.0, width/2.0, -width/2.0, width/2.0]
[…]

### 2020-07-10 plots/movie_5panel/rsync_data.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+rsync -azvp --progress 'cannon:/n/home01/abeane/starbar-correcth/plots/movie_5panel/data/*' data/
+rsync -azvp --progress 'cannon:/n/home01/abeane/starbar-correcth/plots/movie_5panel/data-hose/*' data-hose/
+

### 2020-07-10 plots/movie_5panel/run_projections.py

@@ -1,0 +1,31 @@
+from compute_projections import construct_update_projection_hdf5
+import sys
+
+try:
+	nproc = int(sys.argv[1])
+	print('running with nproc='+str(nproc))
+except:
+	print('running in serial mode')
+	nproc=1
+
+basepath = '../../runs/'
+
+fid_dP_sEQ_g1 = 'fid-dispPoly-fg0.1-SoftEQS0'
+fid_dP_rx_g1 = 'fid-dispPoly-fg0.1-MeshReg'
+fid_dP_rx_g1_dAGB = 'fid-dispPoly-fg0.1-MeshReg-diskAGB'
+fid_dP_rx_g1_dAGB_delay = 'fid-dispPoly-fg0.1-MeshReg-diskAGB-delay1.25'
+fid_dP_rx_g1_Rcore = 'fid-dispPoly-fg0.1-MeshReg-Rcore1.0'
+fid_dP_rx_g1_Rcore_AGB = 'fid-dispPoly-fg0.1-MeshReg-Rcore1.0-diskAGB'
+fid_dP_rx_g1_Rcore14_AGB = 'fid-dispPoly-fg0.1-MeshReg-Rcore1.4-diskAGB'
+
+pair_list = [(fid_dP_sEQ_g1, 'lvl4'), (fid_dP_sEQ_g1, 'lvl3'),
+	  	     (fid_dP_rx_g1, 'lvl4'), (fid_dP_rx_g1, 'lvl3'),
+             (fid_dP_rx_g1_Rcore, 'lvl4'), (fid_dP_rx_g1_Rcore, 'lvl3'),
+             (fid_dP_rx_g1_Rcore_AGB, 'lvl4'), (fid_dP_rx_g1_Rcore_AGB, 'lvl3'),
+             (fid_dP_rx_g1_Rcore14_AGB, 'lvl4')]#, (fid_dP_rx_g1_Rcore14_AGB, 'lvl3')]
+
+name_list = [           p[0] + '-' + p[1] for p in pair_list]
+path_list = [basepath + p[0] + '/' + p[1] for p in pair_list]
+                                           
+for name, path in zip(name_list, path_list):
+    construct_update_projection_hdf5(name, path, nproc=nproc)

### 2020-07-10 plots/toomre/compute_toomre2.py

@@ -1,0 +1,132 @@
+import numpy as np
+import arepo
+import sys
+from tqdm import tqdm
+from scipy.optimize import minimize
+from numba import jit
+
+from joblib import Parallel, delayed
+
+m_p = 1.67262178e-24 # g
+k = 1.38065e-16 # cgs
+
+G = 43007.1
+
+nproc=6
+
+def compute_u_from_T(T):
+    # T in K
+    if T > 1e4:
+        mu = 10.0/17.0 #Assume fully ionised, ignore metals
+    else:
+        mu = 50.0/41.0 #Assume fully neutral, ignore metals
+
+    u = (3.0 * k * T) / (2.0 * mu * m_p) # cgs units
+    u /= 1e10 # code units
+    return u
+
+def _Q1_of_k_(k, kappa, sigma, dens):
+    return (kappa**2 + (k*sigma)**2) / (2. * np.pi * G * k * dens)
+
+@jit(nopython=True)
+def _Q2_of_k_(k, cs, kappa, sigmaR, sigmaStar, sigmaGas):
+    Qg = (kappa**2 + (k*cs)**2) / (2. * np.pi * G * k * sigmaGas)
+    Qs = (kappa**2 + (k*sigmaR)**2) / (2. * np.pi * G * k * sigmaStar)
+    return 1./(1./Qg + 1./Qs)
+
+def _toomre2_(kmin, kmax, cs, kappa, sigmaR, sigmaStar, sigmaGas, returnk=False):
+    fun_list = []
+    x_list = []
+    # print('cs=', cs, 'kappa=', kappa, 'sigmaR=', sigmaR, 'sigmaStar=', sigmaStar, 'sigmaGas=', sigmaGas)
+    for kguess in np.logspace(np.log10(kmin), np.log10(kmax), 10):
+        ans = minimize(_Q2_of_k_, kguess, (cs, kappa, sigmaR, sigmaStar, sigmaGas), bounds=((kmin, kmax),))
+        fun_list.append(ans.fun[0])
+        x_list.append(ans.x)
+
+    fun = np.min(fun_list)
+    x = x_list[np.argmin(fun)]
+
+    if returnk:
+        return fun, x
+    else:
+        return fun
+
+def _toomre_star_(kappa, sigmaR, sigmaStar):
+    return kappa * sigmaR / (3.36 * G * sigmaStar)
+
[…]

### 2022-04-04 paper/old/ms-old.tex
Add old folder to paper

--- 2020-01-19 paper/ms.tex
@@
 % Keep the title short and informative.
-\title[Star Bar]{On Star Formation at Dynamical Resonances of the Bar}
+\title[Bars slowing down]{The Milky Way bar does not have to be slowing down}
 
@@
 \begin{abstract}
+Most disk galaxies harbor a rotating bar-like stellar structure at their
+centers, of varying strengths and sizes. It has long been expected that the
+rate at which galactic bars rotate is slowing down over time due to resonant
+interactions between the bar and the outer disk or dark matter halo. On the
+other hand, the gravitational influence of the bar can drive gas to the center
+of the galaxy, providing a source of angular momentum for the bar, allowing it
+to speed up. Using state-of-the-art numerical simulations (which include a
+resolved cold phase of the interstellar medium) of an isolated disk galaxy, we
+show that the angular momentum increase from gas infall is enough to counter
+the angular momentum loss to the outer disk and halo. This effect does not
+occur in previous generations of feedback and ISM models which rely on wind
+particles to model large-scale feedback.
+
 Long-term, secular processes can be important drivers of present-day galactic
@@
 \section{Introduction}
-Introduction.
+Roughly two-thirds of spiral galaxies are known to host bars in the
+near-infrared, about half of which are strong
+\citep[e.g.][]{2000AJ....119..536E}. Given their ubiquity, it is no surprise
+that the Milky Way is known to host a strong bar. A full understanding of the
+origin of the Galaxy's structure clearly rests on the impact of the bar not
+just in the inner region but also near the Solar circle. Its impact on the gas
+phase is highly discussed in the inner regions of the Galaxy, but studies of
+the bar's impact on star formation at the Solar circle are less developed. In
+this work, we seek to contribute to our understanding of the latter topic.
 
@@
 future.
+
+\subsection{Initial Conditions}\label{ssec:ics}
+The starting point for the construction of our initial conditions is the
+N-body model GALAKOS \citep{2020ApJ...890..117D}. GALAKOS has structural
+parameters and a circular velocity curve which are in reasonable agreement
+with current data. However, GALAKOS is a particularly good starting point
+because as \citet{2020ApJ...890..117D} showed, the disk forms a strong bar
+after ~ 1 Gyr. The bar then grows, until at ~ 2.5 Gyr it has grown and slowed
+down to a pattern speed of ~40 km/s/kpc. Evidence is mounting that the Milky
+Way's bar is long and slow, just like that in GALAKOS.
+
+Including the gas phase was more problematic.
+
+\subsection{Gas Accretion}\label{ssec:gas_acc}
+The method described in Section~\ref{ssec:ics} will succesfully form a bar
+after a few Gyr. However, because the disk is isolated (i.e., the only gas
+present in the initial conditions is in the disk), there is no source of fresh
+gas to resupply the disk. We found in empirical tests that by the time the bar
+had fully formed, the disk had become significantly more quiescent than what
+is expected for the Milky Way, with a global SFR of <~0.5 Msun/yr and a SFR
+surface density at 8 kpc about two orders of magnitude less than expected.
+Such differences prohibit a faithful comparison to the Milky Way.
+
+\subsection{Summary}\label{ssec:ics_summary}
+
+Our goal is now to construct an initial disk galaxy with gas that, like
+GALAKOS, forms a strong bar. The underlying method GALAKOS uses (namely
+MakeNewDisk) can already convert a constant fraction of the initial stellar disk to a
+gas disk which is initially in hydrostatic equilibrium. However, if this setup
+is used na
 

### 2022-04-04 paper/ms.tex
Add class/style files

--- 2020-01-19 paper/ms.tex
@@
-% mnras_template.tex
-%
-% LaTeX template for creating an MNRAS paper
-%
-% v3.0 released 14 May 2015
-% (version numbers match those of mnras.cls)
-%
-% Copyright (C) Royal Astronomical Society 2015
-% Authors:
-% Keith T. Smith (Royal Astronomical Society)
+%\documentstyle[amssymb,12pt,draft,epsf,palatino]{nature-pvd}
+\documentclass{natureprintstyle}
+\bibliographystyle{naturemag}
 
-% Change log
-%
-% v3.0 May 2015
-%    Renamed to match the new package name
-%    Version number matches mnras.cls
-%    A few minor tweaks to wording
-% v1.0 September 2013
-%    Beta testing only - never publicly released
-%    First version: a simple (ish) template for creating an MNRAS paper
+\usepackage{aas_macros}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Basic setup. Most papers should leave these options alone.
-\documentclass[a4paper,fleqn,usenatbib]{mnras}
+\usepackage{epsfig,caption}
+\usepackage{color}
+\usepackage{bm}
+\usepackage{graphicx}
+\usepackage{longtable}
+% \usepackage{amssymb}
+\usepackage{rotating,xcolor}
+\usepackage{hyperref}
+% \usepackage{tgbonum}
 
-% MNRAS is set in Times font. If you don't have this installed (most LaTeX
-% installations will be fine) or prefer the old Computer Modern fonts, comment
-% out the following line
-\usepackage{newtxtext,newtxmath}
-% Depending on your LaTeX fonts installation, you might get better results with one of these:
-%\usepackage{mathptmx}
-%\usepackage{txfonts}
+\usepackage{fontspec}
+\setmainfont{texgyrepagella}[
+  Extension = .otf,
+  UprightFont = *-regular,
+  BoldFont = *-bold,
+  ItalicFont = *-italic,
+  BoldItalicFont = *-bolditalic,
+]
 
-% Use vector fonts, so it zooms properly in on-screen viewing software
-% Don't change these lines unless you know what you are doing
-\usepackage[T1]{fontenc}
-\usepackage{ae,aecompl}
+\usepackage[misc]{ifsym}
 
-%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
+\title{Stellar Bars in Isolated Gas-Rich Galaxies Do Not Slow Down}
 
-% Only include extra packages if you really need them. Common packages are:
-\usepackage{graphicx}	% Including figure files
-\usepackage{amsmath}	% Advanced maths commands
-\usepackage{amssymb}	% Extra maths symbols
+\author{Angus Beane$^{1*}$, Lars Hernquist$^1$, Elena D'Onghia$^2$, et al.}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{document}
 
-%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
-\newcommand{\pc}{\ensuremath{\text{pc}}}
-\newcommand{\kpc}{\ensuremath{\text{kpc}}}
+\maketitle
 
-\newcommand{\Msun}{\ensuremath{\text{M}_{\odot}}}
+\let\thefootnote\relax\footnote{
 
-\newcommand{\beq}{\begin{equation}}
-\newcommand{\eeq}{\end{equation}}
+\begin{affiliations}
+\item Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA
 
-\newcommand{\mwib}{\texttt{mwib}}
+\item Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson, AZ 85721, USA
 
-\newcommand{\arepo}{\textsc{Arepo}}
-\newcommand{\smuggle}{\textsc{SMUGGLE}}
+\item Lol
 
-% Please keep new commands to a minimum, and use \newcommand not \def to avoid
-% overwriting existing commands. Example:
-%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
+$^{*}$ \texttt{\mbox{angus.beane@cfa.harvard.edu}}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%
-
-% Title of the paper, and the short title which is used in the headers.
-% Keep the title short and informative.
-\title[Star Bar]{On Star Formation at Dynamical Resonances of the Bar}
-
-% The list of authors, and the short list which is used in the headers.
-% If you need two or more lines of authors, add an extra line using \newauthor
-\author[A. Beane et al.]{
-% Angus Beane,$^{1}$\thanks{E-mail: abeane@cfa.harvard.edu}
-Angus~Beane$^{1}$,\thanks{E-mail: angus.beane@cfa.harvard.edu}
-Elena~D'Onghia$^{2,3}$,
-Lars~Hernquist$^{1}$,
-and Federico~Marinacci$^{4}$
-\\
-% List of institutions
-% Center for Astrophysics {\normalfont |} Harvard \& Smithsonian, 60 Garden Street, Cambridge, MA 02138, USA
-$^{1}$Center for Astrophysics {\normalfont |} Harvard \& Smithsonian, 60 Garden Street, Cambridge, MA 02138, USA\\
-$^{2}$Department of Astronomy, University of Wisconsin, 475 North Charter Street, Madison, WI 53706, USA\\
-$^{3}$Center for Computational Astrophysics, Flatiron Institute, 162 5th Avenue, New York, NY 10010, USA\\
-$^{4}$Department of Physics \& Astronomy, University of Bologna, via Gobetti 93/2, 40129 Bologna, Italy
-% $^{3}$Department of Physics \& Astronomy, University of Pennsylvania, 209 South 33rd Street, Philadelphia, PA 19104, USA
+\end{affiliations}
 }
 
-% These dates will be filled out by the publisher
-\date{Accepted XXX. Received YYY; in original form ZZZ}
+\vspace{-3.5mm}
+\begin{abstract}
 
-% Enter the current year, for the copyright statements etc.
-\pubyear{2019}
+  Elongated bar-like features are ubiquitous, occuring at the centres of
+  approximately two-thirds of spiral disk galaxies \cite{2000AJ....119..536E,
+  2007ApJ...657..790M}. Due to gravitational interactions between the bar and
+  the other components of galaxies, it is expected that angular momentum and
+  matter will redistribute between galactic components over long (Gyr)
+  timescales in galaxies hosting a bar \cite{1972MNRAS.157....1L,
+  1984MNRAS.209..729T, 1985MNRAS.213..451W}. Previous work has overwhelmingly
+  provided the expectation that, due to these interactions, the bar pattern
+  will slow its rotation over time \cite{1992ApJ...400...80H,
+  2000ApJ...543..704D, 2002MNRAS.330...35A, 2002ApJ...569L..83A,
+  2003MNRAS.341.1179A, 2003MNRAS.346..251O, 2005MNRAS.363..991H,
+  2006ApJ...637..214M, 2007MNRAS.375..460W, 2009ApJ...697..293D}. We have
+  performed a simulation of an isolated galactic disk hosting a strong bar
+  which includes a state-of-the-art model of the interstellar medium. Here we
+  show that in this simulation the bar pattern does not slow down over time,
+  and instead remains at a stable, constant rate of rotation. This behavior
+  has been observed in previous simulations but its explanation has remained
+  elusive.\cite{1993AA...268...65F, 2010ApJ...719.1470V}. We propose an
+  equilibrium mechanism which is consistent with our simulations. This result
+  challenges our expectations for how barred galaxies redistribute angular
+  momentum and matter over Gyr timescales, which is especially relevant for
+  our understanding of how the Milky Way arrived at its present day state.
 
-% Don't change these lines
-\begin{document}
-\label{firstpage}
-\pagerange{\pageref{firstpage}--\pageref{lastpage}}
-\maketitle
-
-% Abstract of the paper
-\begin{abstract}
-Long-term, secular processes can be important drivers of present-day galactic
-properties. The Milky Way is known to host one and perhaps multiple bars whose
-gravitational influence has observable consequences on both stellar and
-gaseous components. It has long been known that along the major axis of a bar
-gas can be driven towards the galactic center, fueling, e.g., the Central
-Molecular Zone and nuclear activity. On the other hand, stars are known to
-congregate near Lagrange points (i.e., minimums in the galactic potential),
-forming known moving groups such as Hercules. We report the results of
-isolated simulations incorporating both a state-of-the-art star formation and
-feedback model (in which a multi-phase interstellar medium is generated
-self-consistently) and a model of the Milky Way which forms a strong bar with
-a pattern speed matched to current observations. We are able to show, for the
-first time, that an enhancement of star formation may occur at the Lagrange
-points. We derive a collisional analogue of the theory of dynamical resonances
-and
 \end{abstract}
 
-% Select between one and six entries from the list of approved keywords.
-% Don't make up new ones.
-\begin{keywords}
-Galaxy: disc -- Galaxy: kinematics and dynamics -- stars: kinematics and dynamics
-\end{keywords}
+\vspace{1cm}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
+\begin{figure*}[b!]%
+\centering
+\includegraphics[width=0.8\textwidth]{fig/fig1.pdf}
+\caption{Surface density projections in an N-body only simulation (upper
+panels) and a simulation which includes the SMUGGLE model (bottom panels). We
+can see that in the N-body run, the bar grows in length and strength. In the
+SMUGGLE run (which includes a gaseous phase and a model for the multi-phase
+interstellar medium), the bar remains at approximately the same length and
+strength over the course of the simulation. The N-body model is identical to
+the GALAKOS model, discussed in the text.}\label{fig:overview}
+\end{figure*}
 
-\section{Introduction}
-Introduction.
+We have performed a simulation of a disk galaxy using the finite-volume
+gravity-hydrodynamics code AREPO \cite{2010MNRAS.401..791S}. In AREPO, the
+fluid is discretized as a moving Voronoi mesh. We use the additional physics
+in the galaxy formation module Stars and MUltiphase Gas in GaLaxiEs (SMUGGLE)
+\cite{2019MNRAS.489.4233M}. SMUGGLE is a comprehensive and self-consistent
+galaxy formation model with a wide range of physical processes, including
+radiative heating/cooling, star formation, and stellar feedback. More detailed
+information on SMUGGLE is given in the Methods section.
 
-\section{Methods}
-\subsection{Numerical Methods and Physical Model}
-We perform our simulations using the moving-mesh code \arepo{}. Gravity is
-solved using a standard oct-tree algorithm \citep{1986Natur.324..446B}.
-Hydrodynamics is modelled with a finite volume solver over an unstructured
-Voronoi mesh allowed to move in a quasi-Lagrangian fashion
-\citep{2010MNRAS.401..791S, 2016MNRAS.455.1134P}. A concise summary of this
-core functionality of \arepo{} is available with its public
-release\footnote{\url{https://arepo-code.org}} \citep{2019arXiv190904667W}.
+Our disk galaxy is a modified version of the GALAKOS
+model\cite{2020ApJ...890..117D}, which consists of a radially exponential and
+vertically isothermal stellar disk along with a
+Hernquist\cite{1990ApJ...356..359H} dark matter halo and bulge. After
+$\sim2.5\,\textrm{Gyr}$ of evolution, the GALAKOS disk forms a bar consistent
+with the Milky Way bar in terms of pattern speed and length
+($\sim40\,\textrm{km}/\textrm{s}/\textrm{kpc}$ and $\sim4.5\,\textrm{kpc}$,
+respectively). We modify this setup by introducing a gas phase after $\sim
+1.5\,\textrm{Gyr}$ of evolution. This initial gas phase includes a hole in the
+center $4\,\textrm{kpc}$ and a constant surface density profile out to
+$9.3\,\textrm{kpc}$. When the gas is added to the simulation, the bar has
+already formed. More details on this configuration are given in the Methods
+section.
 
-In our simulations we will reach $\sim10^3\,\Msun$ mass resolution. As a
-result, the cold phase of the interstellar medium (ISM) is well-resolved
-(\textcolor{red}{cite}), and an appropriate physical model which account for
-the processes occuring at these scales must be used. In this vein, we use the
-state-of-the-art model \smuggle{} \citep{2019MNRAS.489.4233M}. A detailed
-description of the model is presented in \citet{2019MNRAS.489.4233M}, but we
-summarize the salient components here.
+A surface density projection of our simulatons is shown in
+Fig.~\ref{fig:overview}. The upper three panels show the disk in the N-body
+run while the lower three panels show the disk in the SMUGGLE run. Each column
+shows the disk $\sim1\,\textrm{Gyr}$ apart in time. There is a large qualitative
+difference in the evolution of the bar pattern between the two runs. We see
+that in the N-body case, the bar lengthens in time and grows in strength. In
+the SMUGGLE case, the bar pattern retains a similar length and strength
+between panels.
 
-Standard cooling processes are accounted for using tabulated \textsc{cloudy}
-calculations, described in detail in \citet{2013MNRAS.436.3031V}. In addition
-to standard channels, metal line, fine-structure, and molecular cooling
-processes are also accounted for, allowing the gas to reach
-$\sim10\,\text{K}$, as well as self-shielding for high-density gas. Cosmic ray
-and photoelectric heating are also taken into account. Star formation follows
-a standard probabilistic approach \citep{2003MNRAS.339..289S}, with the star
-formation efficiency ($\epsilon$) set to $0.01$ and additional density and
-viriality conditions. Finally, stellar feedback is accounted for in three
-different channels: supernovae, young massive stellar radiation, and stellar
-winds from OB and AGB stars.
+We show the time evolution of different bar properties in Fig.~\ref{fig:prop}.
+In the upper panel, we show the pattern speed over time in the N-body (blue)
+and SMUGGLE (orange) runs. The pattern speed in the N-body case slows down
+while the pattern speed in the SMUGGLE case remains constant. The slowing down
+of the pattern speed in the N-body case is consistent with a long line of
+numerical research on bars in N-body simulations\cite{1992ApJ...400...80H,
+2000ApJ...543..704D, 2002MNRAS.330...35A, 2002ApJ...569L..83A,
+2003MNRAS.341.1179A, 2003MNRAS.346..251O, 2005MNRAS.363..991H,
+2006ApJ...637..214M, 2007MNRAS.375..460W, 2009ApJ...697..293D}. However, in
+the SMUGGLE case the pattern speed remains constant. After the first Gyr of
+evolution, we find that the pattern speed increases by only $\sim10\%$ over
+the next $4\,\textrm{Gyr}$, compared to a $\sim43\%$ decrease in the pattern
+speed for the N-body run over the same interval. As we saw
+qualitatively in Fig.~\ref{fig:overview}, the length of the bar in the N-body
+case grows over time while it remains constant in the SMUGGLE case.
 
-The model used here is similar to the one used first by
-\citet{2011MNRAS.410.1391A, 2013ApJ...770...25A} and by the Feedback In
-Realistic Environments (FIRE) collaboration \citep{2011MNRAS.417..950H,
-2014MNRAS.445..581H, 2018MNRAS.480..800H}, with some key differences discussed
-in \citet{2019MNRAS.489.4233M}. One drawback of the \smuggle{} method is that
-radiation is not self-consistently tracked and evolved, and therefore
-radiation feedback must be implemented in an approximate manner. An extension
-of \smuggle{} which explicitly tracks the radiation field as an additional
-fluid component \citep{2019arXiv191014041K} based upon \areport{}
-\citep{2019MNRAS.485..117K} is available. We plan to use this module in the
-future.
-
-% Standard cooling processes from hydrogen and helium undergoing two-body
-% interactions, Compton cooling of cosmic microwave background photons, and
-% photoionization from a uniform UV background are accounted for using tabulated
-% \textsc{CLOUDY} calculations, described in detail in
-% \citet{2013MNRAS.436.3031V}. In addition to these processes, metal line,
-% fine-structure, and molecular cooling processes are accounted for, in order to
-% allow the gas to reach $\sim 10\,\text{K}$. At high densities, self-shielding
-% of the gas is accounted for using an analytic formula. Cosmic ray and
-% photoelectric heating are also taken into account, important for the thermal
-% stability of the cold and warm phases.
-
-% Star formation follows a standard probabilistic approach
-% \citep{2003MNRAS.339..289S}, where star particles are created and assumed to
-% be a coeval population following a \citet{2001ApJ...554.1274C} initial mass
-% function. Star formation efficiency ($\epsilon$) is assumed to be $0.01$. An
-% additional density threshold of $100\,\text{cm}^{-3}$ and a condition on a gas
-% cell to be sub-virial ($\alpha < 1$) and thus prone to gravitational collapse
-% are applied.
-
-% Finally, stellar feedback is accounted for in three different channels:
-% supernovae, young massive stellar radiation, and stellar winds from OB and AGB
-% stars.
-
-Normally the next section describes the techniques the authors used.
-It is frequently split into subsections, such as Section~\ref{sec:maths} below.
-
-\subsection{Maths}
-\label{sec:maths} % used for referring to this section from elsewhere
-
-Simple mathematics can be inserted into the flow of the text e.g. $2\times3=6$
-or $v=220$\,km\,s$^{-1}$, but more complicated expressions should be entered
-as a numbered equation:
-
-\begin{equation}
-    x=\frac{-b\pm\sqrt{b^2-4ac}}{2a}.
-	\label{eq:quadratic}
-\end{equation}
-
-Refer back to them as e.g. equation~(\ref{eq:quadratic}).
-
-\subsection{Figures and tables}
-
-Figures and tables should be placed at logical positions in the text. Don't
-worry about the exact layout, which will be handled by the publishers.
-
-Figures are referred to as e.g. Fig.~\ref{fig:example_figure}, and tables as
-e.g. Table~\ref{tab:example_table}.
-
-% Example figure
-\begin{figure}
-	% To include a figure from a file named example.*
-	% Allowable file formats are eps or ps if compiling using latex
-	% or pdf, png, jpg if compiling using pdflatex
-	% \includegraphics[width=\columnwidth]{example}
-    \caption{This is an example figure. Captions appear below each figure.
-	Give enough detail for the reader to understand what they're looking at,
-	but leave detailed discussion to the main body of the text.}
-    \label{fig:example_figure}
+The bottom panel of Fig.~\ref{fig:prop} shows the torque exerted on the bar by
+different components. The solid lines indicate the torque on the bar by the
+dark matter halo whereas the dashed line indicates the torque on the bar by
+the gas phase. In the N-body case, the halo exerts a steady negative torque on
+the bar, with an average torque from $1$ to $4\,\textrm{Gyr}$ of $-58.0$ in
+units of $10^{10}M_{\odot}\,(\textrm{km}/\textrm{s})^2$. The halo in the
+SMUGGLE case exerts a similar negative torque on the bar in the first Gyr of
+evolution, but after that the halo exerts a much smaller torque on the bar,
+averaging only $-7.8$ in the same units and over the same interval. The gas in
+the SMUGGLE case exerts a steady positive torque averaging $11.7$ in the same
+units and over the same interval.
+ß
+\begin{figure}[h]%
+\centering
+\includegraphics[width=0.45\textwidth]{fig/fig2.pdf}
+\caption{Various bar properties from the N-body and SMUGGLE runs. The
+\textit{upper panel} show the evolution of the pattern speed. As expected, the
+bar in the N-body run slows down due to interactions between the bar and the
+dark matter halo. However, the bar in the SMUGGLE run does not slow down and
+instead remains at a constant pattern speed. The \textit{middle panel} shows
+the evolution of the bar length. In the N-body case, the bar lengthens. This
+occurs because as the pattern speed drops, bar-like orbits at larger radii are
+possible. Stars are captured on these orbits, lengthening the bar. This
+process does not occur in the SMUGGLE cases since the bar pattern speed is not
+decreasing, and therefore the bar length remains constant. The \textit{lower
+panel} shows the torque on the bar by different components. The solid lines
+show the torque exerted by the halo in both the N-body and SMUGGLE cases. The
+dashed line shows the torque exerted by the gas phase in the SMUGGLE run
+(there is no gas in the N-body run). The first $1.5\,\textrm{Gyr}$ of
+evolution of the N-body model is not shown. Details on how these properties
+are calculated is given in the Methods section.}\label{fig:prop}
 \end{figure}
 
-% Example table
-\begin{table}
+The fact that the halo in the SMUGGLE case exerts a smaller positive torque on
+the bar can be understood in terms of the halo wake mechanism. In the N-body
+case, halo material which is resonant with the bar will form a wake which lags
+behind and exerts a negative torque on the bar, which slows it
+down.\cite{1984MNRAS.209..729T, 1985MNRAS.213..451W,
+1992ApJ...400...80H}\footnote{Since the bar is not a solid body, it is not
+guaranteed that a negative torque will slow it down - e.g. a negative torque
+could shred the bar, reducing its moment of inertia without changing its
+pattern speed. However, the bar seems to empirically respond to a negative
+torque induced by a halo wake by slowing down.} As the bar slows down, the
+location of the resonances in phase space changes, allowing halo material
+newly resonant with the bar to form a new wake. However, the gas is a reliable
+source of positive torque on the bar, speeding the bar up. This stops the
+resonance location from changing such that the halo cannot form a new wake,
+arresting the process by which the halo can slow the bar down.
+
+We can test this interpration by measuring the angle offset between the halo
+wake and the bar. If the wake and the bar are aligned (i.e., there is no angle
+offset), then the wake cannot exert a negative torque on the bar. This angle
+is plotted in the middle panel of Fig.~\ref{fig:wake}, which shows that the
+angle offset is larger in the N-body case than in the SMUGGLE case by about a
+factor of two. The left and right panels of Fig.~\ref{fig:wake} show the halo
+wake with respect to the location of the bar in the N-body (left) and SMUGGLE
+(right) cases.
+
+\begin{figure*}[h]%
 \centering
-	\caption{This is an example table. Captions appear above each table.
-	Remember to define the quantities, symbols and units used.}
-	\label{tab:example_table}
-	\begin{tabular}{lccr} % four columns, alignment for each
-		\hline
-		A & B & C & D\\
-		\hline
-		1 & 2 & 3 & 4\\
-		2 & 4 & 6 & 8\\
-		3 & 5 & 7 & 9\\
-		\hline
-	\end{tabular}
-\end{table}
+\includegraphics[width=0.9\textwidth]{fig/fig3.pdf}
+\caption{The wake excited in the dark matter halo in the N-body case
+(\textit{left panel}) and SMUGGLE case (\textit{right panel}) after
+$2.6\,\textrm{Gyr}$. The \textit{left} and \textit{right panels} show a
+surface density projection in the $x$-$y$ plane of the dark matter halo after
+an axisymmetric average has been subracted. The solid line indicates the
+direction of the bar while the dashed line indicates the direction of the halo
+wake (both measured by taking the second Fourier component within a sphere of
+all material within a radius of $X\,\textrm{kpc}$). The \textit{center panel}
+shows the time evolution of the angle difference between the bar and the halo
+wake, as measured from the second Fourier component. After the first Gyr, the
+angle difference in the SMUGGLE case is smaller than in the N-body case by
+about a factor of two, reflecting how the halo in the SMUGGLE case is unable
+to exert as negative a torque on the bar as in the N-body
+case.}\label{fig:wake}
+\end{figure*}
 
-\section{Conclusions}
+The presence of the gas can arrest the process by which the halo wake forms.
+However, this does not explain why the pattern speed in the SMUGGLE case is
+nearly constant over several Gyr. Naively, it would be a coincidence that the
+bar pattern speed remains constant in the SMUGGLE case, resulting from a
+chance cancellation of the halo and gas torques. However, a constant pattern
+speed in the presence of gas has been observed in a few simulations of barred
+galaxies with gas.\cite{1993AA...268...65F, 2010ApJ...719.1470V}
 
-The last numbered section should briefly summarise what has been done, and describe
-the final conclusions which the authors draw from their work.
+We propose that an equilibrium mechanism is responsible for the pattern speed
+remaining approximately constant. In this scenario, a torque must oppose
+changes in the pattern speed so that when the bar speeds up, a negative torque
+will slow it down and when the bar slows down, a positive torque will speed it
+up. It is simple to explain the first of these - when the pattern speed
+increases, the location of resonances will shift to regions of the halo phase
+space where no wake has been excited yet (e.g., the corotation radius will
+shrink).
 
-\section*{Acknowledgements}
+When the bar slows down, we argue that this induces a larger positive torque
+from the gas phase. Only gas within corotation will flow inwards, while gas
+outside corotation will flow outwards.\cite{2011MNRAS.415.1027H} Since the corotation radius is larger
+for more slowly rotating bars, it follows that more slowly rotating bars
+should be more efficient at driving gas inflows and thus experience a larger
+positive torque from the gas phase. We performed an experiment to test this
+hypothesis by forcing the stellar disk in the SMUGGLE run to rotate at a
+constant angular rate and measuring the torque on the bar by the gas phase at
+different rotation rates. The result of this experiment is shown in
+Fig.~\ref{fig:equil}, which shows that a more slowly rotating bar experiences
+a larger positive torque from the gas.
 
-The Acknowledgements section is not numbered. Here you can thank helpful
-colleagues, acknowledge funding agencies, telescopes and facilities used etc.
-Try to keep it short.
+\begin{figure}[h]
+\centering
+\includegraphics{fig/fig4.pdf}
+\caption{The torque on the bar by the gas when the pattern speed of the bar is
+kept at a constant value. Only gas within the corotation radius is able to
+infall. Since slower bars have larger corotation radii, slower bars experience
+a larger net torque than faster bars. The setup of the simulations used here
+is identical to the SMUGGLE case discussed earlier and in the Methods section,
+except the N-body disk is rotated as a solid body with a constant angular
+velocity.}\label{fig:equil}
+\end{figure}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+We have therefore shown evidence for an equilibrium mechanism at play which
+keeps the pattern speed of the bar constant, resulting from the complex
+interplay between the halo and the gas phase.
 
-%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%
+The implications of this finding are numerous. First, we naturally explain why
+nearly all observed galaxies are fast rotators without requiring the inner
+regions of dark matter halos to be underdense\cite{1998ApJ...493L...5D,
+2000ApJ...543..704D} or introducing new physics.\cite{2021MNRAS.503.2833R,
+2021MNRAS.508..926R} Second, we show that the role of gas is of paramount
+importance in studies which attempt to uncover the nature of dark matter from
+its effect of slowing down the bar.\cite{2021MNRAS.500.4710C,
+2021MNRAS.505.2412C} Third, we provide an explanation for how the Milky Way's
+bar could be both long-lived and a fast rotator, of which there is some
+observational evidence.\cite{2019MNRAS.490.4740B} And finally, we complicate
+the picture of radial mixing expected to sculpt the Milky Way's disk
+\cite{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies upon
+the pattern speed of the bar to change with time.
 
-% The best way to enter references is to use BibTeX:
+Barred galaxies in cosmological simulations of galaxy formation continue to be
+in conflict with observations.\cite{2017MNRAS.469.1054A, 2019MNRAS.483.2721P,
+2021AA...650L..16F} However, the pattern speeds of bars in both cosmological
+simulations and the real universe can be affected by environmental processes
+not included in our simulation -- e.g., satellite
+infall\cite{2011Natur.477..301P}, non-sphericity\cite{2013MNRAS.429.1949A} or
+rotation\cite{2013MNRAS.434.1287S, 2014ApJ...783L..18L, 2018MNRAS.476.1331C,
+2019MNRAS.488.5788C} in the dark matter halo, or perhaps even the gaseous
+circumgalactic medium. Naturally, extending our present work to account for
+such effects is a crucial next step in understanding the formation and
+evolution of galactic bars.
 
-\bibliographystyle{mnras}
-\bibliography{references} % if your bibtex file is called example.bib
+% \begin{figure*}
+% \includegraphics[width=0.5\textwidth]{fig1a.pdf}
+% \includegraphics[width=0.5\textwidth]{fig1b.pdf}
+% \vspace{-4mm}
+% \caption{\textbf{Distribution of stars in the Galactic halo.}  All-sky
+%   Mollweide projection maps of the density of stars at 60 kpc
+%   $<R_{\rm gal}<100$ kpc, smoothed by a Gaussian kernel with
+%   FWHM$ =30^\circ$.  {\bf Left panel:} data based on K giant stars.
+%   {\bf Right panel:} simulation that includes the dynamical response
+%   of the halo to the orbit of the LMC.  Regions near the Galactic
+%   mid-plane ($|b|<10^\circ$) and the LMC and SMC, as well as regions
+%   of high extinction are masked in grey, while the positions of the
+%   LMC and SMC are marked as white stars.  One arm of the Sagittarius
+%   stream has been masked in sky coordinates, as indicated by the grey
+%   region at $b>0^\circ$ and $-180^\circ<l<-150^\circ$.  The
+%   overdensity in the south-west is the local wake caused by the
+%   passage of the LMC.  The overdensity in the North is the collective
+%   response of the Galaxy due to the arrival of the LMC.  Notice that
+%   the dynamic range of the density contrast for the simulated map is
+%   smaller than the observed map.}
+% \vspace{-4mm}
+% \end{figure*}
 
-\appendix
+% \begin{figure}[t!]
+%   \center
+%   \includegraphics[width=0.48\textwidth]{fig2.pdf}
+%   \caption{\textbf{Quantitative comparison between data and models.}
+%     Mean density contrasts in regions encompassing the local wake
+%     ($-60^\circ<b<-20^\circ$, $45^\circ<l<90^\circ$) and the
+%     collective response ($30^\circ<b<60^\circ$,
+%     $240^\circ<l<300^\circ$) are shown along the $x$ and $y$ axes.
+%     There are four simulations spanning a range in LMC masses from
+%     $0.8-2.5\times10^{11}M_\odot$ (small-to-large green circles).  The
+%     smooth model\cite{Rybizki18} is close to zero along both axes as
+%     expected.  The dotted lines indicate uniform distributions (no
+%     density contrast) on the x and y axes.  The density enhancement of
+%     the local wake is measured at a significance of $5\sigma$ (error
+%     bars on the data represent $1\sigma$ Poisson uncertainties) and
+%     the mean density is $1.4\pm0.2$ times larger in the data than in
+%     the fiducial simulation.}
+% \end{figure}
 
-\section{Two-component Toomre instability criterion}
-The Toomre instability criterion for a two-component fluid was first derived
-by \citet{1984ApJ...276..114J}. For a mode of wavenumber $k=2\pi/\lambda$ to
-be stable against gravitational collapse, the criterion for the two fluids in
-an infinitesimally thin disk is that,
-\beq
-Q_2(k) = \left(Q_g^{-1}(k) + Q_s^{-1}(k) \right)^{-1} > 1\text{,}
-\eeq
-where $Q_2(k)$ is the two-component Toomre parameter and,
-\beq
-Q_g(k) = \frac{\kappa^2 + k^2 c_s^2}{2\pi G k \Sigma_{g,0}}\text{,}
-\eeq
-where $\kappa$ is the radial epicyclic frequency, $c_s$ is the sound speed of
-the gas, and $\Sigma_{g,0}$ is the surface density of the gas. $Q_s(k)$ can be
-obtained by replacing $c_s$ by $\sigma_R$ (the radial velocity dispersion) and
-$\Sigma_{g,0}$ by $\Sigma_{s,0}$ (the stellar surface density). Note that the
-familiar one-component Toomre criterion can be obtained by minimizing $Q_g(k)$
-as a function of $k$.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-We denote the minimum of $Q_2(k)$ as a function of $k$ by $Q_2$. We consider
-only physically plausible values of $k$ between $0.06$ and $6000\,\kpc^{-1}$
-(corresponding to modes of wavelength $\lambda\sim100\,\kpc$ to $1\,\pc$). The
-equivalent of the one-component Toomre stability criterion is that $Q_2>1$,
-such that the disk is stable against collapse of modes of all wavelengths.
+%\bibliographystyle{naturemag}
+\bibliography{ref}
 
-\section{\mwib{} resolution levels}
-We set a series of standard resolution levels similar to those in the Aquarius
-convention. Due to the isolated nature of our simulations, it is convenient to
-tune the particle mass to a specific number.
+\begin{addendum}
 
-\begin{table}
-\begin{tabular}{cccc}
-resolution level & $m_{\text{DM}}$      & $m_{\text{b}}$    & $\epsilon$        \\
-                 & $[\,M_{\odot}\,]$    & $[\,M_{\odot}\,]$ & $[\,\text{pc}\,]$ \\
-$5$              & $2.4\times10^6$      & $4.8\times10^5$   & $80$             \\
-$4$              & $3\times10^5$        & $6\times10^4$     & $40$             \\
-$3$              & $3.75 \times 10^{4}$ & $7500$            & $20$             \\
-$2$              & $4690$               & $938$             & $10$            \\
-$1$              & $586$                & $117$             & $5$           
-\end{tabular}
-\end{table}
+\item [Acknowledgements] We thank both referees for their constructive
+  feedback.  C.C. is partially supported by the Packard
+  Foundation.  R.P.N. gratefully acknowledges an Ashford Fellowship
+  and Peirce Fellowship granted by Harvard University.  G.B.  and
+  N.G.-C. are supported by HST grant AR 15004, NASA ATP grant
+  17-ATP17-0006, NSF CAREER AST-1941096.  A.B. acknowledges support
+  from NASA through HST grant HST-GO-15930.  All the simulations were
+  run on El-Gato super computer which was supported by the National
+  Science Foundation under Grant No. 1228509.  We have made use of
+  data from the European Space Agency mission Gaia
+  (http://www.cosmos.esa.int/gaia), processed by the Gaia Data
+  Processing and Analysis Consortium (DPAC; see
+  http://www.cosmos.esa.int/web/gaia/dpac/consortium). Funding for
+  DPAC has been provided by national institutions, in particular the
+  institutions participating in the Gaia Multilateral Agreement.  This
+  publication makes use of data products from the {\it Wide-field
+    Infrared Survey Explorer}, which is a joint project of the
+  University of California, Los Angeles, and the Jet Propulsion
+  Laboratory/California Institute of Technology, and NEOWISE, which is
+  a project of the Jet Propulsion Laboratory/ California Institute of
+  Technology.  {\it WISE} and NEOWISE are funded by the National
+  Aeronautics and Space Administration.
 
-\begin{table}
-\begin{tabular}{cccc}
-resolution level & $m_{\text{DM}}$      & $m_{\text{b}}$    & $\epsilon$        \\
-                 & $[\,M_{\odot}\,]$    & $[\,M_{\odot}\,]$ & $[\,\text{pc}\,]$ \\
-$5$h             & $2.4\times10^6$      & $4.8\times10^5$   & $80$             \\
-$4$h             & $3\times10^5$        & $6\times10^4$     & $40$             \\
-$3$h             & $3.75 \times 10^{4}$ & $7500$            & $20$             \\
-$2$h             & $4690$               & $938$             & $10$            \\
-$1$h             & $586$                & $117$             & $5$           
-\end{tabular}
-\end{table}
+\item[Author Contributions] C.C. and R.P.N. jointly conceived of the
+  project.  C.C. led the analysis of the data.  N.G-C. and
+  G.B. provided the simulation data and aided in its interpretation.
+  All authors contributed to aspects of the analysis and to the
+  writing of the manuscript.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\item[Author  Information]  Reprints and permissions
+    information is available at npg.nature.com/reprintsandpermissions.
+    Correspondence and requests for materials should be addressed to
+    C.C.\ (cconroy@cfa.harvard.edu).
 
-% Don't change these lines
-\bsp	% typesetting comment
-\label{lastpage}
+  \item[Data Availability] The K giant catalog used in this paper is
+    available at \texttt{https://doi.org/10.7910/DVN/2D1H8J}.
+
+  \item[Code Availability] We have opted not to make the code used in
+    this manuscript available because the data reduction and analysis
+    is straightforward and can be easily reproduced following the
+    methods described herein.
+
+\end{addendum}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newpage
+
+\setcounter{page}{1}
+\setcounter{figure}{0}
+\setcounter{table}{0}
+\captionsetup[figure]{labelformat=empty}
+\renewcommand{\thefigure}{Extended Data \arabic{figure}}
+\renewcommand{\thetable}{Extended Data \arabic{table}}
+
+\begin{center}
+{\bf \Large \uppercase{Methods} }
+\end{center}
+
+\noindent
+{\bf SMUGGLE Model}
+\\
+\noindent
+We use the Stars and MUltiphase Gas in GaLaxiEs (SMUGGLE) model
+\cite{2019MNRAS.489.4233M} implemented within the moving-mesh, finite-volume
+hydrodynamics code AREPO \cite{2010MNRAS.401..791S}. The SMUGGLE model
+includes self-gravity, hydrodynamics, radiative heating and cooling, star
+formation, and stellar feedback. Explicit gas cooling and heating of the
+multi-phase interstellar medium is implemented, covering temperature ranges
+between $10$ and $10^8\,\textrm{K}$.
+
+Star formation occurs in cells above a density threshold
+($n_{\textrm{th}}=100\,\textrm{cm}^{-3}$) according to SH03 with a
+star-formation efficiency of $\epsilon = 0.01$. Star formation converts gas
+cells into star particles which represent single stellar populations. For each
+star particle, the deposition of energy, momentum, and mass from stellar winds
+and supernovae is modeled. Photoionization and radiation pressure are modeled
+using an approximate treatment. A more detailed description of this model can
+be found in the flagship SMUGGLE paper.\cite{2019MNRAS.489.4233M}
+
+\vspace{12pt}
+
+\noindent
+{\bf Initial Setup}
+\\
+\noindent
+The initial setup of the galactic disk used in this work follows closely the
+GALAKOS model\cite{2020ApJ...890..117D}, which uses a modified version of
+\texttt{MakeNewDisk}.\cite{2005MNRAS.361..776S} The GALAKOS model has three
+components - a radially exponential and vertically isothermal stellar disk,
+and a stellar bulge and dark matter halo following a Hernquist
+profile.\cite{1990ApJ...356..359H} All N-body runs in this work used the same
+setup parameters as the GALAKOS disk, more details of which can be found in
+the original paper.
+
+The addition of the gas phase was done as follows. The version of
+\texttt{MakeNewDisk} used for the original GALAKOS model can generate a gas
+disk which is radially exponential and in vertical gravito-hydrodynamic
+balance. We modified the radial profile of this code in order to allow us to
+generate a disk with a constant surface density within some radius, and then
+exponentially declining beyond that radius with the scale-length of the
+stellar disk. We used a surface density of $X$ and a maximum radius of $Y$. We
+varied the surface density in TEST.
+
+After generating the gaseous disk in this way, we stitched the gas disk
+together with the GALAKOS N-body disk (and bulge and halo) after the GALAKOS
+disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The purpose of allowing
+the GALAKOS to evolve first for a short period of time is to allow for the bar
+to form unimpacted by the presence of the gas (which would normally disrupt
+the formation of the bar). We made one additional modification when stitching
+the gas disk together with the N-body disk - we created a vacuum within the
+central $X$ pc. This vacuum guards against an initial dramatic infall of gas
+within the bar region, which we found to destroy the bar.
+
+Our setup is initially out of equilibrium, but we found that after
+$\sim300\,\textrm{Myr}$, the entire system has settled into a steady-state
+configuration and initial transients appear not to affect the results after
+this point. The constant surface density of the initial gas disk is important
+for ensuring the gas disk is dense enough in order for comparisons to real
+galaxies to be appropriate. We show the evolution of the surface density
+profile in Fig.~X compared to observational estimates (CITE).
+
+\vspace{12pt}
+
+\noindent
+{\bf Title}
+\\
+\noindent
+Next section.
+
+% \begin{figure}[t!]
+%   \includegraphics[width=0.45\textwidth]{ed_fig1.pdf}
+%   \vspace{-1cm}
+%   \caption{\textbf{Extended Data Figure 1 $|$ Photometric selection of
+%       giants.}  Color-color diagram for stars at high Galactic
+%     latitude ($b>45^\circ$) with {\it Gaia} parallax $\pi<0.1$ mas.
+%     Two sequences are clearly visible, with the upper branch being
+%     associated with giant stars, and the lower with dwarf stars. The
+%     red lines mark the selection boundary used in this work.}
+% \end{figure}
+
+% \begin{figure}[t!]
+%   \includegraphics[width=0.45\textwidth]{ed_fig2.pdf}
+%   \caption{\textbf{Extended Data Figure 2 $|$ Test of Photometric
+%       Distances.}  Comparison of our photometric distances against
+%     literature values for satellite galaxies around the Milky Way
+%     (red) and globular clusters (blue).  Error bars represent the
+%     $1\sigma$ scatter in the photometric distances and dotted lines
+%     mark $\pm20$\% about the one-to-one line.  The dwarf galaxies
+%     include Fornax, Draco, Sculptor, Carina, Ursa Minor, and the LMC
+%     and SMC, and span average metallicities from [Fe/H]$\approx-0.5$
+%     (LMC) to $-2.2$ (Ursa Minor)\cite{McConnachie12}.}
+% \end{figure}
+
+% \begin{figure*}[t!]
+%   \includegraphics[width=0.5\textwidth]{ed_fig3a.pdf}
+%   \includegraphics[width=0.5\textwidth]{ed_fig3b.pdf}
+%   \caption{\textbf{Extended Data Figure 3 $|$ Maps of un-smoothed star
+%       counts.}  Mollweide projection maps of the observed sample of
+%     giants with pixels of size 13.4 sq. deg. {\bf Left panel:} map
+%     with no masking of known structure, either in sky coordinates or
+%     proper motions.  The two main features identified here, the
+%     transient wake and the collective response, are still clearly
+%     visible even in this unfiltered map.  The LMC and SMC appear in
+%     the lower right as a merged region of high density.  Other
+%     features include the stellar disk+bulge in the center, and the
+%     Sagittarius stream both in the north-center and lower left and
+%     right.  {\bf Right panel:} map showing the filtered catalog used
+%     in Fig 1a.}
+% \end{figure*}
+
+% \begin{figure*}[t!]
+%   \center
+%   \includegraphics[width=0.8\textwidth]{ed_fig4.pdf}
+%   \caption{ \textbf{Extended Data Figure 4 $|$ Proper motions of the
+%       halo sample.}  {\bf a:} proper motions of the K giants at
+%     $60<R_{\rm gal}<100$ kpc with $12<W1<15$ and the LMC and SMC
+%     removed via on-sky selection.  {\bf b:} solar reflex-corrected
+%     proper motions.  Notice the much tighter distribution of stars
+%     near (0,0).  In this panel the Sagittarius dSph is visible at
+%     (-2.5,-0.7), LMC and SMC stars not removed by the on-sky selection
+%     are visible as a narrow vertical vertical strip at
+%     $\mu_\alpha\gtrsim0.5$ mas yr$^{-1}$.  The Northern Sagittarius
+%     arm is the large overdensity at (-0.7,0.0), and a Southern arm of
+%     Sagittarius is the diagonal cluster of points at $\mu_\delta<-0.5$
+%     mas yr$^{-1}$.  {\bf c:} reflex-corrected proper motions focusing
+%     on the region of the sky containing the Northern Sagittarius arm
+%     at $l\sim0^\circ$.  The red box indicates our selection for
+%     removing this feature.  {\bf d:} stars not in the selection shown
+%     in panel {\bf c}. Our selection for low proper motion stars is
+%     indicated by a blue circle in {\bf c, d}.}
+% \end{figure*}
+
+% \bibliography{ref2}
+
 \end{document}

### 2022-04-04 paper/ps_enq/sn-article-template/sn-article.tex
Add ps_enq

\documentclass[sn-mathphys]{sn-jnl}% Math and Physical Sciences Reference Style
%%\documentclass[sn-aps]{sn-jnl}% American Physical Society (APS) Reference Style
%%\documentclass[sn-vancouver]{sn-jnl}% Vancouver Reference Style
%%\documentclass[sn-apa]{sn-jnl}% APA Reference Style
%%\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
%%\documentclass[sn-standardnature]{sn-jnl}% Standard Nature Portfolio Reference Style
%%\documentclass[default]{sn-jnl}% Default
%%\documentclass[default,iicol]{sn-jnl}% Default with double column layout

%%%% Standard Packages
%%<additional latex packages if required can be included here>
%%%%

%%%%%=============================================================================%%%%
%%%%  Remarks: This template is provided to aid authors with the preparation
%%%%  of original research articles intended for submission to journals published 
%%%%  by Springer Nature. The guidance has been prepared in partnership with 
%%%%  production teams to conform to Springer Nature technical requirements. 
%%%%  Editorial and presentation requirements differ among journal portfolios and 
%%%%  research disciplines. You may find sections in this template are irrelevant 
%%%%  to your work and are empowered to omit any such section if allowed by the 
%%%%  journal you intend to submit to. The submission guidelines and policies 
%%%%  of the journal take precedence. A detailed User Manual is available in the 
%%%%  template package for technical guidance.
%%%%%=============================================================================%%%%

\usepackage{aas_macros}

\jyear{2021}%

%% as per the requirement new theorem styles can be included as shown below
\theoremstyle{thmstyleone}%
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}% 
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.

\theoremstyle{thmstyletwo}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%

\theoremstyle{thmstylethree}%
\newtheorem{definition}{Definition}%

\raggedbottom
%%\unnumbered% uncomment this for unnumbered level heads

\begin{document}

\title{Stellar Bars in Spiral Galaxies Do Not Slow Down}

\maketitle
Elongated bar-like features are ubiquitous, occurring at the centers of
approximately two-thirds of spiral disk galaxies \cite{2000AJ....119..536E,
2007ApJ...657..790M}. Due to gravitational interactions between the bar and
the other components of galaxies, it is expected that angular momentum and
matter will redistribute between galactic components over long (Gyr)
timescales in galaxies hosting a bar \cite{1972MNRAS.157....1L,
1984MNRAS.209..729T, 1985MNRAS.213..451W}. Previous simulation work has
overwhelmingly provided the expectation that the bar pattern will slow its
rotation over time due to a drag caused by material in the dark matter halo on
orbits resonant with the tumbling bar pattern
\cite[e.g.][]{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
2002ApJ...569L..83A, 2003MNRAS.341.1179A, 2003MNRAS.346..251O,
2005MNRAS.363..991H, 2006ApJ...637..214M, 2007MNRAS.375..460W,
2009ApJ...697..293D}. Simulations have shown that bars should shed enough
angular momentum to be considered ``slow rotators'' in a few Gyr, but most
observed galaxies seem to be ``fast rotators'' \cite{2011MSAIS..18...23C,
2015AA...576A.102A, 2019MNRAS.482.1733G}. We have performed a simulation of an
isolated galactic disk hosting a strong bar which includes a state-of-the-art
model of the interstellar medium. In this simulation, the bar pattern does not
slow down over time, and instead remains at a stable, constant rate of
rotation. Since gas is torqued by the bar to fall towards the center of the
galaxy, it acts to increase the angular momentum of the bar. However, the
pattern speed we measure is nearly constant over many Gyr, suggesting a novel
equilibrium mechanism is at play - we propose such a mechanism consistent with
our simulation. First, a region of the halo phase space is carved out such
that no resonant drag can occur. If the pattern speed decreases, the
corotation radius becomes larger and so more gas is available for infall since
only gas within corotation can infall \cite[e.g.][]{2011MNRAS.415.1027H}. This
speeds the bar up. If the pattern speed increases, new dark matter which is
resonant with the bar becomes available, slowing the bar down. Thus, the
pattern speed must remain constant. The implications of this are numerous.
First, we resolve a long-standing controversy between the observed fast
rotators and the theoretical expectation that bars should be slow. Second, we
show that the role of gas is of paramount importance in studies which attempt
to uncover the nature of dark matter from its effect of slowing down the bar
\cite{2021MNRAS.500.4710C, 2021MNRAS.505.2412C}. Third, we provide an
explanation for how the Milky Way's bar could be both long-lived and a fast
rotator, of which there is some observational evidence
\cite[e.g.][]{2019MNRAS.490.4740B}. And finally, we complicate the picture of
radial mixing expected to sculpt the Milky Way's disk
\cite[e.g.][]{2012MNRAS.420..913B, 2015ApJ...808..132H}. Our work is a
significant advance in our understanding of the dynamics of barred galaxies.

\backmatter

\bibliography{sn-bibliography}% common bib file
%% if required, the content of .bbl file can be included here once bbl is generated
%%\input sn-article.bbl

%% Default %%
%%\input sn-sample-bib.tex%

\end{document}

### 2022-04-04 paper/ms.tex
Add dark matter halo
Added the phrase "dark matter" to some references of the halo where it might be unclear to the reader

@@
 angle difference in the SMUGGLE case is smaller than in the N-body case by
-about a factor of two, reflecting how the halo in the SMUGGLE case is unable
-to exert as negative a torque on the bar as in the N-body
+about a factor of two, reflecting how the dark matter halo in the SMUGGLE case
+is unable to exert as negative a torque on the bar as in the N-body
 case.}\label{fig:wake}
@@
 
-The presence of the gas can arrest the process by which the halo wake forms.
-However, this does not explain why the pattern speed in the SMUGGLE case is
-nearly constant over several Gyr. Naively, it would be a coincidence that the
-bar pattern speed remains constant in the SMUGGLE case, resulting from a
-chance cancellation of the halo and gas torques. However, a constant pattern
-speed in the presence of gas has been observed in a few simulations of barred
-galaxies with gas.\cite{1993AA...268...65F, 2010ApJ...719.1470V}
+The presence of the gas can arrest the process by which the dark matter halo
+wake forms. However, this does not explain why the pattern speed in the
+SMUGGLE case is nearly constant over several Gyr. Naively, it would be a
+coincidence that the bar pattern speed remains constant in the SMUGGLE case,
+resulting from a chance cancellation of the halo and gas torques. However, a
+constant pattern speed in the presence of gas has been observed in a few
+simulations of barred galaxies with gas.\cite{1993AA...268...65F,
+2010ApJ...719.1470V}
 
@@
 up. It is simple to explain the first of these - when the pattern speed
-increases, the location of resonances will shift to regions of the halo phase
-space where no wake has been excited yet (e.g., the corotation radius will
-shrink).
+increases, the location of resonances will shift to regions of the dark matter
+halo phase space where no wake has been excited yet (e.g., the corotation
+radius will shrink).
 
@@
 from the gas phase. Only gas within corotation will flow inwards, while gas
-outside corotation will flow outwards.\cite{2011MNRAS.415.1027H} Since the corotation radius is larger
-for more slowly rotating bars, it follows that more slowly rotating bars
-should be more efficient at driving gas inflows and thus experience a larger
-positive torque from the gas phase. We performed an experiment to test this
-hypothesis by forcing the stellar disk in the SMUGGLE run to rotate at a
-constant angular rate and measuring the torque on the bar by the gas phase at
-different rotation rates. The result of this experiment is shown in
-Fig.~\ref{fig:equil}, which shows that a more slowly rotating bar experiences
-a larger positive torque from the gas.
+outside corotation will flow outwards.\cite{2011MNRAS.415.1027H} Since the
+corotation radius is larger for more slowly rotating bars, it follows that
+more slowly rotating bars should be more efficient at driving gas inflows and
+thus experience a larger positive torque from the gas phase. We performed an
+experiment to test this hypothesis by forcing the stellar disk in the SMUGGLE
+run to rotate at a constant angular rate and measuring the torque on the bar
+by the gas phase at different rotation rates. The result of this experiment is
+shown in Fig.~\ref{fig:equil}, which shows that a more slowly rotating bar
+experiences a larger positive torque from the gas.
 
@@
 keeps the pattern speed of the bar constant, resulting from the complex
-interplay between the halo and the gas phase.
+interplay between the dark matter halo and the gas phase.
 
@@
 After generating the gaseous disk in this way, we stitched the gas disk
-together with the GALAKOS N-body disk (and bulge and halo) after the GALAKOS
-disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The purpose of allowing
-the GALAKOS to evolve first for a short period of time is to allow for the bar
-to form unimpacted by the presence of the gas (which would normally disrupt
-the formation of the bar). We made one additional modification when stitching
-the gas disk together with the N-body disk - we created a vacuum within the
-central $X$ pc. This vacuum guards against an initial dramatic infall of gas
-within the bar region, which we found to destroy the bar.
+together with the GALAKOS N-body disk (and bulge and dark matter halo) after
+the GALAKOS disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The
+purpose of allowing the GALAKOS to evolve first for a short period of time is
+to allow for the bar to form unimpacted by the presence of the gas (which
+would normally disrupt the formation of the bar). We made one additional
+modification when stitching the gas disk together with the N-body disk - we
+created a vacuum within the central $X$ pc. This vacuum guards against an
+initial dramatic infall of gas within the bar region, which we found to
+destroy the bar.
 

### 2022-04-04 paper/ms.tex
Cleanup

@@
 
-\author{Angus Beane$^{1*}$, Lars Hernquist$^1$, Elena D'Onghia$^2$, et al.}
+\author{Angus Beane$^{1*}$, Lars Hernquist$^1$, Elena D'Onghia$^{2,3}$, et al.}
 
@@
 
-\item Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson, AZ 85721, USA
+\item Department of Physics, University of Wisconsin-Madison, Madison, WI, USA
 
-\item Lol
+\item Department of Astronomy, University of Wisconsin-Madison, Madison, WI, USA
 
@@
 
-% \begin{figure*}
-% \includegraphics[width=0.5\textwidth]{fig1a.pdf}
-% \includegraphics[width=0.5\textwidth]{fig1b.pdf}
-% \vspace{-4mm}
-% \caption{\textbf{Distribution of stars in the Galactic halo.}  All-sky
-%   Mollweide projection maps of the density of stars at 60 kpc
-%   $<R_{\rm gal}<100$ kpc, smoothed by a Gaussian kernel with
-%   FWHM$ =30^\circ$.  {\bf Left panel:} data based on K giant stars.
-%   {\bf Right panel:} simulation that includes the dynamical response
-%   of the halo to the orbit of the LMC.  Regions near the Galactic
-%   mid-plane ($|b|<10^\circ$) and the LMC and SMC, as well as regions
-%   of high extinction are masked in grey, while the positions of the
-%   LMC and SMC are marked as white stars.  One arm of the Sagittarius
-%   stream has been masked in sky coordinates, as indicated by the grey
-%   region at $b>0^\circ$ and $-180^\circ<l<-150^\circ$.  The
-%   overdensity in the south-west is the local wake caused by the
-%   passage of the LMC.  The overdensity in the North is the collective
-%   response of the Galaxy due to the arrival of the LMC.  Notice that
-%   the dynamic range of the density contrast for the simulated map is
-%   smaller than the observed map.}
-% \vspace{-4mm}
-% \end{figure*}
-
-% \begin{figure}[t!]
-%   \center
-%   \includegraphics[width=0.48\textwidth]{fig2.pdf}
-%   \caption{\textbf{Quantitative comparison between data and models.}
-%     Mean density contrasts in regions encompassing the local wake
-%     ($-60^\circ<b<-20^\circ$, $45^\circ<l<90^\circ$) and the
-%     collective response ($30^\circ<b<60^\circ$,
-%     $240^\circ<l<300^\circ$) are shown along the $x$ and $y$ axes.
-%     There are four simulations spanning a range in LMC masses from
-%     $0.8-2.5\times10^{11}M_\odot$ (small-to-large green circles).  The
-%     smooth model\cite{Rybizki18} is close to zero along both axes as
-%     expected.  The dotted lines indicate uniform distributions (no
-%     density contrast) on the x and y axes.  The density enhancement of
-%     the local wake is measured at a significance of $5\sigma$ (error
-%     bars on the data represent $1\sigma$ Poisson uncertainties) and
-%     the mean density is $1.4\pm0.2$ times larger in the data than in
-%     the fiducial simulation.}
-% \end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%\bibliographystyle{naturemag}
 \bibliography{ref}
@@
 
-\item [Acknowledgements] We thank both referees for their constructive
-  feedback.  C.C. is partially supported by the Packard
-  Foundation.  R.P.N. gratefully acknowledges an Ashford Fellowship
-  and Peirce Fellowship granted by Harvard University.  G.B.  and
-  N.G.-C. are supported by HST grant AR 15004, NASA ATP grant
-  17-ATP17-0006, NSF CAREER AST-1941096.  A.B. acknowledges support
-  from NASA through HST grant HST-GO-15930.  All the simulations were
-  run on El-Gato super computer which was supported by the National
-  Science Foundation under Grant No. 1228509.  We have made use of
-  data from the European Space Agency mission Gaia
-  (http://www.cosmos.esa.int/gaia), processed by the Gaia Data
-  Processing and Analysis Consortium (DPAC; see
-  http://www.cosmos.esa.int/web/gaia/dpac/consortium). Funding for
-  DPAC has been provided by national institutions, in particular the
-  institutions participating in the Gaia Multilateral Agreement.  This
-  publication makes use of data products from the {\it Wide-field
-    Infrared Survey Explorer}, which is a joint project of the
-  University of California, Los Angeles, and the Jet Propulsion
-  Laboratory/California Institute of Technology, and NEOWISE, which is
-  a project of the Jet Propulsion Laboratory/ California Institute of
-  Technology.  {\it WISE} and NEOWISE are funded by the National
-  Aeronautics and Space Administration.
+\item [Acknowledgements] Foo.
 
-\item[Author Contributions] C.C. and R.P.N. jointly conceived of the
-  project.  C.C. led the analysis of the data.  N.G-C. and
-  G.B. provided the simulation data and aided in its interpretation.
-  All authors contributed to aspects of the analysis and to the
-  writing of the manuscript.
+\item[Author Contributions] Foo.
 
-\item[Author  Information]  Reprints and permissions
-    information is available at npg.nature.com/reprintsandpermissions.
-    Correspondence and requests for materials should be addressed to
-    C.C.\ (cconroy@cfa.harvard.edu).
+  \item[Data Availability] Foo.
 
-  \item[Data Availability] The K giant catalog used in this paper is
-    available at \texttt{https://doi.org/10.7910/DVN/2D1H8J}.
-
-  \item[Code Availability] We have opted not to make the code used in
-    this manuscript available because the data reduction and analysis
-    is straightforward and can be easily reproduced following the
-    methods described herein.
+  \item[Code Availability] Foo.
 
@@
 
-% \begin{figure}[t!]
-%   \includegraphics[width=0.45\textwidth]{ed_fig1.pdf}
-%   \vspace{-1cm}
-%   \caption{\textbf{Extended Data Figure 1 $|$ Photometric selection of
-%       giants.}  Color-color diagram for stars at high Galactic
-%     latitude ($b>45^\circ$) with {\it Gaia} parallax $\pi<0.1$ mas.
-%     Two sequences are clearly visible, with the upper branch being
-%     associated with giant stars, and the lower with dwarf stars. The
-%     red lines mark the selection boundary used in this work.}
-% \end{figure}
-
-% \begin{figure}[t!]
-%   \includegraphics[width=0.45\textwidth]{ed_fig2.pdf}
-%   \caption{\textbf{Extended Data Figure 2 $|$ Test of Photometric
-%       Distances.}  Comparison of our photometric distances against
-%     literature values for satellite galaxies around the Milky Way
-%     (red) and globular clusters (blue).  Error bars represent the
-%     $1\sigma$ scatter in the photometric distances and dotted lines
-%     mark $\pm20$\% about the one-to-one line.  The dwarf galaxies
-%     include Fornax, Draco, Sculptor, Carina, Ursa Minor, and the LMC
-%     and SMC, and span average metallicities from [Fe/H]$\approx-0.5$
-%     (LMC) to $-2.2$ (Ursa Minor)\cite{McConnachie12}.}
-% \end{figure}
-
-% \begin{figure*}[t!]
-%   \includegraphics[width=0.5\textwidth]{ed_fig3a.pdf}
-%   \includegraphics[width=0.5\textwidth]{ed_fig3b.pdf}
-%   \caption{\textbf{Extended Data Figure 3 $|$ Maps of un-smoothed star
-%       counts.}  Mollweide projection maps of the observed sample of
-%     giants with pixels of size 13.4 sq. deg. {\bf Left panel:} map
-%     with no masking of known structure, either in sky coordinates or
-%     proper motions.  The two main features identified here, the
-%     transient wake and the collective response, are still clearly
-%     visible even in this unfiltered map.  The LMC and SMC appear in
-%     the lower right as a merged region of high density.  Other
-%     features include the stellar disk+bulge in the center, and the
-%     Sagittarius stream both in the north-center and lower left and
-%     right.  {\bf Right panel:} map showing the filtered catalog used
-%     in Fig 1a.}
-% \end{figure*}
-
-% \begin{figure*}[t!]
-%   \center
-%   \includegraphics[width=0.8\textwidth]{ed_fig4.pdf}
-%   \caption{ \textbf{Extended Data Figure 4 $|$ Proper motions of the
-%       halo sample.}  {\bf a:} proper motions of the K giants at
-%     $60<R_{\rm gal}<100$ kpc with $12<W1<15$ and the LMC and SMC
-%     removed via on-sky selection.  {\bf b:} solar reflex-corrected
-%     proper motions.  Notice the much tighter distribution of stars
-%     near (0,0).  In this panel the Sagittarius dSph is visible at
-%     (-2.5,-0.7), LMC and SMC stars not removed by the on-sky selection
-%     are visible as a narrow vertical vertical strip at
-%     $\mu_\alpha\gtrsim0.5$ mas yr$^{-1}$.  The Northern Sagittarius
-%     arm is the large overdensity at (-0.7,0.0), and a Southern arm of
-%     Sagittarius is the diagonal cluster of points at $\mu_\delta<-0.5$
-%     mas yr$^{-1}$.  {\bf c:} reflex-corrected proper motions focusing
-%     on the region of the sky containing the Northern Sagittarius arm
-%     at $l\sim0^\circ$.  The red box indicates our selection for
-%     removing this feature.  {\bf d:} stars not in the selection shown
-%     in panel {\bf c}. Our selection for low proper motion stars is
-%     indicated by a blue circle in {\bf c, d}.}
-% \end{figure*}
-
 % \bibliography{ref2}

### 2022-04-05 paper/ms.tex
Small paper updates
Based on Elena's comments

@@
 units and over the same interval.
-ß
+
 \begin{figure}[h]%
@@
 
-The fact that the halo in the SMUGGLE case exerts a smaller positive torque on
-the bar can be understood in terms of the halo wake mechanism. In the N-body
-case, halo material which is resonant with the bar will form a wake which lags
-behind and exerts a negative torque on the bar, which slows it
-down.\cite{1984MNRAS.209..729T, 1985MNRAS.213..451W,
+The fact that the dark matter halo in the SMUGGLE case exerts a smaller
+positive torque on the bar can be understood in terms of the halo wake
+mechanism. In the N-body case, halo material which is resonant with the bar
+will form a wake which lags behind and exerts a negative torque on the bar,
+which slows it down.\cite{1984MNRAS.209..729T, 1985MNRAS.213..451W,
 1992ApJ...400...80H}\footnote{Since the bar is not a solid body, it is not
@@
 simulations of barred galaxies with gas.\cite{1993AA...268...65F,
-2010ApJ...719.1470V}
+2010ApJ...719.1470V} Previous work has argued this is due to the bar torquing
+gas inwards, but no explanation has been given for why it might remain
+constant.
 
@@
 \cite{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies upon
-the pattern speed of the bar to change with time.
+the pattern speed of the bar to change with time (though our work does not
+alter expectations for radial mixing induced by spiral arms\cite{2002MNRAS.336..785S}).
 

### 2022-04-06 paper/ms.tex
Move figures
Just move them all to the end
Add surface density
Adds the time evolution of the surface density

@@
 
-\begin{figure*}[b!]%
-\centering
-\includegraphics[width=0.8\textwidth]{fig/fig1.pdf}
-\caption{Surface density projections in an N-body only simulation (upper
-panels) and a simulation which includes the SMUGGLE model (bottom panels). We
-can see that in the N-body run, the bar grows in length and strength. In the
-SMUGGLE run (which includes a gaseous phase and a model for the multi-phase
-interstellar medium), the bar remains at approximately the same length and
-strength over the course of the simulation. The N-body model is identical to
-the GALAKOS model, discussed in the text.}\label{fig:overview}
-\end{figure*}
-
 We have performed a simulation of a disk galaxy using the finite-volume
@@
 
-\begin{figure}[h]%
-\centering
-\includegraphics[width=0.45\textwidth]{fig/fig2.pdf}
-\caption{Various bar properties from the N-body and SMUGGLE runs. The
-\textit{upper panel} show the evolution of the pattern speed. As expected, the
-bar in the N-body run slows down due to interactions between the bar and the
-dark matter halo. However, the bar in the SMUGGLE run does not slow down and
-instead remains at a constant pattern speed. The \textit{middle panel} shows
-the evolution of the bar length. In the N-body case, the bar lengthens. This
-occurs because as the pattern speed drops, bar-like orbits at larger radii are
-possible. Stars are captured on these orbits, lengthening the bar. This
-process does not occur in the SMUGGLE cases since the bar pattern speed is not
-decreasing, and therefore the bar length remains constant. The \textit{lower
-panel} shows the torque on the bar by different components. The solid lines
-show the torque exerted by the halo in both the N-body and SMUGGLE cases. The
-dashed line shows the torque exerted by the gas phase in the SMUGGLE run
-(there is no gas in the N-body run). The first $1.5\,\textrm{Gyr}$ of
-evolution of the N-body model is not shown. Details on how these properties
-are calculated is given in the Methods section.}\label{fig:prop}
-\end{figure}
-
 The fact that the dark matter halo in the SMUGGLE case exerts a smaller
@@
 (right) cases.
-
-\begin{figure*}[h]%
-\centering
-\includegraphics[width=0.9\textwidth]{fig/fig3.pdf}
-\caption{The wake excited in the dark matter halo in the N-body case
-(\textit{left panel}) and SMUGGLE case (\textit{right panel}) after
-$2.6\,\textrm{Gyr}$. The \textit{left} and \textit{right panels} show a
-surface density projection in the $x$-$y$ plane of the dark matter halo after
-an axisymmetric average has been subracted. The solid line indicates the
-direction of the bar while the dashed line indicates the direction of the halo
-wake (both measured by taking the second Fourier component within a sphere of
-all material within a radius of $X\,\textrm{kpc}$). The \textit{center panel}
-shows the time evolution of the angle difference between the bar and the halo
-wake, as measured from the second Fourier component. After the first Gyr, the
-angle difference in the SMUGGLE case is smaller than in the N-body case by
-about a factor of two, reflecting how the dark matter halo in the SMUGGLE case
-is unable to exert as negative a torque on the bar as in the N-body
-case.}\label{fig:wake}
-\end{figure*}
 
@@
 
-\begin{figure}[h]
-\centering
-\includegraphics{fig/fig4.pdf}
-\caption{The torque on the bar by the gas when the pattern speed of the bar is
-kept at a constant value. Only gas within the corotation radius is able to
-infall. Since slower bars have larger corotation radii, slower bars experience
-a larger net torque than faster bars. The setup of the simulations used here
-is identical to the SMUGGLE case discussed earlier and in the Methods section,
-except the N-body disk is rotated as a solid body with a constant angular
-velocity.}\label{fig:equil}
-\end{figure}
-
 We have therefore shown evidence for an equilibrium mechanism at play which
@@
 
+\begin{figure*}[h]%
+\centering
+\includegraphics[width=0.8\textwidth]{fig/fig1.pdf}
+\caption{Surface density projections in an N-body only simulation (upper
+panels) and a simulation which includes the SMUGGLE model (bottom panels). We
+can see that in the N-body run, the bar grows in length and strength. In the
+SMUGGLE run (which includes a gaseous phase and a model for the multi-phase
+interstellar medium), the bar remains at approximately the same length and
+strength over the course of the simulation. The N-body model is identical to
+the GALAKOS model, discussed in the text.}\label{fig:overview}
+\end{figure*}
+
+\begin{figure}[h]%
+\centering
+\includegraphics[width=0.45\textwidth]{fig/fig2.pdf}
+\caption{Various bar properties from the N-body and SMUGGLE runs. The
+\textit{upper panel} show the evolution of the pattern speed. As expected, the
+bar in the N-body run slows down due to interactions between the bar and the
+dark matter halo. However, the bar in the SMUGGLE run does not slow down and
+instead remains at a constant pattern speed. The \textit{middle panel} shows
+the evolution of the bar length. In the N-body case, the bar lengthens. This
+occurs because as the pattern speed drops, bar-like orbits at larger radii are
+possible. Stars are captured on these orbits, lengthening the bar. This
+process does not occur in the SMUGGLE cases since the bar pattern speed is not
+decreasing, and therefore the bar length remains constant. The \textit{lower
+panel} shows the torque on the bar by different components. The solid lines
+show the torque exerted by the halo in both the N-body and SMUGGLE cases. The
+dashed line shows the torque exerted by the gas phase in the SMUGGLE run
+(there is no gas in the N-body run). The first $1.5\,\textrm{Gyr}$ of
+evolution of the N-body model is not shown. Details on how these properties
+are calculated is given in the Methods section.}\label{fig:prop}
+\end{figure}
+
+\begin{figure*}[h]%
+\centering
+\includegraphics[width=0.9\textwidth]{fig/fig3.pdf}
+\caption{The wake excited in the dark matter halo in the N-body case
+(\textit{left panel}) and SMUGGLE case (\textit{right panel}) after
+$2.6\,\textrm{Gyr}$. The \textit{left} and \textit{right panels} show a
+surface density projection in the $x$-$y$ plane of the dark matter halo after
+an axisymmetric average has been subracted. The solid line indicates the
+direction of the bar while the dashed line indicates the direction of the halo
+wake (both measured by taking the second Fourier component within a sphere of
+all material within a radius of $X\,\textrm{kpc}$). The \textit{center panel}
+shows the time evolution of the angle difference between the bar and the halo
+wake, as measured from the second Fourier component. After the first Gyr, the
+angle difference in the SMUGGLE case is smaller than in the N-body case by
+about a factor of two, reflecting how the dark matter halo in the SMUGGLE case
+is unable to exert as negative a torque on the bar as in the N-body
+case.}\label{fig:wake}
+\end{figure*}
+
+\begin{figure}[h]
+\centering
+\includegraphics{fig/fig4.pdf}
+\caption{The torque on the bar by the gas when the pattern speed of the bar is
+kept at a constant value. Only gas within the corotation radius is able to
+infall. Since slower bars have larger corotation radii, slower bars experience
+a larger net torque than faster bars. The setup of the simulations used here
+is identical to the SMUGGLE case discussed earlier and in the Methods section,
+except the N-body disk is rotated as a solid body with a constant angular
+velocity.}\label{fig:equil}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \captionsetup[figure]{labelformat=empty}
-\renewcommand{\thefigure}{Extended Data \arabic{figure}}
+\renewcommand{\figurename}{Extended Data Figure}
 \renewcommand{\thetable}{Extended Data \arabic{table}}
@@
 balance. We modified the radial profile of this code in order to allow us to
-generate a disk with a constant surface density within some radius, and then
-exponentially declining beyond that radius with the scale-length of the
-stellar disk. We used a surface density of $X$ and a maximum radius of $Y$. We
+generate a disk with a constant surface density within some cut-off radius,
+and then exponentially declining beyond that radius with the scale-length of
+the stellar disk. We used an initial surface density of
+$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\text{kpc}$. We
 varied the surface density in TEST.
@@
 modification when stitching the gas disk together with the N-body disk - we
-created a vacuum within the central $X$ pc. This vacuum guards against an
-initial dramatic infall of gas within the bar region, which we found to
-destroy the bar.
+created a vacuum within the central $4\,\textrm{kpc}$. This vacuum guards
+against an initial dramatic infall of gas within the bar region, which we
+found to destroy the bar.
 
 Our setup is initially out of equilibrium, but we found that after
-$\sim300\,\textrm{Myr}$, the entire system has settled into a steady-state
+$\sim500\,\textrm{Myr}$, the entire system has settled into a steady-state
 configuration and initial transients appear not to affect the results after
@@
 galaxies to be appropriate. We show the evolution of the surface density
-profile in Fig.~X compared to observational estimates (CITE).
+profile in Extended Data Fig.~\ref{fig:surf} compared to observational
+estimates (CITE).
+
+\begin{figure}[h]%
+\centering
+\includegraphics[width=0.45\textwidth]{fig/fig-surf.pdf}
+\caption{The time evolution of the gas surface density (\textit{upper}) and the star formation rate (SFR) surface density (\textit{lower}).  } \label{fig:surf}
+\end{figure}
 

### 2022-04-07 paper/ms.tex
Add circular velocity curve

@@
 
+\usepackage{amsmath}
+
 \usepackage{fontspec}
@@
 \usepackage[misc]{ifsym}
+
+\newcommand{\RCR}{\ensuremath{R_{\textrm{CR}}}}
+\newcommand{\Rot}{\ensuremath{\mathcal{R}}}
 
@@
 the stellar disk. We used an initial surface density of
-$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\text{kpc}$. We
+$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\textrm{kpc}$. We
 varied the surface density in TEST.
@@
 for ensuring the gas disk is dense enough in order for comparisons to real
-galaxies to be appropriate. We show the evolution of the surface density
-profile in Extended Data Fig.~\ref{fig:surf} compared to observational
-estimates (CITE).
+galaxies to be appropriate. We plot the circular velocity curve in Extended
+Data Fig.~\ref{fig:vcirc} compared to observational
+estimates.\cite{2019ApJ...871..120E} Our disk is slightly more massive than
+the Milky Way disk, and in the SMUGGLE run the addition of the gas phase
+results in a slightly higher circular velocity. We also show the evolution of
+the surface density profile in Extended Data Fig.~\ref{fig:surf} compared to
+observational estimates (CITE).
+
+We used a mass resolution of $7.5\times10^3\,M_{\odot}$ for the baryonic
+components (initial stellar disk, stellar bulge, and gas) and a mass
+resolution of $3.75\times10^4\,M_{\odot}$ for the dark matter halo. This mass
+resolution is closest to ``level 3'' in the AURIGA
+simulations.\cite{2017MNRAS.467..179G} This corresponds to approximately
+$6.4\times10^6$ particles in the stellar disk, $1.1\times10^6$ in the bulge,
+$1.2\times10^6$ in the gas disk, and $25.3\times10^6$ in the dark matter halo.
+We used a softening length of $0.02\,\textrm{kpc}$ for all components.
+
+\begin{figure}[h]%
+\centering
+\includegraphics[width=0.45\textwidth]{fig/fig-vcirc.pdf}
+\caption{The circular velocity curve for the N-body run (blue) and the SMUGGLE
+run (orange) compared to observational estimates.\cite{2019ApJ...871..120E} We
+see that the circular velocity curve for both runs is marginally larger than
+the Milky Way's, but still comparable. The SMUGGLE circular velocity curve is
+larger than the N-body curve due to the additional mass in the gas phase.}
+\label{fig:surf}
+\end{figure}
 

### 2022-04-08 paper/ms.tex
Small typo

@@
 the GALAKOS disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The
-purpose of allowing the GALAKOS to evolve first for a short period of time is
-to allow for the bar to form unimpacted by the presence of the gas (which
-would normally disrupt the formation of the bar). We made one additional
-modification when stitching the gas disk together with the N-body disk - we
-created a vacuum within the central $4\,\textrm{kpc}$. This vacuum guards
-against an initial dramatic infall of gas within the bar region, which we
-found to destroy the bar.
+purpose of allowing the GALAKOS disk to evolve first for a short period of
+time is to allow for the bar to form unimpacted by the presence of the gas
+(which would normally disrupt the formation of the bar). We made one
+additional modification when stitching the gas disk together with the N-body
+disk - we created a vacuum within the central $4\,\textrm{kpc}$. This vacuum
+guards against an initial dramatic infall of gas within the bar region, which
+we found to destroy the bar.
 

### 2022-04-08 paper/ms.tex
Add more SMUGGLE details

@@
 \usepackage[misc]{ifsym}
+\usepackage{xcolor}
 
@@
 
+We used the fiducial model parameters, except that we increased the number of
+effective neighbors $N_{\textrm{ngb}}$ for the deposition of feedback from
+$64$ to $512$. We found that a lower value of $N_{\textrm{ngb}}$ resulted in
+inefficient photoionization feedback since the photoionizing budget had not
+been exhausted after deposition into $64$ neighboring cells. We also used an
+updated version of SMUGGLE using a new mechanical feedback routine similar to
+the one described in ref.\cite{2018MNRAS.480..800H} This updated routine is a
+tensor renormalization which ensures linear and angular momentum conservation
+to machine precision.
+
 \vspace{12pt}
@@
 the stellar disk. We used an initial surface density of
-$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\textrm{kpc}$. We
-varied the surface density in TEST.
+$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\textrm{kpc}$.
 

### 2022-04-08 paper/ms.tex
Add bar analysis section

@@
 \noindent
+{\bf Bar analysis}
+\\
+\noindent
+The analysis of various bar properties is performed as follows. First, the
+pattern speed is measured from the angle of the second Fourier component. We measured the second Fourier component by computing,
+\begin{equation}
+\begin{split}
+A_2 &= \sum_i m_i e^{i 2 \phi_i} \\
+A_0 &= \sum_i m_i \textrm{,}
+\end{split}
+\end{equation}
+where $m_i$ and $\phi_i$ are the mass and azimuthal angle of each particle, respectively. We
+computed $A_2$ and $A_0$ in cylindrical bins of width $0.5\,\textrm{kpc}$ from radii of
+$0$ to $30\,\textrm{kpc}$. We defined the angle of the bar $\phi_b$ to be
+twice the angle of complex number $A_2$ as measured in the bin extending from
+a radius of $2.5$ to $3\,\textrm{kpc}$. After correcting for the periodicity
+of $\phi_b$, we measured the pattern speed as the two-sided finite gradient of
+$\phi_b$ as a function of time.
+
+The time evolution of the bar strength, defined as the maximum of
+$\left|A_2/A_0\right|$ as a function of radius, is shown in Extended Data
+Fig.~\ref{fig:strength}. The quantity $\left|A_2/A_0\right|$ varies from $0$
+to $1$, with larger values indicating a stronger bar pattern. We see that in
+the N-body case, $\left|A_2/A_0\right|$ increases over time as the bar pattern
+slows. This is consistent with previous N-body simulations which showed a
+clear correlation between the bar pattern speed and the bar
+strength.\cite{2003MNRAS.341.1179A} In the SMUGGLE case, we see that the bar
+strength has an initial drop but then remains at a roughly constant but
+slightly decreasing strength. This is consistent with the pattern speed in the
+SMUGGLE case being roughly constant or slightly increasing.
+
+\begin{figure}[h]%
+\centering
+\includegraphics[width=0.45\textwidth]{fig/fig-A2.pdf}
+\caption{The time evolution of the bar strength, measured as the second
+Fourier component divided by the zeroth Fourier component (a formula is given
+in the Methods section). We see that in the N-body case (blue) the bar
+strength increases with time, consistent with previous results showing that
+the bar strength increases as bars slow down. In the SMUGGLE case (orange), we
+see that the bar strength remains roughly constant, possibly slightly
+decreasing with time. This is also consistent with the expected relation
+between pattern speed and strength since the bar in this case is not slowing
+down.}
+\label{fig:strength}
+\end{figure}
+
+Computing the length of the bar and the torque on the bar by different
+components requires us to decompose the disk into a component which is trapped
+by the bar and a component which is untrapped. In order to do this, we follow
+closely the technique developed in ref.\cite{2016MNRAS.463.1952P} We analyzed
+the orbit of each star particle (meaning initial disk, bulge, and newly formed
+stars) by extracting the $x$-$y$ positions of the apoapse of each in a frame
+corotating with the bar (apoapses are defined as local maxima in $r$). For
+each apoapse, we searched for the $19$ closest apoapses in time and applied a
+$k$-means clustering algorithm on this set of $20$ points with $k=2$. We then
+computed for each of the two clusters the average angle from the bar
+$\left<\Delta \phi\right>_{0,1}$, the standard deviation in $R$ of the points
+${\sigma_R}_{0,1}$, and the average radius of the cluster
+$\left<R\right>_{0,1}$. At each apoapse, a particle was considered to be in
+the bar if it met the following criteria:
+\begin{equation}
+\textrm{max}\left(\left<\Delta \phi\right>_{0,1}\right) < \pi / 8
+\end{equation}
+\begin{equation}
+\frac{{\sigma_R}_0 + {\sigma_R}_1}{\left<R\right>_0 + \left<R\right>_1} < 0.22
+\end{equation}
+These criterion are slightly different and simplified from the ones used in
+ref.\cite{2016MNRAS.463.1952P}, but we found to empirically work well at
+decomposing the disk into a bar and disk component. In Extended Data
+Fig.~\ref{fig:decomp}, we show an example of this decomposition. The
+\textit{left} panel shows a surface density projection of the stellar disk and
+bulge (including newly formed stars) from the SMUGGLE model after
+$1\,\text{Gyr}$ of evolution in a frame such that the bar is aligned with the
+$x$-axis. The \textit{middle} panel shows a projection of the subset of stars
+that are identified as being trapped in the bar and the \textit{right} panel
+shows a projection of the stars that are not identified as being trapped. The
+fact that the \textit{right} panel is roughly axisymmetric indicates the bar
+decomposition is performing adequately.
+
+\begin{figure*}[h]%
+\centering
+\includegraphics[width=0.95\textwidth]{fig/fig-bar_decomp.pdf}
+\caption{Demonstration of our bar decomposition procedure based on
+ref.\cite{2016MNRAS.463.1952P} The \textit{left panel} shows a surface density
+projection through the stellar component of the N-body simulation (disk and
+bulge). The \textit{middle panel} shows the component of the disk identified
+as being trapped in the bar while the \textit{right panel} shows the component
+of the disk identified as not being trapped in the bar. The fact that the
+untrapped stars form a roughly axisymmetric structure indicates our bar
+decomposition is sufficiently accurate. {\color{red} I will remake this figure
+- I ended up losing the files which stored the bar membership, so I just
+pulled these screenshots from an earlier presentation.}}
+\label{fig:decomp}
+\end{figure*}
+
+After the disk has been decomposed into a trapped and untrapped component, we
+measured the bar length as being the radius $R_b$ which encapsulates $99\%$ of
+the stars identified as being trapped in the bar, allowing for some outliers.
+For the computation of the torque, we used the tree algorithm in
+\texttt{MakeNewDisk} customized to be accessible from \texttt{Python} using
+\texttt{Cython}. This algorithm is based on the \texttt{TREESPH} code. We
+constructed a tree using only the star particles identified as being trapped
+in the bar. We then queried the tree at the locations of all resolution
+elements in the other components and computed the torque of the bar on such
+components. The torque on the bar by the other components is simply the
+negative of the torque on the other components by the bar. A similar method
+was applied in measuring the torque for the disk when the pattern speed is
+kept constant, as is done in Fig.~\ref{fig:equil}.
+
+\vspace{12pt}
+
+\noindent
 {\bf Title}

### 2022-04-19 paper/ms.tex
Add more intro paragraphs

@@
 
+It has long been known that non-axisymmetric features in a stellar disk act to
+redistribute mass and angular momentum.\cite{1972MNRAS.157....1L} The
+interaction between a bar and a spheroid (either a dark matter halo or stellar
+bulge) has received considerable interest.\cite{1984MNRAS.209..729T,
+1985MNRAS.213..451W} A reasonable stellar disk in isolation seems to always
+generate a bar instability,\cite{1971ApJ...168..343H} but the presence of a
+static dark matter halo acts to stabilize the disk against the bar
+instability.\cite{1973ApJ...186..467O, 1976AJ.....81...30H} Later, a number of
+numerical simulations have confirmed that a live dark matter halo generically
+exerts a negative torque on the bar, causing its pattern speed to decrease and
+its length and strength to increase\cite{1992ApJ...400...80H,
+2000ApJ...543..704D, 2002MNRAS.330...35A, 2002ApJ...569L..83A,
+2003MNRAS.341.1179A, 2003MNRAS.346..251O, 2005MNRAS.363..991H,
+2006ApJ...637..214M, 2007MNRAS.375..460W, 2009ApJ...697..293D}. Intuitively,
+the bar excites a wake of resonant material in the dark matter halo in a
+manner similar to dynamical friction. This wake lags and thus exerts a
+negative torque on the bar.
+
+The pre-eminent expectation that the dark matter halo acts to slow the bar
+down conflicts with observational estimates of bar pattern speeds. Bar
+rotation rates are typically classified by the dimensionless ratio
+\begin{equation}
+\Rot = \RCR/R_b\textrm{,}
+\end{equation}
+where \RCR{} is the radius of corotation and $R_b$ is the length of the
+bar.\footnote{The radius of corotation \RCR{} is defined for circular orbits as
+the radius at which the orbital frequency is equal to the pattern speed,
+$\Omega_p$, of a given non-axisymmetric feature. In a galaxy with a constant
+circular velocity $V_c$, it is given by $\RCR = V_c / \Omega_p$.} Galaxies
+with $\Rot < 1.4$ are considered ``fast rotators'' while galaxies with $\Rot >
+1.4$ are considered ``slow rotators.''\cite{2000ApJ...543..704D} Galaxies with
+$\Rot < 1$ are not thought to be stable.\cite{1980AA....81..198C}
+Observational estimates of the pattern speeds of bars indicate that nearly all
+galaxies have $1 < \Rot < 1.4$.\cite{2011MSAIS..18...23C, 2015AA...576A.102A,
+2019MNRAS.482.1733G}
+
+The role of gas on the evolution of the bar is less well-understood. Since the
+gas phase typically only contributes about $10-20\%$ of the mass of a galaxy
+at the present day, one might naively expect it to have a subdominant effect
+on the bar. However, because gas is collisional, it can participate in
+non-resonant angular momentum exchange with the bar. Thus, numerical work has
+shown that the gas phase has an outsized influence on the
+bar.\cite{2010ApJ...719.1470V, 2013MNRAS.429.1949A}
+
 We have performed a simulation of a disk galaxy using the finite-volume
-gravity-hydrodynamics code AREPO \cite{2010MNRAS.401..791S}. In AREPO, the
+gravito-hydrodynamics code AREPO.\cite{2010MNRAS.401..791S} In AREPO, the
 fluid is discretized as a moving Voronoi mesh. We use the additional physics
@@
 radiative heating/cooling, star formation, and stellar feedback. More detailed
-information on SMUGGLE is given in the Methods section.
-
-Our disk galaxy is a modified version of the GALAKOS
-model\cite{2020ApJ...890..117D}, which consists of a radially exponential and
-vertically isothermal stellar disk along with a
-Hernquist\cite{1990ApJ...356..359H} dark matter halo and bulge. After
+information on SMUGGLE is given in the Methods section. Our disk galaxy is a
+modified version of the GALAKOS model\cite{2020ApJ...890..117D}, which
+consists of a stellar disk, stellar bulge, and dark matter halo. After
 $\sim2.5\,\textrm{Gyr}$ of evolution, the GALAKOS disk forms a bar consistent
@@
 ($\sim40\,\textrm{km}/\textrm{s}/\textrm{kpc}$ and $\sim4.5\,\textrm{kpc}$,
-respectively). We modify this setup by introducing a gas phase after $\sim
-1.5\,\textrm{Gyr}$ of evolution. This initial gas phase includes a hole in the
-center $4\,\textrm{kpc}$ and a constant surface density profile out to
-$9.3\,\textrm{kpc}$. When the gas is added to the simulation, the bar has
-already formed. More details on this configuration are given in the Methods
-section.
+respectively). We modify this setup to include a gas phase, the details of
+which is given in the Methods section.
 

### 2022-04-19 paper/ms.tex
Add caption to surface density plot

@@
 \includegraphics[width=0.45\textwidth]{fig/fig-surf.pdf}
-\caption{The time evolution of the gas surface density (\textit{upper}) and the star formation rate (SFR) surface density (\textit{lower}).  } \label{fig:surf}
+\caption{The time evolution of the gas surface density (\textit{upper}) and
+the star formation rate (SFR) surface density (\textit{lower}) at various
+times during the simulation. Colored lines indicate the profiles at selected
+times during the simulation while the horizontal dashed lines indicate
+approximate values for the Milky Way.\cite{2012ARAA..50..531K} We see that the
+gas surface density is slightly higher but within an order of magnitude of the
+Milky Way's typical gas surface density. We see a sharp decrease in the gas
+and SFR surface density along the extent of the bar from $\sim1$ to
+$\sim4\,\textrm{kpc}$, related to the gas inflow in this region. {\color{red}
+I will add the actual points from ref.\cite{2012ARAA..50..531K} (Fig. 7), but
+just placing approx horizontal lines for now. The profiles are roughly flat
+anyways.}}
+\label{fig:surf}
 \end{figure}

### 2022-04-19 paper/ms.tex
Update some references

@@
 larger than the N-body curve due to the additional mass in the gas phase.}
-\label{fig:surf}
+\label{fig:vcirc}
 \end{figure}
@@
 $0$ to $30\,\textrm{kpc}$. We defined the angle of the bar $\phi_b$ to be
-twice the angle of complex number $A_2$ as measured in the bin extending from
+twice the angle of the complex number $A_2$ as measured in the bin extending from
 a radius of $2.5$ to $3\,\textrm{kpc}$. After correcting for the periodicity
@@
 For the computation of the torque, we used the tree algorithm in
-\texttt{MakeNewDisk} customized to be accessible from \texttt{Python} using
-\texttt{Cython}. This algorithm is based on the \texttt{TREESPH} code. We
-constructed a tree using only the star particles identified as being trapped
-in the bar. We then queried the tree at the locations of all resolution
-elements in the other components and computed the torque of the bar on such
-components. The torque on the bar by the other components is simply the
-negative of the torque on the other components by the bar. A similar method
-was applied in measuring the torque for the disk when the pattern speed is
-kept constant, as is done in Fig.~\ref{fig:equil}.
+\texttt{MakeNewDisk}\cite{2005MNRAS.361..776S} customized to be accessible
+from \texttt{Python} using \texttt{Cython}. This algorithm is based on the
+\texttt{TREESPH} code.\cite{1989ApJS...70..419H} We constructed a tree using
+only the star particles identified as being trapped in the bar using an
+opening angle of $0.35$. We then queried the tree at the locations of all
+resolution elements in the other components and computed the torque of the bar
+on such components. The torque on the bar by the other components is simply
+the negative of the torque on the other components by the bar. A similar
+method was applied in measuring the torque for the disk when the pattern speed
+is kept constant, as is done in Fig.~\ref{fig:equil}.
 

### 2022-04-19 paper/ms.tex
Add circular velocity to methods

@@
 for ensuring the gas disk is dense enough in order for comparisons to real
-galaxies to be appropriate. We plot the circular velocity curve in Extended
-Data Fig.~\ref{fig:vcirc} compared to observational
+galaxies to be appropriate.
+
+We computed the circular velocity curve of our model using the \texttt{AGAMA}
+package.\cite{2019MNRAS.482.1525V} We fit the baryonic component (stellar
+disk, bulge, gas, and newly formed stars) with an axisymmetric cylindrical
+spline with $20$ grid points in both the radial and vertical direction
+spanning $0.2$ to $50\,\textrm{kpc}$ in the radial direction and from $0.02$
+to $10\,\textrm{kpc}$ in the vertical direction. We fit the dark matter halo
+using a spherically symmetric multipole fit with a maximum angular harmonic
+coefficient of $l=2$. We plot the circular velocity curve in Extended Data
+Fig.~\ref{fig:vcirc} compared to observational
 estimates.\cite{2019ApJ...871..120E} Our disk is slightly more massive than
@@
 the surface density profile in Extended Data Fig.~\ref{fig:surf} compared to
-observational estimates (CITE).
+observational estimates.\cite{2012ARAA..50..531K}
 

### 2022-04-19 paper/ms.tex
Small updates to intro

@@
 of the pattern speed in the N-body case is consistent with a long line of
-numerical research on bars in N-body simulations\cite{1992ApJ...400...80H,
+numerical research on bars in N-body simulations.\cite{1992ApJ...400...80H,
 2000ApJ...543..704D, 2002MNRAS.330...35A, 2002ApJ...569L..83A,
 2003MNRAS.341.1179A, 2003MNRAS.346..251O, 2005MNRAS.363..991H,
-2006ApJ...637..214M, 2007MNRAS.375..460W, 2009ApJ...697..293D}. However, in
-the SMUGGLE case the pattern speed remains constant. After the first Gyr of
+2006ApJ...637..214M, 2007MNRAS.375..460W, 2009ApJ...697..293D} However, in the
+SMUGGLE case the pattern speed remains constant. After the first Gyr of
 evolution, we find that the pattern speed increases by only $\sim10\%$ over
 the next $4\,\textrm{Gyr}$, compared to a $\sim43\%$ decrease in the pattern
-speed for the N-body run over the same interval. As we saw
-qualitatively in Fig.~\ref{fig:overview}, the length of the bar in the N-body
-case grows over time while it remains constant in the SMUGGLE case.
+speed for the N-body run over the same interval. As we saw qualitatively in
+Fig.~\ref{fig:overview}, the length of the bar in the N-body case grows over
+time while it remains constant in the SMUGGLE case. This is also consistent
+with previous numerical work, which found that bars tend to grow as they slow
+down and the radius of corotation increases.\cite{2000ApJ...543..704D,
+2003MNRAS.341.1179A}
 

### 2022-04-23 paper/ms.tex
Change title
Tom Dame edits
And other small things I noticed

@@
 
-\title{Stellar Bars in Isolated Gas-Rich Galaxies Do Not Slow Down}
+\title{Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down}
 
@@
 
-  Elongated bar-like features are ubiquitous, occuring at the centres of
+  Elongated bar-like features are ubiquitous, occurring at the centers of
   approximately two-thirds of spiral disk galaxies \cite{2000AJ....119..536E,
@@
   1984MNRAS.209..729T, 1985MNRAS.213..451W}. Previous work has overwhelmingly
-  provided the expectation that, due to these interactions, the bar pattern
+  provided the expectation that, due to the interaction between the bar and the halo, the bar pattern
   will slow its rotation over time \cite{1992ApJ...400...80H,
@@
   performed a simulation of an isolated galactic disk hosting a strong bar
-  which includes a state-of-the-art model of the interstellar medium. Here we
+  which includes a state-of-the-art model of the interstellar medium and a live dark matter halo. Here we
   show that in this simulation the bar pattern does not slow down over time,
@@
   has been observed in previous simulations but its explanation has remained
-  elusive.\cite{1993AA...268...65F, 2010ApJ...719.1470V}. We propose an
-  equilibrium mechanism which is consistent with our simulations. This result
-  challenges our expectations for how barred galaxies redistribute angular
-  momentum and matter over Gyr timescales, which is especially relevant for
+  elusive.\cite{1993AA...268...65F, 2010ApJ...719.1470V}.
+We propose that the gas phase of the disk and the dark matter halo act in concert to stabilize the bar pattern and prevent it from slowing down or speeding up.
+This result
+  naturally explains why nearly all observed bars rotate quickly\cite{2011MSAIS..18...23C, 2015AA...576A.102A,
+2019MNRAS.482.1733G} and is especially relevant for
   our understanding of how the Milky Way arrived at its present day state.
@@
 bulge) has received considerable interest.\cite{1984MNRAS.209..729T,
-1985MNRAS.213..451W} A reasonable stellar disk in isolation seems to always
-generate a bar instability,\cite{1971ApJ...168..343H} but the presence of a
-static dark matter halo acts to stabilize the disk against the bar
-instability.\cite{1973ApJ...186..467O, 1976AJ.....81...30H} Later, a number of
+1985MNRAS.213..451W} A stellar disk in isolation seems to always
+generate a bar instability,\cite{1971ApJ...168..343H} but the presence of a strong spherical potential (e.g., from a stellar bulge or dark matter halo) acts to stabilize the disk against the bar
+instability.\cite{1973ApJ...186..467O, 1976AJ.....81...30H} A number of
 numerical simulations have confirmed that a live dark matter halo generically
@@
 
-The pre-eminent expectation that the dark matter halo acts to slow the bar
+The preeminent expectation that the dark matter halo acts to slow the bar
 down conflicts with observational estimates of bar pattern speeds. Bar
@@
 
-A surface density projection of our simulatons is shown in
+A surface density projection of our simulations is shown in
 Fig.~\ref{fig:overview}. The upper three panels show the disk in the N-body
@@
 
-We can test this interpration by measuring the angle offset between the halo
+We can test this interpretation by measuring the angle offset between the halo
 wake and the bar. If the wake and the bar are aligned (i.e., there is no angle
@@
 more slowly rotating bars should be more efficient at driving gas inflows and
-thus experience a larger positive torque from the gas phase. We performed an
-experiment to test this hypothesis by forcing the stellar disk in the SMUGGLE
-run to rotate at a constant angular rate and measuring the torque on the bar
-by the gas phase at different rotation rates. The result of this experiment is
-shown in Fig.~\ref{fig:equil}, which shows that a more slowly rotating bar
-experiences a larger positive torque from the gas.
+thus experience a larger positive torque from the gas phase.
 
-We have therefore shown evidence for an equilibrium mechanism at play which
-keeps the pattern speed of the bar constant, resulting from the complex
-interplay between the dark matter halo and the gas phase.
+We performed an experiment to test this hypothesis by forcing the stellar disk
+in the SMUGGLE run to rotate at a constant angular rate and measuring the
+torque on the bar by the gas phase at different rotation rates. The result of
+this experiment is shown in Fig.~\ref{fig:equil}, which shows that a more
+slowly rotating bar experiences a larger positive torque from the gas. We have
+therefore shown evidence for an equilibrium mechanism which keeps the pattern
+speed of the bar constant, resulting from the complex interplay between the
+dark matter halo and the gas phase.
 
@@
 Barred galaxies in cosmological simulations of galaxy formation continue to be
-in conflict with observations.\cite{2017MNRAS.469.1054A, 2019MNRAS.483.2721P,
+in conflict with observations by producing bars which rotate too slowly.\cite{2017MNRAS.469.1054A, 2019MNRAS.483.2721P,
 2021AA...650L..16F} However, the pattern speeds of bars in both cosmological
@@
 dashed line shows the torque exerted by the gas phase in the SMUGGLE run
-(there is no gas in the N-body run). The first $1.5\,\textrm{Gyr}$ of
-evolution of the N-body model is not shown. Details on how these properties
+(there is no gas in the N-body run). Details on how these properties
 are calculated is given in the Methods section.}\label{fig:prop}
@@
 Star formation occurs in cells above a density threshold
-($n_{\textrm{th}}=100\,\textrm{cm}^{-3}$) according to SH03 with a
-star-formation efficiency of $\epsilon = 0.01$. Star formation converts gas
-cells into star particles which represent single stellar populations. For each
-star particle, the deposition of energy, momentum, and mass from stellar winds
-and supernovae is modeled. Photoionization and radiation pressure are modeled
-using an approximate treatment. A more detailed description of this model can
-be found in the flagship SMUGGLE paper.\cite{2019MNRAS.489.4233M}
+($n_{\textrm{th}}=100\,\textrm{cm}^{-3}$) according to
+ref.\cite{2003MNRAS.339..289S} with a star-formation efficiency of $\epsilon =
+0.01$. Star formation converts gas cells into star particles which represent
+single stellar populations. For each star particle, the deposition of energy,
+momentum, and mass from stellar winds and supernovae is modeled.
+Photo-ionization and radiation pressure are modeled using an approximate
+treatment. A more detailed description of this model can be found in the
+flagship SMUGGLE paper.\cite{2019MNRAS.489.4233M}
 
@@
 $64$ to $512$. We found that a lower value of $N_{\textrm{ngb}}$ resulted in
-inefficient photoionization feedback since the photoionizing budget had not
+inefficient photo-ionization feedback since the photo-ionizing budget had not
 been exhausted after deposition into $64$ neighboring cells. We also used an
@@
 to machine precision.
+
+\vspace{12pt}
+
+\noindent
+{\bf Wind Model}
+\\
+\noindent
+
+In addition to the SMUGGLE model, we considered a simpler model of the interstellar medium based upon
 
@@
 time is to allow for the bar to form unimpacted by the presence of the gas
-(which would normally disrupt the formation of the bar). We made one
-additional modification when stitching the gas disk together with the N-body
-disk - we created a vacuum within the central $4\,\textrm{kpc}$. This vacuum
-guards against an initial dramatic infall of gas within the bar region, which
-we found to destroy the bar.
+(which would normally disrupt the formation of the bar). Throughout this work,
+we consider $t=0$ for the N-body run to be the time at which we added the gas
+phase for the SMUGGLE run (i.e. we ignore the first $1.5\,\textrm{Gyr}$ of
+evolution of the N-body disk when the bar is forming). We made one additional
+modification when stitching the gas disk together with the N-body disk - we
+created a vacuum within the central $4\,\textrm{kpc}$. This vacuum guards
+against an initial dramatic infall of gas within the bar region, which we
+found to destroy the bar.
 
@@
 The analysis of various bar properties is performed as follows. First, the
-pattern speed is measured from the angle of the second Fourier component. We measured the second Fourier component by computing,
+pattern speed is measured from the angle of the second Fourier component. We
+measured the second Fourier component by computing,
 \begin{equation}

### 2022-04-23 paper/ms.tex
Small changes to abstract

@@
   1984MNRAS.209..729T, 1985MNRAS.213..451W}. Previous work has overwhelmingly
-  provided the expectation that, due to the interaction between the bar and the halo, the bar pattern
+  provided the expectation that, due to the interaction between the bar and the dark matter halo, the bar pattern
   will slow its rotation over time \cite{1992ApJ...400...80H,
@@
   2006ApJ...637..214M, 2007MNRAS.375..460W, 2009ApJ...697..293D}. We have
-  performed a simulation of an isolated galactic disk hosting a strong bar
-  which includes a state-of-the-art model of the interstellar medium and a live dark matter halo. Here we
-  show that in this simulation the bar pattern does not slow down over time,
+  performed a simulation of a Milky Way-like galactic disk hosting a strong bar
+  which includes a state-of-the-art model of the interstellar medium and a live dark matter halo. In this simulation the bar pattern does not slow down over time,
   and instead remains at a stable, constant rate of rotation. This behavior
@@
 on the bar. However, because gas is collisional, it can participate in
-non-resonant angular momentum exchange with the bar. Thus, numerical work has
+non-resonant angular momentum exchange with the bar.\cite{2011MNRAS.415.1027H} Thus, numerical work has
 shown that the gas phase has an outsized influence on the

### 2022-04-23 paper/ms.tex
Update indentation

@@
   1984MNRAS.209..729T, 1985MNRAS.213..451W}. Previous work has overwhelmingly
-  provided the expectation that, due to the interaction between the bar and the dark matter halo, the bar pattern
-  will slow its rotation over time \cite{1992ApJ...400...80H,
-  2000ApJ...543..704D, 2002MNRAS.330...35A, 2002ApJ...569L..83A,
-  2003MNRAS.341.1179A, 2003MNRAS.346..251O, 2005MNRAS.363..991H,
-  2006ApJ...637..214M, 2007MNRAS.375..460W, 2009ApJ...697..293D}. We have
-  performed a simulation of a Milky Way-like galactic disk hosting a strong bar
-  which includes a state-of-the-art model of the interstellar medium and a live dark matter halo. In this simulation the bar pattern does not slow down over time,
-  and instead remains at a stable, constant rate of rotation. This behavior
-  has been observed in previous simulations but its explanation has remained
-  elusive.\cite{1993AA...268...65F, 2010ApJ...719.1470V}.
-We propose that the gas phase of the disk and the dark matter halo act in concert to stabilize the bar pattern and prevent it from slowing down or speeding up.
-This result
-  naturally explains why nearly all observed bars rotate quickly\cite{2011MSAIS..18...23C, 2015AA...576A.102A,
-2019MNRAS.482.1733G} and is especially relevant for
-  our understanding of how the Milky Way arrived at its present day state.
+  provided the expectation that, due to the interaction between the bar and
+  the dark matter halo, the bar pattern will slow its rotation over time
+  \cite{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
+  2002ApJ...569L..83A, 2003MNRAS.341.1179A, 2003MNRAS.346..251O,
+  2005MNRAS.363..991H, 2006ApJ...637..214M, 2007MNRAS.375..460W,
+  2009ApJ...697..293D}. We have performed a simulation of a Milky Way-like
+  galactic disk hosting a strong bar which includes a state-of-the-art model
+  of the interstellar medium and a live dark matter halo. In this simulation
+  the bar pattern does not slow down over time, and instead remains at a
+  stable, constant rate of rotation. This behavior has been observed in
+  previous simulations but its explanation has remained
+  elusive.\cite{1993AA...268...65F, 2010ApJ...719.1470V}. We propose that the
+  gas phase of the disk and the dark matter halo act in concert to stabilize
+  the bar pattern and prevent it from slowing down or speeding up. This result
+  naturally explains why nearly all observed bars rotate
+  quickly\cite{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G}
+  and is especially relevant for our understanding of how the Milky Way
+  arrived at its present day state.
 
@@
 bulge) has received considerable interest.\cite{1984MNRAS.209..729T,
-1985MNRAS.213..451W} A stellar disk in isolation seems to always
-generate a bar instability,\cite{1971ApJ...168..343H} but the presence of a strong spherical potential (e.g., from a stellar bulge or dark matter halo) acts to stabilize the disk against the bar
-instability.\cite{1973ApJ...186..467O, 1976AJ.....81...30H} A number of
-numerical simulations have confirmed that a live dark matter halo generically
-exerts a negative torque on the bar, causing its pattern speed to decrease and
-its length and strength to increase\cite{1992ApJ...400...80H,
-2000ApJ...543..704D, 2002MNRAS.330...35A, 2002ApJ...569L..83A,
-2003MNRAS.341.1179A, 2003MNRAS.346..251O, 2005MNRAS.363..991H,
-2006ApJ...637..214M, 2007MNRAS.375..460W, 2009ApJ...697..293D}. Intuitively,
-the bar excites a wake of resonant material in the dark matter halo in a
-manner similar to dynamical friction. This wake lags and thus exerts a
-negative torque on the bar.
+1985MNRAS.213..451W} A stellar disk in isolation seems to always generate a
+bar instability,\cite{1971ApJ...168..343H} but the presence of a strong
+spherical potential (e.g., from a stellar bulge or dark matter halo) acts to
+stabilize the disk against the bar instability.\cite{1973ApJ...186..467O,
+1976AJ.....81...30H} A number of numerical simulations have confirmed that a
+live dark matter halo generically exerts a negative torque on the bar, causing
+its pattern speed to decrease and its length and strength to
+increase\cite{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
+2002ApJ...569L..83A, 2003MNRAS.341.1179A, 2003MNRAS.346..251O,
+2005MNRAS.363..991H, 2006ApJ...637..214M, 2007MNRAS.375..460W,
+2009ApJ...697..293D}. Intuitively, the bar excites a wake of resonant material
+in the dark matter halo in a manner similar to dynamical friction. This wake
+lags and thus exerts a negative torque on the bar.
 
-The preeminent expectation that the dark matter halo acts to slow the bar
-down conflicts with observational estimates of bar pattern speeds. Bar
-rotation rates are typically classified by the dimensionless ratio
+The preeminent expectation that the dark matter halo acts to slow the bar down
+conflicts with observational estimates of bar pattern speeds. Bar rotation
+rates are typically classified by the dimensionless ratio
 \begin{equation}
@@
 where \RCR{} is the radius of corotation and $R_b$ is the length of the
-bar.\footnote{The radius of corotation \RCR{} is defined for circular orbits as
-the radius at which the orbital frequency is equal to the pattern speed,
+bar.\footnote{The radius of corotation \RCR{} is defined for circular orbits
+as the radius at which the orbital frequency is equal to the pattern speed,
 $\Omega_p$, of a given non-axisymmetric feature. In a galaxy with a constant
@@
 on the bar. However, because gas is collisional, it can participate in
-non-resonant angular momentum exchange with the bar.\cite{2011MNRAS.415.1027H} Thus, numerical work has
-shown that the gas phase has an outsized influence on the
-bar.\cite{2010ApJ...719.1470V, 2013MNRAS.429.1949A}
+non-resonant angular momentum exchange with the bar.\cite{2011MNRAS.415.1027H}
+Thus, numerical work has shown that the gas phase has an outsized influence on
+the bar.\cite{2010ApJ...719.1470V, 2013MNRAS.429.1949A}
 
@@
 the pattern speed of the bar to change with time (though our work does not
-alter expectations for radial mixing induced by spiral arms\cite{2002MNRAS.336..785S}).
+alter expectations for radial mixing induced by spiral
+arms\cite{2002MNRAS.336..785S}).
 
 Barred galaxies in cosmological simulations of galaxy formation continue to be
-in conflict with observations by producing bars which rotate too slowly.\cite{2017MNRAS.469.1054A, 2019MNRAS.483.2721P,
-2021AA...650L..16F} However, the pattern speeds of bars in both cosmological
-simulations and the real universe can be affected by environmental processes
-not included in our simulation -- e.g., satellite
-infall\cite{2011Natur.477..301P}, non-sphericity\cite{2013MNRAS.429.1949A} or
-rotation\cite{2013MNRAS.434.1287S, 2014ApJ...783L..18L, 2018MNRAS.476.1331C,
-2019MNRAS.488.5788C} in the dark matter halo, or perhaps even the gaseous
-circumgalactic medium. Naturally, extending our present work to account for
-such effects is a crucial next step in understanding the formation and
-evolution of galactic bars.
+in conflict with observations by producing bars which rotate too
+slowly.\cite{2017MNRAS.469.1054A, 2019MNRAS.483.2721P, 2021AA...650L..16F}
+However, the pattern speeds of bars in both cosmological simulations and the
+real universe can be affected by environmental processes not included in our
+simulation -- e.g., satellite infall\cite{2011Natur.477..301P},
+non-sphericity\cite{2013MNRAS.429.1949A} or rotation\cite{2013MNRAS.434.1287S,
+2014ApJ...783L..18L, 2018MNRAS.476.1331C, 2019MNRAS.488.5788C} in the dark
+matter halo, or perhaps even the gaseous circumgalactic medium. Naturally,
+extending our present work to account for such effects is a crucial next step
+in understanding the formation and evolution of galactic bars.
 
@@
 \noindent
-{\bf Wind Model}
+{\bf Smooth Interstellar Medium Model}
 \\
@@
 \end{equation}
-where $m_i$ and $\phi_i$ are the mass and azimuthal angle of each particle, respectively. We
-computed $A_2$ and $A_0$ in cylindrical bins of width $0.5\,\textrm{kpc}$ from radii of
-$0$ to $30\,\textrm{kpc}$. We defined the angle of the bar $\phi_b$ to be
-twice the angle of the complex number $A_2$ as measured in the bin extending from
-a radius of $2.5$ to $3\,\textrm{kpc}$. After correcting for the periodicity
-of $\phi_b$, we measured the pattern speed as the two-sided finite gradient of
-$\phi_b$ as a function of time.
+where $m_i$ and $\phi_i$ are the mass and azimuthal angle of each particle,
+respectively. We computed $A_2$ and $A_0$ in cylindrical bins of width
+$0.5\,\textrm{kpc}$ from radii of $0$ to $30\,\textrm{kpc}$. We defined the
+angle of the bar $\phi_b$ to be twice the angle of the complex number $A_2$ as
+measured in the bin extending from a radius of $2.5$ to $3\,\textrm{kpc}$.
+After correcting for the periodicity of $\phi_b$, we measured the pattern
+speed as the two-sided finite gradient of $\phi_b$ as a function of time.
 

### 2022-04-23 paper/ms.tex
Add GFM and fg sections
Add sections (in progress) about the SH03 model and the varying gas fractions

@@
 
+We performed two additional tests to explore how our findings depend on the
+assumed model of the interstellar medium and the amount of gas present in the
+system. First, we performed a simulation of the same disk but with a simpler
+model of the interstellar medium\cite{2003MNRAS.339..289S}, closer to standard
+methods used in cosmological simulations of galaxy formation. We find that the
+pattern speed evolution is nearly the same in this case, and so conclude that
+our result is not sensitive to the details of the model for the interstellar
+medium. The details of this experiment is given in the Methods section.
+
+Second, we performed a set of simulations in which we decreased the initial
+gas mass of the disk. We essentially find that there is a certain critical gas
+fraction (maybe gas surface density?) above which our proposed mechanism
+operates. If the gas available is below this threshold, then the pattern speed
+will decrease in a similar fashion to standard N-body simulations.
+Importantly, arbitrarily increasing the gas fraction does not cause the
+pattern speed to continuously increase. This is consistent with our proposed
+equilibrium mechanism and with previous studies of barred galaxies which
+varied the gas fraction.\cite{2010ApJ...719.1470V} The exact critical gas
+fraction above which our mechanism operates is likely dependent on the
+properties of the bar, disk, and dark matter halo -- further work is needed to
+characterize how this gas fraction varies with such properties. Details on
+this test in which we varied the gas mass of the disk is given in the Methods
+section.
+
 The implications of this finding are numerous. First, we naturally explain why
@@
 in understanding the formation and evolution of galactic bars.
+
+We found that below a certain gas fraction, bars should still be able to slow
+down. Therefore, we expect barred spiral galaxies which are exceptionally
+gas-poor to be rotating very slowly. We therefore predict that observations
+which target such galaxies (e.g., lenticular barred
+galaxies\cite{2009ARAA..47..159B}) would find slowly rotating bars. In
+general, we predict that all gas-rich spiral galaxies should rotate quickly
+while gas-poor spiral galaxies should rotate slowly.
 
@@
 
-In addition to the SMUGGLE model, we considered a simpler model of the interstellar medium based upon
+In addition to the SMUGGLE model, we considered a simpler model of the
+interstellar medium based upon ref.\cite{2003MNRAS.339..289S} In this model,
+the multiphase nature of the interstellar medium is modeled in a subgrid
+manner by allowing each resolution element to have a ``cold'' and ``hot''
+component, with the equation of state of the gas suitably modified. Gas is
+allowed to interchange between the ``cold'' and ``hot'' components through
+processes such as cooling and stellar feedback. Cold gas is allowed to undergo
+star formation. The pattern speed evolution of this setup is given in Extended
+Data Fig.~X, which shows that the pattern speed evolves in a similar manner to
+the SMUGGLE case shown in Fig.~\ref{fig:prop}.
 
@@
 \end{figure}
+
+\vspace{12pt}
+
+\noindent
+{\bf Varying Gas Fraction}
+\\
+\noindent
+To be written.
 
@@
 
-\vspace{12pt}
-
-\noindent
-{\bf Title}
-\\
-\noindent
-Next section.
-
 % \bibliography{ref2}

### 2022-04-25 paper/ms.tex
Add Lars' edits

@@
 
-  Elongated bar-like features are ubiquitous, occurring at the centers of
-  approximately two-thirds of spiral disk galaxies \cite{2000AJ....119..536E,
+  Elongated bar-like features are ubiquitous in galaxies, occurring at the centers of
+  approximately two-thirds of spiral disks \cite{2000AJ....119..536E,
   2007ApJ...657..790M}.  Due to gravitational interactions between the bar and
   the other components of galaxies, it is expected that angular momentum and
-  matter will redistribute between galactic components over long (Gyr)
-  timescales in galaxies hosting a bar \cite{1972MNRAS.157....1L,
-  1984MNRAS.209..729T, 1985MNRAS.213..451W}. Previous work has overwhelmingly
-  provided the expectation that, due to the interaction between the bar and
-  the dark matter halo, the bar pattern will slow its rotation over time
+  matter will redistribute over long (Gyr)
+  timescales in barred galaxies \cite{1972MNRAS.157....1L,
+  1984MNRAS.209..729T, 1985MNRAS.213..451W}. Previous work ignoring the gas phase of galaxies has conclusively 
+  demonstrated that bars should slow their rotation over time 
+  due to their interaction with dark matter halos
   \cite{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
@@
   stable, constant rate of rotation. This behavior has been observed in
-  previous simulations but its explanation has remained
-  elusive.\cite{1993AA...268...65F, 2010ApJ...719.1470V}. We propose that the
+  previous simulations using more simplified models for the 
+  interstellar gas but its explanation has remained
+  elusive\cite{1993AA...268...65F, 2010ApJ...719.1470V}. We propose that the
   gas phase of the disk and the dark matter halo act in concert to stabilize
-  the bar pattern and prevent it from slowing down or speeding up. This result
+  the bar pattern speed and prevent a bar from slowing down or speeding up. This result
   naturally explains why nearly all observed bars rotate
-  quickly\cite{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G}
+  rapidly\cite{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G}
   and is especially relevant for our understanding of how the Milky Way
-  arrived at its present day state.
+  arrived at its present state.
 
@@
 bulge) has received considerable interest.\cite{1984MNRAS.209..729T,
-1985MNRAS.213..451W} A stellar disk in isolation seems to always generate a
-bar instability,\cite{1971ApJ...168..343H} but the presence of a strong
+1985MNRAS.213..451W} Stellar disks in isolation are prone to developing bars, \cite{1971ApJ...168..343H} but the presence of a strong
 spherical potential (e.g., from a stellar bulge or dark matter halo) acts to
-stabilize the disk against the bar instability.\cite{1973ApJ...186..467O,
-1976AJ.....81...30H} A number of numerical simulations have confirmed that a
-live dark matter halo generically exerts a negative torque on the bar, causing
+stabilize the disk against bar formation.\cite{1973ApJ...186..467O,
+1976AJ.....81...30H} Many numerical simulations have confirmed that a
+live dark matter halo generically exerts a negative torque on a bar, causing
 its pattern speed to decrease and its length and strength to
@@
 2009ApJ...697..293D}. Intuitively, the bar excites a wake of resonant material
-in the dark matter halo in a manner similar to dynamical friction. This wake
-lags and thus exerts a negative torque on the bar.
+in a dark matter halo in a manner akin to dynamical friction. This wake
+lags a bar and thus exerts a negative torque on it.
 
-The preeminent expectation that the dark matter halo acts to slow the bar down
-conflicts with observational estimates of bar pattern speeds. Bar rotation
+The expectation that a dark matter halo acts to slow down a bar is in
+conflict with observational estimates of bar pattern speeds. Bar rotation
 rates are typically classified by the dimensionless ratio
@@
 non-resonant angular momentum exchange with the bar.\cite{2011MNRAS.415.1027H}
-Thus, numerical work has shown that the gas phase has an outsized influence on
-the bar.\cite{2010ApJ...719.1470V, 2013MNRAS.429.1949A}
+Thus, numerical work has shown that the gas phase can have a stronger influence on
+a bar than its contribution to the mass of a galaxy would suggest.\cite{2010ApJ...719.1470V, 2013MNRAS.429.1949A}
 
@@
 respectively). We modify this setup to include a gas phase, the details of
-which is given in the Methods section.
+which are provided in the Methods section.
 
@@
 that in the N-body case, the bar lengthens in time and grows in strength. In
-the SMUGGLE case, the bar pattern retains a similar length and strength
+the SMUGGLE case, the bar retains a similar length and strength
 between panels.
@@
 and SMUGGLE (orange) runs. The pattern speed in the N-body case slows down
-while the pattern speed in the SMUGGLE case remains constant. The slowing down
+while the pattern speed in the SMUGGLE case remains roughly constant. The slowing down
 of the pattern speed in the N-body case is consistent with a long line of
@@
 Fig.~\ref{fig:overview}, the length of the bar in the N-body case grows over
-time while it remains constant in the SMUGGLE case. This is also consistent
+time while it remains roughly constant in the SMUGGLE case. This is also consistent
 with previous numerical work, which found that bars tend to grow as they slow
@@
 location of the resonances in phase space changes, allowing halo material
-newly resonant with the bar to form a new wake. However, the gas is a reliable
+newly resonant with the bar to participate in forming the  wake. However, 
+the gas is a reliable
 source of positive torque on the bar, speeding the bar up. This stops the
-resonance location from changing such that the halo cannot form a new wake,
+resonance location from changing such that the halo cannot reinforce a wake,
 arresting the process by which the halo can slow the bar down.
@@
 
-The presence of the gas can arrest the process by which the dark matter halo
-wake forms. However, this does not explain why the pattern speed in the
+The presence of the gas can arrest the process by which additional material in
+the dark matter halo can contribute to a
+wake. However, this does not explain why the pattern speed in the
 SMUGGLE case is nearly constant over several Gyr. Naively, it would be a
@@
 our result is not sensitive to the details of the model for the interstellar
-medium. The details of this experiment is given in the Methods section.
+medium. The details of this experiment are given in the Methods section.
 
@@
 characterize how this gas fraction varies with such properties. Details on
-this test in which we varied the gas mass of the disk is given in the Methods
+this test in which we varied the gas mass of the disk are given in the Methods
 section.
 
-The implications of this finding are numerous. First, we naturally explain why
+The implications of our findings are numerous. First, we naturally explain why
 nearly all observed galaxies are fast rotators without requiring the inner
 regions of dark matter halos to be underdense\cite{1998ApJ...493L...5D,
-2000ApJ...543..704D} or introducing new physics.\cite{2021MNRAS.503.2833R,
+2000ApJ...543..704D} or requiring new physics.\cite{2021MNRAS.503.2833R,
 2021MNRAS.508..926R} Second, we show that the role of gas is of paramount

### 2022-04-25 paper/ms.tex
Formatting updates after Lars edits

@@
 
-  Elongated bar-like features are ubiquitous in galaxies, occurring at the centers of
-  approximately two-thirds of spiral disks \cite{2000AJ....119..536E,
-  2007ApJ...657..790M}.  Due to gravitational interactions between the bar and
-  the other components of galaxies, it is expected that angular momentum and
-  matter will redistribute over long (Gyr)
+  Elongated bar-like features are ubiquitous in galaxies, occurring at the
+  centers of approximately two-thirds of spiral
+  disks\cite{2000AJ....119..536E, 2007ApJ...657..790M}.  Due to gravitational
+  interactions between the bar and the other components of galaxies, it is
+  expected that angular momentum and matter will redistribute over long (Gyr)
   timescales in barred galaxies \cite{1972MNRAS.157....1L,
-  1984MNRAS.209..729T, 1985MNRAS.213..451W}. Previous work ignoring the gas phase of galaxies has conclusively 
-  demonstrated that bars should slow their rotation over time 
-  due to their interaction with dark matter halos
+  1984MNRAS.209..729T, 1985MNRAS.213..451W}. Previous work ignoring the gas
+  phase of galaxies has conclusively demonstrated that bars should slow their
+  rotation over time due to their interaction with dark matter halos
   \cite{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
@@
   stable, constant rate of rotation. This behavior has been observed in
-  previous simulations using more simplified models for the 
-  interstellar gas but its explanation has remained
-  elusive\cite{1993AA...268...65F, 2010ApJ...719.1470V}. We propose that the
-  gas phase of the disk and the dark matter halo act in concert to stabilize
-  the bar pattern speed and prevent a bar from slowing down or speeding up. This result
-  naturally explains why nearly all observed bars rotate
-  rapidly\cite{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G}
-  and is especially relevant for our understanding of how the Milky Way
-  arrived at its present state.
+  previous simulations using more simplified models for the interstellar gas
+  but its explanation has remained elusive\cite{1993AA...268...65F,
+  2010ApJ...719.1470V}. We propose that the gas phase of the disk and the dark
+  matter halo act in concert to stabilize the bar pattern speed and prevent a
+  bar from slowing down or speeding up. This result naturally explains why
+  nearly all observed bars rotate rapidly\cite{2011MSAIS..18...23C,
+  2015AA...576A.102A, 2019MNRAS.482.1733G} and is especially relevant for our
+  understanding of how the Milky Way arrived at its present state.
 
@@
 bulge) has received considerable interest.\cite{1984MNRAS.209..729T,
-1985MNRAS.213..451W} Stellar disks in isolation are prone to developing bars, \cite{1971ApJ...168..343H} but the presence of a strong
-spherical potential (e.g., from a stellar bulge or dark matter halo) acts to
-stabilize the disk against bar formation.\cite{1973ApJ...186..467O,
-1976AJ.....81...30H} Many numerical simulations have confirmed that a
-live dark matter halo generically exerts a negative torque on a bar, causing
-its pattern speed to decrease and its length and strength to
-increase\cite{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
-2002ApJ...569L..83A, 2003MNRAS.341.1179A, 2003MNRAS.346..251O,
-2005MNRAS.363..991H, 2006ApJ...637..214M, 2007MNRAS.375..460W,
-2009ApJ...697..293D}. Intuitively, the bar excites a wake of resonant material
-in a dark matter halo in a manner akin to dynamical friction. This wake
-lags a bar and thus exerts a negative torque on it.
+1985MNRAS.213..451W} Stellar disks in isolation are prone to developing bars,
+\cite{1971ApJ...168..343H} but the presence of a strong spherical potential
+(e.g., from a stellar bulge or dark matter halo) acts to stabilize the disk
+against bar formation.\cite{1973ApJ...186..467O, 1976AJ.....81...30H} Many
+numerical simulations have confirmed that a live dark matter halo generically
+exerts a negative torque on a bar, causing its pattern speed to decrease and
+its length and strength to increase\cite{1992ApJ...400...80H,
+2000ApJ...543..704D, 2002MNRAS.330...35A, 2002ApJ...569L..83A,
+2003MNRAS.341.1179A, 2003MNRAS.346..251O, 2005MNRAS.363..991H,
+2006ApJ...637..214M, 2007MNRAS.375..460W, 2009ApJ...697..293D}. Intuitively,
+the bar excites a wake of resonant material in a dark matter halo in a manner
+akin to dynamical friction. This wake lags a bar and thus exerts a negative
+torque on it.
 
@@
 non-resonant angular momentum exchange with the bar.\cite{2011MNRAS.415.1027H}
-Thus, numerical work has shown that the gas phase can have a stronger influence on
-a bar than its contribution to the mass of a galaxy would suggest.\cite{2010ApJ...719.1470V, 2013MNRAS.429.1949A}
+Thus, numerical work has shown that the gas phase can have a stronger
+influence on a bar than its contribution to the mass of a galaxy would
+suggest.\cite{2010ApJ...719.1470V, 2013MNRAS.429.1949A}
 
@@
 run while the lower three panels show the disk in the SMUGGLE run. Each column
-shows the disk $\sim1\,\textrm{Gyr}$ apart in time. There is a large qualitative
-difference in the evolution of the bar pattern between the two runs. We see
-that in the N-body case, the bar lengthens in time and grows in strength. In
-the SMUGGLE case, the bar retains a similar length and strength
+shows the disk $\sim1\,\textrm{Gyr}$ apart in time. There is a large
+qualitative difference in the evolution of the bar pattern between the two
+runs. We see that in the N-body case, the bar lengthens in time and grows in
+strength. In the SMUGGLE case, the bar retains a similar length and strength
 between panels.
@@
 and SMUGGLE (orange) runs. The pattern speed in the N-body case slows down
-while the pattern speed in the SMUGGLE case remains roughly constant. The slowing down
-of the pattern speed in the N-body case is consistent with a long line of
-numerical research on bars in N-body simulations.\cite{1992ApJ...400...80H,
-2000ApJ...543..704D, 2002MNRAS.330...35A, 2002ApJ...569L..83A,
-2003MNRAS.341.1179A, 2003MNRAS.346..251O, 2005MNRAS.363..991H,
-2006ApJ...637..214M, 2007MNRAS.375..460W, 2009ApJ...697..293D} However, in the
-SMUGGLE case the pattern speed remains constant. After the first Gyr of
-evolution, we find that the pattern speed increases by only $\sim10\%$ over
-the next $4\,\textrm{Gyr}$, compared to a $\sim43\%$ decrease in the pattern
-speed for the N-body run over the same interval. As we saw qualitatively in
+while the pattern speed in the SMUGGLE case remains roughly constant. The
+slowing down of the pattern speed in the N-body case is consistent with a long
+line of numerical research on bars in N-body
+simulations.\cite{1992ApJ...400...80H, 2000ApJ...543..704D,
+2002MNRAS.330...35A, 2002ApJ...569L..83A, 2003MNRAS.341.1179A,
+2003MNRAS.346..251O, 2005MNRAS.363..991H, 2006ApJ...637..214M,
+2007MNRAS.375..460W, 2009ApJ...697..293D} However, in the SMUGGLE case the
+pattern speed remains constant. After the first Gyr of evolution, we find that
+the pattern speed increases by only $\sim10\%$ over the next
+$4\,\textrm{Gyr}$, compared to a $\sim43\%$ decrease in the pattern speed for
+the N-body run over the same interval. As we saw qualitatively in
 Fig.~\ref{fig:overview}, the length of the bar in the N-body case grows over
-time while it remains roughly constant in the SMUGGLE case. This is also consistent
-with previous numerical work, which found that bars tend to grow as they slow
-down and the radius of corotation increases.\cite{2000ApJ...543..704D,
-2003MNRAS.341.1179A}
+time while it remains roughly constant in the SMUGGLE case. This is also
+consistent with previous numerical work, which found that bars tend to grow as
+they slow down and the radius of corotation
+increases.\cite{2000ApJ...543..704D, 2003MNRAS.341.1179A}
 
@@
 location of the resonances in phase space changes, allowing halo material
-newly resonant with the bar to participate in forming the  wake. However, 
-the gas is a reliable
-source of positive torque on the bar, speeding the bar up. This stops the
-resonance location from changing such that the halo cannot reinforce a wake,
-arresting the process by which the halo can slow the bar down.
+newly resonant with the bar to participate in forming the wake. However, the
+gas is a reliable source of positive torque on the bar, speeding the bar up.
+This stops the resonance location from changing such that the halo cannot
+reinforce a wake, arresting the process by which the halo can slow the bar
+down.
 
@@
 The presence of the gas can arrest the process by which additional material in
-the dark matter halo can contribute to a
-wake. However, this does not explain why the pattern speed in the
-SMUGGLE case is nearly constant over several Gyr. Naively, it would be a
-coincidence that the bar pattern speed remains constant in the SMUGGLE case,
-resulting from a chance cancellation of the halo and gas torques. However, a
-constant pattern speed in the presence of gas has been observed in a few
-simulations of barred galaxies with gas.\cite{1993AA...268...65F,
-2010ApJ...719.1470V} Previous work has argued this is due to the bar torquing
-gas inwards, but no explanation has been given for why it might remain
-constant.
+the dark matter halo can contribute to a wake. However, this does not explain
+why the pattern speed in the SMUGGLE case is nearly constant over several Gyr.
+Naively, it would be a coincidence that the bar pattern speed remains constant
+in the SMUGGLE case, resulting from a chance cancellation of the halo and gas
+torques. However, a constant pattern speed in the presence of gas has been
+observed in a few simulations of barred galaxies with
+gas.\cite{1993AA...268...65F, 2010ApJ...719.1470V} Previous work has argued
+this is due to the bar torquing gas inwards, but no explanation has been given
+for why it might remain constant.
 

### 2022-04-25 paper/ms.tex
Fix footnotes and figure labels

@@
 % \usepackage{tgbonum}
+\usepackage[symbol]{footmisc}
 
@@
 
-\let\thefootnote\relax\footnote{
+{\let\thefootnote\relax\footnote{
 
@@
 \end{affiliations}
-}
+}}
+
+% \renewcommand{\thefootnote}{\fnsymbol{footnote}}
 
@@
 \setcounter{table}{0}
-\captionsetup[figure]{labelformat=empty}
 \renewcommand{\figurename}{Extended Data Figure}
@@
 modification when stitching the gas disk together with the N-body disk - we
-created a vacuum within the central $4\,\textrm{kpc}$. This vacuum guards
+created a hole within the central $4\,\textrm{kpc}$. This hole guards
 against an initial dramatic infall of gas within the bar region, which we

### 2022-04-26 paper/ms.tex
Add ref to UGC 628
Many cosmetic changes
Moving figures around, adding authors, acknowledgements, etc

@@
 \usepackage[symbol]{footmisc}
+\usepackage{placeins}
 
@@
 \usepackage{xcolor}
+\usepackage[labelfont=bf]{caption}
 
@@
 
-\author{Angus Beane$^{1*}$, Lars Hernquist$^1$, Elena D'Onghia$^{2,3}$, et al.}
+\author{Angus Beane$^{1*}$, Lars Hernquist$^1$, Elena D'Onghia$^{2,3}$,
+Federico Marinacci$^{4}$, Charlie Conroy$^{1}$, Jia Qi$^{5}$, Laura V.
+Sales$^{6}$, Paul Torrey$^{5}$ and Mark Vogelsberger$^{7}$}
 
@@
 \begin{affiliations}
-\item Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA
+% 1
+\item Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA.
 
-\item Department of Physics, University of Wisconsin-Madison, Madison, WI, USA
+% 2
+\item Department of Physics, University of Wisconsin-Madison, Madison, WI, USA.
 
-\item Department of Astronomy, University of Wisconsin-Madison, Madison, WI, USA
+% 3
+\item Department of Astronomy, University of Wisconsin-Madison, Madison, WI, USA.
+
+% 4
+\item Department of Physics \& Astronomy `Augusto Righi', University of Bologna, Bologna, Italy.
+
+% 5
+\item Department of Astronomy, University of Florida, Gainesville, FL, USA.
+
+% 6
+\item Department of Physics \& Astronomy, University of California, Riverside, CA, USA.
+
+% 7
+\item Department of Physics, Massachusetts Institute of Technology, Cambridge, MA, USA.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure*}[h!]%
+\centering
+\includegraphics[width=0.8\textwidth]{fig/fig1.pdf}
+\caption{Surface density projections in an N-body only simulation (upper
+panels) and a simulation which includes the SMUGGLE model (bottom panels). We
+can see that in the N-body run, the bar grows in length and strength. In the
+SMUGGLE run (which includes a gaseous phase and a model for the multi-phase
+interstellar medium), the bar remains at approximately the same length and
+strength over the course of the simulation. The N-body model is identical to
+the GALAKOS model, discussed in the text.}\label{fig:overview}
+\end{figure*}
+
+\begin{figure}[h!]%
+\centering
+\includegraphics[width=0.45\textwidth]{fig/fig2.pdf}
+\caption{Various bar properties from the N-body and SMUGGLE runs. The
+\textit{upper panel} show the evolution of the pattern speed. As expected, the
+bar in the N-body run slows down due to interactions between the bar and the
+dark matter halo. However, the bar in the SMUGGLE run does not slow down and
+instead remains at a constant pattern speed. The \textit{middle panel} shows
+the evolution of the bar length. In the N-body case, the bar lengthens. This
+occurs because as the pattern speed drops, bar-like orbits at larger radii are
+possible. Stars are captured on these orbits, lengthening the bar. This
+process does not occur in the SMUGGLE cases since the bar pattern speed is not
+decreasing, and therefore the bar length remains constant. The \textit{lower
+panel} shows the torque on the bar by different components. The solid lines
+show the torque exerted by the halo in both the N-body and SMUGGLE cases. The
+dashed line shows the torque exerted by the gas phase in the SMUGGLE run
+(there is no gas in the N-body run). Details on how these properties
+are calculated is given in the Methods section.}\label{fig:prop}
+\end{figure}
+
+\begin{figure*}[h!]%
+\centering
+\includegraphics[width=0.9\textwidth]{fig/fig3.pdf}
+\caption{The wake excited in the dark matter halo in the N-body case
+(\textit{left panel}) and SMUGGLE case (\textit{right panel}) after
+$2.6\,\textrm{Gyr}$. The \textit{left} and \textit{right panels} show a
+surface density projection in the $x$-$y$ plane of the dark matter halo after
+an axisymmetric average has been subracted. The solid line indicates the
+direction of the bar while the dashed line indicates the direction of the halo
+wake (both measured by taking the second Fourier component within a sphere of
+all material within a radius of $X\,\textrm{kpc}$). The \textit{center panel}
+shows the time evolution of the angle difference between the bar and the halo
+wake, as measured from the second Fourier component. After the first Gyr, the
+angle difference in the SMUGGLE case is smaller than in the N-body case by
+about a factor of two, reflecting how the dark matter halo in the SMUGGLE case
+is unable to exert as negative a torque on the bar as in the N-body
+case.}\label{fig:wake}
+\end{figure*}
+
+\begin{figure}[h!]
+\centering
+\includegraphics{fig/fig4.pdf}
+\caption{The torque on the bar by the gas when the pattern speed of the bar is
+kept at a constant value. Only gas within the corotation radius is able to
+infall. Since slower bars have larger corotation radii, slower bars experience
+a larger net torque than faster bars. The setup of the simulations used here
+is identical to the SMUGGLE case discussed earlier and in the Methods section,
+except the N-body disk is rotated as a solid body with a constant angular
+velocity.}\label{fig:equil}
+\end{figure}
 
@@
 its effect of slowing down the bar.\cite{2021MNRAS.500.4710C,
-2021MNRAS.505.2412C} Third, we provide an explanation for how the Milky Way's
-bar could be both long-lived and a fast rotator, of which there is some
+2021MNRAS.505.2412C} Third
+ed and a fast rotator, of which there is some
 observational evidence.\cite{2019MNRAS.490.4740B} And finally, we complicate
@@
 which target such galaxies (e.g., lenticular barred
-galaxies\cite{2009ARAA..47..159B}) would find slowly rotating bars. In
-general, we predict that all gas-rich spiral galaxies should rotate quickly
-while gas-poor spiral galaxies should rotate slowly.
-
-\begin{figure*}[h]%
-\centering
-\includegraphics[width=0.8\textwidth]{fig/fig1.pdf}
-\caption{Surface density projections in an N-body only simulation (upper
-panels) and a simulation which includes the SMUGGLE model (bottom panels). We
-can see that in the N-body run, the bar grows in length and strength. In the
-SMUGGLE run (which includes a gaseous phase and a model for the multi-phase
-interstellar medium), the bar remains at approximately the same length and
-strength over the course of the simulation. The N-body model is identical to
-the GALAKOS model, discussed in the text.}\label{fig:overview}
-\end{figure*}
-
-\begin{figure}[h]%
-\centering
-\includegraphics[width=0.45\textwidth]{fig/fig2.pdf}
-\caption{Various bar properties from the N-body and SMUGGLE runs. The
-\textit{upper panel} show the evolution of the pattern speed. As expected, the
-bar in the N-body run slows down due to interactions between the bar and the
-dark matter halo. However, the bar in the SMUGGLE run does not slow down and
-instead remains at a constant pattern speed. The \textit{middle panel} shows
-the evolution of the bar length. In the N-body case, the bar lengthens. This
-occurs because as the pattern speed drops, bar-like orbits at larger radii are
-possible. Stars are captured on these orbits, lengthening the bar. This
-process does not occur in the SMUGGLE cases since the bar pattern speed is not
-decreasing, and therefore the bar length remains constant. The \textit{lower
-panel} shows the torque on the bar by different components. The solid lines
-show the torque exerted by the halo in both the N-body and SMUGGLE cases. The
-dashed line shows the torque exerted by the gas phase in the SMUGGLE run
-(there is no gas in the N-body run). Details on how these properties
-are calculated is given in the Methods section.}\label{fig:prop}
-\end{figure}
-
-\begin{figure*}[h]%
-\centering
-\includegraphics[width=0.9\textwidth]{fig/fig3.pdf}
-\caption{The wake excited in the dark matter halo in the N-body case
-(\textit{left panel}) and SMUGGLE case (\textit{right panel}) after
-$2.6\,\textrm{Gyr}$. The \textit{left} and \textit{right panels} show a
-surface density projection in the $x$-$y$ plane of the dark matter halo after
-an axisymmetric average has been subracted. The solid line indicates the
-direction of the bar while the dashed line indicates the direction of the halo
-wake (both measured by taking the second Fourier component within a sphere of
-all material within a radius of $X\,\textrm{kpc}$). The \textit{center panel}
-shows the time evolution of the angle difference between the bar and the halo
-wake, as measured from the second Fourier component. After the first Gyr, the
-angle difference in the SMUGGLE case is smaller than in the N-body case by
-about a factor of two, reflecting how the dark matter halo in the SMUGGLE case
-is unable to exert as negative a torque on the bar as in the N-body
-case.}\label{fig:wake}
-\end{figure*}
-
-\begin{figure}[h]
-\centering
-\includegraphics{fig/fig4.pdf}
-\caption{The torque on the bar by the gas when the pattern speed of the bar is
-kept at a constant value. Only gas within the corotation radius is able to
-infall. Since slower bars have larger corotation radii, slower bars experience
-a larger net torque than faster bars. The setup of the simulations used here
-is identical to the SMUGGLE case discussed earlier and in the Methods section,
-except the N-body disk is rotated as a solid body with a constant angular
-velocity.}\label{fig:equil}
-\end{figure}
+galaxies\cite{2009ARAA..47..159B}) would find slowly rotating bars. There does
+exist one such example of a galaxy which is known to be a slow rotator -- UGC
+628.\cite{2009AA...499L..25C} This galaxy has been studied in detail by
+ref.\cite{2016MNRAS.463.1751C}, who note that it indeed has a low gas
+fraction. In general, we predict that all gas-rich spiral galaxies should
+rotate quickly while gas-poor spiral galaxies should rotate slowly.
 
@@
 
-\item [Acknowledgements] Foo.
+\item [Acknowledgements] We would like to thank Greg L. Bryan, Drummond B.
+Fielding, Keith Hawkins, Jason A. S. Hunt, Sarah M. R. Jeffreson, Kathryn V.
+Johnston, Julio F. Navarro, David N. Spergel, Martin D. Weinberg, and Yanfei
+Zou for helpful discussions. A.B. would like to thank Todd Phillips for
+helpful discussions. Some computations in this paper were run on the FASRC
+Cannon cluster supported by the FAS Division of Science Research Computing
+Group at Harvard University. Resources supporting this work were also provided
+by the NASA High-End Computing (HEC) Program through the NASA Advanced
+Supercomputing (NAS) Division at Ames Research Center. A.B. was supported by
+the Future Investigators in NASA Earth and Space Science and Technology
+(FINESST) award number 80NSSC20K1536 during the completion of this work. L.H.
+funding. E.D. funding. F.M. funding. C.C. funding. J.Q. funding. L.V.S.
+funding. P.T. funding. M.V. funding.
 
@@
 
-  \item[Code Availability] Foo.
+  \item[Code Availability] A public version of the AREPO code is available at
+  \url{https://arepo-code.org/}.
 
@@
 
+\FloatBarrier
 \newpage
@@
 
-\begin{center}
+\noindent
 {\bf \Large \uppercase{Methods} }
-\end{center}
+
+\bigskip
 
@@
 \noindent
-
 In addition to the SMUGGLE model, we considered a simpler model of the
@@
 
+\begin{figure}[t]%
+\centering
+\includegraphics[width=0.45\textwidth]{fig/fig-vcirc.pdf}
+\caption{The circular velocity curve for the N-body run (blue) and the SMUGGLE
+run (orange) compared to observational estimates.\cite{2019ApJ...871..120E} We
+see that the circular velocity curve for both runs is marginally larger than
+the Milky Way's, but still comparable. The SMUGGLE circular velocity curve is
+larger than the N-body curve due to the additional mass in the gas phase.}
+\label{fig:vcirc}
+\end{figure}
+
 Our setup is initially out of equilibrium, but we found that after
@@
 
-\begin{figure}[h]%
-\centering
-\includegraphics[width=0.45\textwidth]{fig/fig-vcirc.pdf}
-\caption{The circular velocity curve for the N-body run (blue) and the SMUGGLE
-run (orange) compared to observational estimates.\cite{2019ApJ...871..120E} We
-see that the circular velocity curve for both runs is marginally larger than
-the Milky Way's, but still comparable. The SMUGGLE circular velocity curve is
-larger than the N-body curve due to the additional mass in the gas phase.}
-\label{fig:vcirc}
-\end{figure}
-
-\begin{figure}[h]%
+\begin{figure}[t!]%
 \centering
@@
 
-The time evolution of the bar strength, defined as the maximum of
-$\left|A_2/A_0\right|$ as a function of radius, is shown in Extended Data
-Fig.~\ref{fig:strength}. The quantity $\left|A_2/A_0\right|$ varies from $0$
-to $1$, with larger values indicating a stronger bar pattern. We see that in
-the N-body case, $\left|A_2/A_0\right|$ increases over time as the bar pattern
-slows. This is consistent with previous N-body simulations which showed a
-clear correlation between the bar pattern speed and the bar
-strength.\cite{2003MNRAS.341.1179A} In the SMUGGLE case, we see that the bar
-strength has an initial drop but then remains at a roughly constant but
-slightly decreasing strength. This is consistent with the pattern speed in the
-SMUGGLE case being roughly constant or slightly increasing.
-
-\begin{figure}[h]%
+\begin{figure}[t!]%
 \centering
@@
 \end{figure}
+
+\begin{figure*}[t!]%
+\centering
+\includegraphics[width=0.95\textwidth]{fig/fig-bar_decomp.pdf}
+\caption{Demonstration of our bar decomposition procedure based on
+ref.\cite{2016MNRAS.463.1952P} The \textit{left panel} shows a surface density
+projection through the stellar component of the N-body simulation (disk and
+bulge). The \textit{middle panel} shows the component of the disk identified
+as being trapped in the bar while the \textit{right panel} shows the component
+of the disk identified as not being trapped in the bar. The fact that the
+untrapped stars form a roughly axisymmetric structure indicates our bar
+decomposition is sufficiently accurate. {\color{red} I will remake this figure
+- I ended up losing the files which stored the bar membership, so I just
+pulled these screenshots from an earlier presentation.}}
+\label{fig:decomp}
+\end{figure*}
+
+The time evolution of the bar strength, defined as the maximum of
+$\left|A_2/A_0\right|$ as a function of radius, is shown in Extended Data
+Fig.~\ref{fig:strength}. The quantity $\left|A_2/A_0\right|$ varies from $0$
+to $1$, with larger values indicating a stronger bar pattern. We see that in
+the N-body case, $\left|A_2/A_0\right|$ increases over time as the bar pattern
+slows. This is consistent with previous N-body simulations which showed a
+clear correlation between the bar pattern speed and the bar
+strength.\cite{2003MNRAS.341.1179A} In the SMUGGLE case, we see that the bar
+strength has an initial drop but then remains at a roughly constant but
+slightly decreasing strength. This is consistent with the pattern speed in the
+SMUGGLE case being roughly constant or slightly increasing.
 
@@
 
-\begin{figure*}[h]%
-\centering
-\includegraphics[width=0.95\textwidth]{fig/fig-bar_decomp.pdf}
-\caption{Demonstration of our bar decomposition procedure based on
-ref.\cite{2016MNRAS.463.1952P} The \textit{left panel} shows a surface density
-projection through the stellar component of the N-body simulation (disk and
-bulge). The \textit{middle panel} shows the component of the disk identified
-as being trapped in the bar while the \textit{right panel} shows the component
-of the disk identified as not being trapped in the bar. The fact that the
-untrapped stars form a roughly axisymmetric structure indicates our bar
-decomposition is sufficiently accurate. {\color{red} I will remake this figure
-- I ended up losing the files which stored the bar membership, so I just
-pulled these screenshots from an earlier presentation.}}
-\label{fig:decomp}
-\end{figure*}
-
 After the disk has been decomposed into a trapped and untrapped component, we

### 2022-04-26 plots/paper_fig-surf/plot_fig-surf.py
Update surface density plot

@@ -40,6 +40,7 @@ def read_snap(idx, name, lvl, parttype=[2, 3, 4], fields=['Coordinates', 'Masses
 def compute_surface_density(R, mass, Rbins):
     key = np.where(R < Rbins[-1])[0]
     R = R[key]
+    mass = mass[key]
     
     surf_dens = np.zeros(len(Rbins)-1)
     ave_R = np.zeros(len(Rbins)-1)
@@ -117,6 +118,8 @@ def run():
         Rbins = np.logspace(-2, np.log10(20), 25)
 
         mass = sn.part0.mass.value
+        if i==0:
+            r = np.linalg.norm(pos, axis=1)
 
         aveR, surf = compute_surface_density(R, mass, Rbins)
 
@@ -134,11 +137,15 @@ def run():
         key = np.where(sfr > 0)[0]
         Rbins = np.logspace(-2, np.log10(20), 15)
         aveR, surf = compute_surface_density(R[key], sfr[key], Rbins)
+
+        surf[np.logical_or(np.isnan(surf), surf==0.0)] = 1E-9
         print(surf)
 
         ax[1].plot(aveR, (1E9 / 1E6) * surf, label=label_1, c=tb_c[i]) # convert to Msun/Gyr/pc^2
 
-    
+    ax[0].axhline(10, c='k', ls='dashed')
+    ax[1].axhline(10.**(0.5), c='k', ls='dashed')
+
     ax[0].set(yscale='log')
     ax[1].set(yscale='log')
     ax[1].set(xlim=(0, 10))

### 2022-04-26 paper/ms.tex
Replace N-body with macro

@@
 \newcommand{\Rot}{\ensuremath{\mathcal{R}}}
+
+\newcommand{\Nbody}{$N$-body}
 
@@
 \includegraphics[width=0.8\textwidth]{fig/fig1.pdf}
-\caption{Surface density projections in an N-body only simulation (upper
+\caption{Surface density projections in an \Nbody{} only simulation (upper
 panels) and a simulation which includes the SMUGGLE model (bottom panels). We
-can see that in the N-body run, the bar grows in length and strength. In the
+can see that in the \Nbody{} run, the bar grows in length and strength. In the
 SMUGGLE run (which includes a gaseous phase and a model for the multi-phase
 interstellar medium), the bar remains at approximately the same length and
-strength over the course of the simulation. The N-body model is identical to
+strength over the course of the simulation. The \Nbody{} model is identical to
 the GALAKOS model, discussed in the text.}\label{fig:overview}
@@
 \includegraphics[width=0.45\textwidth]{fig/fig2.pdf}
-\caption{Various bar properties from the N-body and SMUGGLE runs. The
+\caption{Various bar properties from the \Nbody{} and SMUGGLE runs. The
 \textit{upper panel} show the evolution of the pattern speed. As expected, the
-bar in the N-body run slows down due to interactions between the bar and the
+bar in the \Nbody{} run slows down due to interactions between the bar and the
 dark matter halo. However, the bar in the SMUGGLE run does not slow down and
 instead remains at a constant pattern speed. The \textit{middle panel} shows
-the evolution of the bar length. In the N-body case, the bar lengthens. This
+the evolution of the bar length. In the \Nbody{} case, the bar lengthens. This
 occurs because as the pattern speed drops, bar-like orbits at larger radii are
@@
 panel} shows the torque on the bar by different components. The solid lines
-show the torque exerted by the halo in both the N-body and SMUGGLE cases. The
+show the torque exerted by the halo in both the \Nbody{} and SMUGGLE cases. The
 dashed line shows the torque exerted by the gas phase in the SMUGGLE run
-(there is no gas in the N-body run). Details on how these properties
+(there is no gas in the \Nbody{} run). Details on how these properties
 are calculated is given in the Methods section.}\label{fig:prop}
@@
 \includegraphics[width=0.9\textwidth]{fig/fig3.pdf}
-\caption{The wake excited in the dark matter halo in the N-body case
+\caption{The wake excited in the dark matter halo in the \Nbody{} case
 (\textit{left panel}) and SMUGGLE case (\textit{right panel}) after
@@
 wake, as measured from the second Fourier component. After the first Gyr, the
-angle difference in the SMUGGLE case is smaller than in the N-body case by
+angle difference in the SMUGGLE case is smaller than in the \Nbody{} case by
 about a factor of two, reflecting how the dark matter halo in the SMUGGLE case
-is unable to exert as negative a torque on the bar as in the N-body
+is unable to exert as negative a torque on the bar as in the \Nbody{}
 case.}\label{fig:wake}
@@
 is identical to the SMUGGLE case discussed earlier and in the Methods section,
-except the N-body disk is rotated as a solid body with a constant angular
+except the \Nbody{} disk is rotated as a solid body with a constant angular
 velocity.}\label{fig:equil}
@@
 A surface density projection of our simulations is shown in
-Fig.~\ref{fig:overview}. The upper three panels show the disk in the N-body
+Fig.~\ref{fig:overview}. The upper three panels show the disk in the \Nbody{}
 run while the lower three panels show the disk in the SMUGGLE run. Each column
@@
 qualitative difference in the evolution of the bar pattern between the two
-runs. We see that in the N-body case, the bar lengthens in time and grows in
+runs. We see that in the \Nbody{} case, the bar lengthens in time and grows in
 strength. In the SMUGGLE case, the bar retains a similar length and strength
@@
 We show the time evolution of different bar properties in Fig.~\ref{fig:prop}.
-In the upper panel, we show the pattern speed over time in the N-body (blue)
-and SMUGGLE (orange) runs. The pattern speed in the N-body case slows down
+In the upper panel, we show the pattern speed over time in the \Nbody{} (blue)
+and SMUGGLE (orange) runs. The pattern speed in the \Nbody{} case slows down
 while the pattern speed in the SMUGGLE case remains roughly constant. The
-slowing down of the pattern speed in the N-body case is consistent with a long
-line of numerical research on bars in N-body
+slowing down of the pattern speed in the \Nbody{} case is consistent with a long
+line of numerical research on bars in \Nbody{}
 simulations.\cite{1992ApJ...400...80H, 2000ApJ...543..704D,
@@
 $4\,\textrm{Gyr}$, compared to a $\sim43\%$ decrease in the pattern speed for
-the N-body run over the same interval. As we saw qualitatively in
-Fig.~\ref{fig:overview}, the length of the bar in the N-body case grows over
+the \Nbody{} run over the same interval. As we saw qualitatively in
+Fig.~\ref{fig:overview}, the length of the bar in the \Nbody{} case grows over
 time while it remains roughly constant in the SMUGGLE case. This is also
@@
 dark matter halo whereas the dashed line indicates the torque on the bar by
-the gas phase. In the N-body case, the halo exerts a steady negative torque on
+the gas phase. In the \Nbody{} case, the halo exerts a steady negative torque on
 the bar, with an average torque from $1$ to $4\,\textrm{Gyr}$ of $-58.0$ in
@@
 positive torque on the bar can be understood in terms of the halo wake
-mechanism. In the N-body case, halo material which is resonant with the bar
+mechanism. In the \Nbody{} case, halo material which is resonant with the bar
 will form a wake which lags behind and exerts a negative torque on the bar,
@@
 is plotted in the middle panel of Fig.~\ref{fig:wake}, which shows that the
-angle offset is larger in the N-body case than in the SMUGGLE case by about a
+angle offset is larger in the \Nbody{} case than in the SMUGGLE case by about a
 factor of two. The left and right panels of Fig.~\ref{fig:wake} show the halo
-wake with respect to the location of the bar in the N-body (left) and SMUGGLE
+wake with respect to the location of the bar in the \Nbody{} (left) and SMUGGLE
 (right) cases.
@@
 operates. If the gas available is below this threshold, then the pattern speed
-will decrease in a similar fashion to standard N-body simulations.
+will decrease in a similar fashion to standard \Nbody{} simulations.
 Importantly, arbitrarily increasing the gas fraction does not cause the
@@
 and a stellar bulge and dark matter halo following a Hernquist
-profile.\cite{1990ApJ...356..359H} All N-body runs in this work used the same
+profile.\cite{1990ApJ...356..359H} All \Nbody{} runs in this work used the same
 setup parameters as the GALAKOS disk, more details of which can be found in
@@
 After generating the gaseous disk in this way, we stitched the gas disk
-together with the GALAKOS N-body disk (and bulge and dark matter halo) after
+together with the GALAKOS \Nbody{} disk (and bulge and dark matter halo) after
 the GALAKOS disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The
@@
 (which would normally disrupt the formation of the bar). Throughout this work,
-we consider $t=0$ for the N-body run to be the time at which we added the gas
+we consider $t=0$ for the \Nbody{} run to be the time at which we added the gas
 phase for the SMUGGLE run (i.e. we ignore the first $1.5\,\textrm{Gyr}$ of
-evolution of the N-body disk when the bar is forming). We made one additional
-modification when stitching the gas disk together with the N-body disk - we
+evolution of the \Nbody{} disk when the bar is forming). We made one additional
+modification when stitching the gas disk together with the \Nbody{} disk - we
 created a hole within the central $4\,\textrm{kpc}$. This hole guards
@@
 \includegraphics[width=0.45\textwidth]{fig/fig-vcirc.pdf}
-\caption{The circular velocity curve for the N-body run (blue) and the SMUGGLE
+\caption{The circular velocity curve for the \Nbody{} run (blue) and the SMUGGLE
 run (orange) compared to observational estimates.\cite{2019ApJ...871..120E} We
@@
 the Milky Way's, but still comparable. The SMUGGLE circular velocity curve is
-larger than the N-body curve due to the additional mass in the gas phase.}
+larger than the \Nbody{} curve due to the additional mass in the gas phase.}
 \label{fig:vcirc}
@@
 Fourier component divided by the zeroth Fourier component (a formula is given
-in the Methods section). We see that in the N-body case (blue) the bar
+in the Methods section). We see that in the \Nbody{} case (blue) the bar
 strength increases with time, consistent with previous results showing that
@@
 ref.\cite{2016MNRAS.463.1952P} The \textit{left panel} shows a surface density
-projection through the stellar component of the N-body simulation (disk and
+projection through the stellar component of the \Nbody{} simulation (disk and
 bulge). The \textit{middle panel} shows the component of the disk identified
@@
 to $1$, with larger values indicating a stronger bar pattern. We see that in
-the N-body case, $\left|A_2/A_0\right|$ increases over time as the bar pattern
-slows. This is consistent with previous N-body simulations which showed a
+the \Nbody{} case, $\left|A_2/A_0\right|$ increases over time as the bar pattern
+slows. This is consistent with previous \Nbody{} simulations which showed a
 clear correlation between the bar pattern speed and the bar

### 2022-04-26 plots/paper_fig-A2/plot_fig-A2.py
Standardize paper figures

@@ -12,7 +12,7 @@ time_conv = 977.79222168
 
 mpl.use('Agg')
 
-rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
 rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 
@@ -27,8 +27,6 @@ tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
 # names
 Nbody = 'Nbody'
 phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
-
-
 
 lvl = 'lvl3'
 
@@ -62,11 +60,12 @@ def moving_average(a, n=3) :
     return ret[n - 1:] / n
 
 def run():
-    fig, ax = plt.subplots(1, 1, figsize=(3, 3))
+    cm = 1/2.54
+    fig, ax = plt.subplots(1, 1, figsize=(9*cm, 7*cm))
     
     fourier = read_fourier(Nbody, lvl)
     t, A2A0 = extract_t_max_A2A0(fourier)
-    ax.plot(t-t[300], A2A0, c=tb_c[0], label='N-body')
+    ax.plot(t-t[300], A2A0, c=tb_c[0], label=r'$N$-body')
 
     fourier = read_fourier(phS2R35, lvl)
     t, A2A0 = extract_t_max_A2A0(fourier)

### 2022-04-26 plots/paper_fig-rot/plot_fig-rot.py

@@ -155,8 +155,8 @@ def run():
         RCR = compute_RCR(agama_pot_SMUGGLE[i], bar_prop_SMUGGLE['pattern_speed'][i])
         RCR_SMUGGLE.append(RCR)
 
-
-    fig, ax = plt.subplots(1, 1, sharex=True, figsize=(3.5, 3.5))
+    cm = 1/2.54
+    fig, ax = plt.subplots(1, 1, sharex=True, figsize=(9*cm, 9*cm))
 
     
 
@@ -173,7 +173,7 @@ def run():
     ax.plot(bar_prop_SMUGGLE['tlist'], rot_SMUGGLE, c=tb_c[1], label='SMUGGLE')#, ls='dashed')
 
     ax.legend(frameon=False)
-    ax.set(ylim=(0, 3), ylabel=r'$\mathcal{R}$', xlabel=r'$t\,[\,\textrm{Gyr}\,]$', xlim=(0, 5))
+    ax.set(ylim=(1.2, 2), ylabel=r'$\mathcal{R}$', xlabel=r'$t\,[\,\textrm{Gyr}\,]$', xlim=(0, 5))
 
     fig.tight_layout()
     fig.savefig('fig-rot.pdf')

### 2022-04-26 plots/paper_fig-surf/plot_fig-surf.py

@@ -14,7 +14,7 @@ from scipy.ndimage import gaussian_filter
 
 mpl.use('Agg')
 
-rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
 rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 
@@ -106,7 +106,8 @@ def compute_dphi(phi_disk, phi_halo):
     return dphi
 
 def run():
-    fig, ax = plt.subplots(2, 1, sharex=True, figsize=(3, 5))
+    cm = 1/2.54
+    fig, ax = plt.subplots(2, 1, sharex=True, figsize=(9*cm, 12*cm))
     idx_list = [0, 100, 200, 400, 600, 800]
     center = np.array([200., 200., 200.])
 
@@ -123,7 +124,7 @@ def run():
 
         aveR, surf = compute_surface_density(R, mass, Rbins)
 
-        if i < 3:
+        if True:
             label_0 = repr(sn.Time.value)
             label_1 = None
         else:
@@ -157,8 +158,8 @@ def run():
     ax[1].set(ylabel=r'$\Sigma_{\textrm{SFR}}\,[\,M_{\odot} / \textrm{Gyr} / \textrm{pc}^2\,]$')
     ax[1].set(xlabel=r'$R\,[\,\textrm{kpc}\,]$')
 
-    ax[0].legend(frameon=False, title=r'$t\,[\,\text{Gyr}\,]$')
-    ax[1].legend(frameon=False)
+    ax[0].legend(frameon=False, title=r'$t\,[\,\text{Gyr}\,]$', ncol=2)
+    # ax[1].legend(frameon=False)
 
     fig.tight_layout()
     fig.savefig('fig-surf.pdf')

### 2022-04-26 plots/paper_fig-vcirc/plot_fig-vcirc.py

@@ -17,7 +17,7 @@ agama.setUnits(mass=1E10, length=1, velocity=1)
 
 mpl.use('Agg')
 
-rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
 rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 
@@ -74,11 +74,12 @@ def compute_vcirc(pot):
     return Rlist, np.sqrt(vcsq)
 
 def run():
-    fig, ax = plt.subplots(1, 1, figsize=(3, 3))
+    cm = 1/2.54
+    fig, ax = plt.subplots(1, 1, figsize=(9*cm, 7*cm))
 
     pot = read_agama_pot(500, Nbody, lvl)
     R, vc = compute_vcirc(pot)
-    ax.plot(R, vc, c=tb_c[0], label='N-body')
+    ax.plot(R, vc, c=tb_c[0], label=r'$N$-body')
 
     pot = read_agama_pot(200, phS2R35, lvl)
     R, vc = compute_vcirc(pot)

### 2022-04-26 plots/paper_fig1/plot_fig1.py

@@ -7,7 +7,7 @@ import matplotlib as mpl
 from matplotlib import rc
 mpl.use('Agg')
 
-rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
 rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 
@@ -109,7 +109,9 @@ def run():
 
     extent = [rng[0][0], rng[0][1], rng[1][0], rng[1][1]]
 
-    fig, ax = plt.subplots(2, 3, sharex=True, sharey=True)
+    cm = 1/2.54
+
+    fig, ax = plt.subplots(2, 3, sharex=True, sharey=True, figsize=(18*cm, 12*cm))
 
     for i in range(len(Nbody_idx)):
         # plot Nbody
@@ -132,7 +134,7 @@ def run():
     ax[0][1].set_title(r'$t=2\,\textrm{Gyr}$')
     ax[0][2].set_title(r'$t=3\,\textrm{Gyr}$')
 
-    ax[0][0].set_ylabel('N-body')
+    ax[0][0].set_ylabel(r'$N$-body')
     ax[1][0].set_ylabel('SMUGGLE')
 
     fig.tight_layout()

### 2022-04-26 plots/paper_fig2/plot_fig2.py

@@ -10,7 +10,7 @@ from scipy.signal import savgol_filter
 
 mpl.use('Agg')
 
-rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
 rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 
@@ -105,12 +105,16 @@ def run():
     bar_prop_Nbody = read_bar_prop(Nbody, lvl)
     bar_prop_SMUGGLE = read_bar_prop(phS2R35, lvl)
 
-    fig, ax = plt.subplots(3, 1, sharex=True, figsize=(3.5, 6))
+    cm = 1/2.54
+    fig, ax = plt.subplots(3, 1, sharex=True, figsize=(8*cm, 12*cm))
 
     # First panel, pattern speed.
     t300 = bar_prop_Nbody['tlist'][300]
-    ax[0].plot(bar_prop_Nbody['tlist'] - t300, bar_prop_Nbody['pattern_speed'], c=tb_c[0], label='N-body')
-    ax[0].plot(bar_prop_SMUGGLE['tlist'], bar_prop_SMUGGLE['pattern_speed'], c=tb_c[1], label='SMUGGLE')
+    ps_N = savgol_filter(bar_prop_Nbody['pattern_speed'], 81, 3)
+    ps_S = savgol_filter(bar_prop_SMUGGLE['pattern_speed'], 81, 3)
+
+    ax[0].plot(bar_prop_Nbody['tlist'] - t300, ps_N, c=tb_c[0], label=r'$N$-body')
+    ax[0].plot(bar_prop_SMUGGLE['tlist'], ps_S, c=tb_c[1], label='SMUGGLE')
 
     ax[0].set(ylim=(0, 60), ylabel=r'$\Omega_p\,[\,\text{km}/\text{s}/\text{kpc}\,]$')
     ax[0].set(xlim=(0, 5))
@@ -159,7 +163,7 @@ def run():
 
     custom_lines = [mpl.lines.Line2D([0], [0], color='k'),
                     mpl.lines.Line2D([0], [0], color='k', ls='dashed')]
-    ax[2].legend(custom_lines, ['halo', 'gas'], frameon=False)
+    ax[2].legend(custom_lines, ['by halo', 'by gas'], frameon=False)
 
 
     fig.tight_layout()
[…]

### 2022-04-26 plots/paper_fig3/plot_fig3.py

@@ -14,7 +14,7 @@ from scipy.ndimage import gaussian_filter
 
 mpl.use('Agg')
 
-rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
 rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 
@@ -153,7 +153,8 @@ def run():
     vmin = -0.002
     vmax = 0.002
 
-    fig, ax = plt.subplots(1, 3, figsize=(8, 2.5))
+    cm = 1/2.54
+    fig, ax = plt.subplots(1, 3, figsize=(18*cm, 6*cm))
 
     # First panel, wake of Nbody.
     sn = read_snap(Nbody_idx, Nbody, lvl, parttype=[1], fields=None)
@@ -180,13 +181,15 @@ def run():
     tN = fourier_Nbody['time']
     tS = fourier_SMUGGLE['time']
 
-    ax[1].plot(tN - tN[300], 180.*dphiN, c=tb_c[0])
-    ax[1].plot(tS, 180.*dphiS, c=tb_c[1])
+    ax[1].plot(tN - tN[300], 180.*dphiN, c=tb_c[0], label=r'$N$-body')
+    ax[1].plot(tS, 180.*dphiS, c=tb_c[1], label='SMUGGLE')
     ax[1].axhline(0.0, c='k')
 
-    ax[1].axvline(tS[520])
+    # ax[1].axvline(tS[520])
 
-    ax[1].set(xlim=(0, 5), ylim=(-35, 70), xlabel=r'$t\,[\,\text{Gyr}\,]$', ylabel=r'$\text{angle difference}\,[\,\text{deg}\,]$')
+    ax[1].set(xlim=(0, 5), ylim=(-20, 40), xlabel=r'$t\,[\,\text{Gyr}\,]$', ylabel=r'$\text{angle difference}\,[\,\text{deg}\,]$')
+
+    ax[1].legend(frameon=False)
 
     # Third panel, wake of SMUGGLE.
     sn = read_snap(SMUGGLE_idx, phS2R35, lvl, parttype=[1], fields=None)

### 2022-04-26 plots/paper_fig4/plot_fig4.py

@@ -10,7 +10,7 @@ from scipy.signal import savgol_filter
 
 mpl.use('Agg')
 
-rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
 rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 
@@ -65,7 +65,8 @@ def read_torque(name, lvl):
     return tlist, tz_halo, tz_not_bar, tz_gas
 
 def run():
-    fig, ax = plt.subplots(1, 1, sharex=True, figsize=(3.5, 3.5))
+    cm = 1/2.54
+    fig, ax = plt.subplots(1, 1, sharex=True, figsize=(9*cm, 7*cm))
 
     # First panel, pattern speed.
     tlist_30, tz_halo_30, _, tz_gas_30 = read_torque(phS2R35, 'lvl3-rot30')
@@ -88,7 +89,7 @@ def run():
 
     # ax.axhline(0, c='k', ls='dashed')
 
-    ax.set(xlim=(29, 41), ylim=(0, 30))
+    ax.set(xlim=(29, 41), ylim=(0, 25))
     ax.set(xticks=rot_list)
     ax.set(xlabel=r'$\Omega_p\,[\,\textrm{km}/\textrm{s}/\textrm{kpc}\,]$')
     ax.set(ylabel=r'$\left< \tau_{\textrm{on bar}} \right>\,[\,10^{10}\,M_{\odot}\,(\text{km}/\text{s})^2\,]$')

### 2022-04-26 paper/ms.tex
Update figures in paper
Makes them all standardized

@@
 \centering
-\includegraphics[width=0.8\textwidth]{fig/fig1.pdf}
+\includegraphics[width=18cm]{fig/fig1.pdf}
 \caption{Surface density projections in an \Nbody{} only simulation (upper
@@
 \centering
-\includegraphics[width=0.45\textwidth]{fig/fig2.pdf}
+\includegraphics[width=9cm]{fig/fig2.pdf}
 \caption{Various bar properties from the \Nbody{} and SMUGGLE runs. The
@@
 \centering
-\includegraphics[width=0.9\textwidth]{fig/fig3.pdf}
+\includegraphics[width=18cm]{fig/fig3.pdf}
 \caption{The wake excited in the dark matter halo in the \Nbody{} case
@@
 \centering
-\includegraphics{fig/fig4.pdf}
+\includegraphics[width=9cm]{fig/fig4.pdf}
 \caption{The torque on the bar by the gas when the pattern speed of the bar is
@@
 \centering
-\includegraphics[width=0.45\textwidth]{fig/fig-vcirc.pdf}
+\includegraphics[width=9cm]{fig/fig-vcirc.pdf}
 \caption{The circular velocity curve for the \Nbody{} run (blue) and the SMUGGLE
-run (orange) compared to observational estimates.\cite{2019ApJ...871..120E} We
-see that the circular velocity curve for both runs is marginally larger than
-the Milky Way's, but still comparable. The SMUGGLE circular velocity curve is
-larger than the \Nbody{} curve due to the additional mass in the gas phase.}
+run (orange) compared to observational estimates for the Milky
+Way.\cite{2019ApJ...871..120E} We see that the circular velocity curve for
+both runs is marginally larger than the Milky Way's, but still comparable. The
+SMUGGLE circular velocity curve is larger than the \Nbody{} curve due to the
+additional mass in the gas phase.}
 \label{fig:vcirc}
@@
 \centering
-\includegraphics[width=0.45\textwidth]{fig/fig-surf.pdf}
+\includegraphics[width=9cm]{fig/fig-surf.pdf}
 \caption{The time evolution of the gas surface density (\textit{upper}) and
@@
 \centering
-\includegraphics[width=0.45\textwidth]{fig/fig-A2.pdf}
-\caption{The time evolution of the bar strength, measured as the second
+\includegraphics[width=9cm]{fig/fig-A2.pdf}
+\caption{The time evolution of the bar strength, measured as the maximum of the second
 Fourier component divided by the zeroth Fourier component (a formula is given
@@
 \centering
-\includegraphics[width=0.95\textwidth]{fig/fig-bar_decomp.pdf}
+\includegraphics[width=18cm]{fig/fig-bar_decomp.pdf}
 \caption{Demonstration of our bar decomposition procedure based on

### 2022-04-26 paper/ms.tex
Update surf dens plot with Evans data

@@
 \newcommand{\Nbody}{$N$-body}
+\newcommand{\Msun}{\ensuremath{M_{\odot}}}
 
@@
 (which would normally disrupt the formation of the bar). Throughout this work,
-we consider $t=0$ for the \Nbody{} run to be the time at which we added the gas
-phase for the SMUGGLE run (i.e. we ignore the first $1.5\,\textrm{Gyr}$ of
-evolution of the \Nbody{} disk when the bar is forming). We made one additional
-modification when stitching the gas disk together with the \Nbody{} disk - we
-created a hole within the central $4\,\textrm{kpc}$. This hole guards
-against an initial dramatic infall of gas within the bar region, which we
-found to destroy the bar.
+we consider $t=0$ for the \Nbody{} run to be the time at which we added the
+gas phase for the SMUGGLE run (i.e. we ignore the first $1.5\,\textrm{Gyr}$ of
+evolution of the \Nbody{} disk when the bar is forming). We made one
+additional modification when stitching the gas disk together with the \Nbody{}
+disk - we created a hole within the central $4\,\textrm{kpc}$. This hole
+guards against an initial dramatic infall of gas within the bar region, which
+we found to destroy the bar. Models of the Milky Way's gas surface density
+profile indicate a reduction in gas in the bar region from $\sim1$ to
+$\sim4\,\textrm{kpc}$,\cite{2022ApJ...929L..18E} and so our practice of
+allowing the gas distribution to have a hole in the central region is
+consistent with our choice to begin the simulations with a bar already formed.
 
@@
 results in a slightly higher circular velocity. We also show the evolution of
-the surface density profile in Extended Data Fig.~\ref{fig:surf} compared to
-observational estimates.\cite{2012ARAA..50..531K}
+the surface density profile in Extended Data Fig.~\ref{fig:surf} We find that
+in our simulation the atomic and molecular gas surface density and the SFR
+surface density is broadly consistent with the expected values for the Milky
+Way.\cite{2022ApJ...929L..18E} The discrepancy between $1$ and
+$4\,\textrm{kpc}$ is likely due to the fact that the distances to molecular
+cloud which underlines this work used a simple kinematic distance based on an
+axisymmetric model of the Milky Way,\cite{2017ApJ...834...57M} which is likely
+not accurate in the bar region where gas has large non-circular velocities.
 
@@
 \includegraphics[width=9cm]{fig/fig-surf.pdf}
-\caption{The time evolution of the gas surface density (\textit{upper}) and
+\caption{The time evolution of the molecular gas surface density (\textit{upper}) and
 the star formation rate (SFR) surface density (\textit{lower}) at various
 times during the simulation. Colored lines indicate the profiles at selected
-times during the simulation while the horizontal dashed lines indicate
-approximate values for the Milky Way.\cite{2012ARAA..50..531K} We see that the
-gas surface density is slightly higher but within an order of magnitude of the
-Milky Way's typical gas surface density. We see a sharp decrease in the gas
-and SFR surface density along the extent of the bar from $\sim1$ to
+times during the simulation while the black dashed lines indicate models for
+the Milky Way which allow the CO-to-H$_2$ conversion factor $X_{\textrm{CO}}$
+to vary with metallicity.\cite{2022ApJ...929L..18E} We see that the molecular
+gas and SFR surface densities are within an order of magnitude of the Milky
+Way's typical values at all times.  We see a sharp decrease in the gas and SFR
+surface densities along the extent of the bar from $\sim1$ to
 $\sim4\,\textrm{kpc}$, related to the gas inflow in this region. {\color{red}
-I will add the actual points from ref.\cite{2012ARAA..50..531K} (Fig. 7), but
-just placing approx horizontal lines for now. The profiles are roughly flat
-anyways.}}
+Waiting on atomic gas data from Juergen Kerp.}}
 \label{fig:surf}
@@
 \end{document}
+
+%

### 2022-04-28 paper/ms.tex
Add plotting details section

@@
 
+\vspace{12pt}
+
+\noindent
+{\bf Plotting Details}
+\\
+\noindent
+We saved snapshots in intervals of $0.005$ in the time units of the
+simulation, $\textrm{kpc}/(\textrm{km}/\textrm{s})$, which is very nearly
+equal to $1\,\textrm{Gyr}$ (it is $\sim0.977\,\textrm{Gyr}$). Therefore,
+throughout this work we referred to the native code time unit as
+$\textrm{Gyr}$. None of our results are sensitive to this choice.
+
+In order to remove numerical noise in several quantities we computed, we
+applied a Savitzky-Golay filter\cite{1964AnaCh..36.1627S} as implemented in
+\texttt{scipy} using a window length of $81$ and polynomial order of $3$. This
+filter was applied to plots of pattern speeds, bar lengths and strengths,
+torques, and angle differences.
+
 % \bibliography{ref2}

### 2022-04-28 paper/ms.tex
Add smooth ISM, fgas models

@@
 \FloatBarrier
-\newpage
+\clearpage
 
@@
 \bigskip
-
-\noindent
-{\bf SMUGGLE Model}
-\\
-\noindent
-We use the Stars and MUltiphase Gas in GaLaxiEs (SMUGGLE) model
-\cite{2019MNRAS.489.4233M} implemented within the moving-mesh, finite-volume
-hydrodynamics code AREPO \cite{2010MNRAS.401..791S}. The SMUGGLE model
-includes self-gravity, hydrodynamics, radiative heating and cooling, star
-formation, and stellar feedback. Explicit gas cooling and heating of the
-multi-phase interstellar medium is implemented, covering temperature ranges
-between $10$ and $10^8\,\textrm{K}$.
-
-Star formation occurs in cells above a density threshold
-($n_{\textrm{th}}=100\,\textrm{cm}^{-3}$) according to
-ref.\cite{2003MNRAS.339..289S} with a star-formation efficiency of $\epsilon =
-0.01$. Star formation converts gas cells into star particles which represent
-single stellar populations. For each star particle, the deposition of energy,
-momentum, and mass from stellar winds and supernovae is modeled.
-Photo-ionization and radiation pressure are modeled using an approximate
-treatment. A more detailed description of this model can be found in the
-flagship SMUGGLE paper.\cite{2019MNRAS.489.4233M}
-
-We used the fiducial model parameters, except that we increased the number of
-effective neighbors $N_{\textrm{ngb}}$ for the deposition of feedback from
-$64$ to $512$. We found that a lower value of $N_{\textrm{ngb}}$ resulted in
-inefficient photo-ionization feedback since the photo-ionizing budget had not
-been exhausted after deposition into $64$ neighboring cells. We also used an
-updated version of SMUGGLE using a new mechanical feedback routine similar to
-the one described in ref.\cite{2018MNRAS.480..800H} This updated routine is a
-tensor renormalization which ensures linear and angular momentum conservation
-to machine precision.
-
-\vspace{12pt}
-
-\noindent
-{\bf Smooth Interstellar Medium Model}
-\\
-\noindent
-In addition to the SMUGGLE model, we considered a simpler model of the
-interstellar medium based upon ref.\cite{2003MNRAS.339..289S} In this model,
-the multiphase nature of the interstellar medium is modeled in a subgrid
-manner by allowing each resolution element to have a ``cold'' and ``hot''
-component, with the equation of state of the gas suitably modified. Gas is
-allowed to interchange between the ``cold'' and ``hot'' components through
-processes such as cooling and stellar feedback. Cold gas is allowed to undergo
-star formation. The pattern speed evolution of this setup is given in Extended
-Data Fig.~X, which shows that the pattern speed evolves in a similar manner to
-the SMUGGLE case shown in Fig.~\ref{fig:prop}.
-
-\vspace{12pt}
 
@@
 
+\begin{figure}[h!]%
+\centering
+\includegraphics[width=9cm]{fig/fig-GFM.pdf}
+\caption{\textbf{Pattern speed evolution of a smooth ISM model.} This
+evolution is shown for the fiducial disk in the \Nbody{} (blue), SMUGGLE
+(orange), and smooth ISM (red) cases. The smooth ISM model is an older model
+for the ISM which treats its multiphase nature in a subgrid
+fashion.\cite{2003MNRAS.339..289S} This fundamentally differs from the SMUGGLE
+model, which explicitly resolve the hot and cold phases of the
+ISM.\cite{2019MNRAS.489.4233M} The pattern speed in the smooth ISM case is
+broadly similar to the evolution in the SMUGGLE case. This shows that the
+stability of the pattern speed is not simply a result of our assumed model for
+the ISM.}
+\label{fig:GFM}
+\end{figure}
+
 Our setup is initially out of equilibrium, but we found that after
@@
 We used a softening length of $0.02\,\textrm{kpc}$ for all components.
+Snapshots were saved at equal intervals of $0.005$ in the time units of the
+simulation ($\textrm{kpc}/(\textrm{km}/\textrm{s})\sim1\,\textrm{Gyr}$).
+
+\begin{figure}[t!]%
+\centering
+\includegraphics[width=9cm]{fig/fig-fgas.pdf}
+\caption{\textbf{Pattern speed evolution with varying gas fractions.} We
+explored the impact of lowering the initial gas surface density of our
+fiducial disk on the evolution of the pattern speed. The surface densities we
+tested of $20$, $15$, $10$, and $5\,\Msun/\textrm{kpc}^2$ correspond to
+initial gas fractions of $16\%$, $10\%$, $7\%$, and $4\%$. We find that
+initial surface densities $20$, $15$, and $10\,\Msun/\textrm{kpc}^2$ result in
+bars which remain at a constant pattern speed, while an initial surface
+density of $5\,\Msun/\textrm{kpc}^2$ results in a bar which slows down in
+roughly the same manner as the \Nbody{} case. }
+\label{fig:fgas}
+\end{figure}
 
@@
 \noindent
+{\bf SMUGGLE Model}
+\\
+\noindent
+We use the Stars and MUltiphase Gas in GaLaxiEs (SMUGGLE) model
+\cite{2019MNRAS.489.4233M} implemented within the moving-mesh, finite-volume
+hydrodynamics code AREPO \cite{2010MNRAS.401..791S}. The SMUGGLE model
+includes self-gravity, hydrodynamics, radiative heating and cooling, star
+formation, and stellar feedback. Explicit gas cooling and heating of the
+multi-phase interstellar medium is implemented, covering temperature ranges
+between $10$ and $10^8\,\textrm{K}$.
+
+Star formation occurs in cells above a density threshold
+($n_{\textrm{th}}=100\,\textrm{cm}^{-3}$) according to
+ref.\cite{2003MNRAS.339..289S} with a star-formation efficiency of $\epsilon =
+0.01$. Star formation converts gas cells into star particles which represent
+single stellar populations. For each star particle, the deposition of energy,
+momentum, and mass from stellar winds and supernovae is modeled.
+Photo-ionization and radiation pressure are modeled using an approximate
+treatment. A more detailed description of this model can be found in the
+flagship SMUGGLE paper.\cite{2019MNRAS.489.4233M}
+
+We used the fiducial model parameters, except that we increased the number of
+effective neighbors $N_{\textrm{ngb}}$ for the deposition of feedback from
+$64$ to $512$. We found that a lower value of $N_{\textrm{ngb}}$ resulted in
+inefficient photo-ionization feedback since the photo-ionizing budget had not
+been exhausted after deposition into $64$ neighboring cells. We also used an
+updated version of SMUGGLE using a new mechanical feedback routine similar to
+the one described in ref.\cite{2018MNRAS.480..800H} This updated routine is a
+tensor renormalization which ensures linear and angular momentum conservation
+to machine precision.
+
+\vspace{12pt}
+
+\noindent
+{\bf Smooth Interstellar Medium Model}
+\\
+\noindent
+In addition to the SMUGGLE model, we considered a simpler model of the
+interstellar medium based upon ref.\cite{2003MNRAS.339..289S} In this model,
+the multiphase nature of the interstellar medium is modeled in a subgrid
+manner by allowing each resolution element to have a ``cold'' and ``hot''
+component, with the equation of state of the gas suitably modified. Gas is
+allowed to interchange between the ``cold'' and ``hot'' components through
+processes such as cooling and stellar feedback. Cold gas is allowed to undergo
+star formation. The pattern speed evolution of this setup is given in Extended
+Data Fig.~\ref{fig:GFM}, which shows that the pattern speed evolves in a
+similar manner to the SMUGGLE case shown in Fig.~\ref{fig:prop}.
+
+\vspace{12pt}
+
+\noindent
 {\bf Varying Gas Fraction}
@@
 \noindent
-To be written.
+We also performed a test in which we varied the initial gas fraction of the
+disk. In our fiducial run, we set the surface density of the gas disk from
+$4\,\textrm{kpc}$ to $\sim9.3\,\textrm{kpc}$ to be $20\,\Msun/\textrm{kpc}^2$.
+We also ran with surface densities of $15$, $10$, and
+$5\,\Msun/\textrm{kpc}^2$. These correspond to initial gas fractions of
+approximately $16\%$, $10\%$, $7\%$, and $4\%$. The pattern speed evolution is
+shown in Extended Data Fig.~\ref{fig:fgas} We find that the bar in disks with
+initial surface densities of $20$, $15$, and $10\,\Msun/\textrm{kpc}$ evolve
+with a constant pattern speed while the bar in a disk with initial surface
+density of $5\,\Msun/\textrm{kpc}^2$ slows down at a similar rate to the \Nbody{} case.
+
+As a result, we conclude that for the disk, bar, and halo properties
+considered in this work, a gas fraction of only approximately $5\%$ is
+necessary in order for the proposed stabilizing mechanism to operate. Furthermore, we co
 

### 2022-04-28 paper/ms.tex
Cosmetic changes, add surf data

@@
 \usepackage[labelfont=bf]{caption}
+\DeclareCaptionLabelSeparator{lsep}{ | } 
+\captionsetup{labelsep=lsep}
 
@@
 
-\author{Angus Beane$^{1*}$, Lars Hernquist$^1$, Elena D'Onghia$^{2,3}$,
-Federico Marinacci$^{4}$, Charlie Conroy$^{1}$, Jia Qi$^{5}$, Laura V.
-Sales$^{6}$, Paul Torrey$^{5}$ and Mark Vogelsberger$^{7}$}
+\author{Angus~Beane$^{1*}$, Lars~Hernquist$^1$, Elena~D'Onghia$^{2,3}$,
+Federico~Marinacci$^{4}$, Charlie~Conroy$^{1}$, Jia~Qi$^{5}$,
+Laura~V.~Sales$^{6}$, Paul~Torrey$^{5}$ and Mark~Vogelsberger$^{7}$
+}
 
@@
   matter halo act in concert to stabilize the bar pattern speed and prevent a
-  bar from slowing down or speeding up. This result naturally explains why
-  nearly all observed bars rotate rapidly\cite{2011MSAIS..18...23C,
-  2015AA...576A.102A, 2019MNRAS.482.1733G} and is especially relevant for our
-  understanding of how the Milky Way arrived at its present state.
+  bar from slowing down or speeding up. We find that in a Milky Way-like disk,
+  a gas fraction of only about \boldmath$5\%$ is necessary for this mechanism to operate.
+  This result naturally explains why nearly all observed bars rotate
+  rapidly\cite{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G}
+  and is especially relevant for our understanding of how the Milky Way
+  arrived at its present state.
 
@@
 \includegraphics[width=18cm]{fig/fig1.pdf}
-\caption{Surface density projections in an \Nbody{} only simulation (upper
-panels) and a simulation which includes the SMUGGLE model (bottom panels). We
-can see that in the \Nbody{} run, the bar grows in length and strength. In the
-SMUGGLE run (which includes a gaseous phase and a model for the multi-phase
-interstellar medium), the bar remains at approximately the same length and
-strength over the course of the simulation. The \Nbody{} model is identical to
-the GALAKOS model, discussed in the text.}\label{fig:overview}
+\caption{\textbf{Projected surface densities of simulated barred galaxies with
+and without the interstellar medium.} The upper panels show an \Nbody{} only
+simulation while the lower panels show a simulation which includes the SMUGGLE
+model for the ISM. Columns show different points in time, separated by
+$1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
+length and strength. In the SMUGGLE run, the bar remains at approximately the
+same length and strength over the course of the simulation. The \Nbody{} model
+is identical to the GALAKOS model, discussed in the text.}\label{fig:overview}
 \end{figure*}
@@
 \includegraphics[width=9cm]{fig/fig2.pdf}
-\caption{Various bar properties from the \Nbody{} and SMUGGLE runs. The
-\textit{upper panel} show the evolution of the pattern speed. As expected, the
-bar in the \Nbody{} run slows down due to interactions between the bar and the
-dark matter halo. However, the bar in the SMUGGLE run does not slow down and
-instead remains at a constant pattern speed. The \textit{middle panel} shows
-the evolution of the bar length. In the \Nbody{} case, the bar lengthens. This
-occurs because as the pattern speed drops, bar-like orbits at larger radii are
-possible. Stars are captured on these orbits, lengthening the bar. This
-process does not occur in the SMUGGLE cases since the bar pattern speed is not
-decreasing, and therefore the bar length remains constant. The \textit{lower
-panel} shows the torque on the bar by different components. The solid lines
-show the torque exerted by the halo in both the \Nbody{} and SMUGGLE cases. The
-dashed line shows the torque exerted by the gas phase in the SMUGGLE run
-(there is no gas in the \Nbody{} run). Details on how these properties
-are calculated is given in the Methods section.}\label{fig:prop}
+\caption{\textbf{Properties of the bar in our \boldmath$N$-body and SMUGGLE
+simulations of a barred galaxy.} The \textit{upper panel} shows the evolution
+of the pattern speed. As expected, the bar in the \Nbody{} run slows down due
+to interactions between the bar and the dark matter halo. However, the bar in
+the SMUGGLE run does not slow down and instead remains at a constant pattern
+speed. The \textit{middle panel} shows the evolution of the bar length. In the
+\Nbody{} case, the bar lengthens. This occurs because as the pattern speed
+drops, bar-like orbits at larger radii are possible. Stars are captured on
+these orbits, lengthening the bar. This process does not occur in the SMUGGLE
+cases since the bar pattern speed is not decreasing, and therefore the bar
+length remains constant. The \textit{lower panel} shows the torque on the bar
+by different components. The solid lines show the torque exerted by the halo
+in both the \Nbody{} and SMUGGLE cases. The dashed line shows the torque
+exerted by the gas phase in the SMUGGLE run (there is no gas in the \Nbody{}
+run). Details on how these properties are calculated is given in the Methods
+section.}\label{fig:prop}
 \end{figure}
@@
 \includegraphics[width=18cm]{fig/fig3.pdf}
-\caption{The wake excited in the dark matter halo in the \Nbody{} case
-(\textit{left panel}) and SMUGGLE case (\textit{right panel}) after
-$2.6\,\textrm{Gyr}$. The \textit{left} and \textit{right panels} show a
-surface density projection in the $x$-$y$ plane of the dark matter halo after
-an axisymmetric average has been subracted. The solid line indicates the
-direction of the bar while the dashed line indicates the direction of the halo
-wake (both measured by taking the second Fourier component within a sphere of
-all material within a radius of $X\,\textrm{kpc}$). The \textit{center panel}
-shows the time evolution of the angle difference between the bar and the halo
-wake, as measured from the second Fourier component. After the first Gyr, the
-angle difference in the SMUGGLE case is smaller than in the \Nbody{} case by
-about a factor of two, reflecting how the dark matter halo in the SMUGGLE case
-is unable to exert as negative a torque on the bar as in the \Nbody{}
-case.}\label{fig:wake}
+\caption{\textbf{The wake excited in the dark matter halo.} The dark matter
+halo wake is shown in the \Nbody{} case (\textit{left panel}) and SMUGGLE case
+(\textit{right panel}) after $2.6\,\textrm{Gyr}$ of evolution. The
+\textit{left} and \textit{right panels} show a surface density projection in
+the $x$-$y$ plane of the dark matter halo after an axisymmetric average has
+been subracted. The solid line indicates the direction of the bar while the
+dashed line indicates the direction of the halo wake (both measured by taking
+the second Fourier component within a sphere of all material within a radius
+of $4\,\textrm{kpc}$). The \textit{center panel} shows the time evolution of
+the angle difference between the bar and the halo wake, as measured from the
+second Fourier component. After the first Gyr, the angle difference in the
+SMUGGLE case is smaller than in the \Nbody{} case by about a factor of two,
+reflecting how the dark matter halo in the SMUGGLE case is unable to exert as
+negative a torque on the bar as in the \Nbody{} case.}\label{fig:wake}
 \end{figure*}
@@
 \includegraphics[width=9cm]{fig/fig4.pdf}
-\caption{The torque on the bar by the gas when the pattern speed of the bar is
-kept at a constant value. Only gas within the corotation radius is able to
-infall. Since slower bars have larger corotation radii, slower bars experience
-a larger net torque than faster bars. The setup of the simulations used here
-is identical to the SMUGGLE case discussed earlier and in the Methods section,
+\caption{\textbf{Average torque exerted by gas on a bar which rotates at a
+fixed pattern speed.} Since only gas within the corotation radius is able to
+infall and slower bars have larger corotation radii, slower bars experience a
+larger net torque than faster bars. The setup of the simulations used here is
+identical to the SMUGGLE case discussed earlier and in the Methods section,
 except the \Nbody{} disk is rotated as a solid body with a constant angular
@@
 Second, we performed a set of simulations in which we decreased the initial
-gas mass of the disk. We essentially find that there is a certain critical gas
-fraction (maybe gas surface density?) above which our proposed mechanism
-operates. If the gas available is below this threshold, then the pattern speed
+gas mass of the disk. We essentially find for our disk, our proposed mechanism
+operates as long as the gas contributes about $5\%$ of the total mass of the
+disk. If the gas available is below this threshold, then the pattern speed
 will decrease in a similar fashion to standard \Nbody{} simulations.
@@
 its effect of slowing down the bar.\cite{2021MNRAS.500.4710C,
-2021MNRAS.505.2412C} Third
-ed and a fast rotator, of which there is some
+2021MNRAS.505.2412C} Third, we provide an explanation for how the Milky Way's
+bar could be both long-lived and a fast rotator, of which there is some
 observational evidence.\cite{2019MNRAS.490.4740B} And finally, we complicate
@@
 galaxies\cite{2009ARAA..47..159B}) would find slowly rotating bars. There does
-exist one such example of a galaxy which is known to be a slow rotator -- UGC
-628.\cite{2009AA...499L..25C} This galaxy has been studied in detail by
-ref.\cite{2016MNRAS.463.1751C}, who note that it indeed has a low gas
-fraction. In general, we predict that all gas-rich spiral galaxies should
-rotate quickly while gas-poor spiral galaxies should rotate slowly.
+exist one such example of a galaxy which is known to be a slow rotator -- the
+low surface brightness galaxy UGC 628.\cite{2009AA...499L..25C} This galaxy
+has been studied in detail by ref.\cite{2016MNRAS.463.1751C}, who note that it
+indeed has a low gas fraction for galaxies of its type. In general, we predict
+that bars in gas-rich spiral galaxies should rotate quickly while bars in
+gas-poor spiral galaxies should rotate slowly.
 
@@
 
-\item [Acknowledgements] We would like to thank Greg L. Bryan, Drummond B.
-Fielding, Keith Hawkins, Jason A. S. Hunt, Sarah M. R. Jeffreson, Kathryn V.
-Johnston, Julio F. Navarro, David N. Spergel, Martin D. Weinberg, and Yanfei
-Zou for helpful discussions. A.B. would like to thank Todd Phillips for
-helpful discussions. Some computations in this paper were run on the FASRC
-Cannon cluster supported by the FAS Division of Science Research Computing
-Group at Harvard University. Resources supporting this work were also provided
-by the NASA High-End Computing (HEC) Program through the NASA Advanced
-Supercomputing (NAS) Division at Ames Research Center. A.B. was supported by
-the Future Investigators in NASA Earth and Space Science and Technology
-(FINESST) award number 80NSSC20K1536 during the completion of this work. L.H.
-funding. E.D. funding. F.M. funding. C.C. funding. J.Q. funding. L.V.S.
-funding. P.T. funding. M.V. funding.
+\item [Acknowledgements] We would like to thank Greg~L. Bryan, Drummond~B.
+Fielding, Keith Hawkins, Jason~A.~S. Hunt, Sarah~M.~R. Jeffreson, Kathryn~V.
+Johnston, Julio~F. Navarro, Joshua~S. Speagle, David~N. Spergel, Martin~D.
+Weinberg, and Yanfei Zou for helpful discussions. A.B. would like to thank
+Todd Phillips for helpful discussions. Some computations in this paper were
+run on the FASRC Cannon cluster supported by the FAS Division of Science
+Research Computing Group at Harvard University. Resources supporting this work
+were also provided by the NASA High-End Computing (HEC) Program through the
+NASA Advanced Supercomputing (NAS) Division at Ames Research Center. A.B. was
+supported by the Future Investigators in NASA Earth and Space Science and
+Technology (FINESST) award number 80NSSC20K1536 during the completion of this
+work. L.H. funding. E.D. funding. F.M. funding. C.C. funding. J.Q. funding.
+L.V.S. funding. P.T. funding. M.V. funding.
 
@@
 \includegraphics[width=9cm]{fig/fig-vcirc.pdf}
-\caption{The circular velocity curve for the \Nbody{} run (blue) and the SMUGGLE
-run (orange) compared to observational estimates for the Milky
+\caption{\textbf{The circular velocity curve of our initial setups.} This
+curve is shown for the \Nbody{} run (blue) and the SMUGGLE run (orange)
+compared to observational estimates for the Milky
 Way.\cite{2019ApJ...871..120E} We see that the circular velocity curve for
@@
 surface density is broadly consistent with the expected values for the Milky
-Way.\cite{2022ApJ...929L..18E} The discrepancy between $1$ and
-$4\,\textrm{kpc}$ is likely due to the fact that the distances to molecular
-cloud which underlines this work used a simple kinematic distance based on an
-axisymmetric model of the Milky Way,\cite{2017ApJ...834...57M} which is likely
-not accurate in the bar region where gas has large non-circular velocities.
+Way.\cite{2008AA...487..951K,2022ApJ...929L..18E} The discrepancy between $1$
+and $4\,\textrm{kpc}$ in the molecular and SFR surface density is likely due
+to the fact that the distances to molecular clouds which underlines this work
+used a simple kinematic distance based on an axisymmetric model of the Milky
+Way,\cite{2017ApJ...834...57M} which is not accurate in the bar region
+where gas has large non-circular velocities.
 
@@
 \includegraphics[width=9cm]{fig/fig-surf.pdf}
-\caption{The time evolution of the molecular gas surface density (\textit{upper}) and
-the star formation rate (SFR) surface density (\textit{lower}) at various
-times during the simulation. Colored lines indicate the profiles at selected
-times during the simulation while the black dashed lines indicate models for
-the Milky Way which allow the CO-to-H$_2$ conversion factor $X_{\textrm{CO}}$
-to vary with metallicity.\cite{2022ApJ...929L..18E} We see that the molecular
-gas and SFR surface densities are within an order of magnitude of the Milky
-Way's typical values at all times.  We see a sharp decrease in the gas and SFR
-surface densities along the extent of the bar from $\sim1$ to
-$\sim4\,\textrm{kpc}$, related to the gas inflow in this region. {\color{red}
-Waiting on atomic gas data from Juergen Kerp.}}
+\caption{\textbf{Evolution of gas and SFR surface density profiles.} The time
+evolution of the atomic gas surface density (\textit{upper}), molecular gas
+surface density (\textit{middle}) and the star formation rate (SFR) surface
+density (\textit{lower}) at various times during our fiducial simulation.
+Colored lines indicate the profiles at selected times during the simulation
+while the black dashed lines indicate observations for the atomic
+gas\cite{2008AA...487..951K} and black dotted lines indicate a model which
+allows the CO-to-H$_2$ conversion factor $X_{\textrm{CO}}$ to vary with
+metallicity.\cite{2022ApJ...929L..18E} Molecular gas surface densities were
+provided separately (N. Evans, private communication). We see that the
+molecular gas and SFR surface densities are within an order of magnitude of
+the Milky Way's typical values at all times.  We see a sharp decrease in the
+gas and SFR surface densities along the extent of the bar from $\sim1$ to
+$\sim4\,\textrm{kpc}$, related to the gas inflow in this region.}
 \label{fig:surf}
@@
 \includegraphics[width=9cm]{fig/fig-A2.pdf}
-\caption{The time evolution of the bar strength, measured as the maximum of the second
-Fourier component divided by the zeroth Fourier component (a formula is given
-in the Methods section). We see that in the \Nbody{} case (blue) the bar
-strength increases with time, consistent with previous results showing that
-the bar strength increases as bars slow down. In the SMUGGLE case (orange), we
-see that the bar strength remains roughly constant, possibly slightly
-decreasing with time. This is also consistent with the expected relation
-between pattern speed and strength since the bar in this case is not slowing
-down.}
+\caption{\textbf{Evolution of bar strength.} The bar strength is measured as
+the maximum of the second Fourier component divided by the zeroth Fourier
+component (a formula is given in the Methods section). We see that in the
+\Nbody{} case (blue) the bar strength increases with time, consistent with
+previous results showing that the bar strength increases as bars slow down. In
+the SMUGGLE case (orange), we see that the bar strength remains roughly
+constant, possibly slightly decreasing with time. This is also consistent with
+the expected relation between pattern speed and strength since the bar in this
+case is not slowing down.}
 \label{fig:strength}
@@
 \includegraphics[width=18cm]{fig/fig-bar_decomp.pdf}
-\caption{Demonstration of our bar decomposition procedure based on
-ref.\cite{2016MNRAS.463.1952P} The \textit{left panel} shows a surface density
-projection through the stellar component of the \Nbody{} simulation (disk and
-bulge). The \textit{middle panel} shows the component of the disk identified
-as being trapped in the bar while the \textit{right panel} shows the component
-of the disk identified as not being trapped in the bar. The fact that the
-untrapped stars form a roughly axisymmetric structure indicates our bar
-decomposition is sufficiently accurate. {\color{red} I will remake this figure
+\caption{\textbf{Disk decomposition into the bar and unbarred disk.} This
+procedure is based on ref.\cite{2016MNRAS.463.1952P} The \textit{left panel}
+shows a surface density projection through the stellar component of the
+\Nbody{} simulation (disk and bulge). The \textit{middle panel} shows the
+component of the disk identified as being trapped in the bar while the
+\textit{right panel} shows the component of the disk identified as not being
+trapped in the bar. The fact that the untrapped stars form a roughly
+axisymmetric structure indicates our bar decomposition is sufficiently
+accurate. {\color{red} I will remake this figure
 - I ended up losing the files which stored the bar membership, so I just

### 2022-04-28 plots/paper_fig-A2/plot_fig-A2.py
Update lots of figures

@@ -65,11 +65,11 @@ def run():
     
     fourier = read_fourier(Nbody, lvl)
     t, A2A0 = extract_t_max_A2A0(fourier)
-    ax.plot(t-t[300], A2A0, c=tb_c[0], label=r'$N$-body')
+    ax.plot(t-t[300], savgol_filter(A2A0, 81, 3), c=tb_c[0], label=r'$N$-body')
 
     fourier = read_fourier(phS2R35, lvl)
     t, A2A0 = extract_t_max_A2A0(fourier)
-    ax.plot(t, A2A0, c=tb_c[1], label='SMUGGLE')
+    ax.plot(t, savgol_filter(A2A0, 81, 3), c=tb_c[1], label='SMUGGLE')
 
     ax.set(xlim=(0, 5), ylim=(0, 0.7))
     ax.set(xlabel=r'$t\,[\,\textrm{Gyr}\,]$', ylabel=r'$\textrm{max}\left(\left|A_2/A_0\right|\right)$')

### 2022-04-28 plots/paper_fig-surf/plot_fig-surf.py

@@ -107,9 +107,13 @@ def compute_dphi(phi_disk, phi_halo):
 
 def run():
     cm = 1/2.54
-    fig, ax = plt.subplots(2, 1, sharex=True, figsize=(9*cm, 12*cm))
+    fig, ax = plt.subplots(3, 1, sharex=True, figsize=(9*cm, 14*cm))
     idx_list = [0, 100, 200, 400, 600, 800]
     center = np.array([200., 200., 200.])
+
+    data_mol = np.genfromtxt('sigmol-G1a-4p5.txt')
+    data_sfr = np.genfromtxt('sigsfr-G1a-4p5.txt')
+    data_atomic = np.genfromtxt('surf_all.out')
 
     for i,idx in enumerate(tqdm(idx_list)):
         sn = read_snap(idx, phS2R35, lvl, parttype=[0], fields=None)
@@ -118,11 +122,13 @@ def run():
         R = np.linalg.norm(pos[:,:2], axis=1)
         Rbins = np.logspace(-2, np.log10(20), 25)
 
-        mass = sn.part0.mass.value
+        mass = sn.part0.mass.value * sn.part0.MolecularHFrac
+        mass_atomic = sn.part0.mass.value * (1-sn.part0.MolecularHFrac)
         if i==0:
             r = np.linalg.norm(pos, axis=1)
 
         aveR, surf = compute_surface_density(R, mass, Rbins)
+        aveR_atomic, surf_atomic = compute_surface_density(R, mass_atomic, Rbins)
 
         if True:
             label_0 = repr(sn.Time.value)
@@ -131,8 +137,8 @@ def run():
             label_0 = None
             label_1 = repr(sn.Time.value)
 
-
-        ax[0].plot(aveR, (1E10 / 1E6) * surf, label=label_0, c=tb_c[i]) # convert to Msun/pc^2
+        ax[0].plot(aveR_atomic, (1E10/1E6) * surf_atomic, label=label_0, c=tb_c[i])
+        ax[1].plot(aveR, (1E10 / 1E6) * surf, label=label_1, c=tb_c[i]) # convert to Msun/pc^2
 
         sfr = sn.part0.sfr.value
         key = np.where(sfr > 0)[0]
[…]

### 2022-04-28 plots/paper_fig3/plot_fig3.py

@@ -150,11 +150,11 @@ def run():
     rng = [[-7.5, 7.5], [-7.5, 7.5]]
     extent = [rng[0][0], rng[0][1], rng[1][0], rng[1][1]]
     
-    vmin = -0.002
-    vmax = 0.002
+    vmin = -0.002 * (1E10/1E6)
+    vmax = 0.002 * (1E10/1E6)
 
     cm = 1/2.54
-    fig, ax = plt.subplots(1, 3, figsize=(18*cm, 6*cm))
+    fig, ax = plt.subplots(1, 4, figsize=(18*cm, 6*cm), gridspec_kw={"width_ratios":[1.1, 1, 1.1, 0.05]})
 
     # First panel, wake of Nbody.
     sn = read_snap(Nbody_idx, Nbody, lvl, parttype=[1], fields=None)
@@ -163,7 +163,7 @@ def run():
     Rbar = bar_prop_Nbody['Rbar'][Nbody_idx]
     print(fourier_Nbody['A2_angle'][Nbody_idx], bar_prop_Nbody['bar_angle'][Nbody_idx], fourier_Nbody['A2_h_angle'][Nbody_idx])
 
-    ax[0].imshow(heatmap.T, origin='lower', vmin=vmin, vmax=vmax, extent=extent, cmap='bwr')
+    ax[0].imshow((1E10/1E6)*heatmap.T, origin='lower', vmin=vmin, vmax=vmax, extent=extent, cmap='bwr')
     ax[0].plot((-Rbar, Rbar), (0.0, 0.0), c='k')
     dphi = fourier_Nbody['A2_h_angle'][Nbody_idx] - fourier_Nbody['A2_angle'][Nbody_idx]
     dphi /= 2.0
@@ -172,7 +172,7 @@ def run():
     ax[0].axes.xaxis.set_visible(False)
     ax[0].axes.yaxis.set_visible(False)
 
-    ax[0].set_aspect('auto')
+    ax[0].set_aspect('equal')
 
     # Second panel, diff in angle.
     dphiN = compute_dphi(fourier_Nbody['A2_angle'], fourier_Nbody['A2_h_angle'])
@@ -180,6 +180,9 @@ def run():
 
     tN = fourier_Nbody['time']
     tS = fourier_SMUGGLE['time']
+
+    dphiN = savgol_filter(dphiN, 81, 3)
[…]

### 2022-05-15 paper/ms.tex
Small changes to paper
Add MV funding
Updates from Overleaf

@@
   bar from slowing down or speeding up. We find that in a Milky Way-like disk,
-  a gas fraction of only about \boldmath$5\%$ is necessary for this mechanism to operate.
-  This result naturally explains why nearly all observed bars rotate
-  rapidly\cite{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G}
-  and is especially relevant for our understanding of how the Milky Way
-  arrived at its present state.
+  a gas fraction of only about \boldmath$5\%$ is necessary for this mechanism
+  to operate. This result naturally explains why nearly all observed bars
+  rotate rapidly\cite{2011MSAIS..18...23C, 2015AA...576A.102A,
+  2019MNRAS.482.1733G} and is especially relevant for our understanding of how
+  the Milky Way arrived at its present state.
 
@@
 different components. The solid lines indicate the torque on the bar by the
-dark matter halo whereas the dashed line indicates the torque on the bar by
+dark matter halo, whereas the dashed line indicates the torque on the bar by
 the gas phase. In the \Nbody{} case, the halo exerts a steady negative torque on
@@
 
-\item [Acknowledgements] We would like to thank Greg~L. Bryan, Drummond~B.
-Fielding, Keith Hawkins, Jason~A.~S. Hunt, Sarah~M.~R. Jeffreson, Kathryn~V.
-Johnston, Julio~F. Navarro, Joshua~S. Speagle, David~N. Spergel, Martin~D.
-Weinberg, and Yanfei Zou for helpful discussions. A.B. would like to thank
-Todd Phillips for helpful discussions. Some computations in this paper were
-run on the FASRC Cannon cluster supported by the FAS Division of Science
-Research Computing Group at Harvard University. Resources supporting this work
-were also provided by the NASA High-End Computing (HEC) Program through the
-NASA Advanced Supercomputing (NAS) Division at Ames Research Center. A.B. was
-supported by the Future Investigators in NASA Earth and Space Science and
-Technology (FINESST) award number 80NSSC20K1536 during the completion of this
-work. L.H. funding. E.D. funding. F.M. funding. C.C. funding. J.Q. funding.
-L.V.S. funding. P.T. funding. M.V. funding.
+\item [Acknowledgements] We would like to thank Greg~L. Bryan, Neal~J. Evans,
+Drummond~B. Fielding, Keith Hawkins, Jason~A.~S. Hunt, Sarah~M.~R. Jeffreson,
+Kathryn~V. Johnston, Peter~M.~W. Kalberla, Jürgen Kerp, Julio~F. Navarro,
+Joshua~S. Speagle, David~N. Spergel, Martin~D. Weinberg, and Yanfei Zou for
+helpful discussions. A.B. would like to thank Todd Phillips for helpful
+discussions. Some computations in this paper were run on the FASRC Cannon
+cluster supported by the FAS Division of Science Research Computing Group at
+Harvard University. Resources supporting this work were also provided by the
+NASA High-End Computing (HEC) Program through the NASA Advanced Supercomputing
+(NAS) Division at Ames Research Center. A.B. was supported by the Future
+Investigators in NASA Earth and Space Science and Technology (FINESST) award
+number 80NSSC20K1536 during the completion of this work. L.H. funding. E.D.
+funding.
+F.M.
+funding.
+C.C.
+funding.
+J.Q.
+acknowledges support from NSF grant AST-2008490.
+L.V.S.
+funding.
+PT acknowledges support from NSF grant AST-1909933, AST-2008490, and NASA ATP Grant 80NSSC20K0502.
+M.V.
+acknowledges support through NASA ATP 19-ATP19-0019,
+19-ATP19-0020, 19-ATP19-0167, and NSF grants AST-1814053, AST-1814259,
+AST-1909831, AST-2007355 and AST-2107724.
 
@@
 strength.\cite{2003MNRAS.341.1179A} In the SMUGGLE case, we see that the bar
-strength has an initial drop but then remains at a roughly constant but
-slightly decreasing strength. This is consistent with the pattern speed in the
+strength has an initial drop but then remains at a roughly constant, but
+slightly decreasing, strength. This is consistent with the pattern speed in the
 SMUGGLE case being roughly constant or slightly increasing.

### 2022-05-15 paper/ms.tex
Changes based on PT comments

@@
 We found that below a certain gas fraction, bars should still be able to slow
-down. Therefore, we expect barred spiral galaxies which are exceptionally
-gas-poor to be rotating very slowly. We therefore predict that observations
-which target such galaxies (e.g., lenticular barred
+down. Therefore, we expect barred spiral galaxies which have been gas-poor for
+extended periods of time to be rotating very slowly. We therefore predict that
+observations which target such galaxies (e.g., lenticular barred
 galaxies\cite{2009ARAA..47..159B}) would find slowly rotating bars. There does
@@
 has been studied in detail by ref.\cite{2016MNRAS.463.1751C}, who note that it
-indeed has a low gas fraction for galaxies of its type. In general, we predict
-that bars in gas-rich spiral galaxies should rotate quickly while bars in
-gas-poor spiral galaxies should rotate slowly.
+indeed has a low gas fraction for galaxies of its type. We predict a general
+trend that bars in gas-rich spiral galaxies should rotate quickly while bars
+in gas-poor spiral galaxies should rotate slowly.
 
@@
 Kathryn~V. Johnston, Peter~M.~W. Kalberla, Jürgen Kerp, Julio~F. Navarro,
-Joshua~S. Speagle, David~N. Spergel, Martin~D. Weinberg, and Yanfei Zou for
-helpful discussions. A.B. would like to thank Todd Phillips for helpful
-discussions. Some computations in this paper were run on the FASRC Cannon
-cluster supported by the FAS Division of Science Research Computing Group at
-Harvard University. Resources supporting this work were also provided by the
-NASA High-End Computing (HEC) Program through the NASA Advanced Supercomputing
-(NAS) Division at Ames Research Center. A.B. was supported by the Future
+Joshua~S. Speagle, Martin~D. Weinberg, and Yanfei Zou for helpful discussions.
+A.B. would like to thank Todd Phillips for helpful discussions. Some
+computations in this paper were run on the FASRC Cannon cluster supported by
+the FAS Division of Science Research Computing Group at Harvard University.
+Resources supporting this work were also provided by the NASA High-End
+Computing (HEC) Program through the NASA Advanced Supercomputing (NAS)
+Division at Ames Research Center. A.B. was supported by the Future
 Investigators in NASA Earth and Space Science and Technology (FINESST) award
 number 80NSSC20K1536 during the completion of this work. L.H. funding. E.D.
-funding.
-F.M.
-funding.
-C.C.
-funding.
-J.Q.
-acknowledges support from NSF grant AST-2008490.
-L.V.S.
-funding.
-PT acknowledges support from NSF grant AST-1909933, AST-2008490, and NASA ATP Grant 80NSSC20K0502.
-M.V.
-acknowledges support through NASA ATP 19-ATP19-0019,
-19-ATP19-0020, 19-ATP19-0167, and NSF grants AST-1814053, AST-1814259,
-AST-1909831, AST-2007355 and AST-2107724.
+funding. F.M. funding. C.C. funding. J.Q. acknowledges support from NSF grant
+AST-2008490. L.V.S. funding. P.T. acknowledges support from NSF grant
+AST-1909933, AST-2008490, and NASA ATP Grant 80NSSC20K0502. M.V. acknowledges
+support through NASA ATP 19-ATP19-0019, 19-ATP19-0020, 19-ATP19-0167, and NSF
+grants AST-1814053, AST-1814259, AST-1909831, AST-2007355 and AST-2107724.
 
@@
 Star formation occurs in cells above a density threshold
-($n_{\textrm{th}}=100\,\textrm{cm}^{-3}$) according to
-ref.\cite{2003MNRAS.339..289S} with a star-formation efficiency of $\epsilon =
-0.01$. Star formation converts gas cells into star particles which represent
-single stellar populations. For each star particle, the deposition of energy,
-momentum, and mass from stellar winds and supernovae is modeled.
+($n_{\textrm{th}}=100\,\textrm{cm}^{-3}$) with a star-formation efficiency of
+$\epsilon = 0.01$. Star formation converts gas cells into star particles which
+represent single stellar populations. For each star particle, the deposition
+of energy, momentum, and mass from stellar winds and supernovae is modeled.
 Photo-ionization and radiation pressure are modeled using an approximate

### 2022-05-15 paper/ms.tex
Add ref to 2020 pattern speed paper

@@
 galaxies have $1 < \Rot < 1.4$.\cite{2011MSAIS..18...23C, 2015AA...576A.102A,
-2019MNRAS.482.1733G}
+2019MNRAS.482.1733G, 2020MNRAS.491.3655G}
 
@@
 section.
+
+We note that ref.\cite{2020MNRAS.491.3655G} found that the rotation parameter
+\Rot{} positively correlates with gas fraction, such that galaxies with higher
+gas fractions are rotating more slowly. However, it is not obvious this is in
+tension with our result since the gas fraction of galaxies correlates with
+other galactic properties.\cite{2009ARAA..47..159B} Furthermore, the
+measurement of pattern speeds is a delicate process still prone to large error
+bars.
 

### 2022-05-15 paper/ms.tex
Clean up Fig 3

@@
 \caption{\textbf{The wake excited in the dark matter halo.} The dark matter
-halo wake is shown in the \Nbody{} case (\textit{left panel}) and SMUGGLE case
+halo wake is shown in the \Nbody{} case (\textit{middle panel}) and SMUGGLE case
 (\textit{right panel}) after $2.6\,\textrm{Gyr}$ of evolution. The
-\textit{left} and \textit{right panels} show a surface density projection in
+\textit{middle} and \textit{right panels} show a surface density projection in
 the $x$-$y$ plane of the dark matter halo after an axisymmetric average has
@@
 the second Fourier component within a sphere of all material within a radius
-of $4\,\textrm{kpc}$). The \textit{center panel} shows the time evolution of
+of $4\,\textrm{kpc}$). The \textit{left panel} shows the time evolution of
 the angle difference between the bar and the halo wake, as measured from the
@@
 offset), then the wake cannot exert a negative torque on the bar. This angle
-is plotted in the middle panel of Fig.~\ref{fig:wake}, which shows that the
+is plotted in the left panel of Fig.~\ref{fig:wake}, which shows that the
 angle offset is larger in the \Nbody{} case than in the SMUGGLE case by about a
-factor of two. The left and right panels of Fig.~\ref{fig:wake} show the halo
-wake with respect to the location of the bar in the \Nbody{} (left) and SMUGGLE
+factor of two. The center and right panels of Fig.~\ref{fig:wake} show the halo
+wake with respect to the location of the bar in the \Nbody{} (center) and SMUGGLE
 (right) cases.

### 2022-05-16 paper/ms.tex
FM and JQ edits

@@
 simulation while the lower panels show a simulation which includes the SMUGGLE
-model for the ISM. Columns show different points in time, separated by
+model for the ISM. Columns show different points in time, separated by about
 $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
@@
 
-A surface density projection of our simulations is shown in
+A face-on surface density projection of our simulations is shown in
 Fig.~\ref{fig:overview}. The upper three panels show the disk in the \Nbody{}
@@
 evolution, but after that the halo exerts a much smaller torque on the bar,
-averaging only $-7.8$ in the same units and over the same interval. The gas in
-the SMUGGLE case exerts a steady positive torque averaging $11.7$ in the same
-units and over the same interval.
+averaging only $-7.8$ in the same units and over the same time interval. The gas in
+the SMUGGLE case exerts a steady positive torque averaging $11.7$ over $1\,\textrm{Gyr}$ in the same
+units.
 
 The fact that the dark matter halo in the SMUGGLE case exerts a smaller
-positive torque on the bar can be understood in terms of the halo wake
+negative torque on the bar can be understood in terms of the halo wake
 mechanism. In the \Nbody{} case, halo material which is resonant with the bar
@@
 torque on the bar by the gas phase at different rotation rates. The result of
-this experiment is shown in Fig.~\ref{fig:equil}, which shows that a more
+this experiment is illustrated in Fig.~\ref{fig:equil}, which shows that a more
 slowly rotating bar experiences a larger positive torque from the gas. We have
-therefore shown evidence for an equilibrium mechanism which keeps the pattern
+therefore presented evidence for an equilibrium mechanism which keeps the pattern
 speed of the bar constant, resulting from the complex interplay between the
@@
 pattern speed to continuously increase. This is consistent with our proposed
-equilibrium mechanism and with previous studies of barred galaxies which
-varied the gas fraction.\cite{2010ApJ...719.1470V} The exact critical gas
+equilibrium mechanism and with previous studies of barred galaxies with varying gas fractions.\cite{2010ApJ...719.1470V} The exact critical gas
 fraction above which our mechanism operates is likely dependent on the
@@
 number 80NSSC20K1536 during the completion of this work. L.H. funding. E.D.
-funding. F.M. funding. C.C. funding. J.Q. acknowledges support from NSF grant
+funding. C.C. funding. J.Q. acknowledges support from NSF grant
 AST-2008490. L.V.S. funding. P.T. acknowledges support from NSF grant
@@
 \cite{2019MNRAS.489.4233M} implemented within the moving-mesh, finite-volume
-hydrodynamics code AREPO \cite{2010MNRAS.401..791S}. The SMUGGLE model
-includes self-gravity, hydrodynamics, radiative heating and cooling, star
+hydrodynamics and gravity code AREPO \cite{2010MNRAS.401..791S}. The SMUGGLE model
+additionally includes radiative heating and cooling, star
 formation, and stellar feedback. Explicit gas cooling and heating of the
@@
 $\epsilon = 0.01$. Star formation converts gas cells into star particles which
-represent single stellar populations. For each star particle, the deposition
-of energy, momentum, and mass from stellar winds and supernovae is modeled.
+represent single stellar populations with an initial mass function according to ref.\cite{2001ApJ...554.1274C} For each star particle, the deposition
+of energy, momentum, mass, and metals from stellar winds and supernovae is modeled.
 Photo-ionization and radiation pressure are modeled using an approximate
@@
 disk. In our fiducial run, we set the surface density of the gas disk from
-$4\,\textrm{kpc}$ to $\sim9.3\,\textrm{kpc}$ to be $20\,\Msun/\textrm{kpc}^2$.
+$4\,\textrm{kpc}$ to $\sim9.3\,\textrm{kpc}$ to be $20\,\Msun/\textrm{pc}^2$.
 We also ran with surface densities of $15$, $10$, and
-$5\,\Msun/\textrm{kpc}^2$. These correspond to initial gas fractions of
+$5\,\Msun/\textrm{pc}^2$. These correspond to initial gas fractions of
 approximately $16\%$, $10\%$, $7\%$, and $4\%$. The pattern speed evolution is
 shown in Extended Data Fig.~\ref{fig:fgas} We find that the bar in disks with
-initial surface densities of $20$, $15$, and $10\,\Msun/\textrm{kpc}$ evolve
+initial surface densities of $20$, $15$, and $10\,\Msun/\textrm{pc}^2$ evolve
 with a constant pattern speed while the bar in a disk with initial surface
@@
 considered in this work, a gas fraction of only approximately $5\%$ is
-necessary in order for the proposed stabilizing mechanism to operate. Furthermore, we co
+necessary in order for the proposed stabilizing mechanism to operate.
 
@@
 the maximum of the second Fourier component divided by the zeroth Fourier
-component (a formula is given in the Methods section). We see that in the
+component. We see that in the
 \Nbody{} case (blue) the bar strength increases with time, consistent with
@@
 These criterion are slightly different and simplified from the ones used in
-ref.\cite{2016MNRAS.463.1952P}, but we found to empirically work well at
+ref.\cite{2016MNRAS.463.1952P}, but we found them to empirically work well at
 decomposing the disk into a bar and disk component. In Extended Data

### 2022-05-19 paper/ms.tex
Update bar decomp figure

@@
 \centering
-\includegraphics[width=18cm]{fig/fig-bar_decomp.pdf}
+\includegraphics[width=18cm]{fig/fig-bar.pdf}
 \caption{\textbf{Disk decomposition into the bar and unbarred disk.} This
@@
 axisymmetric structure indicates our bar decomposition is sufficiently
-accurate. {\color{red} I will remake this figure
-- I ended up losing the files which stored the bar membership, so I just
-pulled these screenshots from an earlier presentation.}}
+accurate.}
 \label{fig:decomp}

### 2022-06-08 plots/fourier_component/compute_fourier_component.py
Clean up fourier computation

@@ -226,6 +226,7 @@ if __name__ == '__main__':
     phgvS2Rc35 = 'phantom-vacuum-Sg20-Rc3.5'
     phgvS15Rc35 = 'phantom-vacuum-Sg15-Rc3.5'
     phgvS10Rc35 = 'phantom-vacuum-Sg10-Rc3.5'
+    phgvS08Rc35 = 'phantom-vacuum-Sg08-Rc3.5'
     phgvS05Rc35 = 'phantom-vacuum-Sg05-Rc3.5'
     
     
@@ -234,25 +235,13 @@ if __name__ == '__main__':
 
     pair_list = [#(fid_dP, 'lvl5'), (fid_dP, 'lvl4'), #(fid_dP, 'lvl3'),
                  (Nbody, 'lvl4'), (Nbody, 'lvl3'), (Nbody, 'lvl2'),
-                 (Nbody, 'lvl3-GMCs'),
-                 #(fid_dP_c1, 'lvl5'), (fid_dP_c1, 'lvl4'), (fid_dP_c1, 'lvl3'),
-                 #(fid_dP2_c1, 'lvl5'), (fid_dP2_c1, 'lvl4'), (fid_dP2_c1, 'lvl3'),
-                 #(fid_dP_c1_bG2, 'lvl5'), (fid_dP_c1_bG2, 'lvl4'), (fid_dP_c1_bG2, 'lvl3'),
-                 #(fid_dP_c1_sp, 'lvl5'), (fid_dP_c1_sp, 'lvl4'), (fid_dP_c1_sp, 'lvl3'),
-                 #(fid_dP_c1_MB, 'lvl5'), (fid_dP_c1_MB, 'lvl4'), (fid_dP_c1_MB, 'lvl3'),
-                 #(phv, 'lvl3'),
-                 #(phgv, 'lvl3'),
-                 #(phgvS1, 'lvl3'),
-                 #(phgvS2, 'lvl3'),
-                 #(phgvS2Rc35, 'lvl3'), (phgvS2Rc35, 'lvl3-isotherm'), 
-                 #(phgvS2Rc35RF, 'lvl4-Paulfix'), (phgvS2Rc35RF, 'lvl3-Paulfix'),
-                 #(phgS1, 'lvl3')]
                  (phgvS2Rc35, 'lvl4'), (phgvS2Rc35, 'lvl3'),
[…]

### 2022-06-08 paper/ms.tex
Small updates

@@
 \newcommand{\Nbody}{$N$-body}
+\newcommand{\SMUGGLE}{SMUGGLE}
 \newcommand{\Msun}{\ensuremath{M_{\odot}}}
@@
 number 80NSSC20K1536 during the completion of this work. L.H. funding. E.D.
-funding. C.C. funding. J.Q. acknowledges support from NSF grant
+funding. J.Q. acknowledges support from NSF grant
 AST-2008490. L.V.S. funding. P.T. acknowledges support from NSF grant
@@
 \includegraphics[width=18cm]{fig/fig-bar.pdf}
-\caption{\textbf{Disk decomposition into the bar and unbarred disk.} This
+\caption{\textbf{Disk decomposition into the barred and unbarred disk.} This
 procedure is based on ref.\cite{2016MNRAS.463.1952P} The \textit{left panel}
-shows a surface density projection through the stellar component of the
-\Nbody{} simulation (disk and bulge). The \textit{middle panel} shows the
+shows a face-on surface density projection through the stellar component of the
+\SMUGGLE{} simulation (disk and bulge) at $t=1\,\textrm{Gyr}$. The \textit{middle panel} shows the
 component of the disk identified as being trapped in the bar while the

### 2022-06-08 plots/fourier_component/compute_fourier_component.py
Add snap700 plots

@@ -236,6 +236,7 @@ if __name__ == '__main__':
     pair_list = [#(fid_dP, 'lvl5'), (fid_dP, 'lvl4'), #(fid_dP, 'lvl3'),
                  (Nbody, 'lvl4'), (Nbody, 'lvl3'), (Nbody, 'lvl2'),
                  (phgvS2Rc35, 'lvl4'), (phgvS2Rc35, 'lvl3'),
+                 (phgvS2Rc35, 'lvl3-snap700'),
                  #(phgvS2Rc35, 'lvl3-rstHalo'), (phgvS2Rc35, 'lvl3-snap700'),
                  #(phgvS2Rc35, 'lvl3-GFM'), (phgvS2Rc35, 'lvl3-isotherm'),
                  #(phgvS2Rc35, 'lvl4-GFM'),

### 2022-06-08 paper/ms.tex
Add snap700 figures
Also moves the record player plot to the Methods

@@
   previous simulations using more simplified models for the interstellar gas
-  but its explanation has remained elusive\cite{1993AA...268...65F,
+  but it has remained unexplained\cite{1993AA...268...65F,
   2010ApJ...719.1470V}. We propose that the gas phase of the disk and the dark
@@
 
-\begin{figure}[h!]
+\begin{figure}[t!]%
 \centering
-\includegraphics[width=9cm]{fig/fig4.pdf}
-\caption{\textbf{Average torque exerted by gas on a bar which rotates at a
-fixed pattern speed.} Since only gas within the corotation radius is able to
-infall and slower bars have larger corotation radii, slower bars experience a
-larger net torque than faster bars. The setup of the simulations used here is
-identical to the SMUGGLE case discussed earlier and in the Methods section,
-except the \Nbody{} disk is rotated as a solid body with a constant angular
-velocity.}\label{fig:equil}
+\includegraphics[width=9cm]{fig/fig-snap700.pdf}
+\caption{\textbf{Pattern speed evolution with a lower initial pattern speed.} We
+tested the evolution of our system when gas is added to the \Nbody{} run at a
+later time, but with all other simulation parameters kept the same. The setup
+is therefore identical to our previous runs just with a bar that is larger,
+stronger, and with a lower pattern speed. We find that the pattern speed
+evolution is very similar to our fiducial case, except that the bar retains
+its original pattern speed. This indicates a mechanism which keeps the bar at
+its formation pattern speed, and that there is not a particular pattern speed
+which the system tends to.}
+\label{fig:snap700}
 \end{figure}
@@
 reinforce a wake, arresting the process by which the halo can slow the bar
-down.
+down. We term this process ``exhaustion.''
 
@@
 We propose that an equilibrium mechanism is responsible for the pattern speed
-remaining approximately constant. In this scenario, a torque must oppose
-changes in the pattern speed so that when the bar speeds up, a negative torque
-will slow it down and when the bar slows down, a positive torque will speed it
-up. It is simple to explain the first of these - when the pattern speed
-increases, the location of resonances will shift to regions of the dark matter
-halo phase space where no wake has been excited yet (e.g., the corotation
-radius will shrink).
+remaining approximately constant. In this scenario, residual negative torque
+from the dark matter halo balances out the positive torque from the gas phase.
+It has been shown when an analytic bar is forced to rotate at a constant
+pattern speed for a few Gyr, the halo exerts almost no torque on the
+bar.\cite{2022MNRAS.513..768C} We instead see over a several Gyr time span
+that the dark matter halo in our simulation is still able to support some
+negative torque.
 
-When the bar slows down, we argue that this induces a larger positive torque
-from the gas phase. Only gas within corotation will flow inwards, while gas
-outside corotation will flow outwards.\cite{2011MNRAS.415.1027H} Since the
-corotation radius is larger for more slowly rotating bars, it follows that
-more slowly rotating bars should be more efficient at driving gas inflows and
-thus experience a larger positive torque from the gas phase.
+We argue that the following occurs. First, the bar is not able to slow down
+quickly enough due to the positive torque of the infalling gas. This causes
+the resonant halo phase space at a particular pattern speed, $\Omega_{p,0}$,
+to become exhausted and no longer able to support a negative torque. Second,
+the gas is still exerting a positive torque on the bar, and therefore the
+pattern speed will again increase. Since at higher pattern speeds the halo has
+not yet been totally exhausted, the halo will once again be able to exert a
+negative torque on the bar. The pattern speed will then settle at a new value
+slightly higher than $\Omega_{p,0}$ where the gas and halo torques cancel.
+Over time, the pattern speed should slowly increase as the halo becomes
+progressively exhausted.
 
-We performed an experiment to test this hypothesis by forcing the stellar disk
-in the SMUGGLE run to rotate at a constant angular rate and measuring the
-torque on the bar by the gas phase at different rotation rates. The result of
-this experiment is illustrated in Fig.~\ref{fig:equil}, which shows that a more
-slowly rotating bar experiences a larger positive torque from the gas. We have
-therefore presented evidence for an equilibrium mechanism which keeps the pattern
-speed of the bar constant, resulting from the complex interplay between the
-dark matter halo and the gas phase.
+A clear prediction of this proposed mechanism is that the constant pattern
+speed a particular galaxy will end up is somewhat arbitrary. In the real
+universe for an isolated galaxy it would be the formation pattern speed of the
+bar while in our simulation it is the pattern speed of the bar when gas is
+added to the system. We tested this by adding gas to the system at a later
+time when the bar has further grown and slowed down with time. In our
+particular test, we added the gas at a time when the pattern speed is
+$\sim30\,\textrm{km}/\textrm{s}/\textrm{kpc}$. As shown in
+Fig.~\ref{fig:snap700}, we find that the pattern speed evolution is very
+similar between the two cases (orange and red lines). When the initial pattern
+speed is lower (red line), the addition of gas does not cause the pattern
+speed to quickly return to the higher value of our fiducial simulation (orange
+line).
 
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width=9cm]{fig/fig4.pdf}
+\caption{\textbf{Average torque exerted by gas on a bar which rotates at a
+fixed pattern speed.} Since only gas within the corotation radius is able to
+infall and slower bars have larger corotation radii, slower bars experience a
+larger net torque than faster bars. The setup of the simulations used here is
+identical to the SMUGGLE case discussed earlier and in the Methods section,
+except the \Nbody{} disk is rotated as a solid body with a constant angular
+velocity.}\label{fig:equil}
+\end{figure}
+
 \begin{figure}[t!]%
@@
 necessary in order for the proposed stabilizing mechanism to operate.
+
+\vspace{12pt}
+
+\noindent
+{\bf Varying pattern speed}
+\\
+\noindent
+When the bar slows down, we argue that this induces a larger positive torque
+from the gas phase. Only gas within corotation will flow inwards, while gas
+outside corotation will flow outwards.\cite{2011MNRAS.415.1027H} Since the
+corotation radius is larger for more slowly rotating bars, it follows that
+more slowly rotating bars should be more efficient at driving gas inflows and
+thus experience a larger positive torque from the gas phase.
+
+We performed an experiment to test this hypothesis by forcing the stellar disk
+in the SMUGGLE run to rotate at a constant angular rate and measuring the
+torque on the bar by the gas phase at different rotation rates. The result of
+this experiment is illustrated in Fig.~\ref{fig:equil}, which shows that a more
+slowly rotating bar experiences a larger positive torque from the gas.
 

### 2022-06-08 paper/ms.tex
Other small changes

@@
 Fig.~\ref{fig:snap700}, we find that the pattern speed evolution is very
-similar between the two cases (orange and red lines). When the initial pattern
+similar between the two cases (orange and red lines). If anything, the system
+with a lower pattern speed seems to speed up more, which is consistent with
+our picture since the stronger bar should experience a larger torque from the
+gas as it is more efficient at driving gas inflows. We also show in the
+Methods that more slowly rotating bars at fixed bar strength are more
+efficient at driving as inflows as well. Nonetheless, when the initial pattern
 speed is lower (red line), the addition of gas does not cause the pattern
@@
 the Milky Way disk, and in the SMUGGLE run the addition of the gas phase
-results in a slightly higher circular velocity. We also show the evolution of
-the surface density profile in Extended Data Fig.~\ref{fig:surf} We find that
-in our simulation the atomic and molecular gas surface density and the SFR
-surface density is broadly consistent with the expected values for the Milky
+results in a slightly higher circular velocity.
+
+We also show the evolution of the surface density profile in Extended Data
+Fig.~\ref{fig:surf} We find that in our simulation the atomic and molecular
+gas surface density and the SFR surface density is broadly consistent with the
+expected values for the Milky
 Way.\cite{2008AA...487..951K,2022ApJ...929L..18E} The discrepancy between $1$
@@
 used a simple kinematic distance based on an axisymmetric model of the Milky
-Way,\cite{2017ApJ...834...57M} which is not accurate in the bar region
-where gas has large non-circular velocities.
+Way,\cite{2017ApJ...834...57M} which is not accurate in the bar region where
+gas has large non-circular velocities.
 

### 2022-06-08 paper/ms.tex
Updates from Overleaf

--- 2022-06-08 paper/ms.tex
@@
 \newcommand{\Nbody}{$N$-body}
-\newcommand{\SMUGGLE}{SMUGGLE}
 \newcommand{\Msun}{\ensuremath{M_{\odot}}}
@@
 number 80NSSC20K1536 during the completion of this work. L.H. funding. E.D.
-funding. J.Q. acknowledges support from NSF grant
-AST-2008490. L.V.S. funding. P.T. acknowledges support from NSF grant
+funding. C.C. funding. J.Q. acknowledges support from NSF grant
+AST-2008490. L.V.S. is grateful for financial support from NASA ATP
+80NSSC20K0566, NSF AST 1817233 and NSF CAREER 1945310
+grants. P.T. acknowledges support from NSF grant
 AST-1909933, AST-2008490, and NASA ATP Grant 80NSSC20K0502. M.V. acknowledges
@@
 \includegraphics[width=18cm]{fig/fig-bar.pdf}
-\caption{\textbf{Disk decomposition into the barred and unbarred disk.} This
+\caption{\textbf{Disk decomposition into the bar and unbarred disk.} This
 procedure is based on ref.\cite{2016MNRAS.463.1952P} The \textit{left panel}
-shows a face-on surface density projection through the stellar component of the
-\SMUGGLE{} simulation (disk and bulge) at $t=1\,\textrm{Gyr}$. The \textit{middle panel} shows the
+shows a surface density projection through the stellar component of the
+\Nbody{} simulation (disk and bulge). The \textit{middle panel} shows the
 component of the disk identified as being trapped in the bar while the

### 2022-06-08 plots/fourier_component/src/main.c
update pdf

@@ -1,0 +1,165 @@
+#include <stdlib.h>
+#include <stdio.h>
+#include <string.h>
+#include <unistd.h>
+#include <hdf5.h>
+
+#define NTYPES 6
+#define IDX(i, j) i*3+j
+
+int Nsnap;
+uint NumPart_Total[NTYPES];
+double MassTable[NTYPES];
+int Nfiles;
+
+struct Part{
+    double Pos[3];
+    double Mass;
+};
+
+void read_header_attribute(hid_t file_id, hid_t DTYPE, char* attr_name, void *buf){
+    hid_t header = H5Gopen(file_id, "/Header", H5P_DEFAULT);
+    hid_t hdf5_attribute = H5Aopen(header, attr_name, H5P_DEFAULT);
+    H5Aread(hdf5_attribute, DTYPE, buf);
+    H5Aclose(hdf5_attribute);
+    H5Gclose(header);
+    return;
+}
+
+void read_parttype(char *output_dir, int snapnum, int ptype, struct Part** output_part){
+    char fname[1000], grp_name[1000];
+    hid_t file_id, grp_id, dset;
+    int i, k;
+    long long Ncum, j;
+    uint NumPart_ThisFile[NTYPES];
+    double *Pos, *Mass;
+
+    *output_part = (struct Part*) malloc(sizeof(struct Part) * NumPart_Total[ptype]);
+
+    Ncum = 0;
+    Nfiles = 1;
+    for(i=0; i<Nfiles; i++){
+        sprintf(fname, "%s/snapdir_%03d/snapshot_%03d.%d.hdf5", output_dir, snapnum, snapnum, i);
+        file_id = H5Fopen(fname, H5F_ACC_RDONLY, H5P_DEFAULT);
+        read_header_attribute(file_id, H5T_NATIVE_UINT, "NumPart_ThisFile", NumPart_ThisFile);
+        Pos = (double *)malloc(sizeof(double) * 3 * NumPart_ThisFile[ptype]);
+        Mass = (double *)malloc(sizeof(double) * NumPart_ThisFile[ptype]);
+
+        sprintf(grp_name, "PartType%d", ptype);
+        printf("grp_name=%s\n", grp_name);
+        grp_id = H5Gopen(file_id, grp_name, H5P_DEFAULT);
+
+        // Read coordinates.
+        dset = H5Dopen(grp_id, "Coordinates", H5P_DEFAULT);
[…]

### 2022-06-08 plots/job_template.sh

@@ -1,0 +1,22 @@
+#!/bin/sh
+#SBATCH -p itc_cluster
+#SBATCH -J nylsis 
+#SBATCH -n 1
+#SBATCH -N 1
+#SBATCH -o OUTPUT_frames.%j.out
+#SBATCH -e ERROR_frames.%j.err
+##SBATCH --exclusive
+#SBATCH --mail-user=angus.beane@cfa.harvard.edu
+#SBATCH --mail-type=BEGIN
+#SBATCH --mail-type=END
+#SBATCH --mail-type=FAIL
+#SBATCH --mem-per-cpu=16000
+##SBATCH -t 4-00:00           # Runtime in D-HH:MM
+#SBATCH -t 7-00:00           # Runtime in D-HH:MM
+
+source ../load-modules.sh
+
+ulimit -c unlimited
+
+python3 
+

### 2022-06-08 plots/movie_5panel/job_movie.sh

@@ -1,0 +1,22 @@
+#!/bin/sh
+#SBATCH -p hernquist,conroy
+#SBATCH -J movie 
+#SBATCH -n 64 
+#SBATCH -N 1
+#SBATCH -o OUTPUT_frames.%j.out
+#SBATCH -e ERROR_frames.%j.err
+##SBATCH --exclusive
+#SBATCH --mail-user=angus.beane@cfa.harvard.edu
+#SBATCH --mail-type=BEGIN
+#SBATCH --mail-type=END
+#SBATCH --mail-type=FAIL
+#SBATCH --mem-per-cpu=3900
+##SBATCH -t 4-00:00           # Runtime in D-HH:MM
+#SBATCH -t 0-04:00           # Runtime in D-HH:MM
+
+source ../load-modules.sh
+
+ulimit -c unlimited
+
+python3 run_projections.py ${SLURM_NTASKS}
+

### 2022-06-08 plots/movie_tp/run_movie.py

@@ -1,0 +1,40 @@
+from make_movie import make_movie
+import sys
+import os
+
+base = sys.argv[1]
+lvl  = sys.argv[2]
+
+width = 30.0
+nres = 256
+
+star_vmin = 1E-3
+star_vmax = 1E0
+# gas_vmin  = 0.1 * star_vmin
+# gas_vmax  = 0.1 * star_vmax
+
+data_dir = 'data/'
+
+name = base + '-' + lvl + '_w' + "{:.01f}".format(width) + '_n' + str(nres)
+proj = data_dir + name + '.hdf5'
+
+mov_starxy = 'movies/' + name + '_star_xy.mp4'
+mov_starxz = 'movies/' + name + '_star_xz.mp4'
+mov_testxy = 'movies/' + name + '_test_xy.mp4'
+mov_testxz = 'movies/' + name + '_test_xz.mp4'
+mov_4panel = 'movies/' + name + '_4panel.mp4'
+
+make_movie(proj, [2, 3], 'xy', mov_starxy, vmin=star_vmin, vmax=star_vmax, plot_time=True)
+make_movie(proj, [2, 3], 'xz', mov_starxz, vmin=star_vmin, vmax=star_vmax)
+make_movie(proj, 5, 'xy', mov_testxy, vmin=star_vmin, vmax=star_vmax)
+make_movie(proj, 5, 'xz', mov_testxz, vmin=star_vmin, vmax=star_vmax, plot_time=True)
+
+if os.path.exists('movies/' + name + '_4panel.mp4'):
+    os.remove('movies/' + name + '_4panel.mp4')
+
+ffmpeg_call = 'ffmpeg -i ' + mov_starxz + ' -i ' + mov_testxz + ' -i ' + mov_starxy + ' -i ' + mov_testxy
+ffmpeg_call = ffmpeg_call + ' -filter_complex "[0:v][1:v]hstack[top];[2:v][3:v]hstack[bottom];[top][bottom]vstack[out]" -map "[out]" '
+ffmpeg_call = ffmpeg_call + mov_4panel
+
+os.system(ffmpeg_call)
+

### 2022-06-08 plots/paper_fig-fgas/plot_fig-fgas.py

@@ -1,0 +1,167 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import arepo
+import h5py as h5
+import matplotlib as mpl
+from matplotlib import rc
+import glob
+from tqdm import tqdm
+from scipy.signal import savgol_filter
+import re
+import astropy.units as u
+
+mpl.use('Agg')
+
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
+rc('text', usetex=True)
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+
+snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
+bprop_path = '/n/home01/abeane/starbar/plots/bar_prop/data/'
+torque_path = '/n/home01/abeane/starbar/plots/torques/data/'
+
+tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
+        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
+
+# names
+Nbody = 'Nbody'
+phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
+phS15R35 = 'phantom-vacuum-Sg15-Rc3.5'
+phS10R35 = 'phantom-vacuum-Sg10-Rc3.5'
+phS08R35 = 'phantom-vacuum-Sg08-Rc3.5'
+phS05R35 = 'phantom-vacuum-Sg05-Rc3.5'
+
+lvl = 'lvl3'
+
+def read_fourier(name, lvl, basepath='/n/home01/abeane/starbar/plots/'):
+    f = h5.File(basepath+'/fourier_component/data/fourier_'+name+'-'+lvl+'.hdf5', mode='r')
+    return f
+
+def read_snap(idx, name, lvl, parttype=[0, 2, 3, 4], fields=['Coordinates', 'Masses']):
+    sn = arepo.Snapshot(snap_path+name+'/'+lvl+'/output', idx, combineFiles=True, parttype=parttype,
+                        fields=fields)
+    return sn
+
+def get_sorted_keys(dat):
+    keys = list(dat.keys())
+    # only keep keys that are snapshot keys
+    keys = [k for k in keys if 'snapshot' in k]
+
+    # extract and sort indices
[…]

### 2022-06-08 plots/paper_fig2/plot_fig2.py

@@ -1,0 +1,225 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import arepo
+import h5py as h5
+import matplotlib as mpl
+from matplotlib import rc
+import glob
+from tqdm import tqdm
+from scipy.signal import savgol_filter
+
+mpl.use('Agg')
+
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
+rc('text', usetex=True)
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+
+snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
+bprop_path = '/n/home01/abeane/starbar/plots/bar_prop/data/'
+torque_path = '/n/home01/abeane/starbar/plots/torques/data/'
+
+tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
+        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
+
+# names
+Nbody = 'Nbody'
+phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
+
+lvl = 'lvl3'
+
+def fix_bar_angle(bar_angle):
+    out = np.zeros(len(bar_angle))
+    out[0] = bar_angle[0]
+
+    for i in range(1, len(bar_angle)):
+        dphi = bar_angle[i] - bar_angle[i-1]
+        if dphi < -np.pi:
+            dphi += 2.*np.pi
+        
+        out[i] = out[i-1] + dphi
+    
+    return out
+
+def read_bar_prop(name, lvl):
+    t = h5.File(bprop_path + 'bar_prop_' + name + '-' + lvl + '.hdf5', mode='r')
+    out = {}
+    for key in t['bar_prop'].keys():
+        out[key] = t['bar_prop'][key][:]
+    
+    out['bar_angle'] = fix_bar_angle(t['bar_angle'][:])
+    t.close()
+
+
+    # print(out)
+    # print(len(out['bar_angle']))
+    # print(len(out['tlist']))
+
+    out['pattern_speed'] = np.gradient(out['bar_angle'], out['tlist'])
+
+
+    return out
+
+def read_torque(name, lvl):
[…]

### 2022-06-08 plots/bar_angle/plot_bar_angle.py
update pdf

@@ -1,0 +1,41 @@
+import matplotlib.pyplot as plt 
+import numpy as np 
+import pickle
+
+def plot_pattern_speed(name_list, c_list, ls_list, fout, n=5):
+    fig, ax = plt.subplots(1, 1)
+
+    for name, c, ls in zip(name_list, c_list, ls_list):
+        fdata = 'data/bar_angle_' + name + '.p'
+        dat = pickle.load(open(fdata, 'rb'))
+
+        time = dat['time']
+        ba, ps = dat['poly_eval'][n]
+        true_ba = dat['bar_angle']
+        phi = dat['phi']
+        t = ax.plot(time, ba/(2.*np.pi), c=c, ls=ls, label=name)
+        
+        c = t[0].get_color()
+        ax.scatter(time, true_ba/(2.*np.pi), c=c, s=0.2)
+        ax.scatter(time, phi/(2.*np.pi), c=c, s=0.2)
+        #ax.scatter(time[300:400], phi[300:400]/(2.*np.pi), c=c, s=0.2)
+
+    ax.set_xlabel('time [Myr]')
+    ax.set_ylabel('bar angle / 2pi')
+    ax.legend(frameon=False)
+
+    fig.tight_layout()
+    fig.savefig(fout)
+
+if __name__ == '__main__':
+    name_list = ['nbody-lvl5', 'nbody-lvl4', 'nbody-lvl3']
+    #name_list = ['nbody-lvl3']
+    c_list = [None, None, None]
+    ls_list = [None, None, None]
+
+    plot_pattern_speed(name_list, c_list, ls_list, 'bar_angle_nbody.pdf')
+
+    name_list = ['wet-lvl5', 'wet-lvl4', 'wet-lvl3']
+    
+    plot_pattern_speed(name_list, c_list, ls_list, 'bar_angle_wet.pdf')
+

### 2022-06-08 plots/movie_wake/job_movie.sh

@@ -1,0 +1,22 @@
+#!/bin/sh
+#SBATCH -p test 
+#SBATCH -J movie 
+#SBATCH -n 24 
+#SBATCH -N 1
+#SBATCH -o OUTPUT_frames.%j.out
+#SBATCH -e ERROR_frames.%j.err
+##SBATCH --exclusive
+#SBATCH --mail-user=angus.beane@cfa.harvard.edu
+#SBATCH --mail-type=BEGIN
+#SBATCH --mail-type=END
+#SBATCH --mail-type=FAIL
+#SBATCH --mem-per-cpu=3900
+##SBATCH -t 4-00:00           # Runtime in D-HH:MM
+#SBATCH -t 0-06:00           # Runtime in D-HH:MM
+
+source ../load-modules.sh
+
+ulimit -c unlimited
+
+python3 run_projections.py ${SLURM_NTASKS}
+

### 2022-06-08 plots/paper_fig-A2/plot_fig-A2.py

@@ -1,0 +1,83 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import arepo
+import h5py as h5
+import matplotlib as mpl
+from matplotlib import rc
+import glob
+from tqdm import tqdm
+from scipy.signal import savgol_filter
+
+time_conv = 977.79222168
+
+mpl.use('Agg')
+
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
+rc('text', usetex=True)
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+
+snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
+bprop_path = '/n/home01/abeane/starbar/plots/bar_prop/data/'
+fourier_path = '/n/home01/abeane/starbar/plots/fourier_component/data/'
+torque_path = '/n/home01/abeane/starbar/plots/torques/data/'
+
+tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
+        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
+
+# names
+Nbody = 'Nbody'
+phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
+
+lvl = 'lvl3'
+
+def read_fourier(name, lvl):
+    t = h5.File(fourier_path + 'fourier_' + name + '-' + lvl + '.hdf5', mode='r')
+    return t
+
+def extract_t_max_A2A0(fourier, debug=False):
+
+    i = 0
+
+    tlist = np.array(fourier['time'])
+    A2A0list = []
+    while 'snapshot_'+"{:03d}".format(i) in fourier.keys():
+        key = 'snapshot_'+"{:03d}".format(i)
+        A0 = np.array(fourier[key]['A0'])
+        A2r = np.array(fourier[key]['A2r'])
+        A2i = np.array(fourier[key]['A2i'])
+    
+        A2 = np.sqrt(A2r*A2r + A2i*A2i)
+        A2A0list.append(np.max(A2/A0))
+        if debug:
+            print(tlist[i], fourier[key]['Rlist'][np.argmax(A2/A0)], np.max(A2/A0))
+        i += 1
+    
[…]

### 2022-06-08 plots/paper_fig-rot/plot_fig-rot.py

@@ -1,0 +1,183 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import arepo
+import h5py as h5
+import matplotlib as mpl
+from matplotlib import rc
+import glob
+from tqdm import tqdm
+from scipy.signal import savgol_filter
+from scipy.optimize import minimize
+import agama
+
+agama.setUnits(mass=1E10, length=1, velocity=1)
+mpl.use('Agg')
+
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
+rc('text', usetex=True)
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+
+snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
+bprop_path = '/n/home01/abeane/starbar/plots/bar_prop/data/'
+torque_path = '/n/home01/abeane/starbar/plots/torques/data/'
+
+tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
+        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
+
+# names
+Nbody = 'Nbody'
+phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
+
+lvl = 'lvl3'
+
+def fix_bar_angle(bar_angle):
+    out = np.zeros(len(bar_angle))
+    out[0] = bar_angle[0]
+
+    for i in range(1, len(bar_angle)):
+        dphi = bar_angle[i] - bar_angle[i-1]
+        if dphi < -np.pi:
+            dphi += 2.*np.pi
+        
+        out[i] = out[i-1] + dphi
+    
+    return out
+
+def read_bar_prop(name, lvl):
+    t = h5.File(bprop_path + 'bar_prop_' + name + '-' + lvl + '.hdf5', mode='r')
+    out = {}
+    for key in t['bar_prop'].keys():
+        out[key] = t['bar_prop'][key][:]
+    
+    out['bar_angle'] = fix_bar_angle(t['bar_angle'][:])
+    t.close()
+
+
+    # print(out)
+    # print(len(out['bar_angle']))
+    # print(len(out['tlist']))
+
+    out['pattern_speed'] = np.gradient(out['bar_angle'], out['tlist'])
+
+
+    return out
+
[…]

### 2022-06-08 plots/paper_fig3/plot_fig3.py

@@ -1,0 +1,252 @@
+from re import S
+import numpy as np
+import matplotlib.pyplot as plt
+import arepo
+import h5py as h5
+import matplotlib as mpl
+from matplotlib import rc
+import glob
+from tqdm import tqdm
+from scipy.signal import savgol_filter
+from numba import njit
+from scipy.interpolate import interp1d
+from scipy.ndimage import gaussian_filter
+
+mpl.use('Agg')
+
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
+rc('text', usetex=True)
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+
+snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
+fourier_path = '/n/home01/abeane/starbar/plots/fourier_sphere/data/'
+bprop_path = '/n/home01/abeane/starbar/plots/bar_prop/data/'
+
+tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
+        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
+
+# names
+Nbody = 'Nbody'
+phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
+
+lvl = 'lvl3'
+
+def read_snap(idx, name, lvl, parttype=[2, 3, 4], fields=['Coordinates', 'Masses']):
+    sn = arepo.Snapshot(snap_path+name+'/'+lvl+'/output', idx, combineFiles=True, parttype=parttype,
+                        fields=fields)
+    return sn
+
+@njit
+def compute_surface_density(R, mass, Rbins):
+    key = np.where(R < Rbins[-1])[0]
+    R = R[key]
+    
+    surf_dens = np.zeros(len(Rbins)-1)
+    ave_R = np.zeros(len(Rbins)-1)
+    N_in_bin = np.zeros(len(Rbins)-1)
+    
+    for i in range(len(R)):
+        for j in range(len(Rbins)-1):
+            if R[i] >= Rbins[j] and R[i] < Rbins[j+1]:
+                ave_R[j] += R[i]
[…]

### 2022-06-08 plots/veldisp/plot_veldisp.py

@@ -1,0 +1,33 @@
+import matplotlib.pyplot as plt 
+import numpy as np 
+import pickle
+
+def plot_pattern_speed(name_list, c_list, ls_list, fout, n=5):
+    fig, ax = plt.subplots(1, 1)
+
+    for name, c, ls in zip(name_list, c_list, ls_list):
+        fdata = 'data/veldisp_' + name + '.p'
+
+        dat = pickle.load(open(fdata, 'rb'))
+
+        time = dat[:,0]
+        sigmaR = dat[:,1]
+        sigmaz = dat[:,3]
+
+        ax.plot(time, sigmaR/sigmaz, c=c, ls=ls, label=name)
+
+
+    ax.set_xlabel('time [Myr]')
+    ax.set_ylabel('sigmaR/sigmaz')
+    ax.legend(frameon=False)
+
+    fig.tight_layout()
+    fig.savefig(fout)
+
+if __name__ == '__main__':
+    name_list = ['nbody-lvl5']#, 'nbody-lvl4', 'nbody-lvl3']
+    c_list = [None, None, None]
+    ls_list = [None, None, None]
+
+    plot_pattern_speed(name_list, c_list, ls_list, 'veldisp_nbody.pdf')
+

### 2022-06-10 paper/ms.tex
Add mock HST images
Also some other clean up

--- 2022-06-08 paper/ms.tex
@@
 \usepackage{amsmath}
+
+\DeclareMathOperator\arcsinh{arcsinh}
 
@@
 \includegraphics[width=18cm]{fig/fig1.pdf}
-\caption{\textbf{Projected surface densities of simulated barred galaxies with
+\caption{\textbf{Synthetic Hubble Space Telescope images of our simulations with
 and without the interstellar medium.} The upper panels show an \Nbody{} only
 simulation while the lower panels show a simulation which includes the SMUGGLE
-model for the ISM. Columns show different points in time, separated by about
-$1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
-length and strength. In the SMUGGLE run, the bar remains at approximately the
-same length and strength over the course of the simulation. The \Nbody{} model
-is identical to the GALAKOS model, discussed in the text.}\label{fig:overview}
+model for the interstellar medium. Each panel is $20\,\textrm{kpc}$ to a side.
+Columns show different points in time, separated by about $1\,\textrm{Gyr}$.
+We can see that in the \Nbody{} run, the bar grows in length and strength. In
+the SMUGGLE run, the bar remains at approximately the same length and strength
+over the course of the simulation. The \Nbody{} model is identical to the
+GALAKOS model, discussed in the text. Details on the production of these
+synthetic images is given in the Methods section.}\label{fig:overview}
 \end{figure*}
@@
 akin to dynamical friction. This wake lags a bar and thus exerts a negative
-torque on it.
+torque on it. For a review of bar dynamics, see ref.\cite{1993RPPh...56..173S}
 
@@
 galaxies\cite{2009ARAA..47..159B}) would find slowly rotating bars. There does
-exist one such example of a galaxy which is known to be a slow rotator -- the
+exist one example of a galaxy which is known to be a slow rotator -- the
 low surface brightness galaxy UGC 628.\cite{2009AA...499L..25C} This galaxy
@@
 Computing (HEC) Program through the NASA Advanced Supercomputing (NAS)
-Division at Ames Research Center. A.B. was supported by the Future
-Investigators in NASA Earth and Space Science and Technology (FINESST) award
-number 80NSSC20K1536 during the completion of this work. L.H. funding. E.D.
-funding. J.Q. acknowledges support from NSF grant
-AST-2008490. L.V.S. funding. P.T. acknowledges support from NSF grant
-AST-1909933, AST-2008490, and NASA ATP Grant 80NSSC20K0502. M.V. acknowledges
-support through NASA ATP 19-ATP19-0019, 19-ATP19-0020, 19-ATP19-0167, and NSF
-grants AST-1814053, AST-1814259, AST-1909831, AST-2007355 and AST-2107724.
+Division at Ames Research Center. This research has made use of the Spanish
+Virtual Observatory (https://svo.cab.inta-csic.es) project funded by
+MCIN/AEI/10.13039/501100011033/ through grant PID2020-112949GB-I00. A.B. was
+supported by the Future Investigators in NASA Earth and Space Science and
+Technology (FINESST) award number 80NSSC20K1536 during the completion of this
+work. L.H. funding. E.D. funding. C.C. funding. J.Q. acknowledges support from
+NSF grant AST-2008490. L.V.S. is grateful for financial support from NASA ATP
+80NSSC20K0566, NSF AST 1817233 and NSF CAREER 1945310 grants. P.T.
+acknowledges support from NSF grant AST-1909933, AST-2008490, and NASA ATP
+Grant 80NSSC20K0502. M.V. acknowledges support through NASA ATP 19-ATP19-0019,
+19-ATP19-0020, 19-ATP19-0167, and NSF grants AST-1814053, AST-1814259,
+AST-1909831, AST-2007355 and AST-2107724.
 
@@
 \noindent
-{\bf Initial Setup}
+{\bf Initial setup}
 \\
@@
 the stellar disk. We used an initial surface density of
-$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\textrm{kpc}$.
+$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\textrm{kpc}$. The
+initial gas disk is generated with a temperature of $10^4\,\textrm{K}$ and
+solar metallicity.
 
@@
 guards against an initial dramatic infall of gas within the bar region, which
-we found to destroy the bar. Models of the Milky Way's gas surface density
-profile indicate a reduction in gas in the bar region from $\sim1$ to
-$\sim4\,\textrm{kpc}$,\cite{2022ApJ...929L..18E} and so our practice of
-allowing the gas distribution to have a hole in the central region is
-consistent with our choice to begin the simulations with a bar already formed.
+we found to destroy the bar. It is not uncommon for observed barrred galaxies
+to have gas deficits in the bar region (though not in the very
+center),\cite{1993RPPh...56..173S} and so our practice of allowing the gas
+distribution to have a hole in the central region is consistent with our
+choice to begin the simulations with a bar already formed.
 
@@
 \noindent
-{\bf SMUGGLE Model}
+{\bf SMUGGLE model}
 \\
@@
 \noindent
-{\bf Smooth Interstellar Medium Model}
+{\bf Smooth interstellar medium model}
 \\
@@
 \noindent
-{\bf Varying Gas Fraction}
+{\bf Varying gas fraction}
 \\
@@
 from \texttt{Python} using \texttt{Cython}. This algorithm is based on the
-\texttt{TREESPH} code.\cite{1989ApJS...70..419H} We constructed a tree using
-only the star particles identified as being trapped in the bar using an
-opening angle of $0.35$. We then queried the tree at the locations of all
+\texttt{TREESPH} code.\cite{1989ApJS...70..419H} We constructed a tree with an opening angle of $0.35$ using
+only the star particles identified as being trapped in the bar. We then queried the tree at the locations of all
 resolution elements in the other components and computed the torque of the bar
@@
 \noindent
-{\bf Plotting Details}
+{\bf Idealized mock light images}
+\\
+\noindent
+We created synthetic mock light images at $1024^2$ resolution in
+Fig.~\ref{fig:overview} for the \textit{Hubble Space Telescope} using a
+post-processed Monte Carlo radiation transfer simulation with the SKIRT9 code
+including secondary emission from dust.\cite{2020AC....3100381C} Emission from
+star particles was calculated assuming a Bruzual-Charlot spectral energy
+distribution\cite{2003MNRAS.344.1000B} and a Chabrier initial stellar mass
+function\cite{2003PASP..115..763C}. The luminosity from each star particle was
+smoothed using a cubic spline kernel with each star particle being assigned a
+smoothing length equal to its distance to its $32$nd closest neighbor. We
+launched $10^9$ photon packets per simulation segment. For star particles
+which were present in the initial simulation (i.e., initial disk and bulge
+particles), we assumed that they have an age of $5\,\textrm{Gyr}$ and solar
+metallicity. For star particles in the SMUGGLE simulation which are newly
+formed, we used their recorded ages and metallicities.
+
+For the SMUGGLE simulations, we included absorption and emission from dust. We
+assumed the THEMIS (The Heterogeneous dust Evolution Model for Interstellar
+Solids) dust model\cite{2017AA...602A..46J} to convert gas metallicity and
+mass density to dust mass density. For the \Nbody{} simulations we ran the
+SKIRT9 code in a mode without an obscuring medium, but otherwise applied the
+same procedure.
+
+We used the profiles for filters F814W, F606W, and F475W on the Advanced
+Camera for Surveys (ACS) instrument for red, green, and blue pixel values,
+respectively. Filtered fluxes were converted to RGB values using an $\arcsinh$
+scaling following ref.\cite{2004PASP..116..133L} Scalings and cutoff values
+were adjusted manually to create visually appealing images. Filter profiles
+were provided by the Spanish Virtual Observatory Filter
+Service.\cite{2012ivoa.rept.1015R, 2020sea..confE.182R}
+
+\vspace{12pt}
+
+\noindent
+{\bf Plotting details}
 \\

### 2022-06-13 paper/ms.tex
Elena edits
Update ms.tex
Adds some final touches for journal submission.

@@
 %\documentstyle[amssymb,12pt,draft,epsf,palatino]{nature-pvd}
-\documentclass{natureprintstyle}
+\documentclass[twoside]{natureprintstyle}
 \bibliographystyle{naturemag}
@@
 \usepackage{placeins}
+\usepackage{amsmath}
 
-\usepackage{amsmath}
+% Adds line numbers, just for submission stage.
+\usepackage[switch]{lineno}
+\linenumbers
+
+% Defines footer format for the date.
+\usepackage{datetime}
+
+\newdateformat{footerfmt}{%
+\THEDAY\,\,\monthname[\THEMONTH]\,\,\THEYEAR}
+
+% Adds current date to the footer.
+\usepackage{fancyhdr}
+% \fancyhf{}
+\renewcommand{\headrulewidth}{0pt}
+\fancyfoot[RO]{\textsc{\footerfmt\today}}
+\fancyfoot[LE]{\textsc{\footerfmt\today}}
+\pagestyle{fancy}
 
@@
 \begin{document}
+
+% Enforces same format on first page as on subsequent pages.
+\makeatletter
+\let\ps@plain\ps@fancy
+\makeatother
 
@@
 
-\item[Author Contributions] Foo.
+\item[Author Contributions] A.B., L.H., and E.D. contributed to the conception
+of the study. A.B. performed the simulations, led the analysis, and wrote the
+manuscript. F.M. led and, J.Q., L.V.S., P.T., and M.V. contributed to the
+development of the SMUGGLE code. All authors contributed to aspects of the
+analysis.
 
-  \item[Data Availability] Foo.
+\item[Data Availability]  Simulation output and initial conditions will be
+made publicly available.
 
 \item[Code Availability] A public version of the AREPO code is available at
-  \url{https://arepo-code.org/}.
+\url{https://arepo-code.org/}. An associated GitHub repository containing
+analysis and plotting scripts will be made publicly available (currently
+private). For the submission stage, we have included a tarball of this
+repository.
 

### 2022-06-14 paper/ms.tex

@@
 \newcommand{\SMUGGLE}{SMUGGLE}
+\newcommand{\AREPO}{\textsc{AREPO}}
 \newcommand{\Msun}{\ensuremath{M_{\odot}}}
@@
 We have performed a simulation of a disk galaxy using the finite-volume
-gravito-hydrodynamics code AREPO.\cite{2010MNRAS.401..791S} In AREPO, the
+gravito-hydrodynamics code \AREPO{}.\cite{2010MNRAS.401..791S} In \AREPO{}, the
 fluid is discretized as a moving Voronoi mesh. We use the additional physics
@@
 
-\item[Code Availability] A public version of the AREPO code is available at
+\item[Code Availability] A public version of the \AREPO{} code is available at
 \url{https://arepo-code.org/}. An associated GitHub repository containing
@@
 \cite{2019MNRAS.489.4233M} implemented within the moving-mesh, finite-volume
-hydrodynamics and gravity code AREPO \cite{2010MNRAS.401..791S}. The SMUGGLE model
+hydrodynamics and gravity code \AREPO{} \cite{2010MNRAS.401..791S}. The SMUGGLE model
 additionally includes radiative heating and cooling, star

### 2022-06-14 paper/ms.tex
Update frozen figs in paper/

@@
 \centering
-\includegraphics[width=18cm]{fig/fig1.pdf}
+\includegraphics[width=18cm]{fig/mockHST.pdf}
 \caption{\textbf{Synthetic Hubble Space Telescope images of our simulations with
@@
 \centering
-\includegraphics[width=9cm]{fig/fig2.pdf}
+\includegraphics[width=9cm]{fig/ps_Rbar_torque.pdf}
 \caption{\textbf{Properties of the bar in our \boldmath$N$-body and SMUGGLE
@@
 \centering
-\includegraphics[width=18cm]{fig/fig3.pdf}
+\includegraphics[width=18cm]{fig/halo_wake.pdf}
 \caption{\textbf{The wake excited in the dark matter halo.} The dark matter
@@
 \centering
-\includegraphics[width=9cm]{fig/fig-snap700.pdf}
+\includegraphics[width=9cm]{fig/ps_late_start.pdf}
 \caption{\textbf{Pattern speed evolution with a lower initial pattern speed.} We
@@
 \centering
-\includegraphics[width=9cm]{fig/fig-vcirc.pdf}
+\includegraphics[width=9cm]{fig/vcirc.pdf}
 \caption{\textbf{The circular velocity curve of our initial setups.} This
@@
 \centering
-\includegraphics[width=9cm]{fig/fig-GFM.pdf}
+\includegraphics[width=9cm]{fig/ps_GFM.pdf}
 \caption{\textbf{Pattern speed evolution of a smooth ISM model.} This
@@
 \centering
-\includegraphics[width=9cm]{fig/fig-fgas.pdf}
+\includegraphics[width=9cm]{fig/ps_fgas.pdf}
 \caption{\textbf{Pattern speed evolution with varying gas fractions.} We
@@
 \centering
-\includegraphics[width=9cm]{fig/fig4.pdf}
+\includegraphics[width=9cm]{fig/torque_ps.pdf}
 \caption{\textbf{Average torque exerted by gas on a bar which rotates at a
@@
 \centering
-\includegraphics[width=9cm]{fig/fig-surf.pdf}
+\includegraphics[width=9cm]{fig/surf_dens.pdf}
 \caption{\textbf{Evolution of gas and SFR surface density profiles.} The time
@@
 \centering
-\includegraphics[width=9cm]{fig/fig-A2.pdf}
+\includegraphics[width=9cm]{fig/A2_over_A0.pdf}
 \caption{\textbf{Evolution of bar strength.} The bar strength is measured as
@@
 \centering
-\includegraphics[width=18cm]{fig/fig-bar.pdf}
+\includegraphics[width=18cm]{fig/bar_decomp.pdf}
 \caption{\textbf{Disk decomposition into the barred and unbarred disk.} This

### 2022-06-16 paper/ms.tex
Update funding info
Update MW bar references

@@
 We have performed a simulation of a disk galaxy using the finite-volume
-gravito-hydrodynamics code \AREPO{}.\cite{2010MNRAS.401..791S} In \AREPO{}, the
-fluid is discretized as a moving Voronoi mesh. We use the additional physics
-in the galaxy formation module Stars and MUltiphase Gas in GaLaxiEs (SMUGGLE)
-\cite{2019MNRAS.489.4233M}. SMUGGLE is a comprehensive and self-consistent
-galaxy formation model with a wide range of physical processes, including
-radiative heating/cooling, star formation, and stellar feedback. More detailed
-information on SMUGGLE is given in the Methods section. Our disk galaxy is a
-modified version of the GALAKOS model\cite{2020ApJ...890..117D}, which
-consists of a stellar disk, stellar bulge, and dark matter halo. After
-$\sim2.5\,\textrm{Gyr}$ of evolution, the GALAKOS disk forms a bar consistent
-with the Milky Way bar in terms of pattern speed and length
-($\sim40\,\textrm{km}/\textrm{s}/\textrm{kpc}$ and $\sim4.5\,\textrm{kpc}$,
-respectively). We modify this setup to include a gas phase, the details of
-which are provided in the Methods section.
+gravito-hydrodynamics code \AREPO{}.\cite{2010MNRAS.401..791S} In \AREPO{},
+the fluid is discretized as a moving Voronoi mesh. We use the additional
+physics in the galaxy formation module Stars and MUltiphase Gas in GaLaxiEs
+(SMUGGLE) \cite{2019MNRAS.489.4233M}. SMUGGLE is a comprehensive and
+self-consistent galaxy formation model with a wide range of physical
+processes, including radiative heating/cooling, star formation, and stellar
+feedback. More detailed information on SMUGGLE is given in the Methods
+section. Our disk galaxy is a modified version of the GALAKOS
+model\cite{2020ApJ...890..117D}, which consists of a stellar disk, stellar
+bulge, and dark matter halo. After $\sim2.5\,\textrm{Gyr}$ of evolution, the
+GALAKOS disk forms a bar consistent with the Milky Way bar in terms of pattern
+speed and length
+($\sim40\,\textrm{km}/\textrm{s}/\textrm{kpc}$\cite{2022arXiv220606207G} and
+$\sim4.5\,\textrm{kpc}$\cite{2015MNRAS.450.4050W}, respectively). We modify
+this setup to include a gas phase, the details of which are provided in the
+Methods section.
 
@@
 Technology (FINESST) award number 80NSSC20K1536 during the completion of this
-work. L.H. funding. E.D. funding. C.C. funding. J.Q. acknowledges support from
-NSF grant AST-2008490. L.V.S. is grateful for financial support from NASA ATP
-80NSSC20K0566, NSF AST 1817233 and NSF CAREER 1945310 grants. P.T.
-acknowledges support from NSF grant AST-1909933, AST-2008490, and NASA ATP
-Grant 80NSSC20K0502. M.V. acknowledges support through NASA ATP 19-ATP19-0019,
-19-ATP19-0020, 19-ATP19-0167, and NSF grants AST-1814053, AST-1814259,
-AST-1909831, AST-2007355 and AST-2107724.
+work. E.D. was partially supported by HST grants HST-AR-16363.001 and
+HST-AR-16602.006-A and by NASA Award NASA 80NSSC22K0761. J.Q. acknowledges
+support from NSF grant AST-2008490. L.V.S. is grateful for financial support
+from NASA ATP 80NSSC20K0566, NSF AST 1817233 and NSF CAREER 1945310 grants.
+P.T. acknowledges support from NSF grant AST-1909933, AST-2008490, and NASA
+ATP Grant 80NSSC20K0502. M.V. acknowledges support through NASA ATP
+19-ATP19-0019, 19-ATP19-0020, 19-ATP19-0167, and NSF grants AST-1814053,
+AST-1814259, AST-1909831, AST-2007355 and AST-2107724.
 

### 2022-06-16 paper/ms.tex
Small Elena edits

@@
 the bar excites a wake of resonant material in a dark matter halo in a manner
-akin to dynamical friction. This wake lags a bar and thus exerts a negative
+akin to dynamical friction. This wake lags the bar and thus exerts a negative
 torque on it. For a review of bar dynamics, see ref.\cite{1993RPPh...56..173S}
@@
 Methods that more slowly rotating bars at fixed bar strength are more
-efficient at driving as inflows as well. Nonetheless, when the initial pattern
+efficient at driving gas inflows as well. Nonetheless, when the initial pattern
 speed is lower (red line), the addition of gas does not cause the pattern

### 2022-06-16 plots/vcirc/plot_vcirc.py
Updates from Overleaf

@@ -1,93 +1,0 @@
-from re import S
-import numpy as np
-import matplotlib.pyplot as plt
-import arepo
-import h5py as h5
-import matplotlib as mpl
-from matplotlib import rc
-import glob
-from tqdm import tqdm
-from scipy.signal import savgol_filter
-from numba import njit
-from scipy.interpolate import interp1d
-from scipy.ndimage import gaussian_filter
-
-import agama
-agama.setUnits(mass=1E10, length=1, velocity=1)
-
-mpl.use('Agg')
-
-rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
-rc('text', usetex=True)
-mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
-
-snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
-agama_pot_path = '/n/home01/abeane/starbar/analysis/agama_pot/data/'
-
-tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
-        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
-
-# names
-Nbody = 'Nbody'
-phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
-
-lvl = 'lvl3'
-
-def read_snap(idx, name, lvl, parttype=[2, 3, 4], fields=['Coordinates', 'Masses']):
-    sn = arepo.Snapshot(snap_path+name+'/'+lvl+'/output', idx, combineFiles=True, parttype=parttype,
-                        fields=fields)
-    return sn
-
-def read_bar_prop(name, lvl):
-    t = h5.File(bprop_path + 'bar_prop_' + name + '-' + lvl + '.hdf5', mode='r')
-    out = {}
-    for key in t['bar_prop'].keys():
-        out[key] = t['bar_prop'][key][:]
-    
-    out['bar_angle'] = t['bar_angle'][:]
-    t.close()
-    
-    return out
-
-def read_agama_pot(idx, name, lvl):
-    fname = agama_pot_path + 'pot_' + name + '-' + lvl + '/pot_' + name + '-' + lvl + '.' + str(idx) + '.txt'
-    return agama.Potential(fname)
-
[…]

### 2022-06-17 paper/ms.tex
Several small text updates

@@
   2010ApJ...719.1470V}. We propose that the gas phase of the disk and the dark
-  matter halo act in concert to stabilize the bar pattern speed and prevent a
+  matter halo act in concert to stabilize the bar pattern speed and prevent the
   bar from slowing down or speeding up. We find that in a Milky Way-like disk,
@@
 the $x$-$y$ plane of the dark matter halo after an axisymmetric average has
-been subracted. The solid line indicates the direction of the bar while the
+been subtracted. The solid line indicates the direction of the bar while the
 dashed line indicates the direction of the halo wake (both measured by taking
@@
 It has long been known that non-axisymmetric features in a stellar disk act to
-redistribute mass and angular momentum.\cite{1972MNRAS.157....1L} The
+redistribute mass and angular momentum\cite{1972MNRAS.157....1L}. The
 interaction between a bar and a spheroid (either a dark matter halo or stellar
-bulge) has received considerable interest.\cite{1984MNRAS.209..729T,
-1985MNRAS.213..451W} Stellar disks in isolation are prone to developing bars,
+bulge) has received considerable interest\cite{1984MNRAS.209..729T,
+1985MNRAS.213..451W}. Stellar disks in isolation are prone to developing bars,
 \cite{1971ApJ...168..343H} but the presence of a strong spherical potential
 (e.g., from a stellar bulge or dark matter halo) acts to stabilize the disk
-against bar formation.\cite{1973ApJ...186..467O, 1976AJ.....81...30H} Many
+against bar formation\cite{1973ApJ...186..467O, 1976AJ.....81...30H}. Many
 numerical simulations have confirmed that a live dark matter halo generically
@@
 with $\Rot < 1.4$ are considered ``fast rotators'' while galaxies with $\Rot >
-1.4$ are considered ``slow rotators.''\cite{2000ApJ...543..704D} Galaxies with
-$\Rot < 1$ are not thought to be stable.\cite{1980AA....81..198C}
+1.4$ are considered ``slow rotators''\cite{2000ApJ...543..704D}. Galaxies with
+$\Rot < 1$ are not thought to be stable\cite{1980AA....81..198C}.
 Observational estimates of the pattern speeds of bars indicate that nearly all
-galaxies have $1 < \Rot < 1.4$.\cite{2011MSAIS..18...23C, 2015AA...576A.102A,
-2019MNRAS.482.1733G, 2020MNRAS.491.3655G}
+galaxies have $1 < \Rot < 1.4$\cite{2011MSAIS..18...23C, 2015AA...576A.102A,
+2019MNRAS.482.1733G, 2020MNRAS.491.3655G}.
 
@@
 on the bar. However, because gas is collisional, it can participate in
-non-resonant angular momentum exchange with the bar.\cite{2011MNRAS.415.1027H}
+non-resonant angular momentum exchange with the bar\cite{2011MNRAS.415.1027H}.
 Thus, numerical work has shown that the gas phase can have a stronger
 influence on a bar than its contribution to the mass of a galaxy would
-suggest.\cite{2010ApJ...719.1470V, 2013MNRAS.429.1949A}
+suggest\cite{2010ApJ...719.1470V, 2013MNRAS.429.1949A}.
 
 We have performed a simulation of a disk galaxy using the finite-volume
-gravito-hydrodynamics code \AREPO{}.\cite{2010MNRAS.401..791S} In \AREPO{},
+gravito-hydrodynamics code \AREPO{}\cite{2010MNRAS.401..791S}. In \AREPO{},
 the fluid is discretized as a moving Voronoi mesh. We use the additional
 physics in the galaxy formation module Stars and MUltiphase Gas in GaLaxiEs
-(SMUGGLE) \cite{2019MNRAS.489.4233M}. SMUGGLE is a comprehensive and
+(SMUGGLE)\cite{2019MNRAS.489.4233M}. SMUGGLE is a comprehensive and
 self-consistent galaxy formation model with a wide range of physical
@@
 
-A face-on surface density projection of our simulations is shown in
+A mock stellar light image using filter profiles from the \textit{Hubble Space Telescope} is shown in
 Fig.~\ref{fig:overview}. The upper three panels show the disk in the \Nbody{}
@@
 line of numerical research on bars in \Nbody{}
-simulations.\cite{1992ApJ...400...80H, 2000ApJ...543..704D,
+simulations\cite{1992ApJ...400...80H, 2000ApJ...543..704D,
 2002MNRAS.330...35A, 2002ApJ...569L..83A, 2003MNRAS.341.1179A,
 2003MNRAS.346..251O, 2005MNRAS.363..991H, 2006ApJ...637..214M,
-2007MNRAS.375..460W, 2009ApJ...697..293D} However, in the SMUGGLE case the
+2007MNRAS.375..460W, 2009ApJ...697..293D}. However, in the SMUGGLE case the
 pattern speed remains constant. After the first Gyr of evolution, we find that
@@
 they slow down and the radius of corotation
-increases.\cite{2000ApJ...543..704D, 2003MNRAS.341.1179A}
+increases\cite{2000ApJ...543..704D, 2003MNRAS.341.1179A}.
 
@@
 will form a wake which lags behind and exerts a negative torque on the bar,
-which slows it down.\cite{1984MNRAS.209..729T, 1985MNRAS.213..451W,
-1992ApJ...400...80H}\footnote{Since the bar is not a solid body, it is not
+slowing it down\cite{1984MNRAS.209..729T, 1985MNRAS.213..451W,
+1992ApJ...400...80H}.\footnote{Since the bar is not a solid body, it is not
 guaranteed that a negative torque will slow it down - e.g. a negative torque
@@
 This stops the resonance location from changing such that the halo cannot
-reinforce a wake, arresting the process by which the halo can slow the bar
+reinforce the wake, arresting the process by which the halo can slow the bar
 down. We term this process ``exhaustion.''
@@
 observed in a few simulations of barred galaxies with
-gas.\cite{1993AA...268...65F, 2010ApJ...719.1470V} Previous work has argued
+gas\cite{1993AA...268...65F, 2010ApJ...719.1470V}. Previous work has argued
 this is due to the bar torquing gas inwards, but no explanation has been given
@@
 pattern speed for a few Gyr, the halo exerts almost no torque on the
-bar.\cite{2022MNRAS.513..768C} We instead see over a several Gyr time span
+bar\cite{2022MNRAS.513..768C}. We instead see over a several Gyr time span
 that the dark matter halo in our simulation is still able to support some
@@
 tension with our result since the gas fraction of galaxies correlates with
-other galactic properties.\cite{2009ARAA..47..159B} Furthermore, the
+other galactic properties\cite{2009ARAA..47..159B}. Furthermore, the
 measurement of pattern speeds is a delicate process still prone to large error
@@
 regions of dark matter halos to be underdense\cite{1998ApJ...493L...5D,
-2000ApJ...543..704D} or requiring new physics.\cite{2021MNRAS.503.2833R,
-2021MNRAS.508..926R} Second, we show that the role of gas is of paramount
+2000ApJ...543..704D} or requiring new physics\cite{2021MNRAS.503.2833R,
+2021MNRAS.508..926R}. Second, we show that the role of gas is of paramount
 importance in studies which attempt to uncover the nature of dark matter from
-its effect of slowing down the bar.\cite{2021MNRAS.500.4710C,
-2021MNRAS.505.2412C} Third, we provide an explanation for how the Milky Way's
+its effect of slowing down the bar\cite{2021MNRAS.500.4710C,
+2021MNRAS.505.2412C}. Third, we provide an explanation for how the Milky Way's
 bar could be both long-lived and a fast rotator, of which there is some
-observational evidence.\cite{2019MNRAS.490.4740B} And finally, we complicate
+observational evidence\cite{2019MNRAS.490.4740B}. And finally, we complicate
 the picture of radial mixing expected to sculpt the Milky Way's disk
@@
 in conflict with observations by producing bars which rotate too
-slowly.\cite{2017MNRAS.469.1054A, 2019MNRAS.483.2721P, 2021AA...650L..16F}
+slowly\cite{2017MNRAS.469.1054A, 2019MNRAS.483.2721P, 2021AA...650L..16F}.
 However, the pattern speeds of bars in both cosmological simulations and the
@@
 exist one example of a galaxy which is known to be a slow rotator -- the
-low surface brightness galaxy UGC 628.\cite{2009AA...499L..25C} This galaxy
+low surface brightness galaxy UGC 628\cite{2009AA...499L..25C}. This galaxy
 has been studied in detail by ref.\cite{2016MNRAS.463.1751C}, who note that it
@@
 
-\item[Data Availability]  Simulation output and initial conditions will be
-made publicly available.
+\item[Data Availability]  Simulation output at a selection of timesteps as
+well as the initial conditions will be made publicly available. Simulation
+output at higher timestep frequency will be available upon request.
 
@@
 private). For the submission stage, we have included a tarball of this
-repository.
+repository (associated file \texttt{gasbar.tar.gz}).
+
+\item[Supplementary Information] Supplementary information is available for
+this manuscript.
 
@@
 GALAKOS model\cite{2020ApJ...890..117D}, which uses a modified version of
-\texttt{MakeNewDisk}.\cite{2005MNRAS.361..776S} The GALAKOS model has three
+the \texttt{MakeNewDisk} code\cite{2005MNRAS.361..776S}. The GALAKOS model has three
 components - a radially exponential and vertically isothermal stellar disk,
 and a stellar bulge and dark matter halo following a Hernquist
-profile.\cite{1990ApJ...356..359H} All \Nbody{} runs in this work used the same
+profile\cite{1990ApJ...356..359H}. All \Nbody{} runs in this work used the same
 setup parameters as the GALAKOS disk, more details of which can be found in
@@
 and then exponentially declining beyond that radius with the scale-length of
-the stellar disk. We used an initial surface density of
+the stellar disk. Our fiducial model used an initial surface density of
 $20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\textrm{kpc}$. The
@@
 guards against an initial dramatic infall of gas within the bar region, which
-we found to destroy the bar. It is not uncommon for observed barrred galaxies
+we found to destroy the bar. It is not uncommon for observed barred galaxies
 to have gas deficits in the bar region (though not in the very
-center),\cite{1993RPPh...56..173S} and so our practice of allowing the gas
+center)\cite{1993RPPh...56..173S}, and so our practice of allowing the gas
 distribution to have a hole in the central region is consistent with our
@@
 compared to observational estimates for the Milky
-Way.\cite{2019ApJ...871..120E} We see that the circular velocity curve for
+Way\cite{2019ApJ...871..120E}. We see that the circular velocity curve for
 both runs is marginally larger than the Milky Way's, but still comparable. The
@@
 for the ISM which treats its multiphase nature in a subgrid
-fashion.\cite{2003MNRAS.339..289S} This fundamentally differs from the SMUGGLE
+fashion\cite{2003MNRAS.339..289S}. This fundamentally differs from the SMUGGLE
 model, which explicitly resolve the hot and cold phases of the
-ISM.\cite{2019MNRAS.489.4233M} The pattern speed in the smooth ISM case is
+ISM\cite{2019MNRAS.489.4233M}. The pattern speed in the smooth ISM case is
 broadly similar to the evolution in the SMUGGLE case. This shows that the
@@
 We computed the circular velocity curve of our model using the \texttt{AGAMA}
-package.\cite{2019MNRAS.482.1525V} We fit the baryonic component (stellar
+package\cite{2019MNRAS.482.1525V}. We fit the baryonic component (stellar
 disk, bulge, gas, and newly formed stars) with an axisymmetric cylindrical
@@
 Fig.~\ref{fig:vcirc} compared to observational
-estimates.\cite{2019ApJ...871..120E} Our disk is slightly more massive than
-the Milky Way disk, and in the SMUGGLE run the addition of the gas phase
-results in a slightly higher circular velocity.
+estimates\cite{2019ApJ...871..120E}. The SMUGGLE disk (which includes
+additional mass in the form of gas) has a slightly higher circular velocity
+than the \Nbody{} disk which, itself, is slightly higher than the
+observational estimates.
 
@@
 expected values for the Milky
-Way.\cite{2008AA...487..951K,2022ApJ...929L..18E} The discrepancy between $1$
+Way\cite{2008AA...487..951K,2022ApJ...929L..18E}. The discrepancy between $1$
 and $4\,\textrm{kpc}$ in the molecular and SFR surface density is likely due
@@
 used a simple kinematic distance based on an axisymmetric model of the Milky
-Way,\cite{2017ApJ...834...57M} which is not accurate in the bar region where
+Way\cite{2017ApJ...834...57M}, which is not accurate in the bar region where
 gas has large non-circular velocities.
@@
 resolution is closest to ``level 3'' in the AURIGA
-simulations.\cite{2017MNRAS.467..179G} This corresponds to approximately
+simulations\cite{2017MNRAS.467..179G}. This corresponds to approximately
 $6.4\times10^6$ particles in the stellar disk, $1.1\times10^6$ in the bulge,
@@
 allows the CO-to-H$_2$ conversion factor $X_{\textrm{CO}}$ to vary with
-metallicity.\cite{2022ApJ...929L..18E} Molecular gas surface densities were
+metallicity\cite{2022ApJ...929L..18E}. Molecular gas surface densities were
 provided separately (N. Evans, private communication). We see that the
@@
 \cite{2019MNRAS.489.4233M} implemented within the moving-mesh, finite-volume
-hydrodynamics and gravity code \AREPO{} \cite{2010MNRAS.401..791S}. The SMUGGLE model
+hydrodynamics and gravity code \AREPO{}\cite{2010MNRAS.401..791S}. The SMUGGLE model
 additionally includes radiative heating and cooling, star
@@
 $\epsilon = 0.01$. Star formation converts gas cells into star particles which
-represent single stellar populations with an initial mass function according to ref.\cite{2001ApJ...554.1274C} For each star particle, the deposition
+represent single stellar populations with a Chabrier initial mass function\cite{2003PASP..115..763C}. For each star particle, the deposition
 of energy, momentum, mass, and metals from stellar winds and supernovae is modeled.
@@
 treatment. A more detailed description of this model can be found in the
-flagship SMUGGLE paper.\cite{2019MNRAS.489.4233M}
+flagship SMUGGLE paper\cite{2019MNRAS.489.4233M}.
 
@@
 from the gas phase. Only gas within corotation will flow inwards, while gas
-outside corotation will flow outwards.\cite{2011MNRAS.415.1027H} Since the
+outside corotation will flow outwards\cite{2011MNRAS.415.1027H}. Since the
 corotation radius is larger for more slowly rotating bars, it follows that
@@
 torque on the bar by the gas phase at different rotation rates. The result of
-this experiment is illustrated in Fig.~\ref{fig:equil}, which shows that a more
+this experiment is illustrated in Extended Data Fig.~\ref{fig:equil}, which shows that a more
 slowly rotating bar experiences a larger positive torque from the gas.
@@
 After correcting for the periodicity of $\phi_b$, we measured the pattern
-speed as the two-sided finite gradient of $\phi_b$ as a function of time.
+speed as one-half the two-sided finite gradient of $\phi_b$ as a function of
+time.
 
@@
 clear correlation between the bar pattern speed and the bar
-strength.\cite{2003MNRAS.341.1179A} In the SMUGGLE case, we see that the bar
+strength\cite{2003MNRAS.341.1179A}. In the SMUGGLE case, we see that the bar
 strength has an initial drop but then remains at a roughly constant, but
@@
 from \texttt{Python} using \texttt{Cython}. This algorithm is based on the
-\texttt{TREESPH} code.\cite{1989ApJS...70..419H} We constructed a tree with an opening angle of $0.35$ using
+\texttt{TREESPH} code\cite{1989ApJS...70..419H}. We constructed a tree with an opening angle of $0.35$ using
 only the star particles identified as being trapped in the bar. We then queried the tree at the locations of all
@@
 post-processed Monte Carlo radiation transfer simulation with the SKIRT9 code
-including secondary emission from dust.\cite{2020AC....3100381C} Emission from
+including secondary emission from dust\cite{2020AC....3100381C}. Emission from
 star particles was calculated assuming a Bruzual-Charlot spectral energy
@@
 were provided by the Spanish Virtual Observatory Filter
-Service.\cite{2012ivoa.rept.1015R, 2020sea..confE.182R}
+Service\cite{2012ivoa.rept.1015R, 2020sea..confE.182R}.
 

### 2022-08-19 paper/ms.tex
Add MNRAS TeX template

@@
-%\documentstyle[amssymb,12pt,draft,epsf,palatino]{nature-pvd}
-\documentclass[twoside]{natureprintstyle}
-\bibliographystyle{naturemag}
+% region preamble
+% mnras_template.tex 
+%
+% LaTeX template for creating an MNRAS paper
+%
+% v3.0 released 14 May 2015
+% (version numbers match those of mnras.cls)
+%
+% Copyright (C) Royal Astronomical Society 2015
+% Authors:
+% Keith T. Smith (Royal Astronomical Society)
 
-\usepackage{aas_macros}
+% Change log
+%
+% v3.0 May 2015
+%    Renamed to match the new package name
+%    Version number matches mnras.cls
+%    A few minor tweaks to wording
+% v1.0 September 2013
+%    Beta testing only - never publicly released
+%    First version: a simple (ish) template for creating an MNRAS paper
 
-\usepackage{epsfig,caption}
-\usepackage{color}
-\usepackage{bm}
-\usepackage{graphicx}
-\usepackage{longtable}
-% \usepackage{amssymb}
-\usepackage{rotating,xcolor}
-\usepackage{hyperref}
-% \usepackage{tgbonum}
-\usepackage[symbol]{footmisc}
-\usepackage{placeins}
-\usepackage{amsmath}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Basic setup. Most papers should leave these options alone.
+\documentclass[fleqn,usenatbib]{mnras}
 
-% Adds line numbers, just for submission stage.
-\usepackage[switch]{lineno}
-\linenumbers
+% MNRAS is set in Times font. If you don't have this installed (most LaTeX
+% installations will be fine) or prefer the old Computer Modern fonts, comment
+% out the following line
+% Depending on your LaTeX fonts installation, you might get better results with one of these:
+%\usepackage{mathptmx}
+%\usepackage{txfonts}
 
-% Defines footer format for the date.
-\usepackage{datetime}
+% Use vector fonts, so it zooms properly in on-screen viewing software
+% Don't change these lines unless you know what you are doing
+\usepackage[T1]{fontenc}
 
-\newdateformat{footerfmt}{%
-\THEDAY\,\,\monthname[\THEMONTH]\,\,\THEYEAR}
+% Allow "Thomas van Noord" and "Simon de Laguarde" and alike to be sorted by "N" and "L" etc. in the bibliography.
+% Write the name in the bibliography as "\VAN{Noord}{Van}{van} Noord, Thomas"
+\DeclareRobustCommand{\VAN}[3]{#2}
+\let\VANthebibliography\thebibliography
+\def\thebibliography{\DeclareRobustCommand{\VAN}[3]{##3}\VANthebibliography}
 
-% Adds current date to the footer.
-\usepackage{fancyhdr}
-% \fancyhf{}
-\renewcommand{\headrulewidth}{0pt}
-\fancyfoot[RO]{\textsc{\footerfmt\today}}
-\fancyfoot[LE]{\textsc{\footerfmt\today}}
-\pagestyle{fancy}
+%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
 
-\DeclareMathOperator\arcsinh{arcsinh}
+% Only include extra packages if you really need them. Common packages are:
+\usepackage{graphicx}	% Including figure files
+\usepackage{amsmath}	% Advanced maths commands
+\usepackage{amssymb}	% Extra maths symbols
+\usepackage{newtxtext,newtxmath}
 
-\usepackage{fontspec}
-\setmainfont{texgyrepagella}[
-  Extension = .otf,
-  UprightFont = *-regular,
-  BoldFont = *-bold,
-  ItalicFont = *-italic,
-  BoldItalicFont = *-bolditalic,
-]
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\usepackage[misc]{ifsym}
-\usepackage{xcolor}
-\usepackage[labelfont=bf]{caption}
-\DeclareCaptionLabelSeparator{lsep}{ | } 
-\captionsetup{labelsep=lsep}
+%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
 
-\newcommand{\RCR}{\ensuremath{R_{\textrm{CR}}}}
-\newcommand{\Rot}{\ensuremath{\mathcal{R}}}
+% Please keep new commands to a minimum, and use \newcommand not \def to avoid
+% overwriting existing commands. Example:
+%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
 
-\newcommand{\Nbody}{$N$-body}
-\newcommand{\SMUGGLE}{SMUGGLE}
-\newcommand{\AREPO}{\textsc{AREPO}}
-\newcommand{\Msun}{\ensuremath{M_{\odot}}}
+% endregion
 
-\title{Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down}
+%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%
 
-\author{Angus~Beane$^{1*}$, Lars~Hernquist$^1$, Elena~D'Onghia$^{2,3}$,
-Federico~Marinacci$^{4}$, Charlie~Conroy$^{1}$, Jia~Qi$^{5}$,
-Laura~V.~Sales$^{6}$, Paul~Torrey$^{5}$ and Mark~Vogelsberger$^{7}$
+% Title of the paper, and the short title which is used in the headers.
+% Keep the title short and informative.
+\title[Stellar Bars and Gas]{Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down}
+
+\author[A. Beane et al.]{Angus Beane,$^{1}$\thanks{E-mail: angus.beane@cfa.harvard.edu (AB)}
+Lars Hernquist,$^{1}$
+Elena~D'Onghia,$^{2,3}$
+Federico~Marinacci,$^{4}$
+Charlie Conroy,$^{1}$
+Jia~Qi,$^{5}$\newauthor
+Laura~V.~Sales,$^{6}$
+Paul~Torrey,$^{5}$
+Mark~Vogelsberger$^{7}$
+\\
+% List of institutions
+$^{1}$Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA\\
+$^{2}$Department of Physics, University of Wisconsin-Madison, Madison, WI, USA\\
+$^{3}$Department of Astronomy, University of Wisconsin-Madison, Madison, WI, USA\\
+$^{4}$Department of Physics \& Astronomy `Augusto Righi', University of Bologna, Bologna, Italy\\
+$^{5}$Department of Astronomy, University of Florida, Gainesville, FL, USA\\
+$^{6}$Department of Physics \& Astronomy, University of California, Riverside, CA, USA\\
+$^{7}$Department of Physics, Massachusetts Institute of Technology, Cambridge, MA, USA\\
 }
 
+% These dates will be filled out by the publisher
+\date{Accepted XXX. Received YYY; in original form ZZZ}
+
+% Enter the current year, for the copyright statements etc.
+\pubyear{2015}
+
+% Don't change these lines
 \begin{document}
-
-% Enforces same format on first page as on subsequent pages.
-\makeatletter
-\let\ps@plain\ps@fancy
-\makeatother
-
+\label{firstpage}
+\pagerange{\pageref{firstpage}--\pageref{lastpage}}
 \maketitle
 
-{\let\thefootnote\relax\footnote{
-
-\begin{affiliations}
-% 1
-\item Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA.
-
-% 2
-\item Department of Physics, University of Wisconsin-Madison, Madison, WI, USA.
-
-% 3
-\item Department of Astronomy, University of Wisconsin-Madison, Madison, WI, USA.
-
-% 4
-\item Department of Physics \& Astronomy `Augusto Righi', University of Bologna, Bologna, Italy.
-
-% 5
-\item Department of Astronomy, University of Florida, Gainesville, FL, USA.
-
-% 6
-\item Department of Physics \& Astronomy, University of California, Riverside, CA, USA.
-
-% 7
-\item Department of Physics, Massachusetts Institute of Technology, Cambridge, MA, USA.
-
-$^{*}$ \texttt{\mbox{angus.beane@cfa.harvard.edu}}
-
-\end{affiliations}
-}}
-
-% \renewcommand{\thefootnote}{\fnsymbol{footnote}}
-
-\vspace{-3.5mm}
+% Abstract of the paper
 \begin{abstract}
-
 	Elongated bar-like features are ubiquitous in galaxies, occurring at the
 	centers of approximately two-thirds of spiral
-  disks\cite{2000AJ....119..536E, 2007ApJ...657..790M}.  Due to gravitational
+	disks \citep{2000AJ....119..536E, 2007ApJ...657..790M}.  Due to gravitational
 	interactions between the bar and the other components of galaxies, it is
 	expected that angular momentum and matter will redistribute over long (Gyr)
-  timescales in barred galaxies \cite{1972MNRAS.157....1L,
+	timescales in barred galaxies \citep{1972MNRAS.157....1L,
 	1984MNRAS.209..729T, 1985MNRAS.213..451W}. Previous work ignoring the gas
@@
 	the Milky Way arrived at its present state.
-
 \end{abstract}
 
-\vspace{1cm}
+% Select between one and six entries from the list of approved keywords.
+% Don't make up new ones.
+\begin{keywords}
+keyword1 -- keyword2 -- keyword3
+\end{keywords}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{figure*}[h!]%
-\centering
-\includegraphics[width=18cm]{fig/mockHST.pdf}
-\caption{\textbf{Synthetic Hubble Space Telescope images of our simulations with
-and without the interstellar medium.} The upper panels show an \Nbody{} only
-simulation while the lower panels show a simulation which includes the SMUGGLE
-model for the interstellar medium. Each panel is $20\,\textrm{kpc}$ to a side.
-Columns show different points in time, separated by about $1\,\textrm{Gyr}$.
-We can see that in the \Nbody{} run, the bar grows in length and strength. In
-the SMUGGLE run, the bar remains at approximately the same length and strength
-over the course of the simulation. The \Nbody{} model is identical to the
-GALAKOS model, discussed in the text. Details on the production of these
-synthetic images is given in the Methods section.}\label{fig:overview}
-\end{figure*}
+%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
 
-\begin{figure}[h!]%
-\centering
-\includegraphics[width=9cm]{fig/ps_Rbar_torque.pdf}
-\caption{\textbf{Properties of the bar in our \boldmath$N$-body and SMUGGLE
-simulations of a barred galaxy.} The \textit{upper panel} shows the evolution
-of the pattern speed. As expected, the bar in the \Nbody{} run slows down due
-to interactions between the bar and the dark matter halo. However, the bar in
-the SMUGGLE run does not slow down and instead remains at a constant pattern
-speed. The \textit{middle panel} shows the evolution of the bar length. In the
-\Nbody{} case, the bar lengthens. This occurs because as the pattern speed
-drops, bar-like orbits at larger radii are possible. Stars are captured on
-these orbits, lengthening the bar. This process does not occur in the SMUGGLE
-cases since the bar pattern speed is not decreasing, and therefore the bar
-length remains constant. The \textit{lower panel} shows the torque on the bar
-by different components. The solid lines show the torque exerted by the halo
-in both the \Nbody{} and SMUGGLE cases. The dashed line shows the torque
-exerted by the gas phase in the SMUGGLE run (there is no gas in the \Nbody{}
-run). Details on how these properties are calculated is given in the Methods
-section.}\label{fig:prop}
-\end{figure}
+\section{Introduction}
 
-\begin{figure*}[h!]%
-\centering
-\includegraphics[width=18cm]{fig/halo_wake.pdf}
-\caption{\textbf{The wake excited in the dark matter halo.} The dark matter
-halo wake is shown in the \Nbody{} case (\textit{middle panel}) and SMUGGLE case
-(\textit{right panel}) after $2.6\,\textrm{Gyr}$ of evolution. The
-\textit{middle} and \textit{right panels} show a surface density projection in
-the $x$-$y$ plane of the dark matter halo after an axisymmetric average has
-been subtracted. The solid line indicates the direction of the bar while the
-dashed line indicates the direction of the halo wake (both measured by taking
-the second Fourier component within a sphere of all material within a radius
-of $4\,\textrm{kpc}$). The \textit{left panel} shows the time evolution of
-the angle difference between the bar and the halo wake, as measured from the
-second Fourier component. After the first Gyr, the angle difference in the
-SMUGGLE case is smaller than in the \Nbody{} case by about a factor of two,
-reflecting how the dark matter halo in the SMUGGLE case is unable to exert as
-negative a torque on the bar as in the \Nbody{} case.}\label{fig:wake}
-\end{figure*}
+\section{Methods, Observations, Simulations etc.}
 
-\begin{figure}[t!]%
-\centering
-\includegraphics[width=9cm]{fig/ps_late_start.pdf}
-\caption{\textbf{Pattern speed evolution with a lower initial pattern speed.} We
-tested the evolution of our system when gas is added to the \Nbody{} run at a
-later time, but with all other simulation parameters kept the same. The setup
-is therefore identical to our previous runs just with a bar that is larger,
-stronger, and with a lower pattern speed. We find that the pattern speed
-evolution is very similar to our fiducial case, except that the bar retains
-its original pattern speed. This indicates a mechanism which keeps the bar at
-its formation pattern speed, and that there is not a particular pattern speed
-which the system tends to.}
-\label{fig:snap700}
-\end{figure}
+% \subsection{Figures and tables}
 
-It has long been known that non-axisymmetric features in a stellar disk act to
-redistribute mass and angular momentum\cite{1972MNRAS.157....1L}. The
-interaction between a bar and a spheroid (either a dark matter halo or stellar
-bulge) has received considerable interest\cite{1984MNRAS.209..729T,
-1985MNRAS.213..451W}. Stellar disks in isolation are prone to developing bars,
-\cite{1971ApJ...168..343H} but the presence of a strong spherical potential
-(e.g., from a stellar bulge or dark matter halo) acts to stabilize the disk
-against bar formation\cite{1973ApJ...186..467O, 1976AJ.....81...30H}. Many
-numerical simulations have confirmed that a live dark matter halo generically
-exerts a negative torque on a bar, causing its pattern speed to decrease and
-its length and strength to increase\cite{1992ApJ...400...80H,
-2000ApJ...543..704D, 2002MNRAS.330...35A, 2002ApJ...569L..83A,
-2003MNRAS.341.1179A, 2003MNRAS.346..251O, 2005MNRAS.363..991H,
-2006ApJ...637..214M, 2007MNRAS.375..460W, 2009ApJ...697..293D}. Intuitively,
-the bar excites a wake of resonant material in a dark matter halo in a manner
-akin to dynamical friction. This wake lags the bar and thus exerts a negative
-torque on it. For a review of bar dynamics, see ref.\cite{1993RPPh...56..173S}
+% Figures and tables should be placed at logical positions in the text. Don't
+% worry about the exact layout, which will be handled by the publishers.
 
-The expectation that a dark matter halo acts to slow down a bar is in
-conflict with observational estimates of bar pattern speeds. Bar rotation
-rates are typically classified by the dimensionless ratio
-\begin{equation}
-\Rot = \RCR/R_b\textrm{,}
-\end{equation}
-where \RCR{} is the radius of corotation and $R_b$ is the length of the
-bar.\footnote{The radius of corotation \RCR{} is defined for circular orbits
-as the radius at which the orbital frequency is equal to the pattern speed,
-$\Omega_p$, of a given non-axisymmetric feature. In a galaxy with a constant
-circular velocity $V_c$, it is given by $\RCR = V_c / \Omega_p$.} Galaxies
-with $\Rot < 1.4$ are considered ``fast rotators'' while galaxies with $\Rot >
-1.4$ are considered ``slow rotators''\cite{2000ApJ...543..704D}. Galaxies with
-$\Rot < 1$ are not thought to be stable\cite{1980AA....81..198C}.
-Observational estimates of the pattern speeds of bars indicate that nearly all
-galaxies have $1 < \Rot < 1.4$\cite{2011MSAIS..18...23C, 2015AA...576A.102A,
-2019MNRAS.482.1733G, 2020MNRAS.491.3655G}.
+% Figures are referred to as e.g. Fig.~\ref{fig:example_figure}, and tables as
+% e.g. Table~\ref{tab:example_table}.
 
-The role of gas on the evolution of the bar is less well-understood. Since the
-gas phase typically only contributes about $10-20\%$ of the mass of a galaxy
-at the present day, one might naively expect it to have a subdominant effect
-on the bar. However, because gas is collisional, it can participate in
-non-resonant angular momentum exchange with the bar\cite{2011MNRAS.415.1027H}.
-Thus, numerical work has shown that the gas phase can have a stronger
-influence on a bar than its contribution to the mass of a galaxy would
-suggest\cite{2010ApJ...719.1470V, 2013MNRAS.429.1949A}.
+% % Example figure
+% \begin{figure}
+% 	% To include a figure from a file named example.*
+% 	% Allowable file formats are eps or ps if compiling using latex
+% 	% or pdf, png, jpg if compiling using pdflatex
+% 	\includegraphics[width=\columnwidth]{example}
+%     \caption{This is an example figure. Captions appear below each figure.
+% 	Give enough detail for the reader to understand what they're looking at,
+% 	but leave detailed discussion to the main body of the text.}
+%     \label{fig:example_figure}
+% \end{figure}
 
-We have performed a simulation of a disk galaxy using the finite-volume
-gravito-hydrodynamics code \AREPO{}\cite{2010MNRAS.401..791S}. In \AREPO{},
-the fluid is discretized as a moving Voronoi mesh. We use the additional
-physics in the galaxy formation module Stars and MUltiphase Gas in GaLaxiEs
-(SMUGGLE)\cite{2019MNRAS.489.4233M}. SMUGGLE is a comprehensive and
-self-consistent galaxy formation model with a wide range of physical
-processes, including radiative heating/cooling, star formation, and stellar
-feedback. More detailed information on SMUGGLE is given in the Methods
-section. Our disk galaxy is a modified version of the GALAKOS
-model\cite{2020ApJ...890..117D}, which consists of a stellar disk, stellar
-bulge, and dark matter halo. After $\sim2.5\,\textrm{Gyr}$ of evolution, the
-GALAKOS disk forms a bar consistent with the Milky Way bar in terms of pattern
-speed and length
-($\sim40\,\textrm{km}/\textrm{s}/\textrm{kpc}$\cite{2022arXiv220606207G} and
-$\sim4.5\,\textrm{kpc}$\cite{2015MNRAS.450.4050W}, respectively). We modify
-this setup to include a gas phase, the details of which are provided in the
-Methods section.
+% % Example table
+% \begin{table}
+% 	\centering
+% 	\caption{This is an example table. Captions appear above each table.
+% 	Remember to define the quantities, symbols and units used.}
+% 	\label{tab:example_table}
+% 	\begin{tabular}{lccr} % four columns, alignment for each
+% 		\hline
+% 		A & B & C & D\\
+% 		\hline
+% 		1 & 2 & 3 & 4\\
+% 		2 & 4 & 6 & 8\\
+% 		3 & 5 & 7 & 9\\
+% 		\hline
+% 	\end{tabular}
+% \end{table}
 
-A mock stellar light image using filter profiles from the \textit{Hubble Space Telescope} is shown in
-Fig.~\ref{fig:overview}. The upper three panels show the disk in the \Nbody{}
-run while the lower three panels show the disk in the SMUGGLE run. Each column
-shows the disk $\sim1\,\textrm{Gyr}$ apart in time. There is a large
-qualitative difference in the evolution of the bar pattern between the two
-runs. We see that in the \Nbody{} case, the bar lengthens in time and grows in
-strength. In the SMUGGLE case, the bar retains a similar length and strength
-between panels.
+\section{Conclusions}
 
-We show the time evolution of different bar properties in Fig.~\ref{fig:prop}.
-In the upper panel, we show the pattern speed over time in the \Nbody{} (blue)
-and SMUGGLE (orange) runs. The pattern speed in the \Nbody{} case slows down
-while the pattern speed in the SMUGGLE case remains roughly constant. The
-slowing down of the pattern speed in the \Nbody{} case is consistent with a long
-line of numerical research on bars in \Nbody{}
-simulations\cite{1992ApJ...400...80H, 2000ApJ...543..704D,
-2002MNRAS.330...35A, 2002ApJ...569L..83A, 2003MNRAS.341.1179A,
-2003MNRAS.346..251O, 2005MNRAS.363..991H, 2006ApJ...637..214M,
-2007MNRAS.375..460W, 2009ApJ...697..293D}. However, in the SMUGGLE case the
-pattern speed remains constant. After the first Gyr of evolution, we find that
-the pattern speed increases by only $\sim10\%$ over the next
-$4\,\textrm{Gyr}$, compared to a $\sim43\%$ decrease in the pattern speed for
-the \Nbody{} run over the same interval. As we saw qualitatively in
-Fig.~\ref{fig:overview}, the length of the bar in the \Nbody{} case grows over
-time while it remains roughly constant in the SMUGGLE case. This is also
-consistent with previous numerical work, which found that bars tend to grow as
-they slow down and the radius of corotation
-increases\cite{2000ApJ...543..704D, 2003MNRAS.341.1179A}.
+The last numbered section should briefly summarise what has been done, and describe
+the final conclusions which the authors draw from their work.
 
-The bottom panel of Fig.~\ref{fig:prop} shows the torque exerted on the bar by
-different components. The solid lines indicate the torque on the bar by the
-dark matter halo, whereas the dashed line indicates the torque on the bar by
-the gas phase. In the \Nbody{} case, the halo exerts a steady negative torque on
-the bar, with an average torque from $1$ to $4\,\textrm{Gyr}$ of $-58.0$ in
-units of $10^{10}M_{\odot}\,(\textrm{km}/\textrm{s})^2$. The halo in the
-SMUGGLE case exerts a similar negative torque on the bar in the first Gyr of
-evolution, but after that the halo exerts a much smaller torque on the bar,
-averaging only $-7.8$ in the same units and over the same time interval. The gas in
-the SMUGGLE case exerts a steady positive torque averaging $11.7$ over $1\,\textrm{Gyr}$ in the same
-units.
+\section*{Acknowledgements}
 
-The fact that the dark matter halo in the SMUGGLE case exerts a smaller
-negative torque on the bar can be understood in terms of the halo wake
-mechanism. In the \Nbody{} case, halo material which is resonant with the bar
-will form a wake which lags behind and exerts a negative torque on the bar,
-slowing it down\cite{1984MNRAS.209..729T, 1985MNRAS.213..451W,
-1992ApJ...400...80H}.\footnote{Since the bar is not a solid body, it is not
-guaranteed that a negative torque will slow it down - e.g. a negative torque
-could shred the bar, reducing its moment of inertia without changing its
-pattern speed. However, the bar seems to empirically respond to a negative
-torque induced by a halo wake by slowing down.} As the bar slows down, the
-location of the resonances in phase space changes, allowing halo material
-newly resonant with the bar to participate in forming the wake. However, the
-gas is a reliable source of positive torque on the bar, speeding the bar up.
-This stops the resonance location from changing such that the halo cannot
-reinforce the wake, arresting the process by which the halo can slow the bar
-down. We term this process ``exhaustion.''
+The Acknowledgements section is not numbered. Here you can thank helpful
+colleagues, acknowledge funding agencies, telescopes and facilities used etc.
+Try to keep it short.
 
-We can test this interpretation by measuring the angle offset between the halo
-wake and the bar. If the wake and the bar are aligned (i.e., there is no angle
-offset), then the wake cannot exert a negative torque on the bar. This angle
-is plotted in the left panel of Fig.~\ref{fig:wake}, which shows that the
-angle offset is larger in the \Nbody{} case than in the SMUGGLE case by about a
-factor of two. The center and right panels of Fig.~\ref{fig:wake} show the halo
-wake with respect to the location of the bar in the \Nbody{} (center) and SMUGGLE
-(right) cases.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Data Availability}
 
-The presence of the gas can arrest the process by which additional material in
-the dark matter halo can contribute to a wake. However, this does not explain
-why the pattern speed in the SMUGGLE case is nearly constant over several Gyr.
-Naively, it would be a coincidence that the bar pattern speed remains constant
-in the SMUGGLE case, resulting from a chance cancellation of the halo and gas
-torques. However, a constant pattern speed in the presence of gas has been
-observed in a few simulations of barred galaxies with
-gas\cite{1993AA...268...65F, 2010ApJ...719.1470V}. Previous work has argued
-this is due to the bar torquing gas inwards, but no explanation has been given
-for why it might remain constant.
+The inclusion of a Data Availability Statement is a requirement for articles published in MNRAS.
+Data Availability Statements provide a standardised format for readers to understand the availability of data underlying the research results described in the article.
+The statement may refer to original data generated in the course of the study or to third-party data analysed in the article.
+The statement should describe and provide means of access, where possible, by linking to the data or providing the required accession numbers for the relevant databases or DOIs.
 
-We propose that an equilibrium mechanism is responsible for the pattern speed
-remaining approximately constant. In this scenario, residual negative torque
-from the dark matter halo balances out the positive torque from the gas phase.
-It has been shown when an analytic bar is forced to rotate at a constant
-pattern speed for a few Gyr, the halo exerts almost no torque on the
-bar\cite{2022MNRAS.513..768C}. We instead see over a several Gyr time span
-that the dark matter halo in our simulation is still able to support some
-negative torque.
+%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%
 
-We argue that the following occurs. First, the bar is not able to slow down
-quickly enough due to the positive torque of the infalling gas. This causes
-the resonant halo phase space at a particular pattern speed, $\Omega_{p,0}$,
-to become exhausted and no longer able to support a negative torque. Second,
-the gas is still exerting a positive torque on the bar, and therefore the
-pattern speed will again increase. Since at higher pattern speeds the halo has
-not yet been totally exhausted, the halo will once again be able to exert a
-negative torque on the bar. The pattern speed will then settle at a new value
-slightly higher than $\Omega_{p,0}$ where the gas and halo torques cancel.
-Over time, the pattern speed should slowly increase as the halo becomes
-progressively exhausted.
+% The best way to enter references is to use BibTeX:
 
-A clear prediction of this proposed mechanism is that the constant pattern
-speed a particular galaxy will end up is somewhat arbitrary. In the real
-universe for an isolated galaxy it would be the formation pattern speed of the
-bar while in our simulation it is the pattern speed of the bar when gas is
-added to the system. We tested this by adding gas to the system at a later
-time when the bar has further grown and slowed down with time. In our
-particular test, we added the gas at a time when the pattern speed is
-$\sim30\,\textrm{km}/\textrm{s}/\textrm{kpc}$. As shown in
-Fig.~\ref{fig:snap700}, we find that the pattern speed evolution is very
-similar between the two cases (orange and red lines). If anything, the system
-with a lower pattern speed seems to speed up more, which is consistent with
-our picture since the stronger bar should experience a larger torque from the
-gas as it is more efficient at driving gas inflows. We also show in the
-Methods that more slowly rotating bars at fixed bar strength are more
-efficient at driving gas inflows as well. Nonetheless, when the initial pattern
-speed is lower (red line), the addition of gas does not cause the pattern
-speed to quickly return to the higher value of our fiducial simulation (orange
-line).
+\bibliographystyle{mnras}
+\bibliography{ref} % if your bibtex file is called example.bib
 
-We performed two additional tests to explore how our findings depend on the
-assumed model of the interstellar medium and the amount of gas present in the
-system. First, we performed a simulation of the same disk but with a simpler
-model of the interstellar medium\cite{2003MNRAS.339..289S}, closer to standard
-methods used in cosmological simulations of galaxy formation. We find that the
-pattern speed evolution is nearly the same in this case, and so conclude that
-our result is not sensitive to the details of the model for the interstellar
-medium. The details of this experiment are given in the Methods section.
+% Alternatively you could enter them by hand, like this:
+% This method is tedious and prone to error if you have lots of references
+%\begin{thebibliography}{99}
+%\bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
+%Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
+%\bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
+%Others S., 2012, Journal of Interesting Stuff, 17, 198
+%\end{thebibliography}
 
-Second, we performed a set of simulations in which we decreased the initial
-gas mass of the disk. We essentially find for our disk, our proposed mechanism
-operates as long as the gas contributes about $5\%$ of the total mass of the
-disk. If the gas available is below this threshold, then the pattern speed
-will decrease in a similar fashion to standard \Nbody{} simulations.
-Importantly, arbitrarily increasing the gas fraction does not cause the
-pattern speed to continuously increase. This is consistent with our proposed
-equilibrium mechanism and with previous studies of barred galaxies with varying gas fractions.\cite{2010ApJ...719.1470V} The exact critical gas
-fraction above which our mechanism operates is likely dependent on the
-properties of the bar, disk, and dark matter halo -- further work is needed to
-characterize how this gas fraction varies with such properties. Details on
-this test in which we varied the gas mass of the disk are given in the Methods
-section.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-We note that ref.\cite{2020MNRAS.491.3655G} found that the rotation parameter
-\Rot{} positively correlates with gas fraction, such that galaxies with higher
-gas fractions are rotating more slowly. However, it is not obvious this is in
-tension with our result since the gas fraction of galaxies correlates with
-other galactic properties\cite{2009ARAA..47..159B}. Furthermore, the
-measurement of pattern speeds is a delicate process still prone to large error
-bars.
+%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%
 
-The implications of our findings are numerous. First, we naturally explain why
-nearly all observed galaxies are fast rotators without requiring the inner
-regions of dark matter halos to be underdense\cite{1998ApJ...493L...5D,
-2000ApJ...543..704D} or requiring new physics\cite{2021MNRAS.503.2833R,
-2021MNRAS.508..926R}. Second, we show that the role of gas is of paramount
-importance in studies which attempt to uncover the nature of dark matter from
-its effect of slowing down the bar\cite{2021MNRAS.500.4710C,
-2021MNRAS.505.2412C}. Third, we provide an explanation for how the Milky Way's
-bar could be both long-lived and a fast rotator, of which there is some
-observational evidence\cite{2019MNRAS.490.4740B}. And finally, we complicate
-the picture of radial mixing expected to sculpt the Milky Way's disk
-\cite{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies upon
-the pattern speed of the bar to change with time (though our work does not
-alter expectations for radial mixing induced by spiral
-arms\cite{2002MNRAS.336..785S}).
+\appendix
 
-Barred galaxies in cosmological simulations of galaxy formation continue to be
-in conflict with observations by producing bars which rotate too
-slowly\cite{2017MNRAS.469.1054A, 2019MNRAS.483.2721P, 2021AA...650L..16F}.
-However, the pattern speeds of bars in both cosmological simulations and the
-real universe can be affected by environmental processes not included in our
-simulation -- e.g., satellite infall\cite{2011Natur.477..301P},
-non-sphericity\cite{2013MNRAS.429.1949A} or rotation\cite{2013MNRAS.434.1287S,
-2014ApJ...783L..18L, 2018MNRAS.476.1331C, 2019MNRAS.488.5788C} in the dark
-matter halo, or perhaps even the gaseous circumgalactic medium. Naturally,
-extending our present work to account for such effects is a crucial next step
-in understanding the formation and evolution of galactic bars.
+\section{Some extra material}
 
-We found that below a certain gas fraction, bars should still be able to slow
-down. Therefore, we expect barred spiral galaxies which have been gas-poor for
-extended periods of time to be rotating very slowly. We therefore predict that
-observations which target such galaxies (e.g., lenticular barred
-galaxies\cite{2009ARAA..47..159B}) would find slowly rotating bars. There does
-exist one example of a galaxy which is known to be a slow rotator -- the
-low surface brightness galaxy UGC 628\cite{2009AA...499L..25C}. This galaxy
-has been studied in detail by ref.\cite{2016MNRAS.463.1751C}, who note that it
-indeed has a low gas fraction for galaxies of its type. We predict a general
-trend that bars in gas-rich spiral galaxies should rotate quickly while bars
-in gas-poor spiral galaxies should rotate slowly.
+If you want to present additional material which would interrupt the flow of the main paper,
+it can be placed in an Appendix which appears after the list of references.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\bibliography{ref}
-
-\begin{addendum}
-
-\item [Acknowledgements] We would like to thank Greg~L. Bryan, Neal~J. Evans,
-Drummond~B. Fielding, Keith Hawkins, Jason~A.~S. Hunt, Sarah~M.~R. Jeffreson,
-Kathryn~V. Johnston, Peter~M.~W. Kalberla, Jürgen Kerp, Julio~F. Navarro,
-Joshua~S. Speagle, Martin~D. Weinberg, and Yanfei Zou for helpful discussions.
-A.B. would like to thank Todd Phillips for helpful discussions. Some
-computations in this paper were run on the FASRC Cannon cluster supported by
-the FAS Division of Science Research Computing Group at Harvard University.
-Resources supporting this work were also provided by the NASA High-End
-Computing (HEC) Program through the NASA Advanced Supercomputing (NAS)
-Division at Ames Research Center. This research has made use of the Spanish
-Virtual Observatory (https://svo.cab.inta-csic.es) project funded by
-MCIN/AEI/10.13039/501100011033/ through grant PID2020-112949GB-I00. A.B. was
-supported by the Future Investigators in NASA Earth and Space Science and
-Technology (FINESST) award number 80NSSC20K1536 during the completion of this
-work. E.D. was partially supported by HST grants HST-AR-16363.001 and
-HST-AR-16602.006-A and by NASA Award NASA 80NSSC22K0761. J.Q. acknowledges
-support from NSF grant AST-2008490. L.V.S. is grateful for financial support
-from NASA ATP 80NSSC20K0566, NSF AST 1817233 and NSF CAREER 1945310 grants.
-P.T. acknowledges support from NSF grant AST-1909933, AST-2008490, and NASA
-ATP Grant 80NSSC20K0502. M.V. acknowledges support through NASA ATP
-19-ATP19-0019, 19-ATP19-0020, 19-ATP19-0167, and NSF grants AST-1814053,
-AST-1814259, AST-1909831, AST-2007355 and AST-2107724.
-
-\item[Author Contributions] A.B., L.H., and E.D. contributed to the conception
-of the study. A.B. performed the simulations, led the analysis, and wrote the
-manuscript. F.M. led and, J.Q., L.V.S., P.T., and M.V. contributed to the
-development of the SMUGGLE code. All authors contributed to aspects of the
-analysis.
-
-\item[Data Availability]  Simulation output at a selection of timesteps as
-well as the initial conditions will be made publicly available. Simulation
-output at higher timestep frequency will be available upon request.
-
-\item[Code Availability] A public version of the \AREPO{} code is available at
-\url{https://arepo-code.org/}. An associated GitHub repository containing
-analysis and plotting scripts will be made publicly available (currently
-private). For the submission stage, we have included a tarball of this
-repository (associated file \texttt{gasbar.tar.gz}).
-
-\item[Supplementary Information] Supplementary information is available for
-this manuscript.
-
-\end{addendum}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\FloatBarrier
-\clearpage
-
-\setcounter{page}{1}
-\setcounter{figure}{0}
-\setcounter{table}{0}
-\renewcommand{\figurename}{Extended Data Figure}
-\renewcommand{\thetable}{Extended Data \arabic{table}}
-
-\noindent
-{\bf \Large \uppercase{Methods} }
-
-\bigskip
-
-\noindent
-{\bf Initial setup}
-\\
-\noindent
-The initial setup of the galactic disk used in this work follows closely the
-GALAKOS model\cite{2020ApJ...890..117D}, which uses a modified version of
-the \texttt{MakeNewDisk} code\cite{2005MNRAS.361..776S}. The GALAKOS model has three
-components - a radially exponential and vertically isothermal stellar disk,
-and a stellar bulge and dark matter halo following a Hernquist
-profile\cite{1990ApJ...356..359H}. All \Nbody{} runs in this work used the same
-setup parameters as the GALAKOS disk, more details of which can be found in
-the original paper.
-
-The addition of the gas phase was done as follows. The version of
-\texttt{MakeNewDisk} used for the original GALAKOS model can generate a gas
-disk which is radially exponential and in vertical gravito-hydrodynamic
-balance. We modified the radial profile of this code in order to allow us to
-generate a disk with a constant surface density within some cut-off radius,
-and then exponentially declining beyond that radius with the scale-length of
-the stellar disk. Our fiducial model used an initial surface density of
-$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\textrm{kpc}$. The
-initial gas disk is generated with a temperature of $10^4\,\textrm{K}$ and
-solar metallicity.
-
-After generating the gaseous disk in this way, we stitched the gas disk
-together with the GALAKOS \Nbody{} disk (and bulge and dark matter halo) after
-the GALAKOS disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The
-purpose of allowing the GALAKOS disk to evolve first for a short period of
-time is to allow for the bar to form unimpacted by the presence of the gas
-(which would normally disrupt the formation of the bar). Throughout this work,
-we consider $t=0$ for the \Nbody{} run to be the time at which we added the
-gas phase for the SMUGGLE run (i.e. we ignore the first $1.5\,\textrm{Gyr}$ of
-evolution of the \Nbody{} disk when the bar is forming). We made one
-additional modification when stitching the gas disk together with the \Nbody{}
-disk - we created a hole within the central $4\,\textrm{kpc}$. This hole
-guards against an initial dramatic infall of gas within the bar region, which
-we found to destroy the bar. It is not uncommon for observed barred galaxies
-to have gas deficits in the bar region (though not in the very
-center)\cite{1993RPPh...56..173S}, and so our practice of allowing the gas
-distribution to have a hole in the central region is consistent with our
-choice to begin the simulations with a bar already formed.
-
-\begin{figure}[t]%
-\centering
-\includegraphics[width=9cm]{fig/vcirc.pdf}
-\caption{\textbf{The circular velocity curve of our initial setups.} This
-curve is shown for the \Nbody{} run (blue) and the SMUGGLE run (orange)
-compared to observational estimates for the Milky
-Way\cite{2019ApJ...871..120E}. We see that the circular velocity curve for
-both runs is marginally larger than the Milky Way's, but still comparable. The
-SMUGGLE circular velocity curve is larger than the \Nbody{} curve due to the
-additional mass in the gas phase.}
-\label{fig:vcirc}
-\end{figure}
-
-\begin{figure}[h!]%
-\centering
-\includegraphics[width=9cm]{fig/ps_GFM.pdf}
-\caption{\textbf{Pattern speed evolution of a smooth ISM model.} This
-evolution is shown for the fiducial disk in the \Nbody{} (blue), SMUGGLE
-(orange), and smooth ISM (red) cases. The smooth ISM model is an older model
-for the ISM which treats its multiphase nature in a subgrid
-fashion\cite{2003MNRAS.339..289S}. This fundamentally differs from the SMUGGLE
-model, which explicitly resolve the hot and cold phases of the
-ISM\cite{2019MNRAS.489.4233M}. The pattern speed in the smooth ISM case is
-broadly similar to the evolution in the SMUGGLE case. This shows that the
-stability of the pattern speed is not simply a result of our assumed model for
-the ISM.}
-\label{fig:GFM}
-\end{figure}
-
-Our setup is initially out of equilibrium, but we found that after
-$\sim500\,\textrm{Myr}$, the entire system has settled into a steady-state
-configuration and initial transients appear not to affect the results after
-this point. The constant surface density of the initial gas disk is important
-for ensuring the gas disk is dense enough in order for comparisons to real
-galaxies to be appropriate.
-
-We computed the circular velocity curve of our model using the \texttt{AGAMA}
-package\cite{2019MNRAS.482.1525V}. We fit the baryonic component (stellar
-disk, bulge, gas, and newly formed stars) with an axisymmetric cylindrical
-spline with $20$ grid points in both the radial and vertical direction
-spanning $0.2$ to $50\,\textrm{kpc}$ in the radial direction and from $0.02$
-to $10\,\textrm{kpc}$ in the vertical direction. We fit the dark matter halo
-using a spherically symmetric multipole fit with a maximum angular harmonic
-coefficient of $l=2$. We plot the circular velocity curve in Extended Data
-Fig.~\ref{fig:vcirc} compared to observational
-estimates\cite{2019ApJ...871..120E}. The SMUGGLE disk (which includes
-additional mass in the form of gas) has a slightly higher circular velocity
-than the \Nbody{} disk which, itself, is slightly higher than the
-observational estimates.
-
-We also show the evolution of the surface density profile in Extended Data
-Fig.~\ref{fig:surf} We find that in our simulation the atomic and molecular
-gas surface density and the SFR surface density is broadly consistent with the
-expected values for the Milky
-Way\cite{2008AA...487..951K,2022ApJ...929L..18E}. The discrepancy between $1$
-and $4\,\textrm{kpc}$ in the molecular and SFR surface density is likely due
-to the fact that the distances to molecular clouds which underlines this work
-used a simple kinematic distance based on an axisymmetric model of the Milky
-Way\cite{2017ApJ...834...57M}, which is not accurate in the bar region where
-gas has large non-circular velocities.
-
-We used a mass resolution of $7.5\times10^3\,M_{\odot}$ for the baryonic
-components (initial stellar disk, stellar bulge, and gas) and a mass
-resolution of $3.75\times10^4\,M_{\odot}$ for the dark matter halo. This mass
-resolution is closest to ``level 3'' in the AURIGA
-simulations\cite{2017MNRAS.467..179G}. This corresponds to approximately
-$6.4\times10^6$ particles in the stellar disk, $1.1\times10^6$ in the bulge,
-$1.2\times10^6$ in the gas disk, and $25.3\times10^6$ in the dark matter halo.
-We used a softening length of $0.02\,\textrm{kpc}$ for all components.
-Snapshots were saved at equal intervals of $0.005$ in the time units of the
-simulation ($\textrm{kpc}/(\textrm{km}/\textrm{s})\sim1\,\textrm{Gyr}$).
-
-\begin{figure}[t!]%
-\centering
-\includegraphics[width=9cm]{fig/ps_fgas.pdf}
-\caption{\textbf{Pattern speed evolution with varying gas fractions.} We
-explored the impact of lowering the initial gas surface density of our
-fiducial disk on the evolution of the pattern speed. The surface densities we
-tested of $20$, $15$, $10$, and $5\,\Msun/\textrm{kpc}^2$ correspond to
-initial gas fractions of $16\%$, $10\%$, $7\%$, and $4\%$. We find that
-initial surface densities $20$, $15$, and $10\,\Msun/\textrm{kpc}^2$ result in
-bars which remain at a constant pattern speed, while an initial surface
-density of $5\,\Msun/\textrm{kpc}^2$ results in a bar which slows down in
-roughly the same manner as the \Nbody{} case. }
-\label{fig:fgas}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=9cm]{fig/torque_ps.pdf}
-\caption{\textbf{Average torque exerted by gas on a bar which rotates at a
-fixed pattern speed.} Since only gas within the corotation radius is able to
-infall and slower bars have larger corotation radii, slower bars experience a
-larger net torque than faster bars. The setup of the simulations used here is
-identical to the SMUGGLE case discussed earlier and in the Methods section,
-except the \Nbody{} disk is rotated as a solid body with a constant angular
-velocity.}\label{fig:equil}
-\end{figure}
-
-\begin{figure}[t!]%
-\centering
-\includegraphics[width=9cm]{fig/surf_dens.pdf}
-\caption{\textbf{Evolution of gas and SFR surface density profiles.} The time
-evolution of the atomic gas surface density (\textit{upper}), molecular gas
-surface density (\textit{middle}) and the star formation rate (SFR) surface
-density (\textit{lower}) at various times during our fiducial simulation.
-Colored lines indicate the profiles at selected times during the simulation
-while the black dashed lines indicate observations for the atomic
-gas\cite{2008AA...487..951K} and black dotted lines indicate a model which
-allows the CO-to-H$_2$ conversion factor $X_{\textrm{CO}}$ to vary with
-metallicity\cite{2022ApJ...929L..18E}. Molecular gas surface densities were
-provided separately (N. Evans, private communication). We see that the
-molecular gas and SFR surface densities are within an order of magnitude of
-the Milky Way's typical values at all times.  We see a sharp decrease in the
-gas and SFR surface densities along the extent of the bar from $\sim1$ to
-$\sim4\,\textrm{kpc}$, related to the gas inflow in this region.}
-\label{fig:surf}
-\end{figure}
-
-\vspace{12pt}
-
-\noindent
-{\bf SMUGGLE model}
-\\
-\noindent
-We use the Stars and MUltiphase Gas in GaLaxiEs (SMUGGLE) model
-\cite{2019MNRAS.489.4233M} implemented within the moving-mesh, finite-volume
-hydrodynamics and gravity code \AREPO{}\cite{2010MNRAS.401..791S}. The SMUGGLE model
-additionally includes radiative heating and cooling, star
-formation, and stellar feedback. Explicit gas cooling and heating of the
-multi-phase interstellar medium is implemented, covering temperature ranges
-between $10$ and $10^8\,\textrm{K}$.
-
-Star formation occurs in cells above a density threshold
-($n_{\textrm{th}}=100\,\textrm{cm}^{-3}$) with a star-formation efficiency of
-$\epsilon = 0.01$. Star formation converts gas cells into star particles which
-represent single stellar populations with a Chabrier initial mass function\cite{2003PASP..115..763C}. For each star particle, the deposition
-of energy, momentum, mass, and metals from stellar winds and supernovae is modeled.
-Photo-ionization and radiation pressure are modeled using an approximate
-treatment. A more detailed description of this model can be found in the
-flagship SMUGGLE paper\cite{2019MNRAS.489.4233M}.
-
-We used the fiducial model parameters, except that we increased the number of
-effective neighbors $N_{\textrm{ngb}}$ for the deposition of feedback from
-$64$ to $512$. We found that a lower value of $N_{\textrm{ngb}}$ resulted in
-inefficient photo-ionization feedback since the photo-ionizing budget had not
-been exhausted after deposition into $64$ neighboring cells. We also used an
-updated version of SMUGGLE using a new mechanical feedback routine similar to
-the one described in ref.\cite{2018MNRAS.480..800H} This updated routine is a
-tensor renormalization which ensures linear and angular momentum conservation
-to machine precision.
-
-\vspace{12pt}
-
-\noindent
-{\bf Smooth interstellar medium model}
-\\
-\noindent
-In addition to the SMUGGLE model, we considered a simpler model of the
-interstellar medium based upon ref.\cite{2003MNRAS.339..289S} In this model,
-the multiphase nature of the interstellar medium is modeled in a subgrid
-manner by allowing each resolution element to have a ``cold'' and ``hot''
-component, with the equation of state of the gas suitably modified. Gas is
-allowed to interchange between the ``cold'' and ``hot'' components through
-processes such as cooling and stellar feedback. Cold gas is allowed to undergo
-star formation. The pattern speed evolution of this setup is given in Extended
-Data Fig.~\ref{fig:GFM}, which shows that the pattern speed evolves in a
-similar manner to the SMUGGLE case shown in Fig.~\ref{fig:prop}.
-
-\vspace{12pt}
-
-\noindent
-{\bf Varying gas fraction}
-\\
-\noindent
-We also performed a test in which we varied the initial gas fraction of the
-disk. In our fiducial run, we set the surface density of the gas disk from
-$4\,\textrm{kpc}$ to $\sim9.3\,\textrm{kpc}$ to be $20\,\Msun/\textrm{pc}^2$.
-We also ran with surface densities of $15$, $10$, and
-$5\,\Msun/\textrm{pc}^2$. These correspond to initial gas fractions of
-approximately $16\%$, $10\%$, $7\%$, and $4\%$. The pattern speed evolution is
-shown in Extended Data Fig.~\ref{fig:fgas} We find that the bar in disks with
-initial surface densities of $20$, $15$, and $10\,\Msun/\textrm{pc}^2$ evolve
-with a constant pattern speed while the bar in a disk with initial surface
-density of $5\,\Msun/\textrm{kpc}^2$ slows down at a similar rate to the \Nbody{} case.
-
-As a result, we conclude that for the disk, bar, and halo properties
-considered in this work, a gas fraction of only approximately $5\%$ is
-necessary in order for the proposed stabilizing mechanism to operate.
-
-\vspace{12pt}
-
-\noindent
-{\bf Varying pattern speed}
-\\
-\noindent
-When the bar slows down, we argue that this induces a larger positive torque
-from the gas phase. Only gas within corotation will flow inwards, while gas
-outside corotation will flow outwards\cite{2011MNRAS.415.1027H}. Since the
-corotation radius is larger for more slowly rotating bars, it follows that
-more slowly rotating bars should be more efficient at driving gas inflows and
-thus experience a larger positive torque from the gas phase.
-
-We performed an experiment to test this hypothesis by forcing the stellar disk
-in the SMUGGLE run to rotate at a constant angular rate and measuring the
-torque on the bar by the gas phase at different rotation rates. The result of
-this experiment is illustrated in Extended Data Fig.~\ref{fig:equil}, which shows that a more
-slowly rotating bar experiences a larger positive torque from the gas.
-
-\vspace{12pt}
-
-\noindent
-{\bf Bar analysis}
-\\
-\noindent
-The analysis of various bar properties is performed as follows. First, the
-pattern speed is measured from the angle of the second Fourier component. We
-measured the second Fourier component by computing,
-\begin{equation}
-\begin{split}
-A_2 &= \sum_i m_i e^{i 2 \phi_i} \\
-A_0 &= \sum_i m_i \textrm{,}
-\end{split}
-\end{equation}
-where $m_i$ and $\phi_i$ are the mass and azimuthal angle of each particle,
-respectively. We computed $A_2$ and $A_0$ in cylindrical bins of width
-$0.5\,\textrm{kpc}$ from radii of $0$ to $30\,\textrm{kpc}$. We defined the
-angle of the bar $\phi_b$ to be twice the angle of the complex number $A_2$ as
-measured in the bin extending from a radius of $2.5$ to $3\,\textrm{kpc}$.
-After correcting for the periodicity of $\phi_b$, we measured the pattern
-speed as one-half the two-sided finite gradient of $\phi_b$ as a function of
-time.
-
-\begin{figure}[t!]%
-\centering
-\includegraphics[width=9cm]{fig/A2_over_A0.pdf}
-\caption{\textbf{Evolution of bar strength.} The bar strength is measured as
-the maximum of the second Fourier component divided by the zeroth Fourier
-component. We see that in the
-\Nbody{} case (blue) the bar strength increases with time, consistent with
-previous results showing that the bar strength increases as bars slow down. In
-the SMUGGLE case (orange), we see that the bar strength remains roughly
-constant, possibly slightly decreasing with time. This is also consistent with
-the expected relation between pattern speed and strength since the bar in this
-case is not slowing down.}
-\label{fig:strength}
-\end{figure}
-
-\begin{figure*}[t!]%
-\centering
-\includegraphics[width=18cm]{fig/bar_decomp.pdf}
-\caption{\textbf{Disk decomposition into the barred and unbarred disk.} This
-procedure is based on ref.\cite{2016MNRAS.463.1952P} The \textit{left panel}
-shows a face-on surface density projection through the stellar component of the
-\SMUGGLE{} simulation (disk and bulge) at $t=1\,\textrm{Gyr}$. The \textit{middle panel} shows the
-component of the disk identified as being trapped in the bar while the
-\textit{right panel} shows the component of the disk identified as not being
-trapped in the bar. The fact that the untrapped stars form a roughly
-axisymmetric structure indicates our bar decomposition is sufficiently
-accurate.}
-\label{fig:decomp}
-\end{figure*}
-
-The time evolution of the bar strength, defined as the maximum of
-$\left|A_2/A_0\right|$ as a function of radius, is shown in Extended Data
-Fig.~\ref{fig:strength}. The quantity $\left|A_2/A_0\right|$ varies from $0$
-to $1$, with larger values indicating a stronger bar pattern. We see that in
-the \Nbody{} case, $\left|A_2/A_0\right|$ increases over time as the bar pattern
-slows. This is consistent with previous \Nbody{} simulations which showed a
-clear correlation between the bar pattern speed and the bar
-strength\cite{2003MNRAS.341.1179A}. In the SMUGGLE case, we see that the bar
-strength has an initial drop but then remains at a roughly constant, but
-slightly decreasing, strength. This is consistent with the pattern speed in the
-SMUGGLE case being roughly constant or slightly increasing.
-
-Computing the length of the bar and the torque on the bar by different
-components requires us to decompose the disk into a component which is trapped
-by the bar and a component which is untrapped. In order to do this, we follow
-closely the technique developed in ref.\cite{2016MNRAS.463.1952P} We analyzed
-the orbit of each star particle (meaning initial disk, bulge, and newly formed
-stars) by extracting the $x$-$y$ positions of the apoapse of each in a frame
-corotating with the bar (apoapses are defined as local maxima in $r$). For
-each apoapse, we searched for the $19$ closest apoapses in time and applied a
-$k$-means clustering algorithm on this set of $20$ points with $k=2$. We then
-computed for each of the two clusters the average angle from the bar
-$\left<\Delta \phi\right>_{0,1}$, the standard deviation in $R$ of the points
-${\sigma_R}_{0,1}$, and the average radius of the cluster
-$\left<R\right>_{0,1}$. At each apoapse, a particle was considered to be in
-the bar if it met the following criteria:
-\begin{equation}
-\textrm{max}\left(\left<\Delta \phi\right>_{0,1}\right) < \pi / 8
-\end{equation}
-\begin{equation}
-\frac{{\sigma_R}_0 + {\sigma_R}_1}{\left<R\right>_0 + \left<R\right>_1} < 0.22
-\end{equation}
-These criterion are slightly different and simplified from the ones used in
-ref.\cite{2016MNRAS.463.1952P}, but we found them to empirically work well at
-decomposing the disk into a bar and disk component. In Extended Data
-Fig.~\ref{fig:decomp}, we show an example of this decomposition. The
-\textit{left} panel shows a surface density projection of the stellar disk and
-bulge (including newly formed stars) from the SMUGGLE model after
-$1\,\text{Gyr}$ of evolution in a frame such that the bar is aligned with the
-$x$-axis. The \textit{middle} panel shows a projection of the subset of stars
-that are identified as being trapped in the bar and the \textit{right} panel
-shows a projection of the stars that are not identified as being trapped. The
-fact that the \textit{right} panel is roughly axisymmetric indicates the bar
-decomposition is performing adequately.
-
-After the disk has been decomposed into a trapped and untrapped component, we
-measured the bar length as being the radius $R_b$ which encapsulates $99\%$ of
-the stars identified as being trapped in the bar, allowing for some outliers.
-For the computation of the torque, we used the tree algorithm in
-\texttt{MakeNewDisk}\cite{2005MNRAS.361..776S} customized to be accessible
-from \texttt{Python} using \texttt{Cython}. This algorithm is based on the
-\texttt{TREESPH} code\cite{1989ApJS...70..419H}. We constructed a tree with an opening angle of $0.35$ using
-only the star particles identified as being trapped in the bar. We then queried the tree at the locations of all
-resolution elements in the other components and computed the torque of the bar
-on such components. The torque on the bar by the other components is simply
-the negative of the torque on the other components by the bar. A similar
-method was applied in measuring the torque for the disk when the pattern speed
-is kept constant, as is done in Fig.~\ref{fig:equil}.
-
-\vspace{12pt}
-
-\noindent
-{\bf Idealized mock light images}
-\\
-\noindent
-We created synthetic mock light images at $1024^2$ resolution in
-Fig.~\ref{fig:overview} for the \textit{Hubble Space Telescope} using a
-post-processed Monte Carlo radiation transfer simulation with the SKIRT9 code
-including secondary emission from dust\cite{2020AC....3100381C}. Emission from
-star particles was calculated assuming a Bruzual-Charlot spectral energy
-distribution\cite{2003MNRAS.344.1000B} and a Chabrier initial stellar mass
-function\cite{2003PASP..115..763C}. The luminosity from each star particle was
-smoothed using a cubic spline kernel with each star particle being assigned a
-smoothing length equal to its distance to its $32$nd closest neighbor. We
-launched $10^9$ photon packets per simulation segment. For star particles
-which were present in the initial simulation (i.e., initial disk and bulge
-particles), we assumed that they have an age of $5\,\textrm{Gyr}$ and solar
-metallicity. For star particles in the SMUGGLE simulation which are newly
-formed, we used their recorded ages and metallicities.
-
-For the SMUGGLE simulations, we included absorption and emission from dust. We
-assumed the THEMIS (The Heterogeneous dust Evolution Model for Interstellar
-Solids) dust model\cite{2017AA...602A..46J} to convert gas metallicity and
-mass density to dust mass density. For the \Nbody{} simulations we ran the
-SKIRT9 code in a mode without an obscuring medium, but otherwise applied the
-same procedure.
-
-We used the profiles for filters F814W, F606W, and F475W on the Advanced
-Camera for Surveys (ACS) instrument for red, green, and blue pixel values,
-respectively. Filtered fluxes were converted to RGB values using an $\arcsinh$
-scaling following ref.\cite{2004PASP..116..133L} Scalings and cutoff values
-were adjusted manually to create visually appealing images. Filter profiles
-were provided by the Spanish Virtual Observatory Filter
-Service\cite{2012ivoa.rept.1015R, 2020sea..confE.182R}.
-
-\vspace{12pt}
-
-\noindent
-{\bf Plotting details}
-\\
-\noindent
-We saved snapshots in intervals of $0.005$ in the time units of the
-simulation, $\textrm{kpc}/(\textrm{km}/\textrm{s})$, which is very nearly
-equal to $1\,\textrm{Gyr}$ (it is $\sim0.977\,\textrm{Gyr}$). Therefore,
-throughout this work we referred to the native code time unit as
-$\textrm{Gyr}$. None of our results are sensitive to this choice.
-
-In order to remove numerical noise in several quantities we computed, we
-applied a Savitzky-Golay filter\cite{1964AnaCh..36.1627S} as implemented in
-\texttt{scipy} using a window length of $81$ and polynomial order of $3$. This
-filter was applied to plots of pattern speeds, bar lengths and strengths,
-torques, and angle differences.
-
-% \bibliography{ref2}
-
+% Don't change these lines
+\bsp	% typesetting comment
+\label{lastpage}
 \end{document}
 
-%
+% End of mnras_template.tex

### 2022-08-22 paper/ms.tex
Move old paper, start writing new paper
Switching from nature to mnras

@@
 
+\newcommand{\RCR}{\ensuremath{R_{\textrm{CR}}}}
+\newcommand{\Rot}{\ensuremath{\mathcal{R}}}
+\newcommand{\Vc}{\ensuremath{V_{\textrm{c}} }}
+\newcommand{\PS}{\ensuremath{\Omega_{\textrm{p}}}}
+\newcommand{\Rb}{\ensuremath{R_{\textrm{b}}}}
+
+\newcommand{\Nbody}{$N$-body}
+
 %%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
@@
 
+% region title
 %%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%
@@
 % Enter the current year, for the copyright statements etc.
-\pubyear{2015}
+\pubyear{2022}
+
+% endregion
 
@@
 \begin{abstract}
-	Elongated bar-like features are ubiquitous in galaxies, occurring at the
-	centers of approximately two-thirds of spiral
-	disks \citep{2000AJ....119..536E, 2007ApJ...657..790M}.  Due to gravitational
-	interactions between the bar and the other components of galaxies, it is
-	expected that angular momentum and matter will redistribute over long (Gyr)
-	timescales in barred galaxies \citep{1972MNRAS.157....1L,
-	1984MNRAS.209..729T, 1985MNRAS.213..451W}. Previous work ignoring the gas
-	phase of galaxies has conclusively demonstrated that bars should slow their
-	rotation over time due to their interaction with dark matter halos
-	\cite{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
-	2002ApJ...569L..83A, 2003MNRAS.341.1179A, 2003MNRAS.346..251O,
-	2005MNRAS.363..991H, 2006ApJ...637..214M, 2007MNRAS.375..460W,
-	2009ApJ...697..293D}. We have performed a simulation of a Milky Way-like
-	galactic disk hosting a strong bar which includes a state-of-the-art model
-	of the interstellar medium and a live dark matter halo. In this simulation
-	the bar pattern does not slow down over time, and instead remains at a
-	stable, constant rate of rotation. This behavior has been observed in
-	previous simulations using more simplified models for the interstellar gas
-	but it has remained unexplained\cite{1993AA...268...65F,
-	2010ApJ...719.1470V}. We propose that the gas phase of the disk and the dark
-	matter halo act in concert to stabilize the bar pattern speed and prevent the
-	bar from slowing down or speeding up. We find that in a Milky Way-like disk,
-	a gas fraction of only about \boldmath$5\%$ is necessary for this mechanism
-	to operate. This result naturally explains why nearly all observed bars
-	rotate rapidly\cite{2011MSAIS..18...23C, 2015AA...576A.102A,
-	2019MNRAS.482.1733G} and is especially relevant for our understanding of how
-	the Milky Way arrived at its present state.
+Elongated bar-like features are ubiquitous in galaxies, occurring at the centers
+of approximately two-thirds of spiral disks.  Due to gravitational interactions
+between the bar and the other components of galaxies, it is expected that
+angular momentum and matter will redistribute over long (Gyr) timescales in
+barred galaxies. Previous work ignoring the gas phase of galaxies has
+conclusively demonstrated that bars should slow their rotation over time due to
+their interaction with dark matter halos. We have performed a simulation of a
+Milky Way-like galactic disk hosting a strong bar which includes a
+state-of-the-art model of the interstellar medium and a live dark matter halo.
+In this simulation the bar pattern does not slow down over time, and instead
+remains at a stable, constant rate of rotation. This behavior has been observed
+in previous simulations using more simplified models for the interstellar gas
+but it has remained unexplained. We propose that the gas phase of the disk and
+the dark matter halo act in concert to stabilize the bar pattern speed and
+prevent the bar from slowing down or speeding up. We find that in a Milky
+Way-like disk, a gas fraction of only about 5\% is necessary for this mechanism
+to operate. This result naturally explains why nearly all observed bars rotate
+rapidly and is especially relevant for our understanding of how the Milky Way
+arrived at its present state.
 \end{abstract}
@@
 \section{Introduction}
+Elongated bar-like features are ubiquitous in galaxies, occurring at the centers
+of approximately two-thirds of spiral disks \citep{2000AJ....119..536E,
+2007ApJ...657..790M} and in our own Milky Way \citep{1957AJ.....62...19J,
+1991ApJ...379..631B}. The ubiquity of bars is not difficult to explain, since
+stellar disks simulated in isolation almost always form bar-like structures
+\citep{1971ApJ...168..343H}. Several studies have shown that a spherical
+potential, such as from a stellar bulge or dark matter halo, acts to stabilize a
+stellar disk against the bar instability \citep{1973ApJ...186..467O,
+1976AJ.....81...30H}.
 
-\section{Methods, Observations, Simulations etc.}
+It is more difficult to reconcile numerical simulations with the observed
+pattern speeds of extragalactic bars. Currently, the best technique for
+measuring the pattern speeds of individual galaxies is the Tremaine-Weinberg
+method \citep{1984ApJ...282L...5T, 2011MSAIS..18...23C}. This method has
+recently been applied to samples of galaxies from the MaNGA survey
+\citep{2019MNRAS.482.1733G, 2020MNRAS.491.3655G}. These recent studies confirm
+what was found in earlier studies -- that nearly all extragalactic bars are fast
+rotators (i.e., they rotate close to their maximum rotation rate)
+
+This is a problem for theoretical simulations, for which there is ample evidence
+that galactic bars should resonantly interact with the dark matter halo, causing
+the bar to slow down over time \citep{1992ApJ...400...80H, 2000ApJ...543..704D,
+2002MNRAS.330...35A, 2002ApJ...569L..83A, 2003MNRAS.341.1179A,
+2003MNRAS.346..251O, 2005MNRAS.363..991H, 2006ApJ...637..214M,
+2007MNRAS.375..460W, 2009ApJ...697..293D}. The physical mechanism of this
+interaction can be understood as an angular form of dynamical friction between
+the bar and the dark matter halo. While studied in detail for the bar
+\citep{1984MNRAS.209..729T, 1985MNRAS.213..451W}, this process is generic for
+any non-axisymmetric disturbance \citep{1972MNRAS.157....1L}. For an old but
+still useful review of bar dynamics, see \citet{1993RPPh...56..173S}
+
+Bar pattern speeds are usually measured using the parameter $\Rot=\RCR/\Rb$,
+where \RCR{} is the corotation radius and \Rb{} is the bar length.\footnote{The
+radius of corotation \RCR{} is defined for circular orbits as the radius at
+which the orbital frequency is equal to the pattern speed, \PS{}, of a given
+non-axisymmetric feature. In a galaxy with a constant circular velocity \Vc{},
+it is given by $\RCR = \Vc / \PS$.} Galaxies with $\Rot < 1.4$ are considered
+``fast rotators'' while galaxies with $\Rot > 1.4$ are considered ``slow
+rotators'' \citep{2000ApJ...543..704D}. Galaxies with $\Rot < 1$ are not thought
+to be stable \citep{1980AA....81..198C}. Observational estimates of the pattern
+speeds of bars indicate that nearly all galaxies have $1 < \Rot <
+1.4$ \citep{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G,
+2020MNRAS.491.3655G}.
+
+While the interaction between the bar and a dark matter halo is well-understood
+theoretically, this is not the case for the interaction between a bar and a
+gaseous disk. Some argue that the gas disk should slow down the bar more, while
+others argue that the tendency of the bar to drive gas inwards means the bar
+should speed up due to the effect of the gas disk.
+
+In this work, we present a simulation of an isolated, gas-rich disk galaxy. Our
+galaxy exhibits some similarities to the Milky Way, described in more detail in
+Section XX. Surprisingly, we find that the pattern speed of the bar in this
+galaxy is constant with time. We find that with even modest gas fractions of
+about 5\%, our isolated galaxy displays no evolution in pattern speed over about
+$8\,\textrm{Gyr}$. This behavior has been observed in a few previous works, but
+we are the first to exhibit that the behavior ought to be present in a large
+fraction of galaxies. We are also the first to provide a physical explanation of
+the behavior with clear observational predictions.
+
+In section 2, \dots
+
+% The expectation that a dark matter halo acts to slow down a bar is in conflict
+% with observational estimates of bar pattern speeds. Bar rotation rates are
+% typically classified by the dimensionless ratio
+% % \begin{equation}
+% % \Rot = \RCR/R_b\textrm{,}
+% % \end{equation}
+% where \RCR{} is the radius of corotation and $R_b$ is the length of the
+% bar.\footnote{The radius of corotation \RCR{} is defined for circular orbits as
+% the radius at which the orbital frequency is equal to the pattern speed,
+% $\Omega_p$, of a given non-axisymmetric feature. In a galaxy with a constant
+% circular velocity $V_c$, it is given by $\RCR = V_c / \Omega_p$.} Galaxies with
+% $\Rot < 1.4$ are considered ``fast rotators'' while galaxies with $\Rot > 1.4$
+% are considered ``slow rotators''\cite{2000ApJ...543..704D}. Galaxies with $\Rot
+% < 1$ are not thought to be stable\cite{1980AA....81..198C}. Observational
+% estimates of the pattern speeds of bars indicate that nearly all galaxies have
+% $1 < \Rot < 1.4$\cite{2011MSAIS..18...23C, 2015AA...576A.102A,
+% 2019MNRAS.482.1733G, 2020MNRAS.491.3655G}.
+
+% The role of gas on the evolution of the bar is less well-understood. Since the
+% gas phase typically only contributes about $10-20\%$ of the mass of a galaxy at
+% the present day, one might naively expect it to have a subdominant effect on the
+% bar. However, because gas is collisional, it can participate in non-resonant
+% angular momentum exchange with the bar\cite{2011MNRAS.415.1027H}. Thus,
+% numerical work has shown that the gas phase can have a stronger influence on a
+% bar than its contribution to the mass of a galaxy would
+% suggest\cite{2010ApJ...719.1470V, 2013MNRAS.429.1949A}.
+
+\section{Methods}
+
+\subsection{Initial Conditions}
+The initial setup of the galactic disk used in this work follows closely the
+GALAKOS model \citep{2020ApJ...890..117D}, which uses a modified version of the
+\texttt{MakeNewDisk} code \citep{2005MNRAS.361..776S}. The GALAKOS model has
+three components - a radially exponential and vertically isothermal stellar
+disk, and a stellar bulge and dark matter halo following a Hernquist profile
+\citep{1990ApJ...356..359H}. All \Nbody{} runs in this work used the same setup
+parameters as the GALAKOS disk, more details of which can be found in the
+original paper.
+
+The addition of the gas phase was done as follows. The version of
+\texttt{MakeNewDisk} used for the original GALAKOS model can generate a gas
+disk which is radially exponential and in vertical gravito-hydrodynamic
+balance. We modified the radial profile of this code in order to allow us to
+generate a disk with a constant surface density within some cut-off radius,
+and then exponentially declining beyond that radius with the scale-length of
+the stellar disk. Our fiducial model used an initial surface density of
+$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\textrm{kpc}$. The
+initial gas disk is generated with a temperature of $10^4\,\textrm{K}$ and
+solar metallicity.
+
+After generating the gaseous disk in this way, we stitched the gas disk
+together with the GALAKOS \Nbody{} disk (and bulge and dark matter halo) after
+the GALAKOS disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The
+purpose of allowing the GALAKOS disk to evolve first for a short period of
+time is to allow for the bar to form unimpacted by the presence of the gas
+(which would normally disrupt the formation of the bar). Throughout this work,
+we consider $t=0$ for the \Nbody{} run to be the time at which we added the
+gas phase for the SMUGGLE run (i.e. we ignore the first $1.5\,\textrm{Gyr}$ of
+evolution of the \Nbody{} disk when the bar is forming). We made one
+additional modification when stitching the gas disk together with the \Nbody{}
+disk - we created a hole within the central $4\,\textrm{kpc}$. This hole
+guards against an initial dramatic infall of gas within the bar region, which
+we found to destroy the bar. It is not uncommon for observed barred galaxies
+to have gas deficits in the bar region (though not in the very
+center)\cite{1993RPPh...56..173S}, and so our practice of allowing the gas
+distribution to have a hole in the central region is consistent with our
+choice to begin the simulations with a bar already formed.
+
+\begin{figure}
+\centering
+\includegraphics[width=9cm]{fig/vcirc.pdf}
+\caption{The circular velocity curve of our initial setups. This
+curve is shown for the \Nbody{} run (blue) and the SMUGGLE run (orange)
+compared to observational estimates for the Milky
+Way \citep{2019ApJ...871..120E}. We see that the circular velocity curve for
+both runs is marginally larger than the Milky Way's, but still comparable. The
+SMUGGLE circular velocity curve is larger than the \Nbody{} curve due to the
+additional mass in the gas phase.}
+\label{fig:vcirc}
+\end{figure}
+
+Our setup is initially out of equilibrium, but we found that after
+$\sim500\,\textrm{Myr}$, the entire system has settled into a steady-state
+configuration and initial transients appear not to affect the results after
+this point. The constant surface density of the initial gas disk is important
+for ensuring the gas disk is dense enough in order for comparisons to real
+galaxies to be appropriate.
+
+We computed the circular velocity curve of our model using the \texttt{AGAMA}
+package\cite{2019MNRAS.482.1525V}. We fit the baryonic component (stellar
+disk, bulge, gas, and newly formed stars) with an axisymmetric cylindrical
+spline with $20$ grid points in both the radial and vertical direction
+spanning $0.2$ to $50\,\textrm{kpc}$ in the radial direction and from $0.02$
+to $10\,\textrm{kpc}$ in the vertical direction. We fit the dark matter halo
+using a spherically symmetric multipole fit with a maximum angular harmonic
+coefficient of $l=2$. We plot the circular velocity curve in Extended Data
+Fig.~\ref{fig:vcirc} compared to observational
+estimates\cite{2019ApJ...871..120E}. The SMUGGLE disk (which includes
+additional mass in the form of gas) has a slightly higher circular velocity
+than the \Nbody{} disk which, itself, is slightly higher than the
+observational estimates.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=9cm]{fig/surf_dens.pdf}
+    \caption{The time evolution of the atomic gas surface density
+    (\textit{upper}), molecular gas surface density (\textit{middle}) and the
+    star formation rate (SFR) surface density (\textit{lower}) at various times
+    during our fiducial simulation. Colored lines indicate the profiles at
+    selected times during the simulation while the black dashed lines indicate
+    observations for the atomic gas \citep{2008AA...487..951K} and black dotted
+    lines indicate a model which allows the CO-to-H$_2$ conversion factor
+    $X_{\textrm{CO}}$ to vary with metallicity \citep{2022ApJ...929L..18E}.
+    Molecular gas surface densities were provided separately (N. Evans, private
+    communication). We see that the molecular gas and SFR surface densities are
+    within an order of magnitude of the Milky Way's typical values at all times.
+    We see a sharp decrease in the gas and SFR surface densities along the
+    extent of the bar from $\sim1$ to $\sim4\,\textrm{kpc}$, related to the gas
+    inflow in this region.}
+    \label{fig:surf}
+\end{figure}
+
+We also show the evolution of the surface density profile in Extended Data
+Fig.~\ref{fig:surf} We find that in our simulation the atomic and molecular
+gas surface density and the SFR surface density is broadly consistent with the
+expected values for the Milky
+Way\cite{2008AA...487..951K,2022ApJ...929L..18E}. The discrepancy between $1$
+and $4\,\textrm{kpc}$ in the molecular and SFR surface density is likely due
+to the fact that the distances to molecular clouds which underlines this work
+used a simple kinematic distance based on an axisymmetric model of the Milky
+Way\cite{2017ApJ...834...57M}, which is not accurate in the bar region where
+gas has large non-circular velocities.
+
+We used a mass resolution of $7.5\times10^3\,M_{\odot}$ for the baryonic
+components (initial stellar disk, stellar bulge, and gas) and a mass
+resolution of $3.75\times10^4\,M_{\odot}$ for the dark matter halo. This mass
+resolution is closest to ``level 3'' in the AURIGA
+simulations\cite{2017MNRAS.467..179G}. This corresponds to approximately
+$6.4\times10^6$ particles in the stellar disk, $1.1\times10^6$ in the bulge,
+$1.2\times10^6$ in the gas disk, and $25.3\times10^6$ in the dark matter halo.
+We used a softening length of $0.02\,\textrm{kpc}$ for all components.
+Snapshots were saved at equal intervals of $0.005$ in the time units of the
+simulation ($\textrm{kpc}/(\textrm{km}/\textrm{s})\sim1\,\textrm{Gyr}$).
 

### 2022-08-22 paper/ms.tex
Add ICs Methods section

@@
 The addition of the gas phase was done as follows. The version of
-\texttt{MakeNewDisk} used for the original GALAKOS model can generate a gas
-disk which is radially exponential and in vertical gravito-hydrodynamic
-balance. We modified the radial profile of this code in order to allow us to
-generate a disk with a constant surface density within some cut-off radius,
-and then exponentially declining beyond that radius with the scale-length of
-the stellar disk. Our fiducial model used an initial surface density of
+\texttt{MakeNewDisk} used for the original GALAKOS model can generate a gas disk
+which is radially exponential and in vertical gravito-hydrodynamic balance. We
+modified the radial profile of this code in order to allow us to generate a disk
+with a constant surface density within some cut-off radius, and then
+exponentially declining beyond that radius with the scale-length of the stellar
+disk. Our fiducial model used an initial surface density of
 $20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\textrm{kpc}$. The
-initial gas disk is generated with a temperature of $10^4\,\textrm{K}$ and
-solar metallicity.
+initial gas disk is generated with a temperature of $10^4\,\textrm{K}$ and solar
+metallicity.
 
-After generating the gaseous disk in this way, we stitched the gas disk
-together with the GALAKOS \Nbody{} disk (and bulge and dark matter halo) after
-the GALAKOS disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The
-purpose of allowing the GALAKOS disk to evolve first for a short period of
-time is to allow for the bar to form unimpacted by the presence of the gas
-(which would normally disrupt the formation of the bar). Throughout this work,
-we consider $t=0$ for the \Nbody{} run to be the time at which we added the
-gas phase for the SMUGGLE run (i.e. we ignore the first $1.5\,\textrm{Gyr}$ of
-evolution of the \Nbody{} disk when the bar is forming). We made one
-additional modification when stitching the gas disk together with the \Nbody{}
-disk - we created a hole within the central $4\,\textrm{kpc}$. This hole
-guards against an initial dramatic infall of gas within the bar region, which
-we found to destroy the bar. It is not uncommon for observed barred galaxies
-to have gas deficits in the bar region (though not in the very
-center)\cite{1993RPPh...56..173S}, and so our practice of allowing the gas
-distribution to have a hole in the central region is consistent with our
-choice to begin the simulations with a bar already formed.
+After generating the gaseous disk in this way, we stitched the gas disk together
+with the GALAKOS \Nbody{} disk (and bulge and dark matter halo) after the
+GALAKOS disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The purpose of
+allowing the GALAKOS disk to evolve first for a short period of time is to allow
+for the bar to form unimpacted by the presence of the gas. We found that
+including the gas before the bar has formed disrupts the formation of the bar,
+as has been seen in previous work \citep[e.g.,][]{2013MNRAS.429.1949A}.
+Throughout this work, we consider $t=0$ for the \Nbody{} run to be the time at
+which we added the gas phase for the SMUGGLE run (i.e., we ignore the first
+$1.5\,\textrm{Gyr}$ of evolution of the \Nbody{} disk when the bar is forming).
+
+We made one additional modification when stitching the gas disk together with
+the \Nbody{} disk - we created a hole within the central $4\,\textrm{kpc}$. This
+hole guards against an initial dramatic infall of gas within the bar region,
+which we found to destroy the bar. It is not uncommon for observed barred
+galaxies to have gas deficits in the bar region \citep[though not in the very
+center;][]{1993RPPh...56..173S}. Therefore, our practice of allowing the gas
+distribution to have a hole in the central region is consistent with our choice
+to begin the simulations with a bar already formed.
 
@@
 \end{figure}
+
+We used a mass resolution of $7.5\times10^3\,M_{\odot}$ for the baryonic
+components (initial stellar disk, stellar bulge, and gas) and a mass resolution
+of $3.75\times10^4\,M_{\odot}$ for the dark matter halo. This mass resolution is
+closest to ``level 3'' in the AURIGA simulations \citep{2017MNRAS.467..179G}.
+This corresponds to approximately $6.4\times10^6$ particles in the stellar disk,
+$1.1\times10^6$ in the bulge, $1.2\times10^6$ in the gas disk, and
+$25.3\times10^6$ in the dark matter halo. We used a softening length of
+$0.02\,\textrm{kpc}$ for all components. Snapshots were saved at equal intervals
+of $0.005$ in the time units of the simulation,
+$\textrm{kpc}/(\textrm{km}/\textrm{s})$.
 

### 2022-08-22 paper/ms.tex
Update ms.tex

@@
 \newcommand{\Nbody}{$N$-body}
+\newcommand{\AREPO}{\textsc{AREPO}}
+\newcommand{\SMUGGLE}{SMUGGLE}
 
@@
 
+\subsection{Numerical Model}
+We use the Stars and MUltiphase Gas in GaLaxiEs (SMUGGLE) model
+\citep{2019MNRAS.489.4233M} implemented within the moving-mesh, finite-volume
+hydrodynamics and gravity code \AREPO{} \citep{2010MNRAS.401..791S}. The SMUGGLE
+model additionally includes radiative heating and cooling, star formation, and
+stellar feedback. Explicit gas cooling and heating of the multi-phase
+interstellar medium is implemented, covering temperature ranges between $10$ and
+$10^8\,\textrm{K}$.
+
+Star formation occurs in cells above a density threshold
+($n_{\textrm{th}}=100\,\textrm{cm}^{-3}$) with a star-formation efficiency of
+$\epsilon = 0.01$. Star formation converts gas cells into star particles which
+represent single stellar populations with a Chabrier initial mass
+function \citep{2003PASP..115..763C}. For each star particle, the deposition of
+energy, momentum, mass, and metals from stellar winds and supernovae is modeled.
+Photo-ionization and radiation pressure are modeled using an approximate
+treatment. A more detailed description of this model can be found in the
+flagship SMUGGLE paper \citep{2019MNRAS.489.4233M}.
+
+We used the fiducial model parameters, except that we increased the number of
+effective neighbors $N_{\textrm{ngb}}$ for the deposition of feedback from
+$64$ to $512$. We found that a lower value of $N_{\textrm{ngb}}$ resulted in
+inefficient photo-ionization feedback since the photo-ionizing budget had not
+been exhausted after deposition into $64$ neighboring cells. We also used an
+updated version of SMUGGLE using a new mechanical feedback routine similar to
+the one described in \citet{2018MNRAS.480..800H}. This updated routine is a
+tensor renormalization which ensures linear and angular momentum conservation
+to machine precision.
+
+In addition to the SMUGGLE model, we considered a simpler model of the
+interstellar medium based upon \citet{2003MNRAS.339..289S}. In this model, the
+multiphase nature of the interstellar medium is modeled in a subgrid manner by
+allowing each resolution element to have a ``cold'' and ``hot'' component, with
+the equation of state of the gas suitably modified. Gas is allowed to
+interchange between the ``cold'' and ``hot'' components through processes such
+as cooling and stellar feedback. Cold gas is allowed to undergo star formation.
+We refer to this model as the ``smooth interstellar medium'' model, and it is
+described in more detail in \citet{2019MNRAS.489.4233M}.
+
+\subsection{Bar Analysis}
+The analysis of various bar properties is performed as follows. First, the
+pattern speed is measured from the angle of the second Fourier component. We
+measured the second Fourier component by computing,
+\begin{equation}
+\begin{split}
+A_2 &= \sum_i m_i e^{i 2 \phi_i} \\
+A_0 &= \sum_i m_i \textrm{,}
+\end{split}
+\end{equation}
+where $m_i$ and $\phi_i$ are the mass and azimuthal angle of each particle,
+respectively. We computed $A_2$ and $A_0$ in cylindrical bins of width
+$0.5\,\textrm{kpc}$ from radii of $0$ to $30\,\textrm{kpc}$. We defined the
+angle of the bar $\phi_b$ to be twice the angle of the complex number $A_2$ as
+measured in the bin extending from a radius of $2.5$ to $3\,\textrm{kpc}$.
+After correcting for the periodicity of $\phi_b$, we measured the pattern
+speed as one-half the two-sided finite gradient of $\phi_b$ as a function of
+time.
+
+\begin{figure}
+\centering
+\includegraphics[width=9cm]{fig/A2_over_A0.pdf}
+\caption{\textbf{Evolution of bar strength.} The bar strength is measured as
+the maximum of the second Fourier component divided by the zeroth Fourier
+component. We see that in the
+\Nbody{} case (blue) the bar strength increases with time, consistent with
+previous results showing that the bar strength increases as bars slow down. In
+the SMUGGLE case (orange), we see that the bar strength remains roughly
+constant, possibly slightly decreasing with time. This is also consistent with
+the expected relation between pattern speed and strength since the bar in this
+case is not slowing down.}
+\label{fig:strength}
+\end{figure}
+
+\begin{figure*}
+\centering
+\includegraphics[width=18cm]{fig/bar_decomp.pdf}
+\caption{\textbf{Disk decomposition into the barred and unbarred disk.} This
+procedure is based on \citet{2016MNRAS.463.1952P} The \textit{left panel} shows
+a face-on surface density projection through the stellar component of the
+\SMUGGLE{} simulation (disk and bulge) at $t=1\,\textrm{Gyr}$. The
+\textit{middle panel} shows the component of the disk identified as being
+trapped in the bar while the \textit{right panel} shows the component of the
+disk identified as not being trapped in the bar. The fact that the untrapped
+stars form a roughly axisymmetric structure indicates our bar decomposition is
+sufficiently accurate.}
+\label{fig:decomp}
+\end{figure*}
+
+The time evolution of the bar strength, defined as the maximum of
+$\left|A_2/A_0\right|$ as a function of radius, is shown in Extended Data
+Fig.~\ref{fig:strength}. The quantity $\left|A_2/A_0\right|$ varies from $0$
+to $1$, with larger values indicating a stronger bar pattern. We see that in
+the \Nbody{} case, $\left|A_2/A_0\right|$ increases over time as the bar pattern
+slows. This is consistent with previous \Nbody{} simulations which showed a
+clear correlation between the bar pattern speed and the bar
+strength\cite{2003MNRAS.341.1179A}. In the SMUGGLE case, we see that the bar
+strength has an initial drop but then remains at a roughly constant, but
+slightly decreasing, strength. This is consistent with the pattern speed in the
+SMUGGLE case being roughly constant or slightly increasing.
+
+Computing the length of the bar and the torque on the bar by different
+components requires us to decompose the disk into a component which is trapped
+by the bar and a component which is untrapped. In order to do this, we follow
+closely the technique developed in ref.\cite{2016MNRAS.463.1952P} We analyzed
+the orbit of each star particle (meaning initial disk, bulge, and newly formed
+stars) by extracting the $x$-$y$ positions of the apoapse of each in a frame
+corotating with the bar (apoapses are defined as local maxima in $r$). For
+each apoapse, we searched for the $19$ closest apoapses in time and applied a
+$k$-means clustering algorithm on this set of $20$ points with $k=2$. We then
+computed for each of the two clusters the average angle from the bar
+$\left<\Delta \phi\right>_{0,1}$, the standard deviation in $R$ of the points
+${\sigma_R}_{0,1}$, and the average radius of the cluster
+$\left<R\right>_{0,1}$. At each apoapse, a particle was considered to be in
+the bar if it met the following criteria:
+\begin{equation}
+\textrm{max}\left(\left<\Delta \phi\right>_{0,1}\right) < \pi / 8
+\end{equation}
+\begin{equation}
+\frac{{\sigma_R}_0 + {\sigma_R}_1}{\left<R\right>_0 + \left<R\right>_1} < 0.22
+\end{equation}
+These criterion are slightly different and simplified from the ones used in
+ref.\cite{2016MNRAS.463.1952P}, but we found them to empirically work well at
+decomposing the disk into a bar and disk component. In Extended Data
+Fig.~\ref{fig:decomp}, we show an example of this decomposition. The
+\textit{left} panel shows a surface density projection of the stellar disk and
+bulge (including newly formed stars) from the SMUGGLE model after
+$1\,\text{Gyr}$ of evolution in a frame such that the bar is aligned with the
+$x$-axis. The \textit{middle} panel shows a projection of the subset of stars
+that are identified as being trapped in the bar and the \textit{right} panel
+shows a projection of the stars that are not identified as being trapped. The
+fact that the \textit{right} panel is roughly axisymmetric indicates the bar
+decomposition is performing adequately.
+
+After the disk has been decomposed into a trapped and untrapped component, we
+measured the bar length as being the radius $R_b$ which encapsulates $99\%$ of
+the stars identified as being trapped in the bar, allowing for some outliers.
+For the computation of the torque, we used the tree algorithm in
+\texttt{MakeNewDisk}\cite{2005MNRAS.361..776S} customized to be accessible
+from \texttt{Python} using \texttt{Cython}. This algorithm is based on the
+\texttt{TREESPH} code\cite{1989ApJS...70..419H}. We constructed a tree with an opening angle of $0.35$ using
+only the star particles identified as being trapped in the bar. We then queried the tree at the locations of all
+resolution elements in the other components and computed the torque of the bar
+on such components. The torque on the bar by the other components is simply
+the negative of the torque on the other components by the bar. A similar
+method was applied in measuring the torque for the disk when the pattern speed
+is kept constant, as is done in Fig.~\ref{fig:equil}.
+
+% \subsection{Idealized Mock Light Images}
+% We created synthetic mock light images at $1024^2$ resolution in
+% Fig.~\ref{fig:overview} for the \textit{Hubble Space Telescope} using a
+% post-processed Monte Carlo radiation transfer simulation with the SKIRT9 code
+% including secondary emission from dust\cite{2020AC....3100381C}. Emission from
+% star particles was calculated assuming a Bruzual-Charlot spectral energy
+% distribution\cite{2003MNRAS.344.1000B} and a Chabrier initial stellar mass
+% function\cite{2003PASP..115..763C}. The luminosity from each star particle was
+% smoothed using a cubic spline kernel with each star particle being assigned a
+% smoothing length equal to its distance to its $32$nd closest neighbor. We
+% launched $10^9$ photon packets per simulation segment. For star particles
+% which were present in the initial simulation (i.e., initial disk and bulge
+% particles), we assumed that they have an age of $5\,\textrm{Gyr}$ and solar
+% metallicity. For star particles in the SMUGGLE simulation which are newly
+% formed, we used their recorded ages and metallicities.
+
+% For the SMUGGLE simulations, we included absorption and emission from dust. We
+% assumed the THEMIS (The Heterogeneous dust Evolution Model for Interstellar
+% Solids) dust model\cite{2017AA...602A..46J} to convert gas metallicity and
+% mass density to dust mass density. For the \Nbody{} simulations we ran the
+% SKIRT9 code in a mode without an obscuring medium, but otherwise applied the
+% same procedure.
+
+% We used the profiles for filters F814W, F606W, and F475W on the Advanced
+% Camera for Surveys (ACS) instrument for red, green, and blue pixel values,
+% respectively. Filtered fluxes were converted to RGB values using an $\arcsinh$
+% scaling following ref.\cite{2004PASP..116..133L} Scalings and cutoff values
+% were adjusted manually to create visually appealing images. Filter profiles
+% were provided by the Spanish Virtual Observatory Filter
+% Service\cite{2012ivoa.rept.1015R, 2020sea..confE.182R}.
+
+\subsection{Plotting Details}
+We saved snapshots in intervals of $0.005$ in the time units of the
+simulation, $\textrm{kpc}/(\textrm{km}/\textrm{s})$, which is very nearly
+equal to $1\,\textrm{Gyr}$ (it is $\sim0.977\,\textrm{Gyr}$). Therefore,
+throughout this work we referred to the native code time unit as
+$\textrm{Gyr}$. None of our results are sensitive to this choice.
+
+In order to remove numerical noise in several quantities we computed, we
+applied a Savitzky-Golay filter \citep{1964AnaCh..36.1627S} as implemented in
+\texttt{scipy} using a window length of $81$ and polynomial order of $3$. This
+filter was applied to plots of pattern speeds, bar lengths and strengths,
+torques, and angle differences.
+
+\section{Discussion}
 Our setup is initially out of equilibrium, but we found that after
@@
 
-We used a mass resolution of $7.5\times10^3\,M_{\odot}$ for the baryonic
-components (initial stellar disk, stellar bulge, and gas) and a mass
-resolution of $3.75\times10^4\,M_{\odot}$ for the dark matter halo. This mass
-resolution is closest to ``level 3'' in the AURIGA
-simulations\cite{2017MNRAS.467..179G}. This corresponds to approximately
-$6.4\times10^6$ particles in the stellar disk, $1.1\times10^6$ in the bulge,
-$1.2\times10^6$ in the gas disk, and $25.3\times10^6$ in the dark matter halo.
-We used a softening length of $0.02\,\textrm{kpc}$ for all components.
-Snapshots were saved at equal intervals of $0.005$ in the time units of the
-simulation ($\textrm{kpc}/(\textrm{km}/\textrm{s})\sim1\,\textrm{Gyr}$).
+The pattern speed evolution of this setup is given in Extended Data
+Fig.~\ref{fig:GFM}, which shows that the pattern speed evolves in a similar
+manner to the SMUGGLE case shown in Fig.~\ref{fig:prop}.
 

### 2022-08-24 paper/ms.tex
Further updates to text

@@
 % region preamble
-% mnras_template.tex 
-%
-% LaTeX template for creating an MNRAS paper
-%
-% v3.0 released 14 May 2015
-% (version numbers match those of mnras.cls)
-%
-% Copyright (C) Royal Astronomical Society 2015
-% Authors:
-% Keith T. Smith (Royal Astronomical Society)
 
-% Change log
-%
-% v3.0 May 2015
-%    Renamed to match the new package name
-%    Version number matches mnras.cls
-%    A few minor tweaks to wording
-% v1.0 September 2013
-%    Beta testing only - never publicly released
-%    First version: a simple (ish) template for creating an MNRAS paper
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Basic setup. Most papers should leave these options alone.
 \documentclass[fleqn,usenatbib]{mnras}
 
-% MNRAS is set in Times font. If you don't have this installed (most LaTeX
-% installations will be fine) or prefer the old Computer Modern fonts, comment
-% out the following line
-% Depending on your LaTeX fonts installation, you might get better results with one of these:
-%\usepackage{mathptmx}
-%\usepackage{txfonts}
-
-% Use vector fonts, so it zooms properly in on-screen viewing software
-% Don't change these lines unless you know what you are doing
 \usepackage[T1]{fontenc}
@@
 
-%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
-
-% Only include extra packages if you really need them. Common packages are:
 \usepackage{graphicx}	% Including figure files
@@
 \newcommand{\SMUGGLE}{SMUGGLE}
+\newcommand{\Msun}{\ensuremath{M_{\odot}}}
 
-%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
-
-% Please keep new commands to a minimum, and use \newcommand not \def to avoid
-% overwriting existing commands. Example:
-%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
+\DeclareMathOperator\arcsinh{arcsinh}
 
@@
 \begin{keywords}
-keyword1 -- keyword2 -- keyword3
+galaxies: bar - galaxies: disc - galaxies: evolution - galaxies: kinematics and
+dynamics
 \end{keywords}
@@
 \section{Introduction}
-Elongated bar-like features are ubiquitous in galaxies, occurring at the centers
-of approximately two-thirds of spiral disks \citep{2000AJ....119..536E,
-2007ApJ...657..790M} and in our own Milky Way \citep{1957AJ.....62...19J,
-1991ApJ...379..631B}. The ubiquity of bars is not difficult to explain, since
-stellar disks simulated in isolation almost always form bar-like structures
-\citep{1971ApJ...168..343H}. Several studies have shown that a spherical
-potential, such as from a stellar bulge or dark matter halo, acts to stabilize a
-stellar disk against the bar instability \citep{1973ApJ...186..467O,
-1976AJ.....81...30H}.
+Approximately two-thirds of spiral disks host an elongated bar-like feature at
+their center \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our own
+Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of bars
+is not difficult to explain, since stellar disks simulated in isolation almost
+always form bar-like structures \citep{1971ApJ...168..343H}. Several studies
+have shown that a spherical potential, such as a stellar bulge or dark matter
+halo, acts to stabilize a stellar disk against the bar instability
+\citep{1973ApJ...186..467O, 1976AJ.....81...30H}.
 
@@
 theoretically, this is not the case for the interaction between a bar and a
-gaseous disk. Some argue that the gas disk should slow down the bar more, while
-others argue that the tendency of the bar to drive gas inwards means the bar
-should speed up due to the effect of the gas disk.
+gaseous disk. Some argue that the gas disk should slow down the bar more
+\citep{2003MNRAS.341.1179A}, while others argue that the tendency of the bar to
+drive gas inwards means the bar should speed up due to the effect of the gas
+disk \citep{2013MNRAS.429.1949A, 2014MNRAS.438L..81A}. Since the gas phase
+typically only contributes about $10-20\%$ of the mass of a galaxy at the
+present day, one might naively expect it to have a subdominant effect on the
+bar. However, because gas is collisional, it can participate in non-resonant
+angular momentum exchange with the bar \citep{2011MNRAS.415.1027H}. Thus,
+numerical work has shown that the gas phase can have a stronger influence on a
+bar than its contribution to the mass of a galaxy would suggest
+\citep{2010ApJ...719.1470V, 2013MNRAS.429.1949A}.
 
-In this work, we present a simulation of an isolated, gas-rich disk galaxy. Our
-galaxy exhibits some similarities to the Milky Way, described in more detail in
-Section XX. Surprisingly, we find that the pattern speed of the bar in this
-galaxy is constant with time. We find that with even modest gas fractions of
-about 5\%, our isolated galaxy displays no evolution in pattern speed over about
-$8\,\textrm{Gyr}$. This behavior has been observed in a few previous works, but
-we are the first to exhibit that the behavior ought to be present in a large
-fraction of galaxies. We are also the first to provide a physical explanation of
-the behavior with clear observational predictions.
+We have performed a simulation of a disk galaxy using the finite-volume,
+gravito-hydrodynamics code AREPO \citep{2010MNRAS.401..791S}. We use the galaxy
+formation code Stars and MUltiphase Gas in GaLaxiEs (SMUGGLE). This disk galaxy
+exhibits almost no evolution in the bar pattern speed over several Gyr when the
+gas phase is accounted for and properly modeled. This behavior has been observed
+in a few previous works \citep{1993AA...268...65F, 2010ApJ...719.1470V,
+2014MNRAS.438L..81A}, but we are the first to exhibit that the behavior ought to
+be present in a large fraction of galaxies. We are also the first to provide a
+physical explanation of the behavior with clear observational predictions. We
+show synthetic \textit{Hubble Space Telescope} images of our barred galaxy in a
+case with and without gas in Fig.~\ref{fig:overview}. We see that the bar grows
+longer and stronger without gas, while it remains at approximately the same
+length and strength when gas is included.
 
-In section 2, \dots
+% In this work, we present a simulation of an isolated, gas-rich disk galaxy. Our
+% galaxy exhibits some similarities to the Milky Way, described in more detail in
+% Section XX. Surprisingly, we find that the pattern speed of the bar in this
+% galaxy is constant with time. We find that with even modest gas fractions of
+% about 5\%, our isolated galaxy displays no evolution in pattern speed over many
+% $\textrm{Gyr}$.
+
+In Section~\ref{sec:methods}, we describe our initial setup, numerical model,
+and details on our bar analysis procedures. In Section~\ref{sec:results}, we
+summarize the main results from our findings. We discuss these findings at more
+length and in the context of previous research in Section~\ref{sec:discussion}
+before concluding in Section~\ref{sec:conclusions}.
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=18cm]{fig/mockHST.pdf}
+    \caption{Synthetic Hubble Space Telescope images of our simulations with and
+    without the interstellar medium. The upper panels show an \Nbody{} only
+    simulation while the lower panels show a simulation which includes the
+    SMUGGLE model for the interstellar medium. Each panel is $20\,\textrm{kpc}$
+    to a side. Columns show different points in time, separated by
+    $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
+    length and strength. In the SMUGGLE run, the bar remains at approximately
+    the same length and strength over the course of the simulation. The \Nbody{}
+    model is identical to the GALAKOS model, discussed in the text. Details on
+    the production of these synthetic images is given in Appendix~\ref{app:hst}.}\label{fig:overview}
+\end{figure*}
 
@@
 \section{Methods}
+\label{sec:methods}
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=9cm]{fig/vcirc.pdf}
-\caption{The circular velocity curve of our initial setups. This
-curve is shown for the \Nbody{} run (blue) and the SMUGGLE run (orange)
-compared to observational estimates for the Milky
-Way \citep{2019ApJ...871..120E}. We see that the circular velocity curve for
-both runs is marginally larger than the Milky Way's, but still comparable. The
-SMUGGLE circular velocity curve is larger than the \Nbody{} curve due to the
-additional mass in the gas phase.}
-\label{fig:vcirc}
-\end{figure}
-
 We used a mass resolution of $7.5\times10^3\,M_{\odot}$ for the baryonic
@@
 $\textrm{kpc}/(\textrm{km}/\textrm{s})$.
+
+Our setup is initially out of equilibrium, but we found that after about
+$500\,\textrm{Myr}$, the system has settled into a steady-state configuration
+and initial transients appear not to affect the results after this point. The
+constant surface density of the initial gas disk is important for ensuring the
+gas disk is dense enough in order for comparisons to real galaxies to be
+appropriate.
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=9cm]{fig/A2_over_A0.pdf}
-\caption{\textbf{Evolution of bar strength.} The bar strength is measured as
-the maximum of the second Fourier component divided by the zeroth Fourier
-component. We see that in the
-\Nbody{} case (blue) the bar strength increases with time, consistent with
-previous results showing that the bar strength increases as bars slow down. In
-the SMUGGLE case (orange), we see that the bar strength remains roughly
-constant, possibly slightly decreasing with time. This is also consistent with
-the expected relation between pattern speed and strength since the bar in this
-case is not slowing down.}
-\label{fig:strength}
-\end{figure}
+In order to compute other properties of the bar, it is necessary to decompose
+the disk into a barred and unbarred component. We achieved this by following
+closely the methods described in \citet{2016MNRAS.463.1952P}. Our implementation
+is described in more detail in Appendix~\ref{app:bardecomp}. After the disk has
+been decomposed into a trapped and untrapped component, we measured the bar
+length as being the radius $R_b$ which encapsulates $99\%$ of the stars
+identified as being trapped in the bar, allowing for some outliers.
 
-\begin{figure*}
-\centering
-\includegraphics[width=18cm]{fig/bar_decomp.pdf}
-\caption{\textbf{Disk decomposition into the barred and unbarred disk.} This
-procedure is based on \citet{2016MNRAS.463.1952P} The \textit{left panel} shows
-a face-on surface density projection through the stellar component of the
-\SMUGGLE{} simulation (disk and bulge) at $t=1\,\textrm{Gyr}$. The
-\textit{middle panel} shows the component of the disk identified as being
-trapped in the bar while the \textit{right panel} shows the component of the
-disk identified as not being trapped in the bar. The fact that the untrapped
-stars form a roughly axisymmetric structure indicates our bar decomposition is
-sufficiently accurate.}
-\label{fig:decomp}
-\end{figure*}
-
-The time evolution of the bar strength, defined as the maximum of
-$\left|A_2/A_0\right|$ as a function of radius, is shown in Extended Data
-Fig.~\ref{fig:strength}. The quantity $\left|A_2/A_0\right|$ varies from $0$
-to $1$, with larger values indicating a stronger bar pattern. We see that in
-the \Nbody{} case, $\left|A_2/A_0\right|$ increases over time as the bar pattern
-slows. This is consistent with previous \Nbody{} simulations which showed a
-clear correlation between the bar pattern speed and the bar
-strength\cite{2003MNRAS.341.1179A}. In the SMUGGLE case, we see that the bar
-strength has an initial drop but then remains at a roughly constant, but
-slightly decreasing, strength. This is consistent with the pattern speed in the
-SMUGGLE case being roughly constant or slightly increasing.
-
-Computing the length of the bar and the torque on the bar by different
-components requires us to decompose the disk into a component which is trapped
-by the bar and a component which is untrapped. In order to do this, we follow
-closely the technique developed in ref.\cite{2016MNRAS.463.1952P} We analyzed
-the orbit of each star particle (meaning initial disk, bulge, and newly formed
-stars) by extracting the $x$-$y$ positions of the apoapse of each in a frame
-corotating with the bar (apoapses are defined as local maxima in $r$). For
-each apoapse, we searched for the $19$ closest apoapses in time and applied a
-$k$-means clustering algorithm on this set of $20$ points with $k=2$. We then
-computed for each of the two clusters the average angle from the bar
-$\left<\Delta \phi\right>_{0,1}$, the standard deviation in $R$ of the points
-${\sigma_R}_{0,1}$, and the average radius of the cluster
-$\left<R\right>_{0,1}$. At each apoapse, a particle was considered to be in
-the bar if it met the following criteria:
-\begin{equation}
-\textrm{max}\left(\left<\Delta \phi\right>_{0,1}\right) < \pi / 8
-\end{equation}
-\begin{equation}
-\frac{{\sigma_R}_0 + {\sigma_R}_1}{\left<R\right>_0 + \left<R\right>_1} < 0.22
-\end{equation}
-These criterion are slightly different and simplified from the ones used in
-ref.\cite{2016MNRAS.463.1952P}, but we found them to empirically work well at
-decomposing the disk into a bar and disk component. In Extended Data
-Fig.~\ref{fig:decomp}, we show an example of this decomposition. The
-\textit{left} panel shows a surface density projection of the stellar disk and
-bulge (including newly formed stars) from the SMUGGLE model after
-$1\,\text{Gyr}$ of evolution in a frame such that the bar is aligned with the
-$x$-axis. The \textit{middle} panel shows a projection of the subset of stars
-that are identified as being trapped in the bar and the \textit{right} panel
-shows a projection of the stars that are not identified as being trapped. The
-fact that the \textit{right} panel is roughly axisymmetric indicates the bar
-decomposition is performing adequately.
-
-After the disk has been decomposed into a trapped and untrapped component, we
-measured the bar length as being the radius $R_b$ which encapsulates $99\%$ of
-the stars identified as being trapped in the bar, allowing for some outliers.
-For the computation of the torque, we used the tree algorithm in
-\texttt{MakeNewDisk}\cite{2005MNRAS.361..776S} customized to be accessible
-from \texttt{Python} using \texttt{Cython}. This algorithm is based on the
-\texttt{TREESPH} code\cite{1989ApJS...70..419H}. We constructed a tree with an opening angle of $0.35$ using
-only the star particles identified as being trapped in the bar. We then queried the tree at the locations of all
-resolution elements in the other components and computed the torque of the bar
-on such components. The torque on the bar by the other components is simply
-the negative of the torque on the other components by the bar. A similar
-method was applied in measuring the torque for the disk when the pattern speed
-is kept constant, as is done in Fig.~\ref{fig:equil}.
-
-% \subsection{Idealized Mock Light Images}
-% We created synthetic mock light images at $1024^2$ resolution in
-% Fig.~\ref{fig:overview} for the \textit{Hubble Space Telescope} using a
-% post-processed Monte Carlo radiation transfer simulation with the SKIRT9 code
-% including secondary emission from dust\cite{2020AC....3100381C}. Emission from
-% star particles was calculated assuming a Bruzual-Charlot spectral energy
-% distribution\cite{2003MNRAS.344.1000B} and a Chabrier initial stellar mass
-% function\cite{2003PASP..115..763C}. The luminosity from each star particle was
-% smoothed using a cubic spline kernel with each star particle being assigned a
-% smoothing length equal to its distance to its $32$nd closest neighbor. We
-% launched $10^9$ photon packets per simulation segment. For star particles
-% which were present in the initial simulation (i.e., initial disk and bulge
-% particles), we assumed that they have an age of $5\,\textrm{Gyr}$ and solar
-% metallicity. For star particles in the SMUGGLE simulation which are newly
-% formed, we used their recorded ages and metallicities.
-
-% For the SMUGGLE simulations, we included absorption and emission from dust. We
-% assumed the THEMIS (The Heterogeneous dust Evolution Model for Interstellar
-% Solids) dust model\cite{2017AA...602A..46J} to convert gas metallicity and
-% mass density to dust mass density. For the \Nbody{} simulations we ran the
-% SKIRT9 code in a mode without an obscuring medium, but otherwise applied the
-% same procedure.
-
-% We used the profiles for filters F814W, F606W, and F475W on the Advanced
-% Camera for Surveys (ACS) instrument for red, green, and blue pixel values,
-% respectively. Filtered fluxes were converted to RGB values using an $\arcsinh$
-% scaling following ref.\cite{2004PASP..116..133L} Scalings and cutoff values
-% were adjusted manually to create visually appealing images. Filter profiles
-% were provided by the Spanish Virtual Observatory Filter
-% Service\cite{2012ivoa.rept.1015R, 2020sea..confE.182R}.
+To compute torques we used the tree algorithm in \texttt{MakeNewDisk}
+\citep{2005MNRAS.361..776S} customized to be accessible from \texttt{Python}
+using \texttt{Cython}. This algorithm is based on the \texttt{TREESPH} code
+\citep{1989ApJS...70..419H}. We constructed a tree with an opening angle of
+$0.35$ using only the star particles identified as being trapped in the bar. We
+then queried the tree at the locations of all resolution elements in the other
+components and computed the torque of the bar on such components. The torque on
+the bar by the other components is simply the negative of the torque on the
+other components by the bar. A similar method was applied in measuring the
+torque for the disk when the pattern speed is kept constant, as is done in
+Fig.~\ref{fig:equil}.
 
@@
 
-\section{Discussion}
-Our setup is initially out of equilibrium, but we found that after
-$\sim500\,\textrm{Myr}$, the entire system has settled into a steady-state
-configuration and initial transients appear not to affect the results after
-this point. The constant surface density of the initial gas disk is important
-for ensuring the gas disk is dense enough in order for comparisons to real
-galaxies to be appropriate.
-
-We computed the circular velocity curve of our model using the \texttt{AGAMA}
-package\cite{2019MNRAS.482.1525V}. We fit the baryonic component (stellar
-disk, bulge, gas, and newly formed stars) with an axisymmetric cylindrical
-spline with $20$ grid points in both the radial and vertical direction
-spanning $0.2$ to $50\,\textrm{kpc}$ in the radial direction and from $0.02$
-to $10\,\textrm{kpc}$ in the vertical direction. We fit the dark matter halo
-using a spherically symmetric multipole fit with a maximum angular harmonic
-coefficient of $l=2$. We plot the circular velocity curve in Extended Data
-Fig.~\ref{fig:vcirc} compared to observational
-estimates\cite{2019ApJ...871..120E}. The SMUGGLE disk (which includes
-additional mass in the form of gas) has a slightly higher circular velocity
-than the \Nbody{} disk which, itself, is slightly higher than the
-observational estimates.
+\section{Results}
+\label{sec:results}
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{fig/ps_Rbar_torque.pdf}
+    \caption{The \textit{upper panel} shows the evolution
+    of the pattern speed. As expected, the bar in the \Nbody{} run slows down due
+    to interactions between the bar and the dark matter halo. However, the bar in
+    the SMUGGLE run does not slow down and instead remains at a constant pattern
+    speed. The \textit{middle panel} shows the evolution of the bar length. In the
+    \Nbody{} case, the bar lengthens. This occurs because as the pattern speed
+    drops, bar-like orbits at larger radii are possible. Stars are captured on
+    these orbits, lengthening the bar. This process does not occur in the SMUGGLE
+    cases since the bar pattern speed is not decreasing, and therefore the bar
+    length remains constant. The \textit{lower panel} shows the torque on the bar
+    by different components. The solid lines show the torque exerted by the halo
+    in both the \Nbody{} and SMUGGLE cases. The dashed line shows the torque
+    exerted by the gas phase in the SMUGGLE run (there is no gas in the \Nbody{}
+    run). Details on how these properties are calculated is given in the Methods
+    section.}\label{fig:prop}
+\end{figure}
 
@@
     \centering
-    \includegraphics[width=9cm]{fig/surf_dens.pdf}
-    \caption{The time evolution of the atomic gas surface density
-    (\textit{upper}), molecular gas surface density (\textit{middle}) and the
-    star formation rate (SFR) surface density (\textit{lower}) at various times
-    during our fiducial simulation. Colored lines indicate the profiles at
-    selected times during the simulation while the black dashed lines indicate
-    observations for the atomic gas \citep{2008AA...487..951K} and black dotted
-    lines indicate a model which allows the CO-to-H$_2$ conversion factor
-    $X_{\textrm{CO}}$ to vary with metallicity \citep{2022ApJ...929L..18E}.
-    Molecular gas surface densities were provided separately (N. Evans, private
-    communication). We see that the molecular gas and SFR surface densities are
-    within an order of magnitude of the Milky Way's typical values at all times.
-    We see a sharp decrease in the gas and SFR surface densities along the
-    extent of the bar from $\sim1$ to $\sim4\,\textrm{kpc}$, related to the gas
-    inflow in this region.}
-    \label{fig:surf}
+    \includegraphics[width=9cm]{fig/A2_over_A0.pdf}
+    \caption{The evolution in bar strength. The bar strength is measured as the
+    maximum of the second Fourier component divided by the zeroth Fourier
+    component. We see that in the \Nbody{} case (blue) the bar strength
+    increases with time, consistent with previous results showing that the bar
+    strength increases as bars slow down. In the SMUGGLE case (orange), we see
+    that the bar strength remains roughly constant, possibly slightly decreasing
+    with time. This is also consistent with the expected relation between
+    pattern speed and strength since the bar in this case is not slowing down.}
+    \label{fig:strength}
 \end{figure}
 
-We also show the evolution of the surface density profile in Extended Data
-Fig.~\ref{fig:surf} We find that in our simulation the atomic and molecular
-gas surface density and the SFR surface density is broadly consistent with the
-expected values for the Milky
-Way\cite{2008AA...487..951K,2022ApJ...929L..18E}. The discrepancy between $1$
-and $4\,\textrm{kpc}$ in the molecular and SFR surface density is likely due
-to the fact that the distances to molecular clouds which underlines this work
-used a simple kinematic distance based on an axisymmetric model of the Milky
-Way\cite{2017ApJ...834...57M}, which is not accurate in the bar region where
-gas has large non-circular velocities.
+We show the time evolution of different bar properties in Fig.~\ref{fig:prop}.
+In the upper panel, we show the pattern speed over time in the \Nbody{} (blue)
+and SMUGGLE (orange) runs. The pattern speed in the \Nbody{} case slows down
+while the pattern speed in the SMUGGLE case remains roughly constant. The
+slowing down of the pattern speed in the \Nbody{} case is consistent with a long
+line of numerical research on bars in \Nbody{} simulations
+\citep{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
+2002ApJ...569L..83A, 2003MNRAS.341.1179A, 2003MNRAS.346..251O,
+2005MNRAS.363..991H, 2006ApJ...637..214M, 2007MNRAS.375..460W,
+2009ApJ...697..293D}.
 
-The pattern speed evolution of this setup is given in Extended Data
-Fig.~\ref{fig:GFM}, which shows that the pattern speed evolves in a similar
-manner to the SMUGGLE case shown in Fig.~\ref{fig:prop}.
+However, in the SMUGGLE case the pattern speed remains constant. After the first
+Gyr of evolution, we find that the pattern speed increases by only $\sim10\%$
+over the next $4\,\textrm{Gyr}$, compared to a $\sim43\%$ decrease in the
+pattern speed for the \Nbody{} run over the same interval. As we saw
+qualitatively in Fig.~\ref{fig:overview}, the middle panel of
+Fig.~\ref{fig:prop} shows that the length of the bar in the \Nbody{} case grows
+over time while it remains roughly constant in the SMUGGLE case. This is also
+consistent with previous numerical work, which found that bars tend to grow as
+they slow down and the radius of corotation increases
+\citep{2000ApJ...543..704D, 2003MNRAS.341.1179A}.
+
+The bottom panel of Fig.~\ref{fig:prop} shows the torque exerted on the bar by
+different components. The solid lines indicate the torque on the bar by the dark
+matter halo, whereas the dashed line indicates the torque on the bar by the gas
+phase. In the \Nbody{} case, the halo exerts a steady negative torque on the
+bar, with an average torque from $1$ to $4\,\textrm{Gyr}$ of $-58.0$ in units of
+$10^{10}M_{\odot}\,(\textrm{km}/\textrm{s})^2$. The halo in the SMUGGLE case
+exerts a similar negative torque on the bar in the first Gyr of evolution, but
+after that the halo exerts a much smaller torque on the bar, averaging only
+$-7.8$ in the same units and over the same time interval. The gas in the SMUGGLE
+case exerts a steady positive torque averaging $11.7$ over $1\,\textrm{Gyr}$ in
+the same units.
+
+The time evolution of the bar strength, defined as the maximum of
+$\left|A_2/A_0\right|$ as a function of radius, is shown in
+Fig.~\ref{fig:strength}. The quantity $\left|A_2/A_0\right|$ varies from $0$ to
+$1$, with larger values indicating a stronger bar pattern. We see that in the
+\Nbody{} case, $\left|A_2/A_0\right|$ increases over time as the bar pattern
+slows. This is consistent with previous \Nbody{} simulations which showed a
+clear correlation between the bar pattern speed and the bar strength
+\citep[e.g.,][]{2003MNRAS.341.1179A}. In the SMUGGLE case, we see that the bar
+strength has an initial drop but then remains at a roughly constant, but
+slightly decreasing, strength. This is consistent with the pattern speed in the
+SMUGGLE case being roughly constant or slightly increasing.
 
@@
 
+\section{Discussion}
+\label{sec:discussion}
+\subsection{Pattern Speed Evolution}
+The lack of evolution in the pattern speed of the SMUGGLE case is intimately
+tied to the sudden decrease in torque exerted on the bar by the dark matter
+halo. We argue that this can be understood in terms of the halo wake mechanism.
+In the \Nbody{} case, halo material which is resonant with the bar will form a
+wake which lags behind and exerts a negative torque on the bar, slowing it down
+\citep{1984MNRAS.209..729T, 1985MNRAS.213..451W,
+1992ApJ...400...80H}.\footnote{Since the bar is not a solid body, it is not
+guaranteed that a negative torque will slow it down - e.g. a negative torque
+could shred the bar, reducing its moment of inertia without changing its pattern
+speed. However, the bar seems to empirically respond to a negative torque
+induced by a halo wake by slowing down.} As the bar slows down, the location of
+the resonances in phase space changes, allowing halo material newly resonant
+with the bar to participate in forming the wake. However, the gas is a reliable
+source of positive torque on the bar, speeding the bar up. This stops the
+resonance location from changing such that the halo cannot reinforce the wake,
+arresting the process by which the halo can slow the bar down. We term this
+process ``exhaustion.''
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=18cm]{fig/halo_wake.pdf}
+    \caption{The wake excited in the dark matter halo. The dark matter halo wake
+    is shown in the \Nbody{} case (\textit{middle panel}) and SMUGGLE case
+    (\textit{right panel}) after $2.6\,\textrm{Gyr}$ of evolution. The
+    \textit{middle} and \textit{right panels} show a surface density projection
+    in the $x$-$y$ plane of the dark matter halo after an axisymmetric average
+    has been subtracted. The solid line indicates the direction of the bar while
+    the dashed line indicates the direction of the halo wake (both measured by
+    taking the second Fourier component within a sphere of all material within a
+    radius of $4\,\textrm{kpc}$). The \textit{left panel} shows the time
+    evolution of the angle difference between the bar and the halo wake, as
+    measured from the second Fourier component. After the first Gyr, the angle
+    difference in the SMUGGLE case is smaller than in the \Nbody{} case by about
+    a factor of two, reflecting how the dark matter halo in the SMUGGLE case is
+    unable to exert as negative a torque on the bar as in the \Nbody{} case.}
+    \label{fig:wake}
+\end{figure*}
+
+We can test this interpretation by measuring the angle offset between the halo
+wake and the bar. If the wake and the bar are aligned (i.e., there is no angle
+offset), then the wake cannot exert a negative torque on the bar. This angle
+is plotted in the left panel of Fig.~\ref{fig:wake}, which shows that the
+angle offset is larger in the \Nbody{} case than in the SMUGGLE case by about a
+factor of two. The center and right panels of Fig.~\ref{fig:wake} show the halo
+wake with respect to the location of the bar in the \Nbody{} (center) and SMUGGLE
+(right) cases at one point in time.
+
+The presence of the gas can arrest the process by which additional material in
+the dark matter halo can contribute to a wake. However, this does not explain
+why the pattern speed in the SMUGGLE case is nearly constant over several Gyr.
+Naively, it would be a coincidence that the bar pattern speed remains constant
+in the SMUGGLE case, resulting from a chance cancellation of the halo and gas
+torques. However, a constant pattern speed in the presence of gas has been
+observed in a few simulations of barred galaxies with
+gas \citep{1993AA...268...65F, 2010ApJ...719.1470V, 2014MNRAS.438L..81A}. Previous
+work has argued this is due to the bar torquing gas inwards, but no explanation
+has been given for why it might remain constant.
+
+We propose that an equilibrium mechanism is responsible for the pattern speed
+remaining approximately constant. In this scenario, residual negative torque
+from the dark matter halo balances out the positive torque from the gas phase.
+It has been shown when an analytic bar is forced to rotate at a constant pattern
+speed for a few Gyr, the halo exerts almost no torque on the bar
+\citep{2022MNRAS.513..768C}. We saw in Fig.~\ref{fig:prop} that the dark matter
+halo in our simulation is still able to support some negative torque over a
+several Gyr time span.
+
+We argue that the following occurs. First, the bar is not able to slow down
+quickly enough due to the positive torque of the infalling gas. This causes
+the resonant halo phase space at a particular pattern speed, $\Omega_{p,0}$,
+to become exhausted and no longer able to support a negative torque. Second,
+the gas is still exerting a positive torque on the bar, and therefore the
+pattern speed will again increase. Since at higher pattern speeds the halo has
+not yet been totally exhausted, the halo will once again be able to exert a
+negative torque on the bar. The pattern speed will then settle at a new value
+slightly higher than $\Omega_{p,0}$ where the gas and halo torques cancel.
+Over time, the pattern speed should slowly increase as the halo becomes
+progressively exhausted.
+
+\subsection{Delayed Gas Injection}
+A clear prediction of our proposed mechanism is that the constant pattern
+speed a particular galaxy will end up is somewhat arbitrary. In the real
+universe for an isolated galaxy it would be the formation pattern speed of the
+bar while in our simulation it is the pattern speed of the bar when gas is
+added to the system. We tested this by adding gas to the system at a later
+time when the bar has further grown and slowed down with time. In our
+particular test, we added the gas at a time when the pattern speed is
+$\sim30\,\textrm{km}/\textrm{s}/\textrm{kpc}$. As shown in
+Fig.~\ref{fig:snap700}, we find that the pattern speed evolution is very
+similar between the two cases (orange and red lines). If anything, the system
+with a lower pattern speed seems to speed up more, which is consistent with
+our picture since the stronger bar should experience a larger torque from the
+gas as it is more efficient at driving gas inflows. We also show in the
+Methods that more slowly rotating bars at fixed bar strength are more
+efficient at driving gas inflows as well. Nonetheless, when the initial pattern
+speed is lower (red line), the addition of gas does not cause the pattern
+speed to quickly return to the higher value of our fiducial simulation (orange
+line).
+
+\begin{figure}[t!]%
+    \centering
+    \includegraphics[width=9cm]{fig/ps_late_start.pdf}
+    \caption{Pattern speed evolution with a lower initial pattern speed. We
+    tested the evolution of our system when gas is added to the \Nbody{} run at
+    a later time, but with all other simulation parameters kept the same. The
+    setup is therefore identical to our previous runs just with a bar that is
+    larger, stronger, and with a lower pattern speed. We find that the pattern
+    speed evolution is very similar to our fiducial case, except that the bar
+    retains its original pattern speed. This indicates a mechanism which keeps
+    the bar at its formation pattern speed, and that there is not a particular
+    pattern speed which the system tends to.}
+    \label{fig:snap700}
+\end{figure}
+
+\subsection{Varying Initial Gas Fractions}
+We also performed a test in which we varied the initial gas fraction of the
+disk. In our fiducial run, we set the surface density of the gas disk from
+$4\,\textrm{kpc}$ to $\sim9.3\,\textrm{kpc}$ to be $20\,\Msun/\textrm{pc}^2$.
+We also ran with surface densities of $15$, $10$, and
+$5\,\Msun/\textrm{pc}^2$. These correspond to initial gas fractions of
+approximately $16\%$, $10\%$, $7\%$, and $4\%$. The pattern speed evolution is
+shown in Extended Data Fig.~\ref{fig:fgas} We find that the bar in disks with
+initial surface densities of $20$, $15$, and $10\,\Msun/\textrm{pc}^2$ evolve
+with a constant pattern speed while the bar in a disk with initial surface
+density of $5\,\Msun/\textrm{kpc}^2$ slows down at a similar rate to the \Nbody{} case.
+
+As a result, we conclude that for the disk, bar, and halo properties
+considered in this work, a gas fraction of only approximately $5\%$ is
+necessary in order for the proposed stabilizing mechanism to operate.
+
+\begin{figure}[t!]
+    \centering
+    \includegraphics[width=9cm]{fig/ps_fgas.pdf}
+    \caption{Pattern speed evolution with varying gas fractions. We explored the
+    impact of lowering the initial gas surface density of our fiducial disk on
+    the evolution of the pattern speed. The surface densities we tested of $20$,
+    $15$, $10$, and $5\,\Msun/\textrm{kpc}^2$ correspond to initial gas
+    fractions of $16\%$, $10\%$, $7\%$, and $4\%$. We find that initial surface
+    densities $20$, $15$, and $10\,\Msun/\textrm{kpc}^2$ result in bars which
+    remain at a constant pattern speed, while an initial surface density of
+    $5\,\Msun/\textrm{kpc}^2$ results in a bar which slows down in roughly the
+    same manner as the \Nbody{} case.}
+    \label{fig:fgas}
+\end{figure}
+
+\subsection{Smooth Interstellar Medium}
+We also performed a simulation of the same disk but with a simpler model of the
+interstellar medium \citep{2003MNRAS.339..289S}, closer to standard methods used
+in cosmological simulations of galaxy formation. We find that the pattern speed
+evolution is nearly the same in this case, and so conclude that our result is
+not sensitive to the details of the model for the interstellar medium. Details
+on the physics of this test is given in Section.~X.
+
+\begin{figure}[h!]%
+    \centering
+    \includegraphics[width=9cm]{fig/ps_GFM.pdf}
+    \caption{Pattern speed evolution of a smooth ISM model. This evolution is
+    shown for the fiducial disk in the \Nbody{} (blue), SMUGGLE (orange), and
+    smooth ISM (red) cases. The smooth ISM model is an older model for the ISM
+    which treats its multiphase nature in a subgrid fashion
+    \citep{2003MNRAS.339..289S}. This fundamentally differs from the SMUGGLE
+    model, which explicitly resolves the hot and cold phases of the ISM
+    \citep{2019MNRAS.489.4233M}. The pattern speed in the smooth ISM case is
+    broadly similar to the evolution in the SMUGGLE case. This shows that the
+    stability of the pattern speed is not simply a result of our assumed model
+    for the ISM.}
+\label{fig:GFM}
+\end{figure}
+
+\subsection{Observations}
+We note that ref.\cite{2020MNRAS.491.3655G} found that the rotation parameter
+\Rot{} positively correlates with gas fraction, such that galaxies with higher
+gas fractions are rotating more slowly. However, it is not obvious this is in
+tension with our result since the gas fraction of galaxies correlates with
+other galactic properties\cite{2009ARAA..47..159B}. Furthermore, the
+measurement of pattern speeds is a delicate process still prone to large error
+bars.
+
+\subsection{Cosmological Simulations}
+Barred galaxies in cosmological simulations of galaxy formation continue to be
+in conflict with observations by producing bars which rotate too
+slowly\cite{2017MNRAS.469.1054A, 2019MNRAS.483.2721P, 2021AA...650L..16F}.
+However, the pattern speeds of bars in both cosmological simulations and the
+real universe can be affected by environmental processes not included in our
+simulation -- e.g., satellite infall\cite{2011Natur.477..301P},
+non-sphericity\cite{2013MNRAS.429.1949A} or rotation\cite{2013MNRAS.434.1287S,
+2014ApJ...783L..18L, 2018MNRAS.476.1331C, 2019MNRAS.488.5788C} in the dark
+matter halo, or perhaps even the gaseous circumgalactic medium. Naturally,
+extending our present work to account for such effects is a crucial next step
+in understanding the formation and evolution of galactic bars.
+
 \section{Conclusions}
+\label{sec:conclusions}
+The implications of our findings are numerous. First, we naturally explain why
+nearly all observed galaxies are fast rotators without requiring the inner
+regions of dark matter halos to be underdense\cite{1998ApJ...493L...5D,
+2000ApJ...543..704D} or requiring new physics\cite{2021MNRAS.503.2833R,
+2021MNRAS.508..926R}. Second, we show that the role of gas is of paramount
+importance in studies which attempt to uncover the nature of dark matter from
+its effect of slowing down the bar\cite{2021MNRAS.500.4710C,
+2021MNRAS.505.2412C}. Third, we provide an explanation for how the Milky Way's
+bar could be both long-lived and a fast rotator, of which there is some
+observational evidence\cite{2019MNRAS.490.4740B}. And finally, we complicate
+the picture of radial mixing expected to sculpt the Milky Way's disk
+\cite{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies upon
+the pattern speed of the bar to change with time (though our work does not
+alter expectations for radial mixing induced by spiral
+arms\cite{2002MNRAS.336..785S}).
 
-The last numbered section should briefly summarise what has been done, and describe
-the final conclusions which the authors draw from their work.
+We found that below a certain gas fraction, bars should still be able to slow
+down. Therefore, we expect barred spiral galaxies which have been gas-poor for
+extended periods of time to be rotating very slowly. We therefore predict that
+observations which target such galaxies (e.g., lenticular barred
+galaxies\cite{2009ARAA..47..159B}) would find slowly rotating bars. There does
+exist one example of a galaxy which is known to be a slow rotator -- the
+low surface brightness galaxy UGC 628\cite{2009AA...499L..25C}. This galaxy
+has been studied in detail by ref.\cite{2016MNRAS.463.1751C}, who note that it
+indeed has a low gas fraction for galaxies of its type. We predict a general
+trend that bars in gas-rich spiral galaxies should rotate quickly while bars
+in gas-poor spiral galaxies should rotate slowly.
 
 \section*{Acknowledgements}
-
-The Acknowledgements section is not numbered. Here you can thank helpful
-colleagues, acknowledge funding agencies, telescopes and facilities used etc.
-Try to keep it short.
+We would like to thank Greg~L. Bryan, Neal~J. Evans, Drummond~B. Fielding, Keith
+Hawkins, Jason~A.~S. Hunt, Sarah~M.~R. Jeffreson, Kathryn~V. Johnston,
+Peter~M.~W. Kalberla, Jürgen Kerp, Julio~F. Navarro, Joshua~S. Speagle,
+Martin~D. Weinberg, and Yanfei Zou for helpful discussions. A.B. would like to
+thank Todd Phillips for helpful discussions. Some computations in this paper
+were run on the FASRC Cannon cluster supported by the FAS Division of Science
+Research Computing Group at Harvard University. Resources supporting this work
+were also provided by the NASA High-End Computing (HEC) Program through the NASA
+Advanced Supercomputing (NAS) Division at Ames Research Center. This research
+has made use of the Spanish Virtual Observatory (https://svo.cab.inta-csic.es)
+project funded by MCIN/AEI/10.13039/501100011033/ through grant
+PID2020-112949GB-I00. A.B. was supported by the Future Investigators in NASA
+Earth and Space Science and Technology (FINESST) award number 80NSSC20K1536
+during the completion of this work. E.D. was partially supported by HST grants
+HST-AR-16363.001 and HST-AR-16602.006-A and by NASA Award NASA 80NSSC22K0761.
+J.Q. acknowledges support from NSF grant AST-2008490. L.V.S. is grateful for
+financial support from NASA ATP 80NSSC20K0566, NSF AST 1817233 and NSF CAREER
+1945310 grants. P.T. acknowledges support from NSF grant AST-1909933,
+AST-2008490, and NASA ATP Grant 80NSSC20K0502. M.V. acknowledges support through
+NASA ATP 19-ATP19-0019, 19-ATP19-0020, 19-ATP19-0167, and NSF grants
+AST-1814053, AST-1814259, AST-1909831, AST-2007355 and AST-2107724.
 
@@
 
-\section{Some extra material}
+\section{Bar Decomposition}
+\label{app:bardecomp}
+\begin{figure*}
+    \centering
+    \includegraphics[width=18cm]{fig/bar_decomp.pdf}
+    \caption{Disk decomposition into the barred and unbarred disk. This
+    procedure is based on \citet{2016MNRAS.463.1952P}. The \textit{left panel}
+    shows a face-on surface density projection through the stellar component of
+    the \SMUGGLE{} simulation (disk and bulge) at $t=1\,\textrm{Gyr}$. The
+    \textit{middle panel} shows the component of the disk identified as being
+    trapped in the bar while the \textit{right panel} shows the component of the
+    disk identified as not being trapped in the bar. The fact that the untrapped
+    stars form a roughly axisymmetric structure indicates our bar decomposition
+    is sufficiently accurate.}
+    \label{fig:decomp}
+\end{figure*}
 
-If you want to present additional material which would interrupt the flow of the main paper,
-it can be placed in an Appendix which appears after the list of references.
+Computing the length of the bar and the torque on the bar by different
+components requires us to decompose the disk into a component which is trapped
+by the bar and a component which is untrapped. In order to do this, we follow
+closely the technique developed in \citet{2016MNRAS.463.1952P}. We analyzed the
+orbit of each star particle (meaning initial disk, bulge, and newly formed
+stars) by extracting the $x$-$y$ positions of the apoapse of each in a frame
+corotating with the bar, where apoapses are defined as local maxima in $r$. For
+each apoapse, we searched for the $19$ closest apoapses in time and applied a
+$k$-means clustering algorithm on this set of $20$ points with $k=2$. We then
+computed for each of the two clusters the average angle from the bar
+$\left<\Delta \phi\right>_{0,1}$, the standard deviation in $R$ of the points
+${\sigma_R}_{0,1}$, and the average radius of the cluster
+$\left<R\right>_{0,1}$. At each apoapse, a particle was considered to be in the
+bar if it met the following criteria:
+\begin{equation}
+\textrm{max}\left(\left<\Delta \phi\right>_{0,1}\right) < \pi / 8
+\end{equation}
+\begin{equation}
+\frac{{\sigma_R}_0 + {\sigma_R}_1}{\left<R\right>_0 + \left<R\right>_1} < 0.22
+\end{equation}
+These criterion are slightly different and simplified from the ones used in \citet{2016MNRAS.463.1952P}, but we found them to empirically work well at
+decomposing the disk into a bar and disk component. In Extended Data
+Fig.~\ref{fig:decomp}, we show an example of this decomposition. The
+\textit{left} panel shows a surface density projection of the stellar disk and
+bulge (including newly formed stars) from the SMUGGLE model after
+$1\,\text{Gyr}$ of evolution in a frame such that the bar is aligned with the
+$x$-axis. The \textit{middle} panel shows a projection of the subset of stars
+that are identified as being trapped in the bar and the \textit{right} panel
+shows a projection of the stars that are not identified as being trapped. The
+fact that the \textit{right} panel is roughly axisymmetric indicates the bar
+decomposition is performing adequately.
+
+\section{Varying Pattern Speed}
+\label{app:varyps}
+When the bar slows down, we argue that this induces a larger positive torque
+from the gas phase. Only gas within corotation will flow inwards, while gas
+outside corotation will flow outwards\cite{2011MNRAS.415.1027H}. Since the
+corotation radius is larger for more slowly rotating bars, it follows that
+more slowly rotating bars should be more efficient at driving gas inflows and
+thus experience a larger positive torque from the gas phase.
+
+We performed an experiment to test this hypothesis by forcing the stellar disk
+in the SMUGGLE run to rotate at a constant angular rate and measuring the
+torque on the bar by the gas phase at different rotation rates. The result of
+this experiment is illustrated in Extended Data Fig.~\ref{fig:equil}, which shows that a more
+slowly rotating bar experiences a larger positive torque from the gas.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=9cm]{fig/torque_ps.pdf}
+    \caption{Average torque exerted by gas on a bar which rotates at a fixed
+    pattern speed. Since only gas within the corotation radius is able to infall
+    and slower bars have larger corotation radii, slower bars experience a
+    larger net torque than faster bars. The setup of the simulations used here
+    is identical to the SMUGGLE case discussed earlier, except the \Nbody{} disk
+    is rotated as a solid body with a constant angular
+    velocity.}
+    \label{fig:equil}
+\end{figure}
+
+\section{Idealized Mock Light Images}
+\label{app:hst}
+We created synthetic mock light images at $1024^2$ resolution in
+Fig.~\ref{fig:overview} for the \textit{Hubble Space Telescope} using a
+post-processed Monte Carlo radiation transfer simulation with the SKIRT9 code
+including secondary emission from dust \citep{2020AC....3100381C}. Emission from
+star particles was calculated assuming a Bruzual-Charlot spectral energy
+distribution \citep{2003MNRAS.344.1000B} and a Chabrier initial stellar mass
+function \citep{2003PASP..115..763C}. The luminosity from each star particle was
+smoothed using a cubic spline kernel with each star particle being assigned a
+smoothing length equal to its distance to its $32$nd closest neighbor. We
+launched $10^9$ photon packets per simulation segment. For star particles
+which were present in the initial simulation (i.e., initial disk and bulge
+particles), we assumed that they have an age of $5\,\textrm{Gyr}$ and solar
+metallicity. For star particles in the SMUGGLE simulation which are newly
+formed, we used their recorded ages and metallicities.
+
+For the SMUGGLE simulations, we included absorption and emission from dust. We
+assumed the THEMIS (The Heterogeneous dust Evolution Model for Interstellar
+Solids) dust model \citep{2017AA...602A..46J} to convert gas metallicity and
+mass density to dust mass density. For the \Nbody{} simulations we ran the
+SKIRT9 code in a mode without an obscuring medium, but otherwise applied the
+same procedure.
+
+We used the profiles for filters F814W, F606W, and F475W on the Advanced
+Camera for Surveys (ACS) instrument for red, green, and blue pixel values,
+respectively. Filtered fluxes were converted to RGB values using an $\arcsinh$
+scaling following \citet{2004PASP..116..133L} Scalings and cutoff values
+were adjusted manually to create visually appealing images. Filter profiles
+were provided by the Spanish Virtual Observatory Filter
+Service \citep{2012ivoa.rept.1015R, 2020sea..confE.182R}.
+
+\section{Comparison to the Milky Way}
+\label{app:milkyway}
+Our fiducial disk model is a relatively good  model of the Milky Way for several
+Gyr. This is uncommon in models of galaxies that include the gas phase of the
+disk but no circumgalactic medium. As mentioned earlier, the pattern speed seems
+to match the observed pattern speed of the Milky Way's bar
+\citep{2019MNRAS.490.4740B}. We briefly summarize some of the other ways our disk is comparable to the Milky Way.
+
+We computed the circular velocity curve of our model using the \texttt{AGAMA}
+package\cite{2019MNRAS.482.1525V}. We fit the baryonic component (stellar
+disk, bulge, gas, and newly formed stars) with an axisymmetric cylindrical
+spline with $20$ grid points in both the radial and vertical direction
+spanning $0.2$ to $50\,\textrm{kpc}$ in the radial direction and from $0.02$
+to $10\,\textrm{kpc}$ in the vertical direction. We fit the dark matter halo
+using a spherically symmetric multipole fit with a maximum angular harmonic
+coefficient of $l=2$. We plot the circular velocity curve in Extended Data
+Fig.~\ref{fig:vcirc} compared to observational
+estimates\cite{2019ApJ...871..120E}. The SMUGGLE disk (which includes
+additional mass in the form of gas) has a slightly higher circular velocity
+than the \Nbody{} disk which, itself, is slightly higher than the
+observational estimates.
+
+We also show the evolution of the surface density profile in Fig.~\ref{fig:surf}
+We find that in our simulation the atomic and molecular gas surface density and
+the SFR surface density is broadly consistent with the expected values for the
+Milky Way\cite{2008AA...487..951K,2022ApJ...929L..18E}. The discrepancy between
+$1$ and $4\,\textrm{kpc}$ in the molecular and SFR surface density is likely due
+to the fact that the distances to molecular clouds which underlines this work
+used a simple kinematic distance based on an axisymmetric model of the Milky
+Way\cite{2017ApJ...834...57M}, which is not accurate in the bar region where gas
+has large non-circular velocities.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=9cm]{fig/surf_dens.pdf}
+    \caption{The time evolution of the atomic gas surface density
+    (\textit{upper}), molecular gas surface density (\textit{middle}) and the
+    star formation rate (SFR) surface density (\textit{lower}) at various times
+    during our fiducial simulation. Colored lines indicate the profiles at
+    selected times during the simulation while the black dashed lines indicate
+    observations for the atomic gas \citep{2008AA...487..951K} and black dotted
+    lines indicate a model which allows the CO-to-H$_2$ conversion factor
+    $X_{\textrm{CO}}$ to vary with metallicity \citep{2022ApJ...929L..18E}.
+    Molecular gas surface densities were provided separately (N. Evans, private
+    communication). We see that the molecular gas and SFR surface densities are
+    within an order of magnitude of the Milky Way's typical values at all times.
+    We see a sharp decrease in the gas and SFR surface densities along the
+    extent of the bar from $\sim1$ to $\sim4\,\textrm{kpc}$, related to the gas
+    inflow in this region.}
+    \label{fig:surf}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=9cm]{fig/vcirc.pdf}
+    \caption{The circular velocity curve of our initial setups. This
+    curve is shown for the \Nbody{} run (blue) and the SMUGGLE run (orange)
+    compared to observational estimates for the Milky
+    Way \citep{2019ApJ...871..120E}. We see that the circular velocity curve for
+    both runs is marginally larger than the Milky Way's, but still comparable. The
+    SMUGGLE circular velocity curve is larger than the \Nbody{} curve due to the
+    additional mass in the gas phase.}
+    \label{fig:vcirc}
+\end{figure}
 
@@
 \label{lastpage}
+
 \end{document}

### 2022-08-24 plots/mockHST/plot_mockHST.py
Update mockHST plot

@@ -17,6 +17,9 @@ snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
 bprop_path = '/n/home01/abeane/starbar/analysis/bar_prop/data/'
 skirt_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/SKIRT/'
 
+columnwidth = 244.0 * 0.035145980349999517 # convert to cm
+textwidth = 508.0 * 0.035145980349999517 # convert to cm
+
 # names
 Nbody = 'Nbody'
 # phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
@@ -119,7 +122,7 @@ def run():
 
     cm = 1/2.54
 
-    fig, ax = plt.subplots(2, 3, sharex=True, sharey=True, figsize=(18*cm, 12*cm))
+    fig, ax = plt.subplots(2, 3, sharex=True, sharey=True, figsize=(textwidth*cm, (12/16)*textwidth*cm))
 
     for i in range(len(Nbody_idx)):
         # plot Nbody

### 2022-08-24 paper/ms.tex

@@
 gravito-hydrodynamics code AREPO \citep{2010MNRAS.401..791S}. We use the galaxy
-formation code Stars and MUltiphase Gas in GaLaxiEs (SMUGGLE). This disk galaxy
-exhibits almost no evolution in the bar pattern speed over several Gyr when the
-gas phase is accounted for and properly modeled. This behavior has been observed
-in a few previous works \citep{1993AA...268...65F, 2010ApJ...719.1470V,
-2014MNRAS.438L..81A}, but we are the first to exhibit that the behavior ought to
-be present in a large fraction of galaxies. We are also the first to provide a
-physical explanation of the behavior with clear observational predictions. We
-show synthetic \textit{Hubble Space Telescope} images of our barred galaxy in a
-case with and without gas in Fig.~\ref{fig:overview}. We see that the bar grows
-longer and stronger without gas, while it remains at approximately the same
-length and strength when gas is included.
+formation model Stars and MUltiphase Gas in GaLaxiEs
+\citep[SMUGGLE;][]{2019MNRAS.489.4233M}. This disk galaxy exhibits almost no
+evolution in the bar pattern speed over several Gyr when the gas phase is
+accounted for and properly modeled. This behavior has been observed in a few
+previous works \citep{1993AA...268...65F, 2010ApJ...719.1470V,
+2014MNRAS.438L..81A}. However, we are the first to exhibit that the behavior
+ought to be present in a large fraction of galaxies, since it occurs in our
+simulation with gas fractions as low as about $5\%$. We are also the first to
+provide a physical explanation of the behavior with clear observational
+predictions.
+
+We show synthetic \textit{Hubble Space Telescope} images of our barred galaxy in
+a case with and without gas in Fig.~\ref{fig:overview}. We see that the bar
+grows longer and stronger without gas, while it remains at approximately the
+same length and strength when gas is included. Details on the production of
+these synthetic images are given in Appendix~\ref{app:hst}.
 
@@
 the equation of state of the gas suitably modified. Gas is allowed to
-interchange between the ``cold'' and ``hot'' components through processes such
-as cooling and stellar feedback. Cold gas is allowed to undergo star formation.
-We refer to this model as the ``smooth interstellar medium'' model, and it is
-described in more detail in \citet{2019MNRAS.489.4233M}.
+interchange between the cold and hot components through processes such as
+cooling and stellar feedback. Cold gas is allowed to undergo star formation. We
+refer to this model as the smooth interstellar medium model, and it is described
+in more detail in \citet{2019MNRAS.489.4233M}.
 
@@
 length as being the radius $R_b$ which encapsulates $99\%$ of the stars
-identified as being trapped in the bar, allowing for some outliers.
+identified as being trapped in the bar.
 
@@
 the bar by the other components is simply the negative of the torque on the
-other components by the bar. A similar method was applied in measuring the
-torque for the disk when the pattern speed is kept constant, as is done in
-Fig.~\ref{fig:equil}.
+other components by the bar.
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{fig/ps_Rbar_torque.pdf}
+    \includegraphics[width=\columnwidth]{fig/ps_torque.pdf}
     \caption{The \textit{upper panel} shows the evolution
@@
     the SMUGGLE run does not slow down and instead remains at a constant pattern
-    speed. The \textit{middle panel} shows the evolution of the bar length. In the
-    \Nbody{} case, the bar lengthens. This occurs because as the pattern speed
-    drops, bar-like orbits at larger radii are possible. Stars are captured on
-    these orbits, lengthening the bar. This process does not occur in the SMUGGLE
-    cases since the bar pattern speed is not decreasing, and therefore the bar
-    length remains constant. The \textit{lower panel} shows the torque on the bar
+    speed. The \textit{lower panel} shows the torque on the bar
     by different components. The solid lines show the torque exerted by the halo
@@
     \centering
-    \includegraphics[width=9cm]{fig/A2_over_A0.pdf}
-    \caption{The evolution in bar strength. The bar strength is measured as the
-    maximum of the second Fourier component divided by the zeroth Fourier
-    component. We see that in the \Nbody{} case (blue) the bar strength
-    increases with time, consistent with previous results showing that the bar
-    strength increases as bars slow down. In the SMUGGLE case (orange), we see
-    that the bar strength remains roughly constant, possibly slightly decreasing
-    with time. This is also consistent with the expected relation between
-    pattern speed and strength since the bar in this case is not slowing down.}
+    \includegraphics[width=9cm]{fig/Rb_A2.pdf}
+    \caption{The evolution in bar length and strength. The \textit{upper panel}
+    shows the evolution of the bar length. In the \Nbody{} case, the bar
+    lengthens. This occurs because as the pattern speed drops, bar-like orbits
+    at larger radii are possible. Stars are captured on these orbits,
+    lengthening the bar. This process does not occur in the SMUGGLE cases since
+    the bar pattern speed is not decreasing, and therefore the bar length
+    remains constant. The bar strength, shown in the \textit{lower panel} is
+    measured as the maximum of the second Fourier component divided by the
+    zeroth Fourier component. We see that in the \Nbody{} case (blue) the bar
+    strength increases with time, consistent with previous results showing that
+    the bar strength increases as bars slow down. In the SMUGGLE case (orange),
+    we see that the bar strength remains roughly constant, possibly slightly
+    decreasing with time. This is also consistent with the expected relation
+    between pattern speed and strength since the bar in this case is not slowing
+    down.}
     \label{fig:strength}
@@
 over the next $4\,\textrm{Gyr}$, compared to a $\sim43\%$ decrease in the
-pattern speed for the \Nbody{} run over the same interval. As we saw
-qualitatively in Fig.~\ref{fig:overview}, the middle panel of
-Fig.~\ref{fig:prop} shows that the length of the bar in the \Nbody{} case grows
-over time while it remains roughly constant in the SMUGGLE case. This is also
-consistent with previous numerical work, which found that bars tend to grow as
-they slow down and the radius of corotation increases
-\citep{2000ApJ...543..704D, 2003MNRAS.341.1179A}.
+pattern speed for the \Nbody{} run over the same interval.
 
@@
 
-The time evolution of the bar strength, defined as the maximum of
-$\left|A_2/A_0\right|$ as a function of radius, is shown in
-Fig.~\ref{fig:strength}. The quantity $\left|A_2/A_0\right|$ varies from $0$ to
-$1$, with larger values indicating a stronger bar pattern. We see that in the
-\Nbody{} case, $\left|A_2/A_0\right|$ increases over time as the bar pattern
-slows. This is consistent with previous \Nbody{} simulations which showed a
-clear correlation between the bar pattern speed and the bar strength
-\citep[e.g.,][]{2003MNRAS.341.1179A}. In the SMUGGLE case, we see that the bar
-strength has an initial drop but then remains at a roughly constant, but
-slightly decreasing, strength. This is consistent with the pattern speed in the
-SMUGGLE case being roughly constant or slightly increasing.
+As we saw qualitatively in Fig.~\ref{fig:overview}, the middle panel of
+Fig.~\ref{fig:strength} shows that the length of the bar in the \Nbody{} case
+grows over time while it remains roughly constant in the SMUGGLE case. This is
+also consistent with previous numerical work, which found that bars tend to grow
+as they slow down and the radius of corotation increases
+\citep{2000ApJ...543..704D, 2003MNRAS.341.1179A}. The time evolution of the bar
+strength, defined as the maximum of $\left|A_2/A_0\right|$ as a function of
+radius, is shown in the lower panel Fig.~\ref{fig:strength}. The quantity
+$\left|A_2/A_0\right|$ varies from $0$ to $1$, with larger values indicating a
+stronger bar pattern. We see that in the \Nbody{} case, $\left|A_2/A_0\right|$
+increases over time as the bar pattern slows. This is consistent with previous
+\Nbody{} simulations which showed a clear correlation between the bar pattern
+speed and the bar strength \citep[e.g.,][]{2003MNRAS.341.1179A}. In the SMUGGLE
+case, we see that the bar strength has an initial drop but then remains at a
+roughly constant, but slightly decreasing, strength. This is consistent with the
+pattern speed in the SMUGGLE case being roughly constant or slightly increasing.
 
@@
 \subsection{Pattern Speed Evolution}
-The lack of evolution in the pattern speed of the SMUGGLE case is intimately
-tied to the sudden decrease in torque exerted on the bar by the dark matter
-halo. We argue that this can be understood in terms of the halo wake mechanism.
-In the \Nbody{} case, halo material which is resonant with the bar will form a
-wake which lags behind and exerts a negative torque on the bar, slowing it down
-\citep{1984MNRAS.209..729T, 1985MNRAS.213..451W,
-1992ApJ...400...80H}.\footnote{Since the bar is not a solid body, it is not
-guaranteed that a negative torque will slow it down - e.g. a negative torque
-could shred the bar, reducing its moment of inertia without changing its pattern
-speed. However, the bar seems to empirically respond to a negative torque
-induced by a halo wake by slowing down.} As the bar slows down, the location of
-the resonances in phase space changes, allowing halo material newly resonant
-with the bar to participate in forming the wake. However, the gas is a reliable
-source of positive torque on the bar, speeding the bar up. This stops the
-resonance location from changing such that the halo cannot reinforce the wake,
-arresting the process by which the halo can slow the bar down. We term this
-process ``exhaustion.''
+The lack of evolution in the pattern speed of the SMUGGLE case (seen in
+Fig.~\ref{fig:prop}) is intimately tied to the sudden decrease in torque exerted
+on the bar by the dark matter halo. We argue that this can be understood in
+terms of the halo wake mechanism. In the \Nbody{} case, halo material which is
+resonant with the bar will form a wake which lags behind and exerts a negative
+torque on the bar, slowing it down \citep{1984MNRAS.209..729T,
+1985MNRAS.213..451W, 1992ApJ...400...80H}.\footnote{Since the bar is not a solid
+body, it is not guaranteed that a negative torque will slow it down - e.g. a
+negative torque could shred the bar, reducing its moment of inertia without
+changing its pattern speed. However, the bar seems to empirically respond to a
+negative torque induced by a halo wake by slowing down.} As the bar slows down,
+the location of the resonances in phase space changes, allowing halo material
+newly resonant with the bar to participate in forming the wake. However, the gas
+is a reliable source of positive torque on the bar, speeding the bar up. This
+stops the resonance location from changing such that the halo cannot reinforce
+the wake, arresting the process by which the halo can slow the bar down. We term
+this process ``exhaustion.''
 
@@
 \subsection{Delayed Gas Injection}
-A clear prediction of our proposed mechanism is that the constant pattern
-speed a particular galaxy will end up is somewhat arbitrary. In the real
-universe for an isolated galaxy it would be the formation pattern speed of the
-bar while in our simulation it is the pattern speed of the bar when gas is
-added to the system. We tested this by adding gas to the system at a later
-time when the bar has further grown and slowed down with time. In our
-particular test, we added the gas at a time when the pattern speed is
+A clear prediction of our proposed mechanism is that the constant pattern speed
+a particular galaxy will end up is somewhat arbitrary. In the real universe for
+an isolated galaxy it would be the formation pattern speed of the bar while in
+our simulation it is the pattern speed of the bar when gas is added to the
+system. We tested this by adding gas to the system at a later time when the bar
+has further grown and slowed down with time. In our particular test, we added
+the gas at a time when the pattern speed is
 $\sim30\,\textrm{km}/\textrm{s}/\textrm{kpc}$. As shown in
-Fig.~\ref{fig:snap700}, we find that the pattern speed evolution is very
-similar between the two cases (orange and red lines). If anything, the system
-with a lower pattern speed seems to speed up more, which is consistent with
-our picture since the stronger bar should experience a larger torque from the
-gas as it is more efficient at driving gas inflows. We also show in the
-Methods that more slowly rotating bars at fixed bar strength are more
-efficient at driving gas inflows as well. Nonetheless, when the initial pattern
-speed is lower (red line), the addition of gas does not cause the pattern
-speed to quickly return to the higher value of our fiducial simulation (orange
-line).
+Fig.~\ref{fig:snap700}, we find that the pattern speed evolution is very similar
+between the two cases (orange and red lines). If anything, the system with a
+lower pattern speed seems to speed up more, which is consistent with our picture
+since the stronger bar should experience a larger torque from the gas as it is
+more efficient at driving gas inflows. We also show in the
+Appendix~\ref{app:varyps} that more slowly rotating bars at fixed bar strength
+are more efficient at driving gas inflows as well. Nonetheless, when the initial
+pattern speed is lower (red line), the addition of gas does not cause the
+pattern speed to quickly return to the higher value of our fiducial simulation
+(orange line).
 
-\begin{figure}[t!]%
+\begin{figure}
     \centering
@@
 approximately $16\%$, $10\%$, $7\%$, and $4\%$. The pattern speed evolution is
-shown in Extended Data Fig.~\ref{fig:fgas} We find that the bar in disks with
+shown in Fig.~\ref{fig:fgas} We find that the bar in disks with
 initial surface densities of $20$, $15$, and $10\,\Msun/\textrm{pc}^2$ evolve
@@
 These criterion are slightly different and simplified from the ones used in \citet{2016MNRAS.463.1952P}, but we found them to empirically work well at
-decomposing the disk into a bar and disk component. In Extended Data
-Fig.~\ref{fig:decomp}, we show an example of this decomposition. The
+decomposing the disk into a bar and disk component. In Fig.~\ref{fig:decomp}, we show an example of this decomposition. The
 \textit{left} panel shows a surface density projection of the stellar disk and
@@
 torque on the bar by the gas phase at different rotation rates. The result of
-this experiment is illustrated in Extended Data Fig.~\ref{fig:equil}, which shows that a more
+this experiment is illustrated in Fig.~\ref{fig:equil}, which shows that a more
 slowly rotating bar experiences a larger positive torque from the gas.
@@
 using a spherically symmetric multipole fit with a maximum angular harmonic
-coefficient of $l=2$. We plot the circular velocity curve in Extended Data
+coefficient of $l=2$. We plot the circular velocity curve in
 Fig.~\ref{fig:vcirc} compared to observational

### 2022-08-24 paper/ms.tex
Updates to text

@@
 \section{Introduction}
+\the\columnwidth
+
+\the\textwidth
+
+% \showthe\textwidth
 Approximately two-thirds of spiral disks host an elongated bar-like feature at
@@
     \centering
-    \includegraphics[width=18cm]{fig/mockHST.pdf}
+    \includegraphics[width=\textwidth]{fig/mockHST.pdf}
     \caption{Synthetic Hubble Space Telescope images of our simulations with and
@@
 \label{app:milkyway}
-Our fiducial disk model is a relatively good  model of the Milky Way for several
-Gyr. This is uncommon in models of galaxies that include the gas phase of the
-disk but no circumgalactic medium. As mentioned earlier, the pattern speed seems
-to match the observed pattern speed of the Milky Way's bar
-\citep{2019MNRAS.490.4740B}. We briefly summarize some of the other ways our disk is comparable to the Milky Way.
+For several Gyr, our fiducial disk exhibits several properties in reasonable
+agreement to the Milky Way. This is uncommon in models of galaxies that include
+the gas phase of the disk but no circumgalactic medium. As mentioned earlier,
+the pattern speed seems to match the observed pattern speed of the Milky Way's
+bar \citep{2019MNRAS.490.4740B}. We briefly summarize some of the other ways our
+disk is comparable to the Milky Way.
 
 We computed the circular velocity curve of our model using the \texttt{AGAMA}
-package\cite{2019MNRAS.482.1525V}. We fit the baryonic component (stellar
-disk, bulge, gas, and newly formed stars) with an axisymmetric cylindrical
-spline with $20$ grid points in both the radial and vertical direction
-spanning $0.2$ to $50\,\textrm{kpc}$ in the radial direction and from $0.02$
-to $10\,\textrm{kpc}$ in the vertical direction. We fit the dark matter halo
-using a spherically symmetric multipole fit with a maximum angular harmonic
-coefficient of $l=2$. We plot the circular velocity curve in
-Fig.~\ref{fig:vcirc} compared to observational
-estimates\cite{2019ApJ...871..120E}. The SMUGGLE disk (which includes
-additional mass in the form of gas) has a slightly higher circular velocity
-than the \Nbody{} disk which, itself, is slightly higher than the
-observational estimates.
+package\cite{2019MNRAS.482.1525V}. We fit the baryonic component (stellar disk,
+bulge, gas, and newly formed stars) with an axisymmetric cylindrical spline with
+$20$ grid points in both the radial and vertical direction spanning $0.2$ to
+$50\,\textrm{kpc}$ in the radial direction and from $0.02$ to $10\,\textrm{kpc}$
+in the vertical direction. We fit the dark matter halo using a spherically
+symmetric multipole fit with a maximum angular harmonic coefficient of $l=2$. We
+plot the circular velocity curve in Fig.~\ref{fig:vcirc} compared to
+observational estimates\cite{2019ApJ...871..120E}. The SMUGGLE disk (which
+includes additional mass in the form of gas) has a slightly higher circular
+velocity than the \Nbody{} disk which, itself, is slightly higher than the
+observational estimates. Overall, though, the circular velocity curves between
+our model and that observed in the Milky Way are broadly consistent.
 

### 2022-08-26 paper/ms.tex
Beef up observation section

@@
 \section{Introduction}
-\the\columnwidth
-
-\the\textwidth
-
-% \showthe\textwidth
+\label{sec:intro}
 Approximately two-thirds of spiral disks host an elongated bar-like feature at
@@
 \subsection{Observations}
-We note that ref.\cite{2020MNRAS.491.3655G} found that the rotation parameter
-\Rot{} positively correlates with gas fraction, such that galaxies with higher
-gas fractions are rotating more slowly. However, it is not obvious this is in
-tension with our result since the gas fraction of galaxies correlates with
-other galactic properties\cite{2009ARAA..47..159B}. Furthermore, the
-measurement of pattern speeds is a delicate process still prone to large error
-bars.
+Observational estimates of the pattern speeds of bars indicate that nearly all
+galaxies have $1 < \Rot < 1.4$ \citep{2011MSAIS..18...23C, 2015AA...576A.102A,
+2019MNRAS.482.1733G, 2020MNRAS.491.3655G}, where $\Rot$ was defined in
+Section~\ref{sec:intro} to be $\Rot\equiv \RCR/\Rb$. This observational fact has
+long been in conflict with the theoretical expectation that bars should slow
+down, increasing \Rot{} \citep[e.g.][]{1984MNRAS.209..729T, 1985MNRAS.213..451W,
+2000ApJ...543..704D}. Explanations for this discrepancy have been given in the
+past. Some have argued that perhaps the central regions of dark matter haloes
+are less dense than we expected from $\Lambda\textrm{CDM}$
+\citep[e.g.][]{2000ApJ...543..704D,2021AA...650L..16F}. Some have argued that
+perhaps bars are recurrent, short-lived phenomena, and that all the bars we see
+in the local universe are very young (CITE). Some have argued that modifications
+to General Relativity ease the tension between the observed universe and
+$\Lambda\textrm{CDM}$ \citep[e.g.][]{2021MNRAS.503.2833R, 2021MNRAS.508..926R}.
+
+Because such a small gas fraction is necessary for our stabilizing mechanism to
+operate ($5\%$ in our Milky Way-like disk), we argue that most galaxies host a
+bar that is not slowing down. This naturally explains why most observed bars are
+fast rotators. However, we acknowledge two instances of reported discrepancies
+between our mechanism and observations.
+
+First, we note that \citet{2020MNRAS.491.3655G} found that the rotation
+parameter \Rot{} positively correlates with gas fraction, such that galaxies
+with higher gas fractions are rotating more slowly. However, it is not obvious
+this is in tension with our result since the gas fraction of galaxies correlates
+with other galactic properties \citep{2009ARAA..47..159B}. Furthermore, the
+measurement of pattern speeds is a delicate process still prone to large errors.
+
+Second, the work of \citet{2021MNRAS.500.4710C} and \citet{2021MNRAS.505.2412C}
+have made indirect measurements of the deceleration of the bar's pattern speed
+from kinematics and chemistry. We point out that these reported measurements are
+not direct measurements of the Milky Way bar's deceleration. Much like the
+simulations in the present work, the simulations of these two works do not
+properly account for the complicated formation process of the Galactic bar,
+which may leave imprints on the present day distribution of stars in spatial,
+kinematic, and chemical space. More investigation is necessary to reconcile the
+present work with these two well-executed manuscripts.
 
@@
 Barred galaxies in cosmological simulations of galaxy formation continue to be
-in conflict with observations by producing bars which rotate too
-slowly\cite{2017MNRAS.469.1054A, 2019MNRAS.483.2721P, 2021AA...650L..16F}.
-However, the pattern speeds of bars in both cosmological simulations and the
-real universe can be affected by environmental processes not included in our
-simulation -- e.g., satellite infall\cite{2011Natur.477..301P},
-non-sphericity\cite{2013MNRAS.429.1949A} or rotation\cite{2013MNRAS.434.1287S,
-2014ApJ...783L..18L, 2018MNRAS.476.1331C, 2019MNRAS.488.5788C} in the dark
-matter halo, or perhaps even the gaseous circumgalactic medium. Naturally,
-extending our present work to account for such effects is a crucial next step
-in understanding the formation and evolution of galactic bars.
+in conflict with observations by producing bars which rotate too slowly
+\citep{2017MNRAS.469.1054A, 2019MNRAS.483.2721P, 2021AA...650L..16F}. However,
+the pattern speeds of bars in both cosmological simulations and the real
+universe can be affected by environmental processes not included in our
+simulation -- e.g., satellite infall \citep{2011Natur.477..301P}, non-sphericity
+\citep{2013MNRAS.429.1949A}, rotation in the dark matter halo
+\citep{2013MNRAS.434.1287S, 2014ApJ...783L..18L, 2018MNRAS.476.1331C,
+2019MNRAS.488.5788C}, or perhaps even the gaseous circumgalactic medium.
+Naturally, extending our present work to account for such effects is a crucial
+next step in understanding the formation and evolution of galactic bars.
 
@@
 nearly all observed galaxies are fast rotators without requiring the inner
-regions of dark matter halos to be underdense\cite{1998ApJ...493L...5D,
-2000ApJ...543..704D} or requiring new physics\cite{2021MNRAS.503.2833R,
+regions of dark matter halos to be underdense \citep{1998ApJ...493L...5D,
+2000ApJ...543..704D} or requiring new physics \citep{2021MNRAS.503.2833R,
 2021MNRAS.508..926R}. Second, we show that the role of gas is of paramount
 importance in studies which attempt to uncover the nature of dark matter from
-its effect of slowing down the bar\cite{2021MNRAS.500.4710C,
+its effect of slowing down the bar \citep{2021MNRAS.500.4710C,
 2021MNRAS.505.2412C}. Third, we provide an explanation for how the Milky Way's
 bar could be both long-lived and a fast rotator, of which there is some
-observational evidence\cite{2019MNRAS.490.4740B}. And finally, we complicate
+observational evidence \citep{2019MNRAS.490.4740B}. And finally, we complicate
 the picture of radial mixing expected to sculpt the Milky Way's disk
-\cite{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies upon
+\citep{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies upon
 the pattern speed of the bar to change with time (though our work does not
 alter expectations for radial mixing induced by spiral
-arms\cite{2002MNRAS.336..785S}).
+arms \citep{2002MNRAS.336..785S}).
 
@@
 extended periods of time to be rotating very slowly. We therefore predict that
-observations which target such galaxies (e.g., lenticular barred
-galaxies\cite{2009ARAA..47..159B}) would find slowly rotating bars. There does
-exist one example of a galaxy which is known to be a slow rotator -- the
-low surface brightness galaxy UGC 628\cite{2009AA...499L..25C}. This galaxy
-has been studied in detail by ref.\cite{2016MNRAS.463.1751C}, who note that it
-indeed has a low gas fraction for galaxies of its type. We predict a general
-trend that bars in gas-rich spiral galaxies should rotate quickly while bars
-in gas-poor spiral galaxies should rotate slowly.
+observations which target such galaxies (e.g., lenticular barred galaxies
+\citep{2009ARAA..47..159B}) would find slowly rotating bars.\footnote{This does
+not mean that we predict \textit{all} gas-poor galaxies should be slowly
+rotating. Indeed, they would need to be gas-deficient for several $\textrm{Gyr}$
+before they would be classified as slow rotators.} There does exist one example
+of a galaxy which is known to be a slow rotator -- the low surface brightness
+galaxy UGC 628 \citep{2009AA...499L..25C}. This galaxy has been studied in
+detail by \citet{2016MNRAS.463.1751C}, who note that it indeed has a low gas
+fraction for galaxies of its type. We predict a general trend that bars in
+gas-rich spiral galaxies should rotate quickly while some bars in gas-poor
+spiral galaxies should rotate slowly.
 

### 2022-08-26 paper/ms.tex
Move figures around

@@
 \label{sec:intro}
-Approximately two-thirds of spiral disks host an elongated bar-like feature at
-their center \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our own
-Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of bars
-is not difficult to explain, since stellar disks simulated in isolation almost
-always form bar-like structures \citep{1971ApJ...168..343H}. Several studies
-have shown that a spherical potential, such as a stellar bulge or dark matter
-halo, acts to stabilize a stellar disk against the bar instability
-\citep{1973ApJ...186..467O, 1976AJ.....81...30H}.
-
-It is more difficult to reconcile numerical simulations with the observed
-pattern speeds of extragalactic bars. Currently, the best technique for
-measuring the pattern speeds of individual galaxies is the Tremaine-Weinberg
-method \citep{1984ApJ...282L...5T, 2011MSAIS..18...23C}. This method has
-recently been applied to samples of galaxies from the MaNGA survey
-\citep{2019MNRAS.482.1733G, 2020MNRAS.491.3655G}. These recent studies confirm
-what was found in earlier studies -- that nearly all extragalactic bars are fast
-rotators (i.e., they rotate close to their maximum rotation rate)
-
-This is a problem for theoretical simulations, for which there is ample evidence
-that galactic bars should resonantly interact with the dark matter halo, causing
-the bar to slow down over time \citep{1992ApJ...400...80H, 2000ApJ...543..704D,
-2002MNRAS.330...35A, 2002ApJ...569L..83A, 2003MNRAS.341.1179A,
-2003MNRAS.346..251O, 2005MNRAS.363..991H, 2006ApJ...637..214M,
-2007MNRAS.375..460W, 2009ApJ...697..293D}. The physical mechanism of this
-interaction can be understood as an angular form of dynamical friction between
-the bar and the dark matter halo. While studied in detail for the bar
-\citep{1984MNRAS.209..729T, 1985MNRAS.213..451W}, this process is generic for
-any non-axisymmetric disturbance \citep{1972MNRAS.157....1L}. For an old but
-still useful review of bar dynamics, see \citet{1993RPPh...56..173S}
-
-Bar pattern speeds are usually measured using the parameter $\Rot=\RCR/\Rb$,
-where \RCR{} is the corotation radius and \Rb{} is the bar length.\footnote{The
-radius of corotation \RCR{} is defined for circular orbits as the radius at
-which the orbital frequency is equal to the pattern speed, \PS{}, of a given
-non-axisymmetric feature. In a galaxy with a constant circular velocity \Vc{},
-it is given by $\RCR = \Vc / \PS$.} Galaxies with $\Rot < 1.4$ are considered
-``fast rotators'' while galaxies with $\Rot > 1.4$ are considered ``slow
-rotators'' \citep{2000ApJ...543..704D}. Galaxies with $\Rot < 1$ are not thought
-to be stable \citep{1980AA....81..198C}. Observational estimates of the pattern
-speeds of bars indicate that nearly all galaxies have $1 < \Rot <
-1.4$ \citep{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G,
-2020MNRAS.491.3655G}.
-
-While the interaction between the bar and a dark matter halo is well-understood
-theoretically, this is not the case for the interaction between a bar and a
-gaseous disk. Some argue that the gas disk should slow down the bar more
-\citep{2003MNRAS.341.1179A}, while others argue that the tendency of the bar to
-drive gas inwards means the bar should speed up due to the effect of the gas
-disk \citep{2013MNRAS.429.1949A, 2014MNRAS.438L..81A}. Since the gas phase
-typically only contributes about $10-20\%$ of the mass of a galaxy at the
-present day, one might naively expect it to have a subdominant effect on the
-bar. However, because gas is collisional, it can participate in non-resonant
-angular momentum exchange with the bar \citep{2011MNRAS.415.1027H}. Thus,
-numerical work has shown that the gas phase can have a stronger influence on a
-bar than its contribution to the mass of a galaxy would suggest
-\citep{2010ApJ...719.1470V, 2013MNRAS.429.1949A}.
-
-We have performed a simulation of a disk galaxy using the finite-volume,
-gravito-hydrodynamics code AREPO \citep{2010MNRAS.401..791S}. We use the galaxy
-formation model Stars and MUltiphase Gas in GaLaxiEs
-\citep[SMUGGLE;][]{2019MNRAS.489.4233M}. This disk galaxy exhibits almost no
-evolution in the bar pattern speed over several Gyr when the gas phase is
-accounted for and properly modeled. This behavior has been observed in a few
-previous works \citep{1993AA...268...65F, 2010ApJ...719.1470V,
-2014MNRAS.438L..81A}. However, we are the first to exhibit that the behavior
-ought to be present in a large fraction of galaxies, since it occurs in our
-simulation with gas fractions as low as about $5\%$. We are also the first to
-provide a physical explanation of the behavior with clear observational
-predictions.
-
-We show synthetic \textit{Hubble Space Telescope} images of our barred galaxy in
-a case with and without gas in Fig.~\ref{fig:overview}. We see that the bar
-grows longer and stronger without gas, while it remains at approximately the
-same length and strength when gas is included. Details on the production of
-these synthetic images are given in Appendix~\ref{app:hst}.
-
-% In this work, we present a simulation of an isolated, gas-rich disk galaxy. Our
-% galaxy exhibits some similarities to the Milky Way, described in more detail in
-% Section XX. Surprisingly, we find that the pattern speed of the bar in this
-% galaxy is constant with time. We find that with even modest gas fractions of
-% about 5\%, our isolated galaxy displays no evolution in pattern speed over many
-% $\textrm{Gyr}$.
-
-In Section~\ref{sec:methods}, we describe our initial setup, numerical model,
-and details on our bar analysis procedures. In Section~\ref{sec:results}, we
-summarize the main results from our findings. We discuss these findings at more
-length and in the context of previous research in Section~\ref{sec:discussion}
-before concluding in Section~\ref{sec:conclusions}.
-
 \begin{figure*}
@@
 \end{figure*}
+
+Approximately two-thirds of spiral disks host an elongated bar-like feature at
+their center \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our own
+Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of bars
+is not difficult to explain, since stellar disks simulated in isolation almost
+always form bar-like structures \citep{1971ApJ...168..343H}. Several studies
+have shown that a spherical potential, such as a stellar bulge or dark matter
+halo, acts to stabilize a stellar disk against the bar instability
+\citep{1973ApJ...186..467O, 1976AJ.....81...30H}.
+
+It is more difficult to reconcile numerical simulations with the observed
+pattern speeds of extragalactic bars. Currently, the best technique for
+measuring the pattern speeds of individual galaxies is the Tremaine-Weinberg
+method \citep{1984ApJ...282L...5T, 2011MSAIS..18...23C}. This method has
+recently been applied to samples of galaxies from the MaNGA survey
+\citep{2019MNRAS.482.1733G, 2020MNRAS.491.3655G}. These recent studies confirm
+what was found in earlier studies -- that nearly all extragalactic bars are fast
+rotators (i.e., they rotate close to their maximum rotation rate)
+
+This is a problem for theoretical simulations, for which there is ample evidence
+that galactic bars should resonantly interact with the dark matter halo, causing
+the bar to slow down over time \citep{1992ApJ...400...80H, 2000ApJ...543..704D,
+2002MNRAS.330...35A, 2002ApJ...569L..83A, 2003MNRAS.341.1179A,
+2003MNRAS.346..251O, 2005MNRAS.363..991H, 2006ApJ...637..214M,
+2007MNRAS.375..460W, 2009ApJ...697..293D}. The physical mechanism of this
+interaction can be understood as an angular form of dynamical friction between
+the bar and the dark matter halo. While studied in detail for the bar
+\citep{1984MNRAS.209..729T, 1985MNRAS.213..451W}, this process is generic for
+any non-axisymmetric disturbance \citep{1972MNRAS.157....1L}. For an old but
+still useful review of bar dynamics, see \citet{1993RPPh...56..173S}
+
+Bar pattern speeds are usually measured using the parameter $\Rot=\RCR/\Rb$,
+where \RCR{} is the corotation radius and \Rb{} is the bar length.\footnote{The
+radius of corotation \RCR{} is defined for circular orbits as the radius at
+which the orbital frequency is equal to the pattern speed, \PS{}, of a given
+non-axisymmetric feature. In a galaxy with a constant circular velocity \Vc{},
+it is given by $\RCR = \Vc / \PS$.} Galaxies with $\Rot < 1.4$ are considered
+``fast rotators'' while galaxies with $\Rot > 1.4$ are considered ``slow
+rotators'' \citep{2000ApJ...543..704D}. Galaxies with $\Rot < 1$ are not thought
+to be stable \citep{1980AA....81..198C}. Observational estimates of the pattern
+speeds of bars indicate that nearly all galaxies have $1 < \Rot <
+1.4$ \citep{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G,
+2020MNRAS.491.3655G}.
+
+While the interaction between the bar and a dark matter halo is well-understood
+theoretically, this is not the case for the interaction between a bar and a
+gaseous disk. Some argue that the gas disk should slow down the bar more
+\citep{2003MNRAS.341.1179A}, while others argue that the tendency of the bar to
+drive gas inwards means the bar should speed up due to the effect of the gas
+disk \citep{2013MNRAS.429.1949A, 2014MNRAS.438L..81A}. Since the gas phase
+typically only contributes about $10-20\%$ of the mass of a galaxy at the
+present day, one might naively expect it to have a subdominant effect on the
+bar. However, because gas is collisional, it can participate in non-resonant
+angular momentum exchange with the bar \citep{2011MNRAS.415.1027H}. Thus,
+numerical work has shown that the gas phase can have a stronger influence on a
+bar than its contribution to the mass of a galaxy would suggest
+\citep{2010ApJ...719.1470V, 2013MNRAS.429.1949A}.
+
+We have performed a simulation of a disk galaxy using the finite-volume,
+gravito-hydrodynamics code AREPO \citep{2010MNRAS.401..791S}. We use the galaxy
+formation model Stars and MUltiphase Gas in GaLaxiEs
+\citep[SMUGGLE;][]{2019MNRAS.489.4233M}. This disk galaxy exhibits almost no
+evolution in the bar pattern speed over several Gyr when the gas phase is
+accounted for and properly modeled. This behavior has been observed in a few
+previous works \citep{1993AA...268...65F, 2010ApJ...719.1470V,
+2014MNRAS.438L..81A}. However, we are the first to exhibit that the behavior
+ought to be present in a large fraction of galaxies, since it occurs in our
+simulation with gas fractions as low as about $5\%$. We are also the first to
+provide a physical explanation of the behavior with clear observational
+predictions.
+
+We show synthetic \textit{Hubble Space Telescope} images of our barred galaxy in
+a case with and without gas in Fig.~\ref{fig:overview}. We see that the bar
+grows longer and stronger without gas, while it remains at approximately the
+same length and strength when gas is included. Details on the production of
+these synthetic images are given in Appendix~\ref{app:hst}.
+
+% In this work, we present a simulation of an isolated, gas-rich disk galaxy. Our
+% galaxy exhibits some similarities to the Milky Way, described in more detail in
+% Section XX. Surprisingly, we find that the pattern speed of the bar in this
+% galaxy is constant with time. We find that with even modest gas fractions of
+% about 5\%, our isolated galaxy displays no evolution in pattern speed over many
+% $\textrm{Gyr}$.
+
+In Section~\ref{sec:methods}, we describe our initial setup, numerical model,
+and details on our bar analysis procedures. In Section~\ref{sec:results}, we
+summarize the main results from our findings. We discuss these findings at more
+length and in the context of previous research in Section~\ref{sec:discussion}
+before concluding in Section~\ref{sec:conclusions}.
 
@@
 \label{sec:methods}
-
-\subsection{Initial Conditions}
-The initial setup of the galactic disk used in this work follows closely the
-GALAKOS model \citep{2020ApJ...890..117D}, which uses a modified version of the
-\texttt{MakeNewDisk} code \citep{2005MNRAS.361..776S}. The GALAKOS model has
-three components - a radially exponential and vertically isothermal stellar
-disk, and a stellar bulge and dark matter halo following a Hernquist profile
-\citep{1990ApJ...356..359H}. All \Nbody{} runs in this work used the same setup
-parameters as the GALAKOS disk, more details of which can be found in the
-original paper.
-
-The addition of the gas phase was done as follows. The version of
-\texttt{MakeNewDisk} used for the original GALAKOS model can generate a gas disk
-which is radially exponential and in vertical gravito-hydrodynamic balance. We
-modified the radial profile of this code in order to allow us to generate a disk
-with a constant surface density within some cut-off radius, and then
-exponentially declining beyond that radius with the scale-length of the stellar
-disk. Our fiducial model used an initial surface density of
-$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\textrm{kpc}$. The
-initial gas disk is generated with a temperature of $10^4\,\textrm{K}$ and solar
-metallicity.
-
-After generating the gaseous disk in this way, we stitched the gas disk together
-with the GALAKOS \Nbody{} disk (and bulge and dark matter halo) after the
-GALAKOS disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The purpose of
-allowing the GALAKOS disk to evolve first for a short period of time is to allow
-for the bar to form unimpacted by the presence of the gas. We found that
-including the gas before the bar has formed disrupts the formation of the bar,
-as has been seen in previous work \citep[e.g.,][]{2013MNRAS.429.1949A}.
-Throughout this work, we consider $t=0$ for the \Nbody{} run to be the time at
-which we added the gas phase for the SMUGGLE run (i.e., we ignore the first
-$1.5\,\textrm{Gyr}$ of evolution of the \Nbody{} disk when the bar is forming).
-
-We made one additional modification when stitching the gas disk together with
-the \Nbody{} disk - we created a hole within the central $4\,\textrm{kpc}$. This
-hole guards against an initial dramatic infall of gas within the bar region,
-which we found to destroy the bar. It is not uncommon for observed barred
-galaxies to have gas deficits in the bar region \citep[though not in the very
-center;][]{1993RPPh...56..173S}. Therefore, our practice of allowing the gas
-distribution to have a hole in the central region is consistent with our choice
-to begin the simulations with a bar already formed.
-
-We used a mass resolution of $7.5\times10^3\,M_{\odot}$ for the baryonic
-components (initial stellar disk, stellar bulge, and gas) and a mass resolution
-of $3.75\times10^4\,M_{\odot}$ for the dark matter halo. This mass resolution is
-closest to ``level 3'' in the AURIGA simulations \citep{2017MNRAS.467..179G}.
-This corresponds to approximately $6.4\times10^6$ particles in the stellar disk,
-$1.1\times10^6$ in the bulge, $1.2\times10^6$ in the gas disk, and
-$25.3\times10^6$ in the dark matter halo. We used a softening length of
-$0.02\,\textrm{kpc}$ for all components. Snapshots were saved at equal intervals
-of $0.005$ in the time units of the simulation,
-$\textrm{kpc}/(\textrm{km}/\textrm{s})$.
-
-Our setup is initially out of equilibrium, but we found that after about
-$500\,\textrm{Myr}$, the system has settled into a steady-state configuration
-and initial transients appear not to affect the results after this point. The
-constant surface density of the initial gas disk is important for ensuring the
-gas disk is dense enough in order for comparisons to real galaxies to be
-appropriate.
-
-\subsection{Numerical Model}
-We use the Stars and MUltiphase Gas in GaLaxiEs (SMUGGLE) model
-\citep{2019MNRAS.489.4233M} implemented within the moving-mesh, finite-volume
-hydrodynamics and gravity code \AREPO{} \citep{2010MNRAS.401..791S}. The SMUGGLE
-model additionally includes radiative heating and cooling, star formation, and
-stellar feedback. Explicit gas cooling and heating of the multi-phase
-interstellar medium is implemented, covering temperature ranges between $10$ and
-$10^8\,\textrm{K}$.
-
-Star formation occurs in cells above a density threshold
-($n_{\textrm{th}}=100\,\textrm{cm}^{-3}$) with a star-formation efficiency of
-$\epsilon = 0.01$. Star formation converts gas cells into star particles which
-represent single stellar populations with a Chabrier initial mass
-function \citep{2003PASP..115..763C}. For each star particle, the deposition of
-energy, momentum, mass, and metals from stellar winds and supernovae is modeled.
-Photo-ionization and radiation pressure are modeled using an approximate
-treatment. A more detailed description of this model can be found in the
-flagship SMUGGLE paper \citep{2019MNRAS.489.4233M}.
-
-We used the fiducial model parameters, except that we increased the number of
-effective neighbors $N_{\textrm{ngb}}$ for the deposition of feedback from
-$64$ to $512$. We found that a lower value of $N_{\textrm{ngb}}$ resulted in
-inefficient photo-ionization feedback since the photo-ionizing budget had not
-been exhausted after deposition into $64$ neighboring cells. We also used an
-updated version of SMUGGLE using a new mechanical feedback routine similar to
-the one described in \citet{2018MNRAS.480..800H}. This updated routine is a
-tensor renormalization which ensures linear and angular momentum conservation
-to machine precision.
-
-In addition to the SMUGGLE model, we considered a simpler model of the
-interstellar medium based upon \citet{2003MNRAS.339..289S}. In this model, the
-multiphase nature of the interstellar medium is modeled in a subgrid manner by
-allowing each resolution element to have a ``cold'' and ``hot'' component, with
-the equation of state of the gas suitably modified. Gas is allowed to
-interchange between the cold and hot components through processes such as
-cooling and stellar feedback. Cold gas is allowed to undergo star formation. We
-refer to this model as the smooth interstellar medium model, and it is described
-in more detail in \citet{2019MNRAS.489.4233M}.
-
-\subsection{Bar Analysis}
-The analysis of various bar properties is performed as follows. First, the
-pattern speed is measured from the angle of the second Fourier component. We
-measured the second Fourier component by computing,
-\begin{equation}
-\begin{split}
-A_2 &= \sum_i m_i e^{i 2 \phi_i} \\
-A_0 &= \sum_i m_i \textrm{,}
-\end{split}
-\end{equation}
-where $m_i$ and $\phi_i$ are the mass and azimuthal angle of each particle,
-respectively. We computed $A_2$ and $A_0$ in cylindrical bins of width
-$0.5\,\textrm{kpc}$ from radii of $0$ to $30\,\textrm{kpc}$. We defined the
-angle of the bar $\phi_b$ to be twice the angle of the complex number $A_2$ as
-measured in the bin extending from a radius of $2.5$ to $3\,\textrm{kpc}$.
-After correcting for the periodicity of $\phi_b$, we measured the pattern
-speed as one-half the two-sided finite gradient of $\phi_b$ as a function of
-time.
-
-In order to compute other properties of the bar, it is necessary to decompose
-the disk into a barred and unbarred component. We achieved this by following
-closely the methods described in \citet{2016MNRAS.463.1952P}. Our implementation
-is described in more detail in Appendix~\ref{app:bardecomp}. After the disk has
-been decomposed into a trapped and untrapped component, we measured the bar
-length as being the radius $R_b$ which encapsulates $99\%$ of the stars
-identified as being trapped in the bar.
-
-To compute torques we used the tree algorithm in \texttt{MakeNewDisk}
-\citep{2005MNRAS.361..776S} customized to be accessible from \texttt{Python}
-using \texttt{Cython}. This algorithm is based on the \texttt{TREESPH} code
-\citep{1989ApJS...70..419H}. We constructed a tree with an opening angle of
-$0.35$ using only the star particles identified as being trapped in the bar. We
-then queried the tree at the locations of all resolution elements in the other
-components and computed the torque of the bar on such components. The torque on
-the bar by the other components is simply the negative of the torque on the
-other components by the bar.
-
-\subsection{Plotting Details}
-We saved snapshots in intervals of $0.005$ in the time units of the
-simulation, $\textrm{kpc}/(\textrm{km}/\textrm{s})$, which is very nearly
-equal to $1\,\textrm{Gyr}$ (it is $\sim0.977\,\textrm{Gyr}$). Therefore,
-throughout this work we referred to the native code time unit as
-$\textrm{Gyr}$. None of our results are sensitive to this choice.
-
-In order to remove numerical noise in several quantities we computed, we
-applied a Savitzky-Golay filter \citep{1964AnaCh..36.1627S} as implemented in
-\texttt{scipy} using a window length of $81$ and polynomial order of $3$. This
-filter was applied to plots of pattern speeds, bar lengths and strengths,
-torques, and angle differences.
-
-\section{Results}
-\label{sec:results}
 \begin{figure}
@@
 \end{figure}
+\subsection{Initial Conditions}
+The initial setup of the galactic disk used in this work follows closely the
+GALAKOS model \citep{2020ApJ...890..117D}, which uses a modified version of the
+\texttt{MakeNewDisk} code \citep{2005MNRAS.361..776S}. The GALAKOS model has
+three components - a radially exponential and vertically isothermal stellar
+disk, and a stellar bulge and dark matter halo following a Hernquist profile
+\citep{1990ApJ...356..359H}. All \Nbody{} runs in this work used the same setup
+parameters as the GALAKOS disk, more details of which can be found in the
+original paper.
+
+The addition of the gas phase was done as follows. The version of
+\texttt{MakeNewDisk} used for the original GALAKOS model can generate a gas disk
+which is radially exponential and in vertical gravito-hydrodynamic balance. We
+modified the radial profile of this code in order to allow us to generate a disk
+with a constant surface density within some cut-off radius, and then
+exponentially declining beyond that radius with the scale-length of the stellar
+disk. Our fiducial model used an initial surface density of
+$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\textrm{kpc}$. The
+initial gas disk is generated with a temperature of $10^4\,\textrm{K}$ and solar
+metallicity.
+
+After generating the gaseous disk in this way, we stitched the gas disk together
+with the GALAKOS \Nbody{} disk (and bulge and dark matter halo) after the
+GALAKOS disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The purpose of
+allowing the GALAKOS disk to evolve first for a short period of time is to allow
+for the bar to form unimpacted by the presence of the gas. We found that
+including the gas before the bar has formed disrupts the formation of the bar,
+as has been seen in previous work \citep[e.g.,][]{2013MNRAS.429.1949A}.
+Throughout this work, we consider $t=0$ for the \Nbody{} run to be the time at
+which we added the gas phase for the SMUGGLE run (i.e., we ignore the first
+$1.5\,\textrm{Gyr}$ of evolution of the \Nbody{} disk when the bar is forming).
+
+We made one additional modification when stitching the gas disk together with
+the \Nbody{} disk - we created a hole within the central $4\,\textrm{kpc}$. This
+hole guards against an initial dramatic infall of gas within the bar region,
+which we found to destroy the bar. It is not uncommon for observed barred
+galaxies to have gas deficits in the bar region \citep[though not in the very
+center;][]{1993RPPh...56..173S}. Therefore, our practice of allowing the gas
+distribution to have a hole in the central region is consistent with our choice
+to begin the simulations with a bar already formed.
+
+We used a mass resolution of $7.5\times10^3\,M_{\odot}$ for the baryonic
+components (initial stellar disk, stellar bulge, and gas) and a mass resolution
+of $3.75\times10^4\,M_{\odot}$ for the dark matter halo. This mass resolution is
+closest to ``level 3'' in the AURIGA simulations \citep{2017MNRAS.467..179G}.
+This corresponds to approximately $6.4\times10^6$ particles in the stellar disk,
+$1.1\times10^6$ in the bulge, $1.2\times10^6$ in the gas disk, and
+$25.3\times10^6$ in the dark matter halo. We used a softening length of
+$0.02\,\textrm{kpc}$ for all components. Snapshots were saved at equal intervals
+of $0.005$ in the time units of the simulation,
+$\textrm{kpc}/(\textrm{km}/\textrm{s})$.
+
+Our setup is initially out of equilibrium, but we found that after about
+$500\,\textrm{Myr}$, the system has settled into a steady-state configuration
+and initial transients appear not to affect the results after this point. The
+constant surface density of the initial gas disk is important for ensuring the
+gas disk is dense enough in order for comparisons to real galaxies to be
+appropriate.
+
+\subsection{Numerical Model}
+We use the Stars and MUltiphase Gas in GaLaxiEs (SMUGGLE) model
+\citep{2019MNRAS.489.4233M} implemented within the moving-mesh, finite-volume
+hydrodynamics and gravity code \AREPO{} \citep{2010MNRAS.401..791S}. The SMUGGLE
+model additionally includes radiative heating and cooling, star formation, and
+stellar feedback. Explicit gas cooling and heating of the multi-phase
+interstellar medium is implemented, covering temperature ranges between $10$ and
+$10^8\,\textrm{K}$.
+
+Star formation occurs in cells above a density threshold
+($n_{\textrm{th}}=100\,\textrm{cm}^{-3}$) with a star-formation efficiency of
+$\epsilon = 0.01$. Star formation converts gas cells into star particles which
+represent single stellar populations with a Chabrier initial mass
+function \citep{2003PASP..115..763C}. For each star particle, the deposition of
+energy, momentum, mass, and metals from stellar winds and supernovae is modeled.
+Photo-ionization and radiation pressure are modeled using an approximate
+treatment. A more detailed description of this model can be found in the
+flagship SMUGGLE paper \citep{2019MNRAS.489.4233M}.
+
+We used the fiducial model parameters, except that we increased the number of
+effective neighbors $N_{\textrm{ngb}}$ for the deposition of feedback from
+$64$ to $512$. We found that a lower value of $N_{\textrm{ngb}}$ resulted in
+inefficient photo-ionization feedback since the photo-ionizing budget had not
+been exhausted after deposition into $64$ neighboring cells. We also used an
+updated version of SMUGGLE using a new mechanical feedback routine similar to
+the one described in \citet{2018MNRAS.480..800H}. This updated routine is a
+tensor renormalization which ensures linear and angular momentum conservation
+to machine precision.
+
+In addition to the SMUGGLE model, we considered a simpler model of the
+interstellar medium based upon \citet{2003MNRAS.339..289S}. In this model, the
+multiphase nature of the interstellar medium is modeled in a subgrid manner by
+allowing each resolution element to have a ``cold'' and ``hot'' component, with
+the equation of state of the gas suitably modified. Gas is allowed to
+interchange between the cold and hot components through processes such as
+cooling and stellar feedback. Cold gas is allowed to undergo star formation. We
+refer to this model as the smooth interstellar medium model, and it is described
+in more detail in \citet{2019MNRAS.489.4233M}.
+
+\subsection{Bar Analysis}
+The analysis of various bar properties is performed as follows. First, the
+pattern speed is measured from the angle of the second Fourier component. We
+measured the second Fourier component by computing,
+\begin{equation}
+\begin{split}
+A_2 &= \sum_i m_i e^{i 2 \phi_i} \\
+A_0 &= \sum_i m_i \textrm{,}
+\end{split}
+\end{equation}
+where $m_i$ and $\phi_i$ are the mass and azimuthal angle of each particle,
+respectively. We computed $A_2$ and $A_0$ in cylindrical bins of width
+$0.5\,\textrm{kpc}$ from radii of $0$ to $30\,\textrm{kpc}$. We defined the
+angle of the bar $\phi_b$ to be twice the angle of the complex number $A_2$ as
+measured in the bin extending from a radius of $2.5$ to $3\,\textrm{kpc}$.
+After correcting for the periodicity of $\phi_b$, we measured the pattern
+speed as one-half the two-sided finite gradient of $\phi_b$ as a function of
+time.
+
+In order to compute other properties of the bar, it is necessary to decompose
+the disk into a barred and unbarred component. We achieved this by following
+closely the methods described in \citet{2016MNRAS.463.1952P}. Our implementation
+is described in more detail in Appendix~\ref{app:bardecomp}. After the disk has
+been decomposed into a trapped and untrapped component, we measured the bar
+length as being the radius $R_b$ which encapsulates $99\%$ of the stars
+identified as being trapped in the bar.
+
+To compute torques we used the tree algorithm in \texttt{MakeNewDisk}
+\citep{2005MNRAS.361..776S} customized to be accessible from \texttt{Python}
+using \texttt{Cython}. This algorithm is based on the \texttt{TREESPH} code
+\citep{1989ApJS...70..419H}. We constructed a tree with an opening angle of
+$0.35$ using only the star particles identified as being trapped in the bar. We
+then queried the tree at the locations of all resolution elements in the other
+components and computed the torque of the bar on such components. The torque on
+the bar by the other components is simply the negative of the torque on the
+other components by the bar.
+
+\subsection{Plotting Details}
+We saved snapshots in intervals of $0.005$ in the time units of the
+simulation, $\textrm{kpc}/(\textrm{km}/\textrm{s})$, which is very nearly
+equal to $1\,\textrm{Gyr}$ (it is $\sim0.977\,\textrm{Gyr}$). Therefore,
+throughout this work we referred to the native code time unit as
+$\textrm{Gyr}$. None of our results are sensitive to this choice.
+
+In order to remove numerical noise in several quantities we computed, we
+applied a Savitzky-Golay filter \citep{1964AnaCh..36.1627S} as implemented in
+\texttt{scipy} using a window length of $81$ and polynomial order of $3$. This
+filter was applied to plots of pattern speeds, bar lengths and strengths,
+torques, and angle differences.
+
+\section{Results}
+\label{sec:results}
 
@@
 pattern speed in the SMUGGLE case being roughly constant or slightly increasing.
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=18cm]{fig/halo_wake.pdf}
+    \caption{The wake excited in the dark matter halo. The dark matter halo wake
+    is shown in the \Nbody{} case (\textit{middle panel}) and SMUGGLE case
+    (\textit{right panel}) after $2.6\,\textrm{Gyr}$ of evolution. The
+    \textit{middle} and \textit{right panels} show a surface density projection
+    in the $x$-$y$ plane of the dark matter halo after an axisymmetric average
+    has been subtracted. The solid line indicates the direction of the bar while
+    the dashed line indicates the direction of the halo wake (both measured by
+    taking the second Fourier component within a sphere of all material within a
+    radius of $4\,\textrm{kpc}$). The \textit{left panel} shows the time
+    evolution of the angle difference between the bar and the halo wake, as
+    measured from the second Fourier component. After the first Gyr, the angle
+    difference in the SMUGGLE case is smaller than in the \Nbody{} case by about
+    a factor of two, reflecting how the dark matter halo in the SMUGGLE case is
+    unable to exert as negative a torque on the bar as in the \Nbody{} case.}
+    \label{fig:wake}
+\end{figure*}
 
@@
 
-\begin{figure*}
-    \centering
-    \includegraphics[width=18cm]{fig/halo_wake.pdf}
-    \caption{The wake excited in the dark matter halo. The dark matter halo wake
-    is shown in the \Nbody{} case (\textit{middle panel}) and SMUGGLE case
-    (\textit{right panel}) after $2.6\,\textrm{Gyr}$ of evolution. The
-    \textit{middle} and \textit{right panels} show a surface density projection
-    in the $x$-$y$ plane of the dark matter halo after an axisymmetric average
-    has been subtracted. The solid line indicates the direction of the bar while
-    the dashed line indicates the direction of the halo wake (both measured by
-    taking the second Fourier component within a sphere of all material within a
-    radius of $4\,\textrm{kpc}$). The \textit{left panel} shows the time
-    evolution of the angle difference between the bar and the halo wake, as
-    measured from the second Fourier component. After the first Gyr, the angle
-    difference in the SMUGGLE case is smaller than in the \Nbody{} case by about
-    a factor of two, reflecting how the dark matter halo in the SMUGGLE case is
-    unable to exert as negative a torque on the bar as in the \Nbody{} case.}
-    \label{fig:wake}
-\end{figure*}
-
 We can test this interpretation by measuring the angle offset between the halo
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=9cm]{fig/ps_late_start.pdf}
-    \caption{Pattern speed evolution with a lower initial pattern speed. We
-    tested the evolution of our system when gas is added to the \Nbody{} run at
-    a later time, but with all other simulation parameters kept the same. The
-    setup is therefore identical to our previous runs just with a bar that is
-    larger, stronger, and with a lower pattern speed. We find that the pattern
-    speed evolution is very similar to our fiducial case, except that the bar
-    retains its original pattern speed. This indicates a mechanism which keeps
-    the bar at its formation pattern speed, and that there is not a particular
-    pattern speed which the system tends to.}
-    \label{fig:snap700}
-\end{figure}
-
 \subsection{Varying Initial Gas Fractions}
@@
 
-\begin{figure}[t!]
-    \centering
-    \includegraphics[width=9cm]{fig/ps_fgas.pdf}
-    \caption{Pattern speed evolution with varying gas fractions. We explored the
-    impact of lowering the initial gas surface density of our fiducial disk on
-    the evolution of the pattern speed. The surface densities we tested of $20$,
-    $15$, $10$, and $5\,\Msun/\textrm{kpc}^2$ correspond to initial gas
-    fractions of $16\%$, $10\%$, $7\%$, and $4\%$. We find that initial surface
-    densities $20$, $15$, and $10\,\Msun/\textrm{kpc}^2$ result in bars which
-    remain at a constant pattern speed, while an initial surface density of
-    $5\,\Msun/\textrm{kpc}^2$ results in a bar which slows down in roughly the
-    same manner as the \Nbody{} case.}
-    \label{fig:fgas}
-\end{figure}
-
 \subsection{Smooth Interstellar Medium}
@@
 on the physics of this test is given in Section.~X.
-
-\begin{figure}[h!]%
-    \centering
-    \includegraphics[width=9cm]{fig/ps_GFM.pdf}
-    \caption{Pattern speed evolution of a smooth ISM model. This evolution is
-    shown for the fiducial disk in the \Nbody{} (blue), SMUGGLE (orange), and
-    smooth ISM (red) cases. The smooth ISM model is an older model for the ISM
-    which treats its multiphase nature in a subgrid fashion
-    \citep{2003MNRAS.339..289S}. This fundamentally differs from the SMUGGLE
-    model, which explicitly resolves the hot and cold phases of the ISM
-    \citep{2019MNRAS.489.4233M}. The pattern speed in the smooth ISM case is
-    broadly similar to the evolution in the SMUGGLE case. This shows that the
-    stability of the pattern speed is not simply a result of our assumed model
-    for the ISM.}
-\label{fig:GFM}
-\end{figure}
 
@@
 measurement of pattern speeds is a delicate process still prone to large errors.
+\begin{figure}
+    \centering
+    \includegraphics[width=9cm]{fig/ps_late_start.pdf}
+    \caption{Pattern speed evolution with a lower initial pattern speed. We
+    tested the evolution of our system when gas is added to the \Nbody{} run at
+    a later time, but with all other simulation parameters kept the same. The
+    setup is therefore identical to our previous runs just with a bar that is
+    larger, stronger, and with a lower pattern speed. We find that the pattern
+    speed evolution is very similar to our fiducial case, except that the bar
+    retains its original pattern speed. This indicates a mechanism which keeps
+    the bar at its formation pattern speed, and that there is not a particular
+    pattern speed which the system tends to.}
+    \label{fig:snap700}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=9cm]{fig/ps_fgas.pdf}
+    \caption{Pattern speed evolution with varying gas fractions. We explored the
+    impact of lowering the initial gas surface density of our fiducial disk on
+    the evolution of the pattern speed. The surface densities we tested of $20$,
+    $15$, $10$, and $5\,\Msun/\textrm{kpc}^2$ correspond to initial gas
+    fractions of $16\%$, $10\%$, $7\%$, and $4\%$. We find that initial surface
+    densities $20$, $15$, and $10\,\Msun/\textrm{kpc}^2$ result in bars which
+    remain at a constant pattern speed, while an initial surface density of
+    $5\,\Msun/\textrm{kpc}^2$ results in a bar which slows down in roughly the
+    same manner as the \Nbody{} case.}
+    \label{fig:fgas}
+\end{figure}
 
@@
 next step in understanding the formation and evolution of galactic bars.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=9cm]{fig/ps_GFM.pdf}
+    \caption{Pattern speed evolution of a smooth ISM model. This evolution is
+    shown for the fiducial disk in the \Nbody{} (blue), SMUGGLE (orange), and
+    smooth ISM (red) cases. The smooth ISM model is an older model for the ISM
+    which treats its multiphase nature in a subgrid fashion
+    \citep{2003MNRAS.339..289S}. This fundamentally differs from the SMUGGLE
+    model, which explicitly resolves the hot and cold phases of the ISM
+    \citep{2019MNRAS.489.4233M}. The pattern speed in the smooth ISM case is
+    broadly similar to the evolution in the SMUGGLE case. This shows that the
+    stability of the pattern speed is not simply a result of our assumed model
+    for the ISM.}
+\label{fig:GFM}
+\end{figure}
 

### 2022-08-29 paper/ms.tex

@@
 Way-like disk, a gas fraction of only about 5\% is necessary for this mechanism
-to operate. This result naturally explains why nearly all observed bars rotate
+to operate. Our result naturally explains why nearly all observed bars rotate
 rapidly and is especially relevant for our understanding of how the Milky Way
@@
 Approximately two-thirds of spiral disks host an elongated bar-like feature at
-their center \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our own
-Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of bars
-is not difficult to explain, since stellar disks simulated in isolation almost
-always form bar-like structures \citep{1971ApJ...168..343H}. Several studies
-have shown that a spherical potential, such as a stellar bulge or dark matter
-halo, acts to stabilize a stellar disk against the bar instability
+their centers \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our
+own Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of
+bars is not difficult to explain, since stellar disks simulated in isolation
+almost always form bar-like structures \citep{1971ApJ...168..343H}. Several
+studies have shown that a spherical potential, such as a stellar bulge or dark
+matter halo, acts to stabilize a stellar disk against bar formtaion
 \citep{1973ApJ...186..467O, 1976AJ.....81...30H}.
@@
 measuring the pattern speeds of individual galaxies is the Tremaine-Weinberg
-method \citep{1984ApJ...282L...5T, 2011MSAIS..18...23C}. This method has
+method \citep{1984ApJ...282L...5T, 2011MSAIS..18...23C}. This approach has
 recently been applied to samples of galaxies from the MaNGA survey
-\citep{2019MNRAS.482.1733G, 2020MNRAS.491.3655G}. These recent studies confirm
-what was found in earlier studies -- that nearly all extragalactic bars are fast
+\citep{2019MNRAS.482.1733G, 2020MNRAS.491.3655G}. These studies confirm
+what was found in earlier works, that nearly all extragalactic bars are fast
 rotators (i.e., they rotate close to their maximum rotation rate)
@@
 disk \citep{2013MNRAS.429.1949A, 2014MNRAS.438L..81A}. Since the gas phase
-typically only contributes about $10-20\%$ of the mass of a galaxy at the
+typically contributes only about $10-20\%$ of the mass of a galaxy at the
 present day, one might naively expect it to have a subdominant effect on the
@@
 previous works \citep{1993AA...268...65F, 2010ApJ...719.1470V,
-2014MNRAS.438L..81A}. However, we are the first to exhibit that the behavior
+2014MNRAS.438L..81A}. However, we are the first to demonstrate that the behavior
 ought to be present in a large fraction of galaxies, since it occurs in our
@@
     \includegraphics[width=\columnwidth]{fig/ps_torque.pdf}
-    \caption{The \textit{upper panel} shows the evolution
-    of the pattern speed. As expected, the bar in the \Nbody{} run slows down due
-    to interactions between the bar and the dark matter halo. However, the bar in
-    the SMUGGLE run does not slow down and instead remains at a constant pattern
-    speed. The \textit{lower panel} shows the torque on the bar
-    by different components. The solid lines show the torque exerted by the halo
-    in both the \Nbody{} and SMUGGLE cases. The dashed line shows the torque
-    exerted by the gas phase in the SMUGGLE run (there is no gas in the \Nbody{}
-    run). Details on how these properties are calculated is given in the Methods
-    section.}\label{fig:prop}
+    \caption{The \textit{upper panel} shows the evolution of the pattern speed.
+    As expected, the bar in the \Nbody{} run slows down due to interactions
+    between the bar and the dark matter halo. However, the bar in the SMUGGLE
+    run does not slow down and instead remains at a constant pattern speed. The
+    \textit{lower panel} shows the torque on the bar by different components.
+    The solid lines indicate the torque exerted by the halo in both the \Nbody{}
+    and SMUGGLE cases. The dashed line is the torque exerted by the gas phase in
+    the SMUGGLE run (there is no gas in the \Nbody{} run). Details on how these
+    properties are calculated is given in the Methods section.}\label{fig:prop}
 \end{figure}
@@
 including the gas before the bar has formed disrupts the formation of the bar,
-as has been seen in previous work \citep[e.g.,][]{2013MNRAS.429.1949A}.
-Throughout this work, we consider $t=0$ for the \Nbody{} run to be the time at
-which we added the gas phase for the SMUGGLE run (i.e., we ignore the first
+as has been seen previously \citep[e.g.,][]{2013MNRAS.429.1949A}. Throughout
+this work, we consider $t=0$ for the \Nbody{} run to be the time at which we
+added the gas phase for the SMUGGLE run (i.e., we ignore the first
 $1.5\,\textrm{Gyr}$ of evolution of the \Nbody{} disk when the bar is forming).
@@
 We made one additional modification when stitching the gas disk together with
-the \Nbody{} disk - we created a hole within the central $4\,\textrm{kpc}$. This
-hole guards against an initial dramatic infall of gas within the bar region,
-which we found to destroy the bar. It is not uncommon for observed barred
-galaxies to have gas deficits in the bar region \citep[though not in the very
-center;][]{1993RPPh...56..173S}. Therefore, our practice of allowing the gas
-distribution to have a hole in the central region is consistent with our choice
-to begin the simulations with a bar already formed.
+the \Nbody{} disk - we created a hole within the central $4\,\textrm{kpc}$ of
+the gas disks. This hole guards against an initial dramatic infall of gas within
+the bar region, which we found to destroy the bar. It is not uncommon for
+observed barred galaxies to have gas deficits in the bar region \citep[though
+not in the very center;][]{1993RPPh...56..173S}. Therefore, our practice of
+allowing the gas distribution to have a hole in the central region is consistent
+with our choice to begin the simulations with a bar already formed. In this
+manner, we are able to study the ensuing self-consistent interaction between the
+bar and the gas, but of course we are unable to explore the origin of bars in
+the presence of the gas.
 
@@
 Our setup is initially out of equilibrium, but we found that after about
-$500\,\textrm{Myr}$, the system has settled into a steady-state configuration
-and initial transients appear not to affect the results after this point. The
-constant surface density of the initial gas disk is important for ensuring the
-gas disk is dense enough in order for comparisons to real galaxies to be
-appropriate.
+$500\,\textrm{Myr}$, the system has settled into a roughly steady-state
+configuration and initial transients appear not to affect the results after this
+point. The constant surface density of the initial gas disk is important for
+ensuring that the gas disk is dense enough in order for comparisons to real
+galaxies to be appropriate.
 
@@
 $\epsilon = 0.01$. Star formation converts gas cells into star particles which
-represent single stellar populations with a Chabrier initial mass
-function \citep{2003PASP..115..763C}. For each star particle, the deposition of
-energy, momentum, mass, and metals from stellar winds and supernovae is modeled.
-Photo-ionization and radiation pressure are modeled using an approximate
+represent single stellar populations with a Chabrier initial mass function
+\citep{2003PASP..115..763C}. For each star particle, the deposition of energy,
+momentum, mass, and metals from stellar winds and supernovae is modeled.
+Photo-ionization and radiation pressure are handled using an approximate
 treatment. A more detailed description of this model can be found in the
@@
 We used the fiducial model parameters, except that we increased the number of
-effective neighbors $N_{\textrm{ngb}}$ for the deposition of feedback from
-$64$ to $512$. We found that a lower value of $N_{\textrm{ngb}}$ resulted in
+effective neighbors $N_{\textrm{ngb}}$ for the deposition of feedback from $64$
+to $512$. We found that a lower value of $N_{\textrm{ngb}}$ resulted in
 inefficient photo-ionization feedback since the photo-ionizing budget had not
-been exhausted after deposition into $64$ neighboring cells. We also used an
+been exhausted after deposition into $64$ neighboring cells. We also employed an
 updated version of SMUGGLE using a new mechanical feedback routine similar to
 the one described in \citet{2018MNRAS.480..800H}. This updated routine is a
-tensor renormalization which ensures linear and angular momentum conservation
-to machine precision.
+tensor renormalization which ensures linear and angular momentum conservation to
+machine precision.
 
 In addition to the SMUGGLE model, we considered a simpler model of the
-interstellar medium based upon \citet{2003MNRAS.339..289S}. In this model, the
-multiphase nature of the interstellar medium is modeled in a subgrid manner by
-allowing each resolution element to have a ``cold'' and ``hot'' component, with
-the equation of state of the gas suitably modified. Gas is allowed to
-interchange between the cold and hot components through processes such as
-cooling and stellar feedback. Cold gas is allowed to undergo star formation. We
-refer to this model as the smooth interstellar medium model, and it is described
-in more detail in \citet{2019MNRAS.489.4233M}.
+interstellar medium based upon \citet{2003MNRAS.339..289S}. In this approach,
+the multiphase nature of the interstellar medium is described in a subgrid
+manner by allowing each resolution element to have a ``cold'' and ``hot''
+component, with the equation of state of the gas suitably modified. Gas is
+allowed to interchange between the cold and hot components through processes
+such as cooling and stellar feedback. Cold gas is allowed to undergo star
+formation. We refer to this model as the smooth interstellar medium model, and
+it is described in more detail in \citet{2019MNRAS.489.4233M}.
 
@@
 
-We show the time evolution of different bar properties in Fig.~\ref{fig:prop}.
+We present the time evolution of different bar properties in Fig.~\ref{fig:prop}.
 In the upper panel, we show the pattern speed over time in the \Nbody{} (blue)
@@
 different components. The solid lines indicate the torque on the bar by the dark
-matter halo, whereas the dashed line indicates the torque on the bar by the gas
-phase. In the \Nbody{} case, the halo exerts a steady negative torque on the
-bar, with an average torque from $1$ to $4\,\textrm{Gyr}$ of $-58.0$ in units of
+matter halo, whereas the dashed line is the torque on the bar by the gas phase.
+In the \Nbody{} case, the halo exerts a steady negative torque on the bar, with
+an average torque from $1$ to $4\,\textrm{Gyr}$ of $-58.0$ in units of
 $10^{10}M_{\odot}\,(\textrm{km}/\textrm{s})^2$. The halo in the SMUGGLE case
 exerts a similar negative torque on the bar in the first Gyr of evolution, but
-after that the halo exerts a much smaller torque on the bar, averaging only
+after that the halo exerts a much weaker torque on the bar, averaging only
 $-7.8$ in the same units and over the same time interval. The gas in the SMUGGLE
@@
 on the bar by the dark matter halo. We argue that this can be understood in
-terms of the halo wake mechanism. In the \Nbody{} case, halo material which is
+terms of the halo wake mechanism. In the \Nbody{} case, halo material that is
 resonant with the bar will form a wake which lags behind and exerts a negative
@@
 interstellar medium \citep{2003MNRAS.339..289S}, closer to standard methods used
-in cosmological simulations of galaxy formation. We find that the pattern speed
-evolution is nearly the same in this case, and so conclude that our result is
-not sensitive to the details of the model for the interstellar medium. Details
-on the physics of this test is given in Section.~X.
+in cosmological simulations of galaxy formation and described in more detail in
+Section~\ref{sec:methods}. The result of this test is presented in
+Fig.~\ref{fig:GFM}. We find that the pattern speed evolution is nearly the same
+in this case, and so conclude that our result is not sensitive to the details of
+the model for the interstellar medium.
 
@@
 perhaps bars are recurrent, short-lived phenomena, and that all the bars we see
-in the local universe are very young (CITE). Some have argued that modifications
-to General Relativity ease the tension between the observed universe and
-$\Lambda\textrm{CDM}$ \citep[e.g.][]{2021MNRAS.503.2833R, 2021MNRAS.508..926R}.
+in the local universe are very young \citep{2002AA...392...83B,
+2005MNRAS.364L..18B}. Some have argued that modifications to General Relativity
+ease the tension between the observed universe and $\Lambda\textrm{CDM}$
+\citep[e.g.][]{2021MNRAS.503.2833R, 2021MNRAS.508..926R}.
 
@@
 \citep{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies upon
-the pattern speed of the bar to change with time (though our work does not
-alter expectations for radial mixing induced by spiral
-arms \citep{2002MNRAS.336..785S}).
+the pattern speed of the bar to change with time Our work does not alter
+expectations for radial mixing induced by spiral arms
+\citep{2002MNRAS.336..785S}.
 
@@
 from the gas phase. Only gas within corotation will flow inwards, while gas
-outside corotation will flow outwards\cite{2011MNRAS.415.1027H}. Since the
+outside corotation will flow outwards \citep{2011MNRAS.415.1027H}. Since the
 corotation radius is larger for more slowly rotating bars, it follows that
@@
 We computed the circular velocity curve of our model using the \texttt{AGAMA}
-package\cite{2019MNRAS.482.1525V}. We fit the baryonic component (stellar disk,
+package \citep{2019MNRAS.482.1525V}. We fit the baryonic component (stellar disk,
 bulge, gas, and newly formed stars) with an axisymmetric cylindrical spline with
@@
 plot the circular velocity curve in Fig.~\ref{fig:vcirc} compared to
-observational estimates\cite{2019ApJ...871..120E}. The SMUGGLE disk (which
+observational estimates \citep{2019ApJ...871..120E}. The SMUGGLE disk (which
 includes additional mass in the form of gas) has a slightly higher circular
@@
 the SFR surface density is broadly consistent with the expected values for the
-Milky Way\cite{2008AA...487..951K,2022ApJ...929L..18E}. The discrepancy between
+Milky Way \citep{2008AA...487..951K,2022ApJ...929L..18E}. The discrepancy between
 $1$ and $4\,\textrm{kpc}$ in the molecular and SFR surface density is likely due
@@
 used a simple kinematic distance based on an axisymmetric model of the Milky
-Way\cite{2017ApJ...834...57M}, which is not accurate in the bar region where gas
+Way \citep{2017ApJ...834...57M}, which is not accurate in the bar region where gas
 has large non-circular velocities.

### 2022-08-30 test_part/plots/plot_tot.py
Add tot sam plot

@@ -1,0 +1,138 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import arepo
+import h5py as h5
+import matplotlib as mpl
+from matplotlib import rc
+import glob
+from tqdm import tqdm
+from scipy.signal import savgol_filter
+import re
+import astropy.units as u
+import matplotlib.pylab as pl
+
+mpl.use('Agg')
+
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
+rc('text', usetex=True)
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+
+columnwidth = 244.0 * 0.035145980349999517 # convert to cm
+textwidth = 508.0 * 0.035145980349999517 # convert to cm
+
+snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
+bprop_path = '/n/home01/abeane/starbar/analysis/bar_prop/data/'
+torque_path = '/n/home01/abeane/starbar/analysis/torques/data/'
+
+tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
+        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
+
+
+
+# names
+Nbody = 'Nbody'
+phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
+phS15R35 = 'phantom-vacuum-Sg15-Rc3.5'
+phS10R35 = 'phantom-vacuum-Sg10-Rc3.5'
+phS05R35 = 'phantom-vacuum-Sg05-Rc3.5'
+
+lvl = 'lvl3'
+lvlGFM = 'lvl3-GFM'
+
+def read_Gout(I):
+    fname = '../runs/output-Gvar/out_G'+str(I)+'.npy'
+    t = np.load(fname, allow_pickle=True).tolist()
+    return t
+
+def read_Iout(I):
+    fname = '../runs/output-Ivar/out_I'+str(I)+'.npy'
+    t = np.load(fname, allow_pickle=True).tolist()
+    return t
+
+def fix_bar_angle(bar_angle):
+    out = np.zeros(len(bar_angle))
+    out[0] = bar_angle[0]
+
+    for i in range(1, len(bar_angle)):
+        dphi = bar_angle[i] - bar_angle[i-1]
[…]

### 2022-08-30 paper/ms.tex
Small manuscript changes

@@
 in previous simulations using more simplified models for the interstellar gas
-but it has remained unexplained. We propose that the gas phase of the disk and
-the dark matter halo act in concert to stabilize the bar pattern speed and
-prevent the bar from slowing down or speeding up. We find that in a Milky
-Way-like disk, a gas fraction of only about 5\% is necessary for this mechanism
-to operate. Our result naturally explains why nearly all observed bars rotate
-rapidly and is especially relevant for our understanding of how the Milky Way
-arrived at its present state.
+but the apparent lack of secular evolution has remained unexplained. We propose
+that the gas phase of the disk and the dark matter halo act in concert to
+stabilize the bar pattern speed and prevent the bar from slowing down or
+speeding up. We find that in a Milky Way-like disk, a gas fraction of only about
+5\% is necessary for this mechanism to operate. Our result naturally explains
+why nearly all observed bars rotate rapidly and is especially relevant for our
+understanding of how the Milky Way arrived at its present state.
 \end{abstract}
@@
 studies have shown that a spherical potential, such as a stellar bulge or dark
-matter halo, acts to stabilize a stellar disk against bar formtaion
-\citep{1973ApJ...186..467O, 1976AJ.....81...30H}.
+matter halo, acts to stabilize stellar disks against bar formtaion
+\citep[e.g.,][]{1973ApJ...186..467O, 1976AJ.....81...30H}.
 

### 2022-08-30 paper/ms.tex
Add sam section

@@
 \newcommand{\Rb}{\ensuremath{R_{\textrm{b}}}}
+\newcommand{\kms}{\ensuremath{\textrm{km}/\textrm{s}}}
 
@@
     the SMUGGLE run (there is no gas in the \Nbody{} run). Details on how these
-    properties are calculated is given in the Methods section.}\label{fig:prop}
+    properties are calculated is given in Section~\ref{ssec:bar_analysis}.}\label{fig:prop}
 \end{figure}
@@
 \subsection{Bar Analysis}
+\label{ssec:bar_analysis}
 The analysis of various bar properties is performed as follows. First, the
@@
     \centering
-    \includegraphics[width=18cm]{fig/halo_wake.pdf}
+    \includegraphics[width=\textwidth]{fig/halo_wake.pdf}
     \caption{The wake excited in the dark matter halo. The dark matter halo wake
@@
 
+\subsection{Semi-Analytic Model}
+We also developed a simple semi-analytic model of a bar-disk-halo system. This
+exercise demonstrates that our proposed mechanism follows from a few simplistic
+assumptions. Our method follows closely the one developed in
+\citet{2022MNRAS.513..768C}. We model the bar-disk-halo system with three
+components: a dark matter Hernquist halo \citep{1990ApJ...356..359H}, a
+Miyamoto-Nagai disk \citep{1975PASJ...27..533M}, and a pure quadrupole bar
+described in \citet{2022MNRAS.513..768C}. The bar and disk components are just
+described by their potential, but we integrate the trajectories of test
+particles drawn from a Hernquist halo. Note that we do not include the
+interactions between these test particles, and so this model is not
+self-consistent. We give our chosen parameter values in Appendix~\ref{app:sam}.
+
+We allow the bar in this model to rotate as a solid body. However, we crucially
+allow the pattern speed of the bar to freely change with time in accordance with
+the torque exerted on the dark matter halo by the bar. In particular, we
+subtract the $z$-component of this torque divided by the moment of inertia of
+the bar from the pattern speed. Since the radius of corotation \RCR{} is a
+parameter in the bar model from \citet{2022MNRAS.513..768C}, we allow the moment
+of inertia of the bar to vary with ${\RCR}^2$. To be more precise, we allow
+\begin{equation}
+I = \frac{I_6}{10^{10} \Msun (\kms)^2} \left( \frac{\RCR}{6\,\textrm{kpc}} \right)^2\textrm{,}
+\end{equation}
+where $I_6$ is a free parameter chosen by the user. We found that allowing $I_6=8$ is a good approximation to our bar model.
+
+In addition to $I_6$, we allowed for another free parameter - the torque from
+the gas phase on the bar, $\tau_{\textrm{gas}}$. This torque is applied to the
+bar in the same way as the torque from the halo is applied. The torque is given
+in code units ($10^{10}\Msun \left(\kms\right)^2$). We allow this torque to
+mildly vary with pattern speed according to the formula,
+\begin{equation}
+\tau = \tau_{\textrm{gas}} + (40\,\kms/\textrm{kpc} - \Omega_p)\textrm{.}
+\end{equation}
+We are motivated to do this because of our result in Appendix~\ref{app:varyps}
+which shows that more slowly rotating bars have higher gas infall rates and thus
+larger positive torques. In the $\tau_{\textrm{gas}}=0$ case, we do not allow
+the gas torque to vary with pattern speed.
+
+We show the effect of varying the gas torque $\tau_{\textrm{gas}}$ from $0$ to
+$20$ in increments of $2$ in Fig.~\ref{fig:sam}. The solid lines indicate the
+semi-analytic model, while the two dashed lines correspond to our fiducial
+simulations introduced earlier. For reference, the average torque exerted by the
+gas phase on the bar in our fiducial simulation was $11.7$ in code units. We see in Fig.~\ref{fig:sam} that we can reproduce the stability of our fiducial
+gas disk simply by including a positive torque on the order of $6$. Our
+semi-analytic model with no gas torque can reproduce the pattern speed evolution
+of the $N$-body case.
+
+We next take the $\tau_{\textrm{gas}}=0$ and $20$ cases from Fig.~\ref{fig:sam}
+and plot the torque evolution. This result, given in Fig.~\ref{fig:sam-torque},
+is comparable to the lower panel of Fig.~\ref{fig:prop}. We find that the
+$\tau_{\textrm{gas}}=0$ case compares favorably to the \Nbody{} case described
+in previous sections. The bar exerts a steady negative torque in this case (blue
+line). When a gas torque is included (orange lines), we find that the halo's
+torque becomes much weaker, similar to what we found in the SMUGGLE case. The
+gas applies a steady positive torque by construction. The exhaustion process is
+reproduced in our simple semi-analytic model.
+
 \subsection{Observations}
@@
 \label{fig:GFM}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{fig/samGvar.pdf}
+    \caption{A comparison between the pattern speeds of our fiducial disk
+    systems and a semi-analytic model. The solid lines indicate the pattern
+    speeds assuming a constant positive torque, varying in increments of $2$
+    from $0$ to $20$. The dashed lines indicate the pattern speed evolution from
+    our fully self-consistent simulations from earlier. We find excellent
+    agreement between our fiducial simulations and our semi-analytic model of a
+    bar-disk-halo system. Torques are given in code units ($10^{10}\Msun
+    \left(\kms\right)^2$).}
+    \label{fig:sam}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{fig/sam_torque.pdf}
+    \caption{The torque exerted on the bar by various components in our
+    semi-analytic model. Solid lines indicate the torque by the halo while the
+    dashed line indicates the torque exerted by the gas phase. We chose two
+    models with $\tau_{\textrm{gas}}=0$ and $20$, which most closely resemble
+    our $N$-body and SMUGGLE disks, respectively. This figure ought to be
+    compared to the lower panel of Fig.~\ref{fig:prop}. Overall, we find good
+    qualitative agreement.}
+    \label{fig:sam-torque}
 \end{figure}
@@
     stars form a roughly axisymmetric structure indicates our bar decomposition
-    is sufficiently accurate.}
+    is sufficiently accurate. We have computed that $76\%$ of the second Fourier
+    component resides in the stars classified as being trapped in the bar.}
     \label{fig:decomp}
@@
 \end{equation}
-These criterion are slightly different and simplified from the ones used in \citet{2016MNRAS.463.1952P}, but we found them to empirically work well at
-decomposing the disk into a bar and disk component. In Fig.~\ref{fig:decomp}, we show an example of this decomposition. The
-\textit{left} panel shows a surface density projection of the stellar disk and
-bulge (including newly formed stars) from the SMUGGLE model after
-$1\,\text{Gyr}$ of evolution in a frame such that the bar is aligned with the
-$x$-axis. The \textit{middle} panel shows a projection of the subset of stars
-that are identified as being trapped in the bar and the \textit{right} panel
-shows a projection of the stars that are not identified as being trapped. The
-fact that the \textit{right} panel is roughly axisymmetric indicates the bar
-decomposition is performing adequately.
+These criterion are slightly different and simplified from the ones used in
+\citet{2016MNRAS.463.1952P}, but we found them to empirically work well at
+decomposing the disk into a bar and disk component. In Fig.~\ref{fig:decomp}, we
+show an example of this decomposition. The \textit{left} panel shows a surface
+density projection of the stellar disk and bulge (including newly formed stars)
+from the SMUGGLE model after $1\,\text{Gyr}$ of evolution in a frame such that
+the bar is aligned with the $x$-axis. The \textit{middle} panel shows a
+projection of the subset of stars that are identified as being trapped in the
+bar and the \textit{right} panel shows a projection of the stars that are not
+identified as being trapped. The fact that the \textit{right} panel is roughly
+axisymmetric indicates the bar decomposition is performing adequately.
+
+We computed the second Fourier component $A_2$ for all particles classified as
+barred and unbarred. We found that $76\%$ of the total $m=2$ Fourier component
+is in the particles classified as barred (i.e.,
+$A_{2,\textrm{bar}}/A_{2,\textrm{tot}}\sim0.76$). This indicates good but not
+perfect classification. Some of this is probably coming from the $m=4$ component
+being classified as unbarred, given that the unbarred component has a somewhat
+boxy shape.
 
@@
 \end{figure}
+
+\section{Semi-Analytic Model Parameters}
+\label{app:sam}
+Our semi-analytic model consisted of a three-component bar-disk-halo system. We
+describe here the parameters we chose for these components. The parameters of
+the disk and halo were chosen to match closely what we used in our fiducial
+simulations. The system can thus be understood as being roughly similar to the
+Milky Way, though no careful analysis has been performed to ensure the closest
+match possible.
+
+For the dark matter halo, we used a Hernquist potential
+\citep{1990ApJ...356..359H} with mass $10^{12}\,\Msun$ and a scale length of
+$26.2\,\textrm{kpc}$. For the stellar disk, we used a MiyaMoto-Nagai disk
+\citep{1975PASJ...27..533M} with mass $4.8\times10^{10}\,\Msun$, radial scale
+length of $2.67\,\textrm{kpc}$, and vertical scale length of
+$0.32\,\textrm{kpc}$. For the bar, we used the quadrupole potential described in
+\citet{2022MNRAS.513..768C}. We used their fiducial parameter values --
+specifically, we set $A=0.02$, $b=0.28$, and $v_c = 235\,\kms$. Our initial
+pattern speed is always set to $40\,\kms/\textrm{kpc}$.
 

### 2022-08-30 paper/ms.tex
Update figure placements
Small updates

@@
 remains at a stable, constant rate of rotation. This behavior has been observed
-in previous simulations using more simplified models for the interstellar gas
+in previous simulations using more simplified models for the interstellar gas,
 but the apparent lack of secular evolution has remained unexplained. We propose
@@
 We have performed a simulation of a disk galaxy using the finite-volume,
-gravito-hydrodynamics code AREPO \citep{2010MNRAS.401..791S}. We use the galaxy
+gravito-hydrodynamics code \AREPO{} \citep{2010MNRAS.401..791S}. We use the galaxy
 formation model Stars and MUltiphase Gas in GaLaxiEs
@@
 
-\section{Results}
-\label{sec:results}
-
-We present the time evolution of different bar properties in Fig.~\ref{fig:prop}.
-In the upper panel, we show the pattern speed over time in the \Nbody{} (blue)
-and SMUGGLE (orange) runs. The pattern speed in the \Nbody{} case slows down
-while the pattern speed in the SMUGGLE case remains roughly constant. The
-slowing down of the pattern speed in the \Nbody{} case is consistent with a long
-line of numerical research on bars in \Nbody{} simulations
-\citep{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
-2002ApJ...569L..83A, 2003MNRAS.341.1179A, 2003MNRAS.346..251O,
-2005MNRAS.363..991H, 2006ApJ...637..214M, 2007MNRAS.375..460W,
-2009ApJ...697..293D}.
-
-However, in the SMUGGLE case the pattern speed remains constant. After the first
-Gyr of evolution, we find that the pattern speed increases by only $\sim10\%$
-over the next $4\,\textrm{Gyr}$, compared to a $\sim43\%$ decrease in the
-pattern speed for the \Nbody{} run over the same interval.
-
-The bottom panel of Fig.~\ref{fig:prop} shows the torque exerted on the bar by
-different components. The solid lines indicate the torque on the bar by the dark
-matter halo, whereas the dashed line is the torque on the bar by the gas phase.
-In the \Nbody{} case, the halo exerts a steady negative torque on the bar, with
-an average torque from $1$ to $4\,\textrm{Gyr}$ of $-58.0$ in units of
-$10^{10}M_{\odot}\,(\textrm{km}/\textrm{s})^2$. The halo in the SMUGGLE case
-exerts a similar negative torque on the bar in the first Gyr of evolution, but
-after that the halo exerts a much weaker torque on the bar, averaging only
-$-7.8$ in the same units and over the same time interval. The gas in the SMUGGLE
-case exerts a steady positive torque averaging $11.7$ over $1\,\textrm{Gyr}$ in
-the same units.
-
-As we saw qualitatively in Fig.~\ref{fig:overview}, the middle panel of
-Fig.~\ref{fig:strength} shows that the length of the bar in the \Nbody{} case
-grows over time while it remains roughly constant in the SMUGGLE case. This is
-also consistent with previous numerical work, which found that bars tend to grow
-as they slow down and the radius of corotation increases
-\citep{2000ApJ...543..704D, 2003MNRAS.341.1179A}. The time evolution of the bar
-strength, defined as the maximum of $\left|A_2/A_0\right|$ as a function of
-radius, is shown in the lower panel Fig.~\ref{fig:strength}. The quantity
-$\left|A_2/A_0\right|$ varies from $0$ to $1$, with larger values indicating a
-stronger bar pattern. We see that in the \Nbody{} case, $\left|A_2/A_0\right|$
-increases over time as the bar pattern slows. This is consistent with previous
-\Nbody{} simulations which showed a clear correlation between the bar pattern
-speed and the bar strength \citep[e.g.,][]{2003MNRAS.341.1179A}. In the SMUGGLE
-case, we see that the bar strength has an initial drop but then remains at a
-roughly constant, but slightly decreasing, strength. This is consistent with the
-pattern speed in the SMUGGLE case being roughly constant or slightly increasing.
-
 \begin{figure*}
@@
 \end{figure*}
+
+\section{Results}
+\label{sec:results}
+
+We present the time evolution of different bar properties in Fig.~\ref{fig:prop}.
+In the upper panel, we show the pattern speed over time in the \Nbody{} (blue)
+and SMUGGLE (orange) runs. The pattern speed in the \Nbody{} case slows down
+while the pattern speed in the SMUGGLE case remains roughly constant. The
+slowing down of the pattern speed in the \Nbody{} case is consistent with a long
+line of numerical research on bars in \Nbody{} simulations
+\citep{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
+2002ApJ...569L..83A, 2003MNRAS.341.1179A, 2003MNRAS.346..251O,
+2005MNRAS.363..991H, 2006ApJ...637..214M, 2007MNRAS.375..460W,
+2009ApJ...697..293D}.
+
+However, in the SMUGGLE case the pattern speed remains constant. After the first
+Gyr of evolution, we find that the pattern speed increases by only $\sim10\%$
+over the next $4\,\textrm{Gyr}$, compared to a $\sim43\%$ decrease in the
+pattern speed for the \Nbody{} run over the same interval.
+
+The bottom panel of Fig.~\ref{fig:prop} shows the torque exerted on the bar by
+different components. The solid lines indicate the torque on the bar by the dark
+matter halo, whereas the dashed line is the torque on the bar by the gas phase.
+In the \Nbody{} case, the halo exerts a steady negative torque on the bar, with
+an average torque from $1$ to $4\,\textrm{Gyr}$ of $-58.0$ in units of
+$10^{10}M_{\odot}\,(\textrm{km}/\textrm{s})^2$. The halo in the SMUGGLE case
+exerts a similar negative torque on the bar in the first Gyr of evolution, but
+after that the halo exerts a much weaker torque on the bar, averaging only
+$-7.8$ in the same units and over the same time interval. The gas in the SMUGGLE
+case exerts a steady positive torque averaging $11.7$ over $1\,\textrm{Gyr}$ in
+the same units.
+
+As we saw qualitatively in Fig.~\ref{fig:overview}, the upper panel of
+Fig.~\ref{fig:strength} shows that the length of the bar in the \Nbody{} case
+grows over time while it remains roughly constant in the SMUGGLE case. This is
+also consistent with previous numerical work, which found that bars tend to grow
+as they slow down and the radius of corotation increases
+\citep{2000ApJ...543..704D, 2003MNRAS.341.1179A}. The time evolution of the bar
+strength, defined as the maximum of $\left|A_2/A_0\right|$ as a function of
+radius, is shown in the lower panel of Fig.~\ref{fig:strength}. The quantity
+$\left|A_2/A_0\right|$ varies from $0$ to $1$, with larger values indicating a
+stronger bar pattern. We see that in the \Nbody{} case, $\left|A_2/A_0\right|$
+increases over time as the bar pattern slows. This is consistent with previous
+\Nbody{} simulations which showed a clear correlation between the bar pattern
+speed and the bar strength \citep[e.g.,][]{2003MNRAS.341.1179A}. In the SMUGGLE
+case, we see that the bar strength has an initial drop but then remains at a
+roughly constant, but slightly decreasing, strength. This is consistent with the
+pattern speed in the SMUGGLE case being roughly constant or slightly increasing.
 
@@
 the wake, arresting the process by which the halo can slow the bar down. We term
-this process ``exhaustion.''
+this process ``halo exhaustion.''
 
@@
 quickly enough due to the positive torque of the infalling gas. This causes
-the resonant halo phase space at a particular pattern speed, $\Omega_{p,0}$,
+the resonant halo phase space at a particular pattern speed, $\Omega_{\textrm{p},0}$,
 to become exhausted and no longer able to support a negative torque. Second,
@@
 negative torque on the bar. The pattern speed will then settle at a new value
-slightly higher than $\Omega_{p,0}$ where the gas and halo torques cancel.
+slightly higher than $\Omega_{\textrm{p},0}$ where the gas and halo torques cancel.
 Over time, the pattern speed should slowly increase as the halo becomes
@@
 \subsection{Varying Initial Gas Fractions}
-We also performed a test in which we varied the initial gas fraction of the
+We performed a test in which we varied the initial gas fraction of the
 disk. In our fiducial run, we set the surface density of the gas disk from
@@
 \subsection{Smooth Interstellar Medium}
-We also performed a simulation of the same disk but with a simpler model of the
+We performed a simulation of the same disk but with a simpler model of the
 interstellar medium \citep{2003MNRAS.339..289S}, closer to standard methods used
@@
 the model for the interstellar medium.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=9cm]{fig/ps_late_start.pdf}
+    \caption{Pattern speed evolution with a lower initial pattern speed. We
+    tested the evolution of our system when gas is added to the \Nbody{} run at
+    a later time, but with all other simulation parameters kept the same. The
+    setup is therefore identical to our previous runs just with a bar that is
+    larger, stronger, and with a lower pattern speed. We find that the pattern
+    speed evolution is very similar to our fiducial case, except that the bar
+    retains its original pattern speed. This indicates a mechanism which keeps
+    the bar at its formation pattern speed, and that there is not a particular
+    pattern speed which the system tends to.}
+    \label{fig:snap700}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=9cm]{fig/ps_fgas.pdf}
+    \caption{Pattern speed evolution with varying gas fractions. We explored the
+    impact of lowering the initial gas surface density of our fiducial disk on
+    the evolution of the pattern speed. The surface densities we tested of $20$,
+    $15$, $10$, and $5\,\Msun/\textrm{kpc}^2$ correspond to initial gas
+    fractions of $16\%$, $10\%$, $7\%$, and $4\%$. We find that initial surface
+    densities $20$, $15$, and $10\,\Msun/\textrm{kpc}^2$ result in bars which
+    remain at a constant pattern speed, while an initial surface density of
+    $5\,\Msun/\textrm{kpc}^2$ results in a bar which slows down in roughly the
+    same manner as the \Nbody{} case.}
+    \label{fig:fgas}
+\end{figure}
 
@@
 subtract the $z$-component of this torque divided by the moment of inertia of
-the bar from the pattern speed. Since the radius of corotation \RCR{} is a
-parameter in the bar model from \citet{2022MNRAS.513..768C}, we allow the moment
-of inertia of the bar to vary with ${\RCR}^2$. To be more precise, we allow
+the bar from the pattern speed at each timestep. Since the radius of corotation
+\RCR{} is a parameter in the bar model from \citet{2022MNRAS.513..768C}, we
+allow the moment of inertia of the bar to vary with ${\RCR}^2$. To be more
+precise, we allow
 \begin{equation}
@@
 \end{equation}
-where $I_6$ is a free parameter chosen by the user. We found that allowing $I_6=8$ is a good approximation to our bar model.
+where $I_6$ is a free parameter chosen by the user. We found that allowing
+$I_6=8$ is a good approximation to our fiducial disk model.
 
@@
 \begin{equation}
-\tau = \tau_{\textrm{gas}} + (40\,\kms/\textrm{kpc} - \Omega_p)\textrm{.}
+\tau = \tau_{\textrm{gas}} + (40 - \Omega_{\textrm{p}})\textrm{,}
 \end{equation}
-We are motivated to do this because of our result in Appendix~\ref{app:varyps}
-which shows that more slowly rotating bars have higher gas infall rates and thus
-larger positive torques. In the $\tau_{\textrm{gas}}=0$ case, we do not allow
-the gas torque to vary with pattern speed.
+where $\Omega_{\textrm{p}}$ is given in $\kms/\textrm{kpc}$. We are motivated to
+do this because of our result in Appendix~\ref{app:varyps} which shows that more
+slowly rotating bars have higher gas infall rates and thus larger positive
+torques. In the $\tau_{\textrm{gas}}=0$ case, we do not allow the gas torque to
+vary with pattern speed.
 
@@
 simulations introduced earlier. For reference, the average torque exerted by the
-gas phase on the bar in our fiducial simulation was $11.7$ in code units. We see in Fig.~\ref{fig:sam} that we can reproduce the stability of our fiducial
-gas disk simply by including a positive torque on the order of $6$. Our
-semi-analytic model with no gas torque can reproduce the pattern speed evolution
-of the $N$-body case.
+gas phase on the bar in our fiducial simulation was $11.7$ in code units. We see
+in Fig.~\ref{fig:sam} that we can reproduce the stability of our fiducial gas
+disk (i.e., its lack of secular evolution) simply by including a positive torque
+on the order of $6$. Our semi-analytic model with no gas torque can reproduce
+the pattern speed evolution of the $N$-body case.
 
@@
 torque becomes much weaker, similar to what we found in the SMUGGLE case. The
-gas applies a steady positive torque by construction. The exhaustion process is
-reproduced in our simple semi-analytic model.
+gas applies a steady positive torque by construction. Therefore, the exhaustion
+process is reproduced in our simple semi-analytic model.
 
@@
 measurement of pattern speeds is a delicate process still prone to large errors.
-\begin{figure}
-    \centering
-    \includegraphics[width=9cm]{fig/ps_late_start.pdf}
-    \caption{Pattern speed evolution with a lower initial pattern speed. We
-    tested the evolution of our system when gas is added to the \Nbody{} run at
-    a later time, but with all other simulation parameters kept the same. The
-    setup is therefore identical to our previous runs just with a bar that is
-    larger, stronger, and with a lower pattern speed. We find that the pattern
-    speed evolution is very similar to our fiducial case, except that the bar
-    retains its original pattern speed. This indicates a mechanism which keeps
-    the bar at its formation pattern speed, and that there is not a particular
-    pattern speed which the system tends to.}
-    \label{fig:snap700}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=9cm]{fig/ps_fgas.pdf}
-    \caption{Pattern speed evolution with varying gas fractions. We explored the
-    impact of lowering the initial gas surface density of our fiducial disk on
-    the evolution of the pattern speed. The surface densities we tested of $20$,
-    $15$, $10$, and $5\,\Msun/\textrm{kpc}^2$ correspond to initial gas
-    fractions of $16\%$, $10\%$, $7\%$, and $4\%$. We find that initial surface
-    densities $20$, $15$, and $10\,\Msun/\textrm{kpc}^2$ result in bars which
-    remain at a constant pattern speed, while an initial surface density of
-    $5\,\Msun/\textrm{kpc}^2$ results in a bar which slows down in roughly the
-    same manner as the \Nbody{} case.}
-    \label{fig:fgas}
-\end{figure}
 
@@
 Naturally, extending our present work to account for such effects is a crucial
-next step in understanding the formation and evolution of galactic bars.
+next step in understanding the formation and evolution of galactic bars. We are
+presently engaged in such an exploration.
 
@@
 \label{sec:conclusions}
+We performed a simulation of a Milky Way-like galactic disk hosting a strong bar
+with a state-of-the-art model for the interstellar medium. We found that the
+pattern speed of the bar in this simulation does not slow down but rather
+remains at a stable, constant pattern speed. We provided a simple semi-analytic
+model which reproduces many of the features from our fiducial disk model.
+
 The implications of our findings are numerous. First, we naturally explain why
@@
 
-\begin{figure}[h!]
+\begin{figure}
     \centering
@@
 pattern speed is always set to $40\,\kms/\textrm{kpc}$.
+
+We integrated our model for $5\,\textrm{Gyr}$ with a timestep of
+$0.01\,\textrm{Gyr}$.
 

### 2022-09-06 paper/ms.tex
Add Federico comments

@@
 We also developed a simple semi-analytic model of a bar-disk-halo system. This
-exercise demonstrates that our proposed mechanism follows from a few simplistic
+exercise demonstrates that our proposed mechanism follows from a few simple
 assumptions. Our method follows closely the one developed in
 \citet{2022MNRAS.513..768C}. We model the bar-disk-halo system with three
-components: a dark matter Hernquist halo \citep{1990ApJ...356..359H}, a
-Miyamoto-Nagai disk \citep{1975PASJ...27..533M}, and a pure quadrupole bar
+components: a dark matter \citet{1990ApJ...356..359H} halo, a
+\citet{1975PASJ...27..533M} disk, and a pure quadrupole bar
 described in \citet{2022MNRAS.513..768C}. The bar and disk components are just
@@
 where $I_6$ is a free parameter chosen by the user. We found that allowing
-$I_6=8$ is a good approximation to our fiducial disk model.
+$I_6=8$ is a good approximation to our fiducial disk model. In code units, the
+moment of inertia of the SMUGGLE bar (i.e., the particles classified as being in
+the bar) is about $2$. This is a factor of $4$ smaller than our fiducial value
+of $I_6=8$, but this is probably due to the fact that the bar does not really
+rotate as a solid body.
 

### 2022-09-07 paper/ms.tex
Add software citations

@@
 \usepackage{newtxtext,newtxmath}
+\usepackage{hyperref}
 
@@
 \newcommand{\Msun}{\ensuremath{M_{\odot}}}
+\newcommand{\Myr}{\ensuremath{\textrm{Myr}}}
 
@@
 spiral galaxies should rotate slowly.
-
+https://github.com/gusbeane/gasbar
 \section*{Acknowledgements}
@@
 \section*{Data Availability}
-
-The inclusion of a Data Availability Statement is a requirement for articles published in MNRAS.
-Data Availability Statements provide a standardised format for readers to understand the availability of data underlying the research results described in the article.
-The statement may refer to original data generated in the course of the study or to third-party data analysed in the article.
-The statement should describe and provide means of access, where possible, by linking to the data or providing the required accession numbers for the relevant databases or DOIs.
+A git repository containing all analysis scripts and notebooks is available at
+\url{https://github.com/gusbeane/gasbar}. Snapshots at a $500\,\Myr$ cadence, as
+well as the initial conditions, are available at . Snapshots at a higher cadence
+are available upon request without expectation of coauthorship.
+
+We made use of the following software packages in {\sc Python} \citep{Python}:
+{\sc agama} \url{https://github.com/GalacticDynamics-Oxford/Agama}, {\sc
+astropy} \citep{astropy:2013, astropy:2018}, {\sc h5py}
+\url{http://www.h5py.org/}, {\sc inspector\_gadget}
+\url{https://bitbucket.org/abauer/inspector_gadget/}, {\sc joblib}
+\url{https://joblib.readthedocs.io/en/latest/}, {\sc matplotlib}
+\citep{Hunter:2007}, {\sc numba} \citep{Lam:2015bfh}, {\sc numpy}
+\citep{harris2020array}, {\sc scipy} \citep{2020SciPy-NMeth}, {\sc tqdm}
+\url{https://tqdm.github.io/}.
 

### 2022-09-07 analysis/bar_orbits/job_bar_orbits.py
update pdf

@@ -1,0 +1,25 @@
+#!/bin/sh
+##SBATCH -p hernquist,conroy,shared,itc_cluster
+#SBATCH -p hernquist,conroy
+##SBATCH --constraint=intel
+#SBATCH -J barorb 
+#SBATCH -n 16
+#SBATCH -N 1
+#SBATCH -o OUTPUT_frames.%j.out
+#SBATCH -e ERROR_frames.%j.err
+##SBATCH --exclusive
+#SBATCH --mail-user=angus.beane@cfa.harvard.edu
+#SBATCH --mail-type=BEGIN
+#SBATCH --mail-type=END
+#SBATCH --mail-type=FAIL
+##SBATCH --mem-per-cpu=3900
+#SBATCH --mem=240G
+##SBATCH -t 4-00:00           # Runtime in D-HH:MM
+#SBATCH -t 7-00:00           # Runtime in D-HH:MM
+
+source ../load-modules.sh
+
+ulimit -c unlimited
+
+python3 compute_bar_orbits.py ${SLURM_NTASKS} $1
+

### 2022-09-07 analysis/fourier_component/job_fourier-lvl2.sh

@@ -1,0 +1,22 @@
+#!/bin/sh
+#SBATCH -p bigmem
+#SBATCH -J fourier 
+#SBATCH -n 48
+#SBATCH -N 1
+#SBATCH -o OUTPUT_frames.%j.out
+#SBATCH -e ERROR_frames.%j.err
+##SBATCH --exclusive
+#SBATCH --mail-user=angus.beane@cfa.harvard.edu
+#SBATCH --mail-type=BEGIN
+#SBATCH --mail-type=END
+#SBATCH --mail-type=FAIL
+#SBATCH --mem=480G
+##SBATCH -t 4-00:00           # Runtime in D-HH:MM
+#SBATCH -t 7-00:00           # Runtime in D-HH:MM
+
+source ../load-modules.sh
+
+ulimit -c unlimited
+
+python3 compute_fourier_component.py ${SLURM_NTASKS} $1
+

### 2022-09-07 analysis/torques/compute_torques-old.py

@@ -1,0 +1,206 @@
+import numpy as np
+import arepo
+import sys
+from tqdm import tqdm
+import astropy.units as u
+import h5py as h5
+import glob
+import os
+from numba import njit
+
+from pyMND.forcetree import construct_tree, force_treeevaluate_loop
+
+from joblib import Parallel, delayed
+
+def compute_acc(pos, tree_pos, tree_mass, theta, soft, num_threads=1, maxnode_fac=2.):
+    tree = construct_tree(tree_pos, tree_mass, theta, soft, maxnode_fac=maxnode_fac)
+    acc = force_treeevaluate_loop(pos, tree, num_threads=num_threads)
+    return acc
+
+    
+    Am_real = np.zeros(nbins)
+    Am_imag = np.zeros(nbins)
+    Rmag = np.zeros(nbins)
+    N_in_bin = np.zeros(nbins)
+    
+    Npart = len(pos)
+    for i in range(Npart):
+        R = np.sqrt(pos[i][0]*pos[i][0] + pos[i][1]*pos[i][1])
+        phi = np.arctan2(pos[i][1], pos[i][0])
+        
+        for j in range(nbins):
+            if R > bins[j] and R < bins[j+1]:
+                Am_real[j] += mass[j]*np.cos(m*phi)
+                Am_imag[j] += mass[j]*np.sin(m*phi)
+                Rmag[j] += R
+                N_in_bin[j] += 1
+    
+    
+    for j in range(nbins):
+        if N_in_bin[j] > 0:
+            Rmag[j] /= N_in_bin[j]
+        else:
+            Rmag[j] = np.nan
+
+    return Rmag, Am_real, Am_imag
+
+def get_cyl(sn, pt, center=np.array([0., 0., 0.])):
+    part = getattr(sn, 'part'+str(pt))
+    pos = part.pos.value - center
+    vel = part.vel.value
+    
+    R = np.linalg.norm(pos[:,:2], axis=1)
+    phi = np.arctan2(pos[:,1], pos[:,0])
+    z = pos[:,2]
+    
+    cosphi = pos[:,0]/R
+    sinphi = pos[:,1]/R
+    
+    vR = cosphi * vel[:,0] + sinphi * vel[:,1]
+    vphi = cosphi * vel[:,1] - sinphi * vel[:,0]
+    vz = vel[:,2]
+    
+    return np.transpose([R, phi, z, vR, vphi, vz])
+
[…]

### 2022-09-07 plots/Rb_A2/plot_Rb_A2.py

@@ -1,0 +1,127 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import arepo
+import h5py as h5
+import matplotlib as mpl
+from matplotlib import rc
+import glob
+from tqdm import tqdm
+from scipy.signal import savgol_filter
+
+time_conv = 977.79222168
+
+columnwidth = 244.0 * 0.035145980349999517 # convert to cm
+textwidth = 508.0 * 0.035145980349999517 # convert to cm
+
+mpl.use('Agg')
+
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
+rc('text', usetex=True)
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+
+basepath = '/n/holylfs05/LABS/hernquist_lab/Users/abeane/gasbar/'
+
+snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
+fourier_path = basepath + '/analysis/fourier_component/data/'
+bprop_path = basepath + '/analysis/bar_prop/data/'
+
+tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
+        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
+
+# names
+Nbody = 'Nbody'
+phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
+
+lvl = 'lvl3'
+
+def read_fourier(name, lvl):
+    t = h5.File(fourier_path + 'fourier_' + name + '-' + lvl + '.hdf5', mode='r')
+    return t
+
+def extract_t_max_A2A0(fourier, debug=False):
+
+    i = 0
+
+    tlist = np.array(fourier['time'])
+    A2A0list = []
+    while 'snapshot_'+"{:03d}".format(i) in fourier.keys():
+        key = 'snapshot_'+"{:03d}".format(i)
+        A0 = np.array(fourier[key]['A0'])
+        A2r = np.array(fourier[key]['A2r'])
+        A2i = np.array(fourier[key]['A2i'])
+    
+        A2 = np.sqrt(A2r*A2r + A2i*A2i)
+        A2A0list.append(np.max(A2/A0))
+        if debug:
[…]

### 2022-09-07 plots/ps_GFM/plot_ps_GFM.py

@@ -1,0 +1,162 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import arepo
+import h5py as h5
+import matplotlib as mpl
+from matplotlib import rc
+import glob
+from tqdm import tqdm
+from scipy.signal import savgol_filter
+import re
+import astropy.units as u
+
+mpl.use('Agg')
+
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
+rc('text', usetex=True)
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+
+columnwidth = 244.0 * 0.035145980349999517 # convert to cm
+textwidth = 508.0 * 0.035145980349999517 # convert to cm
+
+snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
+bprop_path = '/n/home01/abeane/starbar/analysis/bar_prop/data/'
+torque_path = '/n/home01/abeane/starbar/analysis/torques/data/'
+
+tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
+        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
+
+# names
+Nbody = 'Nbody'
+phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
+phS15R35 = 'phantom-vacuum-Sg15-Rc3.5'
+phS10R35 = 'phantom-vacuum-Sg10-Rc3.5'
+phS05R35 = 'phantom-vacuum-Sg05-Rc3.5'
+
+lvl = 'lvl3'
+lvlGFM = 'lvl3-GFM'
+
+def read_fourier(name, lvl, basepath='/n/home01/abeane/starbar/analysis/'):
+    f = h5.File(basepath+'/fourier_component/data/fourier_'+name+'-'+lvl+'.hdf5', mode='r')
+    return f
+
+def read_snap(idx, name, lvl, parttype=[0, 2, 3, 4], fields=['Coordinates', 'Masses']):
+    sn = arepo.Snapshot(snap_path+name+'/'+lvl+'/output', idx, combineFiles=True, parttype=parttype,
+                        fields=fields)
+    return sn
+
+def get_sorted_keys(dat):
+    keys = list(dat.keys())
+    # only keep keys that are snapshot keys
[…]

### 2022-09-07 plots/rot_param/plot_rot_param.py

@@ -1,0 +1,152 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import arepo
+import h5py as h5
+import matplotlib as mpl
+from matplotlib import rc
+import glob
+from tqdm import tqdm
+from scipy.signal import savgol_filter
+from scipy.optimize import minimize
+import agama
+
+agama.setUnits(mass=1E10, length=1, velocity=1)
+mpl.use('Agg')
+
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
+rc('text', usetex=True)
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+
+snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
+bprop_path = '/n/home01/abeane/starbar/analysis/bar_prop/data/'
+torque_path = '/n/home01/abeane/starbar/analysis/torques/data/'
+
+tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
+        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
+
+# names
+Nbody = 'Nbody'
+phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
+
+lvl = 'lvl3'
+
+def fix_bar_angle(bar_angle):
+    out = np.zeros(len(bar_angle))
+    out[0] = bar_angle[0]
+
+    for i in range(1, len(bar_angle)):
+        dphi = bar_angle[i] - bar_angle[i-1]
+        if dphi < -np.pi:
+            dphi += 2.*np.pi
+        
+        out[i] = out[i-1] + dphi
+    
+    return out
+
+def read_bar_prop(name, lvl):
+    t = h5.File(bprop_path + 'bar_prop_' + name + '-' + lvl + '.hdf5', mode='r')
+    out = {}
+    for key in t['bar_prop'].keys():
+        out[key] = t['bar_prop'][key][:]
+    
+    out['bar_angle'] = fix_bar_angle(t['bar_angle'][:])
+    t.close()
+
+
+    # print(out)
+    # print(len(out['bar_angle']))
+    # print(len(out['tlist']))
+
[…]

### 2022-09-07 test_part/ics/GalIC/src/allvars.c

@@ -1,0 +1,474 @@
+/*******************************************************************************
+ * This file is part of the GALIC code developed by D. Yurin and V. Springel.
+ *
+ * Copyright (c) 2014
+ * Denis Yurin (denis.yurin@h-its.org) 
+ * Volker Springel (volker.springel@h-its.org)
+ *******************************************************************************/
+
+/*! \file allvars.h
+ *  \brief declares global variables.
+ *
+ *  This file declares all global variables. Further variables should be added here, and declared as
+ *  'extern'. The actual existence of these variables is provided by the file 'allvars.c'. To produce
+ *  'allvars.c' from 'allvars.h', do the following:
+ *
+ *     - Erase all #define statements
+ *     - add #include "allvars.h"
+ *     - delete all keywords 'extern'
+ *     - delete all struct definitions enclosed in {...}, e.g.
+ *        "extern struct global_data_all_processes {....} All;"
+ *        becomes "struct global_data_all_processes All;"
+ */
+
+#include "allvars.h"
+
+
+
+#ifdef PERIODIC
+MyDouble boxSize, boxHalf;
+
+#ifdef LONG_X
+MyDouble boxSize_X, boxHalf_X;
+#else
+#endif
+#ifdef LONG_Y
+MyDouble boxSize_Y, boxHalf_Y;
+#else
+#endif
+#ifdef LONG_Z
+MyDouble boxSize_Z, boxHalf_Z;
+#else
+#endif
+#endif
+
+#ifdef FIX_PATHSCALE_MPI_STATUS_IGNORE_BUG
+MPI_Status mpistat;
+#endif
+
+/*********************************************************/
+/*  Global variables                                     */
+/*********************************************************/
+
+
+int FG_Nbin, FG_Ngrid;
+double FG_Rmin, FG_Fac, FG_Rin;
+
+double *FG_Pot;
+double *FG_DPotDR;
+double *FG_DPotDz;
+double *FG_Pot_exact;
+double *FG_DPotDR_exact;
+double *FG_DPotDz_exact;
+double *FG_Disp_r[6];
+double *FG_DispZ[6];
+double *FG_DispPhi[6];
+double *FG_Vstream[6];
+double *FG_tilted_vz2[6];
+double *FG_tilted_vR2[6];
+double *FG_tilted_vz2_prime[6];
[…]

### 2022-09-07 test_part/ics/GalIC/src/allvars.h

@@ -1,0 +1,1081 @@
+/*******************************************************************************
+ * This file is part of the GALIC code developed by D. Yurin and V. Springel.
+ *
+ * Copyright (c) 2014
+ * Denis Yurin (denis.yurin@h-its.org) 
+ * Volker Springel (volker.springel@h-its.org)
+ *******************************************************************************/
+/*! \file allvars.h
+ *  \brief declares global variables.
+ *
+ *  This file declares all global variables. Further variables should be added here, and declared as
+ *  'extern'. The actual existence of these variables is provided by the file 'allvars.c'. To proN_Cduce
+ *  'allvars.c' from 'allvars.h', do the following:
+ *
+ *     - Erase all #define statements
+ *     - add #include "allvars.h"
+ *     - delete all keywords 'extern'
+ *     - delete all struct definitions enclosed in {...}, e.g.
+ *        "extern struct global_data_all_processes {....} All;"
+ *        becomes "struct global_data_all_processes All;"
+ */
+#ifndef ALLVARS_H
+#define ALLVARS_H
+
+#include <mpi.h>
+
+#include <stdio.h>
+#include <gsl/gsl_rng.h>
+#include <gsl/gsl_math.h>
+#include <gsl/gsl_integration.h>
+#include <gsl/gsl_spline.h>
+#include <gsl/gsl_errno.h>
+#include <assert.h>
+
+#include "../build/galicconfig.h"
+
+#define TAG_N             10      /*!< Various tags used for labelling MPI messages */
+#define TAG_HEADER        11
+#define TAG_PDATA         12
+#define TAG_SPHDATA       13
+#define TAG_KEY           14
+#define TAG_DMOM          15
+#define TAG_NODELEN       16
+#define TAG_HMAX          17
+#define TAG_GRAV_A        18
+#define TAG_GRAV_B        19
+#define TAG_DIRECT_A      20
+#define TAG_DIRECT_B      21
+#define TAG_HYDRO_A       22
+#define TAG_HYDRO_B       23
[…]

### 2022-09-07 test_part/ics/GalIC/src/domain/domain.c

@@ -1,0 +1,215 @@
+/*******************************************************************************
+ * This file is part of the GALIC code developed by D. Yurin and V. Springel.
+ *
+ * Copyright (c) 2014
+ * Denis Yurin (denis.yurin@h-its.org) 
+ * Volker Springel (volker.springel@h-its.org)
+ *******************************************************************************/
+#include <mpi.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <strings.h>
+#include <math.h>
+
+#include "../allvars.h"
+#include "../proto.h"
+#include "domain.h"
+
+/*! \file domain.c
+ *  \brief code for domain decomposition
+ *
+ *  This file contains the code for the domain decomposition of the
+ *  simulation volume.  The domains are constructed from disjoint subsets
+ *  of the leaves of a fiducial top-level tree that covers the full
+ *  simulation volume. Domain boundaries hence run along tree-node
+ *  divisions of a fiducial global BH tree. As a result of this method, the
+ *  tree force are in principle strictly independent of the way the domains
+ *  are cut. The domain decomposition can be carried out for an arbitrary
+ *  number of CPUs. Individual domains are not cubical, but spatially
+ *  coherent since the leaves are traversed in a Peano-Hilbert order and
+ *  individual domains form segments along this order.  This also ensures
+ *  that each domain has a small surface to volume ratio, which minimizes
+ *  communication.
+ */
+
+
+
+/*! This is the main routine for the domain decomposition.  It acts as a
+ *  driver routine that allocates various temporary buffers, maps the
+ *  particles back onto the periodic box if needed, and then does the
+ *  domain decomposition, and a final Peano-Hilbert order of all particles
+ *  as a tuning measure.
+ */
+void domain_Decomposition(void)
+{
+  mpi_printf("DOMAIN:\n");
+  mpi_printf("DOMAIN: Begin domain decomposition (sync-point %d).\n", All.NumCurrentTiStep);
+
+
+
+  domain_allocate();
+  domain_allocate_lists();
[…]

### 2022-09-07 test_part/ics/GalIC/src/domain/domain.h

@@ -1,0 +1,153 @@
+/*******************************************************************************
+ * This file is part of the GALIC code developed by D. Yurin and V. Springel.
+ *
+ * Copyright (c) 2014
+ * Denis Yurin (denis.yurin@h-its.org) 
+ * Volker Springel (volker.springel@h-its.org)
+ *******************************************************************************/
+#ifndef ALLVARS_H
+#include "../allvars.h"
+#endif
+#ifndef DOMAIN_H
+#define DOMAIN_H
+
+
+extern struct local_topnode_data
+{
+  peanokey Size;		/*!< number of Peano-Hilbert mesh-cells represented by top-level node */
+  peanokey StartKey;		/*!< first Peano-Hilbert key in top-level node */
+  long long Count;		/*!< counts the number of particles in this top-level node */
+  int Daughter;			/*!< index of first daughter cell (out of 8) of top-level node */
+  int Leaf;			/*!< if the node is a leaf, this gives its number when all leaves are traversed in Peano-Hilbert order */
+  int Parent;
+  int PIndex;			/*!< first particle in node */
+}
+ *topNodes, *branchNodes;	/*!< points to the root node of the top-level tree */
+
+ struct domain_count_data
+ {
+   int task;
+   int count;
+   int origintask;
+ };
+
+
+
+extern struct domain_peano_hilbert_data
+ {
+   peanokey key;
+   int index;
+ }
+  *mp;
+
+
+
+extern struct trans_data
+{
+  MyIDType ID;
+  int new_task;
+  int new_index;
+  int wrapped;
+}
+ *trans_table;
+
+extern int N_trans;
+
+extern int Nbranch;
+
+extern double fac_load;
+
+
+extern double totpartcount;
+
+extern struct domain_cost_data
+{
+  int   no;
+  int   Count;        /*!< a table that gives the total number of particles held by each processor */
+}
+ *DomainLeaveNode;
+
+
+
+/*! toGo[partner] gives the number of particles on the current task that have to go to task 'partner'
+ */
+extern int *toGo;
+extern int *toGet;
+extern int *list_NumPart;
+extern int *list_load;
+
+
+
+
+int domain_check_for_local_refine_new(int i, MPI_Comm current_comm);
+int domain_double_to_int(double d);
+double domain_grav_tot_costfactor(int i);
+double domain_hydro_tot_costfactor(int i);
+void domain_init_sum_cost(void);
+void domain_printf(char *buf);
+void domain_report_balance(void);
+int domain_sort_load(const void *a, const void *b);
[…]

### 2022-09-07 test_part/ics/GalIC/src/domain/domain_vars.c

@@ -1,0 +1,60 @@
+/*******************************************************************************
+ * This file is part of the GALIC code developed by D. Yurin and V. Springel.
+ *
+ * Copyright (c) 2014
+ * Denis Yurin (denis.yurin@h-its.org) 
+ * Volker Springel (volker.springel@h-its.org)
+ *******************************************************************************/
+#include <mpi.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <strings.h>
+#include <math.h>
+
+
+#include "../allvars.h"
+#include "../proto.h"
+#include "domain.h"
+
+struct domain_peano_hilbert_data *mp;
+
+struct local_topnode_data *topNodes, *branchNodes;	/*!< points to the root node of the top-level tree */
+
+
+double totpartcount;
+
+struct domain_cost_data *DomainLeaveNode;
+
+double fac_load;
+
+int Nbranch;
+
+/*! toGo[partner] gives the number of particles on the current task that have to go to task 'partner'
+ */
+int *toGo;
+int *toGet;
+int *list_NumPart;
+int *list_load;
+
+
+
+
+void domain_allocate_lists(void)
+{
+  Key = (peanokey *) mymalloc_movable(&Key, "domain_key", (sizeof(peanokey) * All.MaxPart));
+  toGo = (int *) mymalloc_movable(&toGo, "toGo", (sizeof(int) * NTask));
+  toGet = (int *) mymalloc_movable(&toGet, "toGet", (sizeof(int) * NTask));
+  list_NumPart = (int *) mymalloc_movable(&list_NumPart, "list_NumPart", (sizeof(int) * NTask));
+  list_load = (int *) mymalloc_movable(&list_load, "list_load", (sizeof(int) * NTask));
+  DomainLeaveNode = (struct domain_cost_data *) mymalloc_movable(&DomainLeaveNode, "DomainLeaveNode", (MaxTopNodes * sizeof(struct domain_cost_data)));
+}
+
+void domain_free_lists(void)
+{
+  myfree(DomainLeaveNode);
+  myfree(list_load);
+  myfree(list_NumPart);
+  myfree(toGet);
+  myfree(toGo);
+}

### 2022-09-07 test_part/ics/GalIC/src/domain/peano.c

@@ -1,0 +1,715 @@
+/*******************************************************************************
+ * This file is part of the GALIC code developed by D. Yurin and V. Springel.
+ *
+ * Copyright (c) 2014
+ * Denis Yurin (denis.yurin@h-its.org) 
+ * Volker Springel (volker.springel@h-its.org)
+ *******************************************************************************/
+#include <mpi.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <math.h>
+
+#include "../allvars.h"
+#include "../proto.h"
+#include "domain.h"
+
+
+void mysort_peano(void *b, size_t n, size_t s, int (*cmp) (const void *, const void *));
+void peano_hilbert_key_inverse(peanokey key, int bits, int *x, int *y, int *z);
+int peano_compare_key(const void *a, const void *b);
+
+#include <gsl/gsl_heapsort.h>
+
+
+/** Returns the peanokey for the position on the Peano-Hilbert curve
+    that contains pos. */
+peanokey position_to_peanokey(MyDouble pos[3])
+{
+  // First find peano-hilbert key for the position
+  peanokey key = peano_hilbert_key((int) ((pos[0] - DomainCorner[0]) * DomainFac),
+				   (int) ((pos[1] - DomainCorner[1]) * DomainFac),
+				   (int) ((pos[2] - DomainCorner[2]) * DomainFac),
+				   BITS_PER_DIMENSION);
+  return key;
+}
+
+
+/** Returns the topnode index containing the specified peanokey. */
+int peanokey_to_topnode(peanokey key)
+{
+  int node = 0;
+  while(TopNodes[node].Daughter >= 0)
+    {
[…]

### 2022-09-07 test_part/ics/GalIC/src/mpi_utils/checksummed_sendrecv.c

@@ -1,0 +1,278 @@
+/*******************************************************************************
+ * This file is part of the GALIC code developed by D. Yurin and V. Springel.
+ *
+ * Copyright (c) 2014
+ * Denis Yurin (denis.yurin@h-its.org) 
+ * Volker Springel (volker.springel@h-its.org)
+ *******************************************************************************/
+#include <mpi.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <math.h>
+#include <gsl/gsl_math.h>
+
+#include "../allvars.h"
+#include "../proto.h"
+
+#ifdef MPISENDRECV_CHECKSUM
+
+#undef MPI_Sendrecv
+
+
+int MPI_Check_Sendrecv(void *sendbuf, int sendcount, MPI_Datatype sendtype,
+		       int dest, int sendtag, void *recvbufreal, int recvcount,
+		       MPI_Datatype recvtype, int source, int recvtag, MPI_Comm comm, MPI_Status * status)
+{
+  int checksumtag = 1000, errtag = 2000;
+  int i, iter = 0, err_flag, err_flag_imported, size_sendtype, size_recvtype;
+  long long sendCheckSum, recvCheckSum, importedCheckSum;
+  unsigned char *p, *buf, *recvbuf;
+  char msg[500];
+
+  if(dest != source)
+    terminate("destination task different from source task");
+
+  MPI_Type_size(sendtype, &size_sendtype);
+  MPI_Type_size(recvtype, &size_recvtype);
+
+  if(dest == ThisTask)
+    {
+      memcpy(recvbufreal, sendbuf, recvcount * size_recvtype);
+      return 0;
+    }
+
+
+  if(!(buf = mymalloc(recvcount * size_recvtype + 1024)))
+    terminate("not enough memory to allocate the buffer buf");
+
+  for(i = 0, p = buf; i < recvcount * size_recvtype + 1024; i++)
+    *p++ = 255;
+
+  recvbuf = buf + 512;
+
[…]

### 2022-09-07 test_part/ics/GalIC/src/parallel_sort.c

@@ -1,0 +1,719 @@
+/*******************************************************************************
+ * This file is part of the GALIC code developed by D. Yurin and V. Springel.
+ *
+ * Copyright (c) 2014
+ * Denis Yurin (denis.yurin@h-its.org) 
+ * Volker Springel (volker.springel@h-its.org)
+ *******************************************************************************/
+#include <mpi.h>
+#include <stdlib.h>
+#include <stdio.h>
+#include <string.h>
+#include <math.h>
+#include <sys/time.h>
+#include <sys/resource.h>
+#include <unistd.h>
+#include <signal.h>
+#include <gsl/gsl_rng.h>
+
+#include "allvars.h"
+#include "proto.h"
+
+
+#define TRANSFER_SIZE_LIMIT  1000000000
+#define MAX_ITER_PARALLEL_SORT      500
+
+/* Note: For gcc-4.1.2, the compiler produces incorrect code for this routine if optimization level O1 or higher is used.
+ *       In  gcc-4.3.4, this problem is absent.
+ */
+
+#define TAG_TRANSFER  100
+
+/*
+#define CHECK_LOCAL_RANK
+*/
+
+static void serial_sort(char *base, size_t nmemb, size_t size, int (*compar) (const void *, const void *));
+
+static void msort_serial_with_tmp(char *base, size_t n, size_t s, int (*compar) (const void *, const void *),
+				  char *t);
+
+static void get_local_rank(char *element,
+			   size_t tie_braking_rank,
+			   char *base,
+			   size_t nmemb, size_t size, size_t noffs_thistask,
+			   long long left, long long right, size_t * loc, int (*compar) (const void *,
+											 const void *));
+
+#ifdef CHECK_LOCAL_RANK
+static void check_local_rank(char *element,
+			     size_t tie_braking_rank,
+			     char *base,
+			     size_t nmemb, size_t size,
+			     size_t noffs_thistask, long long left, long long right, size_t loc,
+			     int (*compar) (const void *, const void *));
+#endif
+
+static int (*comparfunc) (const void *, const void *);
+static char *median_element_list;
+static size_t element_size;
+
+int parallel_sort_indirect_compare(const void *a, const void *b)
+{
[…]

### 2022-09-07 test_part/ics/GalIC/src/set_particles.c

@@ -1,0 +1,449 @@
+/*******************************************************************************
+ * This file is part of the GALIC code developed by D. Yurin and V. Springel.
+ *
+ * Copyright (c) 2014
+ * Denis Yurin (denis.yurin@h-its.org) 
+ * Volker Springel (volker.springel@h-its.org)
+ *******************************************************************************/
+#include <mpi.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <math.h>
+#include <gsl/gsl_math.h>
+#include <gsl/gsl_rng.h>
+#include <gsl/gsl_randist.h>
+
+#include "allvars.h"
+#include "proto.h"
+
+
+/* This function samples the mass model with particles.
+ * The main job of the code will then be to find the initial velocities.
+ */
+void initialize_particles(void)
+{
+  int n, i, k;
+  double phi, theta, vr;
+  double vsum2 = 0, rsum2 = 0, vsum2_exact = 0;
+  int count_r[6], count_t[6], count_p[6], count_q[6];
+  int tot_count_r[6], tot_count_t[6], tot_count_p[6], tot_count_q[6];
+
+  int nhalo = get_part_count_this_task(All.Halo_N);
+  int ndisk = get_part_count_this_task(All.Disk_N);
+  int nbulge = get_part_count_this_task(All.Bulge_N);
+
+  NumPart = nhalo + ndisk + nbulge;
+
+  MPI_Allreduce(&NumPart, &All.MaxPart, 1, MPI_INT, MPI_MAX, MPI_COMM_WORLD);
+  sumup_large_ints(1, &NumPart, &All.TotNumPart);
+
+  P = (struct particle_data *) mymalloc_movable(&P, "P", All.MaxPart * sizeof(struct particle_data));
+  memset(P, 0, All.MaxPart * sizeof(struct particle_data));
+
+  permutation = (struct permutation_data *) mymalloc_movable(&permutation, "permutation", All.MaxPart * sizeof(struct permutation_data));
+
+  n = 0;
+
+  for(i = 0; i < 6; i++)
[…]

### 2022-09-07 test_part/runs/force.py

@@ -1,0 +1,108 @@
+import numpy as np
+from numba import njit, prange
+
+@njit(parallel=True)
+def Hernquist_force(pos, M, a, G):
+    N = pos.shape[0]
+    mGM = - G * M
+    acc = np.zeros(pos.shape)
+    for i in prange(N):
+        r = np.sqrt(pos[i][0]*pos[i][0] + pos[i][1]*pos[i][1] + pos[i][2]*pos[i][2])
+        acc[i][0] = mGM * pos[i][0] / (r * (a + r)**2)
+        acc[i][1] = mGM * pos[i][1] / (r * (a + r)**2)
+        acc[i][2] = mGM * pos[i][2] / (r * (a + r)**2)
+        
+    return acc
+
+@njit(parallel=True)
+def Hernquist_pot(pos, M, a, G):
+    N = pos.shape[0]
+    mGM = - G * M
+    pot = np.zeros(N)
+    for i in prange(N):
+        r = np.sqrt(pos[i][0]*pos[i][0] + pos[i][1]*pos[i][1] + pos[i][2]*pos[i][2])
+        pot[i] = mGM / (r + a)
+        
+    return pot
+
+@njit
+def _compute_phi_r(A, vc, r, rCR, b):
+    ans = - A * vc**2 / 2
+    ans *= (r/rCR)**2
+    ans *= ((b+1.)/(b+r/rCR))**5.
+    return ans
+
+@njit
+def _compute_partial_phi_r(A, vc, r, rCR, b):
+    term1 = - A * vc**2 * (r/rCR)**2 * ((b+1)/(b+r/rCR))**5
+    term2 = (-A*vc**2/2) * (5 * ((b+1)/(b+r/rCR))**4) * (-((b+1)*rCR)/(r+rCR)**2)
+    
+    return term1 + term2
+
+@njit(parallel=True)
+def bar_pot(pos, A, b, vc, ps, ang):
+    N = pos.shape[0]
+    pot = np.zeros(N)
+    rCR = vc / ps
+    for i in range(N):
+        r = np.sqrt(pos[i][0]*pos[i][0] + pos[i][1]*pos[i][1] + pos[i][2]*pos[i][2])
+        theta = np.arccos(pos[i][2]/r)
+        phi = np.arctan2(pos[i][1], pos[i][0])
+        
+        phi_r = _compute_phi_r(A, vc, r, rCR, b)
+        
[…]

### 2022-09-07 test_part/runs/job_varyG.sh

@@ -1,0 +1,21 @@
+#!/bin/sh
+#SBATCH -p shared,itc_cluster,conroy 
+#SBATCH -J sam_vG
+#SBATCH -n 1
+#SBATCH -N 1 
+#SBATCH --ntasks-per-node=1
+#SBATCH -o logs/OUTPUT.%j.out
+#SBATCH -e logs/ERROR.%j.err
+#SBATCH --mail-user=angus.beane@cfa.harvard.edu
+#SBATCH --mail-type=BEGIN
+#SBATCH --mail-type=END
+#SBATCH --mail-type=FAIL
+#SBATCH --mem-per-cpu=3900
+#SBATCH -t 00-00:30           # Runtime in D-HH:MM
+
+source ./load-modules.sh
+
+export G=$1
+
+python3 run.py 8 $G output-Gvar/out_G${G}.npy
+

### 2022-11-30 paper-ApJ/ms.tex
Add paper-ApJ

--- 2022-09-07 paper/ms.tex
@@
-% region preamble
-
-\documentclass[fleqn,usenatbib]{mnras}
-
-\usepackage[T1]{fontenc}
-
-% Allow "Thomas van Noord" and "Simon de Laguarde" and alike to be sorted by "N" and "L" etc. in the bibliography.
-% Write the name in the bibliography as "\VAN{Noord}{Van}{van} Noord, Thomas"
-\DeclareRobustCommand{\VAN}[3]{#2}
-\let\VANthebibliography\thebibliography
-\def\thebibliography{\DeclareRobustCommand{\VAN}[3]{##3}\VANthebibliography}
+%% Beginning of file 'sample631.tex'
+%%
+%% Modified 2021 March
+%%
+%% This is a sample manuscript marked up using the
+%% AASTeX v6.31 LaTeX 2e macros.
+%%
+%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
+%% (Copyright 2000-2015).  See the classfile for details.
+
+%% AASTeX requires revtex4-1.cls and other external packages such as
+%% latexsym, graphicx, amssymb, longtable, and epsf.  Note that as of 
+%% Oct 2020, APS now uses revtex4.2e for its journals but remember that 
+%% AASTeX v6+ still uses v4.1. All of these external packages should 
+%% already be present in the modern TeX distributions but not always.
+%% For example, revtex4.1 seems to be missing in the linux version of
+%% TexLive 2020. One should be able to get all packages from www.ctan.org.
+%% In particular, revtex v4.1 can be found at 
+%% https://www.ctan.org/pkg/revtex4-1.
+
+%% The first piece of markup in an AASTeX v6.x document is the \documentclass
+%% command. LaTeX will ignore any data that comes before this command. The 
+%% documentclass can take an optional argument to modify the output style.
+%% The command below calls the preprint style which will produce a tightly 
+%% typeset, one-column, single-spaced document.  It is the default and thus
+%% does not need to be explicitly stated.
+%%
+%% using aastex version 6.3
+\documentclass[twocolumn,linenumbers]{aastex631}
+
+%% The default is a single spaced, 10 point font, single spaced article.
+%% There are 5 other style options available via an optional argument. They
+%% can be invoked like this:
+%%
+%% \documentclass[arguments]{aastex631}
+%% 
+%% where the layout options are:
+%%
+%%  twocolumn   : two text columns, 10 point font, single spaced article.
+%%                This is the most compact and represent the final published
+%%                derived PDF copy of the accepted manuscript from the publisher
+%%  manuscript  : one text column, 12 point font, double spaced article.
+%%  preprint    : one text column, 12 point font, single spaced article.  
+%%  preprint2   : two text columns, 12 point font, single spaced article.
+%%  modern      : a stylish, single text column, 12 point font, article with
+%% 		  wider left and right margins. This uses the Daniel
+%% 		  Foreman-Mackey and David Hogg design.
+%%  RNAAS       : Supresses an abstract. Originally for RNAAS manuscripts 
+%%                but now that abstracts are required this is obsolete for
+%%                AAS Journals. Authors might need it for other reasons. DO NOT
+%%                use \begin{abstract} and \end{abstract} with this style.
+%%
+%% Note that you can submit to the AAS Journals in any of these 6 styles.
+%%
+%% There are other optional arguments one can invoke to allow other stylistic
+%% actions. The available options are:
+%%
+%%   astrosymb    : Loads Astrosymb font and define \astrocommands. 
+%%   tighten      : Makes baselineskip slightly smaller, only works with 
+%%                  the twocolumn substyle.
+%%   times        : uses times font instead of the default
+%%   linenumbers  : turn on lineno package.
+%%   trackchanges : required to see the revision mark up and print its output
+%%   longauthor   : Do not use the more compressed footnote style (default) for 
+%%                  the author/collaboration/affiliations. Instead print all
+%%                  affiliation information after each name. Creates a much 
+%%                  longer author list but may be desirable for short 
+%%                  author papers.
+%% twocolappendix : make 2 column appendix.
+%%   anonymous    : Do not show the authors, affiliations and acknowledgments 
+%%                  for dual anonymous review.
+%%
+%% these can be used in any combination, e.g.
+%%
+%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex631}
+%%
+%% AASTeX v6.* now includes \hyperref support. While we have built in specific
+%% defaults into the classfile you can manually override them with the
+%% \hypersetup command. For example,
+%%
+%% \hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
+%%
+%% will change the color of the internal links to red, the links to the
+%% bibliography to green, the file links to cyan, and the external links to
+%% magenta. Additional information on \hyperref options can be found here:
+%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
+%%
+%% Note that in v6.3 "bookmarks" has been changed to "true" in hyperref
+%% to improve the accessibility of the compiled pdf file.
+%%
+%% If you want to create your own macros, you can do so
+%% using \newcommand. Your macros should appear before
+%% the \begin{document} command.
+%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand{\vdag}{(v)^\dagger}
+\newcommand\aastex{AAS\TeX}
+\newcommand\latex{La\TeX}
 
@@
 
-% endregion
-
-% region title
-%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%
-
-% Title of the paper, and the short title which is used in the headers.
-% Keep the title short and informative.
-\title[Stellar Bars and Gas]{Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down}
-
-\author[A. Beane et al.]{Angus Beane,$^{1}$\thanks{E-mail: angus.beane@cfa.harvard.edu (AB)}
-Lars Hernquist,$^{1}$
-Elena~D'Onghia,$^{2,3}$
-Federico~Marinacci,$^{4}$
-Charlie Conroy,$^{1}$
-Jia~Qi,$^{5}$\newauthor
-Laura~V.~Sales,$^{6}$
-Paul~Torrey,$^{5}$
-Mark~Vogelsberger$^{7}$
-\\
-% List of institutions
-$^{1}$Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA\\
-$^{2}$Department of Physics, University of Wisconsin-Madison, Madison, WI, USA\\
-$^{3}$Department of Astronomy, University of Wisconsin-Madison, Madison, WI, USA\\
-$^{4}$Department of Physics \& Astronomy `Augusto Righi', University of Bologna, Bologna, Italy\\
-$^{5}$Department of Astronomy, University of Florida, Gainesville, FL, USA\\
-$^{6}$Department of Physics \& Astronomy, University of California, Riverside, CA, USA\\
-$^{7}$Department of Physics, Massachusetts Institute of Technology, Cambridge, MA, USA\\
-}
-
-% These dates will be filled out by the publisher
-\date{Accepted XXX. Received YYY; in original form ZZZ}
-
-% Enter the current year, for the copyright statements etc.
-\pubyear{2022}
-
-% endregion
-
-% Don't change these lines
+%% Reintroduced the \received and \accepted commands from AASTeX v5.2
+%\received{March 1, 2021}
+%\revised{April 1, 2021}
+%\accepted{\today}
+
+\shorttitle{Stellar Bars and Gas}
+\shortauthors{Beane et al.}
+
+\graphicspath{{./}{figures/}}
+
 \begin{document}
-\label{firstpage}
-\pagerange{\pageref{firstpage}--\pageref{lastpage}}
-\maketitle
-
-% Abstract of the paper
+
+\title{Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down}
+
+%% LaTeX will automatically break titles if they run longer than
+%% one line. However, you may use \\ to force a line break if
+%% you desire. In v6.31 you can include a footnote in the title.
+
+%% A significant change from earlier AASTEX versions is in the structure for 
+%% calling author and affiliations. The change was necessary to implement 
+%% auto-indexing of affiliations which prior was a manual process that could 
+%% easily be tedious in large author manuscripts.
+%%
+%% The \author command is the same as before except it now takes an optional
+%% argument which is the 16 digit ORCID. The syntax is:
+%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
+%%
+%% This will hyperlink the author name to the author's ORCID page. Note that
+%% during compilation, LaTeX will do some limited checking of the format of
+%% the ID to make sure it is valid. If the "orcid-ID.png" image file is 
+%% present or in the LaTeX pathway, the OrcID icon will appear next to
+%% the authors name.
+%%
+%% Use \affiliation for affiliation information. The old \affil is now aliased
+%% to \affiliation. AASTeX v6.31 will automatically index these in the header.
+%% When a duplicate is found its index will be the same as its previous entry.
+%%
+%% Note that \altaffilmark and \altaffiltext have been removed and thus 
+%% can not be used to document secondary affiliations. If they are used latex
+%% will issue a specific error message and quit. Please use multiple 
+%% \affiliation calls for to document more than one affiliation.
+%%
+%% The new \altaffiliation can be used to indicate some secondary information
+%% such as fellowships. This command produces a non-numeric footnote that is
+%% set away from the numeric \affiliation footnotes.  NOTE that if an
+%% \altaffiliation command is used it must come BEFORE the \affiliation call,
+%% right after the \author command, in order to place the footnotes in
+%% the proper location.
+%%
+%% Use \email to set provide email addresses. Each \email will appear on its
+%% own line so you can put multiple email address in one \email call. A new
+%% \correspondingauthor command is available in V6.31 to identify the
+%% corresponding author of the manuscript. It is the author's responsibility
+%% to make sure this name is also in the author list.
+%%
+%% While authors can be grouped inside the same \author and \affiliation
+%% commands it is better to have a single author for each. This allows for
+%% one to exploit all the new benefits and should make book-keeping easier.
+%%
+%% If done correctly the peer review system will be able to
+%% automatically put the author and affiliation information from the manuscript
+%% and save the corresponding author the trouble of entering it by hand.
+
+%\correspondingauthor{August Muench}
+%\email{greg.schwarz@aas.org, gus.muench@aas.org}
+
+\author{Angus Beane}
+\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA}
+
+\author{Lars Hernquist}
+\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA}
+
+\author{Elena D'Onghia}
+\affiliation{Department of Physics, University of Wisconsin-Madison, Madison, WI, USA}
+\affiliation{Department of Astronomy, University of Wisconsin-Madison, Madison, WI, USA}
+
+\author{Federico Marinacci}
+\affiliation{Department of Physics \& Astronomy `Augusto Righi', University of Bologna, Bologna, Italy}
+
+\author{Charlie Conroy}
+\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA}
+
+\author{Jia Qi}
+\affiliation{Department of Astronomy, University of Florida, Gainesville, FL, USA}
+
+\author{Laura V. Sales}
+\affiliation{Department of Physics \& Astronomy, University of California, Riverside, CA, USA}
+
+\author{Paul Torrey}
+\affiliation{Department of Astronomy, University of Florida, Gainesville, FL, USA}
+
+\author{Mark Vogelsberger}
+\affiliation{Department of Physics, Massachusetts Institute of Technology, Cambridge, MA, USA}
+
 \begin{abstract}
-Elongated bar-like features are ubiquitous in galaxies, occurring at the centers
-of approximately two-thirds of spiral disks.  Due to gravitational interactions
-between the bar and the other components of galaxies, it is expected that
-angular momentum and matter will redistribute over long (Gyr) timescales in
-barred galaxies. Previous work ignoring the gas phase of galaxies has
-conclusively demonstrated that bars should slow their rotation over time due to
-their interaction with dark matter halos. We have performed a simulation of a
-Milky Way-like galactic disk hosting a strong bar which includes a
-state-of-the-art model of the interstellar medium and a live dark matter halo.
-In this simulation the bar pattern does not slow down over time, and instead
-remains at a stable, constant rate of rotation. This behavior has been observed
-in previous simulations using more simplified models for the interstellar gas,
-but the apparent lack of secular evolution has remained unexplained. We propose
-that the gas phase of the disk and the dark matter halo act in concert to
-stabilize the bar pattern speed and prevent the bar from slowing down or
-speeding up. We find that in a Milky Way-like disk, a gas fraction of only about
-5\% is necessary for this mechanism to operate. Our result naturally explains
-why nearly all observed bars rotate rapidly and is especially relevant for our
-understanding of how the Milky Way arrived at its present state.
+
+  Elongated bar-like features are ubiquitous in galaxies, occurring at the
+  centers of approximately two-thirds of spiral disks.  Due to gravitational
+  interactions between the bar and the other components of galaxies, it is
+  expected that angular momentum and matter will redistribute over long (Gyr)
+  timescales in barred galaxies. Previous work ignoring the gas phase of
+  galaxies has conclusively demonstrated that bars should slow their rotation
+  over time due to their interaction with dark matter halos. We have performed a
+  simulation of a Milky Way-like galactic disk hosting a strong bar which
+  includes a state-of-the-art model of the interstellar medium and a live dark
+  matter halo. In this simulation the bar pattern does not slow down over time,
+  and instead remains at a stable, constant rate of rotation. This behavior has
+  been observed in previous simulations using more simplified models for the
+  interstellar gas, but the apparent lack of secular evolution has remained
+  unexplained. We propose that the gas phase of the disk and the dark matter
+  halo act in concert to stabilize the bar pattern speed and prevent the bar
+  from slowing down or speeding up. We find that in a Milky Way-like disk, a gas
+  fraction of only about 5\% is necessary for this mechanism to operate. Our
+  result naturally explains why nearly all observed bars rotate rapidly and is
+  especially relevant for our understanding of how the Milky Way arrived at its
+  present state.
+
 \end{abstract}
 
-% Select between one and six entries from the list of approved keywords.
-% Don't make up new ones.
-\begin{keywords}
-galaxies: bar - galaxies: disc - galaxies: evolution - galaxies: kinematics and
-dynamics
-\end{keywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
+%% Keywords should appear after the \end{abstract} command. 
+%% The AAS Journals now uses Unified Astronomy Thesaurus concepts:
+%% https://astrothesaurus.org
+%% You will be asked to selected these concepts during the submission process
+%% but this old "keyword" functionality is maintained in case authors want
+%% to include these concepts in their preprints.
+\keywords{Classical Novae (251) --- Ultraviolet astronomy(1736) --- History of astronomy(1868) --- Interdisciplinary astronomy(804)}
+
+%% From the front matter, we move on to the body of the paper.
+%% Sections are demarcated by \section and \subsection, respectively.
+%% Observe the use of the LaTeX \label
+%% command after the \subsection to give a symbolic KEY to the
+%% subsection for cross-referencing in a \ref command.
+%% You can use LaTeX's \ref and \label commands to keep track of
+%% cross-references to sections, equations, tables, and figures.
+%% That way, if you change the order of any elements, LaTeX will
+%% automatically renumber them.
+%%
+%% We recommend that authors also use the natbib \citep
+%% and \citet commands to identify citations.  The citations are
+%% tied to the reference list via symbolic KEYs. The KEY corresponds
+%% to the KEY in the \bibitem in the reference list below.
 
@@
 what was found in earlier works, that nearly all extragalactic bars are fast
-rotators (i.e., they rotate close to their maximum rotation rate)
+rotators (i.e., they rotate close to their maximum rotation rate).
 
@@
 
-% \subsection{Figures and tables}
-
-% Figures and tables should be placed at logical positions in the text. Don't
-% worry about the exact layout, which will be handled by the publishers.
-
-% Figures are referred to as e.g. Fig.~\ref{fig:example_figure}, and tables as
-% e.g. Table~\ref{tab:example_table}.
-
-% % Example figure
-% \begin{figure}
-% 	% To include a figure from a file named example.*
-% 	% Allowable file formats are eps or ps if compiling using latex
-% 	% or pdf, png, jpg if compiling using pdflatex
-% 	\includegraphics[width=\columnwidth]{example}
-%     \caption{This is an example figure. Captions appear below each figure.
-% 	Give enough detail for the reader to understand what they're looking at,
-% 	but leave detailed discussion to the main body of the text.}
-%     \label{fig:example_figure}
-% \end{figure}
-
-% % Example table
-% \begin{table}
-% 	\centering
-% 	\caption{This is an example table. Captions appear above each table.
-% 	Remember to define the quantities, symbols and units used.}
-% 	\label{tab:example_table}
-% 	\begin{tabular}{lccr} % four columns, alignment for each
-% 		\hline
-% 		A & B & C & D\\
-% 		\hline
-% 		1 & 2 & 3 & 4\\
-% 		2 & 4 & 6 & 8\\
-% 		3 & 5 & 7 & 9\\
-% 		\hline
-% 	\end{tabular}
-% \end{table}
-
 \section{Discussion}
@@
 the bar) is about $2$. This is a factor of $4$ smaller than our fiducial value
-of $I_6=8$, but this is probably due to the fact that the bar does not really
-rotate as a solid body.
+of $I_6=8$, but this is probably due either to the fact that the bar does not
+really rotate as a solid body or that resonantly captured stars contribute to
+the real bar's effective moment of inertia \citep{1985MNRAS.213..451W}.
 
@@
 We next take the $\tau_{\textrm{gas}}=0$ and $20$ cases from Fig.~\ref{fig:sam}
-and plot the torque evolution. This result, given in Fig.~\ref{fig:sam-torque},
+and plot the halo torque evolution. This result, given in Fig.~\ref{fig:sam-torque},
 is comparable to the lower panel of Fig.~\ref{fig:prop}. We find that the
@@
 in conflict with observations by producing bars which rotate too slowly
-\citep{2017MNRAS.469.1054A, 2019MNRAS.483.2721P, 2021AA...650L..16F}. However,
-the pattern speeds of bars in both cosmological simulations and the real
-universe can be affected by environmental processes not included in our
-simulation -- e.g., satellite infall \citep{2011Natur.477..301P}, non-sphericity
-\citep{2013MNRAS.429.1949A}, rotation in the dark matter halo
-\citep{2013MNRAS.434.1287S, 2014ApJ...783L..18L, 2018MNRAS.476.1331C,
-2019MNRAS.488.5788C}, or perhaps even the gaseous circumgalactic medium.
-Naturally, extending our present work to account for such effects is a crucial
-next step in understanding the formation and evolution of galactic bars. We are
-presently engaged in such an exploration.
+\citep{2017MNRAS.469.1054A, 2019MNRAS.483.2721P, 2021AA...650L..16F,
+2022ApJ...940...61F}. However, the pattern speeds of bars in both cosmological
+simulations and the real universe can be affected by environmental processes not
+included in our simulation -- e.g., satellite infall
+\citep{2011Natur.477..301P}, non-sphericity \citep{2013MNRAS.429.1949A},
+rotation in the dark matter halo \citep{2013MNRAS.434.1287S,
+2014ApJ...783L..18L, 2018MNRAS.476.1331C, 2019MNRAS.488.5788C}, or perhaps even
+the gaseous circumgalactic medium. Naturally, extending our present work to
+account for such effects is a crucial next step in understanding the formation
+and evolution of galactic bars. We are presently engaged in such an exploration.
+We note that \citet{2022ApJ...940...61F} argue that bars in the TNG50 simulation
+are too short rather than too slow.
 
@@
 spiral galaxies should rotate slowly.
-https://github.com/gusbeane/gasbar
-\section*{Acknowledgements}
+
+\begin{acknowledgments}
   We would like to thank Greg~L. Bryan, Neal~J. Evans, Drummond~B. Fielding, Keith
@@
   AST-1814053, AST-1814259, AST-1909831, AST-2007355 and AST-2107724.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Data Availability}
-A git repository containing all analysis scripts and notebooks is available at
-\url{https://github.com/gusbeane/gasbar}. Snapshots at a $500\,\Myr$ cadence, as
-well as the initial conditions, are available at . Snapshots at a higher cadence
-are available upon request without expectation of coauthorship.
-
-We made use of the following software packages in {\sc Python} \citep{Python}:
+\end{acknowledgments}
+
+\software{
 {\sc agama} \url{https://github.com/GalacticDynamics-Oxford/Agama}, {\sc
@@
 \citep{harris2020array}, {\sc scipy} \citep{2020SciPy-NMeth}, {\sc tqdm}
-\url{https://tqdm.github.io/}.
-
-%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%
-
-% The best way to enter references is to use BibTeX:
-
-\bibliographystyle{mnras}
-\bibliography{ref} % if your bibtex file is called example.bib
-
-% Alternatively you could enter them by hand, like this:
-% This method is tedious and prone to error if you have lots of references
-%\begin{thebibliography}{99}
-%\bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
-%Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
-%\bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
-%Others S., 2012, Journal of Interesting Stuff, 17, 198
-%\end{thebibliography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%
+\url{https://tqdm.github.io/}
+          }
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Don't change these lines
-\bsp	% typesetting comment
-\label{lastpage}
+%% For this sample we use BibTeX plus aasjournals.bst to generate the
+%% the bibliography. The sample631.bib file was populated from ADS. To
+%% get the citations to show in the compiled file do the following:
+%%
+%% pdflatex sample631.tex
+%% bibtext sample631
+%% pdflatex sample631.tex
+%% pdflatex sample631.tex
+
+\bibliography{ref}{}
+\bibliographystyle{aasjournal}
+
+%% This command is needed to show the entire author+affiliation list when
+%% the collaboration and author truncation commands are used.  It has to
+%% go at the end of the manuscript.
+%\allauthors
+
+%% Include this line if you are using the \added, \replaced, \deleted
+%% commands to see a summary list of all changes at the end of the article.
+%\listofchanges
 
@@
 
-% End of mnras_template.tex
+% End of file `sample631.tex'.

### 2022-11-30 paper-ApJ/ms.tex
Cite Font et al

@@
 to be stable \citep{1980AA....81..198C}. Observational estimates of the pattern
-speeds of bars indicate that nearly all galaxies have $1 < \Rot <
-1.4$ \citep{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G,
-2020MNRAS.491.3655G}.
+speeds of bars indicate that nearly all galaxies have $1 < \Rot < 1.4$
+\citep{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G,
+2020MNRAS.491.3655G}. We note that \citet{2017ApJ...835..279F} argue that the
+pattern speed should be measured relative to a characteristic angular velocity
+of the outer disk.
 

### 2022-11-30 plots/ps_fgas/plot_ps_fgas.py
Fix units in fgas plot

@@ -158,7 +158,7 @@ def run():
     ax.set(ylim=(0, 60), ylabel=r'$\Omega_p\,[\,\text{km}/\text{s}/\text{kpc}\,]$')
     ax.set(xlim=(0, 5), xlabel=r'$t\,[\,\textrm{Gyr}\,]$')
 
-    ax.legend(frameon=False, ncol=2, title=r'$\Sigma_{\textrm{gas}}\,[\,M_{\odot}/\textrm{kpc}^2\,]$')
+    ax.legend(frameon=False, ncol=2, title=r'$\Sigma_{\textrm{gas}}\,[\,M_{\odot}/\textrm{pc}^2\,]$')
 
     fig.tight_layout()
 

### 2022-11-30 paper-ApJ/ms.tex
Fix units, add acknowledgments

@@
 with a constant pattern speed while the bar in a disk with initial surface
-density of $5\,\Msun/\textrm{kpc}^2$ slows down at a similar rate to the \Nbody{} case.
+density of $5\,\Msun/\textrm{pc}^2$ slows down at a similar rate to the \Nbody{} case.
 
@@
     the evolution of the pattern speed. The surface densities we tested of $20$,
-    $15$, $10$, and $5\,\Msun/\textrm{kpc}^2$ correspond to initial gas
+    $15$, $10$, and $5\,\Msun/\textrm{pc}^2$ correspond to initial gas
     fractions of $16\%$, $10\%$, $7\%$, and $4\%$. We find that initial surface
-    densities $20$, $15$, and $10\,\Msun/\textrm{kpc}^2$ result in bars which
+    densities $20$, $15$, and $10\,\Msun/\textrm{pc}^2$ result in bars which
     remain at a constant pattern speed, while an initial surface density of
-    $5\,\Msun/\textrm{kpc}^2$ results in a bar which slows down in roughly the
+    $5\,\Msun/\textrm{pc}^2$ results in a bar which slows down in roughly the
     same manner as the \Nbody{} case.}
@@
 \begin{acknowledgments}
-  We would like to thank Greg~L. Bryan, Neal~J. Evans, Drummond~B. Fielding, Keith
-  Hawkins, Jason~A.~S. Hunt, Sarah~M.~R. Jeffreson, Kathryn~V. Johnston,
-  Peter~M.~W. Kalberla, Jürgen Kerp, Julio~F. Navarro, Joshua~S. Speagle,
-  Martin~D. Weinberg, and Yanfei Zou for helpful discussions. A.B. would like to
-  thank Todd Phillips for helpful discussions. Some computations in this paper
-  were run on the FASRC Cannon cluster supported by the FAS Division of Science
-  Research Computing Group at Harvard University. Resources supporting this work
-  were also provided by the NASA High-End Computing (HEC) Program through the NASA
-  Advanced Supercomputing (NAS) Division at Ames Research Center. This research
-  has made use of the Spanish Virtual Observatory (https://svo.cab.inta-csic.es)
-  project funded by MCIN/AEI/10.13039/501100011033/ through grant
-  PID2020-112949GB-I00. A.B. was supported by the Future Investigators in NASA
-  Earth and Space Science and Technology (FINESST) award number 80NSSC20K1536
-  during the completion of this work. E.D. was partially supported by HST grants
-  HST-AR-16363.001 and HST-AR-16602.006-A and by NASA Award NASA 80NSSC22K0761.
-  J.Q. acknowledges support from NSF grant AST-2008490. L.V.S. is grateful for
-  financial support from NASA ATP 80NSSC20K0566, NSF AST 1817233 and NSF CAREER
-  1945310 grants. P.T. acknowledges support from NSF grant AST-1909933,
-  AST-2008490, and NASA ATP Grant 80NSSC20K0502. M.V. acknowledges support through
-  NASA ATP 19-ATP19-0019, 19-ATP19-0020, 19-ATP19-0167, and NSF grants
-  AST-1814053, AST-1814259, AST-1909831, AST-2007355 and AST-2107724.
+  We would like to thank Greg~L. Bryan, Neal~J. Evans, Drummond~B. Fielding,
+  Keith Hawkins, Jason~A.~S. Hunt, Sarah~M.~R. Jeffreson, Kathryn~V. Johnston,
+  Peter~M.~W. Kalberla, Jürgen Kerp, Julio~F. Navarro, Dylan Nelson, Suchira
+  Sarkar, Joshua~S. Speagle, Martin~D. Weinberg, and Yanfei Zou for helpful
+  discussions. A.B. would like to thank Todd Phillips for helpful discussions.
+  Some computations in this paper were run on the FASRC Cannon cluster supported
+  by the FAS Division of Science Research Computing Group at Harvard University.
+  Resources supporting this work were also provided by the NASA High-End
+  Computing (HEC) Program through the NASA Advanced Supercomputing (NAS)
+  Division at Ames Research Center. This research has made use of the Spanish
+  Virtual Observatory (https://svo.cab.inta-csic.es) project funded by
+  MCIN/AEI/10.13039/501100011033/ through grant PID2020-112949GB-I00. A.B. was
+  supported by the Future Investigators in NASA Earth and Space Science and
+  Technology (FINESST) award number 80NSSC20K1536 during the completion of this
+  work. E.D. was partially supported by HST grants HST-AR-16363.001 and
+  HST-AR-16602.006-A and by NASA Award NASA 80NSSC22K0761. J.Q. acknowledges
+  support from NSF grant AST-2008490. L.V.S. is grateful for financial support
+  from NASA ATP 80NSSC20K0566, NSF AST 1817233 and NSF CAREER 1945310 grants.
+  P.T. acknowledges support from NSF grant AST-1909933, AST-2008490, and NASA
+  ATP Grant 80NSSC20K0502. M.V. acknowledges support through NASA ATP
+  19-ATP19-0019, 19-ATP19-0020, 19-ATP19-0167, and NSF grants AST-1814053,
+  AST-1814259, AST-1909831, AST-2007355 and AST-2107724.
 \end{acknowledgments}

### 2022-11-30 paper-ApJ/ms.tex
Convert SMUGGLE to \SMUGGLE{}

@@
     simulation while the lower panels show a simulation which includes the
-    SMUGGLE model for the interstellar medium. Each panel is $20\,\textrm{kpc}$
+    \SMUGGLE{} model for the interstellar medium. Each panel is $20\,\textrm{kpc}$
     to a side. Columns show different points in time, separated by
     $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
-    length and strength. In the SMUGGLE run, the bar remains at approximately
+    length and strength. In the \SMUGGLE{} run, the bar remains at approximately
     the same length and strength over the course of the simulation. The \Nbody{}
@@
 formation model Stars and MUltiphase Gas in GaLaxiEs
-\citep[SMUGGLE;][]{2019MNRAS.489.4233M}. This disk galaxy exhibits almost no
+\citep[\SMUGGLE{};][]{2019MNRAS.489.4233M}. This disk galaxy exhibits almost no
 evolution in the bar pattern speed over several Gyr when the gas phase is
@@
     As expected, the bar in the \Nbody{} run slows down due to interactions
-    between the bar and the dark matter halo. However, the bar in the SMUGGLE
+    between the bar and the dark matter halo. However, the bar in the \SMUGGLE{}
     run does not slow down and instead remains at a constant pattern speed. The
@@
     The solid lines indicate the torque exerted by the halo in both the \Nbody{}
-    and SMUGGLE cases. The dashed line is the torque exerted by the gas phase in
-    the SMUGGLE run (there is no gas in the \Nbody{} run). Details on how these
+    and \SMUGGLE{} cases. The dashed line is the torque exerted by the gas phase in
+    the \SMUGGLE{} run (there is no gas in the \Nbody{} run). Details on how these
     properties are calculated is given in Section~\ref{ssec:bar_analysis}.}\label{fig:prop}
@@
     at larger radii are possible. Stars are captured on these orbits,
-    lengthening the bar. This process does not occur in the SMUGGLE cases since
+    lengthening the bar. This process does not occur in the \SMUGGLE{} cases since
     the bar pattern speed is not decreasing, and therefore the bar length
@@
     strength increases with time, consistent with previous results showing that
-    the bar strength increases as bars slow down. In the SMUGGLE case (orange),
+    the bar strength increases as bars slow down. In the \SMUGGLE{} case (orange),
     we see that the bar strength remains roughly constant, possibly slightly
@@
 this work, we consider $t=0$ for the \Nbody{} run to be the time at which we
-added the gas phase for the SMUGGLE run (i.e., we ignore the first
+added the gas phase for the \SMUGGLE{} run (i.e., we ignore the first
 $1.5\,\textrm{Gyr}$ of evolution of the \Nbody{} disk when the bar is forming).
@@
 \subsection{Numerical Model}
-We use the Stars and MUltiphase Gas in GaLaxiEs (SMUGGLE) model
+We use the Stars and MUltiphase Gas in GaLaxiEs (\SMUGGLE{}) model
 \citep{2019MNRAS.489.4233M} implemented within the moving-mesh, finite-volume
-hydrodynamics and gravity code \AREPO{} \citep{2010MNRAS.401..791S}. The SMUGGLE
+hydrodynamics and gravity code \AREPO{} \citep{2010MNRAS.401..791S}. The \SMUGGLE{}
 model additionally includes radiative heating and cooling, star formation, and
@@
 treatment. A more detailed description of this model can be found in the
-flagship SMUGGLE paper \citep{2019MNRAS.489.4233M}.
+flagship \SMUGGLE{} paper \citep{2019MNRAS.489.4233M}.
 
@@
 been exhausted after deposition into $64$ neighboring cells. We also employed an
-updated version of SMUGGLE using a new mechanical feedback routine similar to
+updated version of \SMUGGLE{} using a new mechanical feedback routine similar to
 the one described in \citet{2018MNRAS.480..800H}. This updated routine is a
@@
 
-In addition to the SMUGGLE model, we considered a simpler model of the
+In addition to the \SMUGGLE{} model, we considered a simpler model of the
 interstellar medium based upon \citet{2003MNRAS.339..289S}. In this approach,
@@
     \caption{The wake excited in the dark matter halo. The dark matter halo wake
-    is shown in the \Nbody{} case (\textit{middle panel}) and SMUGGLE case
+    is shown in the \Nbody{} case (\textit{middle panel}) and \SMUGGLE{} case
     (\textit{right panel}) after $2.6\,\textrm{Gyr}$ of evolution. The
@@
     measured from the second Fourier component. After the first Gyr, the angle
-    difference in the SMUGGLE case is smaller than in the \Nbody{} case by about
-    a factor of two, reflecting how the dark matter halo in the SMUGGLE case is
+    difference in the \SMUGGLE{} case is smaller than in the \Nbody{} case by about
+    a factor of two, reflecting how the dark matter halo in the \SMUGGLE{} case is
     unable to exert as negative a torque on the bar as in the \Nbody{} case.}
@@
 In the upper panel, we show the pattern speed over time in the \Nbody{} (blue)
-and SMUGGLE (orange) runs. The pattern speed in the \Nbody{} case slows down
-while the pattern speed in the SMUGGLE case remains roughly constant. The
+and \SMUGGLE{} (orange) runs. The pattern speed in the \Nbody{} case slows down
+while the pattern speed in the \SMUGGLE{} case remains roughly constant. The
 slowing down of the pattern speed in the \Nbody{} case is consistent with a long
@@
 
-However, in the SMUGGLE case the pattern speed remains constant. After the first
+However, in the \SMUGGLE{} case the pattern speed remains constant. After the first
 Gyr of evolution, we find that the pattern speed increases by only $\sim10\%$
@@
 an average torque from $1$ to $4\,\textrm{Gyr}$ of $-58.0$ in units of
-$10^{10}M_{\odot}\,(\textrm{km}/\textrm{s})^2$. The halo in the SMUGGLE case
+$10^{10}M_{\odot}\,(\textrm{km}/\textrm{s})^2$. The halo in the \SMUGGLE{} case
 exerts a similar negative torque on the bar in the first Gyr of evolution, but
 after that the halo exerts a much weaker torque on the bar, averaging only
-$-7.8$ in the same units and over the same time interval. The gas in the SMUGGLE
+$-7.8$ in the same units and over the same time interval. The gas in the \SMUGGLE{}
 case exerts a steady positive torque averaging $11.7$ over $1\,\textrm{Gyr}$ in
@@
 Fig.~\ref{fig:strength} shows that the length of the bar in the \Nbody{} case
-grows over time while it remains roughly constant in the SMUGGLE case. This is
+grows over time while it remains roughly constant in the \SMUGGLE{} case. This is
 also consistent with previous numerical work, which found that bars tend to grow
@@
 \Nbody{} simulations which showed a clear correlation between the bar pattern
-speed and the bar strength \citep[e.g.,][]{2003MNRAS.341.1179A}. In the SMUGGLE
+speed and the bar strength \citep[e.g.,][]{2003MNRAS.341.1179A}. In the \SMUGGLE{}
 case, we see that the bar strength has an initial drop but then remains at a
 roughly constant, but slightly decreasing, strength. This is consistent with the
-pattern speed in the SMUGGLE case being roughly constant or slightly increasing.
+pattern speed in the \SMUGGLE{} case being roughly constant or slightly increasing.
 
@@
 \subsection{Pattern Speed Evolution}
-The lack of evolution in the pattern speed of the SMUGGLE case (seen in
+The lack of evolution in the pattern speed of the \SMUGGLE{} case (seen in
 Fig.~\ref{fig:prop}) is intimately tied to the sudden decrease in torque exerted
@@
 is plotted in the left panel of Fig.~\ref{fig:wake}, which shows that the
-angle offset is larger in the \Nbody{} case than in the SMUGGLE case by about a
+angle offset is larger in the \Nbody{} case than in the \SMUGGLE{} case by about a
 factor of two. The center and right panels of Fig.~\ref{fig:wake} show the halo
-wake with respect to the location of the bar in the \Nbody{} (center) and SMUGGLE
+wake with respect to the location of the bar in the \Nbody{} (center) and \SMUGGLE{}
 (right) cases at one point in time.
@@
 the dark matter halo can contribute to a wake. However, this does not explain
-why the pattern speed in the SMUGGLE case is nearly constant over several Gyr.
+why the pattern speed in the \SMUGGLE{} case is nearly constant over several Gyr.
 Naively, it would be a coincidence that the bar pattern speed remains constant
-in the SMUGGLE case, resulting from a chance cancellation of the halo and gas
+in the \SMUGGLE{} case, resulting from a chance cancellation of the halo and gas
 torques. However, a constant pattern speed in the presence of gas has been
@@
 $I_6=8$ is a good approximation to our fiducial disk model. In code units, the
-moment of inertia of the SMUGGLE bar (i.e., the particles classified as being in
+moment of inertia of the \SMUGGLE{} bar (i.e., the particles classified as being in
 the bar) is about $2$. This is a factor of $4$ smaller than our fiducial value
@@
 line). When a gas torque is included (orange lines), we find that the halo's
-torque becomes much weaker, similar to what we found in the SMUGGLE case. The
+torque becomes much weaker, similar to what we found in the \SMUGGLE{} case. The
 gas applies a steady positive torque by construction. Therefore, the exhaustion
@@
     \caption{Pattern speed evolution of a smooth ISM model. This evolution is
-    shown for the fiducial disk in the \Nbody{} (blue), SMUGGLE (orange), and
+    shown for the fiducial disk in the \Nbody{} (blue), \SMUGGLE{} (orange), and
     smooth ISM (red) cases. The smooth ISM model is an older model for the ISM
     which treats its multiphase nature in a subgrid fashion
-    \citep{2003MNRAS.339..289S}. This fundamentally differs from the SMUGGLE
+    \citep{2003MNRAS.339..289S}. This fundamentally differs from the \SMUGGLE{}
     model, which explicitly resolves the hot and cold phases of the ISM
     \citep{2019MNRAS.489.4233M}. The pattern speed in the smooth ISM case is
-    broadly similar to the evolution in the SMUGGLE case. This shows that the
+    broadly similar to the evolution in the \SMUGGLE{} case. This shows that the
     stability of the pattern speed is not simply a result of our assumed model
@@
     models with $\tau_{\textrm{gas}}=0$ and $20$, which most closely resemble
-    our $N$-body and SMUGGLE disks, respectively. This figure ought to be
+    our $N$-body and \SMUGGLE{} disks, respectively. This figure ought to be
     compared to the lower panel of Fig.~\ref{fig:prop}. Overall, we find good
@@
 density projection of the stellar disk and bulge (including newly formed stars)
-from the SMUGGLE model after $1\,\text{Gyr}$ of evolution in a frame such that
+from the \SMUGGLE{} model after $1\,\text{Gyr}$ of evolution in a frame such that
 the bar is aligned with the $x$-axis. The \textit{middle} panel shows a
@@
 We performed an experiment to test this hypothesis by forcing the stellar disk
-in the SMUGGLE run to rotate at a constant angular rate and measuring the
+in the \SMUGGLE{} run to rotate at a constant angular rate and measuring the
 torque on the bar by the gas phase at different rotation rates. The result of
@@
     larger net torque than faster bars. The setup of the simulations used here
-    is identical to the SMUGGLE case discussed earlier, except the \Nbody{} disk
+    is identical to the \SMUGGLE{} case discussed earlier, except the \Nbody{} disk
     is rotated as a solid body with a constant angular
@@
 particles), we assumed that they have an age of $5\,\textrm{Gyr}$ and solar
-metallicity. For star particles in the SMUGGLE simulation which are newly
+metallicity. For star particles in the \SMUGGLE{} simulation which are newly
 formed, we used their recorded ages and metallicities.
 
-For the SMUGGLE simulations, we included absorption and emission from dust. We
+For the \SMUGGLE{} simulations, we included absorption and emission from dust. We
 assumed the THEMIS (The Heterogeneous dust Evolution Model for Interstellar
@@
 plot the circular velocity curve in Fig.~\ref{fig:vcirc} compared to
-observational estimates \citep{2019ApJ...871..120E}. The SMUGGLE disk (which
+observational estimates \citep{2019ApJ...871..120E}. The \SMUGGLE{} disk (which
 includes additional mass in the form of gas) has a slightly higher circular
@@
     \caption{The circular velocity curve of our initial setups. This
-    curve is shown for the \Nbody{} run (blue) and the SMUGGLE run (orange)
+    curve is shown for the \Nbody{} run (blue) and the \SMUGGLE{} run (orange)
     compared to observational estimates for the Milky
@@
     both runs is marginally larger than the Milky Way's, but still comparable. The
-    SMUGGLE circular velocity curve is larger than the \Nbody{} curve due to the
+    \SMUGGLE{} circular velocity curve is larger than the \Nbody{} curve due to the
     additional mass in the gas phase.}

### 2022-11-30 paper-ApJ/ms.tex
Cite Fraser-Mackelvie 2020

@@
 present work with these two well-executed manuscripts.
+
+\citet{2020MNRAS.495.4158F} find that quenched galaxies tend to host longer bars
+than star-forming galaxies. This provides some support for our proposed
+mechanism since there is evidence quenching can occur through gas depletion
+\citep[e.g.][]{2021Natur.597..485W}. However, this correlation could be
+explained simply by the fact that longer bars ought to be more efficient at
+quenching their host galaxies \citep[e.g][]{2015AA...580A.116G}.
 

### 2022-12-13 paper-ApJ/ms.tex
Add some more discussion of past work

@@
 studies have shown that a spherical potential, such as a stellar bulge or dark
-matter halo, acts to stabilize stellar disks against bar formtaion
+matter halo, acts to stabilize stellar disks against bar formation
 \citep[e.g.,][]{1973ApJ...186..467O, 1976AJ.....81...30H}.
@@
 We have performed a simulation of a disk galaxy using the finite-volume,
-gravito-hydrodynamics code \AREPO{} \citep{2010MNRAS.401..791S}. We use the galaxy
-formation model Stars and MUltiphase Gas in GaLaxiEs
+gravito-hydrodynamics code \AREPO{} \citep{2010MNRAS.401..791S}. We use the
+galaxy formation model Stars and MUltiphase Gas in GaLaxiEs
 \citep[\SMUGGLE{};][]{2019MNRAS.489.4233M}. This disk galaxy exhibits almost no
@@
 accounted for and properly modeled. This behavior has been observed in a few
-previous works \citep{1993AA...268...65F, 2010ApJ...719.1470V,
-2014MNRAS.438L..81A}. However, we are the first to demonstrate that the behavior
-ought to be present in a large fraction of galaxies, since it occurs in our
-simulation with gas fractions as low as about $5\%$. We are also the first to
-provide a physical explanation of the behavior with clear observational
-predictions.
+previous works \citep{1993AA...268...65F, 2007ApJ...666..189B,
+2009ApJ...707..218V, 2010ApJ...719.1470V, 2014MNRAS.438L..81A}. However, we are
+the first to demonstrate that the behavior ought to be present in a large
+fraction of galaxies, since it occurs in our simulation with gas fractions as
+low as about $5\%$. We are also the first to provide a physical explanation of
+the behavior with clear observational predictions.
 
@@
 
+This halo exhaustion is similar to an effect called ``metastability'' discussed
+in the literature \citep{2003MNRAS.345..406V, 2006ApJ...639..868S}. We note,
+however, that metastability is a numerical effect arising from low resolution
+and a poor sampling of resonances, and is not expected to occur in the real
+universe.
+
 We can test this interpretation by measuring the angle offset between the halo
@@
 the dark matter halo can contribute to a wake. However, this does not explain
-why the pattern speed in the \SMUGGLE{} case is nearly constant over several Gyr.
-Naively, it would be a coincidence that the bar pattern speed remains constant
-in the \SMUGGLE{} case, resulting from a chance cancellation of the halo and gas
-torques. However, a constant pattern speed in the presence of gas has been
-observed in a few simulations of barred galaxies with
-gas \citep{1993AA...268...65F, 2010ApJ...719.1470V, 2014MNRAS.438L..81A}. Previous
-work has argued this is due to the bar torquing gas inwards, but no explanation
-has been given for why it might remain constant.
+why the pattern speed in the \SMUGGLE{} case is nearly constant over several
+Gyr. Naively, it would be a coincidence that the bar pattern speed remains
+constant in the \SMUGGLE{} case, resulting from a chance cancellation of the
+halo and gas torques. However, a constant pattern speed in the presence of gas
+has been observed in a few simulations of barred galaxies with gas
+\citep{1993AA...268...65F, 2007ApJ...666..189B, 2009ApJ...707..218V,
+2010ApJ...719.1470V, 2014MNRAS.438L..81A}.
+
+\citet{1993AA...268...65F} argue this behavior is due to the steepening of the
+circular velocity curve in the central region as the bar drives gas to the
+center. \citet{2009ApJ...707..218V} argue this behavior occurs when the
+corotation resonance is larger than the disk radius, but we observe the behavior
+when the corotation radius is well within the disk.
+
+% Previous work has argued this is due to the bar torquing
+% gas inwards, but no explanation has been given for why it might remain constant.
 

### 2023-02-16 paper-ApJ/ms.tex
update pdf

@@
 \begin{document}
-
 \title{Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down}
@@
   Elongated bar-like features are ubiquitous in galaxies, occurring at the
-  centers of approximately two-thirds of spiral disks.  Due to gravitational
+  centers of approximately two-thirds of spiral disks in the nearby Universe.  
+  Due to gravitational
   interactions between the bar and the other components of galaxies, it is
@@
 \citep{1984MNRAS.209..729T, 1985MNRAS.213..451W}, this process is generic for
-any non-axisymmetric disturbance \citep{1972MNRAS.157....1L}. For an old but
-still useful review of bar dynamics, see \citet{1993RPPh...56..173S}
+any non-axisymmetric disturbance \citep{1972MNRAS.157....1L}. (For an old but
+still useful review of bar dynamics, see \citet{1993RPPh...56..173S}.)
 
@@
 
+In the last decade, stellar bars have been studied mainly in the
+context of the instability processes that lead to their formation and their
+ability to drive gas toward the galaxy center and contribute to the formation of
+supermassive black holes (SMBH). The primary interest of these studies mainly
+was in the loss of angular momentum of the gas and the associated  gaseous flow
+down to the inner disk, possibly forming a central mass concentration
+\citep{2010ApJ...719.1470V}, and fueling the central SMBH \citep[e.g.][]{1989Natur.338...45S, 1990Natur.345..679S}. A revisit of the more general problem of
+galactic bar properties and formation, including the case of disk galaxies with
+very large gas fraction is timely since it is clear now that galactic disks show
+massive bars already at redshift $z>2$ \citep{2022arXiv221008658G}. At
+that time the universe was 2.5 billion year old and galaxies might have as high
+gas fraction as 80\% \citep{2020ARAA..58..157T}. Furthermore, unlike
+nearby disk galaxies, the high-redshift disks also continuously accrete cold gas
+from the cosmic web, making the formation, stability and evolution of
+non-axisymmetric features a key question to address since they can play a
+fundamental role in the more general problem of disk galaxy evolution.
+
 We have performed a simulation of a disk galaxy using the finite-volume,
@@
 previous works \citep{1993AA...268...65F, 2007ApJ...666..189B,
-2009ApJ...707..218V, 2010ApJ...719.1470V, 2014MNRAS.438L..81A}. However, we are
-the first to demonstrate that the behavior ought to be present in a large
-fraction of galaxies, since it occurs in our simulation with gas fractions as
-low as about $5\%$. We are also the first to provide a physical explanation of
-the behavior with clear observational predictions.
+2009ApJ...707..218V, 2010ApJ...719.1470V, 2014MNRAS.438L..81A}.
+
+Here we show that this phenomenon occurs in galaxies with gas fraction
+as low as about $5\%$, making this feature a likely characteristic of  Milky
+Way-type galaxies. This behavior will be explored in more general disks with
+larger gas fractions than the Milky Way in a forthcoming paper.
+% We are also the first to provide a physical explanation of
+%the behavior with clear observational predictions.
+We also provide a physical explanation of the redistribution of the angular
+momentum among the dark halo wake, the stellar bar, and the disk gas component.
 
@@
 Fig.~\ref{fig:prop}) is intimately tied to the sudden decrease in torque exerted
-on the bar by the dark matter halo. We argue that this can be understood in
-terms of the halo wake mechanism. In the \Nbody{} case, halo material that is
-resonant with the bar will form a wake which lags behind and exerts a negative
-torque on the bar, slowing it down \citep{1984MNRAS.209..729T,
-1985MNRAS.213..451W, 1992ApJ...400...80H}.\footnote{Since the bar is not a solid
+on the bar by the dark matter halo.
+
+%We argue that this can be understood in terms of the halo wake mechanism. 
+We interpret this behavior in terms of the halo wake mechanism. In the
+\Nbody{} case, a well known phenomenon is that the halo material resonant with
+the bar forms a wake, and this wake lags behind \citep{1984MNRAS.209..729T,
+1985MNRAS.213..451W, 1992ApJ...400...80H} and exerts a negative torque on the
+bar, slowing it down (see Fig.~\ref{fig:wake} below).\footnote{Since the bar is not a solid
 body, it is not guaranteed that a negative torque will slow it down - e.g. a
@@
 negative torque induced by a halo wake by slowing down.} As the bar slows down,
-the location of the resonances in phase space changes, allowing halo material
-newly resonant with the bar to participate in forming the wake. However, the gas
-is a reliable source of positive torque on the bar, speeding the bar up. This
-stops the resonance location from changing such that the halo cannot reinforce
-the wake, arresting the process by which the halo can slow the bar down. We term
-this process ``halo exhaustion.''
-
-This halo exhaustion is similar to an effect called ``metastability'' discussed
-in the literature \citep{2003MNRAS.345..406V, 2006ApJ...639..868S}. We note,
-however, that metastability is a numerical effect arising from low resolution
-and a poor sampling of resonances, and is not expected to occur in the real
-universe.
-
-We can test this interpretation by measuring the angle offset between the halo
+the location of the resonances in the phase space changes (see Fig.~12 and Table
+1 in \citet{2020ApJ...890..117D}) allowing halo material newly resonant with the
+bar to participate in the formation of the wake. However, the gas is also a
+reliable source of positive torque on the bar, speeding the bar up. In turn,
+this stops the location of the resonance from changing such that the halo cannot
+reinforce the wake, therefore arresting the process by which the halo can slow
+the bar down. We term this process ``halo exhaustion.'' Note that the resonances
+can be maintained even in the presence of the dissipative gaseous component if
+the gas is continuously accreted from the external environment, as is the case
+for massive disk galaxies at high redshift.
+
+%In the \Nbody{} case, halo material that is
+%resonant with the bar will form a wake which lags behind and exerts a negative 
+%torque on the bar, slowing it down \citep{1984MNRAS.209..729T,
+%1985MNRAS.213..451W, 1992ApJ...400...80H}.\footnote{Since the bar is not a solid
+%body, it is not guaranteed that a negative torque will slow it down - e.g. a
+%negative torque could shred the bar, reducing its moment of inertia without
+%changing its pattern speed. However, the bar seems to empirically respond to a
+%negative torque induced by a halo wake by slowing down.} As the bar slows down,
+%the location of the resonances in phase space changes, allowing halo material
+%newly resonant with the bar to participate in forming the wake. However, the gas
+%is a reliable source of positive torque on the bar, speeding the bar up. This
+%stops the resonance location from changing such that the halo cannot reinforce
+%the wake, arresting the process by which the halo can slow the bar down. We term
+%this process ``halo exhaustion.''
+
+This halo exhaustion process is similar to the ``metastability'' effect, which
+has been previously discussed in the literature \citep{2003MNRAS.345..406V,
+2006ApJ...639..868S}. Finally, we also note that these authors, in particular,
+observe that the effects of numerical resolution in the simulations adopted to
+explore these mechanisms have yet to be fully explored and could play a role in
+the observed phenomenology of the simulations. We plan to address these issues
+in future dedicated work.
+
+We test this interpretation by measuring the angle offset between the halo
 wake and the bar. If the wake and the bar are aligned (i.e., there is no angle
@@
 \citep{1990ApJ...356..359H} with mass $10^{12}\,\Msun$ and a scale length of
-$26.2\,\textrm{kpc}$. For the stellar disk, we used a MiyaMoto-Nagai disk
+$26.2\,\textrm{kpc}$. For the stellar disk, we used a Miyamoto-Nagai disk
 \citep{1975PASJ...27..533M} with mass $4.8\times10^{10}\,\Msun$, radial scale
@@
 \end{document}
-
-% End of file `sample631.tex'.

### 2023-02-16 analysis/in_bar/compute_in_bar.py

@@ -1,0 +1,116 @@
+import numpy as np
+import arepo
+import sys
+from tqdm import tqdm
+import astropy.units as u
+import h5py as h5
+import glob
+import os
+from numba import njit
+import re
+from sklearn.cluster import KMeans
+
+from joblib import Parallel, delayed
+
+def read_snap(path, idx, parttype=[0], fields=['Coordinates', 'Masses', 'Velocities', 'ParticleIDs']):
+    
+    fname = path + '/output'
+    
+    return arepo.Snapshot(fname, idx, parttype=parttype, fields=fields, combineFiles=True)
+
+@njit
+def process_id(ID, bar_metr, t, tlist, nsnap):
+    in_bar = np.full(nsnap, 0, dtype=np.bool_)
+    Naps = t.shape[0]
+    for j,idx in enumerate(range(nsnap)):
+        Tanalyze = tlist[idx]
+        key = np.argmin(np.abs(t - Tanalyze))
+
+        if key==0 or key==Naps-1:
+            continue
+        
+        metr = bar_metr[key]
+        
+        if metr[3] == 0.0:
+            continue
+        
+        c0 = metr[1] < np.pi/8.0
+        c1 = metr[2]/metr[3] < 0.22
+        if c0 and c1:
+            in_bar[j] = 1
+    
+    return in_bar
+
+def _in_bar_one_chunk(prefix_in, prefix_out, name, chunk_idx):
+    fin = prefix_in+'/bar_orbit_'+name+'.' + str(chunk_idx) + '.hdf5'
+    h5in = h5.File(fin, mode='r')
+
+    idx_list = np.array(h5in['idx_list'])
+    nsnap = len(idx_list)
+    
+    id_list = np.array(h5in['id_list']).astype(np.int)
+    tlist = np.array(h5in['tlist'])
+    bar_angle = np.array(h5in['bar_angle'])
+    
+    in_bar = np.full((nsnap, len(id_list)), 0, dtype=np.bool_)
+    
+    for i,ID in enumerate(id_list):
+        bar_metr = np.array(h5in['bar_metrics'][str(ID)])
+        t = tlist[bar_metr[:,0].astype(np.int)]
+        in_bar[:,i] = process_id(ID, bar_metr, t, tlist, nsnap)
+    
+    fout = prefix_out + '/in_bar_' + name + '.' + str(chunk_idx) + '.hdf5'
[…]

### 2023-02-16 analysis/orbit_int/job_orbit_int-SMUGGLE-lvl3-single.sh

@@ -1,0 +1,21 @@
+#!/bin/sh
+#SBATCH -p itc_cluster,shared,conroy
+#SBATCH -J oiS50 
+#SBATCH -c 1
+##SBATCH --array=0-1023
+##SBATCH --array=0-15
+#SBATCH -o logs_SMUGGLE-lvl3/OUTPUT.%A.%a.out
+#SBATCH -e logs_SMUGGLE-lvl3/ERROR.%A.%a.err
+#SBATCH --mail-user=angus.beane@cfa.harvard.edu
+#SBATCH --mail-type=BEGIN
+#SBATCH --mail-type=END
+#SBATCH --mail-type=FAIL
+#SBATCH --mem=12G
+#SBATCH -t 0-04:00           # Runtime in D-HH:MM
+
+source ../load-modules.sh
+
+ulimit -c unlimited
+
+python3 compute_orbit_integrations.py phantom-vacuum-Sg20-Rc3.5 lvl3 300 ${1}
+

### 2023-02-16 analysis/orbit_int/job_orbit_int-SMUGGLE-lvl3.sh

@@ -1,0 +1,21 @@
+#!/bin/sh
+#SBATCH -p itc_cluster,shared,conroy
+#SBATCH -J oiS50 
+#SBATCH -c 1
+#SBATCH --array=0-1023
+##SBATCH --array=0-15
+#SBATCH -o logs_SMUGGLE-lvl3/OUTPUT.%A.%a.out
+#SBATCH -e logs_SMUGGLE-lvl3/ERROR.%A.%a.err
+#SBATCH --mail-user=angus.beane@cfa.harvard.edu
+#SBATCH --mail-type=BEGIN
+#SBATCH --mail-type=END
+#SBATCH --mail-type=FAIL
+#SBATCH --mem=12G
+#SBATCH -t 0-04:00           # Runtime in D-HH:MM
+
+source ../load-modules.sh
+
+ulimit -c unlimited
+
+python3 compute_orbit_integrations.py phantom-vacuum-Sg20-Rc3.5 lvl3 300 ${SLURM_ARRAY_TASK_ID}
+

### 2023-02-16 movies/movie_5panel/compute_projections.py

@@ -1,0 +1,145 @@
+import os
+import glob
+
+import numpy as np 
+import h5py as h5
+from joblib import Parallel, delayed
+
+import arepo
+from tqdm import tqdm
+
+def make_projection_snap(path, snapnum, parttype=[0, 2, 3, 4], 
+                         center=np.array([200, 200, 200]), width=30., nres=256):
+
+    sn = arepo.Snapshot(path+'/output', snapnum, parttype=parttype, 
+                        combineFiles=True, fields=['Coordinates', 'Masses'])
+
+    time = sn.Time
+
+    range_xy = [[center[0] - width/2.0, center[0] + width/2.0], [center[1] - width/2.0, center[1] + width/2.0]]
+    range_xz = [[center[0] - width/2.0, center[0] + width/2.0], [center[2] - width/2.0, center[2] + width/2.0]]
+    range_yz = [[center[1] - width/2.0, center[1] + width/2.0], [center[2] - width/2.0, center[2] + width/2.0]]
+
+    surf = (width/nres)**(2.0)
+
+    heatmap_xy_out = []
+    heatmap_xz_out = []
+    heatmap_yz_out = []
+    for pt in parttype:
+        if sn.NumPart_Total[pt] == 0:
+            heatmap_xy_out.append(np.zeros((nres, nres)))
+            heatmap_xz_out.append(np.zeros((nres, nres)))
+            heatmap_yz_out.append(np.zeros((nres, nres)))
+            continue
+
+        part = getattr(sn, 'part'+str(pt))
+
+        x = part.pos[:,0]
+        y = part.pos[:,1]
+        z = part.pos[:,2]
+
+        xbool = np.logical_and(x > center[0] - width/2.0, x < center[0] + width/2.0)
+        ybool = np.logical_and(y > center[1] - width/2.0, y < center[1] + width/2.0)
+        zbool = np.logical_and(z > center[2] - width/2.0, z < center[2] + width/2.0)
+
+        keys = np.logical_and(np.logical_and(xbool, ybool), zbool)
+
+        if sn.MassTable[pt] > 0:
+            weights = None
+            postfac = sn.MassTable[pt] / surf
+        else:
+            weights = part.Masses[keys] / surf
+            postfac = 1.0
+
+        heatmap_xy, _, _ = np.histogram2d(x[keys], y[keys], bins=(nres, nres), range=range_xy, weights=weights)
[…]

### 2023-02-16 movies/movie_corot/bar_angle.py

@@ -1,0 +1,96 @@
+import numpy as np
+import astropy.units as u
+import re
+
+def get_bar_angle(phi, firstkey):
+    out = np.zeros(len(phi))
+
+    # set the first bar angle
+    first_bar_angle = phi[firstkey]/2.0
+    out[firstkey] = first_bar_angle
+    
+    # set all subsequent angles
+    for i in np.arange(firstkey+1, len(out)):
+        dphi = phi[i] - phi[i-1]
+        if dphi < 0:
+            dphi += 2.*np.pi
+        out[i] = out[i-1] + dphi/2.0
+
+    # set all previous angles to be the bar angle
+    for i in np.arange(0, firstkey):
+        out[i] = first_bar_angle
+
+    return out
+
+def get_sorted_keys(dat):
+    keys = list(dat.keys())
+    # only keep keys that are snapshot keys
+    keys = [k for k in keys if 'snapshot' in k]
+
+    # extract and sort indices
+    indices = [int(re.findall(r'\d?\d?\d\d\d', k)[0]) for k in keys]
+    sorted_arg = np.argsort(indices)
+    keys_sorted = [keys[i] for i in sorted_arg]
+
+    return keys_sorted
+
+def get_A2_angle(dat, keys, Rbin=5):
+    Rlist = np.array([np.array(dat[k]['Rlist']) for k in keys])
+    A2r = np.array([np.array(dat[k]['A2r']) for k in keys])
+    A2i = np.array([np.array(dat[k]['A2i']) for k in keys])
+    phi = np.arctan2(A2i, A2r)
+    phi = phi[:,Rbin]
+    R_at_Rbin = Rlist[:,Rbin]
+    
+    time = np.array(dat['time'])
+
+    return time, R_at_Rbin, phi
+
+def evaluate_polynomial(pfit, n, time, bar_angle_firstkey, firstkey):
+    pfit_n = pfit[n]
+    poly_bar_angle = np.zeros(len(time))
+    poly_pattern_speed = np.zeros(len(time))
+
+    for i in range(n+1):
+        ba = pfit_n[i] * time ** (n-i)
+        poly_bar_angle[firstkey:] += ba[firstkey:]
+        ps = (n-i) * pfit_n[i] * time**(n-1-i)
+        poly_pattern_speed[firstkey:] += ps[firstkey:]
+
+    poly_bar_angle[:firstkey] += bar_angle_firstkey
+
[…]

### 2023-02-16 plots/bar_decomp/plot_bar_decomp.py

@@ -1,0 +1,231 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import arepo
+import h5py as h5
+import matplotlib as mpl
+import glob
+from tqdm import tqdm
+import copy
+
+from matplotlib import rc
+
+mpl.use('Agg')
+
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
+rc('text', usetex=True)
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+
+columnwidth = 244.0 * 0.035145980349999517 # convert to cm
+textwidth = 508.0 * 0.035145980349999517 # convert to cm
+
+snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
+bprop_path = '/n/holylfs05/LABS/hernquist_lab/Users/abeane/gasbar/analysis/bar_prop/data/'
+
+# names
+Nbody = 'Nbody'
+phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
+
+lvl = 'lvl3'
+
+def read_bar_angle(name, lvl):
+    t = h5.File(bprop_path + 'bar_prop_' + name + '-' + lvl + '.hdf5', mode='r')
+    out = t['bar_angle'][:]
+    t.close()
+
+    return out
+
+def get_center(name):
+    if 'Nbody' in name:
+        center = np.array([0., 0., 0.])
+    else:
+        center = np.array([200., 200., 200.])
+
+    return center
+
+def get_mass(lvl):
+    ref = 7.5E3 / 1E10
+    if lvl=='lvl4':
+        return ref * 8
+    if lvl=='lvl3':
+        return ref
+    if lvl=='lvl2':
+        return ref / 8
+    else:
+        print('lvl not recognized')
+        sys.exit(-1)
+
+def rotate_pos(pos, ang):
+
+    Rmat = np.array([[np.cos(ang), -np.sin(ang), 0.0],
+                     [np.sin(ang),  np.cos(ang), 0.0],
+                     [0.0,         0.0,          1.0]])
+    
+    pos = np.swapaxes(pos, 0, 1)
+    pos = np.matmul(Rmat, pos)
+    pos = np.swapaxes(pos, 0, 1)
+    
+    return pos
+
[…]

### 2023-02-16 plots/ps_fgas/plot_ps_fgas.py

@@ -1,0 +1,169 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import arepo
+import h5py as h5
+import matplotlib as mpl
+from matplotlib import rc
+import glob
+from tqdm import tqdm
+from scipy.signal import savgol_filter
+import re
+import astropy.units as u
+
+mpl.use('Agg')
+
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
+rc('text', usetex=True)
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+
+columnwidth = 244.0 * 0.035145980349999517 # convert to cm
+textwidth = 508.0 * 0.035145980349999517 # convert to cm
+
+snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
+bprop_path = '/n/home01/abeane/starbar/analysis/bar_prop/data/'
+torque_path = '/n/home01/abeane/starbar/analysis/torques/data/'
+
+tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
+        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
+
+# names
+Nbody = 'Nbody'
+phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
+phS15R35 = 'phantom-vacuum-Sg15-Rc3.5'
+phS10R35 = 'phantom-vacuum-Sg10-Rc3.5'
+phS08R35 = 'phantom-vacuum-Sg08-Rc3.5'
+phS05R35 = 'phantom-vacuum-Sg05-Rc3.5'
+
+lvl = 'lvl3'
+
+def read_fourier(name, lvl, basepath='/n/home01/abeane/starbar/analysis/'):
+    f = h5.File(basepath+'/fourier_component/data/fourier_'+name+'-'+lvl+'.hdf5', mode='r')
+    return f
+
+def read_snap(idx, name, lvl, parttype=[0, 2, 3, 4], fields=['Coordinates', 'Masses']):
+    sn = arepo.Snapshot(snap_path+name+'/'+lvl+'/output', idx, combineFiles=True, parttype=parttype,
+                        fields=fields)
+    return sn
+
+def get_sorted_keys(dat):
+    keys = list(dat.keys())
[…]

### 2023-02-16 plots/ps_late_start/plot_ps_late_start.py

@@ -1,0 +1,159 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import arepo
+import h5py as h5
+import matplotlib as mpl
+from matplotlib import rc
+import glob
+from tqdm import tqdm
+from scipy.signal import savgol_filter
+import re
+import astropy.units as u
+
+mpl.use('Agg')
+
+columnwidth = 244.0 * 0.035145980349999517 # convert to cm
+textwidth = 508.0 * 0.035145980349999517 # convert to cm
+
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
+rc('text', usetex=True)
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+
+snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
+bprop_path = '/n/home01/abeane/starbar/analysis/bar_prop/data/'
+torque_path = '/n/home01/abeane/starbar/analysis/torques/data/'
+
+tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
+        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
+
+# names
+Nbody = 'Nbody'
+phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
+phS15R35 = 'phantom-vacuum-Sg15-Rc3.5'
+phS10R35 = 'phantom-vacuum-Sg10-Rc3.5'
+phS08R35 = 'phantom-vacuum-Sg08-Rc3.5'
+phS05R35 = 'phantom-vacuum-Sg05-Rc3.5'
+
+lvl = 'lvl3'
+
+def read_fourier(name, lvl, basepath='/n/home01/abeane/starbar/analysis/'):
+    f = h5.File(basepath+'/fourier_component/data/fourier_'+name+'-'+lvl+'.hdf5', mode='r')
+    return f
+
+def read_snap(idx, name, lvl, parttype=[0, 2, 3, 4], fields=['Coordinates', 'Masses']):
+    sn = arepo.Snapshot(snap_path+name+'/'+lvl+'/output', idx, combineFiles=True, parttype=parttype,
+                        fields=fields)
+    return sn
+
+def get_sorted_keys(dat):
+    keys = list(dat.keys())
+    # only keep keys that are snapshot keys
[…]

### 2023-02-16 plots/vcirc/plot_vcirc.py

@@ -1,0 +1,96 @@
+from re import S
+import numpy as np
+import matplotlib.pyplot as plt
+import arepo
+import h5py as h5
+import matplotlib as mpl
+from matplotlib import rc
+import glob
+from tqdm import tqdm
+from scipy.signal import savgol_filter
+from numba import njit
+from scipy.interpolate import interp1d
+from scipy.ndimage import gaussian_filter
+
+import agama
+agama.setUnits(mass=1E10, length=1, velocity=1)
+
+mpl.use('Agg')
+
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
+rc('text', usetex=True)
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+
+columnwidth = 244.0 * 0.035145980349999517 # convert to cm
+textwidth = 508.0 * 0.035145980349999517 # convert to cm
+
+snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
+agama_pot_path = '/n/home01/abeane/starbar/analysis/agama_pot/data/'
+
+tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
+        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
+
+# names
+Nbody = 'Nbody'
+phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
+
+lvl = 'lvl3'
+
+def read_snap(idx, name, lvl, parttype=[2, 3, 4], fields=['Coordinates', 'Masses']):
+    sn = arepo.Snapshot(snap_path+name+'/'+lvl+'/output', idx, combineFiles=True, parttype=parttype,
+                        fields=fields)
+    return sn
+
+def read_bar_prop(name, lvl):
+    t = h5.File(bprop_path + 'bar_prop_' + name + '-' + lvl + '.hdf5', mode='r')
+    out = {}
+    for key in t['bar_prop'].keys():
+        out[key] = t['bar_prop'][key][:]
+    
+    out['bar_angle'] = t['bar_angle'][:]
+    t.close()
+    
+    return out
+
+def read_agama_pot(idx, name, lvl):
[…]

### 2023-02-16 test_part/ics/GalIC/src/allocate.c

@@ -1,0 +1,97 @@
+/*******************************************************************************
+ * This file is part of the GALIC code developed by D. Yurin and V. Springel.
+ *
+ * Copyright (c) 2014
+ * Denis Yurin (denis.yurin@h-its.org) 
+ * Volker Springel (volker.springel@h-its.org)
+ *******************************************************************************/
+#include <mpi.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <math.h>
+
+#include "allvars.h"
+#include "proto.h"
+
+
+
+
+
+
+
+/* This routine allocates memory for
+ * particle storage, both the collisionless and the SPH particles.
+ * The memory for the ordered binary tree of the timeline
+ * is also allocated.
+ */
+void allocate_memory(void)
+{
+  int NTaskTimesThreads;
+
+  NTaskTimesThreads = MaxThreads * NTask;
+
+  Exportflag = (int *) mymalloc("Exportflag", NTaskTimesThreads * sizeof(int));
+  Exportindex = (int *) mymalloc("Exportindex", NTaskTimesThreads * sizeof(int));
+  Exportnodecount = (int *) mymalloc("Exportnodecount", NTaskTimesThreads * sizeof(int));
+
+  Send_count = (int *) mymalloc("Send_count", sizeof(int) * NTaskTimesThreads);
+  Send_offset = (int *) mymalloc("Send_offset", sizeof(int) * NTaskTimesThreads);
+  Recv_count = (int *) mymalloc("Recv_count", sizeof(int) * NTask);
+  Recv_offset = (int *) mymalloc("Recv_offset", sizeof(int) * NTask);
+
+  Send_count_nodes = (int *) mymalloc("Send_count_nodes", sizeof(int) * NTask);
+  Send_offset_nodes = (int *) mymalloc("Send_offset_nodes", sizeof(int) * NTask);
+  Recv_count_nodes = (int *) mymalloc("Recv_count_nodes", sizeof(int) * NTask);
+  Recv_offset_nodes = (int *) mymalloc("Recv_offset_nodes", sizeof(int) * NTask);
+
+  Mesh_Send_count = (int *) mymalloc("Mesh_Send_count", sizeof(int) * NTask);
[…]

### 2023-02-16 test_part/ics/GalIC/src/bulge.c

@@ -1,0 +1,100 @@
+/*******************************************************************************
+ * This file is part of the GALIC code developed by D. Yurin and V. Springel.
+ *
+ * Copyright (c) 2014
+ * Denis Yurin (denis.yurin@h-its.org) 
+ * Volker Springel (volker.springel@h-its.org)
+ *******************************************************************************/
+#include <mpi.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <math.h>
+#include <gsl/gsl_math.h>
+
+#include "allvars.h"
+#include "proto.h"
+
+
+
+
+/* this function returns a new random coordinate for the bulge
+ */
+void bulge_get_fresh_coordinate(double *pos)
+{
+  double r;
+
+  do
+    {
+      double q = gsl_rng_uniform(random_generator);
+
+      if(q > 0)
+        r = All.Bulge_A * (q + sqrt(q)) / (1 - q);
+      else
+        r = 0;
+    }
+  while(r > All.Rmax);
+
+  double phi = gsl_rng_uniform(random_generator) * M_PI * 2;
+  double theta = acos(gsl_rng_uniform(random_generator) * 2 - 1);
+
+  pos[0] = r * sin(theta) * cos(phi);
+  pos[1] = r * sin(theta) * sin(phi);
+  pos[2] = r * cos(theta) / All.BulgeStretch;
+}
+
+
+/* return the bulge density for the given coordinate
+ */
+double bulge_get_density(double *pos)
+{
+  double r = sqrt(pos[0] * pos[0] + pos[1] * pos[1] + All.BulgeStretch * All.BulgeStretch * pos[2] * pos[2]);
+
+  return All.BulgeStretch * All.Bulge_Mass / (2 * M_PI) * All.Bulge_A / (r + 1.0e-6 * All.Bulge_A) / pow(r + All.Bulge_A, 3);
+}
+
+
+/* Note that the other functions below will only be called in a meaningfull for a spherical system */
+
+
+double bulge_get_mass_inside_radius(double r)
+{
+  if(All.Bulge_Mass > 0)
+    return All.Bulge_Mass * pow(r / (r + All.Bulge_A), 2);
+  else
+    return 0;
+}
+
+
+
+double bulge_get_potential(double *pos)
+{
[…]

### 2023-02-16 test_part/ics/GalIC/src/domain/pqueue.c

@@ -1,0 +1,258 @@
+/*******************************************************************************
+ * This file is part of the GALIC code developed by D. Yurin and V. Springel.
+ *
+ * Copyright (c) 2014
+ * Denis Yurin (denis.yurin@h-its.org) 
+ * Volker Springel (volker.springel@h-its.org)
+ *******************************************************************************/
+
+/* V. Springel modified some of the memory allocation calls to inline it with
+ * our internal memory handler.
+ */
+
+
+#include <stdlib.h>
+#include <stdio.h>
+#include <string.h>
+
+#include "pqueue.h"
+#include "../allvars.h"
+#include "../proto.h"
+
+#define left(i)   ((i) << 1)
+#define right(i)  (((i) << 1) + 1)
+#define parent(i) ((i) >> 1)
+
+
+pqueue_t *pqueue_init(size_t n,
+		      pqueue_cmp_pri_f cmppri, pqueue_get_pri_f getpri, pqueue_set_pri_f setpri, pqueue_get_pos_f getpos, pqueue_set_pos_f setpos)
+{
+  pqueue_t *q;
+
+  q = mymalloc("q", sizeof(pqueue_t));
+
+  /* Need to allocate n+1 elements since element 0 isn't used. */
+  q->d = mymalloc("q->d", (n + 1) * sizeof(void *));
+
+  q->size = 1;
+  q->avail = q->step = (n + 1);	/* see comment above about n+1 */
+  q->cmppri = cmppri;
+  q->setpri = setpri;
+  q->getpri = getpri;
+  q->getpos = getpos;
+  q->setpos = setpos;
+
+  return q;
+}
+
+
+void pqueue_free(pqueue_t * q)
+{
+  myfree(q->d);
+  myfree(q);
+}
+
+
+size_t pqueue_size(pqueue_t * q)
+{
+  /* queue element 0 exists but doesn't count since it isn't used. */
+  return (q->size - 1);
+}
+
+
+static void bubble_up(pqueue_t * q, size_t i)
+{
+  size_t parent_node;
+  void *moving_node = q->d[i];
+  pqueue_pri_t moving_pri = q->getpri(moving_node);
+
[…]

### 2023-02-16 test_part/ics/GalIC/src/forcetree/gravtree.c

@@ -1,0 +1,741 @@
+/*******************************************************************************
+ * This file is part of the GALIC code developed by D. Yurin and V. Springel.
+ *
+ * Copyright (c) 2014
+ * Denis Yurin (denis.yurin@h-its.org) 
+ * Volker Springel (volker.springel@h-its.org)
+ *******************************************************************************/
+#include <mpi.h>
+#include <stdlib.h>
+#include <string.h>
+#include <math.h>
+
+#include "../allvars.h"
+#include "../proto.h"
+#include "../domain/domain.h"
+
+/*! \file gravtree.c
+ *  \brief main driver routines for gravitational (short-range) force computation
+ *
+ *  This file contains the code for the gravitational force computation by
+ *  means of the tree algorithm. To this end, a tree force is computed for all
+ *  active local particles, and particles are exported to other processors if
+ *  needed, where they can receive additional force contributions. If the
+ *  TreePM algorithm is enabled, the force computed will only be the
+ *  short-range part.
+ */
+void gravity_tree(void);
+void gravity_primary_loop(int thread_id);
+void gravity_secondary_loop(int thread_id);
+int compare_partlist_task_index(const void *a, const void *b);
+int compare_nodelist_task_index_node(const void *a, const void *b);
+
+
+
+
+/*! \brief main driver routine of tree force calculation
+ *
+ *  This routine handles the whole tree force calculation. First it
+ *  build a new force tree force_treebuild() every timestep. This tree is then
+ *  used to calculate a new tree force for every active particle ( gravity_tree() ).
+ */
+void gravity(void)
+{
+  domain_Decomposition();
+
+  force_treeallocate(NumPart, All.MaxPart);
+  force_treebuild(NumPart, 1);
+
+  gravity_tree();
+
+  myfree(Father);
+  myfree(Nextnode);
+  myfree(Tree_Points);
+  force_treefree();
+
+  domain_free();
+}
+
+
+
+/*! \brief This function computes the gravitational forces for all active particles.
+ *
[…]

### 2023-02-16 test_part/ics/GalIC/src/halo.c

@@ -1,0 +1,178 @@
+/*******************************************************************************
+ * This file is part of the GALIC code developed by D. Yurin and V. Springel.
+ *
+ * Copyright (c) 2014
+ * Denis Yurin (denis.yurin@h-its.org) 
+ * Volker Springel (volker.springel@h-its.org)
+ *******************************************************************************/
+#include <mpi.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <math.h>
+#include <gsl/gsl_math.h>
+
+#include "allvars.h"
+#include "proto.h"
+
+
+/* this file contains auxiliary routines for the description of the halo,
+ * here modeled as a Hernquist sphere
+ */
+
+
+
+
+/* this function returns a new random coordinate for the halo
+ */
+void halo_get_fresh_coordinate(double *pos)
+{
+  double r;
+
+  do
+    {
+      double q = gsl_rng_uniform(random_generator);
+
+      if(q > 0)
+        r = All.Halo_A * (q + sqrt(q)) / (1 - q);
+      else
+        r = 0;
+
+      double phi = gsl_rng_uniform(random_generator) * M_PI * 2;
+      double theta = acos(gsl_rng_uniform(random_generator) * 2 - 1);
+      
+      pos[0] = r * sin(theta) * cos(phi);
+      pos[1] = r * sin(theta) * sin(phi);
+      pos[2] = r * cos(theta) / All.HaloStretch;
+
+      r = sqrt(pos[0]*pos[0] + pos[1]*pos[1] + pos[2]*pos[2]);
+    }
+  while(r > All.Rmax);
+}
+
+
+/* return the dark matter halo density at the given coordinate
+ */
+double halo_get_density(double *pos)
+{
+  double r = sqrt(pos[0] * pos[0] + pos[1] * pos[1] + All.HaloStretch * All.HaloStretch * pos[2] * pos[2]);
+
+  return All.HaloStretch * All.Halo_Mass / (2 * M_PI) * All.Halo_A / (r + 1.0e-6 * All.Halo_A) / pow(r + All.Halo_A, 3);
+}
+
+
+/* Note that the other functions below will only be meaningfully called for a spherical system */
+
+
+
[…]

### 2023-02-16 test_part/ics/GalIC/src/io.c

@@ -1,0 +1,1333 @@
+/*******************************************************************************
+ * This file is part of the GALIC code developed by D. Yurin and V. Springel.
+ *
+ * Copyright (c) 2014
+ * Denis Yurin (denis.yurin@h-its.org) 
+ * Volker Springel (volker.springel@h-its.org)
+ *******************************************************************************/
+#include <mpi.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <stdarg.h>
+#include <string.h>
+#include <math.h>
+#include <errno.h>
+#include <sys/stat.h>
+#include <sys/types.h>
+#include <unistd.h>
+#include <sys/time.h>
+#include <time.h>
+
+#include "allvars.h"
+#include "proto.h"
+
+#ifdef HAVE_HDF5
+#include <hdf5.h>
+void write_header_attributes_in_hdf5(hid_t handle);
+#endif
+
+
+
+
+/*! \file io.c
+ *  \brief Output of a snapshot (or an image file) file to disk.
+ */
+
+static int n_type[6];			/**< contains the local (for a single task) number of particles of each type in the snapshot file */
+static long long ntot_type_all[6];	/**< contains the global number of particles of each type in the snapshot file */
+
+
+
+
+
+void output_density_field(int iter)
+{
+  int nbin;
+  double *dout;
+
+  if(!(iter % All.StepsBetweenDump) == 0)
+    return;
+
+  int num = iter / All.StepsBetweenDump;
+
+  int type, lev;
+
+  for(type = 1; type <= 3; type++)
+    {
+      if(type == 1 && All.Halo_N == 0)
+	continue;
+
+      if(type == 2 && All.Disk_N == 0)
+	continue;
+
+      if(type == 3 && All.Bulge_N == 0)
+	continue;
+
+      char buf[1000];
+      sprintf(buf, "%s/densfield_%d_%03d.dat", All.OutputDir, type, num);
+      FILE *fd = fopen(buf, "w");
+      fwrite(&DG_MaxLevel, sizeof(int), 1, fd);
+      for(lev = DG_MaxLevel; lev >= 0; lev--)
+	{
+	  int nbin = (1 << lev);
+	  fwrite(&nbin, sizeof(int), 1, fd);
+	  fwrite(&DGs_MassTarget[type][STACKOFFSET(lev, 0, 0)], sizeof(double), nbin * nbin, fd);
+	}
+
+      /* target, variable resolution */
[…]

### 2023-02-16 test_part/ics/GalIC/src/orbit_response.c

@@ -1,0 +1,150 @@
+/*******************************************************************************
+ * This file is part of the GALIC code developed by D. Yurin and V. Springel.
+ *
+ * Copyright (c) 2014
+ * Denis Yurin (denis.yurin@h-its.org) 
+ * Volker Springel (volker.springel@h-its.org)
+ *******************************************************************************/
+#include <mpi.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <math.h>
+#include <gsl/gsl_math.h>
+#include <gsl/gsl_rng.h>
+#include <gsl/gsl_randist.h>
+
+
+#include "allvars.h"
+#include "proto.h"
+
+
+/* returns the timestep for the particle with the giving velocity and acceleration
+ */
+double get_timestep(double *pos, double *vel, double *acc, int icell)
+{
+  // double r = sqrt(pos[0]*pos[0] + pos[1]*pos[1] + pos[2]*pos[2]);
+  double v = sqrt(vel[0] * vel[0] + vel[1] * vel[1] + vel[2] * vel[2]);
+  double aa = sqrt(acc[0] * acc[0] + acc[1] * acc[1] + acc[2] * acc[2]);
+
+  double torbit = All.V200 / aa;
+  double tcross = DG_CellSize[icell] / v;
+
+  return dmin(All.TimeStepFactorOrbit * torbit, All.TimeStepFactorCellCross * tcross);
+}
+
+
+/* calculate the density response of a single particle starting from pos[]/vel[],
+ * averaged over time 'timespan'. If timespan=0, the routine determines an
+ * appropriate time itself.
+ */
+double produce_orbit_response_field(double *pos, double *vel, int id, double *mfield, double mass,
+				    double timespan, int *orbitstaken)
+{
+  int i, norbit, icell, flag = 0, iR, iz;
+  double x[3], v[3], a[3], dt, tall, radsign_previous = 0, radsign, fR, fz;
+
+  for(i = 0; i < 3; i++)
+    {
+      x[i] = pos[i];
+      v[i] = vel[i];
[…]

### 2023-02-16 test_part/ics/GalIC/src/system.c

@@ -1,0 +1,348 @@
+/*******************************************************************************
+ * This file is part of the GALIC code developed by D. Yurin and V. Springel.
+ *
+ * Copyright (c) 2014
+ * Denis Yurin (denis.yurin@h-its.org) 
+ * Volker Springel (volker.springel@h-its.org)
+ *******************************************************************************/
+#include <mpi.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <strings.h>
+#include <math.h>
+#include <time.h>
+#include <sys/time.h>
+#include <sys/resource.h>
+#include <sys/types.h>
+#include <unistd.h>
+#include <signal.h>
+#include <gsl/gsl_rng.h>
+#include <fenv.h>
+
+#include "allvars.h"
+#include "proto.h"
+
+int get_thread_num(void)
+{
+#if (NUM_THREADS > 1) /* This enables OpenMP */
+  return omp_get_thread_num();
+#else
+  return 0;
+#endif
+}
+
+
+
+double dabs(double a)
+{
+  if(a < 0)
+    return -a;
+  else
+    return a;
+}
+
+double dmax(double a, double b)
+{
+  if(a > b)
+    return a;
+  else
+    return b;
+}
+
+double dmin(double a, double b)
+{
+  if(a < b)
+    return a;
+  else
+    return b;
+}
+
+int imax(int a, int b)
+{
+  if(a > b)
+    return a;
+  else
+    return b;
+}
+
+int imin(int a, int b)
+{
+  if(a < b)
+    return a;
+  else
+    return b;
+}
+
+
+#ifdef DEBUG_ENABLE_FPU_EXCEPTIONS
+#include <fenv.h>
+void enable_core_dumps_and_fpu_exceptions(void)
+{
+  /* enable floating point exceptions */
+
+  extern int feenableexcept(int __excepts);
+  feenableexcept(FE_DIVBYZERO | FE_INVALID);
+
+
+  /* set core-dump size to infinity */
+  struct rlimit rlim;
+  getrlimit(RLIMIT_CORE, &rlim);
+  rlim.rlim_cur = RLIM_INFINITY;
+  setrlimit(RLIMIT_CORE, &rlim);
+
+  /* MPICH catches the signales SIGSEGV, SIGBUS, and SIGFPE....                                                                   
[…]

### 2023-02-16 test_part/plots/plot_samGvar.py

@@ -1,0 +1,129 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import arepo
+import h5py as h5
+import matplotlib as mpl
+from matplotlib import rc
+import glob
+from tqdm import tqdm
+from scipy.signal import savgol_filter
+import re
+import astropy.units as u
+import matplotlib.pylab as pl
+
+mpl.use('Agg')
+
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
+rc('text', usetex=True)
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+
+columnwidth = 244.0 * 0.035145980349999517 # convert to cm
+textwidth = 508.0 * 0.035145980349999517 # convert to cm
+
+snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
+bprop_path = '/n/home01/abeane/starbar/analysis/bar_prop/data/'
+torque_path = '/n/home01/abeane/starbar/analysis/torques/data/'
+
+tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
+        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
+
+
+
+# names
+Nbody = 'Nbody'
+phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
+phS15R35 = 'phantom-vacuum-Sg15-Rc3.5'
+phS10R35 = 'phantom-vacuum-Sg10-Rc3.5'
+phS05R35 = 'phantom-vacuum-Sg05-Rc3.5'
+
+lvl = 'lvl3'
+lvlGFM = 'lvl3-GFM'
+
+def read_out(I):
+    fname = '../runs/output-Gvar/out_G'+str(I)+'.npy'
+    t = np.load(fname, allow_pickle=True).tolist()
+    return t
+
+def fix_bar_angle(bar_angle):
+    out = np.zeros(len(bar_angle))
+    out[0] = bar_angle[0]
+
+    for i in range(1, len(bar_angle)):
+        dphi = bar_angle[i] - bar_angle[i-1]
+        if dphi < -np.pi:
+            dphi += 2.*np.pi
+        
+        out[i] = out[i-1] + dphi
+    
+    return out
+
+def read_bar_prop(name, lvl):
[…]

### 2023-02-16 test_part/plots/plot_torque.py

@@ -1,0 +1,140 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import arepo
+import h5py as h5
+import matplotlib as mpl
+from matplotlib import rc
+import glob
+from tqdm import tqdm
+from scipy.signal import savgol_filter
+import re
+import astropy.units as u
+import matplotlib.pylab as pl
+
+mpl.use('Agg')
+
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern'], 'size': 8})
+rc('text', usetex=True)
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+
+columnwidth = 244.0 * 0.035145980349999517 # convert to cm
+textwidth = 508.0 * 0.035145980349999517 # convert to cm
+
+snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
+bprop_path = '/n/home01/abeane/starbar/analysis/bar_prop/data/'
+torque_path = '/n/home01/abeane/starbar/analysis/torques/data/'
+
+tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
+        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
+
+
+
+# names
+Nbody = 'Nbody'
+phS2R35 = 'phantom-vacuum-Sg20-Rc3.5'
+phS15R35 = 'phantom-vacuum-Sg15-Rc3.5'
+phS10R35 = 'phantom-vacuum-Sg10-Rc3.5'
+phS05R35 = 'phantom-vacuum-Sg05-Rc3.5'
+
+lvl = 'lvl3'
+lvlGFM = 'lvl3-GFM'
+
+def read_Gout(I):
+    fname = '../runs/output-Gvar/out_G'+str(I)+'.npy'
+    t = np.load(fname, allow_pickle=True).tolist()
+    return t
+
+def read_Iout(I):
+    fname = '../runs/output-Ivar/out_I'+str(I)+'.npy'
+    t = np.load(fname, allow_pickle=True).tolist()
+    return t
+
+def fix_bar_angle(bar_angle):
+    out = np.zeros(len(bar_angle))
+    out[0] = bar_angle[0]
+
+    for i in range(1, len(bar_angle)):
+        dphi = bar_angle[i] - bar_angle[i-1]
[…]

### 2023-02-16 test_part/runs/integrate.py

@@ -1,0 +1,124 @@
+import numpy as np
+from numba import njit, prange
+from numba.core import types
+from numba.experimental import jitclass
+from force import Hernquist_force, bar_force, MiyamotoNagaiDisk
+
+spec = [
+    ('time', types.float64[:]),
+    ('total_torque', types.float64[:,:]),
+    ('torque_gas', types.float64[:]),
+    ('ps', types.float64[:]),
+    ('M', types.float64),
+    ('a', types.float64),
+    ('A', types.float64),
+    ('b', types.float64),
+    ('vc', types.float64),
+    ('mres', types.float64),
+]
+
+
+
+@jitclass(spec)
+class output(object):
+    def __init__(self):
+        pass
+#         key_list = []
+#         for key, _ in spec:
+#             key_list.append(key)
+
+#             self.key_list = key_list
+
+@njit(parallel=True)
+def integrate_particles(pos, vel, mres, M, a, A, b, vc, ps, disk_M, disk_a, disk_b, dt=0.1, Tmax=1.0,
+                        torque_gas=20.0, I_at_rCR6=2.0, G=43007.1, verbose=False):
+    T = 0.0
+    
+    ang = 0.0
+    
+    acc_halo = Hernquist_force(pos, M, a, G)
+    acc_bar = bar_force(pos, A, b, vc, ps, ang)
+    acc_disk = MiyamotoNagaiDisk(pos, disk_M, disk_a, disk_b, G)
+    acc = acc_halo + acc_bar + acc_disk
+    
+    Nint = int(Tmax/dt)+1
+    print('Nint=', Nint)
+    init_ps = ps
+    
+    # Output arrays
+    total_torque = np.zeros((Nint, 3))
+    time_out = np.zeros(Nint)
+    ps_out = np.zeros(Nint)
+    torque_gas_list = np.zeros(Nint)
+    
+    torque_gas0 = torque_gas
+    ps0 = ps
+
+    N = pos.shape[0]
+    
+    for i in range(Nint):
+        # Analysis
+        torque = mres * np.cross(pos, acc)
+        total_torque[i] = np.sum(torque, axis=0)
+        ps_out[i] = ps
+        time_out[i] = T
+        torque_gas_list[i] = torque_gas
+
+        # Print
+        if verbose:
+            print('i=', i, ' T=', T, 'total_torque=', total_torque[i], 'ps=', ps)
+        
+        # First half-kick and drift
+        for j in prange(N):
[…]

### 2023-02-17 paper-ApJ/ms.tex
Add discussion of softening lengths

@@
 %% using aastex version 6.3
-\documentclass[twocolumn,linenumbers]{aastex631}
+\documentclass[twocolumn,linenumbers,trackchanges]{aastex631}
 
@@
 $25.3\times10^6$ in the dark matter halo. We used a softening length of
-$0.02\,\textrm{kpc}$ for all components. Snapshots were saved at equal intervals
-of $0.005$ in the time units of the simulation,
-$\textrm{kpc}/(\textrm{km}/\textrm{s})$.
+$20\,\textrm{pc}$ for all components. \edit1{This softening length is smaller
+than used in the original GALAKOS model ($28\,\textrm{pc}$ in their model, but
+$43.5\,\textrm{pc}$ when scaled to our mass resolution). Our smaller softening
+length is consistent with other resolved ISM models \citep{2018MNRAS.480..800H,
+2019MNRAS.489.4233M}.} Snapshots were saved at equal intervals of $0.005$ in the
+time units of the simulation, $\textrm{kpc}/(\textrm{km}/\textrm{s})$.
 

### 2023-03-16 paper/ms.tex
Add google drive link

@@
-% region preamble
-
-\documentclass[fleqn,usenatbib]{mnras}
-
-\usepackage[T1]{fontenc}
-
-% Allow "Thomas van Noord" and "Simon de Laguarde" and alike to be sorted by "N" and "L" etc. in the bibliography.
-% Write the name in the bibliography as "\VAN{Noord}{Van}{van} Noord, Thomas"
-\DeclareRobustCommand{\VAN}[3]{#2}
-\let\VANthebibliography\thebibliography
-\def\thebibliography{\DeclareRobustCommand{\VAN}[3]{##3}\VANthebibliography}
+%% Beginning of file 'sample631.tex'
+%%
+%% Modified 2021 March
+%%
+%% This is a sample manuscript marked up using the
+%% AASTeX v6.31 LaTeX 2e macros.
+%%
+%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
+%% (Copyright 2000-2015).  See the classfile for details.
+
+%% AASTeX requires revtex4-1.cls and other external packages such as
+%% latexsym, graphicx, amssymb, longtable, and epsf.  Note that as of 
+%% Oct 2020, APS now uses revtex4.2e for its journals but remember that 
+%% AASTeX v6+ still uses v4.1. All of these external packages should 
+%% already be present in the modern TeX distributions but not always.
+%% For example, revtex4.1 seems to be missing in the linux version of
+%% TexLive 2020. One should be able to get all packages from www.ctan.org.
+%% In particular, revtex v4.1 can be found at 
+%% https://www.ctan.org/pkg/revtex4-1.
+
+%% The first piece of markup in an AASTeX v6.x document is the \documentclass
+%% command. LaTeX will ignore any data that comes before this command. The 
+%% documentclass can take an optional argument to modify the output style.
+%% The command below calls the preprint style which will produce a tightly 
+%% typeset, one-column, single-spaced document.  It is the default and thus
+%% does not need to be explicitly stated.
+%%
+%% using aastex version 6.3
+\documentclass[twocolumn,linenumbers,trackchanges]{aastex631}
+
+%% The default is a single spaced, 10 point font, single spaced article.
+%% There are 5 other style options available via an optional argument. They
+%% can be invoked like this:
+%%
+%% \documentclass[arguments]{aastex631}
+%% 
+%% where the layout options are:
+%%
+%%  twocolumn   : two text columns, 10 point font, single spaced article.
+%%                This is the most compact and represent the final published
+%%                derived PDF copy of the accepted manuscript from the publisher
+%%  manuscript  : one text column, 12 point font, double spaced article.
+%%  preprint    : one text column, 12 point font, single spaced article.  
+%%  preprint2   : two text columns, 12 point font, single spaced article.
+%%  modern      : a stylish, single text column, 12 point font, article with
+%% 		  wider left and right margins. This uses the Daniel
+%% 		  Foreman-Mackey and David Hogg design.
+%%  RNAAS       : Supresses an abstract. Originally for RNAAS manuscripts 
+%%                but now that abstracts are required this is obsolete for
+%%                AAS Journals. Authors might need it for other reasons. DO NOT
+%%                use \begin{abstract} and \end{abstract} with this style.
+%%
+%% Note that you can submit to the AAS Journals in any of these 6 styles.
+%%
+%% There are other optional arguments one can invoke to allow other stylistic
+%% actions. The available options are:
+%%
+%%   astrosymb    : Loads Astrosymb font and define \astrocommands. 
+%%   tighten      : Makes baselineskip slightly smaller, only works with 
+%%                  the twocolumn substyle.
+%%   times        : uses times font instead of the default
+%%   linenumbers  : turn on lineno package.
+%%   trackchanges : required to see the revision mark up and print its output
+%%   longauthor   : Do not use the more compressed footnote style (default) for 
+%%                  the author/collaboration/affiliations. Instead print all
+%%                  affiliation information after each name. Creates a much 
+%%                  longer author list but may be desirable for short 
+%%                  author papers.
+%% twocolappendix : make 2 column appendix.
+%%   anonymous    : Do not show the authors, affiliations and acknowledgments 
+%%                  for dual anonymous review.
+%%
+%% these can be used in any combination, e.g.
+%%
+%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex631}
+%%
+%% AASTeX v6.* now includes \hyperref support. While we have built in specific
+%% defaults into the classfile you can manually override them with the
+%% \hypersetup command. For example,
+%%
+%% \hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
+%%
+%% will change the color of the internal links to red, the links to the
+%% bibliography to green, the file links to cyan, and the external links to
+%% magenta. Additional information on \hyperref options can be found here:
+%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
+%%
+%% Note that in v6.3 "bookmarks" has been changed to "true" in hyperref
+%% to improve the accessibility of the compiled pdf file.
+%%
+%% If you want to create your own macros, you can do so
+%% using \newcommand. Your macros should appear before
+%% the \begin{document} command.
+%%
 
@@
 
-% endregion
-
-% region title
-%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%
-
-% Title of the paper, and the short title which is used in the headers.
-% Keep the title short and informative.
-\title[Stellar Bars and Gas]{Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down}
-
-\author[A. Beane et al.]{Angus Beane,$^{1}$\thanks{E-mail: angus.beane@cfa.harvard.edu (AB)}
-Lars Hernquist,$^{1}$
-Elena~D'Onghia,$^{2,3}$
-Federico~Marinacci,$^{4}$
-Charlie Conroy,$^{1}$
-Jia~Qi,$^{5}$\newauthor
-Laura~V.~Sales,$^{6}$
-Paul~Torrey,$^{5}$
-Mark~Vogelsberger$^{7}$
-\\
-% List of institutions
-$^{1}$Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA\\
-$^{2}$Department of Physics, University of Wisconsin-Madison, Madison, WI, USA\\
-$^{3}$Department of Astronomy, University of Wisconsin-Madison, Madison, WI, USA\\
-$^{4}$Department of Physics \& Astronomy `Augusto Righi', University of Bologna, Bologna, Italy\\
-$^{5}$Department of Astronomy, University of Florida, Gainesville, FL, USA\\
-$^{6}$Department of Physics \& Astronomy, University of California, Riverside, CA, USA\\
-$^{7}$Department of Physics, Massachusetts Institute of Technology, Cambridge, MA, USA\\
-}
-
-% These dates will be filled out by the publisher
-\date{Accepted XXX. Received YYY; in original form ZZZ}
-
-% Enter the current year, for the copyright statements etc.
-\pubyear{2022}
-
-% endregion
-
-% Don't change these lines
+%% Reintroduced the \received and \accepted commands from AASTeX v5.2
+%\received{March 1, 2021}
+%\revised{April 1, 2021}
+%\accepted{\today}
+
+\shorttitle{Stellar Bars and Gas}
+\shortauthors{Beane et al.}
+
+\graphicspath{{./}{figures/}}
+
 \begin{document}
-\label{firstpage}
-\pagerange{\pageref{firstpage}--\pageref{lastpage}}
-\maketitle
-
-% Abstract of the paper
+\title{Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down}
+
+%% LaTeX will automatically break titles if they run longer than
+%% one line. However, you may use \\ to force a line break if
+%% you desire. In v6.31 you can include a footnote in the title.
+
+%% A significant change from earlier AASTEX versions is in the structure for 
+%% calling author and affiliations. The change was necessary to implement 
+%% auto-indexing of affiliations which prior was a manual process that could 
+%% easily be tedious in large author manuscripts.
+%%
+%% The \author command is the same as before except it now takes an optional
+%% argument which is the 16 digit ORCID. The syntax is:
+%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
+%%
+%% This will hyperlink the author name to the author's ORCID page. Note that
+%% during compilation, LaTeX will do some limited checking of the format of
+%% the ID to make sure it is valid. If the "orcid-ID.png" image file is 
+%% present or in the LaTeX pathway, the OrcID icon will appear next to
+%% the authors name.
+%%
+%% Use \affiliation for affiliation information. The old \affil is now aliased
+%% to \affiliation. AASTeX v6.31 will automatically index these in the header.
+%% When a duplicate is found its index will be the same as its previous entry.
+%%
+%% Note that \altaffilmark and \altaffiltext have been removed and thus 
+%% can not be used to document secondary affiliations. If they are used latex
+%% will issue a specific error message and quit. Please use multiple 
+%% \affiliation calls for to document more than one affiliation.
+%%
+%% The new \altaffiliation can be used to indicate some secondary information
+%% such as fellowships. This command produces a non-numeric footnote that is
+%% set away from the numeric \affiliation footnotes.  NOTE that if an
+%% \altaffiliation command is used it must come BEFORE the \affiliation call,
+%% right after the \author command, in order to place the footnotes in
+%% the proper location.
+%%
+%% Use \email to set provide email addresses. Each \email will appear on its
+%% own line so you can put multiple email address in one \email call. A new
+%% \correspondingauthor command is available in V6.31 to identify the
+%% corresponding author of the manuscript. It is the author's responsibility
+%% to make sure this name is also in the author list.
+%%
+%% While authors can be grouped inside the same \author and \affiliation
+%% commands it is better to have a single author for each. This allows for
+%% one to exploit all the new benefits and should make book-keeping easier.
+%%
+%% If done correctly the peer review system will be able to
+%% automatically put the author and affiliation information from the manuscript
+%% and save the corresponding author the trouble of entering it by hand.
+
+%\correspondingauthor{August Muench}
+%\email{greg.schwarz@aas.org, gus.muench@aas.org}
+
+\author{Angus Beane}
+\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA}
+
+\author{Lars Hernquist}
+\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA}
+
+\author{Elena D'Onghia}
+\affiliation{Department of Physics, University of Wisconsin-Madison, Madison, WI, USA}
+\affiliation{Department of Astronomy, University of Wisconsin-Madison, Madison, WI, USA}
+
+\author{Federico Marinacci}
+\affiliation{Department of Physics \& Astronomy `Augusto Righi', University of Bologna, Bologna, Italy}
+
+\author{Charlie Conroy}
+\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA}
+
+\author{Jia Qi}
+\affiliation{Department of Astronomy, University of Florida, Gainesville, FL, USA}
+
+\author{Laura V. Sales}
+\affiliation{Department of Physics \& Astronomy, University of California, Riverside, CA, USA}
+
+\author{Paul Torrey}
+\affiliation{Department of Astronomy, University of Florida, Gainesville, FL, USA}
+
+\author{Mark Vogelsberger}
+\affiliation{Department of Physics, Massachusetts Institute of Technology, Cambridge, MA, USA}
+
 \begin{abstract}
-Elongated bar-like features are ubiquitous in galaxies, occurring at the centers
-of approximately two-thirds of spiral disks.  Due to gravitational interactions
-between the bar and the other components of galaxies, it is expected that
-angular momentum and matter will redistribute over long (Gyr) timescales in
-barred galaxies. Previous work ignoring the gas phase of galaxies has
-conclusively demonstrated that bars should slow their rotation over time due to
-their interaction with dark matter halos. We have performed a simulation of a
-Milky Way-like galactic disk hosting a strong bar which includes a
-state-of-the-art model of the interstellar medium and a live dark matter halo.
-In this simulation the bar pattern does not slow down over time, and instead
-remains at a stable, constant rate of rotation. This behavior has been observed
-in previous simulations using more simplified models for the interstellar gas,
-but the apparent lack of secular evolution has remained unexplained. We propose
-that the gas phase of the disk and the dark matter halo act in concert to
-stabilize the bar pattern speed and prevent the bar from slowing down or
-speeding up. We find that in a Milky Way-like disk, a gas fraction of only about
-5\% is necessary for this mechanism to operate. Our result naturally explains
-why nearly all observed bars rotate rapidly and is especially relevant for our
-understanding of how the Milky Way arrived at its present state.
+
+  Elongated bar-like features are ubiquitous in galaxies, occurring at the
+  centers of approximately two-thirds of spiral disks in the nearby Universe.  
+  Due to gravitational
+  interactions between the bar and the other components of galaxies, it is
+  expected that angular momentum and matter will redistribute over long (Gyr)
+  timescales in barred galaxies. Previous work ignoring the gas phase of
+  galaxies has conclusively demonstrated that bars should slow their rotation
+  over time due to their interaction with dark matter halos. We have performed a
+  simulation of a Milky Way-like galactic disk hosting a strong bar which
+  includes a state-of-the-art model of the interstellar medium and a live dark
+  matter halo. In this simulation the bar pattern does not slow down over time,
+  and instead remains at a stable, constant rate of rotation. This behavior has
+  been observed in previous simulations using more simplified models for the
+  interstellar gas, but the apparent lack of secular evolution has remained
+  unexplained. We propose that the gas phase of the disk and the dark matter
+  halo act in concert to stabilize the bar pattern speed and prevent the bar
+  from slowing down or speeding up. We find that in a Milky Way-like disk, a gas
+  fraction of only about 5\% is necessary for this mechanism to operate. Our
+  result naturally explains why nearly all observed bars rotate rapidly and is
+  especially relevant for our understanding of how the Milky Way arrived at its
+  present state.
+
 \end{abstract}
 
-% Select between one and six entries from the list of approved keywords.
-% Don't make up new ones.
-\begin{keywords}
-galaxies: bar - galaxies: disc - galaxies: evolution - galaxies: kinematics and
-dynamics
-\end{keywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
+%% Keywords should appear after the \end{abstract} command. 
+%% The AAS Journals now uses Unified Astronomy Thesaurus concepts:
+%% https://astrothesaurus.org
+%% You will be asked to selected these concepts during the submission process
+%% but this old "keyword" functionality is maintained in case authors want
+%% to include these concepts in their preprints.
+\keywords{Classical Novae (251) --- Ultraviolet astronomy(1736) --- History of astronomy(1868) --- Interdisciplinary astronomy(804)}
+
+%% From the front matter, we move on to the body of the paper.
+%% Sections are demarcated by \section and \subsection, respectively.
+%% Observe the use of the LaTeX \label
+%% command after the \subsection to give a symbolic KEY to the
+%% subsection for cross-referencing in a \ref command.
+%% You can use LaTeX's \ref and \label commands to keep track of
+%% cross-references to sections, equations, tables, and figures.
+%% That way, if you change the order of any elements, LaTeX will
+%% automatically renumber them.
+%%
+%% We recommend that authors also use the natbib \citep
+%% and \citet commands to identify citations.  The citations are
+%% tied to the reference list via symbolic KEYs. The KEY corresponds
+%% to the KEY in the \bibitem in the reference list below.
 
@@
 
+In the last decade, stellar bars have been studied mainly in the
+context of the instability processes that lead to their formation and their
+ability to drive gas toward the galaxy center and contribute to the formation of
+supermassive black holes (SMBH). The primary interest of these studies mainly
+was in the loss of angular momentum of the gas and the associated  gaseous flow
+down to the inner disk, possibly forming a central mass concentration
+\citep{2010ApJ...719.1470V}, and fueling the central SMBH \citep[e.g.][]{1989Natur.338...45S, 1990Natur.345..679S}. A revisit of the more general problem of
+galactic bar properties and formation, including the case of disk galaxies with
+very large gas fraction is timely since it is clear now that galactic disks show
+massive bars already at redshift $z>2$ \citep{2022arXiv221008658G}. At
+that time the universe was 2.5 billion year old and galaxies might have as high
+gas fraction as 80\% \citep{2020ARAA..58..157T}. Furthermore, unlike
+nearby disk galaxies, the high-redshift disks also continuously accrete cold gas
+from the cosmic web, making the formation, stability and evolution of
+non-axisymmetric features a key question to address since they can play a
+fundamental role in the more general problem of disk galaxy evolution.
+
 We have performed a simulation of a disk galaxy using the finite-volume,
-gravito-hydrodynamics code \AREPO{} \citep{2010MNRAS.401..791S}. We use the galaxy
-formation model Stars and MUltiphase Gas in GaLaxiEs
-\citep[SMUGGLE;][]{2019MNRAS.489.4233M}. This disk galaxy exhibits almost no
+gravito-hydrodynamics code \AREPO{} \citep{2010MNRAS.401..791S}. We use the
+galaxy formation model Stars and MUltiphase Gas in GaLaxiEs
+\citep[\SMUGGLE{};][]{2019MNRAS.489.4233M}. This disk galaxy exhibits almost no
 evolution in the bar pattern speed over several Gyr when the gas phase is
 accounted for and properly modeled. This behavior has been observed in a few
-previous works \citep{1993AA...268...65F, 2010ApJ...719.1470V,
-2014MNRAS.438L..81A}. However, we are the first to demonstrate that the behavior
-ought to be present in a large fraction of galaxies, since it occurs in our
-simulation with gas fractions as low as about $5\%$. We are also the first to
-provide a physical explanation of the behavior with clear observational
-predictions.
+previous works \citep{1993AA...268...65F, 2007ApJ...666..189B,
+2009ApJ...707..218V, 2010ApJ...719.1470V, 2014MNRAS.438L..81A}.
+
+Here we show that this phenomenon occurs in galaxies with gas fraction
+as low as about $5\%$, making this feature a likely characteristic of  Milky
+Way-type galaxies. This behavior will be explored in more general disks with
+larger gas fractions than the Milky Way in a forthcoming paper.
+% We are also the first to provide a physical explanation of
+%the behavior with clear observational predictions.
+We also provide a physical explanation of the redistribution of the angular
+momentum among the dark halo wake, the stellar bar, and the disk gas component.
 
@@
 \Nbody{} simulations which showed a clear correlation between the bar pattern
-speed and the bar strength \citep[e.g.,][]{2003MNRAS.341.1179A}. In the SMUGGLE
+speed and the bar strength \citep[e.g.,][]{2003MNRAS.341.1179A}. In the \SMUGGLE{}
 case, we see that the bar strength has an initial drop but then remains at a
 roughly constant, but slightly decreasing, strength. This is consistent with the
-pattern speed in the SMUGGLE case being roughly constant or slightly increasing.
-
-% \subsection{Figures and tables}
-
-% Figures and tables should be placed at logical positions in the text. Don't
-% worry about the exact layout, which will be handled by the publishers.
-
-% Figures are referred to as e.g. Fig.~\ref{fig:example_figure}, and tables as
-% e.g. Table~\ref{tab:example_table}.
-
-% % Example figure
-% \begin{figure}
-% 	% To include a figure from a file named example.*
-% 	% Allowable file formats are eps or ps if compiling using latex
-% 	% or pdf, png, jpg if compiling using pdflatex
-% 	\includegraphics[width=\columnwidth]{example}
-%     \caption{This is an example figure. Captions appear below each figure.
-% 	Give enough detail for the reader to understand what they're looking at,
-% 	but leave detailed discussion to the main body of the text.}
-%     \label{fig:example_figure}
-% \end{figure}
-
-% % Example table
-% \begin{table}
-% 	\centering
-% 	\caption{This is an example table. Captions appear above each table.
-% 	Remember to define the quantities, symbols and units used.}
-% 	\label{tab:example_table}
-% 	\begin{tabular}{lccr} % four columns, alignment for each
-% 		\hline
-% 		A & B & C & D\\
-% 		\hline
-% 		1 & 2 & 3 & 4\\
-% 		2 & 4 & 6 & 8\\
-% 		3 & 5 & 7 & 9\\
-% 		\hline
-% 	\end{tabular}
-% \end{table}
+pattern speed in the \SMUGGLE{} case being roughly constant or slightly increasing.
 
@@
 \subsection{Pattern Speed Evolution}
-The lack of evolution in the pattern speed of the SMUGGLE case (seen in
+The lack of evolution in the pattern speed of the \SMUGGLE{} case (seen in
 Fig.~\ref{fig:prop}) is intimately tied to the sudden decrease in torque exerted
-on the bar by the dark matter halo. We argue that this can be understood in
-terms of the halo wake mechanism. In the \Nbody{} case, halo material that is
-resonant with the bar will form a wake which lags behind and exerts a negative
-torque on the bar, slowing it down \citep{1984MNRAS.209..729T,
-1985MNRAS.213..451W, 1992ApJ...400...80H}.\footnote{Since the bar is not a solid
+on the bar by the dark matter halo.
+
+%We argue that this can be understood in terms of the halo wake mechanism. 
+We interpret this behavior in terms of the halo wake mechanism. In the
+\Nbody{} case, a well known phenomenon is that the halo material resonant with
+the bar forms a wake, and this wake lags behind \citep{1984MNRAS.209..729T,
+1985MNRAS.213..451W, 1992ApJ...400...80H} and exerts a negative torque on the
+bar, slowing it down (see Fig.~\ref{fig:wake} below).\footnote{Since the bar is not a solid
 body, it is not guaranteed that a negative torque will slow it down - e.g. a
@@
 negative torque induced by a halo wake by slowing down.} As the bar slows down,
-the location of the resonances in phase space changes, allowing halo material
-newly resonant with the bar to participate in forming the wake. However, the gas
-is a reliable source of positive torque on the bar, speeding the bar up. This
-stops the resonance location from changing such that the halo cannot reinforce
-the wake, arresting the process by which the halo can slow the bar down. We term
-this process ``halo exhaustion.''
-
-We can test this interpretation by measuring the angle offset between the halo
+the location of the resonances in the phase space changes (see Fig.~12 and Table
+1 in \citet{2020ApJ...890..117D}) allowing halo material newly resonant with the
+bar to participate in the formation of the wake. However, the gas is also a
+reliable source of positive torque on the bar, speeding the bar up. In turn,
+this stops the location of the resonance from changing such that the halo cannot
+reinforce the wake, therefore arresting the process by which the halo can slow
+the bar down. We term this process ``halo exhaustion.'' Note that the resonances
+can be maintained even in the presence of the dissipative gaseous component if
+the gas is continuously accreted from the external environment, as is the case
+for massive disk galaxies at high redshift.
+
+%In the \Nbody{} case, halo material that is
+%resonant with the bar will form a wake which lags behind and exerts a negative 
+%torque on the bar, slowing it down \citep{1984MNRAS.209..729T,
+%1985MNRAS.213..451W, 1992ApJ...400...80H}.\footnote{Since the bar is not a solid
+%body, it is not guaranteed that a negative torque will slow it down - e.g. a
+%negative torque could shred the bar, reducing its moment of inertia without
+%changing its pattern speed. However, the bar seems to empirically respond to a
+%negative torque induced by a halo wake by slowing down.} As the bar slows down,
+%the location of the resonances in phase space changes, allowing halo material
+%newly resonant with the bar to participate in forming the wake. However, the gas
+%is a reliable source of positive torque on the bar, speeding the bar up. This
+%stops the resonance location from changing such that the halo cannot reinforce
+%the wake, arresting the process by which the halo can slow the bar down. We term
+%this process ``halo exhaustion.''
+
+This halo exhaustion process is similar to the ``metastability'' effect, which
+has been previously discussed in the literature \citep{2003MNRAS.345..406V,
+2006ApJ...639..868S}. Finally, we also note that these authors, in particular,
+observe that the effects of numerical resolution in the simulations adopted to
+explore these mechanisms have yet to be fully explored and could play a role in
+the observed phenomenology of the simulations. We plan to address these issues
+in future dedicated work.
+
+We test this interpretation by measuring the angle offset between the halo
 wake and the bar. If the wake and the bar are aligned (i.e., there is no angle
@@
 the dark matter halo can contribute to a wake. However, this does not explain
-why the pattern speed in the SMUGGLE case is nearly constant over several Gyr.
-Naively, it would be a coincidence that the bar pattern speed remains constant
-in the SMUGGLE case, resulting from a chance cancellation of the halo and gas
-torques. However, a constant pattern speed in the presence of gas has been
-observed in a few simulations of barred galaxies with
-gas \citep{1993AA...268...65F, 2010ApJ...719.1470V, 2014MNRAS.438L..81A}. Previous
-work has argued this is due to the bar torquing gas inwards, but no explanation
-has been given for why it might remain constant.
+why the pattern speed in the \SMUGGLE{} case is nearly constant over several
+Gyr. Naively, it would be a coincidence that the bar pattern speed remains
+constant in the \SMUGGLE{} case, resulting from a chance cancellation of the
+halo and gas torques. However, a constant pattern speed in the presence of gas
+has been observed in a few simulations of barred galaxies with gas
+\citep{1993AA...268...65F, 2007ApJ...666..189B, 2009ApJ...707..218V,
+2010ApJ...719.1470V, 2014MNRAS.438L..81A}.
+
+\citet{1993AA...268...65F} argue this behavior is due to the steepening of the
+circular velocity curve in the central region as the bar drives gas to the
+center. \citet{2009ApJ...707..218V} argue this behavior occurs when the
+corotation resonance is larger than the disk radius, but we observe the behavior
+when the corotation radius is well within the disk.
+
+% Previous work has argued this is due to the bar torquing
+% gas inwards, but no explanation has been given for why it might remain constant.
 
@@
 $I_6=8$ is a good approximation to our fiducial disk model. In code units, the
-moment of inertia of the SMUGGLE bar (i.e., the particles classified as being in
+moment of inertia of the \SMUGGLE{} bar (i.e., the particles classified as being in
 the bar) is about $2$. This is a factor of $4$ smaller than our fiducial value
-of $I_6=8$, but this is probably due to the fact that the bar does not really
-rotate as a solid body.
+of $I_6=8$, but this is probably due either to the fact that the bar does not
+really rotate as a solid body or that resonantly captured stars contribute to
+the real bar's effective moment of inertia \citep{1985MNRAS.213..451W}.
 
@@
 
+\citet{2020MNRAS.495.4158F} find that quenched galaxies tend to host longer bars
+than star-forming galaxies. This provides some support for our proposed
+mechanism since there is evidence quenching can occur through gas depletion
+\citep[e.g.][]{2021Natur.597..485W}. However, this correlation could be
+explained simply by the fact that longer bars ought to be more efficient at
+quenching their host galaxies \citep[e.g][]{2015AA...580A.116G}.
+
 \subsection{Cosmological Simulations}
@@
 spiral galaxies should rotate slowly.
-https://github.com/gusbeane/gasbar
-\section*{Acknowledgements}
-We would like to thank Greg~L. Bryan, Neal~J. Evans, Drummond~B. Fielding, Keith
-Hawkins, Jason~A.~S. Hunt, Sarah~M.~R. Jeffreson, Kathryn~V. Johnston,
-Peter~M.~W. Kalberla, Jürgen Kerp, Julio~F. Navarro, Joshua~S. Speagle,
-Martin~D. Weinberg, and Yanfei Zou for helpful discussions. A.B. would like to
-thank Todd Phillips for helpful discussions. Some computations in this paper
-were run on the FASRC Cannon cluster supported by the FAS Division of Science
-Research Computing Group at Harvard University. Resources supporting this work
-were also provided by the NASA High-End Computing (HEC) Program through the NASA
-Advanced Supercomputing (NAS) Division at Ames Research Center. This research
-has made use of the Spanish Virtual Observatory (https://svo.cab.inta-csic.es)
-project funded by MCIN/AEI/10.13039/501100011033/ through grant
-PID2020-112949GB-I00. A.B. was supported by the Future Investigators in NASA
-Earth and Space Science and Technology (FINESST) award number 80NSSC20K1536
-during the completion of this work. E.D. was partially supported by HST grants
-HST-AR-16363.001 and HST-AR-16602.006-A and by NASA Award NASA 80NSSC22K0761.
-J.Q. acknowledges support from NSF grant AST-2008490. L.V.S. is grateful for
-financial support from NASA ATP 80NSSC20K0566, NSF AST 1817233 and NSF CAREER
-1945310 grants. P.T. acknowledges support from NSF grant AST-1909933,
-AST-2008490, and NASA ATP Grant 80NSSC20K0502. M.V. acknowledges support through
-NASA ATP 19-ATP19-0019, 19-ATP19-0020, 19-ATP19-0167, and NSF grants
-AST-1814053, AST-1814259, AST-1909831, AST-2007355 and AST-2107724.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Data Availability}
-A git repository containing all analysis scripts and notebooks is available at
-\url{https://github.com/gusbeane/gasbar}. Snapshots at a $500\,\Myr$ cadence, as
-well as the initial conditions, are available at . Snapshots at a higher cadence
-are available upon request without expectation of coauthorship.
-
-We made use of the following software packages in {\sc Python} \citep{Python}:
+
+\begin{acknowledgments}
+  We would like to thank Greg~L. Bryan, Neal~J. Evans, Drummond~B. Fielding,
+  Keith Hawkins, Jason~A.~S. Hunt, Sarah~M.~R. Jeffreson, Kathryn~V. Johnston,
+  Peter~M.~W. Kalberla, Jürgen Kerp, Julio~F. Navarro, Dylan Nelson, Suchira
+  Sarkar, Joshua~S. Speagle, Martin~D. Weinberg, and Yanfei Zou for helpful
+  discussions. A.B. would like to thank Todd Phillips for helpful discussions.
+  Some computations in this paper were run on the FASRC Cannon cluster supported
+  by the FAS Division of Science Research Computing Group at Harvard University.
+  Resources supporting this work were also provided by the NASA High-End
+  Computing (HEC) Program through the NASA Advanced Supercomputing (NAS)
+  Division at Ames Research Center. This research has made use of the Spanish
+  Virtual Observatory (https://svo.cab.inta-csic.es) project funded by
+  MCIN/AEI/10.13039/501100011033/ through grant PID2020-112949GB-I00. A.B. was
+  supported by the Future Investigators in NASA Earth and Space Science and
+  Technology (FINESST) award number 80NSSC20K1536 during the completion of this
+  work. E.D. was partially supported by HST grants HST-AR-16363.001 and
+  HST-AR-16602.006-A and by NASA Award NASA 80NSSC22K0761. J.Q. acknowledges
+  support from NSF grant AST-2008490. L.V.S. is grateful for financial support
+  from NASA ATP 80NSSC20K0566, NSF AST 1817233 and NSF CAREER 1945310 grants.
+  P.T. acknowledges support from NSF grant AST-1909933, AST-2008490, and NASA
+  ATP Grant 80NSSC20K0502. M.V. acknowledges support through NASA ATP
+  19-ATP19-0019, 19-ATP19-0020, 19-ATP19-0167, and NSF grants AST-1814053,
+  AST-1814259, AST-1909831, AST-2007355 and AST-2107724.
+\end{acknowledgments}
+
+\software{
 {\sc agama} \url{https://github.com/GalacticDynamics-Oxford/Agama}, {\sc
@@
 \url{https://joblib.readthedocs.io/en/latest/}, {\sc matplotlib}
-\citep{Hunter:2007}, {\sc numba} \citep{Lam:2015bfh}, {\sc numpy}
+\citep{Hunter:2007}, {\sc numba} \citep{lam2015numba}, {\sc numpy}
 \citep{harris2020array}, {\sc scipy} \citep{2020SciPy-NMeth}, {\sc tqdm}
-\url{https://tqdm.github.io/}.
-
-%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%
-
-% The best way to enter references is to use BibTeX:
-
-\bibliographystyle{mnras}
-\bibliography{ref} % if your bibtex file is called example.bib
-
-% Alternatively you could enter them by hand, like this:
-% This method is tedious and prone to error if you have lots of references
-%\begin{thebibliography}{99}
-%\bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
-%Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
-%\bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
-%Others S., 2012, Journal of Interesting Stuff, 17, 198
-%\end{thebibliography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%
+\url{https://tqdm.github.io/}
+          }
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Don't change these lines
-\bsp	% typesetting comment
-\label{lastpage}
+%% For this sample we use BibTeX plus aasjournals.bst to generate the
+%% the bibliography. The sample631.bib file was populated from ADS. To
+%% get the citations to show in the compiled file do the following:
+%%
+%% pdflatex sample631.tex
+%% bibtext sample631
+%% pdflatex sample631.tex
+%% pdflatex sample631.tex
+
+\bibliography{ref}{}
+\bibliographystyle{aasjournal}
+
+%% This command is needed to show the entire author+affiliation list when
+%% the collaboration and author truncation commands are used.  It has to
+%% go at the end of the manuscript.
+%\allauthors
+
+%% Include this line if you are using the \added, \replaced, \deleted
+%% commands to see a summary list of all changes at the end of the article.
+%\listofchanges
 
 \end{document}
-
-% End of mnras_template.tex

### 2023-03-16 paper/ms.tex
Fix time of circular velocity curve

@@
 symmetric multipole fit with a maximum angular harmonic coefficient of $l=2$. We
-plot the circular velocity curve in Fig.~\ref{fig:vcirc} compared to
+plot the circular velocity curve {\bf at $t=1\,\textrm{Gyr}$} in Fig.~\ref{fig:vcirc} compared to
 observational estimates \citep{2019ApJ...871..120E}. The \SMUGGLE{} disk (which
@@
     \includegraphics[width=9cm]{fig/vcirc.pdf}
-    \caption{The circular velocity curve of our initial setups. This
-    curve is shown for the \Nbody{} run (blue) and the \SMUGGLE{} run (orange)
-    compared to observational estimates for the Milky
-    Way \citep{2019ApJ...871..120E}. We see that the circular velocity curve for
-    both runs is marginally larger than the Milky Way's, but still comparable. The
-    \SMUGGLE{} circular velocity curve is larger than the \Nbody{} curve due to the
-    additional mass in the gas phase.}
+    \caption{The circular velocity curve of our setups {\bf at $t=1\,\textrm{Gyr}$}.
+    This curve is shown for the \Nbody{} run (blue) and the \SMUGGLE{} run
+    (orange) compared to observational estimates for the Milky Way
+    \citep{2019ApJ...871..120E}. We see that the circular velocity curve for
+    both runs is marginally larger than the Milky Way's, but still comparable.
+    The \SMUGGLE{} circular velocity curve is larger than the \Nbody{} curve due
+    to the additional mass in the gas phase.}
     \label{fig:vcirc}

### 2023-03-23 paper/ms.tex
Add stars instead of gas test

@@
 
+{\bf \subsection{Stars Instead of Gas}
+In the SMUGGLE model considered in this work, we instantaneously added gas to
+the $N$-body system after $1.5\,\textrm{Gyr}$ of evolution. One might wonder if
+this sudden change to the potential is responsible for the stable pattern speed
+evolution. To test whether this is the case, we added mass to the system in the
+same way we did for the SMUGGLE model, but using collisionless particles instead
+of gas. The result of this experiment is shown in Fig.~\ref{fig:ps-star}. While
+there is an offset compared to the pure $N$-body case, we see that the pattern
+speed evolution is broadly consistent with a declining pattern speed. This
+indicates that the gas phase is responsible for the stable pattern speed.}
+
 \begin{figure}
@@
 \label{fig:GFM}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=9cm]{fig/ps_star.pdf}
+    \caption{{\bf Pattern speed evolution of a model in which we instaneously add stars instead of gas to the simulation, with the same density profile as the gas phase.
+The pattern speed evolution in this case is qualitatively similar to that of the $N$-body case, with a slight offset in the pattern speed.
+This test demonstrates that the stable pattern speed evolution in the SMUGGLE case is not simply a consequence of the change in potential imposed in our initial conditions.}}
+\label{fig:ps-star}
 \end{figure}

### 2023-03-23 paper/ms.tex
Add initial gas disk scale height

@@
 
+{\bf We measured the initial scale height of the gas disk in a bin extending from
+$R=8\,\textrm{kpc}$ to $R=8.5\,\textrm{kpc}$. The vertical profile is well-fit
+by a Gaussian with a scale height of $97\,\textrm{pc}$. This is somewhat lower
+than the observed value in the HI disk of $\sim200\,\textrm{kpc}$
+\citep{1995ApJ...448..138M}.}
+
 \begin{figure}

### 2023-03-23 paper/ms.tex
Add discussion of gas corotation

@@
 
+{\bf We also note that since \citet{2011MNRAS.415.1027H} predicts gas outside of
+corotation will flow outward, the bar should exert a positive torque on that
+gas. Indeed, we measured the average torque on gas outside corotation from
+$t=3\,\textrm{Gyr}$ to $5\,\textrm{Gyr}$ to be $0.87$ in code units
+($10^{10}\,M_{\odot}\,(\text{km}/\text{s})^2$). For reference, the average
+torque inside corotation is $-10.8$ over the same time period and in the same
+units. So, while gas outside corotation does experience a positive torque, the
+total torque on the gas phase is still negative.}
+
 \begin{figure}

### 2023-04-04 paper/ms.tex
Small changes from ref report

@@
     simulation while the lower panels show a simulation which includes the
-    \SMUGGLE{} model for the interstellar medium. Each panel is $20\,\textrm{kpc}$
-    to a side. Columns show different points in time, separated by
-    $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
-    length and strength. In the \SMUGGLE{} run, the bar remains at approximately
-    the same length and strength over the course of the simulation. The \Nbody{}
-    model is identical to the GALAKOS model, discussed in the text. Details on
-    the production of these synthetic images is given in Appendix~\ref{app:hst}.}\label{fig:overview}
+    \SMUGGLE{} model for the interstellar medium. Each panel is
+    $20\,\textrm{kpc}$ to a side. Columns show different points in time,
+    separated by $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar
+    grows in length and strength. In the \SMUGGLE{} run, the bar remains at
+    approximately the same length and strength over the course of the
+    simulation. The \Nbody{} model is identical to the GALAKOS model {\bf (with
+    different particle numbers and softening lengths)}, discussed in the text.
+    Details on the production of these synthetic images is given in
+    Appendix~\ref{app:hst}.}\label{fig:overview}
 \end{figure*}
@@
 almost always form bar-like structures \citep{1971ApJ...168..343H}. Several
-studies have shown that a spherical potential, such as a stellar bulge or dark
-matter halo, acts to stabilize stellar disks against bar formation
-\citep[e.g.,][]{1973ApJ...186..467O, 1976AJ.....81...30H}.
+studies have shown that a {\bf hot, centrally concentrated mass distribution},
+such as a stellar bulge or dark matter halo, acts to stabilize stellar disks
+against bar formation \citep[e.g.,][]{1973ApJ...186..467O, 1976AJ.....81...30H}.
 
@@
 
-While the interaction between the bar and a dark matter halo is well-understood
-theoretically, this is not the case for the interaction between a bar and a
-gaseous disk. Some argue that the gas disk should slow down the bar more
-\citep{2003MNRAS.341.1179A}, while others argue that the tendency of the bar to
-drive gas inwards means the bar should speed up due to the effect of the gas
-disk \citep{2013MNRAS.429.1949A, 2014MNRAS.438L..81A}. Since the gas phase
+{\bf While the fact that a bar is slowed down by a dark matter halo is
+well-understood theoretically,} this is not the case for the interaction between
+a bar and a gaseous disk. Some argue that the gas disk should slow down the bar
+more \citep{2003MNRAS.341.1179A}, while others argue that the tendency of the
+bar to drive gas inwards means the bar should speed up due to the effect of the
+gas disk \citep{2013MNRAS.429.1949A, 2014MNRAS.438L..81A}. Since the gas phase
 typically contributes only about $10-20\%$ of the mass of a galaxy at the
@@
 evolution in the bar pattern speed over several Gyr when the gas phase is
-accounted for and properly modeled. This behavior has been observed in a few
-previous works \citep{1993AA...268...65F, 2007ApJ...666..189B,
+accounted for and {\bf robustly} modeled. This behavior has been observed in a
+few previous works \citep{1993AA...268...65F, 2007ApJ...666..189B,
 2009ApJ...707..218V, 2010ApJ...719.1470V, 2014MNRAS.438L..81A}.
@@
 \citep{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies upon
-the pattern speed of the bar to change with time Our work does not alter
+the pattern speed of the bar to change with time. Our work does not alter
 expectations for radial mixing induced by spiral arms
@@
 
-We performed an experiment to test this hypothesis by forcing the stellar disk
-in the \SMUGGLE{} run to rotate at a constant angular rate and measuring the
-torque on the bar by the gas phase at different rotation rates. The result of
-this experiment is illustrated in Fig.~\ref{fig:equil}, which shows that a more
-slowly rotating bar experiences a larger positive torque from the gas.
+{\bf We performed an experiment to test this hypothesis by freezing the stellar
+disk in the \SMUGGLE{} run and forcing it to rotate at a constant angular rate.
+This has the effect of forcing the bar to rotate as a solid body at a constant
+angular rate which we control. The gas is evolved self-consistently with this
+rotating disk.} We measured the torque on the bar by the gas phase at different
+rotation rates. The result of this experiment is illustrated in
+Fig.~\ref{fig:equil}, which shows that a more slowly rotating bar experiences a
+larger positive torque from the gas.
 

### 2023-04-17 plots/fig1/plot_fig1.py
Add new fig1

@@ -7,6 +7,12 @@ from astropy.io import fits
 from scipy.interpolate import interp1d
 import re
 import pickle
+import agama
+
+import astropy.units as u
+from scipy.optimize import minimize
+
+agama.setUnits(mass=1E10, length=1, velocity=1)
 
 from photutils.isophote import EllipseGeometry, Ellipse
 
@@ -49,6 +55,26 @@ def read_fourier(name, lvl='lvl3',
                  basepath='/n/holylfs05/LABS/hernquist_lab/Users/abeane/gasbar/analysis/'):
     f = h5.File(basepath+'/fourier_component/data/fourier_'+name+'-'+lvl+'.hdf5', mode='r')
     return f
+
+def read_agama_pot(idx, name, lvl):
+    base = '/n/holylfs05/LABS/hernquist_lab/Users/abeane/gasbar/analysis/agama_pot/data/'
+    fname = base + 'pot_' + name + '-' + lvl + '/pot_' + name + '-' + lvl + '.' + str(idx) + '.txt'
+    return agama.Potential(fname)
+
+def compute_vc(R, pot):
+    acc = pot.force(R, 0, 0)
+    vcsq = - R * acc[0]
+    return np.sqrt(vcsq)
+
+def _to_minimize(R, pot, omega_p):
+    vc = compute_vc(R, pot)
+    omega = vc / R
+    return np.abs(omega - omega_p)
+
+def compute_RCR(pot, omega_p, Rguess=6):
+    ans = minimize(_to_minimize, Rguess, args=(pot, omega_p))
+
+    return float(ans.x)
 
 def fix_bar_angle(bar_angle):
     out = np.zeros(len(bar_angle))
@@ -134,6 +160,8 @@ def get_A2_angle(dat, keys, Rbin, cum=False):
     R_at_Rbin = Rlist[:,Rbin]
     
     time = np.array(dat['time'])
+    time = time * u.Myr
+    time = time.to_value(u.kpc/(u.km/u.s))
 
     return time, R_at_Rbin, phi
 
@@ -249,6 +277,18 @@ def run():
         snN = read_snap(Nidx, Nbody, 'lvl3')
         snS = read_snap(Sidx, phS2R35, 'lvl3')
         
+        agama_potN = read_agama_pot(Nidx, Nbody, 'lvl3')
+        agama_potS = read_agama_pot(Sidx, phS2R35, 'lvl3')
+        
+        print('N ps=', bangleN['pattern_speed'][Nidx])
[…]

### 2023-04-17 paper/ms.tex
Add radial mixing clarification

@@
 observational evidence \citep{2019MNRAS.490.4740B}. And finally, we complicate
-the picture of radial mixing expected to sculpt the Milky Way's disk
-\citep{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies upon
-the pattern speed of the bar to change with time. Our work does not alter
-expectations for radial mixing induced by spiral arms
+the picture of {\bf stellar} radial mixing expected to sculpt the Milky Way's
+disk \citep{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies
+upon the pattern speed of the bar to change with time. {\bf The radial mixing of
+the gas phase induced by the bar, as predicted in \citet{2011MNRAS.415.1027H},
+might have implications for the radial metallicity gradients of galaxies.} Our
+work does not alter expectations for radial mixing induced by spiral arms
 \citep{2002MNRAS.336..785S}.

### 2023-04-17 paper/ms.tex
Clarify AGAMA fit

@@
 We computed the circular velocity curve of our model using the \texttt{AGAMA}
-package \citep{2019MNRAS.482.1525V}. We fit the baryonic component (stellar disk,
-bulge, gas, and newly formed stars) with an axisymmetric cylindrical spline with
-$20$ grid points in both the radial and vertical direction spanning $0.2$ to
-$50\,\textrm{kpc}$ in the radial direction and from $0.02$ to $10\,\textrm{kpc}$
-in the vertical direction. We fit the dark matter halo using a spherically
-symmetric multipole fit with a maximum angular harmonic coefficient of $l=2$. We
-plot the circular velocity curve {\bf at $t=1\,\textrm{Gyr}$} in Fig.~\ref{fig:vcirc} compared to
-observational estimates \citep{2019ApJ...871..120E}. The \SMUGGLE{} disk (which
-includes additional mass in the form of gas) has a slightly higher circular
-velocity than the \Nbody{} disk which, itself, is slightly higher than the
-observational estimates. Overall, though, the circular velocity curves between
-our model and that observed in the Milky Way are broadly consistent.
+package \citep{2019MNRAS.482.1525V}. We fit the baryonic component (stellar
+disk, bulge, gas, and newly formed stars) with an axisymmetric cylindrical
+spline with $20$ grid points in both the radial and vertical direction spanning
+$0.2$ to $50\,\textrm{kpc}$ in the radial direction and from $0.02$ to
+$10\,\textrm{kpc}$ in the vertical direction. We fit the dark matter halo using
+{\bf an axisymmetric} multipole fit with a maximum angular harmonic coefficient
+of $l=2$, {\bf to account for the compression of the halo by the disk}. We plot
+the circular velocity curve {\bf at $t=1\,\textrm{Gyr}$} in Fig.~\ref{fig:vcirc}
+compared to observational estimates \citep{2019ApJ...871..120E}. The \SMUGGLE{}
+disk (which includes additional mass in the form of gas) has a slightly higher
+circular velocity than the \Nbody{} disk which, itself, is slightly higher than
+the observational estimates. Overall, though, the circular velocity curves
+between our model and that observed in the Milky Way are broadly consistent.
 

### 2023-04-17 paper/ms.tex
Add gas scale height discussion

@@
 
-{\bf We measured the initial scale height of the gas disk in a bin extending from
-$R=8\,\textrm{kpc}$ to $R=8.5\,\textrm{kpc}$. The vertical profile is well-fit
-by a Gaussian with a scale height of $97\,\textrm{pc}$. This is somewhat lower
-than the observed value in the HI disk of $\sim200\,\textrm{kpc}$
-\citep{1995ApJ...448..138M}.}
+{\bf We measured the initial scale height of the atomic gas disk in a bin
+extending from $R=7.5\,\textrm{kpc}$ to $R=8.5\,\textrm{kpc}$. The initial
+vertical profile is well-fit by a Gaussian with a scale height of
+$110\,\textrm{pc}$. At $t=1\,\textrm{Gyr}$, the vertical profile in the same
+radial bin is better fit by an exponential profile with scale height of
+$74\,\textrm{pc}$. These are somewhat lower than the observed value in the HI
+disk of $\sim200\,\textrm{kpc}$ \citep{1995ApJ...448..138M}. This may be caused
+by the model in our simulations not driving enough turbulent pressure, and is an
+interesting avenue of further investigation.}
 

### 2023-04-17 paper/ms.tex
Probably

@@
 Milky Way \citep{2008AA...487..951K,2022ApJ...929L..18E}. The discrepancy between
-$1$ and $4\,\textrm{kpc}$ in the molecular and SFR surface density is likely due
+$1$ and $4\,\textrm{kpc}$ in the molecular and SFR surface density is {\bf probably} due
 to the fact that the distances to molecular clouds which underlines this work

### 2023-04-17 paper/ms.tex
Add discussion of rot rate

@@
 quenching their host galaxies \citep[e.g][]{2015AA...580A.116G}.
+
+{\bf Finally, we mention the evolution of \Rot{} in our simulation. In the
+$N$-body simulation, the bar forms with $\Rot\sim1.6$ at $t=0\,\textrm{Gyr}$,
+which is already well within the slow bar regime. After $5\,\textrm{Gyr}$ of
+evolution, \Rot{} has risen to $\sim1.9$. In the SMUGGLE simulation, the gas is
+added to the $t=0\,\textrm{Gyr}$ snapshot, so it begins with $\Rot{}\sim1.6$. As
+expected, after $5\,\textrm{Gyr}$ of evolution \Rot{} is still $\sim1.6$.
+However, this relies on our measure of the bar length as being the maximum
+radius of all orbits trapped in the bar, which is not an observationally
+accessible measure of bar length. One would need to test different
+observationally possible bar length estimators, such as ellipse fitting
+\citep{1990MNRAS.245..130A, 1999AAS..140....1M, 2002MNRAS.330...35A,
+2006AA...452...97M, 2009AA...495..491A, 2015AA...576A.102A}. This is beyond the
+scope of our current work. Nonetheless, our prediction that \Rot{} is stable in
+the presence of gas is robust. Why the $N$-body simulation forms a bar with
+$\Rot\sim1.6$ is a separate question deserving further attention.}
 

### 2023-04-17 paper/ms.tex
Remove HST and add new fig1

@@
     \centering
-    \includegraphics[width=\textwidth]{fig/mockHST.pdf}
-    \caption{Synthetic Hubble Space Telescope images of our simulations with and
-    without the interstellar medium. The upper panels show an \Nbody{} only
-    simulation while the lower panels show a simulation which includes the
-    \SMUGGLE{} model for the interstellar medium. Each panel is
-    $20\,\textrm{kpc}$ to a side. Columns show different points in time,
-    separated by $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar
-    grows in length and strength. In the \SMUGGLE{} run, the bar remains at
-    approximately the same length and strength over the course of the
-    simulation. The \Nbody{} model is identical to the GALAKOS model {\bf (with
-    different particle numbers and softening lengths)}, discussed in the text.
-    Details on the production of these synthetic images is given in
-    Appendix~\ref{app:hst}.}\label{fig:overview}
+    \includegraphics[width=\textwidth]{fig/fig1.pdf}
+    \caption{{\bf Stellar mass distribution of our simulations with and without
+    the interstellar medium.} The upper panels show an \Nbody{} only simulation
+    while the lower panels show a simulation which includes the \SMUGGLE{} model
+    for the interstellar medium. Each panel is $20\,\textrm{kpc}$ to a side.
+    {\bf The bar length is shown as a solid white bar (details on its
+    computation are given in the text). The corotation radius is shown as a
+    dashed white circle.} Columns show different points in time, separated by
+    $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
+    length and strength. In the \SMUGGLE{} run, the bar remains at approximately
+    the same length and strength over the course of the simulation. The \Nbody{}
+    model is identical to the GALAKOS model {\bf (with different particle
+    numbers and softening lengths)}, discussed in the text.}\label{fig:overview}
 \end{figure*}
@@
 
-We show synthetic \textit{Hubble Space Telescope} images of our barred galaxy in
-a case with and without gas in Fig.~\ref{fig:overview}. We see that the bar
-grows longer and stronger without gas, while it remains at approximately the
-same length and strength when gas is included. Details on the production of
-these synthetic images are given in Appendix~\ref{app:hst}.
+We show {\bf stellar mass distributions} of our barred galaxy in a case with and
+without gas in Fig.~\ref{fig:overview}. We see that the bar grows longer and
+stronger without gas ({\bf bar length shown as a white bar}), while it remains
+at approximately the same length and strength when gas is included. {\bf As the
+bar without gas slows down, the corotation radius (white dashed circle) grows
+larger with time.}
 
@@
 
-\section{Idealized Mock Light Images}
-\label{app:hst}
-We created synthetic mock light images at $1024^2$ resolution in
-Fig.~\ref{fig:overview} for the \textit{Hubble Space Telescope} using a
-post-processed Monte Carlo radiation transfer simulation with the SKIRT9 code
-including secondary emission from dust \citep{2020AC....3100381C}. Emission from
-star particles was calculated assuming a Bruzual-Charlot spectral energy
-distribution \citep{2003MNRAS.344.1000B} and a Chabrier initial stellar mass
-function \citep{2003PASP..115..763C}. The luminosity from each star particle was
-smoothed using a cubic spline kernel with each star particle being assigned a
-smoothing length equal to its distance to its $32$nd closest neighbor. We
-launched $10^9$ photon packets per simulation segment. For star particles
-which were present in the initial simulation (i.e., initial disk and bulge
-particles), we assumed that they have an age of $5\,\textrm{Gyr}$ and solar
-metallicity. For star particles in the \SMUGGLE{} simulation which are newly
-formed, we used their recorded ages and metallicities.
-
-For the \SMUGGLE{} simulations, we included absorption and emission from dust. We
-assumed the THEMIS (The Heterogeneous dust Evolution Model for Interstellar
-Solids) dust model \citep{2017AA...602A..46J} to convert gas metallicity and
-mass density to dust mass density. For the \Nbody{} simulations we ran the
-SKIRT9 code in a mode without an obscuring medium, but otherwise applied the
-same procedure.
-
-We used the profiles for filters F814W, F606W, and F475W on the Advanced
-Camera for Surveys (ACS) instrument for red, green, and blue pixel values,
-respectively. Filtered fluxes were converted to RGB values using an $\arcsinh$
-scaling following \citet{2004PASP..116..133L} Scalings and cutoff values
-were adjusted manually to create visually appealing images. Filter profiles
-were provided by the Spanish Virtual Observatory Filter
-Service \citep{2012ivoa.rept.1015R, 2020sea..confE.182R}.
-
 \section{Comparison to the Milky Way}

### 2023-04-18 plots/ps_torque/plot_ps_torque.py
Make ps unsmoothed default, less torque smoothing

@@ -66,7 +66,7 @@ def read_bar_prop(name, lvl):
     return out
 
 def read_torque(name, lvl):
-    if name == Nbody:
+    if name == phS2R35:
         base = torque_path + 'torques_' + name + '-' + lvl + '-bak' + '/torques_' + name + '-' + lvl + '.'
     else:
         base = torque_path + 'torques_' + name + '-' + lvl + '/torques_' + name + '-' + lvl + '.'
@@ -119,30 +119,17 @@ def run():
 
     cm = 1/2.54
     fig, ax = plt.subplots(2, 1, sharex=True, figsize=(columnwidth*cm, columnwidth*cm))
-    
-    fig_us, ax_us = plt.subplots(2, 1, sharex=True, figsize=(columnwidth*cm, columnwidth*cm))
 
     # First panel, pattern speed.
     t300 = bar_prop_Nbody['tlist'][300]
-    ps_N = savgol_filter(bar_prop_Nbody['pattern_speed'], 81, 3)
-    ps_S = savgol_filter(bar_prop_SMUGGLE['pattern_speed'], 81, 3)
 
-    ax[0].plot(bar_prop_Nbody['tlist'] - t300, ps_N, c=tb_c[0], label=r'$N$-body')
-    ax[0].plot(bar_prop_SMUGGLE['tlist'], ps_S, c=tb_c[1], label='SMUGGLE')
+    ax[0].plot(bar_prop_Nbody['tlist'] - t300, bar_prop_Nbody['pattern_speed'], c=tb_c[0], label=r'$N$-body')
+    ax[0].plot(bar_prop_SMUGGLE['tlist'], bar_prop_SMUGGLE['pattern_speed'], c=tb_c[1], label='SMUGGLE')
     
-
     ax[0].set(ylim=(0, 60), ylabel=r'$\Omega_p\,[\,\text{km}/\text{s}/\text{kpc}\,]$')
     ax[0].set(xlim=(0, 5))
 
     ax[0].legend(frameon=False)
-    
-    ax_us[0].plot(bar_prop_Nbody['tlist'] - t300, bar_prop_Nbody['pattern_speed'], c=tb_c[0], label=r'$N$-body')
-    ax_us[0].plot(bar_prop_SMUGGLE['tlist'], bar_prop_SMUGGLE['pattern_speed'], c=tb_c[1], label='SMUGGLE')
-    
[…]

### 2023-04-18 paper/ms.tex
Add 40pc softening sentence

@@
 length is consistent with other resolved ISM models \citep{2018MNRAS.480..800H,
-2019MNRAS.489.4233M}.} Snapshots were saved at equal intervals of $0.005$ in the
-time units of the simulation, $\textrm{kpc}/(\textrm{km}/\textrm{s})$.
+2019MNRAS.489.4233M}. We reran our fiducial SMUGGLE run with $40\,\textrm{pc}$
+softening and found no difference in the pattern speed evolution of the bar.}
+Snapshots were saved at equal intervals of $0.005$ in the time units of the
+simulation, $\textrm{kpc}/(\textrm{km}/\textrm{s})$.
 

### 2023-04-18 paper/ms.tex
Update SG filter sentence

@@
 
-In order to remove numerical noise in several quantities we computed, we
-applied a Savitzky-Golay filter \citep{1964AnaCh..36.1627S} as implemented in
-\texttt{scipy} using a window length of $81$ and polynomial order of $3$. This
-filter was applied to plots of pattern speeds, bar lengths and strengths,
-torques, and angle differences.
+We applied a Savitzky-Golay filter \citep{1964AnaCh..36.1627S} as implemented in
+\texttt{scipy} using a window length of $21$ and polynomial order of $3$ to the
+plot of torques in order to remove some numerical noise.
 

### 2023-04-18 test_part/runs/integrate.py
Remove Hopkins ps torque dependence

@@ -81,10 +81,10 @@ def integrate_particles(pos, vel, mres, M, a, A, b, vc, ps, disk_M, disk_a, disk
         ps += - total_torque[i][2]/I * dt / 2.0 # constant ps
         ps += torque_gas/I * dt / 2.0
         
-        if torque_gas0 > 0:
-            torque_gas = torque_gas0 + (40-ps)
-        else:
-            torque_gas = torque_gas0
+        #if torque_gas0 > 0:
+        #    torque_gas = torque_gas0 + (40-ps)
+        #else:
+        #    torque_gas = torque_gas0
 
         # Bar drift
         ang += ps * dt

### 2023-04-18 test_part/runs/job_varyG.sh

@@ -1,5 +1,5 @@
 #!/bin/sh
-#SBATCH -p shared,itc_cluster,conroy 
+#SBATCH -p hernquist,shared,itc_cluster,conroy 
 #SBATCH -J sam_vG
 #SBATCH -n 1
 #SBATCH -N 1 

### 2023-04-18 paper/ms.tex
Remove torque varying with ps

@@
 bar in the same way as the torque from the halo is applied. The torque is given
-in code units ($10^{10}\Msun \left(\kms\right)^2$). We allow this torque to
-mildly vary with pattern speed according to the formula,
-\begin{equation}
-\tau = \tau_{\textrm{gas}} + (40 - \Omega_{\textrm{p}})\textrm{,}
-\end{equation}
-where $\Omega_{\textrm{p}}$ is given in $\kms/\textrm{kpc}$. We are motivated to
-do this because of our result in Appendix~\ref{app:varyps} which shows that more
-slowly rotating bars have higher gas infall rates and thus larger positive
-torques. In the $\tau_{\textrm{gas}}=0$ case, we do not allow the gas torque to
-vary with pattern speed.
+in code units ($10^{10}\Msun \left(\kms\right)^2$).
 

### 2023-04-19 plots/halo_wake/plot_halo_wake.py
Update halo wake plot

@@ -22,8 +22,9 @@ rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 
 snap_path = '/n/holystore01/LABS/hernquist_lab/Users/abeane/starbar_runs/runs/'
-fourier_path = '/n/home01/abeane/starbar/analysis/fourier_sphere/data/'
-bprop_path = '/n/home01/abeane/starbar/analysis/bar_prop/data/'
+fourier_path = '/n/holylfs05/LABS/hernquist_lab/Users/abeane/gasbar/analysis/fourier_sphere/data/'
+bprop_path = '/n/holylfs05/LABS/hernquist_lab/Users/abeane/gasbar/analysis/bar_prop/data/'
+in_bar_path = '/n/holylfs05/LABS/hernquist_lab/Users/abeane/gasbar/analysis/in_bar/data/'
 
 tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
         '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
@@ -38,6 +39,22 @@ def read_snap(idx, name, lvl, parttype=[2, 3, 4], fields=['Coordinates', 'Masses
     sn = arepo.Snapshot(snap_path+name+'/'+lvl+'/output', idx, combineFiles=True, parttype=parttype,
                         fields=fields)
     return sn
+
+def read_in_bar(name, lvl, idx, nchunk=256):
+    fname_base = in_bar_path+'in_bar_'+name+'-'+lvl+'/in_bar_'+name+'-'+lvl+'.'
+    
+    in_bar = []
+    
+    for i in tqdm(range(nchunk)):
+        fname = fname_base + str(i) + '.hdf5'
+        
+        t = h5.File(fname, mode='r')
+        in_bar.append(t['PartType1']['in_bar'][idx])
+        t.close()
+    
+    in_bar = np.concatenate(in_bar)
+    
+    return in_bar
 
 @njit
 def compute_surface_density(R, mass, Rbins):
@@ -77,13 +94,17 @@ def rotate_pos(pos, ang):
     
     return pos
 
-def compute_heatmap(sn, center, nres, rng, angle=0.0):
[…]

### 2023-04-19 paper/ms.tex
Add spiral arm argument to barred/unbarred

@@
 is in the particles classified as barred (i.e.,
-$A_{2,\textrm{bar}}/A_{2,\textrm{tot}}\sim0.76$). This indicates good but not
-perfect classification. Some of this is probably coming from the $m=4$ component
-being classified as unbarred, given that the unbarred component has a somewhat
-boxy shape.
+$A_{2,\textrm{bar}}/A_{2,\textrm{tot}}\sim0.76$). Some of this is probably
+coming from the $m=4$ component {\bf (e.g., boxy orbits)} being classified as
+unbarred, {\bf or the presence of weak spiral arms. See also
+\citet{2021MNRAS.500..838P} for more details on the orbit family breakdown.}
 

### 2023-04-19 paper/ms.tex
Add bar mass

@@
     \includegraphics[width=9cm]{fig/Rb_A2.pdf}
-    \caption{The evolution in bar length and strength. The \textit{upper panel}
+    \caption{The evolution in bar length, mass and strength. The \textit{upper panel}
     shows the evolution of the bar length. In the \Nbody{} case, the bar
@@
     at larger radii are possible. Stars are captured on these orbits,
-    lengthening the bar. This process does not occur in the \SMUGGLE{} cases since
-    the bar pattern speed is not decreasing, and therefore the bar length
-    remains constant. The bar strength, shown in the \textit{lower panel} is
-    measured as the maximum of the second Fourier component divided by the
-    zeroth Fourier component. We see that in the \Nbody{} case (blue) the bar
-    strength increases with time, consistent with previous results showing that
-    the bar strength increases as bars slow down. In the \SMUGGLE{} case (orange),
-    we see that the bar strength remains roughly constant, possibly slightly
+    lengthening the bar. This process does not occur in the \SMUGGLE{} cases
+    since the bar pattern speed is not decreasing, and therefore the bar length
+    remains constant. {\bf The bar mass, shown in the \textit{middle panel}, is
+    increasing in the \Nbody{} case as the bar grows. It is decreasing in the
+    \SMUGGLE{} case, indicating a shredding process.} The bar strength, shown in
+    the \textit{lower panel} is measured as the maximum of the second Fourier
+    component divided by the zeroth Fourier component. We see that in the
+    \Nbody{} case (blue) the bar strength increases with time, consistent with
+    previous results showing that the bar strength increases as bars slow down.
+    In the \SMUGGLE{} case (orange), we see that the bar strength slightly
     decreasing with time. This is also consistent with the expected relation
@@
 Fig.~\ref{fig:strength} shows that the length of the bar in the \Nbody{} case
-grows over time while it remains roughly constant in the \SMUGGLE{} case. This is
-also consistent with previous numerical work, which found that bars tend to grow
-as they slow down and the radius of corotation increases
-\citep{2000ApJ...543..704D, 2003MNRAS.341.1179A}. The time evolution of the bar
-strength, defined as the maximum of $\left|A_2/A_0\right|$ as a function of
-radius, is shown in the lower panel of Fig.~\ref{fig:strength}. The quantity
-$\left|A_2/A_0\right|$ varies from $0$ to $1$, with larger values indicating a
-stronger bar pattern. We see that in the \Nbody{} case, $\left|A_2/A_0\right|$
-increases over time as the bar pattern slows. This is consistent with previous
-\Nbody{} simulations which showed a clear correlation between the bar pattern
-speed and the bar strength \citep[e.g.,][]{2003MNRAS.341.1179A}. In the \SMUGGLE{}
-case, we see that the bar strength has an initial drop but then remains at a
-roughly constant, but slightly decreasing, strength. This is consistent with the
-pattern speed in the \SMUGGLE{} case being roughly constant or slightly increasing.
+grows over time while it remains roughly constant in the \SMUGGLE{} case. This
+is also consistent with previous numerical work, which found that bars tend to
+grow as they slow down and the radius of corotation increases
+\citep{2000ApJ...543..704D, 2003MNRAS.341.1179A}. {\bf The middle panel of
+Fig.~\ref{fig:strength} shows the mass of the bar. As the \Nbody{} bar slows
+down and lengthens, it also grows in mass. The \SMUGGLE{} bar, however, loses
+mass over time. This indicates an ongoing shredding process, and highlights the
+fact that bars are not solid bodies and can respond to external torques through
+mass redistribution.}
+
+The time evolution of the bar strength, defined as the maximum of
+$\left|A_2/A_0\right|$ as a function of radius, is shown in the lower panel of
+Fig.~\ref{fig:strength}. The quantity $\left|A_2/A_0\right|$ varies from $0$ to
+$1$, with larger values indicating a stronger bar pattern. We see that in the
+\Nbody{} case, $\left|A_2/A_0\right|$ increases over time as the bar pattern
+slows. This is consistent with previous \Nbody{} simulations which showed a
+clear correlation between the bar pattern speed and the bar strength
+\citep[e.g.,][]{2003MNRAS.341.1179A}. In the \SMUGGLE{} case, we see that the
+bar strength has an initial drop but then remains at a roughly constant, but
+slightly decreasing, strength. This is consistent with the pattern speed in the
+\SMUGGLE{} case being roughly constant or slightly increasing.
 

### 2023-04-19 paper/ms.tex
Add trapped fraction argument

@@
 \texttt{scipy} using a window length of $21$ and polynomial order of $3$ to the
-plot of torques in order to remove some numerical noise.
+plot of torques (Fig.~\ref{fig:prop}) and angle differences
+(Fig.~\ref{fig:wake}) in order to remove some numerical noise.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{fig/halo_trapped.pdf}
+    \caption{The trapped fraction of the dark matter halo within two disk scale
+    lengths ($\sim5.3\,\textrm{kpc}$) for the \Nbody{} and \SMUGGLE{} runs. As
+    the dark matter halo torques the bar, material from the halo is trapped on
+    bar-like orbits. In the \Nbody{} case, the trapped fraction increases with
+    time, indicating the torquing process is active. In the \SMUGGLE{} case, the
+    trapped fraction is nearly constant with time, indicating the torquing
+    process is inactive.}
+    \label{fig:htrap}
+\end{figure}
 
@@
 
-We test this interpretation by measuring the angle offset between the halo
-wake and the bar. If the wake and the bar are aligned (i.e., there is no angle
-offset), then the wake cannot exert a negative torque on the bar. This angle
-is plotted in the left panel of Fig.~\ref{fig:wake}, which shows that the
-angle offset is larger in the \Nbody{} case than in the \SMUGGLE{} case by about a
-factor of two. The center and right panels of Fig.~\ref{fig:wake} show the halo
-wake with respect to the location of the bar in the \Nbody{} (center) and \SMUGGLE{}
-(right) cases at one point in time.
+{\bf We test this interpretation in two ways. First, we measure the trapped
+fraction of the halo. As material in the halo wake torques the bar, that
+material becomes trapped on bar-like orbits \citep[the ``shadow
+bar'';][]{2016MNRAS.463.1952P}. In Fig.~\ref{fig:htrap} we show the halo trapped
+fraction for particles with radii less than two disk scale lengths
+($\sim5.3\,\textrm{kpc}$). In the \Nbody{} case, the trapped fraction increases
+with time, as expected since the halo is actively torquing the bar. In the
+\SMUGGLE{} case, the trapped fraction is constant (or perhaps slightly
+decreasing) with time. This supports our interpertation that the halo wake
+process has shut down in the presence of the gas phase.}
+
+Second, we measure the angle offset between the halo wake and the bar. If the
+wake and the bar are aligned (i.e., there is no angle offset), then the wake
+cannot exert a negative torque on the bar. This angle is plotted in the left
+panel of Fig.~\ref{fig:wake}, which shows that the angle offset is larger in the
+\Nbody{} case than in the \SMUGGLE{} case by about a factor of two. The center
+and right panels of Fig.~\ref{fig:wake} show the halo wake with respect to the
+location of the bar in the \Nbody{} (center) and \SMUGGLE{} (right) cases at one
+point in time.
 

### 2023-04-19 paper/ms.tex
Add 7% gas fraction evol

@@
 
-As a result, we conclude that for the disk, bar, and halo properties
-considered in this work, a gas fraction of only approximately $5\%$ is
-necessary in order for the proposed stabilizing mechanism to operate.
+{\bf After $5\,\textrm{Gyr}$, the $10\,\Msun/\textrm{pc}^2$ simulation has a gas
+fraction of $5.7\%$, but still exhibits constant pattern speed behavior.} As a
+result, we conclude that for the disk, bar, and halo properties considered in
+this work, a gas fraction of only approximately $5\%$ is necessary in order for
+the proposed stabilizing mechanism to operate.
 

### 2023-04-20 paper/ms.tex
Update observational section with lenticular gal

@@
 
+{\bf Lenticular galaxies which lack a signficant gas phase offer an opportunity
+to find slowly rotating bars. It would still take several Gyr for a galaxy
+hosting a fast bar to transition to the slow bar regime, so slow bars should
+only occupy lenticular galaxies which have been lenticular for some time.
+NGC~4277 is one such example, whose bar has been found to rotate with
+$\Rot{}\sim1.8$ \citep{2022AA...664L..10B}. On the other hand, NGC~4264 has a
+fast bar with $\Rot{}\sim1$ \citep{2019MNRAS.488.4972C}. The difference has been
+explained by differences in the dark matter content of the galaxies
+\citep{2023MNRAS.521.2227B}. We offer another explanation based on the timing of
+when gas was stripped from these galaxies.}
+
 \citet{2020MNRAS.495.4158F} find that quenched galaxies tend to host longer bars
@@
 scope of our current work. Nonetheless, our prediction that \Rot{} is stable in
-the presence of gas is robust. Why the $N$-body simulation forms a bar with
-$\Rot\sim1.6$ is a separate question deserving further attention.}
+the presence of gas is robust. Assuming bars form with $\Rot\sim1$, we predict
+this should remain the case with further evolution. Why the $N$-body simulation
+forms a bar with $\Rot\sim1.6$ is a separate question deserving further
+attention.}
 

### 2023-04-20 paper/ms.tex
Add non-lenticular slow rotators

@@
 
+{\bf There are further examples of gas-rich galaxies hosting slow bars. For
+example, UGC~628 \citep[$\Rot\sim2$;][]{2009AA...499L..25C} and NGC~2915
+\citep[$\Rot>1.7$;][]{1999AJ....118.2158B}. UGC~628
+has been studied in detail by \citet{2016MNRAS.463.1751C}, who note that it
+indeed has a low gas fraction for galaxies of its type. NGC~2915 has a gas
+fraction of $0.7$ \citep{2010ApJ...715..656W}, which would seem to be in
+conflict with our prediction that only a $5\%$ gas fraction is necessary to
+arrest the halo slowdown process. However, NGC~2915 has significantly different
+structural properties than the Milky Way-like model we considered in this work.
+In particular, it has a signficantly lower mass ($\sim10^9\Msun$ compared to
+$4.8\times10^{10}\Msun$ in our model). Further work is necessary to see how the
+gas fraction cutoff varies with galactic properties.}
+
 \citet{2020MNRAS.495.4158F} find that quenched galaxies tend to host longer bars
@@
 rotating. Indeed, they would need to be gas-deficient for several $\textrm{Gyr}$
-before they would be classified as slow rotators.} There does exist one example
-of a galaxy which is known to be a slow rotator -- the low surface brightness
-galaxy UGC 628 \citep{2009AA...499L..25C}. This galaxy has been studied in
-detail by \citet{2016MNRAS.463.1751C}, who note that it indeed has a low gas
-fraction for galaxies of its type. We predict a general trend that bars in
-gas-rich spiral galaxies should rotate quickly while some bars in gas-poor
-spiral galaxies should rotate slowly.
+before they would be classified as slow rotators.} {\bf There does exist
+examples of galaxies known to be slow rotators -- the low surface brightness
+galaxy UGC~628 \citep{2009AA...499L..25C}, lenticular galaxy NGC~4277
+\citep{2022AA...664L..10B}, and NGC~2915 \citep{1999AJ....118.2158B}. UGC~628
+has been studied in detail by \citet{2016MNRAS.463.1751C}, who note that it
+indeed has a low gas fraction for galaxies of its type. NGC~2915 has a gas
+fraction of $0.7$ \citep{2010ApJ...715..656W}, which would seem to be in
+conflict with our prediction that only a $5\%$ gas fraction is necessary to
+arrest the halo slowdown process. However, NGC~2915 has significantly different
+structural properties than the Milky Way-like model we considered in this work.
+In particular, it has a signficantly lower mass ($\sim10^9\Msun$ compared to
+$4.8\times10^{10}\Msun$ in our model). Further work is necessary to see how the
+gas fraction cutoff varies with galactic properties. TODO: shorten this} We predict a general trend
+that bars in gas-rich spiral galaxies should rotate quickly while some bars in
+gas-poor spiral galaxies should rotate slowly.
 

### 2023-04-20 paper/ms.tex
Reformat paragraph

@@
 
-In the last decade, stellar bars have been studied mainly in the
-context of the instability processes that lead to their formation and their
-ability to drive gas toward the galaxy center and contribute to the formation of
-supermassive black holes (SMBH). The primary interest of these studies mainly
-was in the loss of angular momentum of the gas and the associated  gaseous flow
-down to the inner disk, possibly forming a central mass concentration
-\citep{2010ApJ...719.1470V}, and fueling the central SMBH \citep[e.g.][]{1989Natur.338...45S, 1990Natur.345..679S}. A revisit of the more general problem of
-galactic bar properties and formation, including the case of disk galaxies with
-very large gas fraction is timely since it is clear now that galactic disks show
-massive bars already at redshift $z>2$ \citep{2022arXiv221008658G}. At
-that time the universe was 2.5 billion year old and galaxies might have as high
-gas fraction as 80\% \citep{2020ARAA..58..157T}. Furthermore, unlike
-nearby disk galaxies, the high-redshift disks also continuously accrete cold gas
-from the cosmic web, making the formation, stability and evolution of
-non-axisymmetric features a key question to address since they can play a
-fundamental role in the more general problem of disk galaxy evolution.
+In the last decade, stellar bars have been studied mainly in the context of the
+instability processes that lead to their formation and their ability to drive
+gas toward the galaxy center and contribute to the formation of supermassive
+black holes (SMBH). The primary interest of these studies mainly was in the loss
+of angular momentum of the gas and the associated  gaseous flow down to the
+inner disk, possibly forming a central mass concentration
+\citep{2010ApJ...719.1470V}, and fueling the central SMBH
+\citep[e.g.][]{1989Natur.338...45S, 1990Natur.345..679S}. A revisit of the more
+general problem of galactic bar properties and formation, including the case of
+disk galaxies with very large gas fraction is timely since it is clear now that
+galactic disks show massive bars already at redshift $z>2$
+\citep{2022arXiv221008658G}. At that time the universe was 2.5 billion year old
+and galaxies might have as high gas fraction as 80\% \citep{2020ARAA..58..157T}.
+Furthermore, unlike nearby disk galaxies, the high-redshift disks also
+continuously accrete cold gas from the cosmic web, making the formation,
+stability and evolution of non-axisymmetric features a key question to address
+since they can play a fundamental role in the more general problem of disk
+galaxy evolution.
 

### 2023-04-20 paper/ms.tex
Move stars instead of gas to appendix

@@
 
-{\bf \subsection{Stars Instead of Gas}
-In the SMUGGLE model considered in this work, we instantaneously added gas to
-the $N$-body system after $1.5\,\textrm{Gyr}$ of evolution. One might wonder if
-this sudden change to the potential is responsible for the stable pattern speed
-evolution. To test whether this is the case, we added mass to the system in the
-same way we did for the SMUGGLE model, but using collisionless particles instead
-of gas. The result of this experiment is shown in Fig.~\ref{fig:ps-star}. While
-there is an offset compared to the pure $N$-body case, we see that the pattern
-speed evolution is broadly consistent with a declining pattern speed. This
-indicates that the gas phase is responsible for the stable pattern speed.}
-
 \begin{figure}
@@
 \label{fig:GFM}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=9cm]{fig/ps_star.pdf}
-    \caption{{\bf Pattern speed evolution of a model in which we instaneously add stars instead of gas to the simulation, with the same density profile as the gas phase.
-The pattern speed evolution in this case is qualitatively similar to that of the $N$-body case, with a slight offset in the pattern speed.
-This test demonstrates that the stable pattern speed evolution in the SMUGGLE case is not simply a consequence of the change in potential imposed in our initial conditions.}}
-\label{fig:ps-star}
 \end{figure}
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=9cm]{fig/ps_star.pdf}
+    \caption{{\bf Pattern speed evolution of a model in which we instaneously
+    add stars instead of gas to the simulation, with the same density profile as
+    the gas phase. The pattern speed evolution in this case is qualitatively
+    similar to that of the $N$-body case, with a slight offset in the pattern
+    speed. This test demonstrates that the stable pattern speed evolution in the
+    SMUGGLE case is not simply a consequence of the change in potential imposed
+    in our initial conditions.}}
+\label{fig:ps-star}
+\end{figure}
+
+{\bf \section{Stars Instead of Gas}
+In the SMUGGLE model considered in this work, we instantaneously added gas to
+the $N$-body system after $1.5\,\textrm{Gyr}$ of evolution. One might wonder if
+this sudden change to the potential is responsible for the stable pattern speed
+evolution. To test whether this is the case, we added mass to the system in the
+same way we did for the SMUGGLE model, but using collisionless particles instead
+of gas. The result of this experiment is shown in Fig.~\ref{fig:ps-star}. While
+there is an offset compared to the pure $N$-body case, we see that the pattern
+speed evolution is broadly consistent with a declining pattern speed. This
+indicates that the gas phase is responsible for the stable pattern speed.}
 \section{Semi-Analytic Model Parameters}

### 2023-04-21 paper/ms.tex
Add citation to Combes and Sanders

@@
 that galactic bars should resonantly interact with the dark matter halo, causing
-the bar to slow down over time \citep{1992ApJ...400...80H, 2000ApJ...543..704D,
+the bar to slow down over time \citep{1981AA....96..164C, 1992ApJ...400...80H, 2000ApJ...543..704D,
 2002MNRAS.330...35A, 2002ApJ...569L..83A, 2003MNRAS.341.1179A,

### 2023-04-21 paper/ms.tex
Expand on discussion of previous literature

@@
 
+\subsection{Previous Simulation Work}
+Substantial work has been devoted to the role of gas in bar dynamics.
+
+To our knowledge the first work on a barred galaxy with a gas component was by
+\citet{1993AA...268...65F}. They found a stable pattern speed when a dissipative
+component was added to the system, with a slight increase in the pattern speed
+near the end of their simulation. However, their model was only evolved for
+$\sim1\,\textrm{Gyr}$, so it is unclear if their bar exhibits a stable pattern
+speed over several Gyr.
+
+\citet{2007ApJ...666..189B} describe a model containing up to $8\%$ gas. Their
+disk is similar to ours (though their dark matter halo is a factor of $10$ less
+massive). They find slowdown behavior up to a gas fraction of $8\%$, with the
+finding that higher gas fractions lead to a reduced slowdown. However, for the
+models which they show pattern speeds they do not include star formation or the
+removal of gas from the center of their disk. In preliminary work, we found
+slowdown behavior in an adiabatic model with similar gas fractions but lacking
+any method for removing gas from the central region (in our fiducial SMUGGLE
+model this is achieved through star formation). We surmise that the removal of
+gas from the central region is an important requisite for stable pattern speeds,
+though a careful torque analysis is necessary to be sure.
+
+\citet{2010ApJ...719.1470V} do find stable pattern speeds in models with gas
+fractions as low as $8\%$ (depending on the force softening used). Crucially,
+their model does contain a routine for removing gas from the central region of
+the disk. These authors state the behavior is bimodal, with a clear stable
+regime and a slowdown regime. These authors make no mention of the process by
+which the halo braking process is arrested, as we propose in the present paper.
+
+\citet{2013MNRAS.429.1949A, 2014MNRAS.438L..81A} find stable evolution in \Rot{}
+for a triaxial halo over $10\,\textrm{Gyr}$ with initial (final) gas fractions
+of $100\%$ ($7\%$), $75\%$ ($6\%$), and $50\%$ ($5\%$). For a model with $20\%$
+($3\%$) gas fraction they find an increasing \Rot{}. In their model with a
+spherical halo they always find increasing \Rot{}, contrary to the present work.
+We note there is a structural difference in the dark matter halo compared to the present work.
+They use a cored isothermal sphere as opposed to our Hernquist halo -- it is not clear to us the impact this would have on the expected These authors make no
+mention of the process by which the halo braking process is arrested, as we
+propose in the present paper.
+
+\citet{2015MNRAS.454.3166A} describe a model containing gas and stars which
+slows down with time. In their model gas is added to the system to target a gas
+fraction of $10\%$. However, their disk is about a factor of $10$ less massive
+than the disk considered in this work. The necessary gas fraction for a stable
+pattern speed probably depends on galaxy properties like mass. It is unclear
+whether $10\%$ is sufficient for their bar to have a stable pattern speed.
+
 \subsection{Cosmological Simulations}

### 2023-04-21 paper/ms.tex
Update formatting

@@
 that galactic bars should resonantly interact with the dark matter halo, causing
-the bar to slow down over time \citep{1981AA....96..164C, 1992ApJ...400...80H, 2000ApJ...543..704D,
-2002MNRAS.330...35A, 2002ApJ...569L..83A, 2003MNRAS.341.1179A,
-2003MNRAS.346..251O, 2005MNRAS.363..991H, 2006ApJ...637..214M,
-2007MNRAS.375..460W, 2009ApJ...697..293D}. The physical mechanism of this
-interaction can be understood as an angular form of dynamical friction between
-the bar and the dark matter halo. While studied in detail for the bar
-\citep{1984MNRAS.209..729T, 1985MNRAS.213..451W}, this process is generic for
-any non-axisymmetric disturbance \citep{1972MNRAS.157....1L}. (For an old but
-still useful review of bar dynamics, see \citet{1993RPPh...56..173S}.)
+the bar to slow down over time \citep{1981AA....96..164C, 1992ApJ...400...80H,
+2000ApJ...543..704D, 2002MNRAS.330...35A, 2002ApJ...569L..83A,
+2003MNRAS.341.1179A, 2003MNRAS.346..251O, 2005MNRAS.363..991H,
+2006ApJ...637..214M, 2007MNRAS.375..460W, 2009ApJ...697..293D}. The physical
+mechanism of this interaction can be understood as an angular form of dynamical
+friction between the bar and the dark matter halo. While studied in detail for
+the bar \citep{1984MNRAS.209..729T, 1985MNRAS.213..451W}, this process is
+generic for any non-axisymmetric disturbance \citep{1972MNRAS.157....1L}. (For
+an old but still useful review of bar dynamics, see
+\citet{1993RPPh...56..173S}.)
 

### 2023-04-21 paper/ms.tex
Bold previous work section

@@
 \subsection{Previous Simulation Work}
-Substantial work has been devoted to the role of gas in bar dynamics.
-
-To our knowledge the first work on a barred galaxy with a gas component was by
-\citet{1993AA...268...65F}. They found a stable pattern speed when a dissipative
-component was added to the system, with a slight increase in the pattern speed
-near the end of their simulation. However, their model was only evolved for
-$\sim1\,\textrm{Gyr}$, so it is unclear if their bar exhibits a stable pattern
-speed over several Gyr.
-
-\citet{2007ApJ...666..189B} describe a model containing up to $8\%$ gas. Their
-disk is similar to ours (though their dark matter halo is a factor of $10$ less
-massive). They find slowdown behavior up to a gas fraction of $8\%$, with the
-finding that higher gas fractions lead to a reduced slowdown. However, for the
-models which they show pattern speeds they do not include star formation or the
-removal of gas from the center of their disk. In preliminary work, we found
+{\bf Substantial work has been devoted to the role of gas in bar dynamics. We
+discuss this and highlight the novel aspects of the present investigation.}
+
+{\bf To our knowledge the first work on a barred galaxy with a gas component was
+by \citet{1993AA...268...65F}. They found a stable pattern speed when a
+dissipative component was added to the system, with a slight increase in the
+pattern speed near the end of their simulation. However, their model was only
+evolved for $\sim1\,\textrm{Gyr}$, so it is unclear if their bar exhibits a
+stable pattern speed over several Gyr.}
+
+{\bf \citet{2007ApJ...666..189B} describe a model containing up to $8\%$ gas.
+Their disk is similar to ours (though their dark matter halo is a factor of $10$
+less massive). They find slowdown behavior up to a gas fraction of $8\%$, with
+the finding that higher gas fractions lead to a reduced slowdown. However, for
+the models which they show pattern speeds they do not include star formation or
+the removal of gas from the center of their disk. In preliminary work, we found
 slowdown behavior in an adiabatic model with similar gas fractions but lacking
@@
 gas from the central region is an important requisite for stable pattern speeds,
-though a careful torque analysis is necessary to be sure.
-
-\citet{2010ApJ...719.1470V} do find stable pattern speeds in models with gas
-fractions as low as $8\%$ (depending on the force softening used). Crucially,
-their model does contain a routine for removing gas from the central region of
-the disk. These authors state the behavior is bimodal, with a clear stable
-regime and a slowdown regime. These authors make no mention of the process by
+though a careful torque analysis is necessary to be sure.}
+
+{\bf \citet{2010ApJ...719.1470V} do find stable pattern speeds in models with
+gas fractions as low as $8\%$ (depending on the force softening used).
+Crucially, their model does contain a routine for removing gas from the central
+region of the disk. These authors state the behavior is bimodal, with a clear
+stable regime and a slowdown regime. These authors make no mention of the
+process by which the halo braking process is arrested, as we propose in the
+present paper.}
+
+{\bf \citet{2013MNRAS.429.1949A, 2014MNRAS.438L..81A} find stable evolution in
+\Rot{} for a triaxial halo over $10\,\textrm{Gyr}$ with initial (final) gas
+fractions of $100\%$ ($7\%$), $75\%$ ($6\%$), and $50\%$ ($5\%$). For a model
+with $20\%$ ($3\%$) gas fraction they find an increasing \Rot{}. In their model
+with a spherical halo they always find increasing \Rot{}, contrary to the
+present work. We note there is a structural difference in their dark matter
+halo. They use a cored isothermal sphere as opposed to our Hernquist halo It is
+not clear to us the impact this would have on the expected torque from the halo
+(i.e., the velocity structure of their halo may allow for more efficient capture
+and thus stronger torquing). These authors make no mention of the process by
 which the halo braking process is arrested, as we propose in the present paper.
-
-\citet{2013MNRAS.429.1949A, 2014MNRAS.438L..81A} find stable evolution in \Rot{}
-for a triaxial halo over $10\,\textrm{Gyr}$ with initial (final) gas fractions
-of $100\%$ ($7\%$), $75\%$ ($6\%$), and $50\%$ ($5\%$). For a model with $20\%$
-($3\%$) gas fraction they find an increasing \Rot{}. In their model with a
-spherical halo they always find increasing \Rot{}, contrary to the present work.
-We note there is a structural difference in the dark matter halo compared to the present work.
-They use a cored isothermal sphere as opposed to our Hernquist halo -- it is not clear to us the impact this would have on the expected These authors make no
-mention of the process by which the halo braking process is arrested, as we
-propose in the present paper.
-
-\citet{2015MNRAS.454.3166A} describe a model containing gas and stars which
+\textcolor{red}{Lars, Elena - I'm most worried about this paragraph. Is the
+argument about cored isothermal sphere reasonable?} }
+
+{\bf \citet{2015MNRAS.454.3166A} describe a model containing gas and stars which
 slows down with time. In their model gas is added to the system to target a gas
@@
 pattern speed probably depends on galaxy properties like mass. It is unclear
-whether $10\%$ is sufficient for their bar to have a stable pattern speed.
+whether $10\%$ is sufficient for their bar to have a stable pattern speed.}
 

### 2023-04-21 paper/ms.tex
Further updates to previous simulation work

@@
 \textcolor{red}{Lars, Elena - I'm most worried about this paragraph. Is the
-argument about cored isothermal sphere reasonable?} }
+argument about cored isothermal sphere reasonable? I'm unsure if the velocity
+structure between a cored isothermal sphere and a Hernquist profile is really
+different enough to matter.} }
+
+{\bf There may also be issues related to structural differences between their
+bars and the bar considered in this work. In our case, we allowed the bar to
+form in an $N$-body run and then added gas after the bar formation. In
+\citet{2013MNRAS.429.1949A, 2014MNRAS.438L..81A} the bar forms from a disk that
+is initially gas-rich. Neither approach is inherently better, but it is known
+that in the latter case the resultant bar strength is weaker for initially
+gas-rich systems \citep[e.g.,][]{2013MNRAS.429.1949A}. Weaker bars should be
+less efficient at driving gas inwards and thus experience less positive torque
+from the gas phase. A direct comparison based on the torque by the gas phase on
+the bar is necessary.}
 

### 2023-04-21 paper/ms.tex
Update discussion of cosmo sims

@@
 \citep{2017MNRAS.469.1054A, 2019MNRAS.483.2721P, 2021AA...650L..16F,
-2022ApJ...940...61F}. However, the pattern speeds of bars in both cosmological
-simulations and the real universe can be affected by environmental processes not
-included in our simulation -- e.g., satellite infall
-\citep{2011Natur.477..301P}, non-sphericity \citep{2013MNRAS.429.1949A},
-rotation in the dark matter halo \citep{2013MNRAS.434.1287S,
-2014ApJ...783L..18L, 2018MNRAS.476.1331C, 2019MNRAS.488.5788C}, or perhaps even
-the gaseous circumgalactic medium. Naturally, extending our present work to
-account for such effects is a crucial next step in understanding the formation
-and evolution of galactic bars. We are presently engaged in such an exploration.
-We note that \citet{2022ApJ...940...61F} argue that bars in the TNG50 simulation
-are too short rather than too slow.
+2022ApJ...940...61F}.\footnote{Though see \citet{2022ApJ...940...61F} who argue
+bars have consistent pattern speeds with observations, but are too short.}
+\citet{2015PASJ...67...63O} describe cosmological zoom simulations of two barred
+Milky Way-like galaxies that both slow down over time. As pointed out by
+\citet{2010ApJ...719.1470V}, the gas fraction cutoff for the stable pattern
+speed regime increases with lower softening lengths. \citet{2015PASJ...67...63O}
+use softening lengths larger than ours by about a factor of $6$. This highlights
+the importance of future work exploring precisely when barred galaxies ought to
+be in the stable regime.
+
+Furthermore, the pattern speeds of bars in both cosmological simulations and the
+real universe can be affected by environmental processes not included in our
+simulation -- e.g., satellite infall \citep{2011Natur.477..301P}, non-sphericity
+\citep{2013MNRAS.429.1949A}, rotation in the dark matter halo
+\citep{2013MNRAS.434.1287S, 2014ApJ...783L..18L, 2018MNRAS.476.1331C,
+2019MNRAS.488.5788C}, or perhaps even the gaseous circumgalactic medium.
+Naturally, extending our present work to account for such effects is a crucial
+next step in understanding the formation and evolution of galactic bars. We are
+presently engaged in such an exploration.
 

### 2023-04-21 paper/ms.tex
Add citation to Fragkoudi

@@
 
+\citet{2021AA...650L..16F} explored the evolution of bars in the Auriga
+cosmological zoom simulations \citep{2017MNRAS.467..179G}. They find \Rot{}
+values consistent with observations. Furthermore, their pattern speed evolution
+shows some apparent periods of stability (see their Fig.~B1). One galaxy, Au27,
+appears to even transition from the stable pattern speed regime to the slowing
+down regime at a lookback time of $\sim1.8\,\textrm{Gyr}$.
+
 Furthermore, the pattern speeds of bars in both cosmological simulations and the

### 2023-04-21 paper/ms.tex
Cite Regan and Teuben
Bold cosmo sims section

--- 2023-04-21 paper/ms.tex
@@
 
-\subsection{Previous Simulation Work}
+\subsection{Previous Idealized Simulation Work}
 {\bf Substantial work has been devoted to the role of gas in bar dynamics. We
@@
 that in the latter case the resultant bar strength is weaker for initially
-gas-rich systems \citep[e.g.,][]{2013MNRAS.429.1949A}. Weaker bars should be
-less efficient at driving gas inwards and thus experience less positive torque
-from the gas phase. A direct comparison based on the torque by the gas phase on
-the bar is necessary.}
+gas-rich systems \citep[e.g.,][]{2013MNRAS.429.1949A}. Weaker bars are less
+efficient at driving gas inwards \citep{2004ApJ...600..595R} and thus should
+experience less positive torque from the gas phase. A direct comparison based on
+the torque by the gas phase on the bar is necessary.}
 
@@
 bars have consistent pattern speeds with observations, but are too short.}
-\citet{2015PASJ...67...63O} describe cosmological zoom simulations of two barred
+{\bf \citet{2015PASJ...67...63O} describe cosmological zoom simulations of two barred
 Milky Way-like galaxies that both slow down over time. As pointed out by
@@
 the importance of future work exploring precisely when barred galaxies ought to
-be in the stable regime.
+be in the stable regime.}
+
+{\bf \citet{2021AA...650L..16F} explored the evolution of bars in the Auriga
+cosmological zoom simulations \citep{2017MNRAS.467..179G}. They find \Rot{}
+values consistent with observations. Furthermore, their pattern speed evolution
+shows some apparent periods of stability (see their Fig.~B1). One galaxy, Au27,
+appears to even transition from the stable pattern speed regime to the slowing
+down regime at a lookback time of $\sim1.8\,\textrm{Gyr}$.}
 

### 2023-04-21 plots/ps_star/plot_ps_star.py
Remove savgol in ps_star

@@ -142,10 +142,10 @@ def run():
     fig, ax = plt.subplots(1, 1, figsize=(columnwidth*cm, (2./3.)*columnwidth*cm))
 
     # First panel, pattern speed.
-    ax.plot(time_Nbody - time_Nbody[300], savgol_filter(ps_Nbody, 81, 3), c=tb_c[0], label=r'$N$-body')
-    ax.plot(time_SMUGGLE20, savgol_filter(ps_SMUGGLE20, 81, 3), c=tb_c[1], label=r'SMUGGLE')
+    ax.plot(time_Nbody - time_Nbody[300], ps_Nbody, c=tb_c[0], label=r'$N$-body')
+    ax.plot(time_SMUGGLE20, ps_SMUGGLE20, c=tb_c[1], label=r'SMUGGLE')
     # ax.plot(time_SMUGGLEs20, savgol_filter(ps_SMUGGLEs20, 21, 3), c=tb_c[2], label=r'SMUGGLE gas converted to stars')
-    ax.plot(time_SMUGGLEs20, savgol_filter(ps_SMUGGLEs20, 81, 3), c=tb_c[2], label=r'SMUGGLE gas converted to stars')
+    ax.plot(time_SMUGGLEs20, ps_SMUGGLEs20, c=tb_c[2], label=r'SMUGGLE gas converted to stars')
 
 
     ax.set(ylim=(0, 60), ylabel=r'$\Omega_p\,[\,\text{km}/\text{s}/\text{kpc}\,]$')

### 2023-04-21 paper/ms.tex
Update plotting details

@@
 \subsection{Plotting Details}
-We saved snapshots in intervals of $0.005$ in the time units of the
-simulation, $\textrm{kpc}/(\textrm{km}/\textrm{s})$, which is very nearly
-equal to $1\,\textrm{Gyr}$ (it is $\sim0.977\,\textrm{Gyr}$). Therefore,
-throughout this work we referred to the native code time unit as
-$\textrm{Gyr}$. None of our results are sensitive to this choice.
-
-We applied a Savitzky-Golay filter \citep{1964AnaCh..36.1627S} as implemented in
-\texttt{scipy} using a window length of $21$ and polynomial order of $3$ to the
-plot of torques (Fig.~\ref{fig:prop}) and angle differences
+We saved snapshots in intervals of $0.005$ in the time units of the simulation,
+$\textrm{kpc}/(\textrm{km}/\textrm{s})$, which is very nearly equal to
+$1\,\textrm{Gyr}$ (it is $\sim0.977\,\textrm{Gyr}$). Therefore, throughout this
+work we referred to the native code time unit as $\textrm{Gyr}$. None of our
+results are sensitive to this choice. We applied a Savitzky-Golay filter
+\citep{1964AnaCh..36.1627S} as implemented in \texttt{scipy} using a window
+length of $21$ and polynomial order of $3$ to the plot of torques
+(Fig.~\ref{fig:prop} and Fig.~\ref{fig:sam-torque}) and angle differences
 (Fig.~\ref{fig:wake}) in order to remove some numerical noise.
@@
 
-\begin{figure}
+\begin{figure*}
     \centering
@@
     \label{fig:equil}
-\end{figure}
-
-\begin{figure}
+\end{figure*}
+
+\begin{figure*}
     \centering
@@
 \label{fig:ps-star}
-\end{figure}
+\end{figure*}
 
@@
     \label{fig:surf}
-\end{figure}
+\end{figure*}
 

### 2023-04-21 paper/ms.tex
Update abstract

@@
   centers of approximately two-thirds of spiral disks in the nearby Universe.  
-  Due to gravitational
-  interactions between the bar and the other components of galaxies, it is
-  expected that angular momentum and matter will redistribute over long (Gyr)
-  timescales in barred galaxies. Previous work ignoring the gas phase of
-  galaxies has conclusively demonstrated that bars should slow their rotation
-  over time due to their interaction with dark matter halos. We have performed a
-  simulation of a Milky Way-like galactic disk hosting a strong bar which
-  includes a state-of-the-art model of the interstellar medium and a live dark
-  matter halo. In this simulation the bar pattern does not slow down over time,
-  and instead remains at a stable, constant rate of rotation. This behavior has
-  been observed in previous simulations using more simplified models for the
-  interstellar gas, but the apparent lack of secular evolution has remained
-  unexplained. We propose that the gas phase of the disk and the dark matter
-  halo act in concert to stabilize the bar pattern speed and prevent the bar
-  from slowing down or speeding up. We find that in a Milky Way-like disk, a gas
-  fraction of only about 5\% is necessary for this mechanism to operate. Our
+  Due to gravitational interactions between the bar and the other components of
+  galaxies, it is expected that angular momentum and matter will redistribute
+  over long (Gyr) timescales in barred galaxies. Previous work ignoring the gas
+  phase of galaxies has conclusively demonstrated that bars should slow their
+  rotation over time due to their interaction with dark matter halos. We have
+  performed a simulation of a Milky Way-like galactic disk hosting a strong bar
+  which includes a state-of-the-art model of the interstellar medium and a live
+  dark matter halo. In this simulation the bar pattern does not slow down over
+  time, and instead remains at a stable, constant rate of rotation. This
+  behavior has been observed in previous simulations using more simplified
+  models for the interstellar gas, but the apparent lack of secular evolution
+  has remained unexplained. {\bf We find that the presence of the gas phase
+  arrests the process by which the dark matter halo slows down a bar, a
+  phenomenon we term halo exhaustion. This exhaustion is responsible for
+  stabilizing the bar pattern speed.} We find that in a Milky Way-like disk, a
+  gas fraction of only about 5\% is necessary for this mechanism to operate. Our
   result naturally explains why nearly all observed bars rotate rapidly and is

### 2023-04-26 paper/ms.tex
Add citation to CALIFA

@@
 recently been applied to samples of galaxies from the MaNGA survey
-\citep{2019MNRAS.482.1733G, 2020MNRAS.491.3655G}. These studies confirm
-what was found in earlier works, that nearly all extragalactic bars are fast
-rotators (i.e., they rotate close to their maximum rotation rate).
+\citep{2019MNRAS.482.1733G, 2020MNRAS.491.3655G} {\bf and CALIFA survey
+\citep{2015AA...576A.102A}}. These studies confirm what was found in earlier
+works, that nearly all extragalactic bars are fast rotators (i.e., they rotate
+close to their maximum rotation rate).
 
@@
 non-axisymmetric feature. In a galaxy with a constant circular velocity \Vc{},
-it is given by $\RCR = \Vc / \PS$.} Galaxies with $\Rot < 1.4$ are considered
-``fast rotators'' while galaxies with $\Rot > 1.4$ are considered ``slow
-rotators'' \citep{2000ApJ...543..704D}. Galaxies with $\Rot < 1$ are not thought
-to be stable \citep{1980AA....81..198C}. Observational estimates of the pattern
-speeds of bars indicate that nearly all galaxies have $1 < \Rot < 1.4$
+it is given by $\RCR = \Vc / \PS$.} {\bf Bars} with $\Rot < 1.4$ are considered
+``fast rotators'' while {\bf bars} with $\Rot > 1.4$ are considered ``slow
+rotators'' \citep{2000ApJ...543..704D}. {\bf Bars} with $\Rot < 1$ are not
+thought to be stable \citep{1980AA....81..198C}. Observational estimates of the
+pattern speeds of bars indicate that nearly all galaxies have $1 < \Rot < 1.4$
 \citep{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G,
@@
 bar than its contribution to the mass of a galaxy would suggest
-\citep{2010ApJ...719.1470V, 2013MNRAS.429.1949A}.
+\citep[e.g.,][]{2010ApJ...719.1470V, 2013MNRAS.429.1949A}.
 
@@
 evolution in the bar pattern speed over several Gyr when the gas phase is
-accounted for and {\bf robustly} modeled. This behavior has been observed in a
-few previous works \citep{1993AA...268...65F, 2007ApJ...666..189B,
-2009ApJ...707..218V, 2010ApJ...719.1470V, 2014MNRAS.438L..81A}.
-
-Here we show that this phenomenon occurs in galaxies with gas fraction
-as low as about $5\%$, making this feature a likely characteristic of  Milky
-Way-type galaxies. This behavior will be explored in more general disks with
-larger gas fractions than the Milky Way in a forthcoming paper.
-% We are also the first to provide a physical explanation of
-%the behavior with clear observational predictions.
-We also provide a physical explanation of the redistribution of the angular
-momentum among the dark halo wake, the stellar bar, and the disk gas component.
+accounted for and {\bf robustly} modeled. This behavior has been observed in
+previous works \citep{1993AA...268...65F, 2007ApJ...666..189B,
+2009ApJ...707..218V, 2010ApJ...719.1470V, 2014MNRAS.438L..81A}. {\bf We propose a new
+physical explanation for this stable behavior.}
+
+{\bf In particular, we find that the presence of a significant gas phase in a
+barred galaxy can arrest the process by which the dark matter halo brakes the
+bar. The constant positive torque on the bar by the gas causes the bar's
+resonance to reside in regions which have become depopulated. In the halo wake
+picture, this is equivalent to no new material being available to reinforce the
+wake. We show that this occurs in Milky Way-like galaxies with gas fractions as
+low as about $5\%$.}
 
@@
     The solid lines indicate the torque exerted by the halo in both the \Nbody{}
-    and \SMUGGLE{} cases. The dashed line is the torque exerted by the gas phase in
-    the \SMUGGLE{} run (there is no gas in the \Nbody{} run). Details on how these
-    properties are calculated is given in Section~\ref{ssec:bar_analysis}.}\label{fig:prop}
+    and \SMUGGLE{} cases. The dashed line is the torque exerted by the gas phase
+    in the \SMUGGLE{} run (there is no gas in the \Nbody{} run). {\bf After
+    $\sim1\,\textrm{Gyr}$ of evolution, the torque by the halo in the \SMUGGLE{}
+    case is severely reduced. We call this halo exhaustion, and discuss its
+    proposed origin in the text.} Details on how these properties are calculated
+    is given in Section~\ref{ssec:bar_analysis}.}\label{fig:prop}
 \end{figure}
@@
 $0.5\,\textrm{kpc}$ from radii of $0$ to $30\,\textrm{kpc}$. We defined the
-angle of the bar $\phi_b$ to be twice the angle of the complex number $A_2$ as
-measured in the bin extending from a radius of $2.5$ to $3\,\textrm{kpc}$.
-After correcting for the periodicity of $\phi_b$, we measured the pattern
-speed as one-half the two-sided finite gradient of $\phi_b$ as a function of
-time.
+angle of the bar $\phi_b$ to be {\bf half} the angle of the complex number $A_2$
+as measured in the bin extending from a radius of $2.5$ to $3\,\textrm{kpc}$.
+After correcting for the periodicity of $\phi_b$, we measured the pattern speed
+as one-half the two-sided finite gradient of $\phi_b$ as a function of time.
 
@@
 the disk into a barred and unbarred component. We achieved this by following
-closely the methods described in \citet{2016MNRAS.463.1952P}. Our implementation
-is described in more detail in Appendix~\ref{app:bardecomp}. After the disk has
-been decomposed into a trapped and untrapped component, we measured the bar
-length as being the radius $R_b$ which encapsulates $99\%$ of the stars
-identified as being trapped in the bar.
+closely the methods described in \citet{2016MNRAS.463.1952P,
+2021MNRAS.500..838P}. Our implementation is described in more detail in
+Appendix~\ref{app:bardecomp}. After the disk has been decomposed into a trapped
+and untrapped component, we measured the bar length as being the radius $R_b$
+which encapsulates $99\%$ of the stars identified as being trapped in the bar.
 
@@
 the bar by the other components is simply the negative of the torque on the
-other components by the bar.
+other components by the bar. A similar analysis using basis function expansions
+was performed in \citet{2019MNRAS.490.3616P}.
 
@@
 work we referred to the native code time unit as $\textrm{Gyr}$. None of our
-results are sensitive to this choice. We applied a Savitzky-Golay filter
+results are sensitive to this choice. {\bf We applied a Savitzky-Golay filter
 \citep{1964AnaCh..36.1627S} as implemented in \texttt{scipy} using a window
@@
 (Fig.~\ref{fig:prop} and Fig.~\ref{fig:sam-torque}) and angle differences
-(Fig.~\ref{fig:wake}) in order to remove some numerical noise.
+(Fig.~\ref{fig:wake}) in order to remove some numerical noise.}
 
@@
     \includegraphics[width=\columnwidth]{fig/halo_trapped.pdf}
-    \caption{The trapped fraction of the dark matter halo within two disk scale
+    \caption{{\bf The trapped fraction of the dark matter halo within two disk scale
     lengths ($\sim5.3\,\textrm{kpc}$) for the \Nbody{} and \SMUGGLE{} runs. As
@@
     trapped fraction is nearly constant with time, indicating the torquing
-    process is inactive.}
+    process is inactive.}}
     \label{fig:htrap}
@@
 reinforce the wake, therefore arresting the process by which the halo can slow
-the bar down. We term this process ``halo exhaustion.'' Note that the resonances
-can be maintained even in the presence of the dissipative gaseous component if
-the gas is continuously accreted from the external environment, as is the case
-for massive disk galaxies at high redshift.
+the bar down. We term this process ``halo exhaustion.''
 
@@
 panel of Fig.~\ref{fig:wake}, which shows that the angle offset is larger in the
-\Nbody{} case than in the \SMUGGLE{} case by about a factor of two. The center
-and right panels of Fig.~\ref{fig:wake} show the halo wake with respect to the
-location of the bar in the \Nbody{} (center) and \SMUGGLE{} (right) cases at one
-point in time.
+\Nbody{} case than in the \SMUGGLE{} case by about a factor of {\bf three}. The
+center and right panels of Fig.~\ref{fig:wake} show the halo wake with respect
+to the location of the bar in the \Nbody{} (center) and \SMUGGLE{} (right) cases
+at one point in time.
 
@@
 \RCR{} is a parameter in the bar model from \citet{2022MNRAS.513..768C}, we
-allow the moment of inertia of the bar to vary with ${\RCR}^2$. To be more
-precise, we allow
+allow the moment of inertia of the bar to vary with ${\RCR}^2$. {\bf This is
+inspired by the fact that the moment of inertia of an ellipsoid scales with the
+sum of the square of its axes.} To be more precise, we allow
 \begin{equation}
@@
 We next take the $\tau_{\textrm{gas}}=0$ and $20$ cases from Fig.~\ref{fig:sam}
-and plot the halo torque evolution. This result, given in Fig.~\ref{fig:sam-torque},
-is comparable to the lower panel of Fig.~\ref{fig:prop}. We find that the
-$\tau_{\textrm{gas}}=0$ case compares favorably to the \Nbody{} case described
-in previous sections. The bar exerts a steady negative torque in this case (blue
-line). When a gas torque is included (orange lines), we find that the halo's
-torque becomes much weaker, similar to what we found in the \SMUGGLE{} case. The
-gas applies a steady positive torque by construction. Therefore, the exhaustion
-process is reproduced in our simple semi-analytic model.
+and plot the halo torque evolution. This result, given in
+Fig.~\ref{fig:sam-torque}, is comparable to the lower panel of
+Fig.~\ref{fig:prop}. We find that the $\tau_{\textrm{gas}}=0$ case compares
+favorably to the \Nbody{} case described in previous sections. The bar exerts a
+steady negative torque in this case (blue line). When a gas torque is included
+(orange lines), we find that the halo's torque becomes much weaker, similar to
+what we found in the \SMUGGLE{} case. The gas applies a steady positive torque
+by construction. Therefore, the exhaustion process is reproduced in our simple
+semi-analytic model. {\bf Curiously, we do not see the bimodal behavior in
+Fig.~\ref{fig:fgas} and described in \citet{2010ApJ...719.1470V}. It is not
+presently clear why this is the case.}
 
@@
 from kinematics and chemistry. We point out that these reported measurements are
-not direct measurements of the Milky Way bar's deceleration. Much like the
-simulations in the present work, the simulations of these two works do not
-properly account for the complicated formation process of the Galactic bar,
-which may leave imprints on the present day distribution of stars in spatial,
-kinematic, and chemical space. More investigation is necessary to reconcile the
-present work with these two well-executed manuscripts.
+not direct measurements of the Milky Way bar's deceleration. {\bf For instance,
+\citet{2021MNRAS.500.4710C} measures the pattern speed based on the asymmetry of
+the Hercules stream, but this can also be produced by spiral arms
+\citep{2018MNRAS.481.3794H}.} Much like the simulations in the present work, the
+simulations of these two works do not properly account for the complicated
+formation process of the Galactic bar, which may leave imprints on the present
+day distribution of stars in spatial, kinematic, and chemical space. More
+investigation is necessary to reconcile the present work with these two
+well-executed manuscripts.
 
@@
 indeed has a low gas fraction for galaxies of its type. NGC~2915 has a gas
-fraction of $0.7$ \citep{2010ApJ...715..656W}, which would seem to be in
+fraction of $70\%$ \citep{2010ApJ...715..656W}, which would seem to be in
 conflict with our prediction that only a $5\%$ gas fraction is necessary to
@@
 scope of our current work. Nonetheless, our prediction that \Rot{} is stable in
-the presence of gas is robust. Assuming bars form with $\Rot\sim1$, we predict
-this should remain the case with further evolution. Why the $N$-body simulation
-forms a bar with $\Rot\sim1.6$ is a separate question deserving further
-attention.}
+the presence of sufficient gas is robust. Assuming bars form with $\Rot\sim1$,
+we predict this should remain the case with further evolution. Why the $N$-body
+simulation forms a bar with $\Rot\sim1.6$ is a separate question deserving
+further attention.}
 
@@
 the finding that higher gas fractions lead to a reduced slowdown. However, for
-the models which they show pattern speeds they do not include star formation or
+the models which they show pattern speeds, they do not include star formation or
 the removal of gas from the center of their disk. In preliminary work, we found
@@
 present work. We note there is a structural difference in their dark matter
-halo. They use a cored isothermal sphere as opposed to our Hernquist halo It is
+halo. They use a cored isothermal sphere as opposed to our Hernquist halo. It is
 not clear to us the impact this would have on the expected torque from the halo
@@
 and thus stronger torquing). These authors make no mention of the process by
-which the halo braking process is arrested, as we propose in the present paper.
-\textcolor{red}{Lars, Elena - I'm most worried about this paragraph. Is the
-argument about cored isothermal sphere reasonable? I'm unsure if the velocity
-structure between a cored isothermal sphere and a Hernquist profile is really
-different enough to matter.} }
+which the halo braking process is arrested, as we propose in the present paper.}
 
@@
 values consistent with observations. Furthermore, their pattern speed evolution
-shows some apparent periods of stability (see their Fig.~B1). One galaxy, Au27,
+shows some apparent periods of stability (see their Fig.~B1). One galaxy, Au26,
 appears to even transition from the stable pattern speed regime to the slowing
@@
     SMUGGLE case is not simply a consequence of the change in potential imposed
-    in our initial conditions.}}
+    in our initial conditions.} \textcolor{red}{This will be updated with a
+    higher resolution run.}}
 \label{fig:ps-star}

### 2023-04-26 paper/ms.tex
Halo material removed sentence
Remove hi res sentence

@@
 to the location of the bar in the \Nbody{} (center) and \SMUGGLE{} (right) cases
-at one point in time.
+at one point in time. {\bf Note that in Fig.~\ref{fig:wake} we have removed the
+halo material trapped in the bar, which exerts no net torque on the bar.}
 
@@
     SMUGGLE case is not simply a consequence of the change in potential imposed
-    in our initial conditions.} \textcolor{red}{This will be updated with a
-    higher resolution run.}}
+    in our initial conditions.}}
 \label{fig:ps-star}

### 2023-05-05 paper/ms.tex
Add Mark refs, Petersen cite

@@
 treatment. A more detailed description of this model can be found in the
-flagship \SMUGGLE{} paper \citep{2019MNRAS.489.4233M}.
+flagship \SMUGGLE{} paper \citep{2019MNRAS.489.4233M}. {\bf A pedagological
+review of cosmological simulations of galaxy formation can be found in
+\citet{2020NatRP...2...42V}. }
 
@@
 as one-half the two-sided finite gradient of $\phi_b$ as a function of time.
+{\bf We note that using the second Fourier mode is a blunt tool, discussed in
+\citet{2019arXiv190308203P}.}
 
@@
 2022ApJ...940...61F}.\footnote{Though see \citet{2022ApJ...940...61F} who argue
-bars have consistent pattern speeds with observations, but are too short.}
-{\bf \citet{2015PASJ...67...63O} describe cosmological zoom simulations of two barred
-Milky Way-like galaxies that both slow down over time. As pointed out by
-\citet{2010ApJ...719.1470V}, the gas fraction cutoff for the stable pattern
-speed regime increases with lower softening lengths. \citet{2015PASJ...67...63O}
-use softening lengths larger than ours by about a factor of $6$. This highlights
-the importance of future work exploring precisely when barred galaxies ought to
-be in the stable regime.}
+bars have consistent pattern speeds with observations, but are too short.} {\bf
+These works examine bars in EAGLE \citep{2015MNRAS.450.1937C,
+2015MNRAS.446..521S}, Illustris \citep{2014Natur.509..177V,
+2014MNRAS.444.1518V}, and Illustris TNG50 \citep{2019MNRAS.490.3196P,
+2019MNRAS.490.3234N}. \citet{2015PASJ...67...63O} describe cosmological zoom
+simulations of two barred Milky Way-like galaxies that both slow down over time.
+As pointed out by \citet{2010ApJ...719.1470V}, the gas fraction cutoff for the
+stable pattern speed regime increases with lower softening lengths.
+\citet{2015PASJ...67...63O} use softening lengths larger than ours by about a
+factor of $6$. This highlights the importance of future work exploring precisely
+when barred galaxies ought to be in the stable regime.}
 

### 2023-05-05 paper/ms.tex
FM comments

@@
   arrests the process by which the dark matter halo slows down a bar, a
-  phenomenon we term halo exhaustion. This exhaustion is responsible for
-  stabilizing the bar pattern speed.} We find that in a Milky Way-like disk, a
-  gas fraction of only about 5\% is necessary for this mechanism to operate. Our
-  result naturally explains why nearly all observed bars rotate rapidly and is
+  phenomenon we term bar locking. This locking is responsible for stabilizing
+  the bar pattern speed.} We find that in a Milky Way-like disk, a gas fraction
+  of only about 5\% is necessary for this mechanism to operate. Our result
+  naturally explains why nearly all observed bars rotate rapidly and is
   especially relevant for our understanding of how the Milky Way arrived at its
@@
     $\sim1\,\textrm{Gyr}$ of evolution, the torque by the halo in the \SMUGGLE{}
-    case is severely reduced. We call this halo exhaustion, and discuss its
-    proposed origin in the text.} Details on how these properties are calculated
-    is given in Section~\ref{ssec:bar_analysis}.}\label{fig:prop}
+    case is severely reduced. We call this bar locking, and discuss its proposed
+    origin in Section~\ref{sec:discussion}. Details on the calculation of the
+    torque and pattern speed is given in
+    Section~\ref{ssec:bar_analysis}.}}\label{fig:prop}
 \end{figure}
@@
 $25.3\times10^6$ in the dark matter halo. We used a softening length of
-$20\,\textrm{pc}$ for all components. \edit1{This softening length is smaller
-than used in the original GALAKOS model ($28\,\textrm{pc}$ in their model, but
-$43.5\,\textrm{pc}$ when scaled to our mass resolution). Our smaller softening
-length is consistent with other resolved ISM models \citep{2018MNRAS.480..800H,
-2019MNRAS.489.4233M}. We reran our fiducial SMUGGLE run with $40\,\textrm{pc}$
-softening and found no difference in the pattern speed evolution of the bar.}
-Snapshots were saved at equal intervals of $0.005$ in the time units of the
-simulation, $\textrm{kpc}/(\textrm{km}/\textrm{s})$.
+$20\,\textrm{pc}$ for all collisionless components. {\bf This softening length
+is smaller than used in the original GALAKOS model ($28\,\textrm{pc}$ in their
+model, but $\sim43.5\,\textrm{pc}$ when scaled to our mass resolution). Our
+smaller softening length is consistent with other resolved ISM models
+\citep{2018MNRAS.480..800H, 2019MNRAS.489.4233M}. However, as a consistency
+check, we reran our fiducial SMUGGLE run with $40\,\textrm{pc}$ softening and
+found no difference in the pattern speed evolution of the bar. For the gas
+component, the softening length is fully adaptive with a softening factor of
+$2.5$ \citep[e.g.,][]{2020ApJS..248...32W}.} Snapshots were saved at equal
+intervals of $0.005$ in the time units of the simulation,
+$\textrm{kpc}/(\textrm{km}/\textrm{s})$.
 
@@
     \includegraphics[width=\columnwidth]{fig/halo_trapped.pdf}
-    \caption{{\bf The trapped fraction of the dark matter halo within two disk scale
-    lengths ($\sim5.3\,\textrm{kpc}$) for the \Nbody{} and \SMUGGLE{} runs. As
+    \caption{{\bf The mass fraction within two disk scale lengths
+    ($\sim5.3\,\textrm{kpc}$) trapped in the \Nbody{} and \SMUGGLE{} runs. As
     the dark matter halo torques the bar, material from the halo is trapped on
     bar-like orbits. In the \Nbody{} case, the trapped fraction increases with
-    time, indicating the torquing process is active. In the \SMUGGLE{} case, the
-    trapped fraction is nearly constant with time, indicating the torquing
-    process is inactive.}}
+    time, indicating the torquing process is active and the bar is unlocked. In
+    the \SMUGGLE{} case, the trapped fraction is nearly constant with time,
+    indicating the torquing process is inactive and the bar is locked.}}
     \label{fig:htrap}
@@
 
-{\bf We test this interpretation in two ways. First, we measure the trapped
-fraction of the halo. As material in the halo wake torques the bar, that
+{\bf We test this interpretation in two ways. First, we measure the fraction of
+mass trapped in the halo. As material in the halo wake torques the bar, that
 material becomes trapped on bar-like orbits \citep[the ``shadow
@@
 ($\sim5.3\,\textrm{kpc}$). In the \Nbody{} case, the trapped fraction increases
-with time, as expected since the halo is actively torquing the bar. In the
-\SMUGGLE{} case, the trapped fraction is constant (or perhaps slightly
-decreasing) with time. This supports our interpertation that the halo wake
-process has shut down in the presence of the gas phase.}
+with time, as expected since the halo is actively torquing the bar (which is,
+therefore, unlocked). In the \SMUGGLE{} case, the trapped fraction is constant
+(or perhaps slightly decreasing) with time (indicating the bar is locked). This
+supports our interpertation that the halo wake process has shut down in the
+presence of the gas phase.}
 
@@
 \subsection{Observations}
+\label{ssec:observations}
 Observational estimates of the pattern speeds of bars indicate that nearly all
@@
 $4.8\times10^{10}\Msun$ in our model). Further work is necessary to see how the
-gas fraction cutoff varies with galactic properties.}
+gas fraction threshold varies with galactic properties.}
 
@@
 gas from the central region is an important requisite for stable pattern speeds,
-though a careful torque analysis is necessary to be sure.}
+though a careful torque analysis is necessary to confirm this hypothesis.}
 
@@
 indeed has a low gas fraction for galaxies of its type. NGC~2915 has a gas
-fraction of $0.7$ \citep{2010ApJ...715..656W}, which would seem to be in
+fraction of $70\%$ \citep{2010ApJ...715..656W}, which would seem to be in
 conflict with our prediction that only a $5\%$ gas fraction is necessary to
 arrest the halo slowdown process. However, NGC~2915 has significantly different
-structural properties than the Milky Way-like model we considered in this work.
-In particular, it has a signficantly lower mass ($\sim10^9\Msun$ compared to
-$4.8\times10^{10}\Msun$ in our model). Further work is necessary to see how the
-gas fraction cutoff varies with galactic properties. TODO: shorten this} We predict a general trend
-that bars in gas-rich spiral galaxies should rotate quickly while some bars in
+structural properties than the Milky Way-like model we considered in this work,
+discussed in Section~\ref{ssec:observations}.} We predict a general trend that
+bars in gas-rich spiral galaxies should rotate quickly while some bars in
 gas-poor spiral galaxies should rotate slowly.
@@
 $74\,\textrm{pc}$. These are somewhat lower than the observed value in the HI
-disk of $\sim200\,\textrm{kpc}$ \citep{1995ApJ...448..138M}. This may be caused
-by the model in our simulations not driving enough turbulent pressure, and is an
-interesting avenue of further investigation.}
+disk of $\sim200\,\textrm{pc}$ \citep{1995ApJ...448..138M, 2017AA...607A.106M}.
+This may be caused by the model in our simulations not driving enough turbulent
+pressure, and is an interesting avenue of further investigation.}
 

### 2023-05-05 paper/ms.tex
Change to bar locking

@@
 reinforce the wake, therefore arresting the process by which the halo can slow
-the bar down. We term this process ``halo exhaustion.''
-
-%In the \Nbody{} case, halo material that is
-%resonant with the bar will form a wake which lags behind and exerts a negative 
-%torque on the bar, slowing it down \citep{1984MNRAS.209..729T,
-%1985MNRAS.213..451W, 1992ApJ...400...80H}.\footnote{Since the bar is not a solid
-%body, it is not guaranteed that a negative torque will slow it down - e.g. a
-%negative torque could shred the bar, reducing its moment of inertia without
-%changing its pattern speed. However, the bar seems to empirically respond to a
-%negative torque induced by a halo wake by slowing down.} As the bar slows down,
-%the location of the resonances in phase space changes, allowing halo material
-%newly resonant with the bar to participate in forming the wake. However, the gas
-%is a reliable source of positive torque on the bar, speeding the bar up. This
-%stops the resonance location from changing such that the halo cannot reinforce
-%the wake, arresting the process by which the halo can slow the bar down. We term
-%this process ``halo exhaustion.''
-
-This halo exhaustion process is similar to the ``metastability'' effect, which
-has been previously discussed in the literature \citep{2003MNRAS.345..406V,
+the bar down. We term this process ``bar locking.''
+
+This bar locking process is similar to the ``metastability'' effect, which has
+been previously discussed in the literature \citep{2003MNRAS.345..406V,
 2006ApJ...639..868S}. Finally, we also note that these authors, in particular,
@@
 We argue that the following occurs. First, the bar is not able to slow down
-quickly enough due to the positive torque of the infalling gas. This causes
-the resonant halo phase space at a particular pattern speed, $\Omega_{\textrm{p},0}$,
-to become exhausted and no longer able to support a negative torque. Second,
-the gas is still exerting a positive torque on the bar, and therefore the
-pattern speed will again increase. Since at higher pattern speeds the halo has
-not yet been totally exhausted, the halo will once again be able to exert a
-negative torque on the bar. The pattern speed will then settle at a new value
-slightly higher than $\Omega_{\textrm{p},0}$ where the gas and halo torques cancel.
-Over time, the pattern speed should slowly increase as the halo becomes
-progressively exhausted.
+quickly enough due to the positive torque of the infalling gas. This causes the
+resonant halo phase space at a particular pattern speed,
+$\Omega_{\textrm{p},0}$, to become {\bf mixed} and no longer able to support a
+negative torque.\footnote{In the halo wake picture, this corresponds to the wake
+becoming fully aligned with the bar.} Second, the gas is still exerting a
+positive torque on the bar, and therefore the pattern speed will again increase.
+Since at higher pattern speeds the halo has not yet been totally {\bf mixed},
+the halo will once again be able to exert a negative torque on the bar. The
+pattern speed will then settle at a new value slightly higher than
+$\Omega_{\textrm{p},0}$ where the gas and halo torques cancel. Over time, the
+pattern speed should slowly increase.
 
@@
 what we found in the \SMUGGLE{} case. The gas applies a steady positive torque
-by construction. Therefore, the exhaustion process is reproduced in our simple
+by construction. Therefore, the locking process is reproduced in our simple
 semi-analytic model. {\bf Curiously, we do not see the bimodal behavior in

### 2023-05-05 paper/ms.tex
Update keywords

@@
 %% to include these concepts in their preprints.
-\keywords{Classical Novae (251) --- Ultraviolet astronomy(1736) --- History of astronomy(1868) --- Interdisciplinary astronomy(804)}
+\keywords{Milky Way Galaxy (1054) --- Milky Way evolution (1052) --- Milky Way dynamics (1051) --- Galaxy dynamics (591) --- Hydrodynamical simulations (767) --- Barred spiral galaxies (136)}
 

### 2023-05-30 paper/ms.tex
Remove all boldface

@@
   models for the interstellar gas, but the apparent lack of secular evolution
-  has remained unexplained. {\bf We find that the presence of the gas phase
+  has remained unexplained. We find that the presence of the gas phase
   arrests the process by which the dark matter halo slows down a bar, a
   phenomenon we term bar locking. This locking is responsible for stabilizing
-  the bar pattern speed.} We find that in a Milky Way-like disk, a gas fraction
+  the bar pattern speed. We find that in a Milky Way-like disk, a gas fraction
   of only about 5\% is necessary for this mechanism to operate. Our result
@@
     \includegraphics[width=\textwidth]{fig/fig1.pdf}
-    \caption{{\bf Stellar mass distribution of our simulations with and without
-    the interstellar medium.} The upper panels show an \Nbody{} only simulation
+    \caption{Stellar mass distribution of our simulations with and without
+    the interstellar medium. The upper panels show an \Nbody{} only simulation
     while the lower panels show a simulation which includes the \SMUGGLE{} model
     for the interstellar medium. Each panel is $20\,\textrm{kpc}$ to a side.
-    {\bf The bar length is shown as a solid white bar (details on its
+    The bar length is shown as a solid white bar (details on its
     computation are given in the text). The corotation radius is shown as a
-    dashed white circle.} Columns show different points in time, separated by
+    dashed white circle. Columns show different points in time, separated by
     $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
@@
     the same length and strength over the course of the simulation. The \Nbody{}
-    model is identical to the GALAKOS model {\bf (with different particle
-    numbers and softening lengths)}, discussed in the text.}\label{fig:overview}
+    model is identical to the GALAKOS model (with different particle
+    numbers and softening lengths), discussed in the text.}\label{fig:overview}
 \end{figure*}
@@
 almost always form bar-like structures \citep{1971ApJ...168..343H}. Several
-studies have shown that a {\bf hot, centrally concentrated mass distribution},
+studies have shown that a hot, centrally concentrated mass distribution,
 such as a stellar bulge or dark matter halo, acts to stabilize stellar disks
@@
 recently been applied to samples of galaxies from the MaNGA survey
-\citep{2019MNRAS.482.1733G, 2020MNRAS.491.3655G} {\bf and CALIFA survey
-\citep{2015AA...576A.102A}}. These studies confirm what was found in earlier
+\citep{2019MNRAS.482.1733G, 2020MNRAS.491.3655G} and CALIFA survey
+\citep{2015AA...576A.102A}. These studies confirm what was found in earlier
 works, that nearly all extragalactic bars are fast rotators (i.e., they rotate
@@
 non-axisymmetric feature. In a galaxy with a constant circular velocity \Vc{},
-it is given by $\RCR = \Vc / \PS$.} {\bf Bars} with $\Rot < 1.4$ are considered
-``fast rotators'' while {\bf bars} with $\Rot > 1.4$ are considered ``slow
-rotators'' \citep{2000ApJ...543..704D}. {\bf Bars} with $\Rot < 1$ are not
+it is given by $\RCR = \Vc / \PS$.} Bars with $\Rot < 1.4$ are considered
+``fast rotators'' while bars with $\Rot > 1.4$ are considered ``slow
+rotators'' \citep{2000ApJ...543..704D}. Bars with $\Rot < 1$ are not
 thought to be stable \citep{1980AA....81..198C}. Observational estimates of the
@@
 
-{\bf While the fact that a bar is slowed down by a dark matter halo is
-well-understood theoretically,} this is not the case for the interaction between
+While the fact that a bar is slowed down by a dark matter halo is
+well-understood theoretically, this is not the case for the interaction between
 a bar and a gaseous disk. Some argue that the gas disk should slow down the bar
@@
 evolution in the bar pattern speed over several Gyr when the gas phase is
-accounted for and {\bf robustly} modeled. This behavior has been observed in
+accounted for and robustly modeled. This behavior has been observed in
 previous works \citep{1993AA...268...65F, 2007ApJ...666..189B,
-2009ApJ...707..218V, 2010ApJ...719.1470V, 2014MNRAS.438L..81A}. {\bf We propose a new
-physical explanation for this stable behavior.}
-
-{\bf In particular, we find that the presence of a significant gas phase in a
+2009ApJ...707..218V, 2010ApJ...719.1470V, 2014MNRAS.438L..81A}. We propose a new
+physical explanation for this stable behavior.
+
+In particular, we find that the presence of a significant gas phase in a
 barred galaxy can arrest the process by which the dark matter halo brakes the
@@
 wake. We show that this occurs in Milky Way-like galaxies with gas fractions as
-low as about $5\%$.}
-
-We show {\bf stellar mass distributions} of our barred galaxy in a case with and
+low as about $5\%$.
+
+We show stellar mass distributions of our barred galaxy in a case with and
 without gas in Fig.~\ref{fig:overview}. We see that the bar grows longer and
-stronger without gas ({\bf bar length shown as a white bar}), while it remains
-at approximately the same length and strength when gas is included. {\bf As the
+stronger without gas (bar length shown as a white bar), while it remains
+at approximately the same length and strength when gas is included. As the
 bar without gas slows down, the corotation radius (white dashed circle) grows
-larger with time.}
+larger with time.
 
@@
     and \SMUGGLE{} cases. The dashed line is the torque exerted by the gas phase
-    in the \SMUGGLE{} run (there is no gas in the \Nbody{} run). {\bf After
+    in the \SMUGGLE{} run (there is no gas in the \Nbody{} run). After
     $\sim1\,\textrm{Gyr}$ of evolution, the torque by the halo in the \SMUGGLE{}
@@
     torque and pattern speed is given in
-    Section~\ref{ssec:bar_analysis}.}}\label{fig:prop}
+    Section~\ref{ssec:bar_analysis}.}\label{fig:prop}
 \end{figure}
@@
     since the bar pattern speed is not decreasing, and therefore the bar length
-    remains constant. {\bf The bar mass, shown in the \textit{middle panel}, is
+    remains constant. The bar mass, shown in the \textit{middle panel}, is
     increasing in the \Nbody{} case as the bar grows. It is decreasing in the
-    \SMUGGLE{} case, indicating a shredding process.} The bar strength, shown in
+    \SMUGGLE{} case, indicating a shredding process. The bar strength, shown in
     the \textit{lower panel} is measured as the maximum of the second Fourier
@@
 $25.3\times10^6$ in the dark matter halo. We used a softening length of
-$20\,\textrm{pc}$ for all collisionless components. {\bf This softening length
+$20\,\textrm{pc}$ for all collisionless components. This softening length
 is smaller than used in the original GALAKOS model ($28\,\textrm{pc}$ in their
@@
 component, the softening length is fully adaptive with a softening factor of
-$2.5$ \citep[e.g.,][]{2020ApJS..248...32W}.} Snapshots were saved at equal
+$2.5$ \citep[e.g.,][]{2020ApJS..248...32W}. Snapshots were saved at equal
 intervals of $0.005$ in the time units of the simulation,
@@
 treatment. A more detailed description of this model can be found in the
-flagship \SMUGGLE{} paper \citep{2019MNRAS.489.4233M}. {\bf A pedagological
+flagship \SMUGGLE{} paper \citep{2019MNRAS.489.4233M}. A pedagological
 review of cosmological simulations of galaxy formation can be found in
-\citet{2020NatRP...2...42V}. }
+\citet{2020NatRP...2...42V}.
 
@@
 $0.5\,\textrm{kpc}$ from radii of $0$ to $30\,\textrm{kpc}$. We defined the
-angle of the bar $\phi_b$ to be {\bf half} the angle of the complex number $A_2$
+angle of the bar $\phi_b$ to be half the angle of the complex number $A_2$
 as measured in the bin extending from a radius of $2.5$ to $3\,\textrm{kpc}$.
@@
 as one-half the two-sided finite gradient of $\phi_b$ as a function of time.
-{\bf We note that using the second Fourier mode is a blunt tool, discussed in
-\citet{2019arXiv190308203P}.}
+We note that using the second Fourier mode is a blunt tool, discussed in
+\citet{2019arXiv190308203P}.
 
@@
 work we referred to the native code time unit as $\textrm{Gyr}$. None of our
-results are sensitive to this choice. {\bf We applied a Savitzky-Golay filter
+results are sensitive to this choice. We applied a Savitzky-Golay filter
 \citep{1964AnaCh..36.1627S} as implemented in \texttt{scipy} using a window
@@
 (Fig.~\ref{fig:prop} and Fig.~\ref{fig:sam-torque}) and angle differences
-(Fig.~\ref{fig:wake}) in order to remove some numerical noise.}
+(Fig.~\ref{fig:wake}) in order to remove some numerical noise.
 
@@
     \includegraphics[width=\columnwidth]{fig/halo_trapped.pdf}
-    \caption{{\bf The mass fraction within two disk scale lengths
+    \caption{The mass fraction within two disk scale lengths
     ($\sim5.3\,\textrm{kpc}$) trapped in the \Nbody{} and \SMUGGLE{} runs. As
@@
     the \SMUGGLE{} case, the trapped fraction is nearly constant with time,
-    indicating the torquing process is inactive and the bar is locked.}}
+    indicating the torquing process is inactive and the bar is locked.}
     \label{fig:htrap}
@@
 grow as they slow down and the radius of corotation increases
-\citep{2000ApJ...543..704D, 2003MNRAS.341.1179A}. {\bf The middle panel of
+\citep{2000ApJ...543..704D, 2003MNRAS.341.1179A}. The middle panel of
 Fig.~\ref{fig:strength} shows the mass of the bar. As the \Nbody{} bar slows
@@
 fact that bars are not solid bodies and can respond to external torques through
-mass redistribution.}
+mass redistribution.
 
@@
 
-{\bf We test this interpretation in two ways. First, we measure the fraction of
+We test this interpretation in two ways. First, we measure the fraction of
 mass trapped in the halo. As material in the halo wake torques the bar, that
@@
 supports our interpertation that the halo wake process has shut down in the
-presence of the gas phase.}
+presence of the gas phase.
 
@@
 panel of Fig.~\ref{fig:wake}, which shows that the angle offset is larger in the
-\Nbody{} case than in the \SMUGGLE{} case by about a factor of {\bf three}. The
+\Nbody{} case than in the \SMUGGLE{} case by about a factor of three. The
 center and right panels of Fig.~\ref{fig:wake} show the halo wake with respect
 to the location of the bar in the \Nbody{} (center) and \SMUGGLE{} (right) cases
-at one point in time. {\bf Note that in Fig.~\ref{fig:wake} we have removed the
-halo material trapped in the bar, which exerts no net torque on the bar.}
+at one point in time. Note that in Fig.~\ref{fig:wake} we have removed the
+halo material trapped in the bar, which exerts no net torque on the bar.
 
@@
 resonant halo phase space at a particular pattern speed,
-$\Omega_{\textrm{p},0}$, to become {\bf mixed} and no longer able to support a
+$\Omega_{\textrm{p},0}$, to become mixed and no longer able to support a
 negative torque.\footnote{In the halo wake picture, this corresponds to the wake
@@
 positive torque on the bar, and therefore the pattern speed will again increase.
-Since at higher pattern speeds the halo has not yet been totally {\bf mixed},
+Since at higher pattern speeds the halo has not yet been totally mixed,
 the halo will once again be able to exert a negative torque on the bar. The
@@
 
-{\bf After $5\,\textrm{Gyr}$, the $10\,\Msun/\textrm{pc}^2$ simulation has a gas
-fraction of $5.7\%$, but still exhibits constant pattern speed behavior.} As a
+After $5\,\textrm{Gyr}$, the $10\,\Msun/\textrm{pc}^2$ simulation has a gas
+fraction of $5.7\%$, but still exhibits constant pattern speed behavior. As a
 result, we conclude that for the disk, bar, and halo properties considered in
@@
 \RCR{} is a parameter in the bar model from \citet{2022MNRAS.513..768C}, we
-allow the moment of inertia of the bar to vary with ${\RCR}^2$. {\bf This is
+allow the moment of inertia of the bar to vary with ${\RCR}^2$. This is
 inspired by the fact that the moment of inertia of an ellipsoid scales with the
-sum of the square of its axes.} To be more precise, we allow
+sum of the square of its axes. To be more precise, we allow
 \begin{equation}
@@
 by construction. Therefore, the locking process is reproduced in our simple
-semi-analytic model. {\bf Curiously, we do not see the bimodal behavior in
+semi-analytic model. Curiously, we do not see the bimodal behavior in
 Fig.~\ref{fig:fgas} and described in \citet{2010ApJ...719.1470V}. It is not
-presently clear why this is the case.}
+presently clear why this is the case.
 
@@
 from kinematics and chemistry. We point out that these reported measurements are
-not direct measurements of the Milky Way bar's deceleration. {\bf For instance,
+not direct measurements of the Milky Way bar's deceleration. For instance,
 \citet{2021MNRAS.500.4710C} measures the pattern speed based on the asymmetry of
 the Hercules stream, but this can also be produced by spiral arms
-\citep{2018MNRAS.481.3794H}.} Much like the simulations in the present work, the
+\citep{2018MNRAS.481.3794H}. Much like the simulations in the present work, the
 simulations of these two works do not properly account for the complicated
@@
 
-{\bf Lenticular galaxies which lack a signficant gas phase offer an opportunity
+Lenticular galaxies which lack a signficant gas phase offer an opportunity
 to find slowly rotating bars. It would still take several Gyr for a galaxy
@@
 \citep{2023MNRAS.521.2227B}. We offer another explanation based on the timing of
-when gas was stripped from these galaxies.}
-
-{\bf There are further examples of gas-rich galaxies hosting slow bars. For
+when gas was stripped from these galaxies.
+
+There are further examples of gas-rich galaxies hosting slow bars. For
 example, UGC~628 \citep[$\Rot\sim2$;][]{2009AA...499L..25C} and NGC~2915
@@
 $4.8\times10^{10}\Msun$ in our model). Further work is necessary to see how the
-gas fraction threshold varies with galactic properties.}
+gas fraction threshold varies with galactic properties.
 
@@
 
-{\bf Finally, we mention the evolution of \Rot{} in our simulation. In the
+Finally, we mention the evolution of \Rot{} in our simulation. In the
 $N$-body simulation, the bar forms with $\Rot\sim1.6$ at $t=0\,\textrm{Gyr}$,
@@
 simulation forms a bar with $\Rot\sim1.6$ is a separate question deserving
-further attention.}
+further attention.
 
 \subsection{Previous Idealized Simulation Work}
-{\bf Substantial work has been devoted to the role of gas in bar dynamics. We
-discuss this and highlight the novel aspects of the present investigation.}
-
-{\bf To our knowledge the first work on a barred galaxy with a gas component was
+Substantial work has been devoted to the role of gas in bar dynamics. We
+discuss this and highlight the novel aspects of the present investigation.
+
+To our knowledge the first work on a barred galaxy with a gas component was
 by \citet{1993AA...268...65F}. They found a stable pattern speed when a
@@
 evolved for $\sim1\,\textrm{Gyr}$, so it is unclear if their bar exhibits a
-stable pattern speed over several Gyr.}
-
-{\bf \citet{2007ApJ...666..189B} describe a model containing up to $8\%$ gas.
+stable pattern speed over several Gyr.
+
+\citet{2007ApJ...666..189B} describe a model containing up to $8\%$ gas.
 Their disk is similar to ours (though their dark matter halo is a factor of $10$
@@
 gas from the central region is an important requisite for stable pattern speeds,
-though a careful torque analysis is necessary to confirm this hypothesis.}
-
-{\bf \citet{2010ApJ...719.1470V} do find stable pattern speeds in models with
+though a careful torque analysis is necessary to confirm this hypothesis.
+
+\citet{2010ApJ...719.1470V} do find stable pattern speeds in models with
 gas fractions as low as $8\%$ (depending on the force softening used).
@@
 process by which the halo braking process is arrested, as we propose in the
-present paper.}
-
-{\bf \citet{2013MNRAS.429.1949A, 2014MNRAS.438L..81A} find stable evolution in
+present paper.
+
+\citet{2013MNRAS.429.1949A, 2014MNRAS.438L..81A} find stable evolution in
 \Rot{} for a triaxial halo over $10\,\textrm{Gyr}$ with initial (final) gas
@@
 and thus stronger torquing). These authors make no mention of the process by
-which the halo braking process is arrested, as we propose in the present paper.}
-
-{\bf There may also be issues related to structural differences between their
+which the halo braking process is arrested, as we propose in the present paper.
+
+There may also be issues related to structural differences between their
 bars and the bar considered in this work. In our case, we allowed the bar to
@@
 experience less positive torque from the gas phase. A direct comparison based on
-the torque by the gas phase on the bar is necessary.}
-
-{\bf \citet{2015MNRAS.454.3166A} describe a model containing gas and stars which
+the torque by the gas phase on the bar is necessary.
+
+\citet{2015MNRAS.454.3166A} describe a model containing gas and stars which
 slows down with time. In their model gas is added to the system to target a gas
@@
 pattern speed probably depends on galaxy properties like mass. It is unclear
-whether $10\%$ is sufficient for their bar to have a stable pattern speed.}
+whether $10\%$ is sufficient for their bar to have a stable pattern speed.
 
@@
 2022ApJ...940...61F}.\footnote{Though see \citet{2022ApJ...940...61F} who argue
-bars have consistent pattern speeds with observations, but are too short.} {\bf
+bars have consistent pattern speeds with observations, but are too short.}
 These works examine bars in EAGLE \citep{2015MNRAS.450.1937C,
@@
 factor of $6$. This highlights the importance of future work exploring precisely
-when barred galaxies ought to be in the stable regime.}
-
-{\bf \citet{2021AA...650L..16F} explored the evolution of bars in the Auriga
+when barred galaxies ought to be in the stable regime.
+
+\citet{2021AA...650L..16F} explored the evolution of bars in the Auriga
 cosmological zoom simulations \citep{2017MNRAS.467..179G}. They find \Rot{}
@@
 appears to even transition from the stable pattern speed regime to the slowing
-down regime at a lookback time of $\sim1.8\,\textrm{Gyr}$.}
+down regime at a lookback time of $\sim1.8\,\textrm{Gyr}$.
 
@@
 observational evidence \citep{2019MNRAS.490.4740B}. And finally, we complicate
-the picture of {\bf stellar} radial mixing expected to sculpt the Milky Way's
+the picture of stellar radial mixing expected to sculpt the Milky Way's
 disk \citep{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies
-upon the pattern speed of the bar to change with time. {\bf The radial mixing of
+upon the pattern speed of the bar to change with time. The radial mixing of
 the gas phase induced by the bar, as predicted in \citet{2011MNRAS.415.1027H},
-might have implications for the radial metallicity gradients of galaxies.} Our
+might have implications for the radial metallicity gradients of galaxies. Our
 work does not alter expectations for radial mixing induced by spiral arms
@@
 rotating. Indeed, they would need to be gas-deficient for several $\textrm{Gyr}$
-before they would be classified as slow rotators.} {\bf There does exist
+before they would be classified as slow rotators.} There does exist
 examples of galaxies known to be slow rotators -- the low surface brightness
@@
 structural properties than the Milky Way-like model we considered in this work,
-discussed in Section~\ref{ssec:observations}.} We predict a general trend that
+discussed in Section~\ref{ssec:observations}. We predict a general trend that
 bars in gas-rich spiral galaxies should rotate quickly while some bars in
@@
 $A_{2,\textrm{bar}}/A_{2,\textrm{tot}}\sim0.76$). Some of this is probably
-coming from the $m=4$ component {\bf (e.g., boxy orbits)} being classified as
-unbarred, {\bf or the presence of weak spiral arms. See also
-\citet{2021MNRAS.500..838P} for more details on the orbit family breakdown.}
+coming from the $m=4$ component (e.g., boxy orbits) being classified as
+unbarred, or the presence of weak spiral arms. See also
+\citet{2021MNRAS.500..838P} for more details on the orbit family breakdown.
 
@@
 
-{\bf We performed an experiment to test this hypothesis by freezing the stellar
+We performed an experiment to test this hypothesis by freezing the stellar
 disk in the \SMUGGLE{} run and forcing it to rotate at a constant angular rate.
@@
 angular rate which we control. The gas is evolved self-consistently with this
-rotating disk.} We measured the torque on the bar by the gas phase at different
+rotating disk. We measured the torque on the bar by the gas phase at different
 rotation rates. The result of this experiment is illustrated in
@@
 
-{\bf We also note that since \citet{2011MNRAS.415.1027H} predicts gas outside of
+We also note that since \citet{2011MNRAS.415.1027H} predicts gas outside of
 corotation will flow outward, the bar should exert a positive torque on that
@@
 units. So, while gas outside corotation does experience a positive torque, the
-total torque on the gas phase is still negative.}
+total torque on the gas phase is still negative.
 
@@
     \includegraphics[width=9cm]{fig/ps_star.pdf}
-    \caption{{\bf Pattern speed evolution of a model in which we instaneously
+    \caption{Pattern speed evolution of a model in which we instaneously
     add stars instead of gas to the simulation, with the same density profile as
@@
     SMUGGLE case is not simply a consequence of the change in potential imposed
-    in our initial conditions.}}
+    in our initial conditions.}
 \label{fig:ps-star}
@@
 
-{\bf \section{Stars Instead of Gas}
+\section{Stars Instead of Gas}
 In the SMUGGLE model considered in this work, we instantaneously added gas to
@@
 speed evolution is broadly consistent with a declining pattern speed. This
-indicates that the gas phase is responsible for the stable pattern speed.}
+indicates that the gas phase is responsible for the stable pattern speed.
 \section{Semi-Analytic Model Parameters}
@@
 $10\,\textrm{kpc}$ in the vertical direction. We fit the dark matter halo using
-{\bf an axisymmetric} multipole fit with a maximum angular harmonic coefficient
-of $l=2$, {\bf to account for the compression of the halo by the disk}. We plot
-the circular velocity curve {\bf at $t=1\,\textrm{Gyr}$} in Fig.~\ref{fig:vcirc}
+an axisymmetric multipole fit with a maximum angular harmonic coefficient
+of $l=2$, to account for the compression of the halo by the disk. We plot
+the circular velocity curve at $t=1\,\textrm{Gyr}$ in Fig.~\ref{fig:vcirc}
 compared to observational estimates \citep{2019ApJ...871..120E}. The \SMUGGLE{}
@@
 Milky Way \citep{2008AA...487..951K,2022ApJ...929L..18E}. The discrepancy between
-$1$ and $4\,\textrm{kpc}$ in the molecular and SFR surface density is {\bf probably} due
+$1$ and $4\,\textrm{kpc}$ in the molecular and SFR surface density is probably due
 to the fact that the distances to molecular clouds which underlines this work
@@
 
-{\bf We measured the initial scale height of the atomic gas disk in a bin
+We measured the initial scale height of the atomic gas disk in a bin
 extending from $R=7.5\,\textrm{kpc}$ to $R=8.5\,\textrm{kpc}$. The initial
@@
 This may be caused by the model in our simulations not driving enough turbulent
-pressure, and is an interesting avenue of further investigation.}
+pressure, and is an interesting avenue of further investigation.
 
@@
     \includegraphics[width=9cm]{fig/vcirc.pdf}
-    \caption{The circular velocity curve of our setups {\bf at $t=1\,\textrm{Gyr}$}.
+    \caption{The circular velocity curve of our setups at $t=1\,\textrm{Gyr}$.
     This curve is shown for the \Nbody{} run (blue) and the \SMUGGLE{} run

### 2023-05-30 paper/ms.tex
Remove shredding term

@@
     increasing in the \Nbody{} case as the bar grows. It is decreasing in the
-    \SMUGGLE{} case, indicating a shredding process. The bar strength, shown in
+    \SMUGGLE{} case, indicating {\bf mass redistribution without patttern speed 
+    changes}. The bar strength, shown in
     the \textit{lower panel} is measured as the maximum of the second Fourier
@@
 down and lengthens, it also grows in mass. The \SMUGGLE{} bar, however, loses
-mass over time. This indicates an ongoing shredding process, and highlights the
+mass over time. This indicates {a change in the bar's angular momentum without 
+a change in pattern speed}, and highlights the
 fact that bars are not solid bodies and can respond to external torques through
@@
 body, it is not guaranteed that a negative torque will slow it down - e.g. a
-negative torque could shred the bar, reducing its moment of inertia without
+negative torque could {\bf reduce the mass of} the bar, reducing its moment of inertia without
 changing its pattern speed. However, the bar seems to empirically respond to a
@@
 reinforce the wake, therefore arresting the process by which the halo can slow
-the bar down. We term this process ``bar locking.''
+the bar down. We term this process {\bf ``bar pattern speed locking,'' or simply
+bar locking for short.}\footnote{{\bf To be more explicit, it is the gas which locks 
+the pattern speed of the bar. The gas does this by forcing the resonant locations 
+of the bar into regions of the halo phase space that can no longer support 
+significant negative torque.}}
 

### 2023-05-30 paper/ms.tex
Fix shred changes

@@
     increasing in the \Nbody{} case as the bar grows. It is decreasing in the
-    \SMUGGLE{} case, indicating {\bf mass redistribution without patttern speed 
-    changes}. The bar strength, shown in
+    \SMUGGLE{} case, indicating {\bf mass redistribution without a change in patttern speed}. The bar strength, shown in
     the \textit{lower panel} is measured as the maximum of the second Fourier
@@
 down and lengthens, it also grows in mass. The \SMUGGLE{} bar, however, loses
-mass over time. This indicates {a change in the bar's angular momentum without 
+mass over time. This indicates {\bf a change in the bar's angular momentum without 
 a change in pattern speed}, and highlights the

### 2023-05-30 paper/ms.tex
Further changes to paper

@@
 \author{Federico Marinacci}
-\affiliation{Department of Physics \& Astronomy `Augusto Righi', University of Bologna, Bologna, Italy}
+\affiliation{Department of Physics \& Astronomy `Augusto Righi', 
+University of Bologna, Bologna, Italy}
 
@@
 %% to include these concepts in their preprints.
-\keywords{Milky Way Galaxy (1054) --- Milky Way evolution (1052) --- Milky Way dynamics (1051) --- Galaxy dynamics (591) --- Hydrodynamical simulations (767) --- Barred spiral galaxies (136)}
+\keywords{Milky Way Galaxy (1054) --- Milky Way evolution (1052) --- Milky Way dynamics (1051)
+ --- Galaxy dynamics (591) --- Hydrodynamical simulations (767) --- Barred spiral galaxies (136)}
 
@@
 
-In particular, we find that the presence of a significant gas phase in a
-barred galaxy can arrest the process by which the dark matter halo brakes the
-bar. The constant positive torque on the bar by the gas causes the bar's
-resonance to reside in regions which have become depopulated. In the halo wake
-picture, this is equivalent to no new material being available to reinforce the
-wake. We show that this occurs in Milky Way-like galaxies with gas fractions as
-low as about $5\%$.
+In particular, we find that the presence of a gas phase in a barred
+galaxy can arrest the process by which the dark matter halo brakes the bar. The
+constant positive torque on the bar by the gas causes the bar's resonance to
+reside in regions which have become depopulated. In the halo wake picture, this
+is equivalent to no new material being available to reinforce the wake. We show
+that this occurs in {\bf a Milky Way-like disk} with gas fractions as low as
+about $5\%$.
 
@@
     increasing in the \Nbody{} case as the bar grows. It is decreasing in the
-    \SMUGGLE{} case, indicating {\bf mass redistribution without a change in patttern speed}. The bar strength, shown in
-    the \textit{lower panel} is measured as the maximum of the second Fourier
-    component divided by the zeroth Fourier component. We see that in the
-    \Nbody{} case (blue) the bar strength increases with time, consistent with
-    previous results showing that the bar strength increases as bars slow down.
-    In the \SMUGGLE{} case (orange), we see that the bar strength slightly
-    decreasing with time. This is also consistent with the expected relation
-    between pattern speed and strength since the bar in this case is not slowing
-    down.}
+    \SMUGGLE{} case, indicating {\bf mass redistribution without a change in
+    pattern speed}. The bar strength, shown in the \textit{lower panel} is
+    measured as the maximum of the second Fourier component divided by the
+    zeroth Fourier component. We see that in the \Nbody{} case (blue) the bar
+    strength increases with time, consistent with previous results showing that
+    the bar strength increases as bars slow down. In the \SMUGGLE{} case
+    (orange), we see that the bar strength slightly decreasing with time. This
+    is also consistent with the expected relation between pattern speed and
+    strength since the bar in this case is not slowing down.}
     \label{fig:strength}
@@
 disk. Our fiducial model used an initial surface density of
-$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\textrm{kpc}$. The
-initial gas disk is generated with a temperature of $10^4\,\textrm{K}$ and solar
-metallicity.
+$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\textrm{kpc}$. {\bf
+This corresponds to an initial gas fraction of $\sim16\%$.} The initial gas disk
+is generated with a temperature of $10^4\,\textrm{K}$ and solar metallicity.
 
@@
 
+{\bf Our method for initializing gas was arrived at after numerous attempts to
+include enough gas in the simulation to be compatible with the Milky Way while
+also not destroying the bar. For example, we tried evolving an exponential gas
+disk adiabatically with the barred disk (i.e., no cooling, star formation, or
+feedback). Because there is no mechanism to remove gas from the central region,
+a large pileup of gas forms in the center.\footnote{{\bf We noticed that the bar
+slows down in this adiabatic model which lacks a mechanism to remove gas from
+the central region. \citet{2007ApJ...666..189B} also noted slowdown behavior in
+models which lacked central gas removal. It appears that gas removal from the
+center (in our model due to star formation) is necessary to stabilize a bar's
+pattern speed.}} Once cooling is turned on, there is a sudden collapse of gas to
+the center, destroying the bar. We don't believe our method is the only nor even
+the best way to include gas. One advantage of our method, though, is that it
+approximates the expectation that in the bar region the gas surface density
+should be significantly reduced \citep[e.g.][]{1993RPPh...56..173S}}
+
 We used a mass resolution of $7.5\times10^3\,M_{\odot}$ for the baryonic
@@
 $25.3\times10^6$ in the dark matter halo. We used a softening length of
-$20\,\textrm{pc}$ for all collisionless components. This softening length
-is smaller than used in the original GALAKOS model ($28\,\textrm{pc}$ in their
+$20\,\textrm{pc}$ for all collisionless components. This softening length is
+smaller than used in the original GALAKOS model ($28\,\textrm{pc}$ in their
 model, but $\sim43.5\,\textrm{pc}$ when scaled to our mass resolution). Our
@@
     \includegraphics[width=\columnwidth]{fig/halo_trapped.pdf}
-    \caption{The mass fraction within two disk scale lengths
+    \caption{The {\bf halo} mass fraction within two disk scale lengths
     ($\sim5.3\,\textrm{kpc}$) trapped in the \Nbody{} and \SMUGGLE{} runs. As
@@
 approximately $16\%$, $10\%$, $7\%$, and $4\%$. The pattern speed evolution is
-shown in Fig.~\ref{fig:fgas} We find that the bar in disks with
+shown in Fig.~\ref{fig:fgas}. We find that the bar in disks with
 initial surface densities of $20$, $15$, and $10\,\Msun/\textrm{pc}^2$ evolve
@@
 this work, a gas fraction of only approximately $5\%$ is necessary in order for
-the proposed stabilizing mechanism to operate.
-
-\subsection{Smooth Interstellar Medium}
-We performed a simulation of the same disk but with a simpler model of the
-interstellar medium \citep{2003MNRAS.339..289S}, closer to standard methods used
-in cosmological simulations of galaxy formation and described in more detail in
-Section~\ref{sec:methods}. The result of this test is presented in
-Fig.~\ref{fig:GFM}. We find that the pattern speed evolution is nearly the same
-in this case, and so conclude that our result is not sensitive to the details of
-the model for the interstellar medium.
+the proposed stabilizing mechanism to operate. {\bf We stress that this gas
+fraction threshold is only for the system considered in this work. Systems with
+different structural parameters may require a different threshold. For example,
+\citet{2015MNRAS.454.3166A} find slowdown behavior in a barred disk $10$ times
+less massive than ours with a gas fraction of $10\%$. To make things more
+complicated, \citet{2010ApJ...719.1470V} find that the gas fraction cutoff
+varies with the softening length used. Determining exactly how the gas fraction
+cutoff varies with these considerations deserves further attention.}
 
@@
 
-\citet{2013MNRAS.429.1949A, 2014MNRAS.438L..81A} find stable evolution in
-\Rot{} for a triaxial halo over $10\,\textrm{Gyr}$ with initial (final) gas
-fractions of $100\%$ ($7\%$), $75\%$ ($6\%$), and $50\%$ ($5\%$). For a model
-with $20\%$ ($3\%$) gas fraction they find an increasing \Rot{}. In their model
-with a spherical halo they always find increasing \Rot{}, contrary to the
-present work. We note there is a structural difference in their dark matter
-halo. They use a cored isothermal sphere as opposed to our Hernquist halo. It is
-not clear to us the impact this would have on the expected torque from the halo
-(i.e., the velocity structure of their halo may allow for more efficient capture
-and thus stronger torquing). These authors make no mention of the process by
-which the halo braking process is arrested, as we propose in the present paper.
+\citet{2013MNRAS.429.1949A, 2014MNRAS.438L..81A} find stable evolution in \Rot{}
+for a triaxial halo over $10\,\textrm{Gyr}$ with initial (final) gas fractions
+of $100\%$ ($7\%$), $75\%$ ($6\%$), and $50\%$ ($5\%$). For a model with $20\%$
+($3\%$) gas fraction they find an increasing \Rot{}. In their model with a
+spherical halo they always find increasing \Rot{}, contrary to the present work.
+We note there is a structural difference in their dark matter halo. They use a
+cored isothermal sphere as opposed to our Hernquist halo. It is not clear to us
+the impact this would have on the expected torque from the halo (i.e., the
+velocity structure of their halo may allow for more efficient capture and thus
+stronger torquing). {\bf The bar locking process (or something similar) which we
+propose in this paper is not mentioned by these authors.}
 
@@
 presently engaged in such an exploration.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=9cm]{fig/ps_GFM.pdf}
-    \caption{Pattern speed evolution of a smooth ISM model. This evolution is
-    shown for the fiducial disk in the \Nbody{} (blue), \SMUGGLE{} (orange), and
-    smooth ISM (red) cases. The smooth ISM model is an older model for the ISM
-    which treats its multiphase nature in a subgrid fashion
-    \citep{2003MNRAS.339..289S}. This fundamentally differs from the \SMUGGLE{}
-    model, which explicitly resolves the hot and cold phases of the ISM
-    \citep{2019MNRAS.489.4233M}. The pattern speed in the smooth ISM case is
-    broadly similar to the evolution in the \SMUGGLE{} case. This shows that the
-    stability of the pattern speed is not simply a result of our assumed model
-    for the ISM.}
-\label{fig:GFM}
-\end{figure}
 
@@
 
+{\bf Snapshots at $500\,\textrm{Myr}$ cadence are publicly available at
+\url{https://drive.google.com/drive/folders/1nZF7mZ98T0QPc2pt7CVCUUdQAm7Q9nK-?usp=sharing}.}
+
 \begin{acknowledgments}
-  We would like to thank Greg~L. Bryan, Neal~J. Evans, Drummond~B. Fielding,
-  Keith Hawkins, Jason~A.~S. Hunt, Sarah~M.~R. Jeffreson, Kathryn~V. Johnston,
+  {\bf We would like to thank the referee, Michael~S. Petersen, for their
+  thorough report which greatly strengthened the arguments in our work.} We
+  would like to thank Greg~L. Bryan, Neal~J. Evans, Drummond~B. Fielding, Keith
+  Hawkins, Jason~A.~S. Hunt, Sarah~M.~R. Jeffreson, Kathryn~V. Johnston,
   Peter~M.~W. Kalberla, Jürgen Kerp, Julio~F. Navarro, Dylan Nelson, Suchira
@@
 indicates that the gas phase is responsible for the stable pattern speed.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=9cm]{fig/ps_GFM.pdf}
+    \caption{Pattern speed evolution of a smooth ISM model. This evolution is
+    shown for the fiducial disk in the \Nbody{} (blue), \SMUGGLE{} (orange), and
+    smooth ISM (red) cases. The smooth ISM model is an older model for the ISM
+    which treats its multiphase nature in a subgrid fashion
+    \citep{2003MNRAS.339..289S}. This fundamentally differs from the \SMUGGLE{}
+    model, which explicitly resolves the hot and cold phases of the ISM
+    \citep{2019MNRAS.489.4233M}. The pattern speed in the smooth ISM case is
+    broadly similar to the evolution in the \SMUGGLE{} case. This shows that the
+    stability of the pattern speed is not simply a result of our assumed model
+    for the ISM.}
+\label{fig:GFM}
+\end{figure}
+
+\section{Smooth Interstellar Medium}
+We performed a simulation of the same disk but with a simpler model of the
+interstellar medium \citep{2003MNRAS.339..289S}, closer to standard methods used
+in cosmological simulations of galaxy formation and described in more detail in
+Section~\ref{sec:methods}. The result of this test is presented in
+Fig.~\ref{fig:GFM}. We find that the pattern speed evolution is nearly the same
+in this case, and so conclude that our result is not sensitive to the details of
+the model for the interstellar medium.
+
 \section{Semi-Analytic Model Parameters}

### 2023-06-04 paper/ms.tex
Remove boldface

@@
 is equivalent to no new material being available to reinforce the wake. We show
-that this occurs in {\bf a Milky Way-like disk} with gas fractions as low as
+that this occurs in a Milky Way-like disk with gas fractions as low as
 about $5\%$.
@@
     increasing in the \Nbody{} case as the bar grows. It is decreasing in the
-    \SMUGGLE{} case, indicating {\bf mass redistribution without a change in
-    pattern speed}. The bar strength, shown in the \textit{lower panel} is
+    \SMUGGLE{} case, indicating mass redistribution without a change in
+    pattern speed. The bar strength, shown in the \textit{lower panel} is
     measured as the maximum of the second Fourier component divided by the
@@
 disk. Our fiducial model used an initial surface density of
-$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\textrm{kpc}$. {\bf
-This corresponds to an initial gas fraction of $\sim16\%$.} The initial gas disk
+$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\textrm{kpc}$.
+This corresponds to an initial gas fraction of $\sim16\%$. The initial gas disk
 is generated with a temperature of $10^4\,\textrm{K}$ and solar metallicity.
@@
 
-{\bf Our method for initializing gas was arrived at after numerous attempts to
+Our method for initializing gas was arrived at after numerous attempts to
 include enough gas in the simulation to be compatible with the Milky Way while
@@
 feedback). Because there is no mechanism to remove gas from the central region,
-a large pileup of gas forms in the center.\footnote{{\bf We noticed that the bar
+a large pileup of gas forms in the center.\footnote{We noticed that the bar
 slows down in this adiabatic model which lacks a mechanism to remove gas from
@@
 center (in our model due to star formation) is necessary to stabilize a bar's
-pattern speed.}} Once cooling is turned on, there is a sudden collapse of gas to
+pattern speed.} Once cooling is turned on, there is a sudden collapse of gas to
 the center, destroying the bar. We don't believe our method is the only nor even
@@
 approximates the expectation that in the bar region the gas surface density
-should be significantly reduced \citep[e.g.][]{1993RPPh...56..173S}}
+should be significantly reduced \citep[e.g.][]{1993RPPh...56..173S}
 
@@
     \includegraphics[width=\columnwidth]{fig/halo_trapped.pdf}
-    \caption{The {\bf halo} mass fraction within two disk scale lengths
+    \caption{The halo mass fraction within two disk scale lengths
     ($\sim5.3\,\textrm{kpc}$) trapped in the \Nbody{} and \SMUGGLE{} runs. As
@@
 down and lengthens, it also grows in mass. The \SMUGGLE{} bar, however, loses
-mass over time. This indicates {\bf a change in the bar's angular momentum without 
-a change in pattern speed}, and highlights the
+mass over time. This indicates a change in the bar's angular momentum without 
+a change in pattern speed, and highlights the
 fact that bars are not solid bodies and can respond to external torques through
@@
 body, it is not guaranteed that a negative torque will slow it down - e.g. a
-negative torque could {\bf reduce the mass of} the bar, reducing its moment of inertia without
+negative torque could reduce the mass of the bar, reducing its moment of inertia without
 changing its pattern speed. However, the bar seems to empirically respond to a
@@
 reinforce the wake, therefore arresting the process by which the halo can slow
-the bar down. We term this process {\bf ``bar pattern speed locking,'' or simply
-bar locking for short.}\footnote{{\bf To be more explicit, it is the gas which locks 
+the bar down. We term this process ``bar pattern speed locking,'' or simply
+bar locking for short.\footnote{ To be more explicit, it is the gas which locks 
 the pattern speed of the bar. The gas does this by forcing the resonant locations 
 of the bar into regions of the halo phase space that can no longer support 
-significant negative torque.}}
+significant negative torque.}
 
@@
 this work, a gas fraction of only approximately $5\%$ is necessary in order for
-the proposed stabilizing mechanism to operate. {\bf We stress that this gas
+the proposed stabilizing mechanism to operate. We stress that this gas
 fraction threshold is only for the system considered in this work. Systems with
@@
 varies with the softening length used. Determining exactly how the gas fraction
-cutoff varies with these considerations deserves further attention.}
+cutoff varies with these considerations deserves further attention.
 
@@
 velocity structure of their halo may allow for more efficient capture and thus
-stronger torquing). {\bf The bar locking process (or something similar) which we
-propose in this paper is not mentioned by these authors.}
+stronger torquing). The bar locking process (or something similar) which we
+propose in this paper is not mentioned by these authors.
 
@@
 
-{\bf Snapshots at $500\,\textrm{Myr}$ cadence are publicly available at
-\url{https://drive.google.com/drive/folders/1nZF7mZ98T0QPc2pt7CVCUUdQAm7Q9nK-?usp=sharing}.}
+Snapshots at $500\,\textrm{Myr}$ cadence are publicly available at
+\url{https://drive.google.com/drive/folders/1nZF7mZ98T0QPc2pt7CVCUUdQAm7Q9nK-?usp=sharing}.
 
 \begin{acknowledgments}
-  {\bf We would like to thank the referee, Michael~S. Petersen, for their
-  thorough report which greatly strengthened the arguments in our work.} We
-  would like to thank Greg~L. Bryan, Neal~J. Evans, Drummond~B. Fielding, Keith
+  We would like to thank Greg~L. Bryan, Neal~J. Evans, Drummond~B. Fielding, Keith
   Hawkins, Jason~A.~S. Hunt, Sarah~M.~R. Jeffreson, Kathryn~V. Johnston,

### 2023-06-04 paper/ms.tex
Add better description of adiabatic experiment

--- 2023-05-30 paper/ms.tex
@@
 feedback). Because there is no mechanism to remove gas from the central region,
-a large pileup of gas forms in the center.\footnote{{\bf We noticed that the bar
-slows down in this adiabatic model which lacks a mechanism to remove gas from
-the central region. \citet{2007ApJ...666..189B} also noted slowdown behavior in
-models which lacked central gas removal. It appears that gas removal from the
-center (in our model due to star formation) is necessary to stabilize a bar's
-pattern speed.}} Once cooling is turned on, there is a sudden collapse of gas to
-the center, destroying the bar. We don't believe our method is the only nor even
-the best way to include gas. One advantage of our method, though, is that it
-approximates the expectation that in the bar region the gas surface density
-should be significantly reduced \citep[e.g.][]{1993RPPh...56..173S}}
+a large, highly pressurized pileup of gas forms in the center.\footnote{{\bf We
+noticed that the bar slows down in this adiabatic model which lacks a mechanism
+to remove gas from the central region. \citet{2007ApJ...666..189B} also noted
+slowdown behavior in models which lacked central gas removal. It appears that
+gas removal from the center (in our model due to star formation) is necessary to
+stabilize a bar's pattern speed.}} We then turned on the full SMUGGLE model. As
+a result, there is a sudden collapse of gas to the center as pressure support is
+lost due to cooling and star formation. This abrupt change in the potential
+destroys the bar.
+
+We don't believe our method is the only nor even the best way to include gas.
+One advantage of our method, though, is that it approximates the expectation
+that in the bar region the gas surface density should be significantly reduced
+\citep[e.g.][]{1993RPPh...56..173S}}
 

### 2023-06-04 paper/ms.tex

--- 2023-06-04 paper/ms.tex
@@
 feedback). Because there is no mechanism to remove gas from the central region,
-a large pileup of gas forms in the center.\footnote{We noticed that the bar
-slows down in this adiabatic model which lacks a mechanism to remove gas from
-the central region. \citet{2007ApJ...666..189B} also noted slowdown behavior in
-models which lacked central gas removal. It appears that gas removal from the
-center (in our model due to star formation) is necessary to stabilize a bar's
-pattern speed.} Once cooling is turned on, there is a sudden collapse of gas to
-the center, destroying the bar. We don't believe our method is the only nor even
-the best way to include gas. One advantage of our method, though, is that it
-approximates the expectation that in the bar region the gas surface density
-should be significantly reduced \citep[e.g.][]{1993RPPh...56..173S}
+a large, highly pressurized pileup of gas forms in the center.\footnote{ We
+noticed that the bar slows down in this adiabatic model which lacks a mechanism
+to remove gas from the central region. \citet{2007ApJ...666..189B} also noted
+slowdown behavior in models which lacked central gas removal. It appears that
+gas removal from the center (in our model due to star formation) is necessary to
+stabilize a bar's pattern speed.} We then turned on the full SMUGGLE model. As
+a result, there is a sudden collapse of gas to the center as pressure support is
+lost due to cooling and star formation. This abrupt change in the potential
+destroys the bar.
+
+We don't believe our method is the only nor even the best way to include gas.
+One advantage of our method, though, is that it approximates the expectation
+that in the bar region the gas surface density should be significantly reduced
+\citep[e.g.][]{1993RPPh...56..173S}
 

### 2023-06-06 paper/ms.tex
Remove line numbers
Add Federico comments

--- 2023-06-04 paper/ms.tex
@@
 
-We don't believe our method is the only nor even the best way to include gas.
+We do not believe our method is the only nor even the best way to include gas.
 One advantage of our method, though, is that it approximates the expectation
@@
 varies with the softening length used. Determining exactly how the gas fraction
-cutoff varies with these considerations deserves further attention.}
+cutoff varies with these properties deserves further attention.}
 
@@
 velocity structure of their halo may allow for more efficient capture and thus
-stronger torquing). {\bf The bar locking process (or something similar) which we
-propose in this paper is not mentioned by these authors.}
+stronger torquing). {\bf The bar locking process (or a similar mechanism) that
+we propose in this paper is not mentioned by these authors.}
 

### 2023-06-06 paper/ms.tex
Update citation of Aumer and Schonrich

@@
 different structural parameters may require a different threshold. For example,
-\citet{2015MNRAS.454.3166A} find slowdown behavior in a barred disk $10$ times
-less massive than ours with a gas fraction of $10\%$. To make things more
-complicated, \citet{2010ApJ...719.1470V} find that the gas fraction cutoff
+\citet{2015MNRAS.454.3166A} find slowdown behavior in a barred disk initially
+$10$ times less massive than ours with a gas fraction of $10\%$. To make things
+more complicated, \citet{2010ApJ...719.1470V} find that the gas fraction cutoff
 varies with the softening length used. Determining exactly how the gas fraction
@@
 slows down with time. In their model gas is added to the system to target a gas
-fraction of $10\%$. However, their disk is about a factor of $10$ less massive
+fraction of $10\%$. {\bf Their disk is initially a factor of $10$ less massive
+than the disk considered in this work, but they grow their disk to be about the
+same mass as ours. They find their bar lengthens and slow down with time, which
+they attribute to the dark matter halo slowdown dominating over the gas inflow.
+However, it may also be the case that a bar in a disk grown in their way
+naturally extends its length as fresh material to be captured by the bar is
+added to the system. A useful experiment would be to grow a barred disk with a
+static dark matter halo to see if its pattern speed evolution mimics theirs, to
+determine if their lengthening and slowing down is truly attributed to the halo
+friction.}
+
+However, their disk is about a factor of $10$ less massive
 than the disk considered in this work. The necessary gas fraction for a stable

### 2023-07-02 plots/halo_wake/plot_halo_wake.py
Correct factor of pi in talk plots

@@ -266,17 +266,17 @@ def run():
     # talk plots
 
     fig, ax = plt.subplots(1, 1, figsize=(3.5, 3.5))
-    ax.plot(tN - tN[300], 180.*dphiN, c=tb_c[0], label='N-body')
+    ax.plot(tN - tN[300], 180.*dphiN/np.pi, c=tb_c[0], label='N-body')
 
     ax.legend(frameon=False)
 
-    ax.set(xlim=(0, 5), ylim=(-35, 70), xlabel=r'$t\,[\,\text{Gyr}\,]$', ylabel=r'$\text{angle difference}\,[\,\text{deg}\,]$')
+    ax.set(xlim=(0, 5), ylim=(-20, 40), xlabel=r'$t\,[\,\text{Gyr}\,]$', ylabel=r'$\text{angle difference}\,[\,\text{deg}\,]$')
     ax.axhline(0, c='k')
 
     fig.tight_layout()
     fig.savefig('talk_fig3a.pdf')
 
-    ax.plot(tS, 180.*dphiS, c=tb_c[1], label='SMUGGLE')
+    ax.plot(tS, 180.*dphiS/np.pi, c=tb_c[1], label='SMUGGLE')
 
     ax.legend(frameon=False)
 
